### 2024-10-03 analysis/paper/access-ice.qmd
Adds draft 1

---
title: "Untitled"
format: html
execute: 
  echo: false
  warning: false
  message: false
filters:
  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
  - ../../_extensions/mloubout/critic-markup/critic-markup.lua
---

```{r setup, include=FALSE}
library(lubridate)
library(ggplot2)
library(data.table)
library(metR)
library(rcdo)

cdo_options_set("-L")
source(here::here("R/functions.R"))
source(here::here("R/datasets.R"))
source(here::here("R/ggplot.R"))
```

# Abstract

balbalbal

# Introduction

{==**Antarctic sea ice general description**. Seasonality, trends.==}

{==**Importance**. Impacts on life. Potential impacts on weather. Impacts on climate. Impacts on ocean circulation.(Justify why it's important to study and understand)==}

{==**Prediction systems.** Why is it important: operations {\>\>Discuss with Phil\<\<}; science needs (we need good models). What is available. Discussion on S1 {\>\>Can we discuss previous S1 results? Laura's paper is not public.\<\<}==}

{==**Objective.** Evaluate S2.==}

# Data and methods

All datasets are regrided using bilinear interpolation to a common grid, which is a polar sterographic grid with approximatedly equal area of 25km².

## Description of the model

ACCESS-S2 is the next seasonal prediction system after ACCESS-S1. It uses virtually the same exact model configuration except for some model versions being different. The main difference are the initial conditions. While S1 used initial conditions from the UK Met Office, S2 uses the Bureau's own system. {== Description of the system ==}{\>\>Atmosphere and ocean. Highlight no ice data assimilation\<\<}

For evaluation we use hindcast for the period 1981--2023{\>\>Check\<\<}. Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.

## Verification datasets

There is not a lot of data on sea ice properties, especially for things like thickness, age, etc. However there are relativelly reliable satellite-derived estimates of sea ice concentration, which esetimates the proportion of each grid area that is covered with ice. These products are not perfect and there still exists a fair bit of observation al uncertainty. To account for this uncertanty we use multiple datasets.

### Bootstrap

{\>\>From https://nsidc.org/data/nsidc-0079/versions/4\<\<} This sea ice concentration data set was derived using measurements from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites. Measurements from the Special Sensor Microwave Imager/Sounder (SSMIS) aboard DMSP-F17 are also included. The data set has been generated using the Advanced Microwave Scanning Radiometer - Earth Observing System (AMSR-E) Bootstrap Algorithm with daily varying tie-points. Daily (every other day prior to July 1987) and monthly data are available for both the north and south polar regions. Data are gridded on the SSM/I polar stereographic grid (25 x 25 km) and provided in two-byte integer format. Data coverage began on 01 November 1978 and is ongoing through the most current processing, with updated data processed several times annually.

Comiso, J.C., Gersten, R.A. 2023. Bootstrap Sea Ice Concentrations from Nimbus-7 SMMR and DMSP SSM/I-SSMIS, Version 4. \[Indicate subset used\]. Boulder, Colorado USA. NASA National Snow and Ice Data Center Distributed Active Archive Center. https://doi.org/10.5067/X5LG68MH013O. \[Date Accessed\] {\>\>Add to zotero\<\<}

### NASA Team

{==missing==}

### CDR

NOAA/NSIDC's Climate Data Record V4 combines the Bootstrap and NASA Team estimates in an attempt to overcome each algorithm's weaknesses. {\>\>More detail on how are they combined and their properties.\<\<} As such, it is is not a trully independend dataset.

Meier, W. N., F. Fetterer, A. K. Windnagel, and S. Stewart. 2021. NOAA/NSIDC Climate Data Record of Passive Microwave Sea Ice Concentration, Version 4. \[Indicate subset used\]. Boulder, Colorado USA. NSIDC: National Snow and Ice Data Center https://doi.org/10.7265/efmz-2t65. \[Date Accessed\] {\>\>Add to zotero\<\<}

### ERA5

{==From the CDS websiste:==} ERA5 is the fifth generation ECMWF reanalysis for the global climate and weather for the past 8 decades. Data is available from 1940 onwards. ERA5 replaces the ERA-Interim reanalysis.

This parameter is the fraction of a grid box which is covered by sea ice. Sea ice can only occur in a grid box which includes ocean or inland water according to the land-sea mask and lake cover, at the resolution being used. This parameter can be known as sea-ice (area) fraction, sea-ice concentration and more generally as sea-ice cover. In ERA5, sea-ice cover is given by two external providers. Before 1979 the HadISST2 dataset is used. From 1979 to August 2007 the OSI SAF (409a) dataset is used and from September 2007 the OSI SAF oper dataset is used. Sea ice is frozen sea water which floats on the surface of the ocean. Sea ice does not include ice which forms on land such as glaciers, icebergs and ice-sheets. It also excludes ice shelves which are anchored on land, but protrude out over the surface of the ocean. These phenomena are not modelled by the IFS. Long-term monitoring of sea ice is important for understanding climate change. Sea ice also affects shipping routes through the polar regions.

{>>Why is ERA5 better or worse than satellites?<<}

## Error measures

FOr evaluation purposes, we use a series of measures.

### Sea ice extent

Sea Ice Extent is defined as the area of ocean covered with at least 15% ice. This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for low sea ice conditions.

Sea Ice Extent is a rough global measure, but a model could have relatively accurate extent of ice but with different distributions. We use two other measures to account for these errors.

We compute Root Mean Square Error of sea ice concentration anomalies.

We also compute the Integrated Ice Edge Error (IIEE) {\>\>CITE\<\<}. This is defined as the area in which the model misspredicts the presence of more than 15% ice. That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.

# Results and discussion

## Reanalysis

### Bias

```{r datasets}
datasets <- list(cdr = CDR(),
                 bt = BT(),
                 era5 = ERA5(),
                 access = S2_reanalysis()
                )
```
```{r extent_daily}
extent_daily <- datasets |>
    lapply(extent) |>
    lapply(\(x) ReadNetCDF(x, "aice")) |>
    rbindlist(idcol = "model") |>
    _[, let(lon = NULL, lat = NULL)]

```

```{r fig-mean-extent}
#| fig-cap: Median and 95% coverage sea ice extent for ACCESS-S2 reanalysis (black) and NSIDC (blue).

extent_daily |>  
  copy() |>
  _[aice == 0, aice := NA] |> 
  # _[, daice := c(diff(aice), 0)] |>
  # _[, daice := daice - median(daice, na.rm = TRUE), by = .(model, yday(time))] |>
  # _[abs(daice) > 5e11 & model == "nsidc",
  #   let(daice = NA,
  #       aice = NA)] |>
  # _[, aice := frollmean(aice, 10, align = "center", na.rm = TRUE), by = model] |>
  _[!(month(time) == 2 & mday(time) == 29)] |> 
  _[, time := update(time, year = 2001)] |> 
  _[, median_ci(aice), by = .(time, model)]  |> 
  ggplot(aes(time, mid)) +
  geom_ribbon(data = \(x) x[model == "access"],
    aes(ymin = low, ymax = high, 
                  color = model, fill = model), alpha = 0.3) +
  geom_line(aes(color = model), linewidth = 1) +
#   scale_color_models +
#   scale_fill_models +
  scale_y_continuous("Extent",
                     labels = labels_extent) +
  scale_x_datetime(NULL, date_labels = "%b", date_breaks = "1 month")
```

```{r climatology}
climatologies <- datasets[c("cdr", "access")] |>
    lapply(\(x) climatology(x)  |> cdo_ymonmean()  |> cdo_execute()) |>
    lapply(\(x) ReadNetCDF(x, "aice")) |>
    rbindlist(idcol = "model") |>
    dcast(time + xgrid + ygrid ~ model, value.var = "aice")
```

```{r fig-bias}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: ACCESS-S2 reanalysis sea ice concentration bias compared with NSIDC sea ice concentration.

climatologies |> 
  ggplot(aes(xgrid, ygrid)) +
  geom_contour_fill(aes(z = access - cdr, fill = after_stat(level)), 
                    breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
  scale_fill_divergent_discretised("ACCESS-S2 bias") +
  geomcoord_antarctica +
  geom_antarctica_fill +
  facet_wrap(~ lubridate::month(time, label = TRUE)) +
  wide_legend
```

### Extent

```{r fig-extent-anom}
#| fig-cap: Sea ice extent anomalies for ACCESS-S2 (black) and NSIDC (blue). Straight lines show a piecewise linear fit with nodes at 2005 and 2012.

extent_daily |>
    copy() |>
    _[aice == 0, aice := NA]  |> 
    _[, .(aice = mean(aice)), by = .(model, time = as.Date(round_date(time, "month")))] |>

  _[, aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE), 
    by = .(month(time), model)] |> 
  ggplot(aes(time, aice)) +
  geom_line(aes(color = model)) +
  geom_smooth(aes(color = model), method = "lm",
              formula = y ~ x +
                I(pmax(x - as.numeric(as.Date("2005-12-01")), 0)) +
                I(pmax(x - as.numeric(as.Date("2012-12-01")), 0)),
              n = 100
  ) +
  scale_y_continuous("Extent",
                     labels = labels_extent) +
  scale_x_date(NULL, expand = c(0, 0)) 
#   scale_color_models

```

{>>Add reginoal extents?<<}

## Hindcast

```{r errors}
errors <- rbind(
    readRDS(here::here("data/derived/rmse.Rds")) |>
        _[version != "S1"], 
    readRDS(here::here("data/derived/iiee.Rds"))
    ) |>
  _[value < 0.01, value := NA] |> 
  _[value > 3e13, value := NA] |> 
  _[, time := as.Date(time)] |> 
  _[, lag := as.numeric(time - time_forecast)]

```

```{r fig-rmse}
#| fig-cap: Median and 95% coverage of sea ice concentration anomalies RMSE as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies.

errors |> 
    _[measure == "rmse"] |>
    _[obs_dataset == "cdr"] |>
  _[lag > 0] |> 
  _[, median_ci(value), 
    by = .(lag, version, measure, month(time_forecast), obs_dataset)] |> 
  ggplot(aes(lag, mid)) +
  geom_ribbon(aes(ymin = low, 
                  ymax = high, 
                  color = version, 
                  fill = version,
                  group = interaction(obs_dataset, version)),
                   alpha = 0.1) +
  geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
  labs(y = "RMSE", 
       x = "Lag") +
#   scale_color_models +
#   scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month))
```

{==Regional?==}

-   Mean extent forecast \<\>
    -   Panantarctic
    -   Regional \<\>

### Comparison with S1

```{r fig-iiee}
#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lag for all forecast initialised on the first of each month for ACCESS-S1 and ACCESS-S2 hindcasts.

errors |> 
    _[measure == "iiee"] |>
    _[obs_dataset == "cdr"] |>
  _[lag > 0] |> 
  _[, month := lubridate::month(time_forecast, label = TRUE)]  |> 
  _[!(month == "Jan" & member == "07")] |> 
  _[, median_ci(value), 
    by = .(lag, version, measure, obs_dataset, month(time_forecast))] |> 
  ggplot(aes(lag, mid)) +
  geom_ribbon(aes(ymin = low, 
                  ymax = high, 
                  color = version, 
                  fill = version), alpha = 0.1) +
  geom_line(aes(color = version), linewidth = 1) +
  scale_y_continuous(labels = labels_extent) +
  labs(y = "Integrated Ice Edge Error", 
       x = "Lag") +
#   scale_color_models +
#   scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month))
```

## Conclusions

### 2024-10-03 R/ggplot.R

@@ -1,0 +1,90 @@
+ggplot2::theme_set(ggplot2::theme_minimal() +
+            ggplot2::theme(panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
+                  legend.position = "bottom",
+                  legend.title.position = "top", 
+                  legend.title = ggplot2::element_text(hjust = 0.5),
+                  legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
+                  legend.key.height = grid::unit(0.75, "lines")
+            ))
+wide_legend <- ggplot2::theme(legend.key.width = grid::unit(1, 'null'))
+
+
+colours_models <- c(
+  S2 = "black",
+  S1 = "#a51d2d",
+  forecast = "black",
+  persistence = "#1a5fb4",
+  nsidc = "#1a5fb4"
+)
+
+labels_models <- c(
+  S2 = "ACCESS-S2",
+  S1 = "ACCESS-S1",
+  forecast = "Forecast",
+  persistence = "Persistence",
+  nsidc = "NSIDC CDRV4"
+)
+
+scale_color_models <- ggplot2::scale_color_manual(NULL,
+                                         values = colours_models,
+                                         labels = labels_models
+)
+
+scale_fill_models <- ggplot2::scale_fill_manual(NULL,
+                                         values = colours_models,
+                                         labels = labels_models
+)
+
+labels_extent <- function(x) {
+  m <- which.max(x)
+  x <- scales::label_number(scale = 1e-12)(x)
+  x[m] <- paste0(x[m], "\nM km²")
+  x
+}
+
+labels_month <- setNames(month.abb, 1:12)
+
+
+topo <- rcdo::cdo_topo(
+    grid = CDR_grid(),
+    ofile = here::here("data/derived", "topo.nc")
+) |>
+    rcdo::cdo_execute(options = c("-f nc")) |>
+    metR::ReadNetCDF(c(z = "topo")) |>
+    _[, .(x = xgrid, y = ygrid, z)]
+
+get <- `$`
+
+sic_projection <- CDR() |>
+    rcdo::cdo_seltimestep(1) |>
+    rcdo::cdo_execute() |>
+    ncdf4::nc_open() |>
+    ncdf4::ncatt_get(varid = "crs") |>
+    get("proj_params")
+
+
+contour <- ggplot2::StatContour$compute_group(topo, breaks = 0)
[…]

### 2024-10-03 analysis/paper/access-ice.qmd
Uses pdf

@@
 title: "Untitled"
-format: html
+format: pdf
 execute: 

### 2024-10-04 scripts/hindcast_extent.R
Computes the extent of hindcasts

@@ -1,0 +1,66 @@
+library(furrr)
+workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 25))
+message(workers, " workers")
+
+plan(multisession, workers = workers)
+
+source(here::here("R/datasets.R"))
+
+get_forecast_times <- function(dir) {
+    years <- list.files(file.path(dir, "e01")) |>
+        strcapture(
+            pattern = "di_a?ice_(\\d{4})\\d{4}_e01.nc",
+            proto = list(time = numeric(1))
+        ) |>
+        range()
+
+    seq(as.Date(paste0(years[1], "-01-01")),
+        as.Date(paste0(years[2], "-01-01")),
+        by = "month"
+    )
+}
+
+forecast_times_s2 <- get_forecast_times("/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/")
+
+all_files <- data.table::CJ(
+    forecast_time = forecast_times_s2,
+    member = 1:9
+)
+
+
+
+furrr::future_pmap(all_files, \(forecast_time, member) {
+    S2_hindcast(forecast_time, member) |>
+        extent() |>
+        metR::ReadNetCDF("aice") |>
+        _[, `:=`(
+            lat = NULL, lon = NULL,
+            forecast_time = forecast_time,
+            member = member
+        )] |>
+        _[]
+}) |>
+    data.table::rbindlist() |>
+    saveRDS("data/derived/S2_hindcast_extent.Rds")
+
+
+forecast_times_s1 <- get_forecast_times("/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/")
+
+all_files <- data.table::CJ(
+    forecast_time = forecast_times_s1,
+    member = 1:9
+)
+
+furrr::future_pmap(all_files, \(forecast_time, member) {
+    S1_hindcast(forecast_time, member) |>
+        extent() |>
+        metR::ReadNetCDF("aice") |>
+        _[, `:=`(
+            lat = NULL, lon = NULL,
+            forecast_time = forecast_time,
+            member = member
+        )] |>
+        _[]
+}) |>
+    data.table::rbindlist() |>
+    saveRDS("data/derived/S1_hindcast_extent.Rds")

### 2024-10-04 scripts/pbs/hindcast_extent.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+#PBS -P dx2
+#PBS -q normal
+#PBS -l ncpus=48
+#PBS -l mem=190GB
+#PBS -l walltime=03:00:00
+#PBS -l wd
+#PBS -l jobfs=400GB
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7
+
+# Load module, always specify version number.
+module load R/4.3.1
+module load cdo/2.4.3
+
+# Must include `#PBS -l storage=scratch/ab12+gdata/yz98` if the job
+# needs access to `/scratch/ab12/` and `/g/data/yz98/`. Details on:
+# https://opus.nci.org.au/display/Help/PBS+Directives+Explained
+
+# Run R application
+export PBS_WORKERS=100
+Rscript scripts/hindcast_extent.R > logs/hindcast_extent.log

### 2024-10-04 scripts/pbs/error_metrics.sh
Updates cdo

@@ -11,7 +11,7 @@
 
 # Load module, always specify version number.
 module load R/4.3.1
-module load cdo/2.0.5
+module load cdo/2.4.3
 
 # Must include `#PBS -l storage=scratch/ab12+gdata/yz98` if the job
 # needs access to `/scratch/ab12/` and `/g/data/yz98/`. Details on:

### 2024-10-04 analysis/paper/access-ice.qmd
Adds extent plot

@@
   message: false
+  cache: true
 filters:
@@
   - ../../_extensions/mloubout/critic-markup/critic-markup.lua
+bibliography: references.bib
 ---
@@
 
+[@comiso2023]
+
 Comiso, J.C., Gersten, R.A.
@@
 
-{>>Why is ERA5 better or worse than satellites?<<}
+{\>\>Why is ERA5 better or worse than satellites?\<\<}
 
@@
 ```
+
 ```{r extent_daily}
@@
 
+```{r}
+extent_climatology <- extent_daily |>  
+  copy() |>
+  _[aice == 0, aice := NA] |> 
+  _[!(month(time) == 2 & mday(time) == 29)] |> 
+  _[, time := update(time, year = 2001)] |> 
+  _[, median_ci(aice), by = .(time, model)]
+
+```
+
 ```{r fig-mean-extent}
@@
 
-extent_daily |>  
-  copy() |>
-  _[aice == 0, aice := NA] |> 
-  # _[, daice := c(diff(aice), 0)] |>
-  # _[, daice := daice - median(daice, na.rm = TRUE), by = .(model, yday(time))] |>
-  # _[abs(daice) > 5e11 & model == "nsidc",
-  #   let(daice = NA,
-  #       aice = NA)] |>
-  # _[, aice := frollmean(aice, 10, align = "center", na.rm = TRUE), by = model] |>
-  _[!(month(time) == 2 & mday(time) == 29)] |> 
-  _[, time := update(time, year = 2001)] |> 
-  _[, median_ci(aice), by = .(time, model)]  |> 
+extent_climatology |>
   ggplot(aes(time, mid)) +
@@
 
-{>>Add reginoal extents?<<}
+{\>\>Add reginoal extents?\<\<}
 
 ## Hindcast
+
+```{r hindcast-extent}
+rbind(S1 = readRDS(here::here("data/derived/S1_hindcast_extent.Rds")),
+      S2 = readRDS(here::here("data/derived/S2_hindcast_extent.Rds")), 
+      idcol = "model") |>
+    _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
+                                lag = as.numeric(as.Date(time) - forecast_time),
+                                model)] |>
+    _[, time := forecast_time + lag] |>
+    _[, time2 := update(time, year = 2001)] |>
+    _[, month := lubridate::month(forecast_time, label = TRUE)] |>
+    # _[]
+    ggplot(aes(time2, mid)) +
+    geom_line(data = extent_climatology[model == "cdr"][, model := NULL], 
+              aes(x = as.Date(time))) +    
+    geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +
+    geom_point(data = \(x) x[lag == 1], aes(fill = month), 
+               shape = 21, size = 3) +
+
+    scale_x_date(NULL, date_breaks = "1 month", 
+                 date_labels = "%b", expand = c(0, 0)) +
+    scale_y_continuous(labels = labels_extent) +
+    scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"), 
+                       aesthetics = c("fill", "color"), guide = "none") +
+    facet_wrap(~ model, ncol = 1)
+```
 
@@
 ## Conclusions
+
+# References
+
+::: {#refs}
+:::

### 2024-10-04 R/functions.R
Moves get_forecast_times to functions.R

@@ -12,6 +12,27 @@ ncview <- function(file) {
     system2("ncview", args = c(file), wait = FALSE)
 }
 
+get_forecast_times <- function(model) {
+  
+  dir <- switch(model,
+    S1 =  "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/",
+    S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/",
+    stop("Model needs to be S1 or S2")
+  )
+  
+  years <- list.files(file.path(dir, "e01")) |>
+    strcapture(
+      pattern = "di_a?ice_(\\d{4})\\d{4}_e01.nc",
+      proto = list(time = numeric(1))
+    ) |>
+    range()
+  
+  seq(as.Date(paste0(years[1], "-01-01")),
+      as.Date(paste0(years[2], "-01-01")),
+      by = "month"
+  )
+}
+
 compute_climatology <- function(dataset,
                                 output,
                                 climatology = 1981:2011,

### 2024-10-04 scripts/error_metrics.R

@@ -151,22 +151,9 @@ compute_metrics <- function(forecast_time, member, version, obs_dataset) {
 }
 
 
-get_forecast_times <- function(dir) {
-    years <- list.files(file.path(dir, "e01")) |>
-        strcapture(
-            pattern = "di_a?ice_(\\d{4})\\d{4}_e01.nc",
-            proto = list(time = numeric(1))
-        ) |>
-        range()
+forecast_times_s1 <- get_forecast_times("S1")
 
-    seq(as.Date(paste0(years[1], "-01-01")),
-        as.Date(paste0(years[2], "-01-01")),
-        by = "month"
-    )
-}
-forecast_times_s1 <- get_forecast_times("/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/")
-
-forecast_times_s2 <- get_forecast_times("/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/")
+forecast_times_s2 <- get_forecast_times("S2")
 
 all_files <- data.table::CJ(
     forecast_time = forecast_times_s1,

### 2024-10-04 scripts/hindcast_extent.R

@@ -4,23 +4,12 @@ message(workers, " workers")
 
 plan(multisession, workers = workers)
 
+source(here::here("R/functions.R"))
 source(here::here("R/datasets.R"))
 
-get_forecast_times <- function(dir) {
-    years <- list.files(file.path(dir, "e01")) |>
-        strcapture(
-            pattern = "di_a?ice_(\\d{4})\\d{4}_e01.nc",
-            proto = list(time = numeric(1))
-        ) |>
-        range()
 
-    seq(as.Date(paste0(years[1], "-01-01")),
-        as.Date(paste0(years[2], "-01-01")),
-        by = "month"
-    )
-}
 
-forecast_times_s2 <- get_forecast_times("/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/")
+forecast_times_s2 <- get_forecast_times("S2")
 
 all_files <- data.table::CJ(
     forecast_time = forecast_times_s2,
@@ -44,7 +33,7 @@ furrr::future_pmap(all_files, \(forecast_time, member) {
     saveRDS("data/derived/S2_hindcast_extent.Rds")
 
 
-forecast_times_s1 <- get_forecast_times("/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/")
+forecast_times_s1 <- get_forecast_times("S1")
 
 all_files <- data.table::CJ(
     forecast_time = forecast_times_s1,

### 2024-10-04 R/ggplot.R
Fixes bad align in labesl with unit

@@ -1,10 +1,27 @@
+element_text_first_line <- function(...) {
+  el <- ggplot2::element_text(...)
+  class(el) <- union("element_text_first_line", class(el))
+  el
+}
+
+# Thanks to the wizard Teun van den Brand 
+# https://fosstodon.org/@teunbrand/113248074175175782
+element_grob.element_text_first_line <- 
+  function(element, label, x = NULL, y = NULL, ...) {
+    twolines <- grepl("\n", label)
+    y <- y - grid::unit(twolines * 0.5, "lines")
+    NextMethod()
+  }
+
+
 ggplot2::theme_set(ggplot2::theme_minimal() +
             ggplot2::theme(panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
                   legend.position = "bottom",
                   legend.title.position = "top", 
                   legend.title = ggplot2::element_text(hjust = 0.5),
                   legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
-                  legend.key.height = grid::unit(0.75, "lines")
+                  legend.key.height = grid::unit(0.75, "lines"),
+                  axis.text.y = element_text_first_line()
             ))
 wide_legend <- ggplot2::theme(legend.key.width = grid::unit(1, 'null'))
 
@@ -41,6 +58,7 @@ labels_extent <- function(x) {
   x[m] <- paste0(x[m], "\nM km²")
   x
 }
+
 
 labels_month <- setNames(month.abb, 1:12)
 

### 2024-10-04 analysis/paper/access-ice.qmd
Drafts bias and anomalies

--- 2024-10-03 analysis/paper/access-ice.qmd
@@
   message: false
+  cache: false
 filters:
@@
   - ../../_extensions/mloubout/critic-markup/critic-markup.lua
+  - ../../_extensions/ute/search-replace/search-replace.lua
+search-replace:
+  +S2: ACCESS-S2
+  +S1: ACCESS-S1
+  +CDR: CDR
+bibliography: references.bib
 ---
@@
 library(rcdo)
+library(tagger)
 
@@
 
-ACCESS-S2 is the next seasonal prediction system after ACCESS-S1.
++S2 is the next saeasonal prediction system after ACCESS-S1.
 It uses virtually the same exact model configuration except for some model versions being different.
@@
 While S1 used initial conditions from the UK Met Office, S2 uses the Bureau's own system.
-{== Description of the system ==}{\>\>Atmosphere and ocean.
-Highlight no ice data assimilation\<\<}
+{== Description of the system ==}{>>Atmosphere and ocean.
+Highlight no ice data assimilation.
+Highligh that the hope is that ice will respond to SST/atmosphere DA<<}
 
-For evaluation we use hindcast for the period 1981--2023{\>\>Check\<\<}. Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
+For evaluation we use hindcst for the period 1981--2023{\>\>Check\<\<}. Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
 
@@
 Data coverage began on 01 November 1978 and is ongoing through the most current processing, with updated data processed several times annually.
+
+[@comiso2023]
 
@@
                  bt = BT(),
-                 era5 = ERA5(),
-                 access = S2_reanalysis()
+                 era5 = ERA5()
 )
 ```
+
 ```{r extent_daily}
@@
   lapply(\(x) ReadNetCDF(x, "aice")) |>
-    rbindlist(idcol = "model") |>
+  rbindlist(idcol = "dataset") |>
   _[, let(lon = NULL, lat = NULL)]
+```
+
+```{r}
+extent_climatology <- extent_daily |>  
+  copy() |>
+  _[aice == 0, aice := NA] |> 
+  _[!(month(time) == 2 & mday(time) == 29)] |> 
+  _[, time := update(time, year = 2001)] |> 
+  _[, median_ci(aice), by = .(time, dataset)]
 
@@
 
-```{r fig-mean-extent}
-#| fig-cap: Median and 95% coverage sea ice extent for ACCESS-S2 reanalysis (black) and NSIDC (blue).
+```{r hindcast_extent}
+hindcast_extent <- rbind(S1 = readRDS(here::here("data/derived/S1_hindcast_extent.Rds")),
+                         S2 = readRDS(here::here("data/derived/S2_hindcast_extent.Rds")), 
+                         idcol = "model") |>
+  _[, model := relevel(factor(model), "S2")] |> 
+  _[, lag := as.numeric(as.Date(time) - forecast_time)] |> 
+  _[lag > 0]
+```
+
+```{r fig-hindcast-extent}
+#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the 
+hindcast_extent |> 
+  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
+                              lag = as.numeric(as.Date(time) - forecast_time),
+                              model)] |>
+  _[, time := forecast_time + lag] |>
+  _[, time2 := update(time, year = 2001)] |>
+  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
+  # _[]
+  ggplot(aes(time2, mid)) +
+  geom_line(data = extent_climatology, 
+            aes(x = as.Date(time),
+                group = dataset)) +    
+  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +
+
+  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
+             aes(fill = month), 
+             shape = 24, size = 2.2) +
+
+  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
+             shape = 21, size = 2) +
+
+  scale_x_date(NULL, date_breaks = "1 month", 
+               date_labels = "%b", expand = c(0, 0)) +
+  scale_y_continuous(NULL, labels = labels_extent) +
+  scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"), 
+                     aesthetics = c("fill", "color"), guide = "none") +
+  facet_wrap(~ model, ncol = 1) +
+  tag_facets() 
+```
+
+@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and `r length(datasets)-1` observational datasets. 
++S2 shows a severe low extent bias, especially in the late summer-early autumn. 
+This is due primarily to a faster and longer melt season between January and March and slower growth during March and April 
+This is then balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
+This bias is common in climate models [ref!] and it's been linked to xxxx??
+
+Comparing +S2 with +S1, the latter has a smaller bias, especially at low lags ([Fig.
+@fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig.
+@fig-mean-growth]).
+At larger lags, +S1's bias in summer and autumn is very similar to +S2`s.
+This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1.
+
+At long lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
+Therefore we can estimate the latter using the hindcasts with the largest possible lag, which is shown in triangles in @fig-hindcast-extent for the same dates as the initial conditions. 
+The difference between the two is the effect of the data assimilation.
+
+The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to both having the same model formulation.
+From June to October, in +S2 circles move away from triangles and towards observations, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and affecting the initial conditions.
+The rest of the year, there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.
+
+```{r fig-mean-growth}
+#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations. Values are smoothed with a 2-degree loess smooth with a 30 day window.
+
+dif_hindcast <- hindcast_extent |> 
+  copy() |> 
+  _[order(time)] |> 
+  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(model, forecast_time, member)] |> 
+  _[, median_ci(dif), by = .(model, time = update(time, year = 2000))]
 
 extent_daily |> 
-  copy() |>
-  _[aice == 0, aice := NA] |> 
-  # _[, daice := c(diff(aice), 0)] |>
-  # _[, daice := daice - median(daice, na.rm = TRUE), by = .(model, yday(time))] |>
-  # _[abs(daice) > 5e11 & model == "nsidc",
-  #   let(daice = NA,
-  #       aice = NA)] |>
-  # _[, aice := frollmean(aice, 10, align = "center", na.rm = TRUE), by = model] |>
-  _[!(month(time) == 2 & mday(time) == 29)] |> 
-  _[, time := update(time, year = 2001)] |> 
-  _[, median_ci(aice), by = .(time, model)]  |> 
+  _[order(time)] |> 
+  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
+  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
   ggplot(aes(time, mid)) +
-  geom_ribbon(data = \(x) x[model == "access"],
-    aes(ymin = low, ymax = high, 
-                  color = model, fill = model), alpha = 0.3) +
-  geom_line(aes(color = model), linewidth = 1) +
-#   scale_color_models +
-#   scale_fill_models +
-  scale_y_continuous("Extent",
-                     labels = labels_extent) +
-  scale_x_datetime(NULL, date_labels = "%b", date_breaks = "1 month")
+  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
+  geom_smooth(aes(group = dataset), span = 30/366, linewidth = 0.5, color = "black", se = FALSE, n = 366)  +
+  geom_smooth(data = dif_hindcast, aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
+  scale_x_datetime(NULL, date_breaks = "1 month", 
+                   date_labels = "%b", expand = c(0, 0)) +
+  scale_y_continuous(NULL, labels = labels_extent)
 ```
@@
 ```{r climatology}
-climatologies <- datasets[c("cdr", "access")] |>
-    lapply(\(x) climatology(x)  |> cdo_ymonmean()  |> cdo_execute()) |>
+climatologies <- list(cdr = datasets[["cdr"]],
+                      S2 = S2_reanalysis()) |>
+  lapply(\(x) climatology(x) |>
+           cdo_ymonmean() |> 
+           cdo_execute()
+  ) |>
   lapply(\(x) ReadNetCDF(x, "aice")) |>
@@
 #| fig-height: 6
-#| fig-cap: ACCESS-S2 reanalysis sea ice concentration bias compared with NSIDC sea ice concentration.
+#| fig-cap: +S2 reanalysis sea ice concentration bias compared with NSIDC sea ice concentration.
 
@@
   ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = access - cdr, fill = after_stat(level)), 
+  geom_contour_fill(aes(z = S2 - cdr, fill = after_stat(level)), 
                     breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
@@
   facet_wrap(~ lubridate::month(time, label = TRUE)) +
-  wide_legend
+  wide_legend +
+  tag_facets()
+```
+@fig-bias shows the difference in monthly mean sea ice concentrations between +CDR and +S2 reanalysis. 
+From October to May, the model underestimates sea ice concentrations pretty much everywhere there is ice except for the deep Weddell Sea in April and May, where sea ice concentrations saturate to 1. 
+In winter, the differences are mostly on the sea ice edge, with slight positive bias in XXX and negative bias around the Indian Ocean sector.
+
+### Anomalies
+
+{==intro connecting subsections==}
+
+```{r fig-extent-anom}
+#| fig-cap: Sea ice extent anomalies for +S1 and +S2 (black) and +CDR (blue).
+
+months_difference <- function(x, y) {
+  lubridate::interval(y, x) %/% months(1)
+}
+
+monthly_extent <- extent_daily |> 
+  _[aice == 0, aice := NA]  |> 
+  _[, .(aice = mean(aice)), by = .(dataset, time = as.Date(round_date(time, "month")))] |>
+  _[, aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE), 
+    by = .(month(time), dataset)]
+
+N <- uniqueN(monthly_extent$time)
+
+hindcast_extent |> 
+  copy() |> 
+  # _[model == "S2"] |> 
+  _[, aice := aice - mean(aice[year(time) %between% c(1981, 2011)]), 
+    by = .(lag, model, time = update(time, year = 2000))] |> 
+  _[, .(aice = mean(aice)), by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
+  _[, lag := months_difference(time, forecast_time) + 1] |> 
+  _[lag %in% c(1, 3, 5, 7)] |> 
+  ggplot(aes(time, aice)) +
+  geom_line(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))][dataset == "cdr"], 
+            aes(group = dataset),
+            color = "#1a5fb4") +
+  geom_smooth(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))][dataset == "cdr"], 
+              aes(group = dataset), color = "#1a5fb4", se = FALSE, span = 10*12/N, n = 366) +
+  geom_line()  +
+  geom_smooth(se = FALSE, span = 10*12/N, n = 366, color = "black") +
+  scale_y_continuous(NULL,
+                     labels = labels_extent) +
+  scale_x_date(NULL, expand = c(0, 0)) +
+  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Lag: ", x)))
 ```
 
-### Extent
+@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lags. 
+The anomalies in this case were computed with respect of the climatology of each lag, which is a way of bias-correction.
+Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skillful even at lag 3. 
++S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.
 
-```{r fig-extent-anom}
-#| fig-cap: Sea ice extent anomalies for ACCESS-S2 (black) and NSIDC (blue). Straight lines show a piecewise linear fit with nodes at 2005 and 2012.
-
-extent_daily |>
-    copy() |>
-    _[aice == 0, aice := NA]  |> 
-    _[, .(aice = mean(aice)), by = .(model, time = as.Date(round_date(time, "month")))] |>
-
-  _[, aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE), 
-    by = .(month(time), model)] |> 
-  ggplot(aes(time, aice)) +
-  geom_line(aes(color = model)) +
-  geom_smooth(aes(color = model), method = "lm",
-              formula = y ~ x +
-                I(pmax(x - as.numeric(as.Date("2005-12-01")), 0)) +
-                I(pmax(x - as.numeric(as.Date("2012-12-01")), 0)),
-              n = 100
-  ) +
-  scale_y_continuous("Extent",
-                     labels = labels_extent) +
-  scale_x_date(NULL, expand = c(0, 0)) 
-#   scale_color_models
-
-```
-
-{>>Add reginoal extents?<<}
+{\>\>Add reginoal extents?\<\<}
 

### 2024-10-05 analysis/paper/access-ice.qmd
Adds RMSE and IIEE

@@
 title: "Untitled"
-format: pdf
+format: gfm
 execute: 
@@
 library(tagger)
+library(patchwork)
 
@@
 For evaluation we use hindcst for the period 1981--2023{\>\>Check\<\<}. Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis. 
+Climatology is smoothed with a 11 day running mean.
 
@@
 
-## Hindcast
+### RMSE
+
+To study +S2 forecasts quantitatively, we compute error measures for all hindcasts started on the 1st of every month.
 
@@
   _[, time := as.Date(time)] |> 
-  _[, lag := as.numeric(time - time_forecast)]
+  _[, lag := as.numeric(time - time_forecast)] |> 
+  _[lag > 0] |> 
+  _[!(month(time_forecast) == 1 & member == "07" & version == "S1")]  # This member is wrong
 
@@
   _[obs_dataset == "cdr"] |>
-  _[lag > 0] |> 
   _[, median_ci(value), 
@@
 
-{==Regional?==}
+@fig-rmse shows the median and 95% range of RMSE of sea ice concentration anomalies for +S2 forecasts compared with a benchmark of persistence. 
+This figure uses +CDR data, but the results are nearly identical compared with ERA5 or Bootstrap. 
+Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a referece?==}.
+Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month. 
+The only exception is the RMSE around Febraury forecasted from June onwards.
 
--   Mean extent forecast \<\>
--   Panantarctic
--   Regional \<\>
+{==Add regional?==}
 
@@
   _[obs_dataset == "cdr"] |>
-  _[lag > 0] |> 
   _[, month := lubridate::month(time_forecast, label = TRUE)]  |> 
-  _[!(month == "Jan" & member == "07")] |> 
   _[, median_ci(value), 
@@
   scale_y_continuous(labels = labels_extent) +
-  labs(y = "Integrated Ice Edge Error", 
+  labs(y = NULL, 
        x = "Lag") +
@@
 ```
+
+To compare +S2 with +S1, we computed the IIEE for both models. 
+This error measure is shown in @fig-iiee for all lags and forecasts initialised at the first of every month.
++S1 has lower error at short lags at all months, with the errors converging as the forecast goes on.
+The time to convergence depends on the month and it can be as fas as a few days in July to as large as several months for forecasts initiated in February and March. 
+Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the the memory of sea ice to initial conditions; at least from October to March when the data assimilated form the other components has little to no influence on sea ice.
+
+Also evident in @fig-iiee is the difference in the error spread at short lags between +S2 and +S1. 
+In all month +S1 shows a small error spread at lag 1, indicating that the error in the initial conditions not only is small, but also fairly constant. 
+This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
+For +S2, this is true only only between July and October, approximately. 
+For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it's not constant.
+
+```{r fig-iiee-variance}
+#| fig-cap: Mean spread of IIEE at different lags for different models.
+
+rbind(mean_variance = errors |>
+        _[measure == "iiee"] |>
+        _[obs_dataset == "cdr"] |>
+        _[, sd(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
+        _[, .(value = mean(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
+
+      variance_of_mean = errors |>
+        _[measure == "iiee"] |>
+        _[obs_dataset == "cdr"] |>
+        _[, mean(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
+        _[, .(value = sd(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
+      idcol = "component") |>
+
+  ggplot(aes(lag, value)) +
+  geom_line(aes(color = version, linetype = component), linewidth = 1) +
+  scale_y_continuous("Variance", 
+                     labels = labels_extent) +
+  # scale_color_models +
+  scale_x_continuous("Lag") +
+  facet_wrap(~ month, labeller = labeller(month = labels_month))
+```
+The large initial error spread could be due either to large spread of ensemble members or due to a large spread of individual forecasts. 
+@fig-iiee-variance splits the IIEE variance for each lag into the mean variance of each individual forecast and the variance of the mean error of each individual forecast, which adds up to the total variance.
+The average variance of each forecast is almost identical between forecast systems in all months. 
+This shows that the ensemble spread of individual forecasts evolves identically, which, again, it's not unexpected because both systems share the same model formulation.
+This also shows that the perturbation scheme in +S2 is comparable to the one in +S1.
+
+On the other hand, the spread of the mean error is always larger in +S2 than +S1. 
+The difference is particularly large at short lags in some months, which coindice with the ones in which the data assimlation scheme is not influencing sea ice initial conditions.
+
+{==This figure is pretty hard to read, so it needs tweaking. Maybe splitting it into two figures, or showing the percentage of varinace explained by one of the components==}
 

### 2024-10-07 analysis/paper/access-ice.qmd
Improves variance split figure

@@
 #| fig-cap: Mean spread of IIEE at different lags for different models. 
+#| fig.height: 9
+#| fig.width: 4
+labels_extent_2 <- function(x) {
+  m <- which.max(x)
+  x <- scales::label_number(scale = (1e-12)^2)(x)
+  x[m] <- paste0(x[m], "\nM km⁴")
+  x
+}
 
@@
         _[obs_dataset == "cdr"] |>
-        _[, sd(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
+        _[, var(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
         _[, .(value = mean(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
@@
         _[, mean(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
-        _[, .(value = sd(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
+        _[, .(value = var(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
       idcol = "component") |>
@@
   ggplot(aes(lag, value)) +
-  geom_line(aes(color = version, linetype = component), linewidth = 1) +
+  geom_line(aes(color = version), linewidth = 1) +
   scale_y_continuous("Variance", 
-                     labels = labels_extent) +
+                     labels = labels_extent_2) +
   # scale_color_models +
   scale_x_continuous("Lag") +
-  facet_wrap(~ month, labeller = labeller(month = labels_month))
+  facet_grid(month ~ component, 
+             labeller = labeller(month = labels_month, 
+                                 component = c(mean_variance = "Mean variance",
+                                               variance_of_mean = "Variance of mean"))) 
 ```
+
 The large initial error spread could be due either to large spread of ensemble members or due to a large spread of individual forecasts. 
@@
 On the other hand, the spread of the mean error is always larger in +S2 than +S1. 
-The difference is particularly large at short lags in some months, which coindice with the ones in which the data assimlation scheme is not influencing sea ice initial conditions.
-
-{==This figure is pretty hard to read, so it needs tweaking. Maybe splitting it into two figures, or showing the percentage of varinace explained by one of the components==}
+The difference is particularly large at short lags in some months, which coindice with the ones in which the data assimilation scheme is not influencing sea ice initial conditions.
 

### 2024-10-07 analysis/paper/access-ice.qmd
Sentence wrap

@@
 bibliography: references.bib
+editor: 
+  markdown: 
+    wrap: sentence
 ---
@@
 library(tagger)
-library(patchwork)
 
@@
 
-{==**Antarctic sea ice general description**. Seasonality, trends.==}
+{==**Antarctic sea ice general description**.
+Seasonality, trends.==}
 
-{==**Importance**. Impacts on life. Potential impacts on weather. Impacts on climate. Impacts on ocean circulation.(Justify why it's important to study and understand)==}
+{==**Importance**.
+Impacts on life.
+Potential impacts on weather.
+Impacts on climate.
+Impacts on ocean circulation.(Justify why it's important to study and understand)==}
 
@@
 While S1 used initial conditions from the UK Met Office, S2 uses the Bureau's own system.
-{== Description of the system ==}{>>Atmosphere and ocean.
+{== Description of the system ==}{\>\>Atmosphere and ocean.
 Highlight no ice data assimilation.
-Highligh that the hope is that ice will respond to SST/atmosphere DA<<}
+Highligh that the hope is that ice will respond to SST/atmosphere DA\<\<}
 
-For evaluation we use hindcst for the period 1981--2023{\>\>Check\<\<}. Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis. 
+For evaluation we use hindcst for the period 1981--2023{\>\>Check\<\<}.
+Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
 Climatology is smoothed with a 11 day running mean.
@@
 There is not a lot of data on sea ice properties, especially for things like thickness, age, etc.
-However there are relativelly reliable satellite-derived estimates of sea ice concentration, which esetimates the proportion of each grid area that is covered with ice.
+However there are relatively reliable satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
 These products are not perfect and there still exists a fair bit of observation al uncertainty.
-To account for this uncertanty we use multiple datasets.
+To account for this uncertainty we use multiple datasets.
 
@@
 
-Comiso, J.C., Gersten, R.A.
-2023.
-Bootstrap Sea Ice Concentrations from Nimbus-7 SMMR and DMSP SSM/I-SSMIS, Version 4.
-\[Indicate subset used\].
-Boulder, Colorado USA.
-NASA National Snow and Ice Data Center Distributed Active Archive Center.
-https://doi.org/10.5067/X5LG68MH013O.
-\[Date Accessed\] {\>\>Add to zotero\<\<}
-
 ### NASA Team
@@
 NOAA/NSIDC's Climate Data Record V4 combines the Bootstrap and NASA Team estimates in an attempt to overcome each algorithm's weaknesses.
-{\>\>More detail on how are they combined and their properties.\<\<} As such, it is is not a trully independend dataset.
+{\>\>More detail on how are they combined and their properties.\<\<} As such, it is is not a truly independent dataset.
 
-Meier, W.
-N., F.
-Fetterer, A.
-K.
-Windnagel, and S.
+Meier, W. N., F. Fetterer, A. K. Windnagel, and S.
 Stewart.
@@
 \[Indicate subset used\].
-Boulder, Colorado USA.
-NSIDC: National Snow and Ice Data Center https://doi.org/10.7265/efmz-2t65.
+Boulder, Colorado USA. NSIDC: National Snow and Ice Data Center https://doi.org/10.7265/efmz-2t65.
 \[Date Accessed\] {\>\>Add to zotero\<\<}
@@
 
-{\>\>Why is ERA5 better or worse than satellites?\<\<}
+{\>\>Why is ERA5 better or worse than satellites?\<
+\<}
 
@@
 
-FOr evaluation purposes, we use a series of measures.
+For evaluation purposes, we use a series of measures.
 
@@
 
-Sea Ice Extent is a rough global measure, but a model could have relatively accurate extent of ice but with different distributions. We use two other measures to account for these errors.
+Sea Ice Extent is a rough global measure, but a model could have relatively accurate extent of ice but with different distributions.
+We use two other measures to account for these errors.
 
@@
 
-We also compute the Integrated Ice Edge Error (IIEE) {\>\>CITE\<\<}.
-This is defined as the area in which the model misspredicts the presence of more than 15% ice.
+We also compute the Integrated Ice Edge Error (IIEE) [@goessling2016].
+This is defined as the area in which the model misspredicts sea ice concentration being above or belog 15% ice.
 That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.
@@
 +S2 shows a severe low extent bias, especially in the late summer-early autumn.
-This is due primarily to a faster and longer melt season between January and March and slower growth during March and April 
-This is then balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
-This bias is common in climate models [ref!] and it's been linked to xxxx??
+This is due primarily to a faster and longer melt season between January and March and slower growth during March and April This is then balanced with faster growth between May and July ([Fig.
+@fig-mean-growth]).
+This bias is common in climate models \[ref!\]
+and it's been linked to xxxx??
 
@@
 @fig-mean-growth]).
-At larger lags, +S1's bias in summer and autumn is very similar to +S2`s.
+At larger lags, +S1's bias in summer and autumn is very similar to +S2\`s.
 This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1.
@@
 ```
+
 @fig-bias shows the difference in monthly mean sea ice concentrations between +CDR and +S2 reanalysis.
@@
 
-{\>\>Add reginoal extents?\<\<}
+{\>\>Add reginoal extents?\<
+\<}
 
@@
 This figure uses +CDR data, but the results are nearly identical compared with ERA5 or Bootstrap.
-Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a referece?==}.
+Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a referece?=
+=}.
 Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
@@
 
-{==Add regional?==}
+{==Add regional?=
+=}
 

### 2024-10-07 analysis/paper/access-ice.qmd
Adds cache back

@@
   message: false
-  cache: false
+  cache: true
 filters:

### 2024-10-07 analysis/paper/access-ice.tex
Tries pdf render

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\IfFileExists{pdfcomment.sty}
{
  \usepackage{pdfcomment}
  \newcommand{\criticmarkupadd}[1]{\pdfmarkupcomment[markup=Highlight,color={0.838431  0.913725  0.734902}]{##1}{}}
  \newcommand{\criticmarkuprm}[1]{\pdfmarkupcomment[markup=Highlight,color={0.887059  0.673725  0.673725}]{##1}{}}
  \newcommand{\criticmarkupmark}[1]{\pdfmarkupcomment[markup=Highlight,color={1  0.99216  0.21961}]{##1}{}}
  \newcommand{\criticmarkupcomm}[1]{\pdfcomment[icon=Comment, color={0.89804  0.69020  0}]{##1}}
}{

  \newcommand{\criticmarkupadd}[1]{\{++{##1}++\}}
  \newcommand{\criticmarkuprm}[1]{\{-{}-{##1}-{}-\}}
  \newcommand{\criticmarkupmark}[1]{\{=={##1}==\}}
  \newcommand{\criticmarkupcomm}[1]{\{>{}>{##1}<{}<\}}
}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Untitled},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Untitled}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
balbalbal
\end{abstract}

\section{Introduction}\label{introduction}

\criticmarkupmark{\textbf{Antarctic sea ice general description}.
Seasonality, trends.}

\criticmarkupmark{\textbf{Importance}. Impacts on life. Potential
impacts on weather. Impacts on climate. Impacts on ocean
circulation.(Justify why it's important to study and understand)}

\criticmarkupmark{\textbf{Prediction systems.} Why is it important:
operations \criticmarkupcomm{Discuss with Phil}; science needs (we need
good models). What is available. Discussion on S1 \criticmarkupcomm{Can
we discuss previous S1 results? Laura's paper is not public.}}

\criticmarkupmark{\textbf{Objective.} Evaluate S2.}

\section{Data and methods}\label{data-and-methods}

All datasets are regrided using bilinear interpolation to a common grid,
which is a polar sterographic grid with approximatedly equal area of
25km².

\subsection{Description of the model}\label{description-of-the-model}

ACCESS-S2 is the next saeasonal prediction system after ACCESS-S1. It
uses virtually the same exact model configuration except for some model
versions being different. The main difference are the initial
conditions. While S1 used initial conditions from the UK Met Office, S2
uses the Bureau's own system. \criticmarkupmark{ Description of the
system }\criticmarkupcomm{Atmosphere and ocean. Highlight no ice data
assimilation. Highligh that the hope is that ice will respond to
SST/atmosphere DA}

For evaluation we use hindcst for the period
1981--2023\criticmarkupcomm{Check}. Anomalies will be taken with respect
to the 1981--2011 climatology computed from the reanalysis. Climatology
is smoothed with a 11 day running mean.

\subsection{Verification datasets}\label{verification-datasets}

There is not a lot of data on sea ice properties, especially for things
like thickness, age, etc. However there are relatively reliable
satellite-derived estimates of sea ice concentration, which estimates
the proportion of each grid area that is covered with ice. These
products are not perfect and there still exists a fair bit of
observation al uncertainty. To account for this uncertainty we use
multiple datasets.

\subsubsection{Bootstrap}\label{bootstrap}

\criticmarkupcomm{From https://nsidc.org/data/nsidc-0079/versions/4}
This sea ice concentration data set was derived using measurements from
the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7
satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors
on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and
-F13 satellites. Measurements from the Special Sensor Microwave
Imager/Sounder (SSMIS) aboard DMSP-F17 are also included. The data set
has been generated using the Advanced Microwave Scanning Radiometer -
Earth Observing System (AMSR-E) Bootstrap Algorithm with daily varying
tie-points. Daily (every other day prior to July 1987) and monthly data
are available for both the north and south polar regions. Data are
gridded on the SSM/I polar stereographic grid (25 x 25 km) and provided
in two-byte integer format. Data coverage began on 01 November 1978 and
is ongoing through the most current processing, with updated data
processed several times annually.

(Comiso 2023)

\subsubsection{NASA Team}\label{nasa-team}

\criticmarkupmark{missing}

\subsubsection{CDR}\label{cdr}

NOAA/NSIDC's Climate Data Record V4 combines the Bootstrap and NASA Team
estimates in an attempt to overcome each algorithm's weaknesses.
\criticmarkupcomm{More detail on how are they combined and their
properties.} As such, it is is not a truly independent dataset.

Meier, W. N., F. Fetterer, A. K. Windnagel, and S. Stewart. 2021.
NOAA/NSIDC Climate Data Record of Passive Microwave Sea Ice
Concentration, Version 4. {[}Indicate subset used{]}. Boulder, Colorado
USA. NSIDC: National Snow and Ice Data Center
https://doi.org/10.7265/efmz-2t65. {[}Date Accessed{]}
\criticmarkupcomm{Add to zotero}

\subsubsection{ERA5}\label{era5}

\criticmarkupmark{From the CDS websiste:} ERA5 is the fifth generation
ECMWF reanalysis for the global climate and weather for the past 8
decades. Data is available from 1940 onwards. ERA5 replaces the
ERA-Interim reanalysis.

This parameter is the fraction of a grid box which is covered by sea
ice. Sea ice can only occur in a grid box which includes ocean or inland
water according to the land-sea mask and lake cover, at the resolution
being used. This parameter can be known as sea-ice (area) fraction,
sea-ice concentration and more generally as sea-ice cover. In ERA5,
sea-ice cover is given by two external providers. Before 1979 the
HadISST2 dataset is used. From 1979 to August 2007 the OSI SAF (409a)
dataset is used and from September 2007 the OSI SAF oper dataset is
used. Sea ice is frozen sea water which floats on the surface of the
ocean. Sea ice does not include ice which forms on land such as
glaciers, icebergs and ice-sheets. It also excludes ice shelves which
are anchored on land, but protrude out over the surface of the ocean.
These phenomena are not modelled by the IFS. Long-term monitoring of sea
ice is important for understanding climate change. Sea ice also affects
shipping routes through the polar regions.

\criticmarkupcomm{Why is ERA5 better or worse than
satellites?\textless{} \textless\}

\subsection{Error measures}\label{error-measures}

For evaluation purposes, we use a series of measures.

\subsubsection{Sea ice extent}\label{sea-ice-extent}

Sea Ice Extent is defined as the area of ocean covered with at least
15\% ice. This threshold is motivated by the limitations in satellite
retrieval, which is increasingly unreliable for low sea ice conditions.

Sea Ice Extent is a rough global measure, but a model could have
relatively accurate extent of ice but with different distributions. We
use two other measures to account for these errors.

We compute Root Mean Square Error of sea ice concentration anomalies.

We also compute the Integrated Ice Edge Error (IIEE) (Goessling et al.
2016). This is defined as the area in which the model misspredicts sea
ice concentration being above or belog 15\% ice. That is, dichotomise
sea ice concentration into areas with more and less than 15\% sea ice
both in the forecast and observations; the IIEE is the area in which
forecast and observations differ.

\section{Results and discussion}\label{results-and-discussion}

\subsection{Reanalysis}\label{reanalysis}

\subsubsection{Bias}\label{bias}

\begin{figure}

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}

}

\caption{\label{fig-hindcast-extent}Median sea ice extent for al
hindcasts initialised the first of the month for ACCESS-S2 and ACCESS-S1
in colours representing the start month. In black, the}

\end{figure}%

Figure~\ref{fig-hindcast-extent} a shows the seasonal cycle of Sea Ice
Extent for the ACCESS-S2 hindcast and 2 observational datasets.
ACCESS-S2 shows a severe low extent bias, especially in the late
summer-early autumn. This is due primarily to a faster and longer melt
season between January and March and slower growth during March and
April This is then balanced with faster growth between May and July
(Fig.~\ref{fig-mean-growth}). This bias is common in climate models
{[}ref!{]} and it's been linked to xxxx??

Comparing ACCESS-S2 with ACCESS-S1, the latter has a smaller bias,
especially at low lags (Fig.~\ref{fig-hindcast-extent}) even though the
typical growth rates are very similar between both models
(Fig.~\ref{fig-mean-growth}). At larger lags, ACCESS-S1's bias in summer
and autumn is very similar to ACCESS-S2`s. This suggests that this lower
sea ice state is closer to the models' equilibrium, indicating that it
is an issue with model formulation that was being corrected by the data
assimilation system in ACCESS-S1.

At long lags, sea ice extent loses most of the initial condition memory
and reverts to the model's preferred equilibrium state. Therefore we can
estimate the latter using the hindcasts with the largest possible lag,
which is shown in triangles in Figure~\ref{fig-hindcast-extent} for the
same dates as the initial conditions. The difference between the two is
the effect of the data assimilation.

The equilibrium of ACCESS-S1 and ACCESS-S2 is very similar (comparing
the triangles in each panel in Fig.~\ref{fig-hindcast-extent}), owing to
both having the same model formulation. From June to October, in
ACCESS-S2 circles move away from triangles and towards observations,
indicating that the information from the ocean and atmosphere data
assimilation is getting to sea ice and affecting the initial conditions.
The rest of the year, there is little if any difference between circles
and triangles in ACCESS-S2, indicating that almost no data assimilation
is taking place and the sea ice component of the model is virtually
free-running.

\begin{figure}

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}

}

\caption{\label{fig-mean-growth}Median daily sea ice extent growth of
ACCESS-S1 and ACCESS-S2 hindcasts and observations. Values are smoothed
with a 2-degree loess smooth with a 30 day window.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-bias-1.pdf}

}

\caption{\label{fig-bias}ACCESS-S2 reanalysis sea ice concentration bias
compared with NSIDC sea ice concentration.}

\end{figure}%

Figure~\ref{fig-bias} shows the difference in monthly mean sea ice
concentrations between CDR and ACCESS-S2 reanalysis. From October to
May, the model underestimates sea ice concentrations pretty much
everywhere there is ice except for the deep Weddell Sea in April and
May, where sea ice concentrations saturate to 1. In winter, the
differences are mostly on the sea ice edge, with slight positive bias in
XXX and negative bias around the Indian Ocean sector.

\subsubsection{Anomalies}\label{anomalies}

\criticmarkupmark{intro connecting subsections}

\begin{figure}

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}

}

\caption{\label{fig-extent-anom}Sea ice extent anomalies for ACCESS-S1
and ACCESS-S2 (black) and CDR (blue).}

\end{figure}%

Figure~\ref{fig-extent-anom} shows monthly sea ice extent anomalies
forecasted at selected lags. The anomalies in this case were computed
with respect of the climatology of each lag, which is a way of
bias-correction. Compared with ACCESS-S1, ACCESS-S2 anomaly forecast is
relatively poor even in the first month, which stays relatively skillful
even at lag 3. ACCESS-S2 shows much bigger variability than
observations, with dramatic lows between 1995 and 2007 and highs between
2007 and 2015.

\criticmarkupcomm{Add reginoal extents?\textless{} \textless\}

\subsubsection{RMSE}\label{rmse}

To study ACCESS-S2 forecasts quantitatively, we compute error measures
for all hindcasts started on the 1st of every month.

\begin{figure}

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-rmse-1.pdf}

}

\caption{\label{fig-rmse}Median and 95\% coverage of sea ice
concentration anomalies RMSE as a function of forecast lag for all
forecast initialised on the first of each month compared with a
reference forecast of persistence of anomalies.}

\end{figure}%

Figure~\ref{fig-rmse} shows the median and 95\% range of RMSE of sea ice
concentration anomalies for ACCESS-S2 forecasts compared with a
benchmark of persistence. This figure uses CDR data, but the results are
nearly identical compared with ERA5 or Bootstrap. Due to errors in the
initial conditions, it is expected that a persistence forecast would be
better than the model forecast at very short lags, but that the
persistence forecast errors would grow faster and eventually surpass the
mode forecast, at which time is statistically useful
\criticmarkupmark{I've got this from CC at the ICTP summer school and
makes sense, but it would be great to have a referece?= =\}. Here the
persistence errors are almost always lower than the ACCESS-S2 forecast,
indicating that the model doesn't have skill at any lag and in any
month. The only exception is the RMSE around Febraury forecasted from
June onwards.

\criticmarkupmark{Add regional?= =\}

\subsubsection{Comparison with S1}\label{comparison-with-s1}

\begin{figure}

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-iiee-1.pdf}

}

\caption{\label{fig-iiee}Median and 95\% coverage of Integrated Ice Edge
Error as a function of forecast lag for all forecast initialised on the
first of each month for ACCESS-S1 and ACCESS-S2 hindcasts.}

\end{figure}%

To compare ACCESS-S2 with ACCESS-S1, we computed the IIEE for both
models. This error measure is shown in Figure~\ref{fig-iiee} for all
lags and forecasts initialised at the first of every month. ACCESS-S1
has lower error at short lags at all months, with the errors converging
as the forecast goes on. The time to convergence depends on the month
and it can be as fas as a few days in July to as large as several months
for forecasts initiated in February and March. Since the only difference
between these forecasts are the initial conditions, this timescale is an
indication of the the memory of sea ice to initial conditions; at least
from October to March when the data assimilated form the other
components has little to no influence on sea ice.

Also evident in Figure~\ref{fig-iiee} is the difference in the error
spread at short lags between ACCESS-S2 and ACCESS-S1. In all month
ACCESS-S1 shows a small error spread at lag 1, indicating that the error
in the initial conditions not only is small, but also fairly constant.
This spread then grows towards a climatological spread as errors
accumulate differently in different forecasts. For ACCESS-S2, this is
true only only between July and October, approximately. For all other
months, the error spread is more or less stable throughout the forecast
window, indicating that not only the initial error is high, but it's not
constant.

\begin{figure}

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-iiee-variance-1.pdf}

}

\caption{\label{fig-iiee-variance}Mean spread of IIEE at different lags
for different models.}

\end{figure}%

The large initial error spread could be due either to large spread of
ensemble members or due to a large spread of individual forecasts.
Figure~\ref{fig-iiee-variance} splits the IIEE variance for each lag
into the mean variance of each individual forecast and the variance of
the mean error of each individual forecast, which adds up to the total
variance. The average variance of each forecast is almost identical
between forecast systems in all months. This shows that the ensemble
spread of individual forecasts evolves identically, which, again, it's
not unexpected because both systems share the same model formulation.
This also shows that the perturbation scheme in ACCESS-S2 is comparable
to the one in ACCESS-S1.

On the other hand, the spread of the mean error is always larger in
ACCESS-S2 than ACCESS-S1. The difference is particularly large at short
lags in some months, which coindice with the ones in which the data
assimilation scheme is not influencing sea ice initial conditions.

\subsection{Conclusions}\label{conclusions}

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-comiso2023}
Comiso, J. 2023. {``Bootstrap {Sea Ice Concentrations} from {Nimbus-7
SMMR} and {DMSP SSM}/{I-SSMIS}. ({NSIDC-0079}, {Version} 4).''} Boulder,
Colorado USA.: {NASA National Snow and Ice Data Center Distributed
Active Archive Center.} \url{https://doi.org/10.5067/X5LG68MH013O}.

\bibitem[\citeproctext]{ref-goessling2016}
Goessling, H. F., S. Tietsche, J. J. Day, E. Hawkins, and T. Jung. 2016.
{``Predictability of the {Arctic} Sea Ice Edge.''} \emph{Geophysical
Research Letters} 43 (4): 1642--50.
\url{https://doi.org/10.1002/2015GL067232}.

\end{CSLReferences}

\end{document}

### 2024-10-07 R/ggplot.R
Adds colours for datasets

@@ -25,21 +25,20 @@ ggplot2::theme_set(ggplot2::theme_minimal() +
             ))
 wide_legend <- ggplot2::theme(legend.key.width = grid::unit(1, 'null'))
 
-
 colours_models <- c(
   S2 = "black",
   S1 = "#a51d2d",
-  forecast = "black",
-  persistence = "#1a5fb4",
-  nsidc = "#1a5fb4"
+  cdr = "#1a5fb4",
+  bt = "#1a5fb4",
+  era5 = "#1a5fb4"
 )
 
 labels_models <- c(
   S2 = "ACCESS-S2",
   S1 = "ACCESS-S1",
-  forecast = "Forecast",
-  persistence = "Persistence",
-  nsidc = "NSIDC CDRV4"
+  cdr = "CDR",
+  bt = "Bootstrap",
+  era5 = "ERA5"
 )
 
 scale_color_models <- ggplot2::scale_color_manual(NULL,
@@ -94,15 +93,3 @@ geomcoord_antarctica <- list(
     geom_antarctica_path
 )
 
-scale_color_models <- ggplot2::scale_color_manual(NULL,
-    values = c(
-        access = "black",
-        hadisst = "#3584e4",
-        nsidc = "#e66100"
-    ),
-    labels = c(
-        access = "ACCESS-S2",
-        hadisst = "HadISST",
-        nsidc = "NSIDC CDRV4"
-    )
-)

### 2024-10-07 analysis/paper/access-ice.qmd
Adds plots for era5

@@
 )
+which_dataset <- "cdr"
 ```
@@
   ggplot(aes(time2, mid)) +
-  geom_line(data = extent_climatology, 
+  geom_line(data = extent_climatology[dataset == which_dataset], 
             aes(x = as.Date(time),
@@
   _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
+  _[dataset == which_dataset] |> 
   ggplot(aes(time, mid)) +
   geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
-  geom_smooth(aes(group = dataset), span = 30/366, linewidth = 0.5, color = "black", se = FALSE, n = 366)  +
-  geom_smooth(data = dif_hindcast, aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
+  geom_smooth(aes(group = dataset, color = dataset),
+              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
+  geom_smooth(data = dif_hindcast,
+              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
   scale_x_datetime(NULL, date_breaks = "1 month", 
                    date_labels = "%b", expand = c(0, 0)) +
-  scale_y_continuous(NULL, labels = labels_extent)
+  scale_y_continuous(NULL, labels = labels_extent) +
+  scale_color_models
+
 ```
@@
 ```{r climatology}
-climatologies <- list(cdr = datasets[["cdr"]],
+climatologies <- c(datasets,
                    S2 = S2_reanalysis()) |>
@@
   lapply(\(x) ReadNetCDF(x, "aice")) |>
-  rbindlist(idcol = "model") |>
-  dcast(time + xgrid + ygrid ~ model, value.var = "aice")
+  rbindlist(idcol = "model") 
 ```
@@
 climatologies |> 
+  _[model %in% c("S2", which_dataset)] |> 
+  dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
+  setnames(which_dataset, "obs") |> 
   ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = S2 - cdr, fill = after_stat(level)), 
+  geom_contour_fill(aes(z = S2 - obs, fill = after_stat(level)), 
                     breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
@@
   ggplot(aes(time, aice)) +
-  geom_line(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))][dataset == "cdr"], 
-            aes(group = dataset),
-            color = "#1a5fb4") +
-  geom_smooth(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))][dataset == "cdr"], 
-              aes(group = dataset), color = "#1a5fb4", se = FALSE, span = 10*12/N, n = 366) +
-  geom_line()  +
-  geom_smooth(se = FALSE, span = 10*12/N, n = 366, color = "black") +
+  geom_line(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
+              _[dataset == which_dataset],
+            aes(group = dataset, color = dataset)) +
+  geom_smooth(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
+                _[dataset == "cdr"], 
+              aes(group = dataset, color = dataset),
+              se = FALSE, span = 10*12/N, n = 366) +
+  geom_line(aes(color = model))  +
+  geom_smooth(aes(color = model), se = FALSE, span = 10*12/N, n = 366) +
   scale_y_continuous(NULL,
@@
   scale_x_date(NULL, expand = c(0, 0)) +
-  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Lag: ", x)))
+  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Lag: ", x))) +
+  scale_color_models
 ```
@@
   _[measure == "rmse"] |>
-  _[obs_dataset == "cdr"] |>
+  _[obs_dataset == which_dataset] |>
   _[, median_ci(value), 
     by = .(lag, version, measure, month(time_forecast), obs_dataset)] |> 
+  _[version == "persistence", version := obs_dataset] |> 
   ggplot(aes(lag, mid)) +
@@
   #   scale_fill_models +
-  facet_wrap(~ month, labeller = labeller(month = labels_month))
+  facet_wrap(~ month, labeller = labeller(month = labels_month)) +
+  scale_color_models +
+  scale_fill_models
 ```
@@
   _[measure == "iiee"] |>
-  _[obs_dataset == "cdr"] |>
+  _[obs_dataset == which_dataset] |>
   _[, month := lubridate::month(time_forecast, label = TRUE)]  |> 
@@
   scale_y_continuous(labels = labels_extent) +
+  scale_color_models +
+  scale_fill_models +
   labs(y = NULL, 
@@
         _[measure == "iiee"] |>
-        _[obs_dataset == "cdr"] |>
+        _[obs_dataset == which_dataset] |>
         _[, var(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
@@
         _[measure == "iiee"] |>
-        _[obs_dataset == "cdr"] |>
+        _[obs_dataset == which_dataset] |>
         _[, mean(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
@@
   scale_x_continuous("Lag") +
+  scale_color_models +
   facet_grid(month ~ component, 
@@
 :::
+
+```{r dataset_era}
+which_dataset <- "era5"
+```
+
+```{r}
+<<fig-hindcast-extent>>
+```
+
+```{r}
+<<fig-mean-growth>>
+```
+
+```{r}
+<<fig-bias>>
+```
+
+```{r}
+<<<fig-extent-anom>>
+```
+
+```{r}
+<<fig-rmse>>
+```
+
+```{r}
+<<fig-iiee>>
+```
+
+```{r}
+<<fig-iiee-variance>>
+```

### 2024-10-07 analysis/paper/access-ice.qmd
Adds better lag indicators

@@
   scale_color_models
-
 ```
@@
                     breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
-  scale_fill_divergent_discretised("ACCESS-S2 bias") +
+  scale_fill_divergent_discretised(paste0("Bias with respect to ", labels_models[which_dataset])) +
   geomcoord_antarctica +
@@
             aes(group = dataset, color = dataset)) +
-  geom_smooth(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
-                _[dataset == "cdr"], 
-              aes(group = dataset, color = dataset),
-              se = FALSE, span = 10*12/N, n = 366) +
+  # geom_smooth(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
+  #               _[dataset == "cdr"], 
+  #             aes(group = dataset, color = dataset),
+  #             se = FALSE, span = 10*12/N, n = 366) +
   geom_line(aes(color = model))  +
-  geom_smooth(aes(color = model), se = FALSE, span = 10*12/N, n = 366) +
+  # geom_smooth(aes(color = model), se = FALSE, span = 10*12/N, n = 366) +
   scale_y_continuous(NULL,
@@
 The anomalies in this case were computed with respect of the climatology of each lag, which is a way of bias-correction.
-Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skillful even at lag 3.
+Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at lag 3.
 +S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.
 
-{\>\>Add reginoal extents?\<
+{\>\>Add regional extents?\<
 \<}
@@
 #| fig-cap: Median and 95% coverage of sea ice concentration anomalies RMSE as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies.
+#| fig.height: 5
+#| fig.width: 9
+
+month_from_lag <- function(day = 1, step = 2) {
+  force(day)
+  force(step)
+  function(x) {
+    unique(x[, .(lag, month)])[, time := make_date(month = month, day = 1) + lag] |> 
+      _[mday(time) == day, .(lag, month, 
+                             month2 = lubridate::month(time, label = TRUE)
+      )] |> 
+      _[, .SD[((seq_len(.N) + 1) %% step) == 0], by = month]
+  }
+}
 
@@
   ggplot(aes(lag, mid)) +
+
+  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
+             colour = "gray50", linewidth = 0.1) +
+
+  geom_text(data = month_from_lag(15, 1), aes(y = Inf, label = month2), 
+            size = 2.5,
+            vjust = 1) +
+
+  geom_text(data = month_from_lag(1, 1), aes(y = -Inf, label = lag), 
+            size = 2.5,
+            vjust = 0) +
+
   geom_ribbon(aes(ymin = low, 
@@
   geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
+  scale_x_continuous(breaks = NULL) +
   labs(y = "RMSE", 
        x = "Lag") +
+  scale_color_models +
+  scale_fill_models +
   #   scale_color_models +
   #   scale_fill_models +
-  facet_wrap(~ month, labeller = labeller(month = labels_month)) +
-  scale_color_models +
-  scale_fill_models
+  facet_wrap(~ month, labeller = labeller(month = labels_month)) 
 ```
@@
   ggplot(aes(lag, mid)) +
+
+  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
+             colour = "gray50", linewidth = 0.1) +
+
+  geom_text(data = month_from_lag(15, 2), aes(y = Inf, label = month2), 
+            size = 2.5,
+            vjust = 1) +
+
+  geom_text(data = month_from_lag(1, 2), aes(y = -Inf, label = lag), 
+            size = 2.5,
+            vjust = 0) +
+
   geom_ribbon(aes(ymin = low, 
@@
   scale_y_continuous(labels = labels_extent) +
+  scale_x_continuous(breaks = NULL) +
   scale_color_models +
@@
 
+# Suplementary figures
+
+## ERA5
+
 ```{r dataset_era}

### 2024-10-07 analysis/paper/access-ice.qmd
Re-renders without cache

@@
   message: false
-  cache: true
+  cache: false
 filters:
@@
   ggplot(aes(lag, value)) +
+
+  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
+             colour = "gray50", linewidth = 0.1) +
+
+  geom_text(data = month_from_lag(15, 2), aes(y = Inf, label = month2), 
+            size = 2.5,
+            vjust = 1) +
+
+  geom_text(data = month_from_lag(1, 2), aes(y = -Inf, label = lag), 
+            size = 2.5,
+            vjust = 0) +
+
   geom_line(aes(color = version), linewidth = 1) +
@@
   # scale_color_models +
-  scale_x_continuous("Lag") +
+  scale_x_continuous("Lag", breaks = NULL) +
   scale_color_models +

### 2024-10-11 R/datasets.R
Adds OSI

@@ -1,294 +1,371 @@
 source(here::here("R/functions.R"))
 
-cdo_apply <- function(ifile, operation, ofile = NULL) {
-    rcdo::cdo_operator("apply", params = c("operation"), Inf, 1) |>
-        rcdo::cdo(
-            input = list(ifile),
-            params = list(operation = operation),
-            output = ofile
-        )
+cdo_apply <- function(ifile, op, ofile = NULL) {
+  rcdo::cdo_operator("apply", params = c("op"), Inf, 1) |>
+    rcdo::cdo(
+      input = list(ifile),
+      params = list(op = op),
+      output = ofile
+    )
 }
 
 CDR_grid <- function() {
-    here::here("data/raw/nsidc_grid.txt")
+  here::here("data/raw/nsidc_grid.txt")
 }
 
 CDR <- function() {
-    file <- here::here("data/raw/cdr.nc")
-    years <- 1981:2023
-
-    if (file.exists(file)) {
-        return(file)
-    }
-
-    urls <- paste0(
-        "https://polarwatch.noaa.gov/erddap/griddap/nsidcG02202v4sh1day.nc?cdr_seaice_conc[(",
-        years,
-        "-01-01T00:00:00Z):1:(",
-        years,
-        "-12-31T00:00:00Z)][(4350000.0):1:(-3950000.0)][(-3950000.0):1:(3950000.0)]"
-    )
-
-    files <- replicate(length(urls), tempfile())
-    res <- curl::multi_download(urls, files, resume = TRUE)
-
-    files |>
-        rcdo::cdo_mergetime() |>
-        cdo_chname("cdr_seaice_conc", "aice") |>
-        cdo_del29feb() |>
-        rcdo::cdo_setgrid(CDR_grid()) |>
-        rcdo::cdo_execute(output = file, options = c("-L", "-O"))
-
-    remove(files)
+  file <- here::here("data/raw/cdr.nc")
+  years <- 1981:2023
+  
+  if (file.exists(file)) {
     return(file)
+  }
+  
+  urls <- paste0(
+    "https://polarwatch.noaa.gov/erddap/griddap/nsidcG02202v4sh1day.nc?cdr_seaice_conc[(",
+    years,
+    "-01-01T00:00:00Z):1:(",
+    years,
+    "-12-31T00:00:00Z)][(4350000.0):1:(-3950000.0)][(-3950000.0):1:(3950000.0)]"
+  )
+  
[…]

### 2024-10-11 R/functions.R

@@ -77,6 +77,20 @@ climatology <- function(dataset) {
     return(file)
 }
 
+
+anomalies <- function(dataset, climatology) {
+  file  <- here::here("data/derived/anomalies", name_from_dataset(dataset))
+  dir.create(dirname(file), FALSE, TRUE)
+
+  if (file.exists(file)) {
+    return(file)
+  }
+  cdo_ydaysub(cdo_del29feb(dataset), cdo_del29feb(climatology)) |>
+    cdo_execute(output = file, options = "-L")
+}
+
+
+
 name_from_dataset <- function(dataset) {
     infohash <- digest::digest(file.info(dataset)[c("size", "mtime")])
     paste0(tools::file_path_sans_ext(basename(dataset)), "_", infohash, ".nc")

### 2024-10-11 scripts/error_metrics.R

@@ -20,19 +20,17 @@ obs <- list(
         bootstrap = BT() |>
             force(),
         era5 = ERA5() |>
+            force(),
+        osi = OSI() |>
             force()
     )
 )
 
-climatologies <- list(
-    cdr = CDR(),
-    bootstrap = BT(),
-    era5 = ERA5()
-) |>
+climatologies <- obs$data |>
     furrr::future_map(climatology)
 
 obs$anomalies <- furrr::future_map(names(climatologies), \(n) {
-    get_anomalies(obs$data[[n]], climatologies[[n]])
+    anomalies(obs$data[[n]], climatologies[[n]])
 }) |>
     setNames(names(climatologies))
 
@@ -68,8 +66,9 @@ cdo_persist <- function(file, file2, init_time) {
 
 
 get_part <- memoise::memoise(function(file, first_time, last_time) {
-    outfile <- tempfile(tmpdir = "data/temp/part/")
-    dir.create("data/temp/part", showWarnings = FALSE, recursive = TRUE)
+    outfile <- tempfile(tmpdir = here::here("data/temp/part", 
+                                            tools::file_path_sans_ext(basename(file))))
+    dir.create(dirname(outfile), showWarnings = FALSE, recursive = TRUE)
 
     file |>
         cdo_seldate(
@@ -77,7 +76,7 @@ get_part <- memoise::memoise(function(file, first_time, last_time) {
             enddate = as.character(last_time)
         ) |>
         cdo_execute(options = "-L", output = outfile)
-}, cache = cachem::cache_disk("data/temp/cache"))
+}, cache = cachem::cache_disk(here::here("data/temp/cache")))
 
 compute_metrics <- function(forecast_time, member, version, obs_dataset) {
     # message(i)
@@ -117,12 +116,14 @@ compute_metrics <- function(forecast_time, member, version, obs_dataset) {
     observation_part_anomaly <- get_part(obs$anomalies[[obs_dataset]], first_time, last_time)
 
     nc <- ncdf4::nc_open(observation_part)
-    times_obs <- metR:::.parse_time(time = nc$dim$time$vals, units = nc$dim$time$units, calendar = nc$dim$time$calendar)
+    times_obs <- metR:::.parse_time(time = nc$dim$time$vals, 
+                                    units = nc$dim$time$units, 
+                                    calendar = nc$dim$time$calendar)
     ncdf4::nc_close(nc)
 
     if (!file.exists(rmse_out)) {
         file |>
[…]

### 2024-10-11 analysis/paper/access-ice.qmd
Writes S2+S1 info

@@
 
-All datasets are regrided using bilinear interpolation to a common grid, which is a polar sterographic grid with approximatedly equal area of 25km².
+## +S2
 
-## Description of the model
++S2 [@wedd2022] is the Australian Bureau of Meteorology's current seasonal prediction system that replaced +S1 [@hudson2017] in October 2021. 
+Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0  [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
+The atmosphere has a N216 horizontal resolution(~60km in the mid-latitudes) with 85 levels. 
+The land model uses the same horizontal grid with 4 soil levels. 
+The ocean component has a 1/4º resolution with 75 vertical levels. 
+The ice component --based on CICE version 4.1-- has the same resolution than the ocean and 5 sea ice categories as well as an open water category.
 
-+S2 is the next saeasonal prediction system after ACCESS-S1.
-It uses virtually the same exact model configuration except for some model versions being different.
-The main difference are the initial conditions.
-While S1 used initial conditions from the UK Met Office, S2 uses the Bureau's own system.
-{== Description of the system ==}{\>\>Atmosphere and ocean.
-Highlight no ice data assimilation.
-Highligh that the hope is that ice will respond to SST/atmosphere DA\<\<}
+Both systems took atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts. 
+The main difference between the two are the ocean and ice initial conditions. 
++S1 initial conditions came from from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) DA scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
++S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022. 
+This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
+SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are larger than 0ºC.
+Relevant for this work, sea ice concentrations are not assimilated.
 
-For evaluation we use hindcst for the period 1981--2023{\>\>Check\<\<}.
+For evaluation we use hindcast for the period 1981--2023{\>\>Check\<\<}.
 Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
@@
 However there are relatively reliable satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
-These products are not perfect and there still exists a fair bit of observation al uncertainty.
-To account for this uncertainty we use multiple datasets.
+
+We use NOAA/NSIDC's Climate Data Record V4 as the primary verification dataset. 
+Its sea ice concentration data are estimated taking the maximum value of the NASA Team [ref] and NASA Bootstrap [ref] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
+Both algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
+The data has a resolution of 25 by 25km and daily from 1978 onwards.
+
+These data are not perfect, there still exists a fair bit of observational uncertainty.
+We include other sources of data to take into acount the observational uncertainty.
+
+However, we find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
+Nevertheless, we provide figures using ERA5
+
+The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [ref] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR.
+We also use
+
+Copernicus Climate Change Service (C3S) (2020): Sea ice concentration daily gridded data from 1978 to present derived from satellite observations.
+Copernicus Climate Change Service (C3S) Climate Data Store (CDS).
+DOI: 10.24381/cds.3cd8b812 (Accessed on DD-MMM-YYYY)
+
+{==From the CDS websiste:==} ERA5 is the fifth generation ECMWF reanalysis for the global climate and weather for the past 8 decades.
+Data is available from 1940 onwards.
+ERA5 replaces the ERA-Interim reanalysis.
+
+This parameter is the fraction of a grid box which is covered by sea ice.
+Sea ice can only occur in a grid box which includes ocean or inland water according to the land-sea mask and lake cover, at the resolution being used.
+This parameter can be known as sea-ice (area) fraction, sea-ice concentration and more generally as sea-ice cover.
+In ERA5, sea-ice cover is given by two external providers.
+Before 1979 the HadISST2 dataset is used.
+From 1979 to August 2007 the OSI SAF (409a) dataset is used and from September 2007 the OSI SAF oper dataset is used.
+Sea ice is frozen sea water which floats on the surface of the ocean.
+Sea ice does not include ice which forms on land such as glaciers, icebergs and ice-sheets.
+It also excludes ice shelves which are anchored on land, but protrude out over the surface of the ocean.
+These phenomena are not modelled by the IFS.
+Long-term monitoring of sea ice is important for understanding climate change.
+Sea ice also affects shipping routes through the polar regions.
+
+{\>\>Why is ERA5 better or worse than satellites?\<
+\<}
 
@@
 
-{==From the CDS websiste:==} ERA5 is the fifth generation ECMWF reanalysis for the global climate and weather for the past 8 decades.
-Data is available from 1940 onwards.
-ERA5 replaces the ERA-Interim reanalysis.
-
-This parameter is the fraction of a grid box which is covered by sea ice.
-Sea ice can only occur in a grid box which includes ocean or inland water according to the land-sea mask and lake cover, at the resolution being used.
-This parameter can be known as sea-ice (area) fraction, sea-ice concentration and more generally as sea-ice cover.
-In ERA5, sea-ice cover is given by two external providers.
-Before 1979 the HadISST2 dataset is used.
-From 1979 to August 2007 the OSI SAF (409a) dataset is used and from September 2007 the OSI SAF oper dataset is used.
-Sea ice is frozen sea water which floats on the surface of the ocean.
-Sea ice does not include ice which forms on land such as glaciers, icebergs and ice-sheets.
-It also excludes ice shelves which are anchored on land, but protrude out over the surface of the ocean.
-These phenomena are not modelled by the IFS.
-Long-term monitoring of sea ice is important for understanding climate change.
-Sea ice also affects shipping routes through the polar regions.
-
-{\>\>Why is ERA5 better or worse than satellites?\<
-\<}
+All datasets are regrided using bilinear interpolation to a common grid, which is a polar stereographic grid with approximately equal area of 25km².
 
@@
                  bt = BT(),
+                 osi = OSI(),
                  era5 = ERA5()
@@
 
-## ERA5
+<!-- ## ERA5 -->
 
-```{r dataset_era}
-which_dataset <- "era5"
-```
+<!-- ```{r dataset_era} -->
+<!-- which_dataset <- "era5" -->
+<!-- ``` -->
 
-```{r}
-<<fig-hindcast-extent>>
-```
+<!-- ```{r fig-hindcast-extent-era5} -->
+<!-- which_dataset <- "era5" -->
+<!-- <<fig-hindcast-extent>> -->
+<!-- ``` -->
 
-```{r}
-<<fig-mean-growth>>
-```
+<!-- ```{r fig-hindcast-extent-bt} -->
+<!-- which_dataset <- "bt" -->
+<!-- <<fig-hindcast-extent>> -->
+<!-- ``` -->
 
-```{r}
-<<fig-bias>>
-```
+<!-- ```{r fig-mean-growth} -->
+<!-- <<fig-mean-growth>> -->
+<!-- ``` -->
 
-```{r}
-<<<fig-extent-anom>>
-```
+<!-- ```{r} -->
+<!-- <<fig-bias>> -->
+<!-- ``` -->
 
-```{r}
-<<fig-rmse>>
-```
+<!-- ```{r} -->
+<!-- <<<fig-extent-anom>> -->
+<!-- ``` -->
 
-```{r}
-<<fig-iiee>>
-```
+<!-- ```{r} -->
+<!-- <<fig-rmse>> -->
+<!-- ``` -->
 
-```{r}
-<<fig-iiee-variance>>
-```
+<!-- ```{r} -->
+<!-- <<fig-iiee>> -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- <<fig-iiee-variance>> -->
+<!-- ``` -->

### 2024-10-11 R/ggplot.R
Renders

@@ -30,7 +30,8 @@ colours_models <- c(
   S1 = "#a51d2d",
   cdr = "#1a5fb4",
   bt = "#1a5fb4",
-  era5 = "#1a5fb4"
+  era5 = "#1a5fb4",
+  osi = "#1a5fb4"
 )
 
 labels_models <- c(
@@ -38,7 +39,8 @@ labels_models <- c(
   S1 = "ACCESS-S1",
   cdr = "CDR",
   bt = "Bootstrap",
-  era5 = "ERA5"
+  era5 = "ERA5",
+  osi = "OSI"
 )
 
 scale_color_models <- ggplot2::scale_color_manual(NULL,

### 2024-10-17 analysis/paper/access-ice.qmd
Adds introdcuction

@@
   +S1: ACCESS-S1
-  +CDR: CDR
+  +CDR: NOAA/NSIDC CDRV4
 bibliography: references.bib
@@
 
+.....
+
+Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice had a slightly increasing trend that models systematically failed to reproduce and puzzled researchers \cite{hobbs2016}.
+Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change.
+[ref: that CESM run with forced winds].
+
 {==**Importance**.
@@
 
-{==**Prediction systems.** Why is it important: operations {\>\>Discuss with Phil\<\<}; science needs (we need good models).
+Antarctic sea ice is in a remote location but the potential impacts of its variability and trends extend far into the lower latitudes, affecting ocean and atmospheric circulation, ocean carbon uptake and biological processes.
+It's likely that a reduction in sea ice extent would reduce the temperature gradient between high and low latitudes (and vice versa), which in turn influences the strength and location of the jet, which is a crucial component of weather events.
+For the northern hemisphere, it has been hypothesised that the reduction in sea ice in the northern hemisphere has lead to a weaker, wavier jet, increasing the frequency of extreme events [refs].
+Less is known about the effects of Antarctic sea ice variability {==is this true?
+It feels that way from the (lack of) literature.
+Is there a reference to substantiate this claim?==} on weather timescales.
+On climate scales, it is thought that the signal of a reduction in Antarctic sea ice is like a "mini global warming" and experiments also predict a weakening and northerly shift of the jet [@ayres2022; @england2020].
+
+{==Relevant papers to read:
+
+\cite{ayres2019}
+\cite{england2018}
+\cite{raphael2011}
+\cite{raphael2003a}
+
+==}
+
+{==**Prediction systems.** Why is it important: operations {\>\>Discuss with Phil\<\<}; science needs (we need good models).\cite{desilva2020}\cite{wagner2020}
 What is available.
@@
 Laura's paper is not public.\<\<}==}
+
+Correctly modelling Antarctic sea ice is not only necessary for process understanding and climate projections to inform adaptation strategies, but accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent [@desilva2020 @wagner2020].
+However, Antarctic sea ice forecasts have been challenging both due to model deficiencies and observational uncertainty [@zampieri2019]{==check if this is true, hehe.==}. 
+Dynamical seasonal forecasts of Summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.
+
+{==**S2** Introduce S2==}
 
@@
 
+```{r datasets}
+datasets <- list(cdr = CDR(),
+                 bt = BT(),
+                 osi = OSI(),
+                 era5 = ERA5()
+)
+which_dataset <- "cdr"
+```
+
 There is not a lot of data on sea ice properties, especially for things like thickness, age, etc.
@@
 
-We use NOAA/NSIDC's Climate Data Record V4 as the primary verification dataset. 
-Its sea ice concentration data are estimated taking the maximum value of the NASA Team [ref] and NASA Bootstrap [ref] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
+These data are not perfect. \cite{meier2019}
+Sensor sensitivity and resolution. 
+Weather filtering.
+Land 
+To account for observational uncertainty we performed the verification on other data sources. 
+We find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
+
+We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset. 
+Its sea ice concentration data are estimated taking the maximum value of the NASA Team [ref] and NASA Bootstrap [@comiso2023] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
 Both algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
@@
 
-These data are not perfect, there still exists a fair bit of observational uncertainty.
-We include other sources of data to take into acount the observational uncertainty.
+Results based on two other different datasets are provided as supplementary material:
 
-However, we find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
-Nevertheless, we provide figures using ERA5
+1. The NASA Bootstrap [@comiso2023] product is... it is not an independent dataset from NOAA/NSIDC CDRV4.
 
-The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [ref] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR.
-We also use
-
-Copernicus Climate Change Service (C3S) (2020): Sea ice concentration daily gridded data from 1978 to present derived from satellite observations.
-Copernicus Climate Change Service (C3S) Climate Data Store (CDS).
-DOI: 10.24381/cds.3cd8b812 (Accessed on DD-MMM-YYYY)
-
-{==From the CDS websiste:==} ERA5 is the fifth generation ECMWF reanalysis for the global climate and weather for the past 8 decades.
-Data is available from 1940 onwards.
-ERA5 replaces the ERA-Interim reanalysis.
-
-This parameter is the fraction of a grid box which is covered by sea ice.
-Sea ice can only occur in a grid box which includes ocean or inland water according to the land-sea mask and lake cover, at the resolution being used.
-This parameter can be known as sea-ice (area) fraction, sea-ice concentration and more generally as sea-ice cover.
-In ERA5, sea-ice cover is given by two external providers.
-Before 1979 the HadISST2 dataset is used.
-From 1979 to August 2007 the OSI SAF (409a) dataset is used and from September 2007 the OSI SAF oper dataset is used.
-Sea ice is frozen sea water which floats on the surface of the ocean.
-Sea ice does not include ice which forms on land such as glaciers, icebergs and ice-sheets.
-It also excludes ice shelves which are anchored on land, but protrude out over the surface of the ocean.
-These phenomena are not modelled by the IFS.
-Long-term monitoring of sea ice is important for understanding climate change.
-Sea ice also affects shipping routes through the polar regions.
-
-{\>\>Why is ERA5 better or worse than satellites?\<
-\<}
-
-### Bootstrap
-
-{\>\>From https://nsidc.org/data/nsidc-0079/versions/4\<\<} This sea ice concentration data set was derived using measurements from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
-Measurements from the Special Sensor Microwave Imager/Sounder (SSMIS) aboard DMSP-F17 are also included.
-The data set has been generated using the Advanced Microwave Scanning Radiometer - Earth Observing System (AMSR-E) Bootstrap Algorithm with daily varying tie-points.
-Daily (every other day prior to July 1987) and monthly data are available for both the north and south polar regions.
-Data are gridded on the SSM/I polar stereographic grid (25 x 25 km) and provided in two-byte integer format.
-Data coverage began on 01 November 1978 and is ongoing through the most current processing, with updated data processed several times annually.
-
-[@comiso2023]
-
-### NASA Team
-
-{==missing==}
-
-### CDR
-
-NOAA/NSIDC's Climate Data Record V4 combines the Bootstrap and NASA Team estimates in an attempt to overcome each algorithm's weaknesses.
-{\>\>More detail on how are they combined and their properties.\<\<} As such, it is is not a truly independent dataset.
-
-Meier, W. N., F. Fetterer, A. K. Windnagel, and S.
-Stewart.
-2021.
-NOAA/NSIDC Climate Data Record of Passive Microwave Sea Ice Concentration, Version 4.
-\[Indicate subset used\].
-Boulder, Colorado USA. NSIDC: National Snow and Ice Data Center https://doi.org/10.7265/efmz-2t65.
-\[Date Accessed\] {\>\>Add to zotero\<\<}
-
-### ERA5
-
-All datasets are regrided using bilinear interpolation to a common grid, which is a polar stereographic grid with approximately equal area of 25km².
+1.
+The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [ref] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
 
@@
 Sea Ice Extent is defined as the area of ocean covered with at least 15% ice.
-This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for low sea ice conditions.
+This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea ice conditions.
 
@@
 We also compute the Integrated Ice Edge Error (IIEE) [@goessling2016].
-This is defined as the area in which the model misspredicts sea ice concentration being above or belog 15% ice.
+This is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
 That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.
@@
 ### Bias
-
-```{r datasets}
-datasets <- list(cdr = CDR(),
-                 bt = BT(),
-                 osi = OSI(),
-                 era5 = ERA5()
-)
-which_dataset <- "cdr"
-```
 
@@
 ```{r fig-hindcast-extent}
-#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the 
+#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
 hindcast_extent |> 
@@
 
-@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and `r length(datasets)-1` observational datasets.
+@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR
 +S2 shows a severe low extent bias, especially in the late summer-early autumn.
@@
 
+Melt overestimation during the melt season is very common in models [@massonnet2023].
+
 Comparing +S2 with +S1, the latter has a smaller bias, especially at low lags ([Fig.
@@
 @fig-mean-growth]).
-At larger lags, +S1's bias in summer and autumn is very similar to +S2\`s.
+At larger lags, +S1's bias in summer and autumn is very similar to +S2's.
 This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1.
@@
 
-{\>\>Add regional extents?\<
-\<}
-
 ### RMSE
@@
 Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
-The only exception is the RMSE around Febraury forecasted from June onwards.
-
-{==Add regional?=
-=}
+The only exception is the RMSE around February forecasted from June onwards.
 
@@
 
+```{r}
+t_test <- function(x, y, ...) {
+
+  test <- t.test(x, y, ...)
+
+  list(low = test[["conf.int"]][1],
+       high = test[["conf.int"]][2], 
+       estimate = -diff(test$estimate), 
+       p.value = test[["p.value"]]
+  )
+}
+
+max_lag <- errors[version == "S1", max(lag)]
+
+dif <- errors |> 
+  na.omit() |> 
+  _[obs_dataset == which_dataset] |>
+  _[measure == "iiee"] |> 
+  _[version != "persistence"] |> 
+  _[lag <= max_lag] |> 
+  dcast(time + time_forecast + member + lag ~ version, value.var = "value") |> 
+  _[, t_test(na.omit(S1), na.omit(S2)), 
+    by = .(lag, month(time_forecast))] |> 
+  _[p.value > 0.01] |> 
+  _[, .SD[which.min(lag)], by = month] |> 
+  _[order(month), .(month, lag)]
+
+labels_month_significant <- setNames(paste0(month.abb, " (", dif$lag, ")"),
+                                     1:12)
+```
+
 ```{r fig-iiee}
-#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lag for all forecast initialised on the first of each month for ACCESS-S1 and ACCESS-S2 hindcasts.
+#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lag for all forecast initialised on the first of each month for +S1 and +S2 hindcasts.
+For each month, the number in parenthesis indicates the minimum lag at which the the mean error of each model is not statistically significant at a 1% level using a two-sisded t-test.
 
@@
   _[obs_dataset == which_dataset] |>
-  _[, month := lubridate::month(time_forecast, label = TRUE)]  |> 
+  # _[, month := lubridate::month(time_forecast, label = TRUE)]  |> 
   _[, median_ci(value), 
@@
   geom_line(aes(color = version), linewidth = 1) +
+
+  # geom_errorbar(data = \(x) x[dif, on = .NATURAL][version == "S1"], 
+  #              aes(x = lag, 
+  #                  ymin = mid - (high - low), ymax = mid + (high - low))) +
+
+  # geom_point(data = \(x) x[dif, on = .NATURAL][version == "S1"],
+  # aes(lag, mid), size = 3, shape = 21) +
+
   scale_y_continuous(labels = labels_extent) +
@@
   #   scale_fill_models +
-  facet_wrap(~ month, labeller = labeller(month = labels_month)) 
+  facet_wrap(~ month, labeller = labeller(month = labels_month_significant)) 
 ```
@@
 +S1 has lower error at short lags at all months, with the errors converging as the forecast goes on.
-The time to convergence depends on the month and it can be as fas as a few days in July to as large as several months for forecasts initiated in February and March.
+The time to convergence depends on the month and it can be as fas as two weeks June and July to as large as 160 days for forecasts initialised in February. 
 Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the the memory of sea ice to initial conditions; at least from October to March when the data assimilated form the other components has little to no influence on sea ice.

### 2024-10-21 analysis/initial_conditions/initial_conditions.qmd
Add exploration of IC for experiments

---
title: "Untitled"
format: revealjs
---

```{r setup}
library(lubridate)
library(ggplot2)
library(data.table)
library(metR)
library(rcdo)

cdo_options_set("-L")
source(here::here("R/functions.R"))
source(here::here("R/datasets.R"))
source(here::here("R/ggplot.R"))
```

## Experiments

Idea:

1. Use initial conditions from a very average day/month.
2. Run experiments with changing sea ice initial condtions: 
  1. Experiments with high initial sea ice
  2. Experiments with low sea ice.

## Identifying "average" starting points

1. Compute an error metric between dates and the climatology and chose the lowest. 
2. Three posible metrics:
  1. RMSE 
  2. IIEE
  3. Extent difference (IIEE, is better overall because it also includes location error, so I discard this)

---

```{r}
s2_monthly <- S2_reanalysis() |> 
  cdo_monmean()

s2_monthly_clim <- S2_reanalysis() |> 
  climatology() |> 
  cdo_monmean()

rmse_clim <- cdo_ymonsub(s2_monthly, s2_monthly_clim) |> 
  cdo_sqr() |> 
  cdo_fldmean() |> 
  cdo_sqrt() |> 
  cdo_chname("aice", "rmse") |> 
  cdo_execute() |> 
  ReadNetCDF("rmse") |> 
  _[, .(time, rmse)] |>
  melt(id.vars = "time")

```

```{r}
s2_monthly_edge <- s2_monthly |> 
  cdo_gtc(0.15) |> 
  cdo_execute()

s2_monthly_clim_edge <- s2_monthly_clim |> 
  cdo_gtc(0.15) |> 
  cdo_execute()

iiee_clim <- cdo_ymonne(s2_monthly_edge, s2_monthly_clim_edge) |>
  cdo_fldint() |>
  cdo_chname("aice", "iiee") |> 
  cdo_execute() |> 
  ReadNetCDF("iiee") |> 
  _[, .(time, iiee)] |> 
  melt(id.vars = "time")

```

```{r}
#| fig.cap: Standardised value of each errror.

rbind(iiee_clim, rmse_clim) |> 
  # _[, anomaly := value/median(value), by = .(variable, month(time))] |>
  _[, value := value/sd(value, na.rm = TRUE), by = .(variable)] |>
  ggplot(aes(time, value)) +
  geom_line(aes(color = variable)) +
  labs(color = "Error") +
  facet_wrap(~ month(time))
```

---

```{r}
#| fig.cap: Relationship between IIEE and RMSE with respeceet to climatology for each month. 
rbind(iiee_clim, rmse_clim) |> 
  # _[, anomaly := value/median(value), by = .(variable, month(time))] |>
  # _[, anomaly := as.numeric(scale(anomaly)), by = .(variable)] |> 
  dcast(time ~ variable, value.var = "value") |> 
  ggplot(aes(iiee, rmse)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ month(time))
```

```{r}
min_error <- rbind(iiee_clim, rmse_clim)  |> 
  copy() |> 
  _[, month := month(time)] |> 
  _[, .SD[which.min(value)], by = .(variable, month)]

s2_climatology <- s2_monthly_clim |> 
  cdo_execute() |> 
  ReadNetCDF(c(clim = "aice")) |> 
  setnames("time", "clim_time")

min_error_data <- s2_monthly |> 
  cdo_execute() |> 
   ReadNetCDF("aice", subset = list(time = as.list(min_error$time))) |> 
  _[, clim_time := update(time, year = 2000)] |> 
  _[min_error, on = .NATURAL, allow.cartesian = TRUE]

```

---

```{r}
min_error_data |> 
  _[variable == "iiee"] |> 
  _[, label := paste0(formatC(month(time), width = "2", flag = "0"), " - ", as.character(time))] |> 
  s2_climatology[i = _, on = .NATURAL] |> 
  ggplot(aes(xgrid, ygrid)) +
  geom_contour_fill(aes(z = aice - clim, fill = after_stat(level))) +
  geom_contour2(aes(z = aice), breaks = 0.15) +
  geom_contour2(aes(z = clim), breaks = 0.15, linetype = 2) +
  scale_fill_divergent_discretised() +
  geomcoord_antarctica +
  geom_antarctica_path +
  facet_wrap(~ label) +
  wide_legend +
  labs(title = "iiee")
```

---

```{r}
min_error_data |> 
  _[variable == "rmse"] |> 
  _[, label := paste0(formatC(month(time), width = "2", flag = "0"), " - ", as.character(time))] |> 
  s2_climatology[i = _, on = .NATURAL] |> 
  ggplot(aes(xgrid, ygrid)) +
  geom_contour_fill(aes(z = aice - clim, fill = after_stat(level))) +
  geom_contour2(aes(z = aice), breaks = 0.15) +
  geom_contour2(aes(z = clim), breaks = 0.15, linetype = 2) +
  scale_fill_divergent_discretised() +
  geomcoord_antarctica +
  geom_antarctica_path +
  facet_wrap(~ label) +
  wide_legend +
  labs(title = "rmse")
```

### 2024-10-21 analysis/initial_conditions/initial_conditions.qmd
More IC

@@
 format: revealjs
+execute:
+  cache: true
 ---
@@
 ```{r setup}
+#| cache: false
 library(lubridate)
@@
 2.  Run experiments with changing sea ice initial condtions:
-  1. Experiments with high initial sea ice
-  2. Experiments with low sea ice.
+3.  Experiments with high initial sea ice
+4.  Experiments with low sea ice.
 
@@
 1.  Compute an error metric between dates and the climatology and chose the lowest.
-2. Three posible metrics:
+2.  Three possible metrics:
     1.  RMSE
@@
 
----
+------------------------------------------------------------------------
 
@@
 
----
-
+------------------------------------------------------------------------
+```{r}
+breaks <- seq(-0.9, 0.9, by = 0.1)
+```
 ```{r}
@@
   _[variable == "iiee"] |> 
-  _[, label := paste0(formatC(month(time), width = "2", flag = "0"), " - ", as.character(time))] |> 
+  _[, label := paste0(formatC(month(time), width = "2", flag = "0"), " - ", as.character(as.Date(time)))] |> 
   s2_climatology[i = _, on = .NATURAL] |> 
   ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = aice - clim, fill = after_stat(level))) +
+  geom_contour_fill(aes(z = aice - clim, fill = after_stat(level)), breaks = breaks) +
   geom_contour2(aes(z = aice), breaks = 0.15) +
@@
   geom_antarctica_path +
-  facet_wrap(~ label) +
+  facet_wrap(~ label, ncol = 6) +
   wide_legend +
@@
   _[variable == "rmse"] |> 
-  _[, label := paste0(formatC(month(time), width = "2", flag = "0"), " - ", as.character(time))] |> 
+  _[, label := paste0(formatC(month(time), width = "2", flag = "0"), " - ", as.character(as.Date(time)))] |> 
   s2_climatology[i = _, on = .NATURAL] |> 
   ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = aice - clim, fill = after_stat(level))) +
+  geom_contour_fill(aes(z = aice - clim, fill = after_stat(level)), breaks = breaks) +
   geom_contour2(aes(z = aice), breaks = 0.15) +

### 2024-10-24 analysis/paper/access-ice.qmd
A bit more introduction

@@
 
-{==**Antarctic sea ice general description**.
-Seasonality, trends.==}
-
-.....
-
-Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice had a slightly increasing trend that models systematically failed to reproduce and puzzled researchers \cite{hobbs2016}.
+Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice had a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016].
 Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change.
@@
 
-{==**Importance**.
-Impacts on life.
-Potential impacts on weather.
-Impacts on climate.
-Impacts on ocean circulation.(Justify why it's important to study and understand)==}
+Antarctic sea ice is in a remote location but the potential impacts of its variability and trends extend far into the lower latitudes, affecting ocean and atmospheric circulation, ocean carbon uptake and biological processes.
+It's likely that a reduction in sea ice extent would reduce the temperature gradient between high and low latitudes (and vice versa), which in turn influences the strength and location of the eddy-driven jet, which is a crucial component of weather events.
+For the northern hemisphere, it has been hypothesised that the reduction in sea ice in the northern hemisphere has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015].
+There is some evidence of this effect in the Southern Hemisphere. 
+Models show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in Springs following a low sea ice maximum. 
+But models don't fully agree on the exact phase of the zonal wave response and observations don't show a clear signal [@rea2024]. 
+The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location [@ayres2019]. 
+It is also small compared with the direct effect of CO2 increase and Sea Surface Temperature warming [@ayres2019].
+On climate scales, it is thought that the signal of a reduction in Antarctic sea ice is like a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].
 
-Antarctic sea ice is in a remote location but the potential impacts of its variability and trends extend far into the lower latitudes, affecting ocean and atmospheric circulation, ocean carbon uptake and biological processes.
-It's likely that a reduction in sea ice extent would reduce the temperature gradient between high and low latitudes (and vice versa), which in turn influences the strength and location of the jet, which is a crucial component of weather events.
-For the northern hemisphere, it has been hypothesised that the reduction in sea ice in the northern hemisphere has lead to a weaker, wavier jet, increasing the frequency of extreme events [refs].
-Less is known about the effects of Antarctic sea ice variability {==is this true?
-It feels that way from the (lack of) literature.
-Is there a reference to substantiate this claim?==} on weather timescales.
-On climate scales, it is thought that the signal of a reduction in Antarctic sea ice is like a "mini global warming" and experiments also predict a weakening and northerly shift of the jet [@ayres2022; @england2020].
-
-{==Relevant papers to read:
-
-\cite{ayres2019}
-\cite{england2018}
-\cite{raphael2011}
-\cite{raphael2003a}
-
-==}
-
-{==**Prediction systems.** Why is it important: operations {\>\>Discuss with Phil\<\<}; science needs (we need good models).\cite{desilva2020}\cite{wagner2020}
-What is available.
-Discussion on S1 {\>\>Can we discuss previous S1 results?
-Laura's paper is not public.\<\<}==}
-
-Correctly modelling Antarctic sea ice is not only necessary for process understanding and climate projections to inform adaptation strategies, but accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent [@desilva2020 @wagner2020].
-However, Antarctic sea ice forecasts have been challenging both due to model deficiencies and observational uncertainty [@zampieri2019]{==check if this is true, hehe.==}. 
+Correctly modelling Antarctic sea ice is not only necessary for process understanding and climate projections to inform adaptation strategies. 
+Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent [@desilva2020 @wagner2020]. 
+However, Antarctic sea ice forecasts have been challenging both due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019]. 
 Dynamical seasonal forecasts of Summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.
@@
 
-```{r}
+```{r extent_climatology} 
 extent_climatology <- extent_daily |>  
@@
   _[, lag := as.numeric(as.Date(time) - forecast_time)] |> 
-  _[lag > 0]
+  _[lag > 0] |> 
+  _[!(month(forecast_time) == 1 & member == 7 & model == "S1")]  # This member is wrong
 ```
@@
 @fig-mean-growth]).
-This bias is common in climate models \[ref!\]
-and it's been linked to xxxx??
-
-Melt overestimation during the melt season is very common in models [@massonnet2023].
+Many S2S systems exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
 
@@
 This figure uses +CDR data, but the results are nearly identical compared with ERA5 or Bootstrap.
-Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a referece?=
+Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a reference?=
 =}.

### 2024-10-25 analysis/paper/access-ice.qmd
Reverse bias colours

@@
                     breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
-  scale_fill_divergent_discretised(paste0("Bias with respect to ", labels_models[which_dataset])) +
+  scale_fill_divergent_discretised(paste0("Bias with respect to ",
+                                          labels_models[which_dataset]),
+                                   low = scales::muted("red"),
+                                   high = scales::muted("blue")) +
   geomcoord_antarctica +

### 2024-11-04 R/functions.R
Experiments

@@ -109,7 +109,7 @@ extent <- function(dataset) {
 }
 
 remap_cdr <- function(file) {
-    nsidc_grid <- "data/raw/nsidc_grid.txt"
+    nsidc_grid <- here::here("data/raw/nsidc_grid.txt")
 
     rcdo::cdo_remapbil(file, nsidc_grid)
 }

### 2024-11-12 R/ggplot.R
Adds hindcast climatology

@@ -89,9 +89,9 @@ geom_antarctica_fill <- ggplot2::geom_polygon(data = contour, ggplot2::aes(x, y,
 
 geomcoord_antarctica <- list(
     NULL,
-    ggplot2::coord_sf(crs = sic_projection, lims_method = "box"),
-    ggplot2::scale_x_continuous(name = NULL, expand = c(0, 0)),
-    ggplot2::scale_y_continuous(name = NULL, expand = c(0, 0)),
+    ggplot2::coord_sf(crs = sic_projection, lims_method = "box", label_axes =  "----"),
+    ggplot2::scale_x_continuous(name = NULL, expand = c(0, 0), labels = NULL),
+    ggplot2::scale_y_continuous(name = NULL, expand = c(0, 0), labels = NULL),
     geom_antarctica_path
 )
 

### 2024-11-12 scripts/hindcast_climatology.R

@@ -1,0 +1,26 @@
+library(furrr)
+
+plan(multisession, workers = 12)
+
+source(here::here("R/datasets.R"))
+
+times <- get_forecast_times("S2")
+
+nsidc_grid <- here::here("data/raw/nsidc_grid.txt")
+
+
+months <- 1:12
+
+furrr::future_walk(months, \(month) {
+  month <- formatC(month, width = 2, flag = "0")
+  files <- here::here(glue::glue("data/derived/hindcast_ensmean/di_aice_*{month}01_emm.nc")) |> 
+    Sys.glob()
+  files_chr <- paste0(files, collapse = " ")
+  
+  out <- here::here(glue::glue("data/derived/hindcast_climatology/mo_aice_1981{month}01_mean.nc"))
+  dir.create(dirname(out), showWarnings = FALSE, recursive = TRUE)
+  glue::glue("cdo -L -O -s ensmean [ -monmean : {files_chr} ] {out}") |> 
+    system()
+})
+
+

### 2024-11-12 scripts/pbs/download_era5.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+#PBS -P lo70
+#PBS -q copyq
+#PBS -l ncpus=1
+#PBS -l mem=64GB
+#PBS -l walltime=01:00:00
+#PBS -l wd
+#PBS -l jobfs=400GB
+
+# Load module, always specify version number.
+module load R/4.3.1
+
+Rscript scripts/download_era5.R > logs/downlaod_era5.log

### 2024-11-12 scripts/pbs/hindcast_climatology.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+#PBS -P dx2
+#PBS -q normal
+#PBS -l ncpus=48
+#PBS -l mem=190GB
+#PBS -l walltime=03:00:00
+#PBS -l wd
+#PBS -l jobfs=400GB
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7
+
+# Load module, always specify version number.
+module load R/4.3.1
+module load cdo/2.4.3
+
+# Must include `#PBS -l storage=scratch/ab12+gdata/yz98` if the job
+# needs access to `/scratch/ab12/` and `/g/data/yz98/`. Details on:
+# https://opus.nci.org.au/display/Help/PBS+Directives+Explained
+
+# Run R application
+SCRIPT=hindcast_climatology.R
+Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log

### 2024-11-12 analysis/poster_aarc/poster_figures.Rmd
Adds poster figures

--- 2024-10-24 analysis/paper/access-ice.qmd
@@
 ---
-title: "Untitled"
-format: gfm
-execute: 
-  echo: false
-  warning: false
-  message: false
-  cache: false
-filters:
-  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
-  - ../../_extensions/mloubout/critic-markup/critic-markup.lua
-  - ../../_extensions/ute/search-replace/search-replace.lua
-search-replace:
-  +S2: ACCESS-S2
-  +S1: ACCESS-S1
-  +CDR: NOAA/NSIDC CDRV4
-bibliography: references.bib
-editor: 
-  markdown: 
-    wrap: sentence
+title: "Poster figures"
+output: 
+  html_document:
+    self_contained: false  
+date: "2024-11-11"
 ---
@@
 source(here::here("R/ggplot.R"))
+
 ```
 
-# Abstract
+```{r}
 
-balbalbal
+my_theme <- function(base_size = 20) {
+  ggplot2::theme_minimal(base_size = base_size) +
+            ggplot2::theme(panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
+                  legend.position = "bottom",
+                  legend.title.position = "top", 
+                  legend.title = ggplot2::element_text(hjust = 0.5),
+                  legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
+                  legend.key.height = grid::unit(0.75, "lines"),
+                  axis.text.y = element_text_first_line())
+}
 
-# Introduction
+ggplot2::theme_set(my_theme())
 
-Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice had a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016].
-Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change.
-[ref: that CESM run with forced winds].
-
-Antarctic sea ice is in a remote location but the potential impacts of its variability and trends extend far into the lower latitudes, affecting ocean and atmospheric circulation, ocean carbon uptake and biological processes.
-It's likely that a reduction in sea ice extent would reduce the temperature gradient between high and low latitudes (and vice versa), which in turn influences the strength and location of the eddy-driven jet, which is a crucial component of weather events.
-For the northern hemisphere, it has been hypothesised that the reduction in sea ice in the northern hemisphere has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015].
-There is some evidence of this effect in the Southern Hemisphere. 
-Models show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in Springs following a low sea ice maximum. 
-But models don't fully agree on the exact phase of the zonal wave response and observations don't show a clear signal [@rea2024]. 
-The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location [@ayres2019]. 
-It is also small compared with the direct effect of CO2 increase and Sea Surface Temperature warming [@ayres2019].
-On climate scales, it is thought that the signal of a reduction in Antarctic sea ice is like a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].
-
-Correctly modelling Antarctic sea ice is not only necessary for process understanding and climate projections to inform adaptation strategies. 
-Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent [@desilva2020 @wagner2020]. 
-However, Antarctic sea ice forecasts have been challenging both due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019]. 
-Dynamical seasonal forecasts of Summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.
-
-{==**S2** Introduce S2==}
-
-{==**Objective.** Evaluate S2.==}
-
-# Data and methods
-
-## +S2
-
-+S2 [@wedd2022] is the Australian Bureau of Meteorology's current seasonal prediction system that replaced +S1 [@hudson2017] in October 2021. 
-Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0  [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
-The atmosphere has a N216 horizontal resolution(~60km in the mid-latitudes) with 85 levels. 
-The land model uses the same horizontal grid with 4 soil levels. 
-The ocean component has a 1/4º resolution with 75 vertical levels. 
-The ice component --based on CICE version 4.1-- has the same resolution than the ocean and 5 sea ice categories as well as an open water category.
-
-Both systems took atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts. 
-The main difference between the two are the ocean and ice initial conditions. 
-+S1 initial conditions came from from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) DA scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
-+S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022. 
-This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
-SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are larger than 0ºC.
-Relevant for this work, sea ice concentrations are not assimilated.
-
-For evaluation we use hindcast for the period 1981--2023{\>\>Check\<\<}.
-Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
-Climatology is smoothed with a 11 day running mean.
-
-## Verification datasets
+```
 
@@
 ```
-
-There is not a lot of data on sea ice properties, especially for things like thickness, age, etc.
-However there are relatively reliable satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
-
-These data are not perfect. \cite{meier2019}
-Sensor sensitivity and resolution. 
-Weather filtering.
-Land 
-To account for observational uncertainty we performed the verification on other data sources. 
-We find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
-
-We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset. 
-Its sea ice concentration data are estimated taking the maximum value of the NASA Team [ref] and NASA Bootstrap [@comiso2023] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
-Both algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
-The data has a resolution of 25 by 25km and daily from 1978 onwards.
-
-Results based on two other different datasets are provided as supplementary material:
-
-1. The NASA Bootstrap [@comiso2023] product is... it is not an independent dataset from NOAA/NSIDC CDRV4.
-
-1.
-The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [ref] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
-
-## Error measures
-
-For evaluation purposes, we use a series of measures.
-
-### Sea ice extent
-
-Sea Ice Extent is defined as the area of ocean covered with at least 15% ice.
-This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea ice conditions.
-
-Sea Ice Extent is a rough global measure, but a model could have relatively accurate extent of ice but with different distributions.
-We use two other measures to account for these errors.
-
-We compute Root Mean Square Error of sea ice concentration anomalies.
-
-We also compute the Integrated Ice Edge Error (IIEE) [@goessling2016].
-This is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
-That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.
-
-# Results and discussion
-
-## Reanalysis
-
-### Bias
 
@@
 
-```{r fig-hindcast-extent}
+```{r fig-hindcast-extent, fig.width=2306/150, fig.height=1352/150}
 #| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
@@
             aes(x = as.Date(time),
-                group = dataset)) + 
-  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +
+                group = dataset),
+            linewidth = 1) + 
+  geom_line(aes(color = month, group = interaction(forecast_time, time != time2)),
+            linewidth = 1) +
 
@@
              aes(fill = month), 
-             shape = 24, size = 2.2) +
+             shape = 24, size = 4) +
 
   geom_point(data = \(x) x[lag == 1], aes(fill = month), 
-             shape = 21, size = 2) +
+             shape = 21, size = 4) +
 
@@
   tag_facets() 
-```
-
-@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR
-+S2 shows a severe low extent bias, especially in the late summer-early autumn.
-This is due primarily to a faster and longer melt season between January and March and slower growth during March and April This is then balanced with faster growth between May and July ([Fig.
-@fig-mean-growth]).
-Many S2S systems exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
-
-Comparing +S2 with +S1, the latter has a smaller bias, especially at low lags ([Fig.
-@fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig.
-@fig-mean-growth]).
-At larger lags, +S1's bias in summer and autumn is very similar to +S2's.
-This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1.
-
-At long lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
-Therefore we can estimate the latter using the hindcasts with the largest possible lag, which is shown in triangles in @fig-hindcast-extent for the same dates as the initial conditions.
-The difference between the two is the effect of the data assimilation.
-
-The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to both having the same model formulation.
-From June to October, in +S2 circles move away from triangles and towards observations, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and affecting the initial conditions.
-The rest of the year, there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.
-
-```{r fig-mean-growth}
-#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations. Values are smoothed with a 2-degree loess smooth with a 30 day window.
-
-dif_hindcast <- hindcast_extent |> 
-  copy() |> 
-  _[order(time)] |> 
-  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(model, forecast_time, member)] |> 
-  _[, median_ci(dif), by = .(model, time = update(time, year = 2000))]
-
-extent_daily |> 
-  _[order(time)] |> 
-  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
-  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
-  _[dataset == which_dataset] |> 
-  ggplot(aes(time, mid)) +
-  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
-  geom_smooth(aes(group = dataset, color = dataset),
-              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
-  geom_smooth(data = dif_hindcast,
-              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
-  scale_x_datetime(NULL, date_breaks = "1 month", 
-                   date_labels = "%b", expand = c(0, 0)) +
-  scale_y_continuous(NULL, labels = labels_extent) +
-  scale_color_models
 ```
@@
 
-```{r fig-bias}
-#| fig-width: 9
-#| fig-height: 6
-#| fig-cap: +S2 reanalysis sea ice concentration bias compared with NSIDC sea ice concentration.
+```{r hindcast_clim}
+zeropad <- function(x, n = 2) formatC(x, width = 2, flag = "0")
+monnb <- function(d) { lt <- as.POSIXlt(as.Date(d, origin="1900-01-01")); 
+lt$year*12 + lt$mon } 
+mondf <- function(d1, d2) { monnb(d2) - monnb(d1) }
 
-climatologies |> 
-  _[model %in% c("S2", which_dataset)] |> 
-  dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
-  setnames(which_dataset, "obs") |> 
+months <- 1:12
+hindcast_clim <- lapply(setNames(months, months), \(month) {
+  here::here(glue::glue("data/derived/hindcast_climatology/mo_aice_1981{zeropad(month)}01_mean.nc")) |> 
+    ReadNetCDF("aice")
+}) |> 
+  rbindlist(idcol = "forecast_month") |> 
+  _[, forecast_month := as.numeric(forecast_month)] |> 
+  _[, forecast_time := lubridate::make_datetime(1981, forecast_month, 01)] |> 
+  _[, time := lubridate::floor_date(time, "month")] |> 
+  _[, lag := mondf(forecast_time, time)] |> 
+  _[lag == 0] |> 
+  _[, time2 := update(time, year = 2000)]
+
+obs_clim <- CDR() |> 
+  climatology() |> 
+  ReadNetCDF(c(obs = "aice")) |> 
+  setnames("time", "time2")
+
+hindcast_clim <- hindcast_clim |> 
+  merge(obs_clim) |> 
+  _[, month := month(time)] 
+```
+
+```{r fig-bias, fig.height=2306/150, fig.width=2306/150}
+#| fig-cap: 1-month lag +S2 forecast bias compared with NSIDC.
+
+hindcast_clim |> 
+  # _[time == time[1]] |> 
   ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = S2 - obs, fill = after_stat(level)), 
+  geom_contour_fill(aes(z = aice - obs, fill = after_stat(level)), 
                     breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
-  scale_fill_divergent_discretised(paste0("Bias with respect to ", labels_models[which_dataset])) +
+  scale_fill_divergent_discretised(NULL,
+                                   low = scales::muted("red"),
+                                   high = scales::muted("blue")) +
   geomcoord_antarctica +
@@
   wide_legend +
-  tag_facets()
+  tag_facets() +
+  theme(panel.ontop = TRUE, 
+        panel.background = element_blank(),
+        panel.grid = element_line(colour = alpha("gray50", 0.5), linewidth = 0.4))
 ```
@@
 In winter, the differences are mostly on the sea ice edge, with slight positive bias in XXX and negative bias around the Indian Ocean sector.
-
-### Anomalies
-
-{==intro connecting subsections==}
-
-```{r fig-extent-anom}
-#| fig-cap: Sea ice extent anomalies for +S1 and +S2 (black) and +CDR (blue).
-
-months_difference <- function(x, y) {
-  lubridate::interval(y, x) %/% months(1)
-}
-
-monthly_extent <- extent_daily |> 
-  _[aice == 0, aice := NA]  |> 
-  _[, .(aice = mean(aice)), by = .(dataset, time = as.Date(round_date(time, "month")))] |>
-  _[, aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE), 
-    by = .(month(time), dataset)]
-
-N <- uniqueN(monthly_extent$time)
-
-hindcast_extent |> 
-  copy() |> 
-  # _[model == "S2"] |> 
-  _[, aice := aice - mean(aice[year(time) %between% c(1981, 2011)]), 
-    by = .(lag, model, time = update(time, year = 2000))] |> 
-  _[, .(aice = mean(aice)), by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
-  _[, lag := months_difference(time, forecast_time) + 1] |> 
-  _[lag %in% c(1, 3, 5, 7)] |> 
-  ggplot(aes(time, aice)) +
-  geom_line(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
-              _[dataset == which_dataset],
-            aes(group = dataset, color = dataset)) +
-  # geom_smooth(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
-  #               _[dataset == "cdr"], 
-  #             aes(group = dataset, color = dataset),
-  #             se = FALSE, span = 10*12/N, n = 366) +
-  geom_line(aes(color = model))  +
-  # geom_smooth(aes(color = model), se = FALSE, span = 10*12/N, n = 366) +
-  scale_y_continuous(NULL,
-                     labels = labels_extent) +
-  scale_x_date(NULL, expand = c(0, 0)) +
-  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Lag: ", x))) +
-  scale_color_models
-```
-
-@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lags.
-The anomalies in this case were computed with respect of the climatology of each lag, which is a way of bias-correction.
-Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at lag 3.
-+S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.
 
@@
 
-```{r fig-rmse}
+```{r fig-rmse, fig.height=1524/150, fig.width=2306/150}
 #| fig-cap: Median and 95% coverage of sea ice concentration anomalies RMSE as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies.
@@
   geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
-  scale_x_continuous(breaks = NULL) +
+  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
   labs(y = "RMSE", 
@@
 ```
-
-@fig-rmse shows the median and 95% range of RMSE of sea ice concentration anomalies for +S2 forecasts compared with a benchmark of persistence.
-This figure uses +CDR data, but the results are nearly identical compared with ERA5 or Bootstrap.
-Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a reference?=
-=}.
-Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
-The only exception is the RMSE around February forecasted from June onwards.
 
@@
 
-```{r fig-iiee}
+```{r fig-iiee, fig.height=1524/150, fig.width=2306/150}
 #| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lag for all forecast initialised on the first of each month for +S1 and +S2 hindcasts.
@@
   scale_y_continuous(labels = labels_extent) +
-  scale_x_continuous(breaks = NULL) +
+  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
   scale_color_models +
@@
 ```
-
-To compare +S2 with +S1, we computed the IIEE for both models.
-This error measure is shown in @fig-iiee for all lags and forecasts initialised at the first of every month.
-+S1 has lower error at short lags at all months, with the errors converging as the forecast goes on.
-The time to convergence depends on the month and it can be as fas as two weeks June and July to as large as 160 days for forecasts initialised in February. 
-Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the the memory of sea ice to initial conditions; at least from October to March when the data assimilated form the other components has little to no influence on sea ice.
-
-Also evident in @fig-iiee is the difference in the error spread at short lags between +S2 and +S1.
-In all month +S1 shows a small error spread at lag 1, indicating that the error in the initial conditions not only is small, but also fairly constant.
-This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
-For +S2, this is true only only between July and October, approximately.
-For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it's not constant.
-
-```{r fig-iiee-variance}
-#| fig-cap: Mean spread of IIEE at different lags for different models. 
-#| fig.height: 9
-#| fig.width: 4
-labels_extent_2 <- function(x) {
-  m <- which.max(x)
-  x <- scales::label_number(scale = (1e-12)^2)(x)
-  x[m] <- paste0(x[m], "\nM km⁴")
-  x
-}
-
-rbind(mean_variance = errors |>
-        _[measure == "iiee"] |>
-        _[obs_dataset == which_dataset] |>
-        _[, var(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
-        _[, .(value = mean(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
-
-      variance_of_mean = errors |>
-        _[measure == "iiee"] |>
-        _[obs_dataset == which_dataset] |>
-        _[, mean(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
-        _[, .(value = var(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
-      idcol = "component") |>
-
-  ggplot(aes(lag, value)) +
-
-  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
-             colour = "gray50", linewidth = 0.1) +
-
-  geom_text(data = month_from_lag(15, 2), aes(y = Inf, label = month2), 
-            size = 2.5,
-            vjust = 1) +
-
-  geom_text(data = month_from_lag(1, 2), aes(y = -Inf, label = lag), 
-            size = 2.5,
-            vjust = 0) +
-
-  geom_line(aes(color = version), linewidth = 1) +
-  scale_y_continuous("Variance", 
-                     labels = labels_extent_2) +
-  # scale_color_models +
-  scale_x_continuous("Lag", breaks = NULL) +
-  scale_color_models +
-  facet_grid(month ~ component, 
-             labeller = labeller(month = labels_month, 
-                                 component = c(mean_variance = "Mean variance",
-                                               variance_of_mean = "Variance of mean"))) 
-```
-
-The large initial error spread could be due either to large spread of ensemble members or due to a large spread of individual forecasts.
-@fig-iiee-variance splits the IIEE variance for each lag into the mean variance of each individual forecast and the variance of the mean error of each individual forecast, which adds up to the total variance.
-The average variance of each forecast is almost identical between forecast systems in all months.
-This shows that the ensemble spread of individual forecasts evolves identically, which, again, it's not unexpected because both systems share the same model formulation.
-This also shows that the perturbation scheme in +S2 is comparable to the one in +S1.
-
-On the other hand, the spread of the mean error is always larger in +S2 than +S1.
-The difference is particularly large at short lags in some months, which coindice with the ones in which the data assimilation scheme is not influencing sea ice initial conditions.
-
-## Conclusions
-
-# References
-
-::: {#refs}
-:::
-
-# Suplementary figures
-
-<!-- ## ERA5 -->
-
-<!-- ```{r dataset_era} -->
-<!-- which_dataset <- "era5" -->
-<!-- ``` -->
-
-<!-- ```{r fig-hindcast-extent-era5} -->
-<!-- which_dataset <- "era5" -->
-<!-- <<fig-hindcast-extent>> -->
-<!-- ``` -->
-
-<!-- ```{r fig-hindcast-extent-bt} -->
-<!-- which_dataset <- "bt" -->
-<!-- <<fig-hindcast-extent>> -->
-<!-- ``` -->
-
-<!-- ```{r fig-mean-growth} -->
-<!-- <<fig-mean-growth>> -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- <<fig-bias>> -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- <<<fig-extent-anom>> -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- <<fig-rmse>> -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- <<fig-iiee>> -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- <<fig-iiee-variance>> -->
-<!-- ``` -->

### 2024-12-06 analysis/paper/access-ice.qmd
Adds rmse by longitude

--- 2024-10-24 analysis/paper/access-ice.qmd
@@
 Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change.
-[ref: that CESM run with forced winds].
+\[ref: that CESM run with forced winds\].
 
@@
 
-Correctly modelling Antarctic sea ice is not only necessary for process understanding and climate projections to inform adaptation strategies. 
+Correctly modelling Antarctic sea undsen sector and the Amundsen-Bellingshausen sector (Fice is not only necessary for process understanding and climate projections to inform adaptation strategies.
 Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent [@desilva2020 @wagner2020].
@@
 Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
-The atmosphere has a N216 horizontal resolution(~60km in the mid-latitudes) with 85 levels. 
+The atmosphere has a N216 horizontal resolution(\~60km in the mid-latitudes) with 85 levels.
 The land model uses the same horizontal grid with 4 soil levels.
@@
 
-These data are not perfect. \cite{meier2019}
-Sensor sensitivity and resolution. 
+These data are not perfect.
+\cite{meier2019} Sensor sensitivity and resolution.
 Weather filtering.
-Land 
-To account for observational uncertainty we performed the verification on other data sources. 
+Land To account for observational uncertainty we performed the verification on other data sources.
 We find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
@@
 We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset.
-Its sea ice concentration data are estimated taking the maximum value of the NASA Team [ref] and NASA Bootstrap [@comiso2023] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
+Its sea ice concentration data are estimated taking the maximum value of the NASA Team \[ref\] and NASA Bootstrap [@comiso2023] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
 Both algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
@@
 
-1.
-The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [ref] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
+2.
+The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) \[ref\] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
 
@@
 
+```{r area_daily}
+area_daily <- datasets |>
+  lapply(area) |>
+  lapply(\(x) ReadNetCDF(x, "aice")) |>
+  rbindlist(idcol = "dataset") |>
+  _[, let(lon = NULL, lat = NULL)]
+```
+
+```{r area_climatology}
+area_climatology <- area_daily |>  
+  copy() |>
+  _[aice == 0, aice := NA] |> 
+  _[!(month(time) == 2 & mday(time) == 29)] |> 
+  _[, time := update(time, year = 2001)] |> 
+  _[, median_ci(aice), by = .(time, dataset)]
+
+```
+
 ```{r hindcast_extent}
@@
                          S2 = readRDS(here::here("data/derived/S2_hindcast_extent.Rds")), 
+                         idcol = "model") |>
+  _[, model := relevel(factor(model), "S2")] |> 
+  _[, lag := as.numeric(as.Date(time) - forecast_time)] |> 
+  _[lag > 0] |> 
+  _[!(month(forecast_time) == 1 & member == 7 & model == "S1")]  # This member is wrong
+```
+
+```{r}
+hindcast_area <- rbind(S1 = readRDS(here::here("data/derived/S1_hindcast_area.Rds")),
+                       S2 = readRDS(here::here("data/derived/S2_hindcast_area.Rds")), 
                        idcol = "model") |>
@@
 
-@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR
-+S2 shows a severe low extent bias, especially in the late summer-early autumn.
+@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR +S2 shows a severe low extent bias, especially in the late summer-early autumn.
 This is due primarily to a faster and longer melt season between January and March and slower growth during March and April This is then balanced with faster growth between May and July ([Fig.
@@
 
+```{r hindcast_clim}
+zeropad <- function(x, n = 2) formatC(x, width = 2, flag = "0")
+monnb <- function(d) { lt <- as.POSIXlt(as.Date(d, origin="1900-01-01")); 
+lt$year*12 + lt$mon } 
+mondf <- function(d1, d2) { monnb(d2) - monnb(d1) }
+
+months <- 1:12
+hindcast_clim <- lapply(setNames(months, months), \(month) {
+  here::here(glue::glue("data/derived/hindcast_climatology/mo_aice_1981{zeropad(month)}01_mean.nc")) |> 
+    ReadNetCDF("aice")
+}) |> 
+  rbindlist(idcol = "forecast_month") |> 
+  _[, forecast_month := as.numeric(forecast_month)] |> 
+  _[, forecast_time := lubridate::make_datetime(1981, forecast_month, 01)] |> 
+  _[, time := lubridate::floor_date(time, "month")] |> 
+  _[, lag := mondf(forecast_time, time)] |> 
+  _[lag == 0] |> 
+  _[, time2 := update(time, year = 2000)]
+
+obs_clim <- CDR() |> 
+  climatology() |> 
+  ReadNetCDF(c(obs = "aice")) |> 
+  setnames("time", "time2")
+
+hindcast_clim <- hindcast_clim |> 
+  merge(obs_clim) |> 
+  _[, month := month(time)] 
+```
+
 ```{r fig-bias}
@@
 #| fig-height: 6
-#| fig-cap: +S2 reanalysis sea ice concentration bias compared with NSIDC sea ice concentration.
+#| fig-cap: 1-month lag +S2 forecast bias compared with NSIDC.
 
-climatologies |> 
-  _[model %in% c("S2", which_dataset)] |> 
-  dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
-  setnames(which_dataset, "obs") |> 
+hindcast_clim |> 
+  # dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
+  # setnames(which_dataset, "obs") |>
   ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = S2 - obs, fill = after_stat(level)), 
+  geom_contour_fill(aes(z = aice - obs, fill = after_stat(level)), 
                     breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
-  scale_fill_divergent_discretised(paste0("Bias with respect to ", labels_models[which_dataset])) +
+  scale_fill_divergent_discretised(paste0("Bias with respect to ",
+                                          labels_models[which_dataset]),
+                                   low = scales::muted("red"),
+                                   high = scales::muted("blue")) +
   geomcoord_antarctica +
@@
 Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
-The only exception is the RMSE around February forecasted from June onwards.
+The only exception is the RMSE around February.
+
+```{r}
+topo_lonlat <- rcdo::cdo_topo(grid = "r180x90") |>
+  rcdo::cdo_execute(options = c("-f nc")) |>
+  metR::ReadNetCDF(c(z = "topo"), subset = list(lat = c(-90, -60))) |> 
+  ggperiodic::periodic(lon = c(0, 360)) 
+```
+
+```{r rmse_lon}
+rmse_lon <- here::here("data/derived/rmse_lon.Rds") |> 
+  readRDS() |> 
+  _[version != "S1"] |> 
+  _[, lag := as.numeric(as.Date(time) - time_forecast)]
+
+rmse_lon[, base := value[version == "persistence"], 
+         by = .(time, lon, lag, obs_dataset)]
+
+rmse_lon_mean <- rmse_lon |> 
+  _[version != "persistence"] |>
+  _[, .(dif = median(value - base, na.rm = TRUE)),
+    by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))]
+```
+
+```{r fig-rmse_lon, fig.height=7, fig.width=10}
+#| fig.cap: Meadian difference between RMSE of +S2 forecasts and persistence forecast at longitudinal bands.
+
+rmse_lon_mean |> 
+  _[obs_dataset == which_dataset] |> 
+  ggperiodic::periodic(lon = c(0, 360)) |> 
+  ggplot(aes(lon, lag)) +
+  geom_contour_fill(aes(z = dif, fill = after_stat(level)), binwidth = 0.025)  +
+  geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.4),  binwidth = 0.025) +
+  geom_contour2(data = topo_lonlat,
+                aes(lon, scales::rescale(lat, c(1, 150)), z = z),
+                breaks = 0) +
+  geom_hline(data = month_from_lag(1, 1), aes(yintercept = lag),
+             colour = "gray50", linewidth = 0.1) +
+
+  geom_text(data = month_from_lag(15, 1), aes(x = 361, label = month2),
+            size = 2.5, 
+            hjust = 0) +
+
+  geom_text(data = month_from_lag(1, 1), aes(x = -2, label = lag),
+            size = 2.5,
+            hjust = 1) +
+  scale_fill_divergent_discretised() +
+  scale_x_longitude(ticks = 60, expand = c(0.08, 0)) +
+
+  scale_y_continuous(breaks = NULL, expand = c(0, 0)) +
+  labs(fill = NULL, 
+       y = "Lag") +
+  wide_legend +
+  facet_wrap(~ month, labeller = labeller(month = labels_month)) +
+  theme(panel.background = element_blank()) +
+  tag_facets(position = list(x = 0.1, y = 1-.02))
+```
+
+Even though +S2 is no better than persistence at forecasting pan-Antarctic sea ice concentration anomalies, the quality of the forecast might vary between regions. 
+To analyse the spatial distribution of the model error, we computed ice concentration RMSE on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]` degrees thick.
+The median difference between the forecast and persistence RMSE is shown on @fig-rmse_lon, where negative values indicate that the model has lower RMSE than the benchmark.
+
+The skill shown by +S2 at February-March forecasts is evident by a band of negative values across almost all longitudes except around and east of the Antarctic Peninsula, where most of the Summer ice is left.
+This suggests that the positive skill comes from the model correctly forecasting no ice where there is never ice. 
+However, there are other regions of skilful forecasts.
+
+Forecasts initialised on January 1^st^ ([Fig.
+@fig-rmse_lon] panel a) shows RMSE values below persistence at lags larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea.
+These two regions are also relatively well forecasted from other months. 
+Notably, May forecasts of Weddell sea ice concentration anomalies ([Fig. @fig-rmse_lon] panel e) are skillful after about 20 days and remain so until December. 
+The Weddell sea region is also the region with the maximum erro --arround June irregardless of initialisation date.
 
@@
 
+------------------------------------------------------------------------
+
+We evaluated the sea
+
+The ACCESS-S2 model, the Australian Bureau of Meteorology's current seasonal prediction system, shows a significant low extent bias in its sea ice extent predictions, particularly during late summer and early autumn.
+This bias is attributed to a faster and longer melt season between January and March, and slower growth between March and April.
+This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems with the model's thermodynamic representation.
+
+Comparison with the previous ACCESS-S1 model reveals that while both models share the same formulation and thus similar equilibrium states, ACCESS-S1 had a smaller bias, especially at shorter forecast lead times.
+This difference suggests that the data assimilation system in ACCESS-S1 was correcting for this low sea ice bias.
+The analysis suggests that the data assimilation in ACCESS-S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea ice component runs virtually free, reverting to its biased equilibrium state.
+
+Further evaluation of ACCESS-S2's performance in forecasting sea ice anomalies reveals a poor performance even at short lead times, exhibiting much larger variability than observed.
+Quantitative analysis using root mean square error (RMSE) of sea ice concentration anomalies shows that ACCESS-S2 generally lacks skill at any lead time and for any month, with its performance often worse than a simple persistence forecast.
+
+Comparison with ACCESS-S1 using the Integrated Ice Edge Error (IIEE) metric confirms that ACCESS-S1 consistently demonstrates lower errors at short lead times, with the errors converging between the two models as the forecast progresses.
+This convergence timescale, ranging from two weeks to 160 days depending on the month, is indicative of the sea ice memory of initial conditions, particularly during October to March when data assimilation has minimal influence.
+
+Analysis of error spread further reveals that ACCESS-S1 consistently exhibits a smaller error spread at short lead times, indicating both smaller and more consistent initial condition errors.
+While both models show similar ensemble spread in individual forecasts, likely due to their shared model formulation and comparable perturbation schemes, ACCESS-S2 shows a larger spread in the mean error of individual forecasts, particularly at short lead times when data assimilation is less influential on sea ice initial conditions.
+
+------------------------------------------------------------------------
+
 # References
@@
 <!-- ```{r dataset_era} -->
+
 <!-- which_dataset <- "era5" -->
+
 <!-- ``` -->
@@
 <!-- ```{r fig-hindcast-extent-era5} -->
+
 <!-- which_dataset <- "era5" -->
+
 <!-- <<fig-hindcast-extent>> -->
+
 <!-- ``` -->
@@
 <!-- ```{r fig-hindcast-extent-bt} -->
+
 <!-- which_dataset <- "bt" -->
+
 <!-- <<fig-hindcast-extent>> -->
+
 <!-- ``` -->
@@
 <!-- ```{r fig-mean-growth} -->
+
 <!-- <<fig-mean-growth>> -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- <<fig-bias>> -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- <<<fig-extent-anom>> -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- <<fig-rmse>> -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- <<fig-iiee>> -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- <<fig-iiee-variance>> -->
+
 <!-- ``` -->

### 2024-12-06 R/ggplot.R

@@ -72,20 +72,18 @@ topo <- rcdo::cdo_topo(
     metR::ReadNetCDF(c(z = "topo")) |>
     _[, .(x = xgrid, y = ygrid, z)]
 
-get <- `$`
-
 sic_projection <- CDR() |>
     rcdo::cdo_seltimestep(1) |>
     rcdo::cdo_execute() |>
     ncdf4::nc_open() |>
     ncdf4::ncatt_get(varid = "crs") |>
-    get("proj_params")
+    _[["proj_params"]]
 
 
-contour <- ggplot2::StatContour$compute_group(topo, breaks = 0)
-geom_antarctica_path <- ggplot2::geom_path(data = contour, ggplot2::aes(x, y, group = group), inherit.aes = FALSE, colour = "black")
+contour_antarctica <- ggplot2::StatContour$compute_group(topo, breaks = 0)
+geom_antarctica_path <- ggplot2::geom_path(data = contour_antarctica, ggplot2::aes(x, y, group = group), inherit.aes = FALSE, colour = "black")
 
-geom_antarctica_fill <- ggplot2::geom_polygon(data = contour, ggplot2::aes(x, y, group = group), inherit.aes = FALSE, colour = "black", fill = "#FAFAFA")
+geom_antarctica_fill <- ggplot2::geom_polygon(data = contour_antarctica, ggplot2::aes(x, y, group = group), inherit.aes = FALSE, colour = "black", fill = "#FAFAFA")
 
 geomcoord_antarctica <- list(
     NULL,

### 2024-12-06 scripts/hindcast_extent.R

@@ -1,55 +1,50 @@
 library(furrr)
-workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 25))
+workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 2))
 message(workers, " workers")
 
-plan(multisession, workers = workers)
+plan(multicore, workers = workers)
 
 source(here::here("R/functions.R"))
 source(here::here("R/datasets.R"))
 
-
-
-forecast_times_s2 <- get_forecast_times("S2")
-
-all_files <- data.table::CJ(
-    forecast_time = forecast_times_s2,
+run <- function(model, measure) {
+  message(glue::glue("computing {measure} for {model}"))
+  forecast_times <- get_forecast_times(model)
+  
+  all_files <- data.table::CJ(
+    forecast_time = forecast_times,
     member = 1:9
-)
-
-
-
-furrr::future_pmap(all_files, \(forecast_time, member) {
-    S2_hindcast(forecast_time, member) |>
-        extent() |>
-        metR::ReadNetCDF("aice") |>
-        _[, `:=`(
-            lat = NULL, lon = NULL,
-            forecast_time = forecast_time,
-            member = member
-        )] |>
-        _[]
-}) |>
+  )
+  
+  hindcast_fun <- match.fun(glue::glue("{model}_hindcast"))
+  
+  fun_apply <- match.fun(measure)
+  
+  file <- here::here(glue::glue("data/derived/{model}_hindcast_{measure}.Rds"))
+  if (file.exists(file)) {
+    return(file)
+  }
+  
+  furrr::future_pmap(all_files, \(forecast_time, member) {
+    hindcast_fun(forecast_time, member) |>
+      fun_apply() |> 
+      metR::ReadNetCDF("aice") |>
+      _[, let(lat = NULL, lon = NULL,
+              forecast_time = forecast_time,
+              member = member)] |>
+      _[]
+  }) |>
     data.table::rbindlist() |>
-    saveRDS("data/derived/S2_hindcast_extent.Rds")
+    saveRDS(file)
+  
+  return(file)
+}
 
 
-forecast_times_s1 <- get_forecast_times("S1")
 
-all_files <- data.table::CJ(
-    forecast_time = forecast_times_s1,
-    member = 1:9
-)
+models <- c("S2", "S1")
+measures <- c("extent", "area")
 
-furrr::future_pmap(all_files, \(forecast_time, member) {
[…]

### 2024-12-06 scripts/pbs/error_metrics.sh

@@ -1,13 +1,13 @@
 #!/bin/bash
 
-#PBS -P dx2
+#PBS -P lo70
 #PBS -q normal
 #PBS -l ncpus=48
-#PBS -l mem=190GB
-#PBS -l walltime=03:00:00
+#PBS -l mem=128GB
+#PBS -l walltime=1:00:00
 #PBS -l wd
 #PBS -l jobfs=400GB
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-export PBS_WORKERS=100
-Rscript scripts/error_metrics.R > logs/error_metrics.log
+export PBS_WORKERS=48
+Rscript scripts/error_metrics.R 

### 2024-12-06 scripts/pbs/read_errors.sh

@@ -2,9 +2,9 @@
 
 #PBS -P lo70
 #PBS -q normal
-#PBS -l ncpus=48
-#PBS -l mem=190GB
-#PBS -l walltime=03:00:00
+#PBS -l ncpus=192
+#PBS -l mem=380GB
+#PBS -l walltime=10:00:00
 #PBS -l wd
 #PBS -l jobfs=400GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-export PBS_WORKERS=100
-Rscript scripts/hindcast_extent.R > logs/hindcast_extent.log
+export PBS_WORKERS=192
+Rscript scripts/read_errors.R

### 2024-12-06 scripts/read_errors.R

@@ -1,0 +1,81 @@
+sink(here::here("logs/read_errors"))
+print("hello!")
+library(furrr)
+workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 7))
+message(workers, " workers")
+
+plan(multicore, workers = workers)
+
+read_measures <- function(dir) {
+  files <- list.files(dir, include.dirs = FALSE, recursive = TRUE, full.names = TRUE) 
+  
+  dates <- paste0(here::here("data/derived"), "/(\\w*)/(\\w*)/(\\w*)/di_aice_(\\d{8})_e([\\dm]{2}).nc") |>
+    utils::strcapture(files,
+                      proto = list(
+                        measure = character(),
+                        version = character(),
+                        obs_dataset = character(),
+                        time_forecast = character(),
+                        member = character()
+                      ), perl = TRUE
+    ) |>
+    data.table::as.data.table() |>
+    _[, time_forecast := as.Date(time_forecast, format = "%Y%m%d")] |>
+    _[]
+  
+  # rmse_lon is 360 times larger than the other measures, so it 
+  # just doesnt fit into memory. :( 
+  if (dates$measure[1] == "rmse_lon") {
+    which_read <- dates[, which(version == "S2" & 
+                                  obs_dataset == "cdr" & member %in% c("01", "02", "03"))]
+    files <- files[which_read]
+    dates <- dates[which_read]
+  }
+  
+  future_map(seq_along(files), \(i) {
+    ncfile <- ncdf4::nc_open(files[[i]])
+    on.exit(ncdf4::nc_close(ncfile))
+    data <- try(metR::ReadNetCDF(ncfile, vars = c(value = "aice")), silent = TRUE)
+    
+    if (inherits(data, "try-error")) {
+      file.remove(files[[i]])
+      warning("file ", basename(files[[i]]), " deleted")
+      return(NULL)
+    }
+    
+    data |>
+      _[, let(lat = NULL)] |>
+      cbind(dates[i]) |>
+      _[]
+  }) |>
+    data.table::rbindlist()
+}
+
+
+rmse_dir <- here::here("data/derived", "rmse")
+
+rmse_lon_dir <- here::here("data/derived", "rmse_lon")
+
+iiee_dir <- here::here("data/derived", "iiee")
+
+gc(verbose = TRUE)
[…]

### 2024-12-13 analysis/paper/access-ice.qmd

@@
 ---
-title: "Untitled"
-format: gfm
+title: Evaluation of +S2 sea ice forecast and some other better title
+pdf-engine: pdflatex
+journal:
+  name: tc
+author:
+  - given_name: Elio 
+    surname: Campitelli
+    affiliation: "1,2"
+    email: elio.campitelli@monash.edu
+    corresponding: true
+  - given_name: Ariaan 
+    surname: Purich
+    affiliation: "1,2"
+  - given_name: Julie
+    surname: Arblaster
+    affiliation: "1,2"
+  - given_name: Eun-Pa
+    surname: Lim
+    affiliation: "3"
+  - given_name: Matthew
+    surname: Wheeler
+    affiliation: "3"
+  - given_name: Phillip
+    surname: Reid
+    affiliation: "3"
+
+affiliation:
+  - code: 1
+    address: School of Earth, Atmosphere and Environment, Monash University, Australia
+  - code: 2
+    address: Securing Antarctica’s Environmental Future, Monash University, Australia
+  - code: 3
+    address: Bureau of Meteorology
+
+abstract: |
+  The abstract goes here.
+  It can also be on _multiple lines_.
+bibliography: references.bib
+running:
+  title: +S2 sea ice forecast
+  author: Campitelli et al.
+# This section is mandatory even if you declare that no competing interests are present.
+competinginterests: |
+  The authors declare no competing interests.
+# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
+# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
+copyrightstatement: |
+  The author's copyright for this publication is transferred to institution/company. 
+### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
+### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
+### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
+availability:
+  #code: |
+  #  use this to add a statement when having only software code available
+  #data: |
+  #  use this to add a statement when having only data sets available
+  codedata: |
+    use this to add a statement when having data sets and software code available
+  sample: |
+    use this section when having geoscientific samples available
+videosupplement: |
+  use this section when having video supplements available
+authorcontribution: |
+  Contributions
+disclaimer: |
+  We like Copernicus.
+acknowledgements: |
+  Acknowledgements will gohere
+appendix: |
+  \section{Figures and tables in appendices}
+  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
+  \subsection{Option 1}
+  If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
+  They will be correctly named automatically.
+  \subsection{Option 2}
+  If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
+
+  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
+  `\appendixfigures` needs to be added in front of appendix figures
+  `\appendixtables` needs to be added in front of appendix tables
+
+  Please add `\clearpage` between each table and/or figure.
+format: copernicus-pdf
+
 execute: 
@@
   message: false
-  cache: false
+  cache: refresh
 filters:
   - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
-  - ../../_extensions/mloubout/critic-markup/critic-markup.lua
+  # - ../../_extensions/mloubout/critic-markup/critic-markup.lua
   - ../../_extensions/ute/search-replace/search-replace.lua
@@
   +CDR: NOAA/NSIDC CDRV4
-bibliography: references.bib
-editor: 
-  markdown: 
-    wrap: sentence
 ---
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, cache = FALSE}
 library(lubridate)
@@
 cdo_options_set("-L")
+```
+
+```{r helpers, cache = FALSE}
 source(here::here("R/functions.R"))
@@
 
-Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice had a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016].
-Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change.
-\[ref: that CESM run with forced winds\].
+Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016].
+Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change [@espinosa2024].
 
-Antarctic sea ice is in a remote location but the potential impacts of its variability and trends extend far into the lower latitudes, affecting ocean and atmospheric circulation, ocean carbon uptake and biological processes.
-It's likely that a reduction in sea ice extent would reduce the temperature gradient between high and low latitudes (and vice versa), which in turn influences the strength and location of the eddy-driven jet, which is a crucial component of weather events.
-For the northern hemisphere, it has been hypothesised that the reduction in sea ice in the northern hemisphere has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015].
+Antarctic sea ice is in a remote location but the potential impacts of its variability extends far into the lower latitudes, affecting the circulation of both the ocean and the atmosphere, ocean carbon uptake and biological processes.
+It's likely that a reduction in sea ice extent would reduce the meridional temperature gradient which in turn would reduce the strength and location of the eddy-driven jet (and vice versa for an increase in sea ice), which is a crucial component of weather events.
+In the northern hemisphere, it has been hypothesised that the reduction in Arctic sea ice has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015].
 There is some evidence of this effect in the Southern Hemisphere.
@@
 But models don't fully agree on the exact phase of the zonal wave response and observations don't show a clear signal [@rea2024].
-The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location [@ayres2019].
-It is also small compared with the direct effect of CO2 increase and Sea Surface Temperature warming [@ayres2019].
-On climate scales, it is thought that the signal of a reduction in Antarctic sea ice is like a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].
+The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location and relatively small compared with the direct effect of CO2 increase and Sea Surface Temperature (SST) warming [@ayres2019].
+On climate scales, it is thought that the temperature signal of a reduction in Antarctic sea ice mirrors a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].
 
-Correctly modelling Antarctic sea undsen sector and the Amundsen-Bellingshausen sector (Fice is not only necessary for process understanding and climate projections to inform adaptation strategies.
-Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent [@desilva2020 @wagner2020].
-However, Antarctic sea ice forecasts have been challenging both due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
-Dynamical seasonal forecasts of Summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.
+Correctly modelling Antarctic sea  is not only necessary for process understanding and climate projections to inform adaptation strategies.
+Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent such as scientific missions, fisheries and tourism [@desilva2020 @wagner2020].
+However, producing accurate Antarctic sea ice forecasts has been challenging due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
+Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.
 
-{==**S2** Introduce S2==}
-
-{==**Objective.** Evaluate S2.==}
+In this work we evaluate sea ice forecasts produced by +S2, the Australian Bureau of Meteorology subseasonal-to-seasonal prediction system. 
+This evaluation will inform future work with the model as a reseach tool and explore the potential of using it as a sea ice forecasting tool.
 
@@
 
-+S2 [@wedd2022] is the Australian Bureau of Meteorology's current seasonal prediction system that replaced +S1 [@hudson2017] in October 2021.
-Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
-The atmosphere has a N216 horizontal resolution(\~60km in the mid-latitudes) with 85 levels.
++S2 [@wedd2022] is the Australian Bureau of Meteorology current subseasonal-to-seasonal prediction system which replaced +S1 [@hudson2017] in October 2021.
+Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015] model components.
+The atmosphere has a N216 horizontal resolution (\~60km in the mid-latitudes) with 85 levels.
 The land model uses the same horizontal grid with 4 soil levels.
@@
 
-Both systems took atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts.
+Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts.
 The main difference between the two are the ocean and ice initial conditions.
-+S1 initial conditions came from from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) DA scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
++S1 initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
 +S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022.
 This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
-SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are larger than 0ºC.
+SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are over 0ºC.
 Relevant for this work, sea ice concentrations are not assimilated.
 
-For evaluation we use hindcast for the period 1981--2023{\>\>Check\<\<}.
-Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
-Climatology is smoothed with a 11 day running mean.
+The +S2 hindcast used in this study runs for the period 1981--2018.
+Each forecast consists of 9 ensemble members, 3 of which run for 279 days and 6 for 42 days.
+Anomalies are be taken with respect to the 1981--2011 climatology computed from the daily +S2 reanalysis and smoothed with a 11 day running mean.
+
+The +S1 hindcast runs for the period 1990--2012, with 7-member forecasts integrated for 217 days. 
+No reanalysis is available so we are not able to compute a climatology for this forecast, therefore we don't show metrics based on anomalies.
 
@@
 datasets <- list(cdr = CDR(),
-                 bt = BT(),
-                 osi = OSI(),
-                 era5 = ERA5()
-)
+                 osi = OSI())
 which_dataset <- "cdr"
@@
 
-There is not a lot of data on sea ice properties, especially for things like thickness, age, etc.
-However there are relatively reliable satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
-
-These data are not perfect.
-\cite{meier2019} Sensor sensitivity and resolution.
-Weather filtering.
-Land To account for observational uncertainty we performed the verification on other data sources.
-We find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
+For verification we use satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
+Different datasets derived using different algorithms provide different estimates, each with its own biases and uncertainties. 
+@meier2019 estimated that the inter-product uncertainty of sea ice extent is of the order of 1 million km^2^. 
+As will be shown below, this spread is minimal compared with the errors in the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
 
 We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset.
-Its sea ice concentration data are estimated taking the maximum value of the NASA Team \[ref\] and NASA Bootstrap [@comiso2023] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
-Both algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
+It takes the maximum value of the NASA Team [@cavalieri1984] and NASA Bootstrap [@comiso2023] sea ice concentration products to reduce their low concentration bias [@meier2014; @meier2021].
+Both source algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
 The data has a resolution of 25 by 25km and daily from 1978 onwards.
 
-Results based on two other different datasets are provided as supplementary material:
-
-1.  The NASA Bootstrap [@comiso2023] product is... it is not an independent dataset from NOAA/NSIDC CDRV4.
-
-2.
-The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) \[ref\] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
+The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [@OSISAF] is another satellite-derived sea ice concentration product.
+It is based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
+Figures prepared with this dataset are provided in the supplementary material.
 
@@
 
-### Sea ice extent
-
 Sea Ice Extent is defined as the area of ocean covered with at least 15% ice.
@@
 
-Sea Ice Extent is a rough global measure, but a model could have relatively accurate extent of ice but with different distributions.
-We use two other measures to account for these errors.
+Pan-Antarctic Sea Ice Extent serves as a rough global measure of the amount of sea ice but it doesn't take into account the spatial distribution. 
+A model could have relatively accurate extent of ice but with different distributions.
+To account for location errors, we computed the Root Mean Squared Error (RMSE) of sea ice concentration anomalies and the Integrated Ice Edge Error [IIEE; @goessling2016].
 
-We compute Root Mean Square Error of sea ice concentration anomalies.
+We compute RMSE as the square root of the average squared differences between forecasted and observed sea ice concentration anomalies. 
+We compute a pan-Antarctic RMSE by averaging over the whole domain, and also a zonally-varying RMSE by averaging over 24 15° longitude slices around Antarctica.
 
-We also compute the Integrated Ice Edge Error (IIEE) [@goessling2016].
-This is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
+The IIEE is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
 That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.
+
+All error measures were computed on the +CDR domain grid and projection to which model output was bilinearly interpolated.
 
@@
 @fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR +S2 shows a severe low extent bias, especially in the late summer-early autumn.
-This is due primarily to a faster and longer melt season between January and March and slower growth during March and April This is then balanced with faster growth between May and July ([Fig.
-@fig-mean-growth]).
-Many S2S systems exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
+This is due primarily to faster melt during a longer melt season between January and March and slower growth during March and April. 
+This is then balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
+Many sea ice models exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
 
-Comparing +S2 with +S1, the latter has a smaller bias, especially at low lags ([Fig.
-@fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig.
-@fig-mean-growth]).
++S1 has a smaller low bias than +S2, especially at low lags ([Fig. @fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig. @fig-mean-growth]).
 At larger lags, +S1's bias in summer and autumn is very similar to +S2's.
-This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1.
+This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1 and not in +S2.
 
-At long lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
+At large lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
 Therefore we can estimate the latter using the hindcasts with the largest possible lag, which is shown in triangles in @fig-hindcast-extent for the same dates as the initial conditions.
@@
 
-The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to both having the same model formulation.
-From June to October, in +S2 circles move away from triangles and towards observations, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and affecting the initial conditions.
+The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to them having the same model formulation.
+From June to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and improving the initial conditions.
 The rest of the year, there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.
+In +S1, on the other hand, circles are very close to observations year-round.
 
@@
                    date_labels = "%b", expand = c(0, 0)) +
-  scale_y_continuous(NULL, labels = labels_extent) +
+  scale_y_continuous(NULL, labels = \(x) paste0(labels_extent(x), "/day")) +
   scale_color_models
-```
-
-```{r climatology}
-climatologies <- c(datasets,
-                   S2 = S2_reanalysis()) |>
-  lapply(\(x) climatology(x) |>
-           cdo_ymonmean() |> 
-           cdo_execute()
-  ) |>
-  lapply(\(x) ReadNetCDF(x, "aice")) |>
-  rbindlist(idcol = "model") 
 ```
@@
   climatology() |> 
+  cdo_ymonmean() |>
+  cdo_execute() |> 
   ReadNetCDF(c(obs = "aice")) |> 
-  setnames("time", "time2")
+  setnames("time", "time2") |> 
+  _[, time2 := update(time2, day = 1)]
 
@@
 #| fig-height: 6
-#| fig-cap: 1-month lag +S2 forecast bias compared with NSIDC.
+#| fig-cap: Mean sea ice concentrations 1-month lag monthly mean +S2 forecast bias compared with NSIDC.
 
@@
 
-@fig-bias shows the difference in monthly mean sea ice concentrations between +CDR and +S2 reanalysis.
-From October to May, the model underestimates sea ice concentrations pretty much everywhere there is ice except for the deep Weddell Sea in April and May, where sea ice concentrations saturate to 1.
-In winter, the differences are mostly on the sea ice edge, with slight positive bias in XXX and negative bias around the Indian Ocean sector.
+@fig-bias shows the difference of monthly mean sea ice concentrations between +CDR and +S2 hindcasts at the shortest lag. 
+From October to May, the model underestimates sea ice concentrations pretty much everywhere there is ice except for the deep Weddell Sea in April and May, where sea ice concentrations saturate to 1 both in the observations and forecasts.
+In winter, the differences are mostly on the sea ice edge, with light positive bias in the African sector of East Antarctica and negative bias around the Indian Ocean sector which partially compensate, resulting in the near-zero extent bias seen in those months ([Fig.
+@fig-hindcast-extent]).
 
 ### Anomalies
-
-{==intro connecting subsections==}
 
@@
 @fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lags.
-The anomalies in this case were computed with respect of the climatology of each lag, which is a way of bias-correction.
+The anomalies in this case were computed with respect of the climatology of each lag, which works as a bias-correction method. 
 Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at lag 3.
@@
 
+```{r clim_std}
+clim_std <- datasets |> 
+  lapply(\(x) x |> 
+           anomalies() |> 
+           cdo_sqr() |> 
+           cdo_fldmean() |> 
+           cdo_sqrt() |> 
+           cdo_execute() |> 
+           ReadNetCDF(c(std = "aice")) |> 
+           _[, let(lat = NULL, lon = NULL)]
+  ) |> 
+  rbindlist(idcol = "obs_dataset") |> 
+  _[, time := as.Date(time)] |> 
+  _[std == 0, std := NA] |> 
+  _[std < 0.05, std := NA]
+
+```
+
 ```{r fig-rmse}
@@
   _[measure == "rmse"] |>
+  _[, time2 := update(time, year = 2000)] |> 
+  # merge(clim_std) |>
   _[obs_dataset ==  which_dataset] |>
@@
   geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
+
+  # geom_line(aes(y = clim_std)) +
   scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
@@
 @fig-rmse shows the median and 95% range of RMSE of sea ice concentration anomalies for +S2 forecasts compared with a benchmark of persistence.
-This figure uses +CDR data, but the results are nearly identical compared with ERA5 or Bootstrap.
-Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a reference?=
-=}.
+Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, after when it is statistically useful.
 Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
-The only exception is the RMSE around February.
+The only exception is around February, where the model has lower RMSE than the persistence forecast at virtually every lag.
 
@@
 
-```{r rmse_lon}
-rmse_lon <- here::here("data/derived/rmse_lon.Rds") |> 
+```{r rmse_lon_mean}
+rmse_lon_mean <- here::here("data/derived/rmse_lon.Rds") |> 
   readRDS() |> 
   _[version != "S1"] |> 
-  _[, lag := as.numeric(as.Date(time) - time_forecast)]
-
-rmse_lon[, base := value[version == "persistence"], 
-         by = .(time, lon, lag, obs_dataset)]
-
-rmse_lon_mean <- rmse_lon |> 
+  _[, lag := as.numeric(as.Date(time) - time_forecast)] |> 
+  _[, base := value[version == "persistence"], 
+    by = .(time, lon, lag, obs_dataset)] |> 
   _[version != "persistence"] |>
@@
   facet_wrap(~ month, labeller = labeller(month = labels_month)) +
-  theme(panel.background = element_blank()) +
-  tag_facets(position = list(x = 0.1, y = 1-.02))
+  theme(panel.background = element_blank(), 
+        tagger.panel.tag.background = element_rect(colour = "white", fill = "white")) +
+  tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1))
 ```
@@
 Even though +S2 is no better than persistence at forecasting pan-Antarctic sea ice concentration anomalies, the quality of the forecast might vary between regions. 
-To analyse the spatial distribution of the model error, we computed ice concentration RMSE on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]` degrees thick.
+To analyse the spatial distribution of the model error, we computed sea ice concentration anomalies RMSE on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]`º wide.
 The median difference between the forecast and persistence RMSE is shown on @fig-rmse_lon, where negative values indicate that the model has lower RMSE than the benchmark.
 
-The skill shown by +S2 at February-March forecasts is evident by a band of negative values across almost all longitudes except around and east of the Antarctic Peninsula, where most of the Summer ice is left.
-This suggests that the positive skill comes from the model correctly forecasting no ice where there is never ice. 
+The skill shown by +S2 at February-March forecasts is evidenced by a band of negative values across almost all longitudes. 
+The only region where February-March sea ice is not well forecasted is east of the Antarctic Peninsula, which is the only region with consistent summer sea ice.
+This suggests that the positive skill around the summer minimum comes from the model correctly forecasting no ice where no ice is ever found. 
 However, there are other regions of skilful forecasts.
@@
 Forecasts initialised on January 1^st^ ([Fig.
-@fig-rmse_lon] panel a) shows RMSE values below persistence at lags larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea.
+@fig-rmse_lon] panel a) show RMSE values below persistence at lags larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea.
 These two regions are also relatively well forecasted from other months. 
-Notably, May forecasts of Weddell sea ice concentration anomalies ([Fig. @fig-rmse_lon] panel e) are skillful after about 20 days and remain so until December. 
-The Weddell sea region is also the region with the maximum erro --arround June irregardless of initialisation date.
+Notably, May forecasts of Weddell sea ice concentration anomalies ([Fig. @fig-rmse_lon] panel e) are skilful after about 20 days and remain so until December. 
+The Weddell Sea is also the region with the maximum error --around June regardless of initialisation date.
 
@@
 This error measure is shown in @fig-iiee for all lags and forecasts initialised at the first of every month.
-+S1 has lower error at short lags at all months, with the errors converging as the forecast goes on.
-The time to convergence depends on the month and it can be as fas as two weeks June and July to as large as 160 days for forecasts initialised in February.
-Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the the memory of sea ice to initial conditions; at least from October to March when the data assimilated form the other components has little to no influence on sea ice.
++S1 has lower error at short lags at all months, with the errors converging to +S2 as the forecast goes on.
+The time to convergence depends on the month and it can be as short as two weeks June and July to as long as 160 days for forecasts initialised in February.
+Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the the memory of sea ice to initial conditions --at least from October to March when the data assimilated form the other components has little to no influence on sea ice.
 
 Also evident in @fig-iiee is the difference in the error spread at short lags between +S2 and +S1.
-In all month +S1 shows a small error spread at lag 1, indicating that the error in the initial conditions not only is small, but also fairly constant.
+In all months +S1 has much narrower error spread at lag 1. 
+The error in the initial conditions is not only small, but also fairly constant.
 This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
@@
 #| fig-cap: Mean spread of IIEE at different lags for different models. 
-#| fig.height: 9
-#| fig.width: 4
+#| fig.height: 6
+#| fig.width: 8
 labels_extent_2 <- function(x) {
@@
   scale_color_models +
-facet_grid(month ~ component, 
+  ggh4x::facet_nested_wrap(month ~ component, ncol = 6, 
+                           remove_labels = "all",
              labeller = labeller(month = labels_month, 
@@
 
-The large initial error spread could be due either to large spread of ensemble members or due to a large spread of individual forecasts.
+The large initial error spread could be due either to large spread of ensemble members or to large spread of individual forecasts.
 @fig-iiee-variance splits the IIEE variance for each lag into the mean variance of each individual forecast and the variance of the mean error of each individual forecast, which adds up to the total variance.
@@
 
-------------------------------------------------------------------------
+Sea ice forecasts from the +S2 model show a significant low extent bias, particularly during late summer and early autumn. 
+This bias is attributed to a faster and longer melt season between January and March, and slower growth between March and April.
+This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems either with the model's thermodynamic representation or with short wave radiation forcing, as shown in other climate models.
+Even though +S2 shares the same model formulation with +S1, the latter does not suffer from this bias. 
+This is due to the assimilation of sea ice concentration anomalies into the initial conditions, which successful corrects for the negative bias in the model. 
+The analysis suggests that the data assimilation system in +S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea ice component runs virtually free, reverting to its biased equilibrium state.
 
-We evaluated the sea
+Analysis of the error spread shows that +S2 initial conditions from December to May not only have large errors, but that the initial error spread is very large compared with +S1. 
+This spread is not due to the perturbation scheme, since the mean error variance for individual forecasts is low and comparable with +S1. 
+Instead, it is due to large variance of the mean error of individual forecasts, which is comparable to the climatology spread. 
+This is further evidence that individual initial conditions are not being affected by the data assimilation scheme.
 
-The ACCESS-S2 model, the Australian Bureau of Meteorology's current seasonal prediction system, shows a significant low extent bias in its sea ice extent predictions, particularly during late summer and early autumn.
-This bias is attributed to a faster and longer melt season between January and March, and slower growth between March and April.
-This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems with the model's thermodynamic representation.
+Based on the observation that +S2 sea ice initial conditions are essentially not initialised, comparing its forecasts with +S1's allows us to estimate the time-scale for which initial conditions are important.
+This highlights February initial conditions as a crucial for determining sea ice evolution at least up to late June. 
+Arctic sea ice forecast also show greater sensitivity to initial conditions in (boreal) summer compared with (boreal) winter [@day2014; @bunzel2016], so similar mechanism might be playing a role.
 
-Comparison with the previous ACCESS-S1 model reveals that while both models share the same formulation and thus similar equilibrium states, ACCESS-S1 had a smaller bias, especially at shorter forecast lead times.
-This difference suggests that the data assimilation system in ACCESS-S1 was correcting for this low sea ice bias.
-The analysis suggests that the data assimilation in ACCESS-S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea ice component runs virtually free, reverting to its biased equilibrium state.
+Although forecasts are not skilful for forecasting panantarctic sea ice concentrations, there are some areas where the model does show skill. 
+The Weddell Sea, and the Ross Sea to a lesser extent are particularly well forecasted between June and November for forecasts initialised from January to June. 
+These regions have been recognised as a regions of high predictability thanks to persistent and eastwardly advected upper ocean heat content anomalies [@bushuk2021] 
+Sea surface temperatures are assimilated by +S2 only in areas with temperatures greater than 0°C and errors don't show a clear easterly-propagating signal, so it is not clear if +S2 is leveraging the same source of predictability.
 
-Further evaluation of ACCESS-S2's performance in forecasting sea ice anomalies reveals a poor performance even at short lead times, exhibiting much larger variability than observed.
-Quantitative analysis using root mean square error (RMSE) of sea ice concentration anomalies shows that ACCESS-S2 generally lacks skill at any lead time and for any month, with its performance often worse than a simple persistence forecast.
-
-Comparison with ACCESS-S1 using the Integrated Ice Edge Error (IIEE) metric confirms that ACCESS-S1 consistently demonstrates lower errors at short lead times, with the errors converging between the two models as the forecast progresses.
-This convergence timescale, ranging from two weeks to 160 days depending on the month, is indicative of the sea ice memory of initial conditions, particularly during October to March when data assimilation has minimal influence.
-
-Analysis of error spread further reveals that ACCESS-S1 consistently exhibits a smaller error spread at short lead times, indicating both smaller and more consistent initial condition errors.
-While both models show similar ensemble spread in individual forecasts, likely due to their shared model formulation and comparable perturbation schemes, ACCESS-S2 shows a larger spread in the mean error of individual forecasts, particularly at short lead times when data assimilation is less influential on sea ice initial conditions.
-
-------------------------------------------------------------------------
+-----
 

### 2025-02-19 scripts/build_ensemble.R
Computes climatology

@@ -54,61 +54,3 @@ build_ensemble_S2 <- function(forecasts, nominal_date) {
 furrr::future_imap(dates, build_ensemble_S2) |> 
   data.table::rbindlist() |> 
   saveRDS(here::here("data/derived/hindcast_s2.Rds"))
-
-library(rcdo)
-source(here::here("R/datasets.R"))
-
-future::plan("multicore", workers = 24)
-
-dates <- get_forecast_times("S2")
-
-dates <- dates[lubridate::day(dates) %in% 1]
-
-dates <- lapply(dates, \(x) x - c(2, 1, 0)) |> 
-  setNames(dates)
-
-# The first time doesn't have previous days
-dates <- dates[-1]
-
-
-process_ice <- function(file, shift, file_out) {
-  if (file.exists(file_out)) return(file_out)
-
-  file |> 
-    rcdo::cdo_selname("aice") |>
-    remap_cdr() |>
-    cdo_shifttime(shiftValue = paste0(shift, "day")) |> 
-    cdo_del29feb() |> 
-    cdo_execute(output = file_out)
-}
-
-get_file <- function(forecast_time, member) {
-  paste0(
-    "/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/e0",
-    member,
-    "/di_aice_",
-    format(as.Date(forecast_time), "%Y%m%d"),
-    "_e0",
-    member,
-    ".nc"
-  )  
-}
-
-build_ensemble_S2 <- function(forecasts, nominal_date) {
-  folder <- here::here("data/derived/hindcast/S2", nominal_date)
-  dir.create(folder, showWarnings = FALSE, recursive = TRUE)
-  
-  data <- data.table::CJ(original_forecast_date = forecasts, original_member = 1:3) |> 
-    _[, file := get_file(original_forecast_date, original_member)] |> 
-    _[, member := 1:9] |> 
-    _[, forecast_date := as.Date(nominal_date)] |> 
-    _[, file_out := file.path(folder, paste0(formatC(member, width = 2, flag = "0"), ".nc"))] |>
-    _[, process_ice(file, as.numeric(forecast_date - original_forecast_date), file_out),
-      by = .(forecast_date, member, original_forecast_date, original_member)]
-}
-
-
-furrr::future_imap(dates, build_ensemble_S2) |> 
-  data.table::rbindlist() |> 
-  saveRDS(here::here("data/derived/hindcast_s2.Rds"))
-

### 2025-02-19 scripts/hindcast_climatology.R

@@ -1,25 +1,53 @@
-library(furrr)
-
-plan(multisession, workers = 12)
-
+library(rcdo)
+library(data.table)
+library(metR)
 source(here::here("R/datasets.R"))
 
-times <- get_forecast_times("S2")
+future::plan("multicore", workers = 12)
 
-nsidc_grid <- here::here("data/raw/nsidc_grid.txt")
+file <- get_forecast_times("S2") |> 
+  Filter(f = \(x) lubridate::day(x) == 1 & lubridate::month(x) == 1) |> 
+  _[2] |> 
+  hindcast(model = "S2", members = 1)
 
-months <- 1:12
+ndays <- as.numeric(cdo_ntime(file) |> 
+             cdo_execute())
 
-furrr::future_walk(months, \(month) {
-  month <- formatC(month, width = 2, flag = "0")
-  files <- here::here(glue::glue("data/derived/hindcast_ensmean/di_aice_*{month}01_emm.nc")) |> 
-    Sys.glob()
-  files_chr <- paste0(files, collapse = " ")
+bytes_per_day <- file.size(file)/ndays
+
+year_clim <- c(1990, 2012)
+
+compute_clim <- function(model, month) {
+  file <- here::here("data/derived/climatology", model, pad_number(month), "em.nc")
+  dir.create(dirname(file), showWarnings = FALSE, recursive = TRUE)
   
-  out <- here::here(glue::glue("data/derived/hindcast_climatology/mo_aice_1981{month}01_mean.nc"))
-  dir.create(dirname(out), showWarnings = FALSE, recursive = TRUE)
-  glue::glue("cdo -L -O -s ensmean [ -monmean : {files_chr} ] {out}") |> 
-    system()
-})
+  dates <- get_forecast_times(model) |> 
+    Filter(f = \(x) lubridate::day(x) == 1 & lubridate::month(x) == month & lubridate::year(x) %between% year_clim) 
+  
+  # the 7th member has some weird jumps and inconsistencies. 
+  if (model == "S1" && month == 1) {
+    members <- c(1:6, 8:9)
+  } else {
+    members <- 1:9
+  }
+  forecasts <- hindcast(dates, model, 1)
+  
+  # Some forecast have less than the correct number of days. 
+  # Throw them out
+  size <- forecasts |> 
+    file.size()
+  
+  ndays <- floor(size/bytes_per_day)
+  
+  forecasts[ndays >= 215] |> 
[…]

### 2025-02-19 scripts/pbs/build_ensemble.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
@@ -9,7 +9,7 @@
 #PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
 #PBS -N build_ensemble
 
 # Load module, always specify version number.

### 2025-02-19 scripts/pbs/hindcast_climatology.sh

@@ -1,13 +1,16 @@
 #!/bin/bash
 
-#PBS -P dx2
+#PBS -P lo70
 #PBS -q normal
 #PBS -l ncpus=48
-#PBS -l mem=190GB
-#PBS -l walltime=03:00:00
+#PBS -l mem=64GB
+#PBS -l walltime=01:00:00
 #PBS -l wd
-#PBS -l jobfs=400GB
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7
+#PBS -l jobfs=20GB
+#PBS -o logs/
+#PBS -e logs/
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
+#PBS -N hindcast_climatolgy
 
 # Load module, always specify version number.
 module load R/4.3.1

### 2025-02-19 scripts/hindcast_extent.R
Computes extent of climatology

@@ -1,5 +1,5 @@
 library(furrr)
-workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 2))
+workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 7))
 message(workers, " workers")
 
 plan(multicore, workers = workers)
@@ -9,42 +9,45 @@ source(here::here("R/datasets.R"))
 
 run <- function(model, measure) {
   message(glue::glue("computing {measure} for {model}"))
-  forecast_times <- get_forecast_times(model)
+  forecast_times <- get_forecast_times(model) |> 
+    Filter(f = \(x) data.table::mday(x) == 1)
   
   all_files <- data.table::CJ(
     forecast_time = forecast_times,
     member = 1:9
-  )
-  
-  hindcast_fun <- match.fun(glue::glue("{model}_hindcast"))
+  ) |> 
+    _[, file := hindcast(forecast_time, model, member), by = .(forecast_time, member)] |> 
+    na.omit()
   
   fun_apply <- match.fun(measure)
   
-  file <- here::here(glue::glue("data/derived/{model}_hindcast_{measure}.Rds"))
-  if (file.exists(file)) {
-    return(file)
-  }
+  files <- furrr::future_pmap_chr(all_files, \(forecast_time, member, file) {
+    extent <- file |> 
+      fun_apply()
+    
+    extent2 <- try(metR::GlanceNetCDF(extent, "aice"))
+    
+    if (inherits(extent2, "try-error")) {
+      file.remove(extent)
+      extent <- file |> 
+        fun_apply() 
+    }
+    extent
+  }) 
   
-  furrr::future_pmap(all_files, \(forecast_time, member) {
-    hindcast_fun(forecast_time, member) |>
-      fun_apply() |> 
-      metR::ReadNetCDF("aice") |>
-      _[, let(lat = NULL, lon = NULL,
-              forecast_time = forecast_time,
-              member = member)] |>
-      _[]
-  }) |>
-    data.table::rbindlist() |>
-    saveRDS(file)
-  
-  return(file)
+  all_files[, file := files]
 }
 
 
 
 models <- c("S2", "S1")
-measures <- c("extent", "area")
+measures <- c("extent")
 
-data.table::CJ(model = models, measure = measures) |> 
-  purrr::pmap(run)
+files <- data.table::CJ(model = models, measure = measures) |> 
+  _[order(-model)] |> 
[…]

### 2025-02-19 scripts/pbs/hindcast_extent.sh

@@ -2,12 +2,16 @@
 
 #PBS -P lo70
 #PBS -q normal
-#PBS -l ncpus=48
-#PBS -l mem=190GB
+#PBS -l ncpus=7
+#PBS -l mem=64GB
 #PBS -l walltime=03:00:00
 #PBS -l wd
-#PBS -l jobfs=400GB
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5
+#PBS -l jobfs=100GB
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
+#PBS -N hindcast_extent
+#PBS -o logs/
+#PBS -e logs/
+
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -18,5 +22,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-export PBS_WORKERS=100
+export PBS_WORKERS=4
 Rscript scripts/hindcast_extent.R > logs/hindcast_extent.log

### 2025-02-19 scripts/error_metrics.R
Computes error metrics

@@ -1,13 +1,19 @@
-logfile <- here::here("logs/error_metrics.log")
+logfile <- here::here("logs/error_metrics.log")q
 log <- function(text) {
-  write(paste0(format(lubridate::now()), ": ", text), logfile, append = TRUE)
+  messg <- paste0(format(lubridate::now()), ": ", text)
+  if (interactive()) {
+    message(messg)
+  } else {
+    write(messg, logfile, append = TRUE)    
+  }
 }
 on.exit(log("exiting"))
 writeLines("", logfile)
 log("Booting up")
+
 library(furrr)
 
-workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 25))
+workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 2))
 plan_type <- "multicore"
 log(glue::glue("setting up {workers} workers in a {plan_type} plan"))
 
@@ -20,8 +26,8 @@ cdo_options_set("-L")
 library(data.table)
 library(lubridate)
 library(ncdf4)
-source("R/functions.R")
-source("R/datasets.R")
+source(here::here("R/functions.R"))
+source(here::here("R/datasets.R"))
 
 # Force evaluation now and not in the workers.
 obs <- list(
@@ -150,13 +156,9 @@ cdo_rmse_lon <- function(file, obs, n = 15, output) {
 compute_metrics <- function(forecast_time, member, version, obs_dataset) {
   tick <- lubridate::now()
   cdo_options_set(c("-L"))
-  # file <- files[i, ]
   
-  if (version == "S2") {
-    file <- S2_hindcast(forecast_time, member)
-  } else if (version == "S1") {
-    file <- S1_hindcast(forecast_time, member)
-  }
+  file <- hindcast(forecast_time, model = version, members = member)
+  model_climatology <- here::here("data/derived/climatology", version, pad_number(month(forecast_time)), "em.nc")
   
   file_template <- paste0("di_aice_", format(forecast_time, "%Y%m%d"), "_e0", member, ".nc")
   
@@ -197,16 +199,16 @@ compute_metrics <- function(forecast_time, member, version, obs_dataset) {
   
   if (!file.exists(rmse_out)) {
     file |>
-      cdo_ydaysub(S2_reanalysis() |> climatology()) |>
+      cdo_ydaysub(model_climatology) |>
       cdo_rmse(observation_part_anomaly) |>
[…]

### 2025-02-19 scripts/pbs/error_metrics.sh

@@ -3,20 +3,19 @@
 #PBS -P lo70
 #PBS -q normal
 #PBS -l ncpus=48
-#PBS -l mem=128GB
-#PBS -l walltime=1:00:00
+#PBS -l mem=64GB
+#PBS -l walltime=8:00:00
 #PBS -l wd
-#PBS -l jobfs=400GB
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5
+#PBS -l jobfs=100GB
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
+#PBS -N error_metrics
+#PBS -o logs/
+#PBS -e logs/
 
 # Load module, always specify version number.
 module load R/4.3.1
 module load cdo/2.4.3
 
-# Must include `#PBS -l storage=scratch/ab12+gdata/yz98` if the job
-# needs access to `/scratch/ab12/` and `/g/data/yz98/`. Details on:
-# https://opus.nci.org.au/display/Help/PBS+Directives+Explained
-
 # Run R application
-export PBS_WORKERS=48
+export PBS_WORKERS=24
 Rscript scripts/error_metrics.R 

### 2025-02-19 R/datasets.R
Adds... stuff

@@ -220,7 +220,7 @@ S2_reanalysis <- function() {
     rcdo::cdo_execute(output = file, options = "-L")
 }
 
-hindcast <- function(model, forecast_times, members = 1:9) {
+hindcast <- function(forecast_times, model, members = 1:9) {
   if (model == "S2") {
     S2_hindcast(forecast_times, members)
   } else {

### 2025-02-19 R/functions.R

@@ -176,3 +176,24 @@ cdo_extent <- function(file, threshhold = 0.15) {
         rcdo::cdo_gtc(threshhold) |>
         rcdo::cdo_fldint()
 }
+
+
+
+pad_number <- function(x, pad = 2) formatC(x, width = pad, flag = "0")
+
+
+
+correlate <- function(x, y, signif = 2, sep = "\ ", ...) {
+  correlation <- stats::cor.test(x, y, ...)
+  out <- with(correlation,
+              list(estimate = estimate,
+                   p.value = p.value,
+                   low = conf.int[1],
+                   hig = conf.int[2]))
+  
+  out$text <- with(out,
+                   paste0(round(estimate, signif), sep, "(CI:\ ",
+                          round(low, signif), "\ --\ ", round(hig, signif), ")"))
+  
+  out
+}

### 2025-02-19 analysis/paper/access-ice.qmd
Integrates some changes into figures

@@
 ```{r hindcast_extent}
-hindcast_extent <- rbind(S1 = readRDS(here::here("data/derived/S1_hindcast_extent.Rds")),
-                         S2 = readRDS(here::here("data/derived/S2_hindcast_extent.Rds")), 
-                         idcol = "model") |>
+hindcast_extent <- fread(here::here("data/derived/hindcast_extent.csv")) |> 
   _[, model := relevel(factor(model), "S2")] |> 
-  _[, lag := as.numeric(as.Date(time) - forecast_time)] |> 
-  _[lag > 0] |> 
-  _[!(month(forecast_time) == 1 & member == 7 & model == "S1")]  # This member is wrong
-```
-
-```{r}
-hindcast_area <- rbind(S1 = readRDS(here::here("data/derived/S1_hindcast_area.Rds")),
-                       S2 = readRDS(here::here("data/derived/S2_hindcast_area.Rds")), 
-                       idcol = "model") |>
-  _[, model := relevel(factor(model), "S2")] |> 
-  _[, lag := as.numeric(as.Date(time) - forecast_time)] |> 
+  _[, lag := as.numeric(as.Date(time) - as.Date(forecast_time))] |> 
   _[lag > 0] |> 
@@
 months <- 1:12
+
 hindcast_clim <- lapply(setNames(months, months), \(month) {
-  here::here(glue::glue("data/derived/hindcast_climatology/mo_aice_1981{zeropad(month)}01_mean.nc")) |> 
+  here::here(glue::glue("data/derived/climatology/S2/{zeropad(month)}/em.nc")) |> 
+    cdo_selmonth(month) |> 
+    cdo_execute() |> 
     ReadNetCDF("aice")
@@
   rbindlist(idcol = "forecast_month") |> 
-  _[, forecast_month := as.numeric(forecast_month)] |> 
-  _[, forecast_time := lubridate::make_datetime(1981, forecast_month, 01)] |> 
-  _[, time := lubridate::floor_date(time, "month")] |> 
-  _[, lag := mondf(forecast_time, time)] |> 
-  _[lag == 0] |> 
-  _[, time2 := update(time, year = 2000)]
+  _[, forecast_month := as.numeric(forecast_month)]
+
+hindcast_clim <- hindcast_clim |> 
+  copy() |> 
+  _[, .(aice = mean(aice)), by = .(month(time), ygrid, xgrid, forecast_month)] |> 
+  _[forecast_month == month]
 
@@
   ReadNetCDF(c(obs = "aice")) |> 
-  setnames("time", "time2") |> 
-  _[, time2 := update(time2, day = 1)]
+  _[, month := month(time)]
 
 hindcast_clim <- hindcast_clim |> 
-  merge(obs_clim) |> 
-  _[, month := month(time)] 
+  merge(obs_clim)
 ```
@@
   geom_antarctica_fill +
-  facet_wrap(~ lubridate::month(time, label = TRUE)) +
+  facet_wrap(~ month) +
   wide_legend +
@@
 monthly_extent <- extent_daily |> 
+  _[dataset == which_dataset] |> 
   _[aice == 0, aice := NA]  |> 
@@
 
+s1_range <- hindcast_extent[model == "S1"][, range(time)]
+s2_range <- hindcast_extent[model == "S2"][, range(time)]
+
 hindcast_extent |> 
   copy() |> 
-  # _[model == "S2"] |> 
   _[, aice := aice - mean(aice[year(time) %between% c(1981, 2011)]), 
     by = .(lag, model, time = update(time, year = 2000))] |> 
-  _[, .(aice = mean(aice)), by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
-  _[, lag := months_difference(time, forecast_time) + 1] |> 
-  _[lag %in% c(1, 3, 5, 7)] |> 
-  ggplot(aes(time, aice)) +
-  geom_line(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
-              _[dataset == which_dataset],
-            aes(group = dataset, color = dataset)) +
-  # geom_smooth(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
-  #               _[dataset == "cdr"], 
-  #             aes(group = dataset, color = dataset),
-  #             se = FALSE, span = 10*12/N, n = 366) +
-  geom_line(aes(color = model))  +
-  # geom_smooth(aes(color = model), se = FALSE, span = 10*12/N, n = 366) +
+  _[, .(forecast = mean(aice)), by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
+  _[, lag := months_difference(time, forecast_time)] |> 
+  _[, merge(.SD, monthly_extent, all = TRUE), by = .(model, lag)] |> 
+  _[lag %in% (c(1, 3, 5, 7) - 1)] |>
+  _[time %between% s2_range] |> 
+  ggplot(aes(time)) +
+  geom_line(aes(y = aice, group = dataset, color = dataset)) +
+  geom_line(aes(y = forecast, color = model))  +
+  geom_text(data = \(x) x[time %between% s1_range, correlate(aice, forecast, use = "complete.obs"), by = .(lag, model)],
+            aes(label = text), x = -Inf, y = Inf, 
+            vjust = 1.2, hjust = -0.2) +
   scale_y_continuous(NULL,
@@
   scale_x_date(NULL, expand = c(0, 0)) +
-  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Lag: ", x))) +
+  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Month: ", x))) +
   scale_color_models

### 2025-02-24 analysis/paper/access-ice.qmd
udaptes error metrics

@@
 errors <- rbind(
-  readRDS(here::here("data/derived/rmse.Rds")) |>
-    _[version != "S1"], 
+  readRDS(here::here("data/derived/rmse.Rds")),
   readRDS(here::here("data/derived/iiee.Rds"))
@@
 ```
+
+::: notes
+
+Use climatology
+
+monthly rmse/corelation at gridpoints per lead time.
+
+:::
 

### 2025-02-24 analysis/paper/access-ice.tex
udaptes error metrics

@@
 %
-\documentclass[
-  letterpaper,
-  DIV=11,
-  numbers=noendperiod]{scrartcl}
+\documentclass[tc, manuscript]{copernicus}
 
@@
 \setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
@@
 \makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+
+\usepackage[inline]{trackchanges}
 \makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
 \makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-\KOMAoption{captions}{tableheading}
+\makeatletter
+\makeatother
 \makeatletter
@@
 \makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\IfFileExists{pdfcomment.sty}
-{
-  \usepackage{pdfcomment}
-  \newcommand{\criticmarkupadd}[1]{\pdfmarkupcomment[markup=Highlight,color={0.838431  0.913725  0.734902}]{##1}{}}
-  \newcommand{\criticmarkuprm}[1]{\pdfmarkupcomment[markup=Highlight,color={0.887059  0.673725  0.673725}]{##1}{}}
-  \newcommand{\criticmarkupmark}[1]{\pdfmarkupcomment[markup=Highlight,color={1  0.99216  0.21961}]{##1}{}}
-  \newcommand{\criticmarkupcomm}[1]{\pdfcomment[icon=Comment, color={0.89804  0.69020  0}]{##1}}
-}{
-
-  \newcommand{\criticmarkupadd}[1]{\{++{##1}++\}}
-  \newcommand{\criticmarkuprm}[1]{\{-{}-{##1}-{}-\}}
-  \newcommand{\criticmarkupmark}[1]{\{=={##1}==\}}
-  \newcommand{\criticmarkupcomm}[1]{\{>{}>{##1}<{}<\}}
-}
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
 \ifLuaTeX
@@
 \fi
-\usepackage{bookmark}
-
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={Untitled},
+  pdftitle={Evaluation of +S2 sea ice forecast and some other better title},
   colorlinks=true,
@@
   citecolor={Blue},
-  urlcolor={Blue},
+  urlcolor={red},
   pdfcreator={LaTeX via pandoc}}
 
-\title{Untitled}
-\author{}
-\date{}
+\title{Evaluation of +S2 sea ice forecast and some other better title}
+
+\Author[1,2]{Elio}{Campitelli}
+\Author[1,2]{Ariaan}{Purich}
+\Author[1,2]{Julie}{Arblaster}
+\Author[3]{Eun-Pa}{Lim}
+\Author[3]{Matthew}{Wheeler}
+\Author[3]{Phillip}{Reid}
+
+\affil[1]{School of Earth, Atmosphere and Environment, Monash
+University, Australia}
+\affil[2]{Securing Antarctica's Environmental Future, Monash University,
+Australia}
+\affil[3]{Bureau of Meteorology}
+
+\runningtitle{+S2 sea ice forecast}
+
+\runningauthor{Campitelli et al.}
+
+\correspondence{Elio\ Campitelli\ (elio.campitelli@monash.edu)}
 
 \begin{document}
+
+\received{}
+\pubdiscuss{} %% only important for two-stage journals
+\revised{}
+\accepted{}
+\published{}
+
+%% These dates will be inserted by Copernicus Publications during the typesetting process.
+
+\firstpage{1}
 \maketitle
 \begin{abstract}
-balbalbal
+The abstract goes here. It can also be on \emph{multiple lines}.
 \end{abstract}
+\copyrightstatement{The author's copyright for this publication is
+transferred to institution/company.}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, frame hidden, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-\criticmarkupmark{\textbf{Antarctic sea ice general description}.
-Seasonality, trends.}
+Unlike Arctic sea ice, which has been steadily retreating at least since
+the start of satellite records in the early 80s, Antarctic sea ice has
+experienced a slightly increasing trend that models systematically
+failed to reproduce and puzzled researchers \citep{hobbs2016}. Then, in
+2016 Antarctic sea ice extent dropped precipitously and has been at low
+and record low levels since, highlighting even more our lack of
+understanding of Antarctic sea ice variability and change
+\citep{espinosa2024}.
 
-\criticmarkupmark{\textbf{Importance}. Impacts on life. Potential
-impacts on weather. Impacts on climate. Impacts on ocean
-circulation.(Justify why it's important to study and understand)}
+Antarctic sea ice is in a remote location but the potential impacts of
+its variability extends far into the lower latitudes, affecting the
+circulation of both the ocean and the atmosphere, ocean carbon uptake
+and biological processes. It's likely that a reduction in sea ice extent
+would reduce the meridional temperature gradient which in turn would
+reduce the strength and location of the eddy-driven jet (and vice versa
+for an increase in sea ice), which is a crucial component of weather
+events. In the northern hemisphere, it has been hypothesised that the
+reduction in Arctic sea ice has lead to a weaker, wavier jet, increasing
+the frequency of extreme events \citep{barnes2015}. There is some
+evidence of this effect in the Southern Hemisphere. Models show a
+weakened vortex in the middle atmosphere, increased vertical wave flux
+and greater zonal wave 1 amplitude in Springs following a low sea ice
+maximum. But models don't fully agree on the exact phase of the zonal
+wave response and observations don't show a clear signal
+\citep{rea2024}. The jet response to sea ice loss is also very model
+dependent, particularly sensitive to its the climatological location and
+relatively small compared with the direct effect of CO2 increase and Sea
+Surface Temperature (SST) warming \citep{ayres2019}. On climate scales,
+it is thought that the temperature signal of a reduction in Antarctic
+sea ice mirrors a ``mini global warming'' and experiments also predict a
+weakening and northerly shift of the eddy-driven
+\citep{raphael2011, england2018, england2020, ayres2022}.
 
-\criticmarkupmark{\textbf{Prediction systems.} Why is it important:
-operations \criticmarkupcomm{Discuss with Phil}; science needs (we need
-good models). What is available. Discussion on S1 \criticmarkupcomm{Can
-we discuss previous S1 results? Laura's paper is not public.}}
+Correctly modelling Antarctic sea is not only necessary for process
+understanding and climate projections to inform adaptation strategies.
+Accurate seasonal to subseasonal forecast are crucial for operations in
+and around the Antarctic continent such as scientific missions,
+fisheries and tourism \citep[ \citet{wagner2020}]{desilva2020}. However,
+producing accurate Antarctic sea ice forecasts has been challenging due
+to model biases and inherent large variability and complexity and it has
+lagged behind Arctic sea ice forecasts \citep{zampieri2019}. Dynamical
+seasonal forecasts of summer Antarctic sea ice have been shown to
+perform worse than relatively simple statistical methods
+\citep{massonnet2023}, which also underscores the need for better
+understanding of sea ice dynamics.
 
-\criticmarkupmark{\textbf{Objective.} Evaluate S2.}
+In this work we evaluate sea ice forecasts produced by ACCESS-S2, the
+Australian Bureau of Meteorology subseasonal-to-seasonal prediction
+system. This evaluation will inform future work with the model as a
+reseach tool and explore the potential of using it as a sea ice
+forecasting tool.
 
-\section{Data and methods}\label{data-and-methods}
+\hypertarget{data-and-methods}{%
+\section{Data and methods}\label{data-and-methods}}
 
-All datasets are regrided using bilinear interpolation to a common grid,
-which is a polar sterographic grid with approximatedly equal area of
-25km².
+\hypertarget{s2}{%
+\subsection{ACCESS-S2}\label{s2}}
 
-\subsection{Description of the model}\label{description-of-the-model}
+ACCESS-S2 \citep{wedd2022} is the Australian Bureau of Meteorology
+current subseasonal-to-seasonal prediction system which replaced
+ACCESS-S1 \citep{hudson2017} in October 2021. Both ACCESS-S2 and
+ACCESS-S1 consist on the Global Atmosphere 6.0 (GA6)
+\citep{williams2015, waters2017}, Global Land 6.0
+\citep{best2011, waters2017}, Global Ocean 5.0
+\citep{madec2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
+\citet{rae2015}{]} model components. The atmosphere has a N216
+horizontal resolution (\textasciitilde60km in the mid-latitudes) with 85
+levels. The land model uses the same horizontal grid with 4 soil levels.
+The ocean component has a 1/4º resolution with 75 vertical levels. The
+ice component --based on CICE version 4.1-- has the same resolution than
+the ocean and 5 sea ice categories as well as an open water category.
 
-ACCESS-S2 is the next saeasonal prediction system after ACCESS-S1. It
-uses virtually the same exact model configuration except for some model
-versions being different. The main difference are the initial
-conditions. While S1 used initial conditions from the UK Met Office, S2
-uses the Bureau's own system. \criticmarkupmark{ Description of the
-system }\criticmarkupcomm{Atmosphere and ocean. Highlight no ice data
-assimilation. Highligh that the hope is that ice will respond to
-SST/atmosphere DA}
+Both systems take atmospheric initial conditions derived from
+ERA-interim \citep{dee2011} for the hindcast and from the Bureau's
+operational analysis for real-time forecasts. The main difference
+between the two are the ocean and ice initial conditions. ACCESS-S1
+initial conditions come from the Met Office FOAM system, which uses a
+multivariate, incremental three-dimensional variational (3D-Var),
+first-guess-at-appropriate-time (FGAT) data assimilation scheme
+\citep{waters2015} and assimilates sea surface temperature (SST), sea
+surface height (SSH), in situ temperature and salinity profiles, and sea
+ice concentration. ACCESS-S2, instead, runs from initial conditions
+generated by the in-house data assimilation scheme described in
+\citet{wedd2022}. This scheme is a weakly coupled ensemble optimal
+interpolation method and assimilates temperature and salinity profiles
+from EN4 \citep{good2013} for the hindcast and from the WMO Global
+Telecommunication System (GTS) and both the Coriolis and USGODAE Global
+Data Assembly Centers (GDACs) for the real-time forecast. SSTs are
+nudged to Reynolds OISSTv2.1 \citep{reynolds2007} for the hindcast and
+to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and
+Beggs 2008) for the real-time forecast in areas where SSTs are over 0ºC.
+Relevant for this work, sea ice concentrations are not assimilated.
 
-For evaluation we use hindcst for the period
-1981--2023\criticmarkupcomm{Check}. Anomalies will be taken with respect
-to the 1981--2011 climatology computed from the reanalysis. Climatology
-is smoothed with a 11 day running mean.
+The ACCESS-S2 hindcast used in this study runs for the period
+1981--2018. Each forecast consists of 9 ensemble members, 3 of which run
+for 279 days and 6 for 42 days. Anomalies are be taken with respect to
+the 1981--2011 climatology computed from the daily ACCESS-S2 reanalysis
+and smoothed with a 11 day running mean.
 
-\subsection{Verification datasets}\label{verification-datasets}
+The ACCESS-S1 hindcast runs for the period 1990--2012, with 7-member
+forecasts integrated for 217 days. No reanalysis is available so we are
+not able to compute a climatology for this forecast, therefore we don't
+show metrics based on anomalies.
 
-There is not a lot of data on sea ice properties, especially for things
-like thickness, age, etc. However there are relatively reliable
-satellite-derived estimates of sea ice concentration, which estimates
-the proportion of each grid area that is covered with ice. These
-products are not perfect and there still exists a fair bit of
-observation al uncertainty. To account for this uncertainty we use
-multiple datasets.
+\hypertarget{verification-datasets}{%
+\subsection{Verification datasets}\label{verification-datasets}}
 
-\subsubsection{Bootstrap}\label{bootstrap}
+For verification we use satellite-derived estimates of sea ice
+concentration, which estimates the proportion of each grid area that is
+covered with ice. Different datasets derived using different algorithms
+provide different estimates, each with its own biases and uncertainties.
+\citet{meier2019} estimated that the inter-product uncertainty of sea
+ice extent is of the order of 1 million km\textsuperscript{2}. As will
+be shown below, this spread is minimal compared with the errors in the
+ACCESS-S2 and ACCESS-S1 forecasts, so the conclusions are independent of
+the dataset used.
 
-\criticmarkupcomm{From https://nsidc.org/data/nsidc-0079/versions/4}
-This sea ice concentration data set was derived using measurements from
-the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7
+We use NOAA/NSIDC's Climate Data Record V4 {[}CDR; \citet{meier2014}{]}
+as the primary verification dataset. It takes the maximum value of the
+NASA Team \citep{cavalieri1984} and NASA Bootstrap \citep{comiso2023}
+sea ice concentration products to reduce their low concentration bias
+\citep{meier2014, meier2021}. Both source algorithms use data from the
+Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7
 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors
 on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and
--F13 satellites. Measurements from the Special Sensor Microwave
-Imager/Sounder (SSMIS) aboard DMSP-F17 are also included. The data set
-has been generated using the Advanced Microwave Scanning Radiometer -
-Earth Observing System (AMSR-E) Bootstrap Algorithm with daily varying
-tie-points. Daily (every other day prior to July 1987) and monthly data
-are available for both the north and south polar regions. Data are
-gridded on the SSM/I polar stereographic grid (25 x 25 km) and provided
-in two-byte integer format. Data coverage began on 01 November 1978 and
-is ongoing through the most current processing, with updated data
-processed several times annually.
+-F13 satellites. The data has a resolution of 25 by 25km and daily from
+1978 onwards.
 
-(Comiso 2023)
+The European Organisation for the Exploitation of Meteorological
+Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility
+(OSI SAF) \citep{OSISAF} is another satellite-derived sea ice
+concentration product. It is based on mostly the same sensors as the NOA
+CDR but computed independently using different algorithms. Figures
+prepared with this dataset are provided in the supplementary material.
 
-\subsubsection{NASA Team}\label{nasa-team}
-
-\criticmarkupmark{missing}
-
-\subsubsection{CDR}\label{cdr}
-
-NOAA/NSIDC's Climate Data Record V4 combines the Bootstrap and NASA Team
-estimates in an attempt to overcome each algorithm's weaknesses.
-\criticmarkupcomm{More detail on how are they combined and their
-properties.} As such, it is is not a truly independent dataset.
-
-Meier, W. N., F. Fetterer, A. K. Windnagel, and S. Stewart. 2021.
-NOAA/NSIDC Climate Data Record of Passive Microwave Sea Ice
-Concentration, Version 4. {[}Indicate subset used{]}. Boulder, Colorado
-USA. NSIDC: National Snow and Ice Data Center
-https://doi.org/10.7265/efmz-2t65. {[}Date Accessed{]}
-\criticmarkupcomm{Add to zotero}
-
-\subsubsection{ERA5}\label{era5}
-
-\criticmarkupmark{From the CDS websiste:} ERA5 is the fifth generation
-ECMWF reanalysis for the global climate and weather for the past 8
-decades. Data is available from 1940 onwards. ERA5 replaces the
-ERA-Interim reanalysis.
-
-This parameter is the fraction of a grid box which is covered by sea
-ice. Sea ice can only occur in a grid box which includes ocean or inland
-water according to the land-sea mask and lake cover, at the resolution
-being used. This parameter can be known as sea-ice (area) fraction,
-sea-ice concentration and more generally as sea-ice cover. In ERA5,
-sea-ice cover is given by two external providers. Before 1979 the
-HadISST2 dataset is used. From 1979 to August 2007 the OSI SAF (409a)
-dataset is used and from September 2007 the OSI SAF oper dataset is
-used. Sea ice is frozen sea water which floats on the surface of the
-ocean. Sea ice does not include ice which forms on land such as
-glaciers, icebergs and ice-sheets. It also excludes ice shelves which
-are anchored on land, but protrude out over the surface of the ocean.
-These phenomena are not modelled by the IFS. Long-term monitoring of sea
-ice is important for understanding climate change. Sea ice also affects
-shipping routes through the polar regions.
-
-\criticmarkupcomm{Why is ERA5 better or worse than
-satellites?\textless{} \textless\}
-
-\subsection{Error measures}\label{error-measures}
+\hypertarget{error-measures}{%
+\subsection{Error measures}\label{error-measures}}
 
@@
 
-\subsubsection{Sea ice extent}\label{sea-ice-extent}
-
 Sea Ice Extent is defined as the area of ocean covered with at least
 15\% ice. This threshold is motivated by the limitations in satellite
-retrieval, which is increasingly unreliable for low sea ice conditions.
+retrieval, which is increasingly unreliable for lower sea ice
+conditions.
 
-Sea Ice Extent is a rough global measure, but a model could have
-relatively accurate extent of ice but with different distributions. We
-use two other measures to account for these errors.
+Pan-Antarctic Sea Ice Extent serves as a rough global measure of the
+amount of sea ice but it doesn't take into account the spatial
+distribution. A model could have relatively accurate extent of ice but
+with different distributions. To account for location errors, we
+computed the Root Mean Squared Error (RMSE) of sea ice concentration
+anomalies and the Integrated Ice Edge Error {[}IIEE;
+\citet{goessling2016}{]}.
 
-We compute Root Mean Square Error of sea ice concentration anomalies.
+We compute RMSE as the square root of the average squared differences
+between forecasted and observed sea ice concentration anomalies. We
+compute a pan-Antarctic RMSE by averaging over the whole domain, and
+also a zonally-varying RMSE by averaging over 24 15° longitude slices
+around Antarctica.
 
-We also compute the Integrated Ice Edge Error (IIEE) (Goessling et al.
-2016). This is defined as the area in which the model misspredicts sea
-ice concentration being above or belog 15\% ice. That is, dichotomise
-sea ice concentration into areas with more and less than 15\% sea ice
-both in the forecast and observations; the IIEE is the area in which
-forecast and observations differ.
+The IIEE is defined as the area in which the model misspredicts sea ice
+concentration being above or belong 15\% ice. That is, dichotomise sea
+ice concentration into areas with more and less than 15\% sea ice both
+in the forecast and observations; the IIEE is the area in which forecast
+and observations differ.
 
-\section{Results and discussion}\label{results-and-discussion}
+All error measures were computed on the NOAA/NSIDC CDRV4 domain grid and
+projection to which model output was bilinearly interpolated.
 
-\subsection{Reanalysis}\label{reanalysis}
+\hypertarget{results-and-discussion}{%
+\section{Results and discussion}\label{results-and-discussion}}
 
-\subsubsection{Bias}\label{bias}
+\hypertarget{reanalysis}{%
+\subsection{Reanalysis}\label{reanalysis}}
+
+\hypertarget{bias}{%
+\subsubsection{Bias}\label{bias}}
 
@@
 
-\centering{
-
-\includegraphics{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}
+{\centering \includegraphics{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}
 
@@
 hindcasts initialised the first of the month for ACCESS-S2 and ACCESS-S1
-in colours representing the start month. In black, the}
+in colours representing the start month. In black, the median sea ice
+extent of NOAA/NSIDC CDRV4.}
 
-\end{figure}%
+\end{figure}
 
 Figure~\ref{fig-hindcast-extent} a shows the seasonal cycle of Sea Ice
-Extent for the ACCESS-S2 hindcast and 2 observational datasets.
-ACCESS-S2 shows a severe low extent bias, especially in the late
-summer-early autumn. This is due primarily to a faster and longer melt
-season between January and March and slower growth during March and
-April This is then balanced with faster growth between May and July
-(Fig.~\ref{fig-mean-growth}). This bias is common in climate models
-{[}ref!{]} and it's been linked to xxxx??
+Extent for the ACCESS-S2 hindcast and NOAA/NSIDC CDRV4 ACCESS-S2 shows a
+severe low extent bias, especially in the late summer-early autumn. This
+is due primarily to faster melt during a longer melt season between
+January and March and slower growth during March and April. This is then
+balanced with faster growth between May and July
+(Fig.~\ref{fig-mean-growth}). Many sea ice models exhibit this
+systematic underestimation during the sea ice minimum and early freezing
+season \citep{massonnet2023} and could indicate problems in the
+representation of thermodynamics in the model \citep{zampieri2019}.
 
-Comparing ACCESS-S2 with ACCESS-S1, the latter has a smaller bias,
-especially at low lags (Fig.~\ref{fig-hindcast-extent}) even though the
-typical growth rates are very similar between both models
-(Fig.~\ref{fig-mean-growth}). At larger lags, ACCESS-S1's bias in summer
-and autumn is very similar to ACCESS-S2`s. This suggests that this lower
-sea ice state is closer to the models' equilibrium, indicating that it
-is an issue with model formulation that was being corrected by the data
-assimilation system in ACCESS-S1.
+ACCESS-S1 has a smaller low bias than ACCESS-S2, especially at low lags
+(Fig.~\ref{fig-hindcast-extent}) even though the typical growth rates
+are very similar between both models (Fig.~\ref{fig-mean-growth}). At
+larger lags, ACCESS-S1's bias in summer and autumn is very similar to
+ACCESS-S2's. This suggests that this lower sea ice state is closer to
+the models' equilibrium, indicating that it is an issue with model
+formulation that was being corrected by the data assimilation system in
+ACCESS-S1 and not in ACCESS-S2.
 
-At long lags, sea ice extent loses most of the initial condition memory
+At large lags, sea ice extent loses most of the initial condition memory
 and reverts to the model's preferred equilibrium state. Therefore we can
@@
 the triangles in each panel in Fig.~\ref{fig-hindcast-extent}), owing to
-both having the same model formulation. From June to October, in
-ACCESS-S2 circles move away from triangles and towards observations,
+them having the same model formulation. From June to October, in
+ACCESS-S2 circles are closer to observations than to the triangles,
 indicating that the information from the ocean and atmosphere data
-assimilation is getting to sea ice and affecting the initial conditions.
+assimilation is getting to sea ice and improving the initial conditions.
 The rest of the year, there is little if any difference between circles
@@
 is taking place and the sea ice component of the model is virtually
-free-running.
+free-running. In ACCESS-S1, on the other hand, circles are very close to
+observations year-round.
 
@@
 
-\centering{
-
-\includegraphics{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}
+{\centering \includegraphics{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
 
@@
 
-\centering{
-
-\includegraphics{access-ice_files/figure-pdf/fig-bias-1.pdf}
+{\centering \includegraphics{access-ice_files/figure-pdf/fig-bias-1.pdf}
 
@@
 
-\caption{\label{fig-bias}ACCESS-S2 reanalysis sea ice concentration bias
-compared with NSIDC sea ice concentration.}
+\caption{\label{fig-bias}Mean sea ice concentrations 1-month lag monthly
+mean ACCESS-S2 forecast bias compared with NSIDC.}
 
-\end{figure}%
+\end{figure}
 
-Figure~\ref{fig-bias} shows the difference in monthly mean sea ice
-concentrations between CDR and ACCESS-S2 reanalysis. From October to
-May, the model underestimates sea ice concentrations pretty much
-everywhere there is ice except for the deep Weddell Sea in April and
-May, where sea ice concentrations saturate to 1. In winter, the
-differences are mostly on the sea ice edge, with slight positive bias in
-XXX and negative bias around the Indian Ocean sector.
+Figure~\ref{fig-bias} shows the difference of monthly mean sea ice
+concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
+shortest lag. From October to May, the model underestimates sea ice
+concentrations pretty much everywhere there is ice except for the deep
+Weddell Sea in April and May, where sea ice concentrations saturate to 1
+both in the observations and forecasts. In winter, the differences are
+mostly on the sea ice edge, with light positive bias in the African
+sector of East Antarctica and negative bias around the Indian Ocean
+sector which partially compensate, resulting in the near-zero extent
+bias seen in those months (Fig.~\ref{fig-hindcast-extent}).
 
-\subsubsection{Anomalies}\label{anomalies}
-
-\criticmarkupmark{intro connecting subsections}
+\hypertarget{anomalies}{%
+\subsubsection{Anomalies}\label{anomalies}}
 
@@
 
-\centering{
-
-\includegraphics{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}
+{\centering \includegraphics{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}
 
@@
 \caption{\label{fig-extent-anom}Sea ice extent anomalies for ACCESS-S1
-and ACCESS-S2 (black) and CDR (blue).}
+and ACCESS-S2 (black) and NOAA/NSIDC CDRV4 (blue).}
 
-\end{figure}%
+\end{figure}
 
@@
 forecasted at selected lags. The anomalies in this case were computed
-with respect of the climatology of each lag, which is a way of
-bias-correction. Compared with ACCESS-S1, ACCESS-S2 anomaly forecast is
-relatively poor even in the first month, which stays relatively skillful
-even at lag 3. ACCESS-S2 shows much bigger variability than
-observations, with dramatic lows between 1995 and 2007 and highs between
-2007 and 2015.
+with respect of the climatology of each lag, which works as a
+bias-correction method. Compared with ACCESS-S1, ACCESS-S2 anomaly
+forecast is relatively poor even in the first month, which stays
+relatively skilful even at lag 3. ACCESS-S2 shows much bigger
+variability than observations, with dramatic lows between 1995 and 2007
+and highs between 2007 and 2015.
 
-\criticmarkupcomm{Add reginoal extents?\textless{} \textless\}
-
-\subsubsection{RMSE}\label{rmse}
+\hypertarget{rmse}{%
+\subsubsection{RMSE}\label{rmse}}
 
@@
 
-\centering{
-
-\includegraphics{access-ice_files/figure-pdf/fig-rmse-1.pdf}
+{\centering \includegraphics{access-ice_files/figure-pdf/fig-rmse-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
+
+Use climatology
+
+monthly rmse/corelation at gridpoints per lead time.
 
@@
 concentration anomalies for ACCESS-S2 forecasts compared with a
-benchmark of persistence. This figure uses CDR data, but the results are
-nearly identical compared with ERA5 or Bootstrap. Due to errors in the
-initial conditions, it is expected that a persistence forecast would be
-better than the model forecast at very short lags, but that the
-persistence forecast errors would grow faster and eventually surpass the
-mode forecast, at which time is statistically useful
-\criticmarkupmark{I've got this from CC at the ICTP summer school and
-makes sense, but it would be great to have a referece?= =\}. Here the
-persistence errors are almost always lower than the ACCESS-S2 forecast,
-indicating that the model doesn't have skill at any lag and in any
-month. The only exception is the RMSE around Febraury forecasted from
-June onwards.
-
-\criticmarkupmark{Add regional?= =\}
-
-\subsubsection{Comparison with S1}\label{comparison-with-s1}
+benchmark of persistence. Due to errors in the initial conditions, it is
+expected that a persistence forecast would be better than the model
+forecast at very short lags, but that the persistence forecast errors
+would grow faster and eventually surpass the mode forecast, after when
+it is statistically useful. Here the persistence errors are almost
+always lower than the ACCESS-S2 forecast, indicating that the model
+doesn't have skill at any lag and in any month. The only exception is
+around February, where the model has lower RMSE than the persistence
+forecast at virtually every lag.
 
@@
 
-\centering{
+{\centering \includegraphics{access-ice_files/figure-pdf/fig-rmse_lon-1.pdf}
 
-\includegraphics{access-ice_files/figure-pdf/fig-iiee-1.pdf}
+}
+
+\caption{\label{fig-rmse_lon}Meadian difference between RMSE of
+ACCESS-S2 forecasts and persistence forecast at longitudinal bands.}
+
+\end{figure}
+
+Even though ACCESS-S2 is no better than persistence at forecasting
+pan-Antarctic sea ice concentration anomalies, the quality of the
+forecast might vary between regions. To analyse the spatial distribution
+of the model error, we computed sea ice concentration anomalies RMSE on
+15 meridional slices 24º wide. The median difference between the
+forecast and persistence RMSE is shown on Figure~\ref{fig-rmse_lon},
+where negative values indicate that the model has lower RMSE than the
+benchmark.
+
+The skill shown by ACCESS-S2 at February-March forecasts is evidenced by
+a band of negative values across almost all longitudes. The only region
+where February-March sea ice is not well forecasted is east of the
+Antarctic Peninsula, which is the only region with consistent summer sea
+ice. This suggests that the positive skill around the summer minimum
+comes from the model correctly forecasting no ice where no ice is ever
+found. However, there are other regions of skilful forecasts.
+
+Forecasts initialised on January 1\textsuperscript{st}
+(Fig.~\ref{fig-rmse_lon} panel a) show RMSE values below persistence at
+lags larger than 180 days (corresponding to July through September) in
+the Ross and Weddell Sea. These two regions are also relatively well
+forecasted from other months. Notably, May forecasts of Weddell sea ice
+concentration anomalies (Fig.~\ref{fig-rmse_lon} panel e) are skilful
+after about 20 days and remain so until December. The Weddell Sea is
+also the region with the maximum error --around June regardless of
+initialisation date.
+
+\hypertarget{comparison-with-s1}{%
+\subsubsection{Comparison with S1}\label{comparison-with-s1}}
+
+\begin{figure}
+
+{\centering \includegraphics{access-ice_files/figure-pdf/fig-iiee-1.pdf}
 
@@
 Error as a function of forecast lag for all forecast initialised on the
-first of each month for ACCESS-S1 and ACCESS-S2 hindcasts.}
+first of each month for ACCESS-S1 and ACCESS-S2 hindcasts. For each
+month, the number in parenthesis indicates the minimum lag at which the
+the mean error of each model is not statistically significant at a 1\%
+level using a two-sisded t-test.}
 
-\end{figure}%
+\end{figure}
 
@@
 has lower error at short lags at all months, with the errors converging
-as the forecast goes on. The time to convergence depends on the month
-and it can be as fas as a few days in July to as large as several months
-for forecasts initiated in February and March. Since the only difference
-between these forecasts are the initial conditions, this timescale is an
-indication of the the memory of sea ice to initial conditions; at least
-from October to March when the data assimilated form the other
-components has little to no influence on sea ice.
+to ACCESS-S2 as the forecast goes on. The time to convergence depends on
+the month and it can be as short as two weeks June and July to as long
+as 160 days for forecasts initialised in February. Since the only
+difference between these forecasts are the initial conditions, this
+timescale is an indication of the the memory of sea ice to initial
+conditions --at least from October to March when the data assimilated
+form the other components has little to no influence on sea ice.
 
 Also evident in Figure~\ref{fig-iiee} is the difference in the error
-spread at short lags between ACCESS-S2 and ACCESS-S1. In all month
-ACCESS-S1 shows a small error spread at lag 1, indicating that the error
-in the initial conditions not only is small, but also fairly constant.
-This spread then grows towards a climatological spread as errors
-accumulate differently in different forecasts. For ACCESS-S2, this is
-true only only between July and October, approximately. For all other
-months, the error spread is more or less stable throughout the forecast
-window, indicating that not only the initial error is high, but it's not
+spread at short lags between ACCESS-S2 and ACCESS-S1. In all months
+ACCESS-S1 has much narrower error spread at lag 1. The error in the
+initial conditions is not only small, but also fairly constant. This
+spread then grows towards a climatological spread as errors accumulate
+differently in different forecasts. For ACCESS-S2, this is true only
+only between July and October, approximately. For all other months, the
+error spread is more or less stable throughout the forecast window,
+indicating that not only the initial error is high, but it's not
 constant.
@@
 
-\centering{
-
-\includegraphics{access-ice_files/figure-pdf/fig-iiee-variance-1.pdf}
+{\centering \includegraphics{access-ice_files/figure-pdf/fig-iiee-variance-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
 
 The large initial error spread could be due either to large spread of
-ensemble members or due to a large spread of individual forecasts.
+ensemble members or to large spread of individual forecasts.
 Figure~\ref{fig-iiee-variance} splits the IIEE variance for each lag
@@
 
-\subsection{Conclusions}\label{conclusions}
+\hypertarget{conclusions}{%
+\subsection{Conclusions}\label{conclusions}}
 
-\section{References}\label{references}
+Sea ice forecasts from the ACCESS-S2 model show a significant low extent
+bias, particularly during late summer and early autumn. This bias is
+attributed to a faster and longer melt season between January and March,
+and slower growth between March and April. This underestimation during
+the minimum and early freezing season is a common issue in many
+seasonal-to-subseasonal (S2S) systems, suggesting potential problems
+either with the model's thermodynamic representation or with short wave
+radiation forcing, as shown in other climate models. Even though
+ACCESS-S2 shares the same model formulation with ACCESS-S1, the latter
+does not suffer from this bias. This is due to the assimilation of sea
+ice concentration anomalies into the initial conditions, which
+successful corrects for the negative bias in the model. The analysis
+suggests that the data assimilation system in ACCESS-S2 is only
+effectively influencing sea ice initial conditions from June to October,
+while the rest of the year, the sea ice component runs virtually free,
+reverting to its biased equilibrium state.
 
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-comiso2023}
-Comiso, J. 2023. {``Bootstrap {Sea Ice Concentrations} from {Nimbus-7
-SMMR} and {DMSP SSM}/{I-SSMIS}. ({NSIDC-0079}, {Version} 4).''} Boulder,
-Colorado USA.: {NASA National Snow and Ice Data Center Distributed
-Active Archive Center.} \url{https://doi.org/10.5067/X5LG68MH013O}.
+Analysis of the error spread shows that ACCESS-S2 initial conditions
+from December to May not only have large errors, but that the initial
+error spread is very large compared with ACCESS-S1. This spread is not
+due to the perturbation scheme, since the mean error variance for
+individual forecasts is low and comparable with ACCESS-S1. Instead, it
+is due to large variance of the mean error of individual forecasts,
+which is comparable to the climatology spread. This is further evidence
+that individual initial conditions are not being affected by the data
+assimilation scheme.
 
-\bibitem[\citeproctext]{ref-goessling2016}
-Goessling, H. F., S. Tietsche, J. J. Day, E. Hawkins, and T. Jung. 2016.
-{``Predictability of the {Arctic} Sea Ice Edge.''} \emph{Geophysical
-Research Letters} 43 (4): 1642--50.
-\url{https://doi.org/10.1002/2015GL067232}.
+Based on the observation that ACCESS-S2 sea ice initial conditions are
+essentially not initialised, comparing its forecasts with ACCESS-S1's
+allows us to estimate the time-scale for which initial conditions are
+important. This highlights February initial conditions as a crucial for
+determining sea ice evolution at least up to late June. Arctic sea ice
+forecast also show greater sensitivity to initial conditions in (boreal)
+summer compared with (boreal) winter \citep{day2014, bunzel2016}, so
+similar mechanism might be playing a role.
 
-\end{CSLReferences}
+Although forecasts are not skilful for forecasting panantarctic sea ice
+concentrations, there are some areas where the model does show skill.
+The Weddell Sea, and the Ross Sea to a lesser extent are particularly
+well forecasted between June and November for forecasts initialised from
+January to June. These regions have been recognised as a regions of high
+predictability thanks to persistent and eastwardly advected upper ocean
+heat content anomalies \citep{bushuk2021} Sea surface temperatures are
+assimilated by ACCESS-S2 only in areas with temperatures greater than
+0°C and errors don't show a clear easterly-propagating signal, so it is
+not clear if ACCESS-S2 is leveraging the same source of predictability.
 
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\renewcommand{\bibsection}{}
+\bibliography{references.bib}
+
+\hypertarget{suplementary-figures}{%
+\section{Suplementary figures}\label{suplementary-figures}}
+
+\codedataavailability{use this to add a statement when having data sets
+and software code
+available} %% use this section when having data sets and software code available
+
+\sampleavailability{use this section when having geoscientific samples
+available} %% use this section when having geoscientific samples available
+
+\videosupplement{use this section when having video supplements
+available} %% use this section when having geoscientific samples available
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
+\section{Figures and tables in appendices}
+
+Regarding figures and tables in appendices, the following two options
+are possible depending on your general handling of figures and tables in
+the manuscript environment:
+
+\subsection{Option 1}
+
+If you sorted all figures and tables into the sections of the text,
+please also sort the appendix figures and appendix tables into the
+respective appendix sections. They will be correctly named
+automatically.
+
+\subsection{Option 2}
+
+If you put all figures after the reference list, please insert appendix
+tables and figures after the normal tables and figures.
+
+To rename them correctly to A1, A2, etc., please add the following
+commands in front of them: \texttt{\textbackslash{}appendixfigures}
+needs to be added in front of appendix figures
+\texttt{\textbackslash{}appendixtables} needs to be added in front of
+appendix tables
+
+Please add \texttt{\textbackslash{}clearpage} between each table and/or
+figure.
+\noappendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontribution{Contributions} %% optional section
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\competinginterests{The authors declare no competing
+interests.} %% this section is mandatory even if you declare that no competing interests are present
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\disclaimer{We like Copernicus.} %% optional section
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{acknowledgements}
+Acknowledgements will gohere
+\end{acknowledgements}
+
+\bibliographystyle{copernicus}
 \end{document}

### 2025-02-24 scripts/error_metrics.R

@@ -1,4 +1,4 @@
-logfile <- here::here("logs/error_metrics.log")q
+logfile <- here::here("logs/error_metrics.log")
 log <- function(text) {
   messg <- paste0(format(lubridate::now()), ": ", text)
   if (interactive()) {
@@ -158,6 +158,11 @@ compute_metrics <- function(forecast_time, member, version, obs_dataset) {
   cdo_options_set(c("-L"))
   
   file <- hindcast(forecast_time, model = version, members = member)
+
+  if (length(file) == 0) {
+    stop(glue::glue("No file found for model {version} at forecast time {forecast_time} member {member}"))
+  }
+  
   model_climatology <- here::here("data/derived/climatology", version, pad_number(month(forecast_time)), "em.nc")
   
   file_template <- paste0("di_aice_", format(forecast_time, "%Y%m%d"), "_e0", member, ".nc")
@@ -258,7 +263,8 @@ forecast_times_s1 <- get_forecast_times("S1") |>
   Filter(f = \(x) mday(x) == 1) 
 
 forecast_times_s2 <- get_forecast_times("S2") |> 
-  Filter(f = \(x) mday(x) == 1) 
+  Filter(f = \(x) mday(x) == 1) |> 
+  Filter(f = Negate(\(x) year(x) == 1981 & month(x) == 1)) 
 
 all_files <- data.table::CJ(
   forecast_time = forecast_times_s1,

### 2025-02-24 scripts/pbs/error_metrics.sh

@@ -2,7 +2,7 @@
 
 #PBS -P lo70
 #PBS -q normal
-#PBS -l ncpus=48
+#PBS -l ncpus=32
 #PBS -l mem=64GB
 #PBS -l walltime=8:00:00
 #PBS -l wd
@@ -17,5 +17,5 @@ module load R/4.3.1
 module load cdo/2.4.3
 
 # Run R application
-export PBS_WORKERS=24
+export PBS_WORKERS=28
 Rscript scripts/error_metrics.R 

### 2025-02-25 analysis/poster_ncar/presentation_figures.Rmd
works on ncar figures

--- 2025-02-24 analysis/paper/access-ice.qmd
@@
 ---
-title: Evaluation of +S2 sea ice forecast and some other better title
-pdf-engine: pdflatex
-journal:
-  name: tc
-author:
-  - given_name: Elio 
-    surname: Campitelli
-    affiliation: "1,2"
-    email: elio.campitelli@monash.edu
-    corresponding: true
-  - given_name: Ariaan 
-    surname: Purich
-    affiliation: "1,2"
-  - given_name: Julie
-    surname: Arblaster
-    affiliation: "1,2"
-  - given_name: Eun-Pa
-    surname: Lim
-    affiliation: "3"
-  - given_name: Matthew
-    surname: Wheeler
-    affiliation: "3"
-  - given_name: Phillip
-    surname: Reid
-    affiliation: "3"
-
-affiliation:
-  - code: 1
-    address: School of Earth, Atmosphere and Environment, Monash University, Australia
-  - code: 2
-    address: Securing Antarctica’s Environmental Future, Monash University, Australia
-  - code: 3
-    address: Bureau of Meteorology
-
-abstract: |
-  The abstract goes here.
-  It can also be on _multiple lines_.
-bibliography: references.bib
-running:
-  title: +S2 sea ice forecast
-  author: Campitelli et al.
-# This section is mandatory even if you declare that no competing interests are present.
-competinginterests: |
-  The authors declare no competing interests.
-# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
-# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
-copyrightstatement: |
-  The author's copyright for this publication is transferred to institution/company. 
-### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
-### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
-### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
-availability:
-  #code: |
-  #  use this to add a statement when having only software code available
-  #data: |
-  #  use this to add a statement when having only data sets available
-  codedata: |
-    use this to add a statement when having data sets and software code available
-  sample: |
-    use this section when having geoscientific samples available
-videosupplement: |
-  use this section when having video supplements available
-authorcontribution: |
-  Contributions
-disclaimer: |
-  We like Copernicus.
-acknowledgements: |
-  Acknowledgements will gohere
-appendix: |
-  \section{Figures and tables in appendices}
-  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
-  \subsection{Option 1}
-  If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
-  They will be correctly named automatically.
-  \subsection{Option 2}
-  If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
-
-  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
-  `\appendixfigures` needs to be added in front of appendix figures
-  `\appendixtables` needs to be added in front of appendix tables
-
-  Please add `\clearpage` between each table and/or figure.
-format: copernicus-pdf
-
-execute: 
-  echo: false
-  warning: false
-  message: false
-  cache: refresh
-filters:
-  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
-  # - ../../_extensions/mloubout/critic-markup/critic-markup.lua
-  - ../../_extensions/ute/search-replace/search-replace.lua
-search-replace:
-  +S2: ACCESS-S2
-  +S1: ACCESS-S1
-  +CDR: NOAA/NSIDC CDRV4
+title: "Poster figures"
+output: 
+  html_document:
+    self_contained: false  
+date: "2024-11-11"
 ---
@@
 
-# Abstract
-
-balbalbal
-
-# Introduction
-
-Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016].
-Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change [@espinosa2024].
-
-Antarctic sea ice is in a remote location but the potential impacts of its variability extends far into the lower latitudes, affecting the circulation of both the ocean and the atmosphere, ocean carbon uptake and biological processes.
-It's likely that a reduction in sea ice extent would reduce the meridional temperature gradient which in turn would reduce the strength and location of the eddy-driven jet (and vice versa for an increase in sea ice), which is a crucial component of weather events.
-In the northern hemisphere, it has been hypothesised that the reduction in Arctic sea ice has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015].
-There is some evidence of this effect in the Southern Hemisphere.
-Models show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in Springs following a low sea ice maximum.
-But models don't fully agree on the exact phase of the zonal wave response and observations don't show a clear signal [@rea2024].
-The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location and relatively small compared with the direct effect of CO2 increase and Sea Surface Temperature (SST) warming [@ayres2019].
-On climate scales, it is thought that the temperature signal of a reduction in Antarctic sea ice mirrors a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].
-
-Correctly modelling Antarctic sea  is not only necessary for process understanding and climate projections to inform adaptation strategies.
-Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent such as scientific missions, fisheries and tourism [@desilva2020 @wagner2020].
-However, producing accurate Antarctic sea ice forecasts has been challenging due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
-Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.
-
-In this work we evaluate sea ice forecasts produced by +S2, the Australian Bureau of Meteorology subseasonal-to-seasonal prediction system. 
-This evaluation will inform future work with the model as a reseach tool and explore the potential of using it as a sea ice forecasting tool.
-
-# Data and methods
-
-## +S2
-
-+S2 [@wedd2022] is the Australian Bureau of Meteorology current subseasonal-to-seasonal prediction system which replaced +S1 [@hudson2017] in October 2021.
-Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015] model components.
-The atmosphere has a N216 horizontal resolution (\~60km in the mid-latitudes) with 85 levels.
-The land model uses the same horizontal grid with 4 soil levels.
-The ocean component has a 1/4º resolution with 75 vertical levels.
-The ice component --based on CICE version 4.1-- has the same resolution than the ocean and 5 sea ice categories as well as an open water category.
-
-Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts.
-The main difference between the two are the ocean and ice initial conditions.
-+S1 initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
-+S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022.
-This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
-SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are over 0ºC.
-Relevant for this work, sea ice concentrations are not assimilated.
-
-The +S2 hindcast used in this study runs for the period 1981--2018.
-Each forecast consists of 9 ensemble members, 3 of which run for 279 days and 6 for 42 days.
-Anomalies are be taken with respect to the 1981--2011 climatology computed from the daily +S2 reanalysis and smoothed with a 11 day running mean.
-
-The +S1 hindcast runs for the period 1990--2012, with 7-member forecasts integrated for 217 days. 
-No reanalysis is available so we are not able to compute a climatology for this forecast, therefore we don't show metrics based on anomalies.
-
-## Verification datasets
-
 ```{r datasets}
-datasets <- list(cdr = CDR(),
-                 osi = OSI())
+datasets <- list(cdr = CDR())
 which_dataset <- "cdr"
 ```
-
-For verification we use satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
-Different datasets derived using different algorithms provide different estimates, each with its own biases and uncertainties. 
-@meier2019 estimated that the inter-product uncertainty of sea ice extent is of the order of 1 million km^2^. 
-As will be shown below, this spread is minimal compared with the errors in the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
-
-We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset.
-It takes the maximum value of the NASA Team [@cavalieri1984] and NASA Bootstrap [@comiso2023] sea ice concentration products to reduce their low concentration bias [@meier2014; @meier2021].
-Both source algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
-The data has a resolution of 25 by 25km and daily from 1978 onwards.
-
-The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [@OSISAF] is another satellite-derived sea ice concentration product.
-It is based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
-Figures prepared with this dataset are provided in the supplementary material.
-
-## Error measures
-
-For evaluation purposes, we use a series of measures.
-
-Sea Ice Extent is defined as the area of ocean covered with at least 15% ice.
-This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea ice conditions.
-
-Pan-Antarctic Sea Ice Extent serves as a rough global measure of the amount of sea ice but it doesn't take into account the spatial distribution. 
-A model could have relatively accurate extent of ice but with different distributions.
-To account for location errors, we computed the Root Mean Squared Error (RMSE) of sea ice concentration anomalies and the Integrated Ice Edge Error [IIEE; @goessling2016].
-
-We compute RMSE as the square root of the average squared differences between forecasted and observed sea ice concentration anomalies. 
-We compute a pan-Antarctic RMSE by averaging over the whole domain, and also a zonally-varying RMSE by averaging over 24 15° longitude slices around Antarctica.
-
-The IIEE is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
-That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.
-
-All error measures were computed on the +CDR domain grid and projection to which model output was bilinearly interpolated.
-
-# Results and discussion
-
-## Reanalysis
 
@@
 extent_climatology <- extent_daily |>  
-  copy() |>
-  _[aice == 0, aice := NA] |> 
-  _[!(month(time) == 2 & mday(time) == 29)] |> 
-  _[, time := update(time, year = 2001)] |> 
-  _[, median_ci(aice), by = .(time, dataset)]
-
-```
-
-```{r area_daily}
-area_daily <- datasets |>
-  lapply(area) |>
-  lapply(\(x) ReadNetCDF(x, "aice")) |>
-  rbindlist(idcol = "dataset") |>
-  _[, let(lon = NULL, lat = NULL)]
-```
-
-```{r area_climatology}
-area_climatology <- area_daily |>  
   copy() |>
@@
   scale_y_continuous(NULL, labels = labels_extent) +
-  scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"), 
+  scale_color_manual(values = colorRampPalette(gglgbtq::palette_lgbtq("nonbinary")[c(1, 3, 4, 1)])(12),
                      aesthetics = c("fill", "color"), guide = "none") +
@@
 
-@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR +S2 shows a severe low extent bias, especially in the late summer-early autumn.
-This is due primarily to faster melt during a longer melt season between January and March and slower growth during March and April. 
-This is then balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
-Many sea ice models exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
+```{r}
+hindcast_extent |> 
+  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
+                              lag = as.numeric(as.Date(time) - forecast_time),
+                              model)] |>
+  _[, time := forecast_time + lag] |>
+  _[, time2 := update(time, year = 2001)] |>
+  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
+  _[copy(extent_climatology)[, time := as.Date(time)], on = c(time2 = "time")] |> 
+  _[] |> 
+  ggplot(aes(time2, mid - i.mid)) +
+  geom_hline(yintercept = 0) +
+  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +
 
-+S1 has a smaller low bias than +S2, especially at low lags ([Fig. @fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig. @fig-mean-growth]).
-At larger lags, +S1's bias in summer and autumn is very similar to +S2's.
-This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1 and not in +S2.
+  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
+             aes(fill = month), 
+             shape = 24, size = 2.2) +
 
-At large lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
-Therefore we can estimate the latter using the hindcasts with the largest possible lag, which is shown in triangles in @fig-hindcast-extent for the same dates as the initial conditions.
-The difference between the two is the effect of the data assimilation.
+  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
+             shape = 21, size = 2) +
 
-The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to them having the same model formulation.
-From June to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and improving the initial conditions.
-The rest of the year, there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.
-In +S1, on the other hand, circles are very close to observations year-round.
-
-```{r fig-mean-growth}
-#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations. Values are smoothed with a 2-degree loess smooth with a 30 day window.
-
-dif_hindcast <- hindcast_extent |> 
-  copy() |> 
-  _[order(time)] |> 
-  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(model, forecast_time, member)] |> 
-  _[, median_ci(dif), by = .(model, time = update(time, year = 2000))]
-
-extent_daily |> 
-  _[order(time)] |> 
-  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
-  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
-  _[dataset == which_dataset] |> 
-  ggplot(aes(time, mid)) +
-  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
-  geom_smooth(aes(group = dataset, color = dataset),
-              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
-  geom_smooth(data = dif_hindcast,
-              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
-  scale_x_datetime(NULL, date_breaks = "1 month", 
+  scale_x_date(NULL, date_breaks = "1 month", 
                date_labels = "%b", expand = c(0, 0)) +
-  scale_y_continuous(NULL, labels = \(x) paste0(labels_extent(x), "/day")) +
-  scale_color_models
+  scale_y_continuous(NULL, labels = labels_extent) +
+  scale_color_manual(values = colorRampPalette(gglgbtq::palette_lgbtq("nonbinary")[c(1, 3, 4, 1)])(12),
+                     aesthetics = c("fill", "color"), guide = "none") +
+  facet_wrap(~ model, ncol = 1) +
+  tag_facets() 
 ```
@@
 
+```{r}
+trans_pink <- gglgbtq::palette_lgbtq("transgender")[2] |> 
+  colorspace::darken(0.2)
+trans_blue <- gglgbtq::palette_lgbtq("transgender")[1] |> 
+  colorspace::darken(0.2)
+
+```
+
 ```{r fig-bias}
@@
 hindcast_clim |> 
+  # _[month == 1] |>
   # dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
@@
                     breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
+  # geom_contour2(aes(z = aice - obs), breaks = AnchorBreaks(binwidth = 0.1, exclude = 0),
+  #               linewidth = 0.1) +
   scale_fill_divergent_discretised(paste0("Bias with respect to ",
                                           labels_models[which_dataset]),
-                                   low = scales::muted("red"),
-                                   high = scales::muted("blue")) +
+                                   low = trans_pink,
+                                   high = trans_blue) +
   geomcoord_antarctica +
   geom_antarctica_fill +
-  facet_wrap(~ month) +
+  facet_wrap(~ month, 
+             labeller = labeller(month = setNames(months, month.abb)),
+             nrow = 2) +
   wide_legend +
@@
 
-@fig-bias shows the difference of monthly mean sea ice concentrations between +CDR and +S2 hindcasts at the shortest lag. 
-From October to May, the model underestimates sea ice concentrations pretty much everywhere there is ice except for the deep Weddell Sea in April and May, where sea ice concentrations saturate to 1 both in the observations and forecasts.
-In winter, the differences are mostly on the sea ice edge, with light positive bias in the African sector of East Antarctica and negative bias around the Indian Ocean sector which partially compensate, resulting in the near-zero extent bias seen in those months ([Fig.
-@fig-hindcast-extent]).
+### Anomalies
 
-### Anomalies
+```{r}
+
+nb_yellow <- gglgbtq::palette_lgbtq("nonbinary")[1] |> 
+  colorspace::darken(0.3)
+nb_purple <- gglgbtq::palette_lgbtq("nonbinary")[3]
+colours_models <- c(
+  S2 = nb_purple,
+  S1 = nb_yellow,
+  cdr = "black",
+  persistence = "black",
+  climatology = "gray50"
+)
+
+labels_models <- c(
+  S2 = "ACCESS-S2",
+  S1 = "ACCESS-S1",
+  cdr = "CDR",
+  bt = "Bootstrap",
+  era5 = "ERA5",
+  osi = "OSI",
+  persistence = "Persistence",
+  climatology = "Climatology"
+)
+
+scale_color_models <-   ggplot2::scale_color_manual(NULL,
+                                                    values = colours_models,
+                                                    labels = labels_models
+)
+
+scale_fill_models <-   ggplot2::scale_fill_manual(NULL,
+                                                  values = colours_models,
+                                                  labels = labels_models
+)
+```
 
@@
   geom_line(aes(y = forecast, color = model))  +
-  geom_text(data = \(x) x[time %between% s1_range, correlate(aice, forecast, use = "complete.obs"), by = .(lag, model)],
-            aes(label = text), x = -Inf, y = Inf, 
+  geom_text(data = \(x) x[time %between% s1_range, 
+                          .(text = sqrt(mean((aice - forecast)^2, na.rm = TRUE))), by = .(lag, model)],
+            aes(label = paste0(scales::number(text, scale = 1e-12), " M km²")), x = -Inf, y = Inf,
             vjust = 1.2, hjust = -0.2) +
@@
   scale_color_models
+
 ```
 
-@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lags.
-The anomalies in this case were computed with respect of the climatology of each lag, which works as a bias-correction method. 
-Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at lag 3.
-+S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.
-
 ### RMSE
-
-To study +S2 forecasts quantitatively, we compute error measures for all hindcasts started on the 1st of every month.
 
@@
            cdo_execute() |> 
-           ReadNetCDF(c(std = "aice")) |> 
+           ReadNetCDF(c(climatology = "aice")) |> 
            _[, let(lat = NULL, lon = NULL)]
@@
   _[, time := as.Date(time)] |> 
-  _[std == 0, std := NA] |> 
-  _[std < 0.05, std := NA]
+  _[climatology == 0, climatology := NA] |> 
+  _[climatology < 0.05, climatology := NA]
 
+```
+
+```{r}
+clim_std2 <- clim_std[errors, on = .NATURAL] |> 
+  _[, let(climatology = NULL,
+          value = climatology)] |> 
+  na.omit() |> 
+  _[, version := "climatology"]
 ```
@@
 
-errors |> 
+g <- clim_std2 |> 
+  rbind(errors) |>
+  _[obs_dataset == which_dataset] |> 
   _[measure == "rmse"] |>
   _[, time2 := update(time, year = 2000)] |> 
-  # merge(clim_std) |>
-  _[obs_dataset ==  which_dataset] |>
   _[, median_ci(value),
     by = .(lag, version, measure, month(time_forecast), obs_dataset)] |>
-  _[version == "persistence", version := obs_dataset] |> 
   ggplot(aes(lag, mid)) +
@@
 
-  # geom_line(aes(y = clim_std)) +
   scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
@@
   scale_fill_models +
-  #   scale_color_models +
-  #   scale_fill_models +
   facet_wrap(~ month, labeller = labeller(month = labels_month)) 
@@
 
-::: notes
+```{r}
+g %+% g$data[version != "climatology"]
+```
 
-Use climatology
+```{r}
+g %+% g$data[version != "persistence"]
+```
 
-monthly rmse/corelation at gridpoints per lead time.
+```{r}
+p <- g
 
-:::
+p$layers <- p$layers[-4]
+p
+```
 
-@fig-rmse shows the median and 95% range of RMSE of sea ice concentration anomalies for +S2 forecasts compared with a benchmark of persistence.
-Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, after when it is statistically useful.
-Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
-The only exception is around February, where the model has lower RMSE than the persistence forecast at virtually every lag.
+```{r}
+p <- g +
+  coord_cartesian(xlim = c(0, 100))
+
+p$layers <- p$layers[-4]
+p
+```
 
@@
 
+topo_inset <- list(topo_lonlat) |>
+  rep(4) |>
+  rbindlist(idcol = "month") |>
+  ggplot(aes(lon, lat)) +
+  geom_contour_fill(aes(z = z),
+                    breaks = c(0, Inf), fill = "gray90") +
+  geom_contour2(aes(z = z),
+                breaks = 0) +
+  facet_wrap(~month, labeller = labeller(mohth = rep("", 4)), nrow = 1) +
+  scale_y_latitude(limits = c(-90, -60)) +
+  scale_x_longitude(ticks = 60, expand = c(0.08, 0)) +
+  theme_void() +
+  theme(strip.background = element_blank(),
+        strip.text = element_blank(),
+        aspect.ratio = 0.1,
+        plot.margin = margin(c(-50, 0, 0, 0)))
+
+theme_margin <- theme_get()$plot.margin
+theme_margin[3] <- margin(-10)[1]
+
+binwidth <- 0.02
 rmse_lon_mean |> 
   _[obs_dataset == which_dataset] |> 
+  # _[month %in% 1:4] |> 
   ggperiodic::periodic(lon = c(0, 360)) |> 
   ggplot(aes(lon, lag)) +
-  geom_contour_fill(aes(z = dif, fill = after_stat(level)), binwidth = 0.025)  +
-  geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.4),  binwidth = 0.025) +
-  geom_contour2(data = topo_lonlat,
-                aes(lon, scales::rescale(lat, c(1, 150)), z = z),
-                breaks = 0) +
+  geom_contour_fill(aes(z = dif, fill = after_stat(level)),
+                    breaks = AnchorBreaks(0, exclude = 0, binwidth = binwidth))  +
+  geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.4),
+                      breaks = AnchorBreaks(0, exclude = 0, binwidth = binwidth)) +
+
   geom_hline(data = month_from_lag(1, 1), aes(yintercept = lag),
@@
             hjust = 1) +
-  scale_fill_divergent_discretised() +
+  # scale_fill_divergent_discretised() +
+  scale_fill_divergent_discretised(high = trans_pink,
+                                   low  = trans_blue) +
   scale_x_longitude(ticks = 60, expand = c(0.08, 0)) +
@@
   theme(panel.background = element_blank(), 
-        tagger.panel.tag.background = element_rect(colour = "white", fill = "white")) +
-  tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1))
+        tagger.panel.tag.background = element_rect(colour = "white", fill = "white"), 
+        plot.margin = theme_margin) +
+  tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1)) +
+
+  topo_inset +
+
+  plot_layout(ncol = 1, guides = "collect", heights = c(10, 1), axes = "collect")
 ```
-
-Even though +S2 is no better than persistence at forecasting pan-Antarctic sea ice concentration anomalies, the quality of the forecast might vary between regions. 
-To analyse the spatial distribution of the model error, we computed sea ice concentration anomalies RMSE on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]`º wide.
-The median difference between the forecast and persistence RMSE is shown on @fig-rmse_lon, where negative values indicate that the model has lower RMSE than the benchmark.
-
-The skill shown by +S2 at February-March forecasts is evidenced by a band of negative values across almost all longitudes. 
-The only region where February-March sea ice is not well forecasted is east of the Antarctic Peninsula, which is the only region with consistent summer sea ice.
-This suggests that the positive skill around the summer minimum comes from the model correctly forecasting no ice where no ice is ever found. 
-However, there are other regions of skilful forecasts.
-
-Forecasts initialised on January 1^st^ ([Fig.
-@fig-rmse_lon] panel a) show RMSE values below persistence at lags larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea.
-These two regions are also relatively well forecasted from other months. 
-Notably, May forecasts of Weddell sea ice concentration anomalies ([Fig. @fig-rmse_lon] panel e) are skilful after about 20 days and remain so until December. 
-The Weddell Sea is also the region with the maximum error --around June regardless of initialisation date.
 
@@
 
-To compare +S2 with +S1, we computed the IIEE for both models.
-This error measure is shown in @fig-iiee for all lags and forecasts initialised at the first of every month.
-+S1 has lower error at short lags at all months, with the errors converging to +S2 as the forecast goes on.
-The time to convergence depends on the month and it can be as short as two weeks June and July to as long as 160 days for forecasts initialised in February.
-Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the the memory of sea ice to initial conditions --at least from October to March when the data assimilated form the other components has little to no influence on sea ice.
-
-Also evident in @fig-iiee is the difference in the error spread at short lags between +S2 and +S1.
-In all months +S1 has much narrower error spread at lag 1. 
-The error in the initial conditions is not only small, but also fairly constant.
-This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
-For +S2, this is true only only between July and October, approximately.
-For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it's not constant.
-
-```{r fig-iiee-variance}
-#| fig-cap: Mean spread of IIEE at different lags for different models. 
-#| fig.height: 6
-#| fig.width: 8
-labels_extent_2 <- function(x) {
-  m <- which.max(x)
-  x <- scales::label_number(scale = (1e-12)^2)(x)
-  x[m] <- paste0(x[m], "\nM km⁴")
-  x
-}
-
-rbind(mean_variance = errors |>
-        _[measure == "iiee"] |>
-        _[obs_dataset == which_dataset] |>
-        _[, var(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
-        _[, .(value = mean(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
-
-      variance_of_mean = errors |>
-        _[measure == "iiee"] |>
-        _[obs_dataset == which_dataset] |>
-        _[, mean(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
-        _[, .(value = var(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
-      idcol = "component") |>
-
-  ggplot(aes(lag, value)) +
-
-  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
-             colour = "gray50", linewidth = 0.1) +
-
-  geom_text(data = month_from_lag(15, 2), aes(y = Inf, label = month2), 
-            size = 2.5,
-            vjust = 1) +
-
-  geom_text(data = month_from_lag(1, 2), aes(y = -Inf, label = lag), 
-            size = 2.5,
-            vjust = 0) +
-
-  geom_line(aes(color = version), linewidth = 1) +
-  scale_y_continuous("Variance", 
-                     labels = labels_extent_2) +
-  # scale_color_models +
-  scale_x_continuous("Lag", breaks = NULL) +
-  scale_color_models +
-  ggh4x::facet_nested_wrap(month ~ component, ncol = 6, 
-                           remove_labels = "all",
-             labeller = labeller(month = labels_month, 
-                                 component = c(mean_variance = "Mean variance",
-                                               variance_of_mean = "Variance of mean"))) 
+```{r}
+last_plot() +
+  coord_cartesian(xlim = c(0, 100))
 ```
-
-The large initial error spread could be due either to large spread of ensemble members or to large spread of individual forecasts.
-@fig-iiee-variance splits the IIEE variance for each lag into the mean variance of each individual forecast and the variance of the mean error of each individual forecast, which adds up to the total variance.
-The average variance of each forecast is almost identical between forecast systems in all months.
-This shows that the ensemble spread of individual forecasts evolves identically, which, again, it's not unexpected because both systems share the same model formulation.
-This also shows that the perturbation scheme in +S2 is comparable to the one in +S1.
-
-On the other hand, the spread of the mean error is always larger in +S2 than +S1.
-The difference is particularly large at short lags in some months, which coindice with the ones in which the data assimilation scheme is not influencing sea ice initial conditions.
-
-## Conclusions
-
-Sea ice forecasts from the +S2 model show a significant low extent bias, particularly during late summer and early autumn. 
-This bias is attributed to a faster and longer melt season between January and March, and slower growth between March and April.
-This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems either with the model's thermodynamic representation or with short wave radiation forcing, as shown in other climate models.
-Even though +S2 shares the same model formulation with +S1, the latter does not suffer from this bias. 
-This is due to the assimilation of sea ice concentration anomalies into the initial conditions, which successful corrects for the negative bias in the model. 
-The analysis suggests that the data assimilation system in +S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea ice component runs virtually free, reverting to its biased equilibrium state.
-
-Analysis of the error spread shows that +S2 initial conditions from December to May not only have large errors, but that the initial error spread is very large compared with +S1. 
-This spread is not due to the perturbation scheme, since the mean error variance for individual forecasts is low and comparable with +S1. 
-Instead, it is due to large variance of the mean error of individual forecasts, which is comparable to the climatology spread. 
-This is further evidence that individual initial conditions are not being affected by the data assimilation scheme.
-
-Based on the observation that +S2 sea ice initial conditions are essentially not initialised, comparing its forecasts with +S1's allows us to estimate the time-scale for which initial conditions are important.
-This highlights February initial conditions as a crucial for determining sea ice evolution at least up to late June. 
-Arctic sea ice forecast also show greater sensitivity to initial conditions in (boreal) summer compared with (boreal) winter [@day2014; @bunzel2016], so similar mechanism might be playing a role.
-
-Although forecasts are not skilful for forecasting panantarctic sea ice concentrations, there are some areas where the model does show skill. 
-The Weddell Sea, and the Ross Sea to a lesser extent are particularly well forecasted between June and November for forecasts initialised from January to June. 
-These regions have been recognised as a regions of high predictability thanks to persistent and eastwardly advected upper ocean heat content anomalies [@bushuk2021] 
-Sea surface temperatures are assimilated by +S2 only in areas with temperatures greater than 0°C and errors don't show a clear easterly-propagating signal, so it is not clear if +S2 is leveraging the same source of predictability.
-
------
-
-# References
-
-::: {#refs}
-:::
-
-# Suplementary figures
-
-<!-- ## ERA5 -->
-
-<!-- ```{r dataset_era} -->
-
-<!-- which_dataset <- "era5" -->
-
-<!-- ``` -->
-
-<!-- ```{r fig-hindcast-extent-era5} -->
-
-<!-- which_dataset <- "era5" -->
-
-<!-- <<fig-hindcast-extent>> -->
-
-<!-- ``` -->
-
-<!-- ```{r fig-hindcast-extent-bt} -->
-
-<!-- which_dataset <- "bt" -->
-
-<!-- <<fig-hindcast-extent>> -->
-
-<!-- ``` -->
-
-<!-- ```{r fig-mean-growth} -->
-
-<!-- <<fig-mean-growth>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<fig-bias>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<<fig-extent-anom>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<fig-rmse>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<fig-iiee>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<fig-iiee-variance>> -->
-
-<!-- ``` -->

### 2025-02-25 R/ggplot.R

@@ -53,10 +53,10 @@ scale_fill_models <- ggplot2::scale_fill_manual(NULL,
                                          labels = labels_models
 )
 
-labels_extent <- function(x) {
+labels_extent <- function(x, sep = "\n") {
   m <- which.max(x)
   x <- scales::label_number(scale = 1e-12)(x)
-  x[m] <- paste0(x[m], "\nM km²")
+  x[m] <- paste0(x[m], sep, "M km²")
   x
 }
 

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
More figures for ncar

@@
 source(here::here("R/ggplot.R"))
+```
+
+```{r}
+
+nb_yellow <- gglgbtq::palette_lgbtq("nonbinary")[1] |> 
+  colorspace::darken(0.3)
+nb_purple <- gglgbtq::palette_lgbtq("nonbinary")[3]
+colours_models <- c(
+  S2 = nb_purple,
+  S1 = nb_yellow,
+  cdr = "black",
+  persistence = "black",
+  climatology = "gray50"
+)
+
+labels_models <- c(
+  S2 = "ACCESS-S2",
+  S1 = "ACCESS-S1",
+  cdr = "CDR",
+  bt = "Bootstrap",
+  era5 = "ERA5",
+  osi = "OSI",
+  persistence = "Persistence",
+  climatology = "Climatology"
+)
+
+scale_color_models <-   ggplot2::scale_color_manual(NULL,
+                                                    values = colours_models,
+                                                    labels = labels_models
+)
+
+scale_fill_models <-   ggplot2::scale_fill_manual(NULL,
+                                                  values = colours_models,
+                                                  labels = labels_models
+)
+
+trans_pink <- gglgbtq::palette_lgbtq("transgender")[2] |> 
+  colorspace::darken(0.2)
+trans_blue <- gglgbtq::palette_lgbtq("transgender")[1] |> 
+  colorspace::darken(0.2)
+
 ```
@@
 
-```{r fig-hindcast-extent}
+```{r}
+cm <- 2.54
+```
+
+```{r fig-hindcast-extent, fig.height=11.99/cm, fig.width=23.59/cm}
 #| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
@@
 
-```{r}
+```{r hindcast_extent-bias, fig.height=11.99/cm, fig.width=23.59/cm}
 hindcast_extent |> 
@@
 
+```{r extent-delta, fig.height=11.99/cm, fig.width=19.39/cm}
+
+dif_hindcast <- hindcast_extent |> 
+  copy() |> 
+  _[order(time)] |> 
+  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(model, forecast_time, member)] |> 
+  _[, median_ci(dif), by = .(model, time = update(time, year = 2000))]
+
+extent_daily |> 
+  _[order(time)] |> 
+  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
+  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
+  _[dataset == which_dataset] |> 
+  ggplot(aes(time, mid)) +
+  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
+  geom_smooth(aes(group = dataset, color = dataset),
+              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
+  geom_smooth(data = dif_hindcast,
+              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
+  scale_x_datetime(NULL, date_breaks = "1 month", 
+                   date_labels = "%b", expand = c(0, 0)) +
+  scale_y_continuous(NULL, labels = \(x) labels_extent(x, units = "M km²/day")) +
+  scale_color_models
+```
+
 ```{r hindcast_clim}
@@
 lt$year*12 + lt$mon } 
-mondf <- function(d1, d2) { monnb(d2) - monnb(d1) }
+month_differnece <- function(d1, d2) { monnb(d2) - monnb(d1) }
+
+month_diff <- function(month1, month2) {
+  d <- (month2 - month1)
+
+  (d %% 12)
+
+}
 
@@
 
-hindcast_clim <- lapply(setNames(months, months), \(month) {
-  here::here(glue::glue("data/derived/climatology/S2/{zeropad(month)}/em.nc")) |> 
-    cdo_selmonth(month) |> 
+hindcast_clim <- purrr::pmap(CJ(month = 1:12, version = c("S2", "S1")), \(month, version) {
+  here::here(glue::glue("data/derived/climatology/{version}/{zeropad(month)}/em.nc")) |> 
+    # cdo_selmonth(month) |> 
+    cdo_monmean() |> 
     cdo_execute() |> 
-    ReadNetCDF("aice")
+    ReadNetCDF("aice") |> 
+    _[, let(forecast_month = month, 
+            version = version)] 
 }) |> 
-  rbindlist(idcol = "forecast_month") |> 
-  _[, forecast_month := as.numeric(forecast_month)]
-
-hindcast_clim <- hindcast_clim |> 
-  copy() |> 
-  _[, .(aice = mean(aice)), by = .(month(time), ygrid, xgrid, forecast_month)] |> 
-  _[forecast_month == month]
+  rbindlist() |> 
+  _[, forecast_month := as.numeric(forecast_month)] |> 
+  _[, forecast_time := update(time, month = forecast_month)] |> 
+  _[, time := as.Date(update(time, day = 1))]
 
@@
   ReadNetCDF(c(obs = "aice")) |> 
-  _[, month := month(time)]
+  _[, time := as.Date(update(time, day = 1))]
 
 hindcast_clim <- hindcast_clim |> 
-  merge(obs_clim)
-```
+  merge(obs_clim) |> 
+  _[, month := month(time)] |>
+  _[, month_diff := month_differnece(time[1], forecast_time[1]), by = .(forecast_month, time)]
 
-```{r}
-trans_pink <- gglgbtq::palette_lgbtq("transgender")[2] |> 
-  colorspace::darken(0.2)
-trans_blue <- gglgbtq::palette_lgbtq("transgender")[1] |> 
-  colorspace::darken(0.2)
+hindcast_clim[, month_diff := month_diff(forecast_month[1], month[1]), by = .(forecast_month, month)]
 
@@
 ```{r fig-bias}
-#| fig-width: 9
-#| fig-height: 6
 #| fig-cap: Mean sea ice concentrations 1-month lag monthly mean +S2 forecast bias compared with NSIDC.
 
+bias_plot <- function(hindcast_clim, which_version, which_lead) {
   hindcast_clim |> 
-  # _[month == 1] |>
+    _[version == which_version] |>
+    _[month_diff == which_lead] |> 
     # dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
@@
     #               linewidth = 0.1) +
-  scale_fill_divergent_discretised(paste0("Bias with respect to ",
-                                          labels_models[which_dataset]),
+    scale_fill_divergent_discretised(paste0(which_version, " bias at leadtime ", which_lead, " months"),
                                      low = trans_pink,
@@
     tag_facets()
+}
+
+```
+
+```{r, fig.width=22.15/cm, fig.height=12.3/cm}
+hindcast_clim |> 
+  bias_plot("S2", 0)
+```
+
+```{r}
+hindcast_clim |> 
+  bias_plot("S1", 0)
+```
+
+```{r}
+hindcast_clim |> 
+  bias_plot("S2", 6)
+
+```
+
+```{r}
+hindcast_clim |> 
+  bias_plot("S1", 6)
 ```
@@
 
-```{r}
-
-nb_yellow <- gglgbtq::palette_lgbtq("nonbinary")[1] |> 
-  colorspace::darken(0.3)
-nb_purple <- gglgbtq::palette_lgbtq("nonbinary")[3]
-colours_models <- c(
-  S2 = nb_purple,
-  S1 = nb_yellow,
-  cdr = "black",
-  persistence = "black",
-  climatology = "gray50"
-)
-
-labels_models <- c(
-  S2 = "ACCESS-S2",
-  S1 = "ACCESS-S1",
-  cdr = "CDR",
-  bt = "Bootstrap",
-  era5 = "ERA5",
-  osi = "OSI",
-  persistence = "Persistence",
-  climatology = "Climatology"
-)
-
-scale_color_models <-   ggplot2::scale_color_manual(NULL,
-                                                    values = colours_models,
-                                                    labels = labels_models
-)
-
-scale_fill_models <-   ggplot2::scale_fill_manual(NULL,
-                                                  values = colours_models,
-                                                  labels = labels_models
-)
-```
-
-```{r fig-extent-anom}
+```{r fig-extent-anom, fig.height=11.99/cm, fig.width=23.22/cm}
 #| fig-cap: Sea ice extent anomalies for +S1 and +S2 (black) and +CDR (blue).
@@
 
-```{r}
+```{rrmse_persistence, fig.width=19.43/cm, fig.height=11.99/cm}
 g %+% g$data[version != "climatology"]
@@
 
-```{r}
+```{rrmse_climatology, fig.width=19.43/cm, fig.height=11.99/cm}
 g %+% g$data[version != "persistence"]
@@
 ```{r}
-p <- g
+# ACHTUNG: we have something going on with the 29th of february.
 
-p$layers <- p$layers[-4]
-p
+clim_std2 |> 
+  rbind(errors) |> 
+  _[measure == "rmse"] |> 
+  _[obs_dataset == which_dataset] |> 
+  _[, .(rmse = median(value, na.rm = TRUE)), 
+    by = .(time = update(time, year = 2001), 
+           time_forecast = update(time_forecast, year = 2001), 
+           lag,
+           version)] |> 
+  # _[version == "climatology"] |>
+  ggplot(aes(time, rmse)) +
+  geom_line(data = \(x) x[version == "climatology"],
+            aes(group = interaction(time_forecast, month(time_forecast) <= month(time)))) +
+  # geom_line(data = \(x) x[version == "S2"],
+  #           aes(color = factor(month(time_forecast)),
+  #               group = interaction(time_forecast, month(time_forecast) <= month(time)))) +
+  # geom_point(data = \(x) x[version == "S2" & lag ==1], aes(color =  factor(month(time_forecast)))) +
+  scale_color_manual(values = colorRampPalette(gglgbtq::palette_lgbtq("nonbinary")[c(1, 3, 4, 1)])(12),
+                     aesthetics = c("fill", "color"), guide = "none") +
+  scale_y_continuous(limits = c(0, NA)) +
+  geom_vline(xintercept = as.Date("2001-02-28"))
 ```
 
-```{r}
+```{r, fig.width=19.43/cm, fig.height=11.99/cm}
 p <- g +

### 2025-02-27 R/ggplot.R

@@ -53,10 +53,10 @@ scale_fill_models <- ggplot2::scale_fill_manual(NULL,
                                          labels = labels_models
 )
 
-labels_extent <- function(x, sep = "\n") {
+labels_extent <- function(x, sep = "\n", units = "M km²") {
   m <- which.max(x)
   x <- scales::label_number(scale = 1e-12)(x)
-  x[m] <- paste0(x[m], sep, "M km²")
+  x[m] <- paste0(x[m], sep, units)
   x
 }
 

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
Fiigures

@@
 library(patchwork)
+
+knitr::opts_chunk$set(cache = TRUE)
 
@@
 
-```{r extent-delta, fig.height=11.99/cm, fig.width=19.39/cm}
+```{r extent-delta, fig.height=11.99/cm, fig.width=23.59/cm}
 
@@
     #               linewidth = 0.1) +
-    scale_fill_divergent_discretised(paste0(which_version, " bias at leadtime ", which_lead, " months"),
+    scale_fill_divergent_discretised(paste0(which_version, " bias at ", which_lead, " months at lead time"),
                                      low = trans_pink,
@@
 
-```{r, fig.width=22.15/cm, fig.height=12.3/cm}
+```{r}
+width_bias <- 22.15/cm
+height_bias <- 12.3/cm
+```
+
+```{r bias_S2_0, fig.width=width_bias, fig.height=height_bias}
 hindcast_clim |> 
@@
 
-```{r}
+```{r bias_S1_0, fig.width=width_bias, fig.height=height_bias}
 hindcast_clim |> 
@@
 
-```{r}
+lead
+```{r bias_S2_6, fig.width=width_bias, fig.height=height_bias}
 hindcast_clim |> 
@@
 
-```{r}
+```{r bias_S1_6, fig.width=width_bias, fig.height=height_bias}
 hindcast_clim |> 
@@
   labs(y = "RMSE", 
-       x = "Lag") +
+       x = "Lead time") +
   scale_color_models +
@@
 
-```{rrmse_persistence, fig.width=19.43/cm, fig.height=11.99/cm}
+```{r}
+width_errors <- 19.43/cm
+height_errors <- 11.99/cm
+```
+
+```{rrmse_persistence, fig.width=width_errors, fig.height=height_errors}
 g %+% g$data[version != "climatology"]
@@
 
-```{rrmse_climatology, fig.width=19.43/cm, fig.height=11.99/cm}
+```{rrmse_climatology, fig.width=width_errors, fig.height=height_errors}
 g %+% g$data[version != "persistence"]
@@
 
-```{r, fig.width=19.43/cm, fig.height=11.99/cm}
+```{r error_zoom, fig.width=width_errors, fig.height=height_errors}
 p <- g +
@@
 
-```{r fig-rmse_lon, fig.height=7, fig.width=10}
+```{r fig-rmse_lon, fig.height=11.99/cm, fig.width=23.1/cm}
 #| fig.cap: Meadian difference between RMSE of +S2 forecasts and persistence forecast at longitudinal bands.
@@
   labs(fill = NULL, 
-       y = "Lag") +
+       y = "Lead time") +
   wide_legend +
@@
 ```
-
-### Comparison with S1
-
-```{r}
-t_test <- function(x, y, ...) {
-
-  test <- t.test(x, y, ...)
-
-  list(low = test[["conf.int"]][1],
-       high = test[["conf.int"]][2], 
-       estimate = -diff(test$estimate), 
-       p.value = test[["p.value"]]
-  )
-}
-
-max_lag <- errors[version == "S1", max(lag)]
-
-dif <- errors |> 
-  na.omit() |> 
-  _[obs_dataset == which_dataset] |>
-  _[measure == "iiee"] |> 
-  _[version != "persistence"] |> 
-  _[lag <= max_lag] |> 
-  dcast(time + time_forecast + member + lag ~ version, value.var = "value") |> 
-  _[, t_test(na.omit(S1), na.omit(S2)), 
-    by = .(lag, month(time_forecast))] |> 
-  _[p.value > 0.01] |> 
-  _[, .SD[which.min(lag)], by = month] |> 
-  _[order(month), .(month, lag)]
-
-labels_month_significant <- setNames(paste0(month.abb, " (", dif$lag, ")"),
-                                     1:12)
-```
-
-```{r fig-iiee}
-#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lag for all forecast initialised on the first of each month for +S1 and +S2 hindcasts.
-For each month, the number in parenthesis indicates the minimum lag at which the the mean error of each model is not statistically significant at a 1% level using a two-sisded t-test.
-
-errors |> 
-  _[measure == "iiee"] |>
-  _[obs_dataset == which_dataset] |>
-  # _[, month := lubridate::month(time_forecast, label = TRUE)]  |> 
-  _[, median_ci(value), 
-    by = .(lag, version, measure, obs_dataset, month(time_forecast))] |> 
-  ggplot(aes(lag, mid)) +
-
-  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
-             colour = "gray50", linewidth = 0.1) +
-
-  geom_text(data = month_from_lag(15, 2), aes(y = Inf, label = month2), 
-            size = 2.5,
-            vjust = 1) +
-
-  geom_text(data = month_from_lag(1, 2), aes(y = -Inf, label = lag), 
-            size = 2.5,
-            vjust = 0) +
-
-  geom_ribbon(aes(ymin = low, 
-                  ymax = high, 
-                  color = version, 
-                  fill = version), alpha = 0.1) +
-  geom_line(aes(color = version), linewidth = 1) +
-
-  # geom_errorbar(data = \(x) x[dif, on = .NATURAL][version == "S1"], 
-  #              aes(x = lag, 
-  #                  ymin = mid - (high - low), ymax = mid + (high - low))) +
-
-  # geom_point(data = \(x) x[dif, on = .NATURAL][version == "S1"],
-  # aes(lag, mid), size = 3, shape = 21) +
-
-  scale_y_continuous(labels = labels_extent) +
-  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
-  scale_color_models +
-  scale_fill_models +
-  labs(y = NULL, 
-       x = "Lag") +
-  #   scale_color_models +
-  #   scale_fill_models +
-  facet_wrap(~ month, labeller = labeller(month = labels_month_significant)) 
-```
-
-```{r}
-last_plot() +
-  coord_cartesian(xlim = c(0, 100))
-```

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
Modifies delta

@@
 
-```{r extent-delta, fig.height=11.99/cm, fig.width=23.59/cm}
+```{r extent-delta, fig.height=11.99/cm, fig.width=24.45/cm}
 
@@
   scale_y_continuous(NULL, labels = \(x) labels_extent(x, units = "M km²/day")) +
+  theme(legend.position = "inside", legend.position.inside = c(0.9, 0.9), 
+        legend.direction = "vertical") +
   scale_color_models

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
More delta tweaks
Changes breaks in lon_rmse plots

@@
 
-```{r extent-delta, fig.height=11.99/cm, fig.width=24.45/cm}
+```{r extent-delta, fig.height=11.99/cm, fig.width=24.09/cm}
 
@@
 
-```{r fig-bias}
+```{r bias_plot}
 #| fig-cap: Mean sea ice concentrations 1-month lag monthly mean +S2 forecast bias compared with NSIDC.
@@
 bias_plot <- function(hindcast_clim, which_version, which_lead) {
+
+  breaks <- seq(-1, 1, by = 0.1)
+  breaks <- breaks[breaks != 0]
   hindcast_clim |> 
@@
     geom_contour_fill(aes(z = aice - obs, fill = after_stat(level)), 
-                      breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
+                      breaks = breaks) +
     # geom_contour2(aes(z = aice - obs), breaks = AnchorBreaks(binwidth = 0.1, exclude = 0),
@@
 
-```{r bias_S2_0, fig.width=width_bias, fig.height=height_bias}
+```{r bias_S2_0, fig.width=width_bias, fig.height=height_bias, dependson="bias_plot"}
+
 hindcast_clim |> 
@@
 
-```{r bias_S1_0, fig.width=width_bias, fig.height=height_bias}
+```{r bias_S1_0, fig.width=width_bias, fig.height=height_bias, dependson="bias_plot"}
+
 hindcast_clim |> 
@@
 lead
-```{r bias_S2_6, fig.width=width_bias, fig.height=height_bias}
+```{r bias_S2_6, fig.width=width_bias, fig.height=height_bias, dependson="bias_plot"}
 hindcast_clim |> 
@@
 
-```{r bias_S1_6, fig.width=width_bias, fig.height=height_bias}
+```{r bias_S1_6, fig.width=width_bias, fig.height=height_bias, dependson="bias_plot"}
 hindcast_clim |> 
   bias_plot("S1", 6)
+
 ```

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
Names rmse figures

@@
     facet_wrap(~ month, 
-               labeller = labeller(month = setNames(months, month.abb)),
+               labeller = labeller(month = setNames(1:12, month.abb)),
                nrow = 2) +
@@
 
-```{rrmse_persistence, fig.width=width_errors, fig.height=height_errors}
+```{r rmse_persistence, fig.width=width_errors, fig.height=height_errors}
 g %+% g$data[version != "climatology"]
@@
 
-```{rrmse_climatology, fig.width=width_errors, fig.height=height_errors}
+```{r rmse_climatology, fig.width=width_errors, fig.height=height_errors}
 g %+% g$data[version != "persistence"]

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
CHanges size climatology
More figure tweaks

@@
     facet_wrap(~ month, 
-               labeller = labeller(month = setNames(1:12, month.abb)),
+               labeller = labeller(month = labels_month),
                nrow = 2) +
@@
 ```{r}
-width_errors <- 19.43/cm
+width_errors <- 23.87/cm
 height_errors <- 11.99/cm
@@
 
-```{r fig-rmse_lon, fig.height=11.99/cm, fig.width=23.1/cm}
+```{r fig-rmse_lon, fig.height=12.6/cm, fig.width=23.1/cm}
 #| fig.cap: Meadian difference between RMSE of +S2 forecasts and persistence forecast at longitudinal bands.

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
Tweaks

@@
         strip.text = element_blank(),
-        aspect.ratio = 0.1,
-        plot.margin = margin(c(-50, 0, 0, 0)))
-
-theme_margin <- theme_get()$plot.margin
-theme_margin[3] <- margin(-10)[1]
+        aspect.ratio = 0.1)
 
@@
   theme(panel.background = element_blank(), 
-        tagger.panel.tag.background = element_rect(colour = "white", fill = "white"), 
-        plot.margin = theme_margin) +
+        tagger.panel.tag.background = element_rect(colour = "white", fill = "white")) +
   tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1)) +
+  coord_cartesian(clip = "off") +
 

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
Improves delta

@@
 
-```{r extent-delta, fig.height=11.99/cm, fig.width=24.09/cm}
-
+```{r extent-delta-compute, fig.height=11.99/cm, fig.width=24.09/cm}
 dif_hindcast <- hindcast_extent |> 
@@
 
-extent_daily |> 
+dif_hindcast <- rbind(
+  dif_hindcast |> 
+    copy() |> 
+    _[, time := update(time, year = 1999)],
+  dif_hindcast,
+  dif_hindcast |> 
+    copy() |> 
+    _[, time := update(time, year = 2001)]
+)
+
+dif_extent <- extent_daily |> 
   _[order(time)] |> 
@@
   _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
-  _[dataset == which_dataset] |> 
+  _[dataset == which_dataset]
+
+dif_extent <- rbind(
+  dif_extent |> 
+    copy() |> 
+    _[, time := update(time, year = 1999)],
+  dif_extent,
+  dif_extent |> 
+    copy() |> 
+    _[, time := update(time, year = 2001)]
+)
+```
+
+```{r extent-delta, fig.height=11.99/cm, fig.width=24.09/cm}
+n = 366*3
+
+dif_extent |> 
   ggplot(aes(time, mid)) +
@@
   geom_smooth(aes(group = dataset, color = dataset), 
-              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
+              method = "loess",
+              span = 30/n, linewidth = 0.5, se = FALSE, n = n)  +
+  # geom_line() +
   geom_smooth(data = dif_hindcast,
-              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
+              method = "loess",
+              aes(color = model), span = 30/n, linewidth = 0.5, se = FALSE, n = n) +
   scale_x_datetime(NULL, date_breaks = "1 month", 
@@
         legend.direction = "vertical") +
-  scale_color_models
+  scale_color_models +
+  coord_cartesian(xlim = as.POSIXct(c("2000-01-01", "2000-12-31"), tz = "UTC"))
 ```

### 2025-03-07 analysis/paper/access-ice.qmd
Changes antarctica map and other stuff

@@
 The +S2 hindcast used in this study runs for the period 1981--2018.
-Each forecast consists of 9 ensemble members, 3 of which run for 279 days and 6 for 42 days.
-Anomalies are be taken with respect to the 1981--2011 climatology computed from the daily +S2 reanalysis and smoothed with a 11 day running mean.
+Each forecast consists of 9 ensemble members built from three consecutive 3-member forecasts initialised at the first of every month and the two previous days and run from 279 days. 
+The +S1 hindcast is built in the same manner and running for the period 1990--2012.
 
-The +S1 hindcast runs for the period 1990--2012, with 7-member forecasts integrated for 217 days. 
-No reanalysis is available so we are not able to compute a climatology for this forecast, therefore we don't show metrics based on anomalies.
+Anomalies for each hindcast are be taken with respect to a forecast-dependent climatology for the period 1990--2012. 
+This serves as a first-order correction of model drift.
 
@@
 
-```{r extent_climatology}
-extent_climatology <- extent_daily |>  
-  copy() |>
-  _[aice == 0, aice := NA] |> 
-  _[!(month(time) == 2 & mday(time) == 29)] |> 
-  _[, time := update(time, year = 2001)] |> 
-  _[, median_ci(aice), by = .(time, dataset)]
-
-```
-
-```{r area_daily}
-area_daily <- datasets |>
-  lapply(area) |>
-  lapply(\(x) ReadNetCDF(x, "aice")) |>
-  rbindlist(idcol = "dataset") |>
-  _[, let(lon = NULL, lat = NULL)]
-```
-
-```{r area_climatology}
-area_climatology <- area_daily |>  
-  copy() |>
-  _[aice == 0, aice := NA] |> 
-  _[!(month(time) == 2 & mday(time) == 29)] |> 
-  _[, time := update(time, year = 2001)] |> 
-  _[, median_ci(aice), by = .(time, dataset)]
-
-```
-
 ```{r hindcast_extent}
@@
 
+```{r extent_climatology}
+ranges <- hindcast_extent[, .(range = list(range(time))), by = model]
+
+extent_climatology <- purrr::pmap(ranges, \(model, range) {
+  extent_daily |>  
+    _[time %between% range] |> 
+    _[aice == 0, aice := NA] |> 
+    _[!(month(time) == 2 & mday(time) == 29)] |> 
+    _[, time := update(time, year = 2001)] |> 
+    _[, median_ci(aice), by = .(time, dataset)] |> 
+    _[, model := model] |> 
+    _[]
+}) |> 
+  rbindlist()
+```
+
+```{r range_max_lag}
+range_max_lag <- hindcast_extent[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)] |> 
+  _[, .(max = max(lag),
+        min = min(lag)),
+    by = model]
+```
+
 ```{r fig-hindcast-extent}
-#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
-hindcast_extent |> 
+#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month with the median sea ice extent of +CDR in the left column and the median difference in the right column.
+The +CDR climatology is computed in the period corresponding to each hindcast.
+Circles represent the initial conditions at the first of every month and triangles represent the median value at the first of every month forecasted with the largest possible lead time.
+Colours indicate the initialisation month of the forecast.
+
+hindcast_bias_extent <- hindcast_extent |> 
   _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
@@
   _[, month := lubridate::month(forecast_time, label = TRUE)] |> 
-  # _[]
   ggplot(aes(time2, mid)) +
@@
                      aesthetics = c("fill", "color"), guide = "none") +
-  facet_wrap(~ model, ncol = 1) +
+  facet_wrap(~ model, ncol = 1, labeller = labeller(model = labels_models)) +
   tag_facets()
+
+hindcast_bias_extent_dif <- copy(hindcast_bias_extent)
+
+hindcast_bias_extent_dif$layers[[1]] <- geom_hline(yintercept = 0) 
+hindcast_bias_extent_dif$data <- extent_daily[hindcast_extent, on = c("time")] |> 
+  _[, median_ci(i.aice - aice), 
+    by = .(forecast_time = update(forecast_time, year = 2001),
+           lag = as.numeric(as.Date(time) - forecast_time),
+           model)] |> 
+  _[, time := forecast_time + lag] |>
+  _[, time2 := update(time, year = 2001)] |>
+  _[, month := lubridate::month(forecast_time, label = TRUE)]
+
+hindcast_bias_extent +
+  hindcast_bias_extent_dif
+
 ```
 
-@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR +S2 shows a severe low extent bias, especially in the late summer-early autumn.
+@fig-hindcast-extent a shows the median sea ice extent (left column) and median difference with respect of +CDR for the +S2 and +S1 hindcasts.
+Median extent of initial conditions at the first of every month are indicated with circles, while median extent at the same date but forecasted with the largest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
+At this large lead time, the the information of the initial conditions is essentially lost and the forecast reverts to each model's preferred equilibrium state.
+
+Both models' equilibrium state (triangles) show low bias of sea ice extent, particularly in the late-autumn and winter months.
 This is due primarily to faster melt during a longer melt season between January and March and slower growth during March and April. 
-This is then balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
+This is then partially balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
 Many sea ice models exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
+It is not surprising that both forecasting systems converge to a similar climatology because they share the same model configuration.
 
-+S1 has a smaller low bias than +S2, especially at low lags ([Fig. @fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig. @fig-mean-growth]).
-At larger lags, +S1's bias in summer and autumn is very similar to +S2's.
-This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1 and not in +S2.
++S2 initial conditions (circles) also have a low extent bias, especially in the late summer-early autumn, while +S1 initial conditions are very close to observations. 
+The faster growth between May and July reduces the low bias in +S2 for forecasts initialised from April to June, but it adds a high bias in +S1.
 
-At large lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
-Therefore we can estimate the latter using the hindcasts with the largest possible lag, which is shown in triangles in @fig-hindcast-extent for the same dates as the initial conditions.
-The difference between the two is the effect of the data assimilation.
+The difference between the initial conditions and the model preferred state can be attributed to the effect of data assimilation, which in +S2 is due solely to atmospheric and oceanic data assimilation.
+From June to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
+The rest of the year there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.
 
-The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to them having the same model formulation.
-From June to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and improving the initial conditions.
-The rest of the year, there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.
-In +S1, on the other hand, circles are very close to observations year-round.
-
-```{r fig-mean-growth}
-#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations. Values are smoothed with a 2-degree loess smooth with a 30 day window.
-
+```{r extent-delta-compute, fig.height=11.99/cm, fig.width=24.09/cm}
 dif_hindcast <- hindcast_extent |> 
@@
   _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
-  _[dataset == which_dataset] |> 
+  _[dataset == which_dataset]
+
+dif_extent <- rbind(
+  dif_extent |> 
+    copy() |> 
+    _[, time := update(time, year = 1999)],
+  dif_extent,
+  dif_extent |> 
+    copy() |> 
+    _[, time := update(time, year = 2001)]
+)
+```
+
+```{r fig-mean-growth}
+#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations. Values are smoothed with a 2-degree loess smooth with a 30 day window.
+
+n = 366*3
+dif_extent |> 
   ggplot(aes(time, mid)) +
@@
   geom_smooth(aes(group = dataset, color = dataset), 
-              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
+              method = "loess",
+              span = 30/n, linewidth = 0.5, se = FALSE, n = n)  +
+  # geom_line() +
   geom_smooth(data = dif_hindcast,
-              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
+              method = "loess",
+              aes(color = model), span = 30/n, linewidth = 0.5, se = FALSE, n = n) +
   scale_x_datetime(NULL, date_breaks = "1 month", 
                    date_labels = "%b", expand = c(0, 0)) +
-  scale_y_continuous(NULL, labels = \(x) paste0(labels_extent(x), "/day")) +
-  scale_color_models
+  scale_y_continuous(NULL, labels = \(x) labels_extent(x, units = "M km²/day")) +
+  scale_color_models +
+  coord_cartesian(xlim = as.POSIXct(c("2000-01-01", "2000-12-31"), tz = "UTC"))
 ```
@@
 
+ice_shelves_nsicd <- fread(here::here("data/derived/antarctic_ice_shelves_nsidc.csv"))
+
 hindcast_clim |> 
@@
                                    high = scales::muted("blue")) +
+  # geom_polygon(data = ice_shelves_nsicd, aes(X, Y, group = group), 
+  #              colour = "black", fill = "#FAFAFA") +
   geomcoord_antarctica +
@@
 
-errors |> 
+clim_std2 |> 
+  rbind(errors) |>
   _[measure == "rmse"] |>
@@
     by = .(lag, version, measure, month(time_forecast), obs_dataset)] |>
-  _[version == "persistence", version := obs_dataset] |> 
   ggplot(aes(lag, mid)) +
@@
 
-  geom_ribbon(aes(ymin = low, 
-                  ymax = high, 
-                  color = version, 
-                  fill = version,
-                  group = interaction(obs_dataset, version)),
-              alpha = 0.1) +
+  # geom_ribbon(aes(ymin = low, 
+  #                 ymax = high, 
+  #                 color = version, 
+  #                 fill = version,
+  #                 group = interaction(obs_dataset, version)),
+  #             alpha = 0.1) +
   geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
@@
 ::: notes
-
-Use climatology
 
@@
 ```{r}
-topo_lonlat <- rcdo::cdo_topo(grid = "r180x90") |>
+topo_lonlat <- rcdo::cdo_topo(grid = "r360x180") |>
   rcdo::cdo_execute(options = c("-f nc")) |>
@@
   ggperiodic::periodic(lon = c(0, 360))
+
+topo_lonlat <- here::here("data/raw/ETOPO.nc") |> 
+  ReadNetCDF() |> 
+  na.omit() 
 ```
@@
 
+binwidth <- 0.015
+
 rmse_lon_mean |> 
@@
   ggplot(aes(lon, lag)) +
-  geom_contour_fill(aes(z = dif, fill = after_stat(level)), binwidth = 0.025)  +
-  geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.4),  binwidth = 0.025) +
+  geom_contour_fill(aes(z = dif, fill = after_stat(level)),
+                    breaks = AnchorBreaks(0, exclude = 0, binwidth = binwidth))  +
+  geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.4),
+                      breaks = AnchorBreaks(0, exclude = 0, binwidth = binwidth)) +
+
+  geom_contour_fill(data = topo_lonlat,
+                    aes(lon, scales::rescale(lat, c(-60, 1)), 
+                        z = z, 
+                        fill = NULL),
+                    fill = "#FAFAFA",
+                    breaks = c(0, Inf)) +
+
   geom_contour2(data = topo_lonlat,
-                aes(lon, scales::rescale(lat, c(1, 150)), z = z),
+                aes(lon, scales::rescale(lat, c(-60, 1)), z = z),
                 breaks = 0) +
@@
   facet_wrap(~ month, labeller = labeller(month = labels_month)) +
-  theme(panel.background = element_blank(), 
+  theme(panel.background = element_blank(), panel.grid = element_blank(),
         tagger.panel.tag.background = element_rect(colour = "white", fill = "white")) +

### 2025-03-07 R/ggplot.R

@@ -15,15 +15,16 @@ element_grob.element_text_first_line <-
 
 
 ggplot2::theme_set(ggplot2::theme_minimal() +
-            ggplot2::theme(panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
-                  legend.position = "bottom",
-                  legend.title.position = "top", 
-                  legend.title = ggplot2::element_text(hjust = 0.5),
-                  legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
-                  legend.key.height = grid::unit(0.75, "lines"),
-                  axis.text.y = element_text_first_line()
-            ))
+                     ggplot2::theme(panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
+                                    legend.position = "bottom",
+                                    legend.title.position = "top", 
+                                    legend.title = ggplot2::element_text(hjust = 0.5),
+                                    legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
+                                    legend.key.height = grid::unit(0.75, "lines"),
+                                    axis.text.y = element_text_first_line()
+                     ))
 wide_legend <- ggplot2::theme(legend.key.width = grid::unit(1, 'null'))
+
 
 colours_models <- c(
   S2 = "black",
@@ -31,7 +32,9 @@ colours_models <- c(
   cdr = "#1a5fb4",
   bt = "#1a5fb4",
   era5 = "#1a5fb4",
-  osi = "#1a5fb4"
+  osi = "#1a5fb4",
+  persistence = "black",
+  climatology = "gray50"
 )
 
 labels_models <- c(
@@ -40,17 +43,19 @@ labels_models <- c(
   cdr = "CDR",
   bt = "Bootstrap",
   era5 = "ERA5",
-  osi = "OSI"
+  osi = "OSI",
+  persistence = "Persistence",
+  climatology = "Climatology"
 )
 
 scale_color_models <- ggplot2::scale_color_manual(NULL,
-                                         values = colours_models,
-                                         labels = labels_models
+                                                  values = colours_models,
+                                                  labels = labels_models
 )
 
 scale_fill_models <- ggplot2::scale_fill_manual(NULL,
[…]

### 2025-03-07 scripts/get_topo.R

@@ -1,0 +1,19 @@
+library(rcdo)
+# https://www.ngdc.noaa.gov/thredds/dodsC/global/ETOPO2022/60s/60s_surface_elev_netcdf/ETOPO_2022_v1_60s_N90W180_surface.nc.html
+url <- "https://www.ngdc.noaa.gov/thredds/dodsC/global/ETOPO2022/60s/60s_surface_elev_netcdf/ETOPO_2022_v1_60s_N90W180_surface.nc?lat[0:1:1800],lon[0:1:21599],z[0:10:1800][0:1:21599],crs"
+
+etopo_file <- here::here("data/raw/ETOPO.nc")
+
+url |> 
+  cdo_copy() |> 
+  cdo_remapmean("r360x180") |> 
+  cdo_execute(etopo_file, options = c("-f nc"))
+
+metR::ReadNetCDF(etopo_file) |> 
+  data.table::copy() |> 
+  # _[, z := Smooth2D(lon, lat, z)] |> 
+  ggplot(aes(lon, lat)) +
+  geom_contour_fill(aes(z = z)) +
+  geom_contour2(aes(z = z), breaks = 0)
+
+

### 2025-03-07 analysis/paper/access-ice.qmd
Remvoes 95% reange

@@
   geom_line(aes(y = forecast, color = model))  +
-  geom_text(data = \(x) x[time %between% s1_range, correlate(aice, forecast, use = "complete.obs"), by = .(lag, model)],
-            aes(label = text), x = -Inf, y = Inf, 
+  geom_text(data = \(x) x[time %between% s1_range, 
+                          .(text = sqrt(mean((aice - forecast)^2, na.rm = TRUE))), 
+                          by = .(lag, model)],
+            aes(label = paste0(scales::number(text, scale = 1e-12), " M km²")), x = -Inf, y = Inf,
             vjust = 1.2, hjust = -0.2) +
@@
 ```{r fig-rmse}
-#| fig-cap: Median and 95% coverage of sea ice concentration anomalies RMSE as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies.
+#| fig-cap: Median RMSE of sea ice concentration anomalies as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies.
 #| fig.height: 5

### 2025-03-07 analysis/paper/access-ice.qmd
Tidy things up

@@
 
-```{r extent-delta-compute, fig.height=11.99/cm, fig.width=24.09/cm}
+```{r extent-delta-compute}
 dif_hindcast <- hindcast_extent |> 
@@
 
-```{r}
+```{r clim_std2}
 clim_std2 <- clim_std[errors, on = .NATURAL] |> 
@@
 
-```{r}
-topo_lonlat <- rcdo::cdo_topo(grid = "r360x180") |>
-  rcdo::cdo_execute(options = c("-f nc")) |>
-  metR::ReadNetCDF(c(z = "topo"), subset = list(lat = c(-90, -60))) |> 
-  ggperiodic::periodic(lon = c(0, 360))
-
+```{r topo_lonlat}
 topo_lonlat <- here::here("data/raw/ETOPO.nc") |> 
@@
 
-```{r}
+```{r t_test}
 t_test <- function(x, y, ...) {

### 2025-03-07 analysis/paper/access-ice.tex
renders

@@
 transferred to institution/company.}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, frame hidden, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 The ACCESS-S2 hindcast used in this study runs for the period
-1981--2018. Each forecast consists of 9 ensemble members, 3 of which run
-for 279 days and 6 for 42 days. Anomalies are be taken with respect to
-the 1981--2011 climatology computed from the daily ACCESS-S2 reanalysis
-and smoothed with a 11 day running mean.
+1981--2018. Each forecast consists of 9 ensemble members built from
+three consecutive 3-member forecasts initialised at the first of every
+month and the two previous days and run from 279 days. The ACCESS-S1
+hindcast is built in the same manner and running for the period
+1990--2012.
 
-The ACCESS-S1 hindcast runs for the period 1990--2012, with 7-member
-forecasts integrated for 217 days. No reanalysis is available so we are
-not able to compute a climatology for this forecast, therefore we don't
-show metrics based on anomalies.
+Anomalies for each hindcast are be taken with respect to a
+forecast-dependent climatology for the period 1990--2012. This serves as
+a first-order correction of model drift.
 
@@
 hindcasts initialised the first of the month for ACCESS-S2 and ACCESS-S1
-in colours representing the start month. In black, the median sea ice
-extent of NOAA/NSIDC CDRV4.}
+in colours representing the start month with the median sea ice extent
+of NOAA/NSIDC CDRV4 in the left column and the median difference in the
+right column. The NOAA/NSIDC CDRV4 climatology is computed in the period
+corresponding to each hindcast. Circles represent the initial conditions
+at the first of every month and triangles represent the median value at
+the first of every month forecasted with the largest possible lead time.
+Colours indicate the initialisation month of the forecast.}
 
@@
 
-Figure~\ref{fig-hindcast-extent} a shows the seasonal cycle of Sea Ice
-Extent for the ACCESS-S2 hindcast and NOAA/NSIDC CDRV4 ACCESS-S2 shows a
-severe low extent bias, especially in the late summer-early autumn. This
-is due primarily to faster melt during a longer melt season between
-January and March and slower growth during March and April. This is then
+Figure~\ref{fig-hindcast-extent} a shows the median sea ice extent (left
+column) and median difference with respect of NOAA/NSIDC CDRV4 for the
+ACCESS-S2 and ACCESS-S1 hindcasts. Median extent of initial conditions
+at the first of every month are indicated with circles, while median
+extent at the same date but forecasted with the largest lead time
+possible for each model (between 274 and 277 days for ACCESS-S2 and
+between 213 and 216 days for ACCESS-S1). At this large lead time, the
+the information of the initial conditions is essentially lost and the
+forecast reverts to each model's preferred equilibrium state.
+
+Both models' equilibrium state (triangles) show low bias of sea ice
+extent, particularly in the late-autumn and winter months. This is due
+primarily to faster melt during a longer melt season between January and
+March and slower growth during March and April. This is then partially
 balanced with faster growth between May and July
@@
 season \citep{massonnet2023} and could indicate problems in the
-representation of thermodynamics in the model \citep{zampieri2019}.
+representation of thermodynamics in the model \citep{zampieri2019}. It
+is not surprising that both forecasting systems converge to a similar
+climatology because they share the same model configuration.
 
-ACCESS-S1 has a smaller low bias than ACCESS-S2, especially at low lags
-(Fig.~\ref{fig-hindcast-extent}) even though the typical growth rates
-are very similar between both models (Fig.~\ref{fig-mean-growth}). At
-larger lags, ACCESS-S1's bias in summer and autumn is very similar to
-ACCESS-S2's. This suggests that this lower sea ice state is closer to
-the models' equilibrium, indicating that it is an issue with model
-formulation that was being corrected by the data assimilation system in
-ACCESS-S1 and not in ACCESS-S2.
+ACCESS-S2 initial conditions (circles) also have a low extent bias,
+especially in the late summer-early autumn, while ACCESS-S1 initial
+conditions are very close to observations. The faster growth between May
+and July reduces the low bias in ACCESS-S2 for forecasts initialised
+from April to June, but it adds a high bias in ACCESS-S1.
 
-At large lags, sea ice extent loses most of the initial condition memory
-and reverts to the model's preferred equilibrium state. Therefore we can
-estimate the latter using the hindcasts with the largest possible lag,
-which is shown in triangles in Figure~\ref{fig-hindcast-extent} for the
-same dates as the initial conditions. The difference between the two is
-the effect of the data assimilation.
-
-The equilibrium of ACCESS-S1 and ACCESS-S2 is very similar (comparing
-the triangles in each panel in Fig.~\ref{fig-hindcast-extent}), owing to
-them having the same model formulation. From June to October, in
-ACCESS-S2 circles are closer to observations than to the triangles,
-indicating that the information from the ocean and atmosphere data
-assimilation is getting to sea ice and improving the initial conditions.
-The rest of the year, there is little if any difference between circles
-and triangles in ACCESS-S2, indicating that almost no data assimilation
-is taking place and the sea ice component of the model is virtually
-free-running. In ACCESS-S1, on the other hand, circles are very close to
-observations year-round.
+The difference between the initial conditions and the model preferred
+state can be attributed to the effect of data assimilation, which in
+ACCESS-S2 is due solely to atmospheric and oceanic data assimilation.
+From June to October, in ACCESS-S2 circles are closer to observations
+than to the triangles, indicating that the information from the ocean
+and atmosphere data assimilation is affecting sea ice and improving the
+initial conditions. The rest of the year there is little if any
+difference between circles and triangles in ACCESS-S2, indicating that
+almost no data assimilation is taking place and the sea ice component of
+the model is virtually free-running.
 
@@
 
-\caption{\label{fig-rmse}Median and 95\% coverage of sea ice
-concentration anomalies RMSE as a function of forecast lag for all
-forecast initialised on the first of each month compared with a
-reference forecast of persistence of anomalies.}
+\caption{\label{fig-rmse}Median RMSE of sea ice concentration anomalies
+as a function of forecast lag for all forecast initialised on the first
+of each month compared with a reference forecast of persistence of
+anomalies.}
 
 \end{figure}
-
-Use climatology
 

### 2025-03-07 analysis/poster_ncar/presentation_figures.Rmd
Figures for NCAR

@@
   _[lag > 0] |> 
-  _[!(month(forecast_time) == 1 & member == 7 & model == "S1")]  # This member is wrong
+  _[!(month(forecast_time) == 1 & member == 7 & model == "S1")]  # This member is wrong =
 ```
@@
 
+```{r}
+
+hindcast_extent |> 
+  copy() |> 
+  _[, aice := aice - mean(aice[year(time) %between% c(1981, 2011)]), 
+    by = .(lag, model, time = update(time, year = 2000))] |> 
+  _[, .(forecast = mean(aice)), by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
+  _[, lag := months_difference(time, forecast_time)] |> 
+  _[, merge(.SD, monthly_extent, all = TRUE), by = .(model, lag)] |> 
+  _[time %between% s1_range] |> 
+  _[, .(text = sqrt(mean((aice - forecast)^2, na.rm = TRUE))), by = .(lag, model, dataset)] |> 
+  ggplot(aes(lag, text)) +
+  geom_line(aes(color = model)) +
+  facet_wrap(~dataset)
+
+```
+
+```{r}
+
+hindcast_extent |> 
+  copy() |> 
+  setnames("aice", "forecast") |> 
+  _[, .(forecast = mean(forecast)), by = .(time, forecast_time, lag, model)] |>
+  merge(extent_daily) |> 
+  _[dataset == which_dataset] |> 
+  _[, clim := year(time) %between% c(1981, 2011)] |> 
+  _[, let(aice = aice - mean(aice[clim == TRUE]),
+          forecast = forecast - mean(forecast[clim == TRUE])),
+    by = .(dataset, model, yday(time), forecast_time)] |>
+  _[time %between% s1_range] |> 
+  _[] |> 
+  _[, .(rmse = sqrt(mean((aice - forecast)^2, na.rm = TRUE))), by = .(model, lag, dataset)] |> 
+  ggplot(aes(lag, rmse)) +
+  geom_line(aes(color = model)) +
+  scale_y_continuous(labels = labels_extent)
+
+```
+
 ### RMSE
@@
   _[, version := "climatology"]
+```
+
+```{r}
+clim_std |> 
+  _[obs_dataset == which_dataset] |> 
+  _[, .(value = median(climatology, na.rm = TRUE)), by = .(time = update(time, year = 2000))] |> 
+  ggplot(aes(time, value)) +
+  geom_line()
 ```
@@
   rbind(errors) |>
+  # errors
   _[measure == "rmse"] |> 
@@
   geom_vline(xintercept = as.Date("2001-02-28"))
+```
+
+```{r}
+clim_std2[version == "climatology"] |> 
+  _[obs_dataset == which_dataset] |> 
+  _[member == "01"] |>
+  _[, .(rmse = median(value, na.rm = TRUE)), 
+    by = .(time = update(time, year = 2001), 
+           time_forecast = update(time_forecast, year = 2001), 
+           lag,
+           version)] |>
+  ggplot(aes(time, rmse)) +
+  geom_line()
+
 ```

### 2025-04-18 analysis/paper/access-ice.tex
Transfers all comments

@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 \usepackage[inline]{trackchanges}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 transferred to institution/company.}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 the start of satellite records in the early 80s, Antarctic sea ice has
-experienced a slightly increasing trend that models systematically
-failed to reproduce and puzzled researchers \citep{hobbs2016}. Then, in
-2016 Antarctic sea ice extent dropped precipitously and has been at low
-and record low levels since, highlighting even more our lack of
-understanding of Antarctic sea ice variability and change
-\citep{espinosa2024}.
+experienced a slightly increasing trend up to 2015, which models
+systematically failed to reproduce and puzzled researchers
+\citep{hobbs2016}. Then, in 2016 Antarctic sea ice extent dropped
+precipitously \citep{turner2017a} and has been at low and record low
+levels since \citep{purich2023, hobs2024, gilbert2024, josey2024}, even
+more highlighting our lack of understanding of Antarctic sea ice
+variability and long-term change \citep{espinosa2024}.
 
-Antarctic sea ice is in a remote location but the potential impacts of
+Antarctic sea ice is in a remote location, but the potential impacts of
 its variability extends far into the lower latitudes, affecting the
 circulation of both the ocean and the atmosphere, ocean carbon uptake
-and biological processes. It's likely that a reduction in sea ice extent
-would reduce the meridional temperature gradient which in turn would
-reduce the strength and location of the eddy-driven jet (and vice versa
-for an increase in sea ice), which is a crucial component of weather
-events. In the northern hemisphere, it has been hypothesised that the
-reduction in Arctic sea ice has lead to a weaker, wavier jet, increasing
-the frequency of extreme events \citep{barnes2015}. There is some
-evidence of this effect in the Southern Hemisphere. Models show a
-weakened vortex in the middle atmosphere, increased vertical wave flux
-and greater zonal wave 1 amplitude in Springs following a low sea ice
-maximum. But models don't fully agree on the exact phase of the zonal
-wave response and observations don't show a clear signal
-\citep{rea2024}. The jet response to sea ice loss is also very model
-dependent, particularly sensitive to its the climatological location and
-relatively small compared with the direct effect of CO2 increase and Sea
-Surface Temperature (SST) warming \citep{ayres2019}. On climate scales,
-it is thought that the temperature signal of a reduction in Antarctic
-sea ice mirrors a ``mini global warming'' and experiments also predict a
-weakening and northerly shift of the eddy-driven
+and biological processes. It is likely that a reduction in sea ice
+extent would reduce the meridional temperature gradient which, in turn,
+would reduce the strength and location of the eddy-driven jet (and the
+opposite circulation changes with an increase in sea ice), which is a
+crucial control on extratropical weather. In the northern hemisphere
+(NH), it has been hypothesised that the reduction in Arctic sea ice has
+led to a weaker, wavier jet, increasing the frequency of extreme events
+\citep{barnes2015}. There is some evidence of this effect in the
+southern hemisphere (SH). Models show a weakened vortex in the middle
+atmosphere, increased vertical wave flux and greater zonal wave 1
+amplitude in austral spring following a low sea ice maximum. But models
+do not fully agree on the exact phase of the zonal wave response, and
+observations do not show a clear signal \citep{rea2024}. The jet
+response to sea ice loss is also very model dependent, particularly
+sensitive to its climatological location, and relatively small compared
+with the direct effect of CO2 increase and sea surface temperature (SST)
+warming \citep{ayres2019}. On climate scales, it is thought that the
+temperature signal of a reduction in Antarctic sea ice mirrors a ``mini
+global warming'', and experiments also predict a weakening and
+equatorward shift of the SH eddy-driven jet
 \citep{raphael2011, england2018, england2020, ayres2022}.
@@
 Correctly modelling Antarctic sea is not only necessary for process
-understanding and climate projections to inform adaptation strategies.
-Accurate seasonal to subseasonal forecast are crucial for operations in
-and around the Antarctic continent such as scientific missions,
-fisheries and tourism \citep[ \citet{wagner2020}]{desilva2020}. However,
-producing accurate Antarctic sea ice forecasts has been challenging due
-to model biases and inherent large variability and complexity and it has
-lagged behind Arctic sea ice forecasts \citep{zampieri2019}. Dynamical
-seasonal forecasts of summer Antarctic sea ice have been shown to
-perform worse than relatively simple statistical methods
-\citep{massonnet2023}, which also underscores the need for better
-understanding of sea ice dynamics.
+understanding and climate projections to inform adaptation strategies
+but also for accurate seasonal to sub-seasonal forecasts, which are
+crucial for operations in and around the Antarctic continent such as
+scientific missions, fisheries and tourism \citep[
+\citet{wagner2020}]{desilva2020}. However, producing accurate Antarctic
+sea ice forecasts has been challenging due to model biases, inherent
+large variability and complexity, and it has lagged behind Arctic sea
+ice forecasts \citep{zampieri2019}. Dynamical seasonal forecasts of
+summer Antarctic sea ice have been shown to perform worse than
+relatively simple statistical methods \citep{massonnet2023}, which also
+underscores the need for better understanding of sea ice dynamics.
 
-In this work we evaluate sea ice forecasts produced by ACCESS-S2, the
-Australian Bureau of Meteorology subseasonal-to-seasonal prediction
-system. This evaluation will inform future work with the model as a
-reseach tool and explore the potential of using it as a sea ice
-forecasting tool.
+In this work we evaluate sea ice forecasts produced by the Australian
+Community Climate and Earth System Simulator -- seasonal version2
+(ACCESS-S2), which is the Australian Bureau of Meteorology (BoM)'s
+current operational subseasonal-to-seasonal prediction system, and
+compare its forecast skill with that of the former system ACCESS-S1.
+This evaluation will inform future work with the prediction system as a
+research tool and explore the potential of using its sea-ice forecasts
+for decision-making.
 
-\hypertarget{data-and-methods}{%
-\section{Data and methods}\label{data-and-methods}}
+\section{Data and methods}\label{data-and-methods}
 
-\hypertarget{s2}{%
-\subsection{ACCESS-S2}\label{s2}}
+\subsection{ACCESS-S2}\label{s2}
 
-ACCESS-S2 \citep{wedd2022} is the Australian Bureau of Meteorology
-current subseasonal-to-seasonal prediction system which replaced
-ACCESS-S1 \citep{hudson2017} in October 2021. Both ACCESS-S2 and
-ACCESS-S1 consist on the Global Atmosphere 6.0 (GA6)
-\citep{williams2015, waters2017}, Global Land 6.0
+ACCESS-S2 \citep{wedd2022} became operational in October 2021 by
+replacing the ACCESS-S1 system \citep{hudson2017}.. The model components
+of both ACCESS-S2 and ACCESS-S1 consist of the Global Atmosphere 6.0
+(GA6) \citep{williams2015, waters2017}, Global Land 6.0
 \citep{best2011, waters2017}, Global Ocean 5.0
 \citep{madec2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
-\citet{rae2015}{]} model components. The atmosphere has a N216
-horizontal resolution (\textasciitilde60km in the mid-latitudes) with 85
-levels. The land model uses the same horizontal grid with 4 soil levels.
-The ocean component has a 1/4º resolution with 75 vertical levels. The
-ice component --based on CICE version 4.1-- has the same resolution than
-the ocean and 5 sea ice categories as well as an open water category.
+\citet{rae2015}{]}. The atmosphere has a N216 horizontal resolution
+(\textasciitilde60km in the mid-latitudes) with 85 vertical levels. The
+land model uses the same horizontal grid with 4 soil levels. The ocean
+component has a 1/4º resolution with 75 vertical levels. The sea ice
+component, based on CICE version 4.1, has the same resolution as the
+ocean and 5 sea ice categories as well as an open water category.
 
 Both systems take atmospheric initial conditions derived from
-ERA-interim \citep{dee2011} for the hindcast and from the Bureau's
+ERA-interim \citep{dee2011} for their hindcasts and from the Bureau's
 operational analysis for real-time forecasts. The main difference
-between the two are the ocean and ice initial conditions. ACCESS-S1
-initial conditions come from the Met Office FOAM system, which uses a
-multivariate, incremental three-dimensional variational (3D-Var),
+between the two are the ocean and sea ice initial conditions. ACCESS-S1
+ocean initial conditions come from the Met Office FOAM system, which
+uses a multivariate, incremental three-dimensional variational (3D-Var),
 first-guess-at-appropriate-time (FGAT) data assimilation scheme
@@
 surface height (SSH), in situ temperature and salinity profiles, and sea
-ice concentration. ACCESS-S2, instead, runs from initial conditions
-generated by the in-house data assimilation scheme described in
+ice concentration. ACCESS-S2, on the other hand, runs from initial
+conditions generated by the BoM data assimilation scheme described in
 \citet{wedd2022}. This scheme is a weakly coupled ensemble optimal
 interpolation method and assimilates temperature and salinity profiles
-from EN4 \citep{good2013} for the hindcast and from the WMO Global
-Telecommunication System (GTS) and both the Coriolis and USGODAE Global
-Data Assembly Centers (GDACs) for the real-time forecast. SSTs are
-nudged to Reynolds OISSTv2.1 \citep{reynolds2007} for the hindcast and
-to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and
-Beggs 2008) for the real-time forecast in areas where SSTs are over 0ºC.
-Relevant for this work, sea ice concentrations are not assimilated.
+from EN4 \citep{good2013} for the hindcasts and from the WMO Global
+Telecommunication System (GTS) and the Coriolis and USGODAE Global Data
+Assembly Centers (GDACs) for the real-time forecasts. SSTs are nudged to
+Reynolds OISSTv2.1 \citep{reynolds2007} for the hindcasts and to the
+Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs
+2008) for the real-time forecasts in areas where SSTs are over 0ºC.
+Relevant for this work, sea ice concentrations are not assimilated in
+ACCESS-S2.
 
@@
 
-\hypertarget{verification-datasets}{%
-\subsection{Verification datasets}\label{verification-datasets}}
+\subsection{Verification datasets}\label{verification-datasets}
 
@@
 covered with ice. Different datasets derived using different algorithms
-provide different estimates, each with its own biases and uncertainties.
-\citet{meier2019} estimated that the inter-product uncertainty of sea
-ice extent is of the order of 1 million km\textsuperscript{2}. As will
-be shown below, this spread is minimal compared with the errors in the
-ACCESS-S2 and ACCESS-S1 forecasts, so the conclusions are independent of
-the dataset used.
+provide different estimates, each with their own biases and
+uncertainties. \citet{meier2019} estimated that the inter-product
+uncertainty of sea ice extent is of the order of 1 million \(km^2\). As
+will be shown below, this spread is minimal compared with the typical
+errors in the ACCESS-S2 and ACCESS-S1 forecasts, so the conclusions are
+independent of the dataset used.
 
@@
 on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and
--F13 satellites. The data has a resolution of 25 by 25km and daily from
+-F13 satellites. The data has a resolution of 25 by 25 km and daily from
 1978 onwards.
@@
 (OSI SAF) \citep{OSISAF} is another satellite-derived sea ice
-concentration product. It is based on mostly the same sensors as the NOA
-CDR but computed independently using different algorithms. Figures
+concentration product. It is based on mostly the same sensors as the
+NOAA CDR but computed independently using different algorithms. Figures
 prepared with this dataset are provided in the supplementary material.
 
-\hypertarget{error-measures}{%
-\subsection{Error measures}\label{error-measures}}
+\subsection{Error measures}\label{error-measures}
 
@@
 
-Sea Ice Extent is defined as the area of ocean covered with at least
+Sea-ice extent is defined as the area of ocean covered with at least
 15\% ice. This threshold is motivated by the limitations in satellite
-retrieval, which is increasingly unreliable for lower sea ice
-conditions.
+retrieval, which is increasingly unreliable for lower sea ice conditions
+(cite -- probably something from NSIDC).
 
-Pan-Antarctic Sea Ice Extent serves as a rough global measure of the
-amount of sea ice but it doesn't take into account the spatial
-distribution. A model could have relatively accurate extent of ice but
-with different distributions. To account for location errors, we
-computed the Root Mean Squared Error (RMSE) of sea ice concentration
-anomalies and the Integrated Ice Edge Error {[}IIEE;
+Pan-Antarctic (net) sea ice extent serves as a rough hemispheric measure
+of the amount of sea ice, but it does not take into account the spatial
+distribution. A model could have relatively accurate extent of the net
+ice but with different regional distributions. To account for location
+errors, we computed the Root Mean Squared Error (RMSE) of sea ice
+concentration anomalies and the Integrated Ice Edge Error {[}IIEE;
 \citet{goessling2016}{]}.
@@
 between forecasted and observed sea ice concentration anomalies. We
-compute a pan-Antarctic RMSE by averaging over the whole domain, and
-also a zonally-varying RMSE by averaging over 24 15° longitude slices
-around Antarctica.
+compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC
+CDRV4 southern hemisphere domain, and also a zonally-varying RMSE by
+averaging over twenty-four 15° longitude slices around Antarctica.
 
-The IIEE is defined as the area in which the model misspredicts sea ice
-concentration being above or belong 15\% ice. That is, dichotomise sea
+The IIEE is defined as the area where the model miss-predicts sea ice
+concentration being above or below 15\% ice. That is, dichotomise sea
 ice concentration into areas with more and less than 15\% sea ice both
-in the forecast and observations; the IIEE is the area in which forecast
-and observations differ.
+in the forecast and observations.
 
@@
 
-\hypertarget{results-and-discussion}{%
-\section{Results and discussion}\label{results-and-discussion}}
+\section{Results and discussion}\label{results-and-discussion}
 
-\hypertarget{reanalysis}{%
-\subsection{Reanalysis}\label{reanalysis}}
-
-\hypertarget{bias}{%
-\subsubsection{Bias}\label{bias}}
+\subsection{Bias}\label{bias}
 
@@
 
-{\centering \includegraphics{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}}
 
@@
 
-\end{figure}
+\end{figure}%
 
@@
 column) and median difference with respect of NOAA/NSIDC CDRV4 for the
-ACCESS-S2 and ACCESS-S1 hindcasts. Median extent of initial conditions
-at the first of every month are indicated with circles, while median
-extent at the same date but forecasted with the largest lead time
-possible for each model (between 274 and 277 days for ACCESS-S2 and
-between 213 and 216 days for ACCESS-S1). At this large lead time, the
-the information of the initial conditions is essentially lost and the
-forecast reverts to each model's preferred equilibrium state.
+ACCESS-S2 and ACCESS-S1 hindcasts (right column). Median extent of
+initial conditions at the first of every month are indicated with
+circles, while median extent at the same date but forecasted with the
+largest lead time possible for each model (between 274 and 277 days for
+ACCESS-S2 and between 213 and 216 days for ACCESS-S1). At this large
+lead time, the the information of the initial conditions is essentially
+lost and the forecast reverts to each model's preferred equilibrium
+state.
 
-Both models' equilibrium state (triangles) show low bias of sea ice
-extent, particularly in the late-autumn and winter months. This is due
-primarily to faster melt during a longer melt season between January and
-March and slower growth during March and April. This is then partially
-balanced with faster growth between May and July
+Both models' equilibrium state (triangles) show a negative bias of sea
+ice extent, particularly in the late-autumn and winter months. This is
+due primarily to faster melt during a longer melt season between January
+and March and slower growth during March and April. This is then
+partially balanced with faster growth between May and July
 (Fig.~\ref{fig-mean-growth}). Many sea ice models exhibit this
@@
 is not surprising that both forecasting systems converge to a similar
-climatology because they share the same model configuration.
+climatology because they share the same model formulation.
 
-ACCESS-S2 initial conditions (circles) also have a low extent bias,
+ACCESS-S2 initial conditions (circles) also have a negative bias,
 especially in the late summer-early autumn, while ACCESS-S1 initial
 conditions are very close to observations. The faster growth between May
-and July reduces the low bias in ACCESS-S2 for forecasts initialised
-from April to June, but it adds a high bias in ACCESS-S1.
+and July reduces the negative bias in ACCESS-S2 for forecasts
+initialised from April to June, but it adds a positive bias in
+ACCESS-S1.
 
@@
 difference between circles and triangles in ACCESS-S2, indicating that
-almost no data assimilation is taking place and the sea ice component of
-the model is virtually free-running.
+the ocean and atmosphere data assimilation is not affecting seas ice and
+that this component of the model is virtually free-running.
 
@@
 
-{\centering \includegraphics{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}}
 
@@
 ACCESS-S1 and ACCESS-S2 hindcasts and observations. Values are smoothed
-with a 2-degree loess smooth with a 30 day window.}
+with a 11-day running mean.}
 
-\end{figure}
+\end{figure}%
 
@@
 
-{\centering \includegraphics{access-ice_files/figure-pdf/fig-bias-1.pdf}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1.pdf}}
 
@@
 shortest lag. From October to May, the model underestimates sea ice
-concentrations pretty much everywhere there is ice except for the deep
-Weddell Sea in April and May, where sea ice concentrations saturate to 1
-both in the observations and forecasts. In winter, the differences are
-mostly on the sea ice edge, with light positive bias in the African
-sector of East Antarctica and negative bias around the Indian Ocean
-sector which partially compensate, resulting in the near-zero extent
-bias seen in those months (Fig.~\ref{fig-hindcast-extent}).
+concentrations in most regions except for the inner Weddell Sea in April
+and May, where sea ice concentrations saturate to 1 both in the
+observations and forecasts. In winter, the differences are mostly on the
+sea ice edge with light positive bias in the African sector of East
+Antarctica and negative bias around the Indian Ocean sector which
+partially compensate, resulting in the near-zero extent bias seen in
+those months (Fig.~\ref{fig-hindcast-extent}).
 
-\hypertarget{anomalies}{%
-\subsubsection{Anomalies}\label{anomalies}}
+\subsection{RMSE}\label{rmse}
 
@@
 
-{\centering \includegraphics{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}}
 
@@
 
-\caption{\label{fig-extent-anom}Sea ice extent anomalies for ACCESS-S1
-and ACCESS-S2 (black) and NOAA/NSIDC CDRV4 (blue).}
+\caption{\label{fig-extent-anom}Monthly mean sea ice extent anomalies
+for ACCESS-S1 and ACCESS-S2 (black) and NOAA/NSIDC CDRV4 (blue)
+forecasted at selected lead times. The RMSE during the overlapping
+period (1990--2013) is shown on the top left.}
 
-\end{figure}
+\end{figure}%
 
 Figure~\ref{fig-extent-anom} shows monthly sea ice extent anomalies
-forecasted at selected lags. The anomalies in this case were computed
-with respect of the climatology of each lag, which works as a
-bias-correction method. Compared with ACCESS-S1, ACCESS-S2 anomaly
-forecast is relatively poor even in the first month, which stays
+forecasted at selected lead times. Compared with ACCESS-S1, ACCESS-S2
+anomaly forecast is relatively poor even in the first month, which stays
 relatively skilful even at lag 3. ACCESS-S2 shows much bigger
@@
 
-\hypertarget{rmse}{%
-\subsubsection{RMSE}\label{rmse}}
-
-To study ACCESS-S2 forecasts quantitatively, we compute error measures
+To assess ACCESS-S2 forecasts quantitatively, we compute error measures
 for all hindcasts started on the 1st of every month.
@@
 
-{\centering \includegraphics{access-ice_files/figure-pdf/fig-rmse-1.pdf}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse-1.pdf}}
 
@@
 
-Figure~\ref{fig-rmse} shows the median and 95\% range of RMSE of sea ice
-concentration anomalies for ACCESS-S2 forecasts compared with a
-benchmark of persistence. Due to errors in the initial conditions, it is
-expected that a persistence forecast would be better than the model
-forecast at very short lags, but that the persistence forecast errors
-would grow faster and eventually surpass the mode forecast, after when
-it is statistically useful. Here the persistence errors are almost
-always lower than the ACCESS-S2 forecast, indicating that the model
-doesn't have skill at any lag and in any month. The only exception is
-around February, where the model has lower RMSE than the persistence
-forecast at virtually every lag.
+Figure~\ref{fig-rmse} shows the median RMSE of sea ice concentration
+anomalies for ACCESS-S2 and ACCESS-S1 hindcasts compared with a
+benchmark of persistence and climatology. Due to errors in the initial
+conditions, it is expected that a persistence forecast would be better
+than the model forecast at very short lead times, but that the
+persistence forecast errors would grow faster and eventually surpass the
+model forecast, after when the model is statistically useful. Here the
+persistence errors are almost always lower than the ACCESS-S2 forecast,
+indicating that the model does not have skill at any lead time and in
+any month. The only exception is around February, where the model has
+lower RMSE than the persistence forecast at virtually every lag.
 
@@
 
-{\centering \includegraphics{access-ice_files/figure-pdf/fig-rmse_lon-1.pdf}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1.pdf}}
 
@@
 
-\caption{\label{fig-rmse_lon}Meadian difference between RMSE of
-ACCESS-S2 forecasts and persistence forecast at longitudinal bands.}
+\caption{\label{fig-rmse_lon}Median difference between RMSE of ACCESS-S2
+forecasts and persistence forecast computed on
+\texttt{r\ uniqueN(rmse\_lon\_mean\$lon)} meridional slices
+\texttt{r\ diff(unique(rmse\_lon\_mean\$lon)){[}1{]}}º wide.}
 
-\end{figure}
+\end{figure}%
 
@@
 
-Forecasts initialised on January 1\textsuperscript{st}
-(Fig.~\ref{fig-rmse_lon} panel a) show RMSE values below persistence at
-lags larger than 180 days (corresponding to July through September) in
-the Ross and Weddell Sea. These two regions are also relatively well
-forecasted from other months. Notably, May forecasts of Weddell sea ice
-concentration anomalies (Fig.~\ref{fig-rmse_lon} panel e) are skilful
-after about 20 days and remain so until December. The Weddell Sea is
-also the region with the maximum error --around June regardless of
-initialisation date.
+Forecasts initialised on January 1st (Fig.~\ref{fig-rmse_lon} panel a)
+show RMSE values below persistence at lags larger than 180 days
+(corresponding to July through September) in the Ross and Weddell Sea.
+These two regions are also relatively well forecasted from other months.
+Notably, May forecasts of Weddell sea ice concentration anomalies
+(Fig.~\ref{fig-rmse_lon} panel e) are skilful after about 20 days and
+remain so until December. The Weddell Sea is also the region with the
+maximum error --around June regardless of initialisation date.
 
-\hypertarget{comparison-with-s1}{%
-\subsubsection{Comparison with S1}\label{comparison-with-s1}}
+\subsubsection{Comparison with S1}\label{comparison-with-s1}
 
@@
 
-{\centering \includegraphics{access-ice_files/figure-pdf/fig-iiee-1.pdf}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-iiee-1.pdf}}
 
@@
 
-\end{figure}
+\end{figure}%
 
 To compare ACCESS-S2 with ACCESS-S1, we computed the IIEE for both
-models. This error measure is shown in Figure~\ref{fig-iiee} for all
-lags and forecasts initialised at the first of every month. ACCESS-S1
-has lower error at short lags at all months, with the errors converging
-to ACCESS-S2 as the forecast goes on. The time to convergence depends on
-the month and it can be as short as two weeks June and July to as long
-as 160 days for forecasts initialised in February. Since the only
-difference between these forecasts are the initial conditions, this
-timescale is an indication of the the memory of sea ice to initial
-conditions --at least from October to March when the data assimilated
-form the other components has little to no influence on sea ice.
+models. This error measure is shown in Figure~\ref{fig-iiee} for
+forecasts initialised at the first of every month at all lead times.
+ACCESS-S1 has lower error at lead times up to 60 days at all months,
+with the errors converging to ACCESS-S2 as the forecast goes on. The
+time to convergence depends on the month and it can be as short as two
+weeks June and July to as long as 160 days for forecasts initialised in
+February. Since the only difference between these forecasts are the
+initial conditions, this timescale is an indication of the memory of sea
+ice to initial conditions --at least from October to March when the data
+assimilated form the other components has little to no influence on sea
+ice.
 
@@
 spread at short lags between ACCESS-S2 and ACCESS-S1. In all months
-ACCESS-S1 has much narrower error spread at lag 1. The error in the
-initial conditions is not only small, but also fairly constant. This
-spread then grows towards a climatological spread as errors accumulate
-differently in different forecasts. For ACCESS-S2, this is true only
-only between July and October, approximately. For all other months, the
-error spread is more or less stable throughout the forecast window,
-indicating that not only the initial error is high, but it's not
-constant.
+ACCESS-S1 has much narrower error spread at the first day of the
+forecasts. The error in the initial conditions is not only small, but
+also fairly constant. This spread then grows towards a climatological
+spread as errors accumulate differently in different forecasts. For
+ACCESS-S2, this is true only between July and October, approximately.
+For all other months, the error spread is more or less stable throughout
+the forecast window, indicating that not only the initial error is high,
+but it is not constant.
 
@@
 
-{\centering \includegraphics{access-ice_files/figure-pdf/fig-iiee-variance-1.pdf}
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-iiee-variance-1.pdf}}
 
@@
 ensemble members or to large spread of individual forecasts.
-Figure~\ref{fig-iiee-variance} splits the IIEE variance for each lag
-into the mean variance of each individual forecast and the variance of
-the mean error of each individual forecast, which adds up to the total
-variance. The average variance of each forecast is almost identical
-between forecast systems in all months. This shows that the ensemble
-spread of individual forecasts evolves identically, which, again, it's
-not unexpected because both systems share the same model formulation.
-This also shows that the perturbation scheme in ACCESS-S2 is comparable
-to the one in ACCESS-S1.
+Figure~\ref{fig-iiee-variance} splits the IIEE variance for each lead
+time into the mean variance of each individual forecast and the variance
+of the mean error of each individual forecast, which adds up to the
+total variance. The average variance of each forecast is almost
+identical between the two forecast systems in all months. This shows
+that the ensemble spread of individual forecasts evolves identically,
+which, again, is not unexpected because both systems share the same
+model formulation. This also shows that the perturbation scheme in
+ACCESS-S2 is comparable to the one in ACCESS-S1.
 
@@
 ACCESS-S2 than ACCESS-S1. The difference is particularly large at short
-lags in some months, which coindice with the ones in which the data
-assimilation scheme is not influencing sea ice initial conditions.
+lead times in some months, which coincide with the ones in which the
+data assimilation scheme is not influencing sea ice initial conditions.
 
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
+\subsection{Conclusions}\label{conclusions}
 
@@
 does not suffer from this bias. This is due to the assimilation of sea
-ice concentration anomalies into the initial conditions, which
-successful corrects for the negative bias in the model. The analysis
-suggests that the data assimilation system in ACCESS-S2 is only
-effectively influencing sea ice initial conditions from June to October,
-while the rest of the year, the sea ice component runs virtually free,
-reverting to its biased equilibrium state.
+ice concentrations into the initial conditions, which successful
+corrects for the negative bias in the model. Our analysis suggests that
+the data assimilation system in ACCESS-S2 is only effectively
+influencing sea ice initial conditions from June to October, while the
+rest of the year, the sea ice component runs virtually free, reverting
+to its biased equilibrium state.
 
@@
 determining sea ice evolution at least up to late June. Arctic sea ice
-forecast also show greater sensitivity to initial conditions in (boreal)
-summer compared with (boreal) winter \citep{day2014, bunzel2016}, so
+forecast also show greater sensitivity to initial conditions in boreal
+summer compared with boreal winter \citep{day2014, bunzel2016}, so
 similar mechanism might be playing a role.
 
-Although forecasts are not skilful for forecasting panantarctic sea ice
+Although forecasts are not skilful for forecasting pan-Antarctic sea ice
 concentrations, there are some areas where the model does show skill.
@@
 predictability thanks to persistent and eastwardly advected upper ocean
-heat content anomalies \citep{bushuk2021} Sea surface temperatures are
+heat content anomalies \citep{bushuk2021}. Sea surface temperatures are
 assimilated by ACCESS-S2 only in areas with temperatures greater than
 0°C and errors don't show a clear easterly-propagating signal, so it is
-not clear if ACCESS-S2 is leveraging the same source of predictability.
+not clear if ACCESS-S2 is leveraging the same source of predictability..
 
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\section{References}\label{references}
 
@@
 
-\hypertarget{suplementary-figures}{%
-\section{Suplementary figures}\label{suplementary-figures}}
+\section{Suplementary figures}\label{suplementary-figures}
 
@@
 \bibliographystyle{copernicus}
+
 \end{document}

### 2025-04-18 R/ggplot.R

@@ -25,14 +25,18 @@ ggplot2::theme_set(ggplot2::theme_minimal() +
                      ))
 wide_legend <- ggplot2::theme(legend.key.width = grid::unit(1, 'null'))
 
+trans_pink <- "#D07C8E"
+trans_blue <- "#0AA3CE"
+
+
 
 colours_models <- c(
-  S2 = "black",
-  S1 = "#a51d2d",
-  cdr = "#1a5fb4",
-  bt = "#1a5fb4",
-  era5 = "#1a5fb4",
-  osi = "#1a5fb4",
+  S1 = "#ACA40A",
+  S2 = "#9C59D1",
+  cdr = "black",
+  bt = "black",
+  era5 = "black",
+  osi = "black",
   persistence = "black",
   climatology = "gray50"
 )

### 2025-05-07 analysis/paper/access-ice.qmd
Revision

@@
   Please add `\clearpage` between each table and/or figure.
-format: copernicus-pdf
+format: docx
 
@@
   message: false
-  cache: refresh
+  cache: true
 filters:
@@
 
-Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016].
-Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change [@espinosa2024].
+Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend up to 2015, which models systematically failed to reproduce and puzzled researchers [@hobbs2016].
+Then, in 2016 Antarctic sea ice extent dropped precipitously [@turner2017a] and has been at low and record low levels since [@purich2023; @hobs2024; @gilbert2024; @josey2024], even more highlighting our lack of understanding of Antarctic sea ice variability and long-term change [@espinosa2024].
 
-Antarctic sea ice is in a remote location but the potential impacts of its variability extends far into the lower latitudes, affecting the circulation of both the ocean and the atmosphere, ocean carbon uptake and biological processes.
-It's likely that a reduction in sea ice extent would reduce the meridional temperature gradient which in turn would reduce the strength and location of the eddy-driven jet (and vice versa for an increase in sea ice), which is a crucial component of weather events.
-In the northern hemisphere, it has been hypothesised that the reduction in Arctic sea ice has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015].
+Antarctic sea ice covers the polar ocean, yet the potential impacts of its variability extend far into the mid and lower latitudes, affecting the circulation of both the ocean and the atmosphere, ocean carbon uptake and biological processes [@fogwill2020, @england2020, @swadling2023].
+In coupled climate model experiments, a reduction in sea ice extent has been shown to reduce the meridional temperature gradient, in turn reducing the eddy-driven jet strength and shifting it equatorward [@england2018 @england2020; @ayres2019; @ayres2022] and vice versa for an increase in sea ice [@raphael2003a; @raphael2011], which is a crucial control on extratropical weather.
+In the Northern Hemisphere, it has been hypothesised that the reduction in Arctic sea ice has led to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015; @smith2022b]. 
 There is some evidence of this effect in the Southern Hemisphere.
-Models show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in Springs following a low sea ice maximum.
-But models don't fully agree on the exact phase of the zonal wave response and observations don't show a clear signal [@rea2024].
-The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location and relatively small compared with the direct effect of CO2 increase and Sea Surface Temperature (SST) warming [@ayres2019].
-On climate scales, it is thought that the temperature signal of a reduction in Antarctic sea ice mirrors a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].
+For example, model simulations show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in austral spring following lower than average sea ice extent at in early winter [@rea2024, @song2025].
+But models different models disagree on the exact phase of the zonal wave response, and observations do not show a clear signal [@rea2024].
+The magnitude of the eddy-driven jet response  to sea ice loss has been shown to be model dependent, and is sensitive to its climatological location [@ayres2019].
+Further, sea ice induced changes in the jet have been found to be  relatively small compared with the jet-response to the direct effect of CO2 increase and global sea surface temperature (SST) warming [@ayres2022].
 
-Correctly modelling Antarctic sea  is not only necessary for process understanding and climate projections to inform adaptation strategies.
-Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent such as scientific missions, fisheries and tourism [@desilva2020 @wagner2020].
-However, producing accurate Antarctic sea ice forecasts has been challenging due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
-Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.
+Correctly modelling Antarctic sea is not only necessary for process understanding and climate projections to inform adaptation strategies but also for accurate seasonal to sub-seasonal forecasts, which are crucial for operations in and around the Antarctic continent such as scientific missions, fisheries and tourism [@desilva2020 @wagner2020].
+However, producing accurate Antarctic sea ice forecasts has been challenging due to model biases, inherent large variability and complexity, and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
+Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics, and drivers.
 
-In this work we evaluate sea ice forecasts produced by +S2, the Australian Bureau of Meteorology subseasonal-to-seasonal prediction system. 
-This evaluation will inform future work with the model as a reseach tool and explore the potential of using it as a sea ice forecasting tool.
+In this work we evaluate sea ice forecasts produced by the Australian Community Climate and Earth System Simulator -- seasonal version2 (+S2), which is the Australian Bureau of Meteorology (BoM)'s current operational subseasonal-to-seasonal prediction system, and compare its forecast skill with that of the former system +S1.
+This evaluation will inform future work with the prediction system as a research tool and explore the potential of using its sea-ice forecasts for decision-making.
 
@@
 
-+S2 [@wedd2022] is the Australian Bureau of Meteorology current subseasonal-to-seasonal prediction system which replaced +S1 [@hudson2017] in October 2021.
-Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015] model components.
-The atmosphere has a N216 horizontal resolution (\~60km in the mid-latitudes) with 85 levels.
++S2 [@wedd2022] became operational in October 2021 by replacing the +S1 system [@hudson2017]..
+The model components of both +S2 and +S1 consist of the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
+The atmosphere has a N216 horizontal resolution (\~60km in the mid-latitudes) with 85 vertical levels.
 The land model uses the same horizontal grid with 4 soil levels.
 The ocean component has a 1/4º resolution with 75 vertical levels.
-The ice component --based on CICE version 4.1-- has the same resolution than the ocean and 5 sea ice categories as well as an open water category.
+The sea ice component, based on CICE version 4.1, has the same resolution as the ocean and 5 sea ice categories as well as an open water category.
 
-Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts.
-The main difference between the two are the ocean and ice initial conditions.
-+S1 initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
-+S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022.
-This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
-SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are over 0ºC.
-Relevant for this work, sea ice concentrations are not assimilated.
+Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for their hindcasts and from the Bureau's operational analysis for real-time forecasts.
+The main difference between the two are the ocean and sea ice initial conditions.
++S1 ocean initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
++S2, on the other hand, runs from initial conditions generated by the BoM data assimilation scheme described in @wedd2022.
+This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcasts and from the WMO Global Telecommunication System (GTS) and the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecasts.
+SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcasts and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecasts in areas where SSTs are over 0ºC.
+Relevant for this work, sea ice concentrations are not assimilated in +S2.
 
@@
 Each forecast consists of 9 ensemble members built from three consecutive 3-member forecasts initialised at the first of every month and the two previous days and run from 279 days.
-The +S1 hindcast is built in the same manner and running for the period 1990--2012.
+The +S1 hindcast is built in the same manner (albeit with 217-day runs) for the period 1990--2012.
 
@@
 For verification we use satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
-Different datasets derived using different algorithms provide different estimates, each with its own biases and uncertainties. 
-@meier2019 estimated that the inter-product uncertainty of sea ice extent is of the order of 1 million km^2^. 
-As will be shown below, this spread is minimal compared with the errors in the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
+Different datasets derived using different algorithms provide different estimates, each with their own biases and uncertainties.
+@meier2019 estimated that the inter-product uncertainty of sea ice extent is of the order of 1 million $km^2$.
+As will be shown below, this spread is minimal compared with the typical errors in the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
 
@@
 Both source algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
-The data has a resolution of 25 by 25km and daily from 1978 onwards.
+The data has a resolution of 25 by 25 km and daily from 1978 onwards.
 
 The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [@OSISAF] is another satellite-derived sea ice concentration product.
-It is based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.
-Figures prepared with this dataset are provided in the supplementary material.
+It is based on mostly the same sensors as the NOAA CDR but computed independently using different algorithms.
+Figures prepared with this dataset are provided in the supplementary material and don't differ significantly from the ones prepared using CDR.
 
@@
 
-Sea Ice Extent is defined as the area of ocean covered with at least 15% ice.
-This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea ice conditions.
+Sea-ice extent is defined as the area of ocean covered with at least 15% ice.
+This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea ice conditions (cite -- probably something from NSIDC).
 
-Pan-Antarctic Sea Ice Extent serves as a rough global measure of the amount of sea ice but it doesn't take into account the spatial distribution. 
-A model could have relatively accurate extent of ice but with different distributions.
+Pan-Antarctic (net) sea ice extent serves as a rough hemispheric measure of the amount of sea ice, but it does not take into account the spatial distribution.
+A model could have relatively accurate extent of the net ice but with different regional distributions.
 To account for location errors, we computed the Root Mean Squared Error (RMSE) of sea ice concentration anomalies and the Integrated Ice Edge Error [IIEE; @goessling2016].
@@
 We compute RMSE as the square root of the average squared differences between forecasted and observed sea ice concentration anomalies.
-We compute a pan-Antarctic RMSE by averaging over the whole domain, and also a zonally-varying RMSE by averaging over 24 15° longitude slices around Antarctica.
+We compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC CDRV4 southern hemisphere domain, and also a zonally-varying RMSE computed over twenty-four 15° longitude slices around Antarctica.
 
-The IIEE is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
-That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.
+The IIEE is defined as the area where the model miss-predicts sea ice concentration being above or below 15% ice.
+That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations.
 
-All error measures were computed on the +CDR domain grid and projection to which model output was bilinearly interpolated.
+All error measures were computed on the NOAA/NSIDC CDRV4 domain grid and projection to which model output was bilinearly interpolated.
 
@@
 
-## Reanalysis
-
-### Bias
+## Bias
 
@@
 ```{r fig-hindcast-extent}
-#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month with the median sea ice extent of +CDR in the left column and the median difference in the right column.
+#| fig-cap: Pan-antarctic median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month with the median sea ice extent of +CDR in the left column and the median difference in the right column.
 The +CDR climatology is computed in the period corresponding to each hindcast.
@@
 
-@fig-hindcast-extent a shows the median sea ice extent (left column) and median difference with respect of +CDR for the +S2 and +S1 hindcasts.
+@fig-hindcast-extent a shows the median sea ice extent (left column) and median difference with respect of +CDR for the +S2 and +S1 hindcasts (right column).
 Median extent of initial conditions at the first of every month are indicated with circles, while median extent at the same date but forecasted with the largest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
@@
 
-Both models' equilibrium state (triangles) show low bias of sea ice extent, particularly in the late-autumn and winter months.
++S2 initial conditions (circles) show an overall negative bias, especially in the late summer-early autumn, while +S1 initial conditions are very close to observations.
+Both models' equilibrium state (triangles) show a negative bias of sea ice extent, particularly in the late-autumn and winter months.
 This is due primarily to faster melt during a longer melt season between January and March and slower growth during March and April.
@@
 Many sea ice models exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
-It is not surprising that both forecasting systems converge to a similar climatology because they share the same model configuration.
-
-+S2 initial conditions (circles) also have a low extent bias, especially in the late summer-early autumn, while +S1 initial conditions are very close to observations. 
-The faster growth between May and July reduces the low bias in +S2 for forecasts initialised from April to June, but it adds a high bias in +S1.
+It is not surprising that both forecasting systems converge to a similar climatology because they share the same model formulation.
 
@@
 From June to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
-The rest of the year there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.
+The rest of the year there is little if any difference between circles and triangles in +S2, indicating that the ocean and atmosphere data assimilation is not affecting seas ice and that this component of the model is virtually free-running.
 
@@
 ```{r fig-mean-growth}
-#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations. Values are smoothed with a 2-degree loess smooth with a 30 day window.
+#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations computed as the median daily differences in sea ice extent for each date and all lead times.
+Values are smoothed with a 11-day running mean.
 
-n = 366*3
+n <- 366*3
+w <- 11
 dif_extent |> 
+  copy() |> 
+  _[, mid := frollmean(mid, n = w, align = "center"), by = dataset] |> 
   ggplot(aes(time, mid)) +
   geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
-  geom_smooth(aes(group = dataset, color = dataset), 
-              method = "loess",
-              span = 30/n, linewidth = 0.5, se = FALSE, n = n)  +
+  geom_line(aes(color = dataset)) +
+  # geom_smooth(aes(group = dataset, color = dataset), 
+  #             method = "loess",
+  #             span = w/n, linewidth = 0.5, se = FALSE, n = n)  +
   # geom_line() +
-  geom_smooth(data = dif_hindcast,
-              method = "loess",
-              aes(color = model), span = 30/n, linewidth = 0.5, se = FALSE, n = n) +
+
+  geom_line(data = dif_hindcast |> 
+              copy() |> 
+              _[, mid := frollmean(mid, n = w, align = "center"), by = model],
+            aes(color = model)) +
+  # geom_smooth(data = dif_hindcast,
+  #             method = "loess",
+  #             aes(color = model), span = w/n, linewidth = 0.5, se = FALSE, n = n) +
   scale_x_datetime(NULL, date_breaks = "1 month", 
@@
 #| fig-height: 6
-#| fig-cap: Mean sea ice concentrations 1-month lag monthly mean +S2 forecast bias compared with NSIDC.
+#| fig-cap: Mean difffernece between monthly sea ice concentrations at 0-month lead time +S2 forecast and observations.
 
@@
                                           labels_models[which_dataset]),
-                                   low = scales::muted("red"),
-                                   high = scales::muted("blue")) +
+                                   low = trans_pink,
+                                   high = trans_blue) +
   # geom_polygon(data = ice_shelves_nsicd, aes(X, Y, group = group), 
@@
   geom_antarctica_fill +
-  facet_wrap(~ month) +
+  facet_wrap(~ month, labeller = labeller(month = labels_month)) +
   wide_legend +
@@
 
-@fig-bias shows the difference of monthly mean sea ice concentrations between +CDR and +S2 hindcasts at the shortest lag. 
-From October to May, the model underestimates sea ice concentrations pretty much everywhere there is ice except for the deep Weddell Sea in April and May, where sea ice concentrations saturate to 1 both in the observations and forecasts.
-In winter, the differences are mostly on the sea ice edge, with light positive bias in the African sector of East Antarctica and negative bias around the Indian Ocean sector which partially compensate, resulting in the near-zero extent bias seen in those months ([Fig.
+@fig-bias shows the difference of monthly mean sea ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest lead time
+From October to May, the model underestimates sea ice concentrations in most regions except for the inner Weddell Sea in April and May, where sea ice concentrations saturate to 1 both in the observations and forecasts.
+In winter, the differences are mostly on the sea ice edge with light positive bias in the African sector of East Antarctica and negative bias around the Indian Ocean sector which partially compensate, resulting in the near-zero extent bias seen in those months ([Fig.
 @fig-hindcast-extent]).
 
-### Anomalies
+## RMSE
 
 ```{r fig-extent-anom}
-#| fig-cap: Sea ice extent anomalies for +S1 and +S2 (black) and +CDR (blue).
+#| fig-cap: Monthly mean sea ice extent anomalies for +S1 and +S2 and observations (black) forecasted at selected lead times.
+The RMSE during the overlapping period (1990--2013) is shown on the top left.
 
@@
 
-@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lags.
-The anomalies in this case were computed with respect of the climatology of each lag, which works as a bias-correction method. 
-Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at lag 3.
+@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lead times.
+Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at a lead time of 3 months.
 +S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.
 
-### RMSE
-
-To study +S2 forecasts quantitatively, we compute error measures for all hindcasts started on the 1st of every month.
+To assess +S2 forecasts quantitatively, we compute error measures for all hindcasts started on the 1st of every month.
 
@@
 ```{r fig-rmse}
-#| fig-cap: Median RMSE of sea ice concentration anomalies as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies.
+#| fig-cap: Median RMSE of sea ice concentration anomalies as a function of forecast lead time for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies (black) and climatology (gray).
 #| fig.height: 5
@@
   labs(y = "RMSE", 
-       x = "Lag") +
+       x = "Lead time (days)") +
   scale_color_models +
@@
 
-::: notes
-
-monthly rmse/corelation at gridpoints per lead time.
-
-:::
-
-@fig-rmse shows the median and 95% range of RMSE of sea ice concentration anomalies for +S2 forecasts compared with a benchmark of persistence.
-Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, after when it is statistically useful.
-Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
-The only exception is around February, where the model has lower RMSE than the persistence forecast at virtually every lag.
+@fig-rmse shows the median RMSE of sea ice concentration anomalies for +S2 and +S1 hindcasts compared with a benchmark of persistence and climatology.
+Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lead times, but that the persistence forecast errors would grow faster and eventually surpass the model forecast, after when the model is statistically useful.
+Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model does not have skill at any lead time and in any month.
+The only exception is around February, where the model has lower RMSE than the persistence forecast at virtually every lead time.
 
@@
 ```{r fig-rmse_lon, fig.height=7, fig.width=10}
-#| fig.cap: Meadian difference between RMSE of +S2 forecasts and persistence forecast at longitudinal bands.
+#| fig.cap: Median difference between RMSE of +S2 forecasts and persistence forecast computed on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]`º wide.
+Antarctica's coastline is shown at the bottom of each panel for reference.
 
@@
             hjust = 1) +
-  scale_fill_divergent_discretised() +
+  scale_fill_divergent_discretised(low = trans_blue, high = trans_pink) +
   scale_x_longitude(ticks = 60, expand = c(0.08, 0)) +
@@
   labs(fill = NULL, 
-       y = "Lag") +
+       y = "Lead time (days)") +
   wide_legend +
@@
 
-The skill shown by +S2 at February-March forecasts is evidenced by a band of negative values across almost all longitudes. 
+The skill shown by +S2 at February-March forecasts is observed
+
+by a band of negative values across almost all longitudes.
 The only region where February-March sea ice is not well forecasted is east of the Antarctic Peninsula, which is the only region with consistent summer sea ice.
@@
 
-Forecasts initialised on January 1^st^ ([Fig.
-@fig-rmse_lon] panel a) show RMSE values below persistence at lags larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea.
+Forecasts initialised on January 1st ([Fig.
+@fig-rmse_lon] panel a) show RMSE values below persistence at lead times larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea.
 These two regions are also relatively well forecasted from other months.
@@
 ```{r fig-iiee}
-#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lag for all forecast initialised on the first of each month for +S1 and +S2 hindcasts.
-For each month, the number in parenthesis indicates the minimum lag at which the the mean error of each model is not statistically significant at a 1% level using a two-sisded t-test.
+#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lead time for all forecast initialised on the first of each month for +S1 and +S2 hindcasts.
+For each month, the number in parenthesis indicates the minimum lead time at which the the mean error of each model is not statistically significant at a 1% level using a two-sisded t-test.
 
@@
   labs(y = NULL, 
-       x = "Lag") +
+       x = "Lead time (days)") +
   #   scale_color_models +
@@
 To compare +S2 with +S1, we computed the IIEE for both models.
-This error measure is shown in @fig-iiee for all lags and forecasts initialised at the first of every month.
-+S1 has lower error at short lags at all months, with the errors converging to +S2 as the forecast goes on.
+This error measure is shown in @fig-iiee for forecasts initialised at the first of every month at all lead times.
++S1 has lower error at lead times up to 60 days at all months, with the errors converging to +S2 as the forecast goes on.
 The time to convergence depends on the month and it can be as short as two weeks June and July to as long as 160 days for forecasts initialised in February.
-Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the the memory of sea ice to initial conditions --at least from October to March when the data assimilated form the other components has little to no influence on sea ice.
+Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the memory of sea ice to initial conditions --at least from October to March when the data assimilated form the other components has little to no influence on sea ice.
 
-Also evident in @fig-iiee is the difference in the error spread at short lags between +S2 and +S1.
-In all months +S1 has much narrower error spread at lag 1. 
+Also evident in @fig-iiee is the difference in the error spread at short lead times between +S2 and +S1.
+In all months +S1 has much narrower error spread at the first day of the forecasts.
 The error in the initial conditions is not only small, but also fairly constant.
 This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
-For +S2, this is true only only between July and October, approximately.
-For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it's not constant.
+For +S2, this is true only between July and October.
+For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it is not constant.
 
 ```{r fig-iiee-variance}
-#| fig-cap: Mean spread of IIEE at different lags for different models. 
+#| fig-cap: Mean spread of IIEE at different lead times for different models. 
 #| fig.height: 6
@@
   # scale_color_models +
-  scale_x_continuous("Lag", breaks = NULL) +
+  scale_x_continuous("Lead time (days)", breaks = NULL) +
   scale_color_models +
@@
 The large initial error spread could be due either to large spread of ensemble members or to large spread of individual forecasts.
-@fig-iiee-variance splits the IIEE variance for each lag into the mean variance of each individual forecast and the variance of the mean error of each individual forecast, which adds up to the total variance.
-The average variance of each forecast is almost identical between forecast systems in all months.
-This shows that the ensemble spread of individual forecasts evolves identically, which, again, it's not unexpected because both systems share the same model formulation.
+@fig-iiee-variance splits the IIEE variance for each lead time into the mean variance of each individual forecast and the variance of the mean error of each individual forecast, which adds up to the total variance.
+The average variance of each forecast is almost identical between the two forecast systems in all months.
+This shows that the ensemble spread of individual forecasts evolves identically, which, again, is not unexpected because both systems share the same model formulation.
 This also shows that the perturbation scheme in +S2 is comparable to the one in +S1.
@@
 On the other hand, the spread of the mean error is always larger in +S2 than +S1.
-The difference is particularly large at short lags in some months, which coindice with the ones in which the data assimilation scheme is not influencing sea ice initial conditions.
+The difference is particularly large at short lead times in some months, which coincide with the ones in which the data assimilation scheme is not influencing sea ice initial conditions.
 
@@
 This bias is attributed to a faster and longer melt season between January and March, and slower growth between March and April.
-This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems either with the model's thermodynamic representation or with short wave radiation forcing, as shown in other climate models.
+This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems either with the model's thermodynamic representation or with short wave radiation forcing, as shown in other climate models [@zampieri2019; @roach2020].
 Even though +S2 shares the same model formulation with +S1, the latter does not suffer from this bias.
-This is due to the assimilation of sea ice concentration anomalies into the initial conditions, which successful corrects for the negative bias in the model. 
-The analysis suggests that the data assimilation system in +S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea ice component runs virtually free, reverting to its biased equilibrium state.
+This is due to the assimilation of sea ice concentrations into the initial conditions, which successful corrects for the negative bias in the model.
+Our analysis suggests that the data assimilation system in +S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea ice component runs virtually free, reverting to its biased equilibrium state.
 
@@
 This highlights February initial conditions as a crucial for determining sea ice evolution at least up to late June.
-Arctic sea ice forecast also show greater sensitivity to initial conditions in (boreal) summer compared with (boreal) winter [@day2014; @bunzel2016], so similar mechanism might be playing a role.
+Arctic sea ice forecast also show greater sensitivity to initial conditions in boreal summer compared with boreal winter [@day2014; @bunzel2016], so similar mechanism might be playing a role.
 
-Although forecasts are not skilful for forecasting panantarctic sea ice concentrations, there are some areas where the model does show skill. 
+Although forecasts are not skilful for forecasting pan-Antarctic sea ice concentrations, there are some areas where the model does show skill.
 The Weddell Sea, and the Ross Sea to a lesser extent are particularly well forecasted between June and November for forecasts initialised from January to June.
-These regions have been recognised as a regions of high predictability thanks to persistent and eastwardly advected upper ocean heat content anomalies [@bushuk2021] 
-Sea surface temperatures are assimilated by +S2 only in areas with temperatures greater than 0°C and errors don't show a clear easterly-propagating signal, so it is not clear if +S2 is leveraging the same source of predictability.
+These regions have been recognised as a regions of high predictability thanks to persistent and eastwardly advected upper ocean heat content anomalies [@bushuk2021].
+Sea surface temperatures are assimilated by +S2 only in areas with temperatures greater than 0°C and errors don't show a clear easterly-propagating signal, so it is not clear if +S2 is leveraging the same source of predictability..
 
------
+------------------------------------------------------------------------
 

### 2025-05-07 analysis/paper/access-ice.tex
Revision

@@
 
-Antarctic sea ice is in a remote location, but the potential impacts of
-its variability extends far into the lower latitudes, affecting the
-circulation of both the ocean and the atmosphere, ocean carbon uptake
-and biological processes. It is likely that a reduction in sea ice
-extent would reduce the meridional temperature gradient which, in turn,
-would reduce the strength and location of the eddy-driven jet (and the
-opposite circulation changes with an increase in sea ice), which is a
-crucial control on extratropical weather. In the northern hemisphere
-(NH), it has been hypothesised that the reduction in Arctic sea ice has
-led to a weaker, wavier jet, increasing the frequency of extreme events
-\citep{barnes2015}. There is some evidence of this effect in the
-southern hemisphere (SH). Models show a weakened vortex in the middle
-atmosphere, increased vertical wave flux and greater zonal wave 1
-amplitude in austral spring following a low sea ice maximum. But models
-do not fully agree on the exact phase of the zonal wave response, and
-observations do not show a clear signal \citep{rea2024}. The jet
-response to sea ice loss is also very model dependent, particularly
-sensitive to its climatological location, and relatively small compared
-with the direct effect of CO2 increase and sea surface temperature (SST)
-warming \citep{ayres2019}. On climate scales, it is thought that the
-temperature signal of a reduction in Antarctic sea ice mirrors a ``mini
-global warming'', and experiments also predict a weakening and
-equatorward shift of the SH eddy-driven jet
-\citep{raphael2011, england2018, england2020, ayres2022}.
+Antarctic sea ice covers the polar ocean, yet the potential impacts of
+its variability extend far into the mid and lower latitudes, affecting
+the circulation of both the ocean and the atmosphere, ocean carbon
+uptake and biological processes \citep[\citet{england2020},
+\citet{swadling2023}]{fogwill2020}. In coupled climate model
+experiments, a reduction in sea ice extent has been shown to reduce the
+meridional temperature gradient, in turn reducing the eddy-driven jet
+strength and shifting it equatorward \citetext{\citealp[
+\citet{england2020}]{england2018}; \citealp{ayres2019}; \citealp{ayres2022}}
+and vice versa for an increase in sea ice
+\citep{raphael2003a, raphael2011}, which is a crucial control on
+extratropical weather. In the Northern Hemisphere, it has been
+hypothesised that the reduction in Arctic sea ice has led to a weaker,
+wavier jet, increasing the frequency of extreme events
+\citep{barnes2015, smith2022b}. There is some evidence of this effect in
+the Southern Hemisphere. For example, model simulations show a weakened
+vortex in the middle atmosphere, increased vertical wave flux and
+greater zonal wave 1 amplitude in austral spring following lower than
+average sea ice extent at in early winter
+\citep[\citet{song2025}]{rea2024}. But models different models disagree
+on the exact phase of the zonal wave response, and observations do not
+show a clear signal \citep{rea2024}. The magnitude of the eddy-driven
+jet response to sea ice loss has been shown to be model dependent, and
+is sensitive to its climatological location \citep{ayres2019}. Further,
+sea ice induced changes in the jet have been found to be relatively
+small compared with the jet-response to the direct effect of CO2
+increase and global sea surface temperature (SST) warming
+\citep{ayres2022}.
 
@@
 relatively simple statistical methods \citep{massonnet2023}, which also
-underscores the need for better understanding of sea ice dynamics.
+underscores the need for better understanding of sea ice dynamics, and
+drivers.
 
@@
 month and the two previous days and run from 279 days. The ACCESS-S1
-hindcast is built in the same manner and running for the period
-1990--2012.
+hindcast is built in the same manner (albeit with 217-day runs) for the
+period 1990--2012.
 
@@
 NOAA CDR but computed independently using different algorithms. Figures
-prepared with this dataset are provided in the supplementary material.
+prepared with this dataset are provided in the supplementary material
+and don't differ significantly from the ones prepared using CDR.
 
@@
 compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC
-CDRV4 southern hemisphere domain, and also a zonally-varying RMSE by
-averaging over twenty-four 15° longitude slices around Antarctica.
+CDRV4 southern hemisphere domain, and also a zonally-varying RMSE
+computed over twenty-four 15° longitude slices around Antarctica.
 
@@
 
-\caption{\label{fig-hindcast-extent}Median sea ice extent for al
-hindcasts initialised the first of the month for ACCESS-S2 and ACCESS-S1
-in colours representing the start month with the median sea ice extent
-of NOAA/NSIDC CDRV4 in the left column and the median difference in the
-right column. The NOAA/NSIDC CDRV4 climatology is computed in the period
-corresponding to each hindcast. Circles represent the initial conditions
-at the first of every month and triangles represent the median value at
-the first of every month forecasted with the largest possible lead time.
-Colours indicate the initialisation month of the forecast.}
+\caption{\label{fig-hindcast-extent}Pan-antarctic median sea ice extent
+for al hindcasts initialised the first of the month for ACCESS-S2 and
+ACCESS-S1 in colours representing the start month with the median sea
+ice extent of NOAA/NSIDC CDRV4 in the left column and the median
+difference in the right column. The NOAA/NSIDC CDRV4 climatology is
+computed in the period corresponding to each hindcast. Circles represent
+the initial conditions at the first of every month and triangles
+represent the median value at the first of every month forecasted with
+the largest possible lead time. Colours indicate the initialisation
+month of the forecast.}
 
@@
 
-Both models' equilibrium state (triangles) show a negative bias of sea
-ice extent, particularly in the late-autumn and winter months. This is
-due primarily to faster melt during a longer melt season between January
-and March and slower growth during March and April. This is then
-partially balanced with faster growth between May and July
-(Fig.~\ref{fig-mean-growth}). Many sea ice models exhibit this
-systematic underestimation during the sea ice minimum and early freezing
-season \citep{massonnet2023} and could indicate problems in the
-representation of thermodynamics in the model \citep{zampieri2019}. It
-is not surprising that both forecasting systems converge to a similar
-climatology because they share the same model formulation.
-
-ACCESS-S2 initial conditions (circles) also have a negative bias,
+ACCESS-S2 initial conditions (circles) show an overall negative bias,
 especially in the late summer-early autumn, while ACCESS-S1 initial
-conditions are very close to observations. The faster growth between May
-and July reduces the negative bias in ACCESS-S2 for forecasts
-initialised from April to June, but it adds a positive bias in
-ACCESS-S1.
+conditions are very close to observations. Both models' equilibrium
+state (triangles) show a negative bias of sea ice extent, particularly
+in the late-autumn and winter months. This is due primarily to faster
+melt during a longer melt season between January and March and slower
+growth during March and April. This is then partially balanced with
+faster growth between May and July (Fig.~\ref{fig-mean-growth}). Many
+sea ice models exhibit this systematic underestimation during the sea
+ice minimum and early freezing season \citep{massonnet2023} and could
+indicate problems in the representation of thermodynamics in the model
+\citep{zampieri2019}. It is not surprising that both forecasting systems
+converge to a similar climatology because they share the same model
+formulation.
 
@@
 \caption{\label{fig-mean-growth}Median daily sea ice extent growth of
-ACCESS-S1 and ACCESS-S2 hindcasts and observations. Values are smoothed
-with a 11-day running mean.}
+ACCESS-S1 and ACCESS-S2 hindcasts and observations computed as the
+median daily differences in sea ice extent for each date and all lead
+times. Values are smoothed with a 11-day running mean.}
 
@@
 
-\caption{\label{fig-bias}Mean sea ice concentrations 1-month lag monthly
-mean ACCESS-S2 forecast bias compared with NSIDC.}
+\caption{\label{fig-bias}Mean difffernece between monthly sea ice
+concentrations at 0-month lead time ACCESS-S2 forecast and
+observations.}
 
@@
 concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
-shortest lag. From October to May, the model underestimates sea ice
+shortest lead time From October to May, the model underestimates sea ice
 concentrations in most regions except for the inner Weddell Sea in April
@@
 \caption{\label{fig-extent-anom}Monthly mean sea ice extent anomalies
-for ACCESS-S1 and ACCESS-S2 (black) and NOAA/NSIDC CDRV4 (blue)
-forecasted at selected lead times. The RMSE during the overlapping
-period (1990--2013) is shown on the top left.}
+for ACCESS-S1 and ACCESS-S2 and observations (black) forecasted at
+selected lead times. The RMSE during the overlapping period (1990--2013)
+is shown on the top left.}
 
@@
 anomaly forecast is relatively poor even in the first month, which stays
-relatively skilful even at lag 3. ACCESS-S2 shows much bigger
-variability than observations, with dramatic lows between 1995 and 2007
-and highs between 2007 and 2015.
+relatively skilful even at a lead time of 3 months. ACCESS-S2 shows much
+bigger variability than observations, with dramatic lows between 1995
+and 2007 and highs between 2007 and 2015.
 
@@
 \caption{\label{fig-rmse}Median RMSE of sea ice concentration anomalies
-as a function of forecast lag for all forecast initialised on the first
-of each month compared with a reference forecast of persistence of
-anomalies.}
+as a function of forecast lead time for all forecast initialised on the
+first of each month compared with a reference forecast of persistence of
+anomalies (black) and climatology (gray).}
 
 \end{figure}%
-
-monthly rmse/corelation at gridpoints per lead time.
 
@@
 any month. The only exception is around February, where the model has
-lower RMSE than the persistence forecast at virtually every lag.
+lower RMSE than the persistence forecast at virtually every lead time.
 
@@
 \texttt{r\ uniqueN(rmse\_lon\_mean\$lon)} meridional slices
-\texttt{r\ diff(unique(rmse\_lon\_mean\$lon)){[}1{]}}º wide.}
+\texttt{r\ diff(unique(rmse\_lon\_mean\$lon)){[}1{]}}º wide.
+Antarctica's coastline is shown at the bottom of each panel for
+reference.}
 
@@
 
-The skill shown by ACCESS-S2 at February-March forecasts is evidenced by
-a band of negative values across almost all longitudes. The only region
-where February-March sea ice is not well forecasted is east of the
-Antarctic Peninsula, which is the only region with consistent summer sea
-ice. This suggests that the positive skill around the summer minimum
+The skill shown by ACCESS-S2 at February-March forecasts is observed
+
+by a band of negative values across almost all longitudes. The only
+region where February-March sea ice is not well forecasted is east of
+the Antarctic Peninsula, which is the only region with consistent summer
+sea ice. This suggests that the positive skill around the summer minimum
 comes from the model correctly forecasting no ice where no ice is ever
@@
 Forecasts initialised on January 1st (Fig.~\ref{fig-rmse_lon} panel a)
-show RMSE values below persistence at lags larger than 180 days
+show RMSE values below persistence at lead times larger than 180 days
 (corresponding to July through September) in the Ross and Weddell Sea.
@@
 \caption{\label{fig-iiee}Median and 95\% coverage of Integrated Ice Edge
-Error as a function of forecast lag for all forecast initialised on the
-first of each month for ACCESS-S1 and ACCESS-S2 hindcasts. For each
-month, the number in parenthesis indicates the minimum lag at which the
-the mean error of each model is not statistically significant at a 1\%
-level using a two-sisded t-test.}
+Error as a function of forecast lead time for all forecast initialised
+on the first of each month for ACCESS-S1 and ACCESS-S2 hindcasts. For
+each month, the number in parenthesis indicates the minimum lead time at
+which the the mean error of each model is not statistically significant
+at a 1\% level using a two-sisded t-test.}
 
@@
 Also evident in Figure~\ref{fig-iiee} is the difference in the error
-spread at short lags between ACCESS-S2 and ACCESS-S1. In all months
-ACCESS-S1 has much narrower error spread at the first day of the
+spread at short lead times between ACCESS-S2 and ACCESS-S1. In all
+months ACCESS-S1 has much narrower error spread at the first day of the
 forecasts. The error in the initial conditions is not only small, but
@@
 spread as errors accumulate differently in different forecasts. For
-ACCESS-S2, this is true only between July and October, approximately.
-For all other months, the error spread is more or less stable throughout
-the forecast window, indicating that not only the initial error is high,
-but it is not constant.
+ACCESS-S2, this is true only between July and October. For all other
+months, the error spread is more or less stable throughout the forecast
+window, indicating that not only the initial error is high, but it is
+not constant.
 
@@
 
-\caption{\label{fig-iiee-variance}Mean spread of IIEE at different lags
-for different models.}
+\caption{\label{fig-iiee-variance}Mean spread of IIEE at different lead
+times for different models.}
 
@@
 either with the model's thermodynamic representation or with short wave
-radiation forcing, as shown in other climate models. Even though
-ACCESS-S2 shares the same model formulation with ACCESS-S1, the latter
-does not suffer from this bias. This is due to the assimilation of sea
-ice concentrations into the initial conditions, which successful
-corrects for the negative bias in the model. Our analysis suggests that
-the data assimilation system in ACCESS-S2 is only effectively
-influencing sea ice initial conditions from June to October, while the
-rest of the year, the sea ice component runs virtually free, reverting
-to its biased equilibrium state.
+radiation forcing, as shown in other climate models
+\citep{zampieri2019, roach2020}. Even though ACCESS-S2 shares the same
+model formulation with ACCESS-S1, the latter does not suffer from this
+bias. This is due to the assimilation of sea ice concentrations into the
+initial conditions, which successful corrects for the negative bias in
+the model. Our analysis suggests that the data assimilation system in
+ACCESS-S2 is only effectively influencing sea ice initial conditions
+from June to October, while the rest of the year, the sea ice component
+runs virtually free, reverting to its biased equilibrium state.
 

### 2025-05-09 analysis/sources.R
Re render

@@ -1,0 +1,126 @@
+library(ggplot2)
+
+og_extent <- BT() |>
+    rcdo::cdo_gtc(0.15) |>
+    rcdo::cdo_fldint() |>
+    rcdo::cdo_execute(options = c("-L"))
+
+cdr_extent <- CDR_BT() |>
+    rcdo::cdo_gtc(0.15) |>
+    rcdo::cdo_fldint() |>
+    rcdo::cdo_execute(options = c("-L"))
+
+# era5_extent <- ERA5() |>
+#     rcdo::cdo_gtc(0.15) |>
+#     rcdo::cdo_fldint() |>
+#     rcdo::cdo_execute(options = c("-L"))
+
+
+extents <- metR::ReadNetCDF(og_extent,
+    vars = c(og_bt = "aice")
+) |>
+    _[
+        metR::ReadNetCDF(cdr_extent,
+            vars = c(cdr_bt = "aice")
+        ),
+        on = .NATURAL
+    ] |>
+    # _[
+    #     metR::ReadNetCDF(era5_extent,
+    #         vars = c(era5 = "aice")
+    #     ) |>
+    #         data.table::setnames("valid_time", "time"),
+    #     on = .NATURAL
+    # ] |>
+    _[, lon := NULL] |>
+    _[, lat := NULL]
+
+extents |>
+    data.table::copy() |>
+    _[, dif := og_bt - cdr_bt] |>
+    data.table::melt(id.vars = "time", value.name = "extent") |>
+    _[extent == 0, extent := NA] |>
+    # _[, value := value - mean(value, na.rm = TRUE), by = .(data.table::yday(time))] |>
+    _[data.table::year(time) == 2023] |>
+    ggplot(aes(time, extent)) +
+    geom_line(aes(color = variable))
+
+extents |>
+    data.table::copy() |>
+    _[, dif := og_bt - cdr_bt] |>
+    data.table::melt(id.vars = "time", value.name = "extent") |>
+    _[extent == 0, extent := NA] |>
+    # _[, value := value - mean(value, na.rm = TRUE), by = .(data.table::yday(time))] |>
+    _[data.table::year(time) == 2000] |>
+    ggplot(aes(time, extent)) +
+    geom_line(aes(color = variable))
+
+
+extents |>
+    data.table::copy() |>
+    _[, dif := og_bt - cdr_bt] |>
+    data.table::melt(id.vars = "time", value.name = "extent") |>
[…]

### 2025-05-09 analysis/paper/access-ice.qmd
New render

@@
 ```{r fig-hindcast-extent}
-#| fig-cap: Pan-antarctic median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month with the median sea ice extent of +CDR in the left column and the median difference in the right column.
+#| fig-cap: Pan-Antarctic median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month with the median sea ice extent of +CDR in the left column and the median difference in the right column.
 The +CDR climatology is computed in the period corresponding to each hindcast.
@@
 #| fig-height: 6
-#| fig-cap: Mean difffernece between monthly sea ice concentrations at 0-month lead time +S2 forecast and observations.
+#| fig-cap: Mean difference between monthly sea ice concentrations at 0-month lead time +S2 forecast and observations.
 
@@
 #| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lead time for all forecast initialised on the first of each month for +S1 and +S2 hindcasts.
-For each month, the number in parenthesis indicates the minimum lead time at which the the mean error of each model is not statistically significant at a 1% level using a two-sisded t-test.
+For each month, the number in parenthesis indicates the minimum lead time at which the the mean error of each model is not statistically significant at a 1% level using a two-sided t-test.
 
@@
 
-# Suplementary figures
+# Supplementary figures
 

### 2025-05-12 analysis/paper/access-ice.qmd
render

@@
 Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend up to 2015, which models systematically failed to reproduce and puzzled researchers [@hobbs2016].
-Then, in 2016 Antarctic sea ice extent dropped precipitously [@turner2017a] and has been at low and record low levels since [@purich2023; @hobs2024; @gilbert2024; @josey2024], even more highlighting our lack of understanding of Antarctic sea ice variability and long-term change [@espinosa2024].
+Then, in 2016 Antarctic sea ice extent dropped precipitously [@turner2017a] and has been at low and record low levels since [@purich2023; @hobbs2024; @gilbert2024; @josey2024], even more highlighting our lack of understanding of Antarctic sea ice variability and long-term change [@espinosa2024].
 
@@
 In coupled climate model experiments, a reduction in sea ice extent has been shown to reduce the meridional temperature gradient, in turn reducing the eddy-driven jet strength and shifting it equatorward [@england2018 @england2020; @ayres2019; @ayres2022] and vice versa for an increase in sea ice [@raphael2003a; @raphael2011], which is a crucial control on extratropical weather.
-In the Northern Hemisphere, it has been hypothesised that the reduction in Arctic sea ice has led to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015; @smith2022b]. 
+In the Northern Hemisphere, it has been hypothesised that the reduction in Arctic sea ice has led to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015; @smith2022]. 
 There is some evidence of this effect in the Southern Hemisphere.
@@
 +S2 [@wedd2022] became operational in October 2021 by replacing the +S1 system [@hudson2017]..
-The model components of both +S2 and +S1 consist of the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
+The model components of both +S2 and +S1 consist of the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@gurvan2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
 The atmosphere has a N216 horizontal resolution (\~60km in the mid-latitudes) with 85 vertical levels.

### 2025-05-12 analysis/paper/access-ice.qmd
Some light editing

@@
 
-Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend up to 2015, which models systematically failed to reproduce and puzzled researchers [@hobbs2016].
+Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend up to 2015. 
+This trend puzzled researchers and was not reproduced by models [@hobbs2016].
 Then, in 2016 Antarctic sea ice extent dropped precipitously [@turner2017a] and has been at low and record low levels since [@purich2023; @hobbs2024; @gilbert2024; @josey2024], even more highlighting our lack of understanding of Antarctic sea ice variability and long-term change [@espinosa2024].
@@
 Antarctic sea ice covers the polar ocean, yet the potential impacts of its variability extend far into the mid and lower latitudes, affecting the circulation of both the ocean and the atmosphere, ocean carbon uptake and biological processes [@fogwill2020, @england2020, @swadling2023].
-In coupled climate model experiments, a reduction in sea ice extent has been shown to reduce the meridional temperature gradient, in turn reducing the eddy-driven jet strength and shifting it equatorward [@england2018 @england2020; @ayres2019; @ayres2022] and vice versa for an increase in sea ice [@raphael2003a; @raphael2011], which is a crucial control on extratropical weather.
+In coupled climate model experiments, a reduction in sea ice extent has been shown to reduce the meridional temperature gradient, in turn reducing the eddy-driven jet strength and shifting it equatorward [@england2018; @england2020; @ayres2019; @ayres2022] and vice versa for an increase in sea ice [@raphael2003a; @raphael2011], which is a crucial control on extratropical weather.
 In the Northern Hemisphere, it has been hypothesised that the reduction in Arctic sea ice has led to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015; @smith2022]. 
 There is some evidence of this effect in the Southern Hemisphere.
-For example, model simulations show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in austral spring following lower than average sea ice extent at in early winter [@rea2024, @song2025].
-But models different models disagree on the exact phase of the zonal wave response, and observations do not show a clear signal [@rea2024].
+For example, model simulations show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in austral spring following lower than average sea ice extent in early winter [@rea2024, @song2025].
+But models disagree on the exact phase of the zonal wave response, and observations do not show a clear signal [@rea2024].
 The magnitude of the eddy-driven jet response  to sea ice loss has been shown to be model dependent, and is sensitive to its climatological location [@ayres2019].
@@
 
-Correctly modelling Antarctic sea is not only necessary for process understanding and climate projections to inform adaptation strategies but also for accurate seasonal to sub-seasonal forecasts, which are crucial for operations in and around the Antarctic continent such as scientific missions, fisheries and tourism [@desilva2020 @wagner2020].
+Correctly modelling Antarctic sea ice is essential for understanding processes and improving climate projections to inform adaptation strategies.
+Accurate seasonal to sub-seasonal forecasts are also crucial for operations in and around the Antarctic continent, including scientific missions, fisheries, and tourism [@desilva2020; @wagner2020].
 However, producing accurate Antarctic sea ice forecasts has been challenging due to model biases, inherent large variability and complexity, and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
@@
 
-Anomalies for each hindcast are be taken with respect to a forecast-dependent climatology for the period 1990--2012.
+Anomalies for each hindcast are taken with respect to a climatology specific to each initalisation date for the period 1990--2012. 
 This serves as a first-order correction of model drift.
@@
 
-Sea-ice extent is defined as the area of ocean covered with at least 15% ice.
+Sea-ice extent is defined as the area of the ocean covered with at least 15% ice.
 This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea ice conditions (cite -- probably something from NSIDC).
@@
 Pan-Antarctic (net) sea ice extent serves as a rough hemispheric measure of the amount of sea ice, but it does not take into account the spatial distribution.
-A model could have relatively accurate extent of the net ice but with different regional distributions.
+A model could have a relatively accurate extent of the net ice but with different regional distributions.
 To account for location errors, we computed the Root Mean Squared Error (RMSE) of sea ice concentration anomalies and the Integrated Ice Edge Error [IIEE; @goessling2016].
@@
 From June to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
-The rest of the year there is little if any difference between circles and triangles in +S2, indicating that the ocean and atmosphere data assimilation is not affecting seas ice and that this component of the model is virtually free-running.
+The rest of the year there is little if any difference between circles and triangles in +S2, indicating that the ocean and atmosphere data assimilation is not affecting sea ice and that this component of the model is virtually free-running.
 
@@
 #| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lead time for all forecast initialised on the first of each month for +S1 and +S2 hindcasts.
-For each month, the number in parenthesis indicates the minimum lead time at which the the mean error of each model is not statistically significant at a 1% level using a two-sided t-test.
+For each month, the number in parenthesis indicates the minimum lead time at which the mean error of each model is not statistically significant at a 1% level using a two-sided t-test.
 
@@
 The time to convergence depends on the month and it can be as short as two weeks June and July to as long as 160 days for forecasts initialised in February.
-Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the memory of sea ice to initial conditions --at least from October to March when the data assimilated form the other components has little to no influence on sea ice.
+Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the memory of sea ice to initial conditions --at least from October to March when the data assimilated from the other components has little to no influence on sea ice.
 
@@
 Even though +S2 shares the same model formulation with +S1, the latter does not suffer from this bias.
-This is due to the assimilation of sea ice concentrations into the initial conditions, which successful corrects for the negative bias in the model.
+This is due to the assimilation of sea ice concentrations into the initial conditions, which successfully corrects for the negative bias in the model.
 Our analysis suggests that the data assimilation system in +S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea ice component runs virtually free, reverting to its biased equilibrium state.
@@
 Based on the observation that +S2 sea ice initial conditions are essentially not initialised, comparing its forecasts with +S1's allows us to estimate the time-scale for which initial conditions are important.
-This highlights February initial conditions as a crucial for determining sea ice evolution at least up to late June.
-Arctic sea ice forecast also show greater sensitivity to initial conditions in boreal summer compared with boreal winter [@day2014; @bunzel2016], so similar mechanism might be playing a role.
+This highlights February initial conditions as crucial for determining sea ice evolution at least up to late June.
+Arctic sea ice forecasts also show greater sensitivity to initial conditions in boreal summer compared with boreal winter [@day2014; @bunzel2016], so a similar mechanism might be playing a role.
 
@@
 The Weddell Sea, and the Ross Sea to a lesser extent are particularly well forecasted between June and November for forecasts initialised from January to June.
-These regions have been recognised as a regions of high predictability thanks to persistent and eastwardly advected upper ocean heat content anomalies [@bushuk2021].
+These regions have been recognised as a region of high predictability thanks to persistent and eastwardly advected upper ocean heat content anomalies [@bushuk2021].
 Sea surface temperatures are assimilated by +S2 only in areas with temperatures greater than 0°C and errors don't show a clear easterly-propagating signal, so it is not clear if +S2 is leveraging the same source of predictability..

### 2025-07-09 R/datasets.R
Fixes OSI download

@@ -284,7 +284,7 @@ OSI <- function() {
   dir <- here::here("data/temp/osi/")
   dir.create(dir, FALSE, TRUE)
   requests <- lapply(years, \(year) {
-    raw_file <-  paste0("OSI_", year)
+    raw_file <-  paste0("OSI_", year, ".zip")
     
     if (file.exists(here::here(dir, raw_file))) {
       return(NULL)
@@ -307,7 +307,7 @@ OSI <- function() {
   })
   requests <- Filter(Negate(is.null), requests) 
   
-  if (length(requests != 0)) {
+  if (length(requests) != 0) {
     res <- ecmwfr::wf_request_batch(requests, 
                                     retry = 1,
                                     workers = 20,
@@ -316,7 +316,7 @@ OSI <- function() {
   
   files_year <- vapply(years, FUN.VALUE = character(1), \(year) {
     
-    in_file <- file.path(dir, paste0("OSI_", year))
+    in_file <- file.path(dir, paste0("OSI_", year, ".zip"))
     out_file <- file.path(dir, paste0(year, "merged"))
     
     if (file.exists(out_file)) {

### 2025-07-09 analysis/paper/access-ice.qmd
Filters bad extent values in observations

@@
   rbindlist(idcol = "dataset") |>
-  _[, let(lon = NULL, lat = NULL)]
+  _[, let(lon = NULL, lat = NULL)] |> 
+  _[, anom  := aice - mean(aice, na.rm = TRUE), by  = .(yday(time), dataset)] |>
+  # some values before 2010 are zero or extremely low. likely due to missing values.
+  _[year(time) < 2010 & anom < -2.5e12, aice := NA] |>  
+  _[, anom := NULL]
 ```

### 2025-07-09 analysis/paper/access-ice.qmd
Simplifies hindcast-extent plot.

@@
 
-hindcast_bias_extent <- hindcast_extent |> 
-  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
-                              lag = as.numeric(as.Date(time) - forecast_time),
-                              model)] |> 
+hindcast_extent[, .(model, forecast_time, member, time, Forecast = aice)] |> 
+  extent_daily[i = _, on = c("time")] |> 
+  _[, Difference := Forecast - aice] |> 
+  _[, aice := NULL] |> 
+  tidyfast::dt_pivot_longer(cols = c(Forecast, Difference)) |> 
+  _[, median_ci(value), 
+    by = .(dataset, 
+           model, 
+           measure = name,
+           forecast_time = update(forecast_time, year = 2001),
+           lag = as.numeric(as.Date(time) - forecast_time))] |> 
   _[, time := forecast_time + lag] |>
@@
   _[, month := lubridate::month(forecast_time, label = TRUE)] |> 
+  _[, model := factor(model, levels = c("S1", "S2"))] |> 
+  _[, measure := factor(measure, levels = c("Forecast", "Difference"))] |>
+
   ggplot(aes(time2, mid)) +
-  geom_line(data = extent_climatology[dataset == which_dataset], 
+  geom_line(data = extent_climatology[dataset == which_dataset] |> 
+              _[, measure := factor("Forecast", levels = c("Forecast", "Difference"))], 
             aes(x = as.Date(time),
                 group = dataset)) +
-  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +
 
-  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
-             aes(fill = month), 
+  geom_hline(data = data.table(y = 0, measure = factor("Difference", levels = c("Forecast", "Difference"))), aes(yintercept = 0)) +
+  geom_line(aes(colour = month, 
+                group = interaction(forecast_time, time != time2))) +
+  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], 
+                           by = .(forecast_time, model, measure)], 
+             aes(fill = lubridate::month(time2, label = TRUE)), 
              shape = 24, size = 2.2) +
-
-  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
+  geom_point(data = \(x) x[lag == 1], 
+             aes(fill = lubridate::month(time2, label = TRUE)), 
              shape = 21, size = 2) +
@@
                      aesthetics = c("fill", "color"), guide = "none") +
-  facet_wrap(~ model, ncol = 1, labeller = labeller(model = labels_models)) +
-  tag_facets()
-
-hindcast_bias_extent_dif <- copy(hindcast_bias_extent)
-
-hindcast_bias_extent_dif$layers[[1]] <- geom_hline(yintercept = 0) 
-hindcast_bias_extent_dif$data <- extent_daily[hindcast_extent, on = c("time")] |> 
-  _[, median_ci(i.aice - aice), 
-    by = .(forecast_time = update(forecast_time, year = 2001),
-           lag = as.numeric(as.Date(time) - forecast_time),
-           model)] |> 
-  _[, time := forecast_time + lag] |>
-  _[, time2 := update(time, year = 2001)] |>
-  _[, month := lubridate::month(forecast_time, label = TRUE)]
-
-hindcast_bias_extent +
-  hindcast_bias_extent_dif
-
+  facet_grid(measure ~ model, scales = "free_y")
 ```

### 2025-10-06 analysis/paper/access-ice.qmd
Third draft

--- 2025-07-09 analysis/paper/access-ice.qmd
@@
 
-abstract: |
-  The abstract goes here.
-  It can also be on _multiple lines_.
+abstract: The abstract goes here.
 bibliography: references.bib
@@
 # This section is mandatory even if you declare that no competing interests are present.
-competinginterests: |
-  The authors declare no competing interests.
+competinginterests: The authors declare no competing interests.
 # See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
 # Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
-copyrightstatement: |
-  The author's copyright for this publication is transferred to institution/company. 
+copyrightstatement: The author's copyright for this publication is transferred to institution/company. 
 ### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
@@
   #  use this to add a statement when having only data sets available
-  codedata: |
-    use this to add a statement when having data sets and software code available
-  sample: |
-    use this section when having geoscientific samples available
-videosupplement: |
-  use this section when having video supplements available
-authorcontribution: |
-  Contributions
-disclaimer: |
-  We like Copernicus.
-acknowledgements: |
-  Acknowledgements will gohere
+  codedata: use this to add a statement when having data sets and software code available
+  sample: use this section when having geoscientific samples available
+  videosupplement: use this section when having video supplements available
+  authorcontribution: Contributions
+  disclaimer: We like Copernicus.
+  acknowledgements: Acknowledgements will gohere
   appendix: |
@@
     Please add `\clearpage` between each table and/or figure.
-format: docx
+
+format: pdf
 
@@
 library(patchwork)
+library(mirai)
 
 cdo_options_set("-L")
+cdo_cache_set(here::here("data/temp/cache"))
 ```
@@
 
+```{r mirai_daemons}
+if (!mirai::daemons_set()) invisible(daemons(6))
+```
+
 # Abstract
@@
 
-Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice has experienced a slightly increasing trend up to 2015. 
-This trend puzzled researchers and was not reproduced by models [@hobbs2016].
-Then, in 2016 Antarctic sea ice extent dropped precipitously [@turner2017a] and has been at low and record low levels since [@purich2023; @hobbs2024; @gilbert2024; @josey2024], even more highlighting our lack of understanding of Antarctic sea ice variability and long-term change [@espinosa2024].
-
-Antarctic sea ice covers the polar ocean, yet the potential impacts of its variability extend far into the mid and lower latitudes, affecting the circulation of both the ocean and the atmosphere, ocean carbon uptake and biological processes [@fogwill2020, @england2020, @swadling2023].
-In coupled climate model experiments, a reduction in sea ice extent has been shown to reduce the meridional temperature gradient, in turn reducing the eddy-driven jet strength and shifting it equatorward [@england2018; @england2020; @ayres2019; @ayres2022] and vice versa for an increase in sea ice [@raphael2003a; @raphael2011], which is a crucial control on extratropical weather.
-In the Northern Hemisphere, it has been hypothesised that the reduction in Arctic sea ice has led to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015; @smith2022]. 
-There is some evidence of this effect in the Southern Hemisphere.
-For example, model simulations show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in austral spring following lower than average sea ice extent in early winter [@rea2024, @song2025].
-But models disagree on the exact phase of the zonal wave response, and observations do not show a clear signal [@rea2024].
-The magnitude of the eddy-driven jet response  to sea ice loss has been shown to be model dependent, and is sensitive to its climatological location [@ayres2019].
-Further, sea ice induced changes in the jet have been found to be  relatively small compared with the jet-response to the direct effect of CO2 increase and global sea surface temperature (SST) warming [@ayres2022].
-
 Correctly modelling Antarctic sea ice is essential for understanding processes and improving climate projections to inform adaptation strategies. 
-Accurate seasonal to sub-seasonal forecasts are also crucial for operations in and around the Antarctic continent, including scientific missions, fisheries, and tourism [@desilva2020; @wagner2020].
-However, producing accurate Antarctic sea ice forecasts has been challenging due to model biases, inherent large variability and complexity, and it has lagged behind Arctic sea ice forecasts [@zampieri2019].
-Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics, and drivers.
-
-In this work we evaluate sea ice forecasts produced by the Australian Community Climate and Earth System Simulator -- seasonal version2 (+S2), which is the Australian Bureau of Meteorology (BoM)'s current operational subseasonal-to-seasonal prediction system, and compare its forecast skill with that of the former system +S1.
-This evaluation will inform future work with the prediction system as a research tool and explore the potential of using its sea-ice forecasts for decision-making.
+Accurate seasonal to sub-seasonal forecasts are also crucial for operation contingency planning in and around the Antarctic continent, including scientific missions, fisheries, and tourism [@desilva2020; @wagner2020].
+Getting sea-ice right might also help improve weather forecasts over and away from sea-ice regions [@rinke2006; @wang2024 @semmler2016].
+
+However, producing accurate Antarctic sea-ice forecasts has been challenging due to model biases, inherent large variability and complexity, and it has lagged behind Arctic sea-ice forecasts [@zampieri2019;  @gao2024a].
+Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023] and machine learning approaches [e.g.
+@dong2024; @lin2025], which also underscores the need for better understanding of sea-ice dynamics, and drivers of its variability.
+
+Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea-ice initial conditions affect the quality of the forecast and in which timescales.
+Exploring seasonal predictions of Arctic sea-ice, @guemas2016 found that sea-ice initial conditions are only important in Autumn to predict Summer sea-ice, but the impact wasn't as dramatic when predicting Winter sea-ice.
+@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea-ice forecast initialised in July but not when initialised in January.
+
+For the Antarctic, @holland2013 studied the initial-value predictability of Antarctic sea-ice in the CCSESM3 model. 
+They found that sea-ice and ocean initial conditions provide predictive information to forecast sea-ice edge location several months in advance and that some predictability is retained for up to two years thanks to ocean heat content anomalies that are advected eastward.
+This is in contrast with @marchi2020, who ran perfect model experiments to argue that uncertainty in the atmosphere state and evolution is the main driver of uncertainty in Antarctic sea-ice extent prediction, with sea-ice and ocean initial conditions having lesser importance.
+More recently, @morioka2022 studied decadal forecasts of Antarctic sea-ice and found that initialising ocean and sea-ice improved the correlation between simulated and observed sea-ice concentration evolution in the Amundsen–Bellingshausen Sea.
+It is hard to compare these studies since they are based on forecasts initialised at different times of the year: @holland2013 ran 20 ensemble members initialised on the 1st of January of a particular year, @marchi2020 ran forecasts from the 1st of March and 1st of September and @morioka2022 run forecasts only from the 1st of March.
+@marchi2020 also used a coupled ocean–sea-ice model instead of a fully coupled model like @holland2013 did.
+
+In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the Australian Community Climate and Earth System Simulator – Seasonal (ACCES-S) to its version two. 
+The change in version was focused on using ocean, sea-ice and land initial conditions generated by the BoM instead of depending on the UK’ Met Office. 
+Crucially, ACCESS-S2 does not assimilate sea-ice observations, and instead sea-ice is only affected by the ocean and atmospheric data assimilation via the coupled integration.
+
+Since model configuration is identical between the two systems, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year.
+This can be an opportunity to test the effect of sea-ice initial conditions on the forecast of sea-ice concentrations.
+
+In this study we compare sea-ice hindcasts produced by ACCESS-S1 and ACCESS-S2. 
+We focus on seasonality of errors and biases and the effect of the data assimilation system. 
+This comparison will also inform future work with the prediction system as a research tool to better understand the dynamics and variability of the Antarctic sea ice and its impacts on the climate system as well as to explore the potential of using its sea-ice forecasts for decision-making.
+The work will also serve as a benchmark for future prediction systems to attempt to improve upon.
 
@@
     _[, time := update(time, year = 2001)] |> 
-    _[, median_ci(aice), by = .(time, dataset)] |> 
+    _[, average(aice), by = .(time, dataset)] |> 
     _[, model := model] |> 
@@
 ```{r range_max_lag}
-range_max_lag <- hindcast_extent[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)] |> 
+range_max_lag <- hindcast_extent[mday(time) == 2, .SD[which.max(lag)], 
+                                 by = .(forecast_time, model)] |> 
   _[, .(max = max(lag),
@@
 
-hindcast_bias_extent <- hindcast_extent |> 
-  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
-                              lag = as.numeric(as.Date(time) - forecast_time),
-                              model)] |> 
+hindcast_extent[, .(model, forecast_time, member, time, Forecast = aice)] |> 
+  extent_daily[i = _, on = c("time")] |> 
+  _[, Difference := Forecast - aice] |> 
+  _[, aice := NULL] |> 
+  tidyfast::dt_pivot_longer(cols = c(Forecast, Difference)) |> 
+  _[, average(value), 
+    by = .(dataset, 
+           model, 
+           measure = name,
+           forecast_time = update(forecast_time, year = 2001),
+           lag = as.numeric(as.Date(time) - forecast_time))] |> 
   _[, time := forecast_time + lag] |>
@@
   _[, month := lubridate::month(forecast_time, label = TRUE)] |> 
-  ggplot(aes(time2, mid)) +
-  geom_line(data = extent_climatology[dataset == which_dataset], 
-            aes(x = as.Date(time),
+  _[, model := factor(model, levels = c("S1", "S2"))] |> 
+  _[, measure := factor(measure, levels = c("Forecast", "Difference"))] |>
+
+  ggplot(aes(time2, estimate)) +
+  geom_line(data = extent_climatology[dataset == which_dataset] |> 
+              _[, measure := factor("Forecast", levels = c("Forecast", "Difference"))], 
+            aes(x = as.Date(time), colour = dataset, 
                 group = dataset)) +
-  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +
-
-  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
-             aes(fill = month), 
+
+  geom_hline(data = data.table(y = 0, measure = factor("Difference", levels = c("Forecast", "Difference"))), aes(yintercept = 0)) +
+  geom_line(aes(group = interaction(forecast_time, time != time2), colour = model)) +
+  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], 
+                           by = .(forecast_time, model, measure)], 
+             aes(fill = model), 
              shape = 24, size = 2.2) +
-
-  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
+  geom_point(data = \(x) x[lag == 1], 
+             aes(fill = model), 
              shape = 21, size = 2) +
@@
   scale_y_continuous(NULL, labels = labels_extent) +
-  scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"), 
-                     aesthetics = c("fill", "color"), guide = "none") +
-  facet_wrap(~ model, ncol = 1, labeller = labeller(model = labels_models)) +
-  tag_facets()
-
-hindcast_bias_extent_dif <- copy(hindcast_bias_extent)
-
-hindcast_bias_extent_dif$layers[[1]] <- geom_hline(yintercept = 0) 
-hindcast_bias_extent_dif$data <- extent_daily[hindcast_extent, on = c("time")] |> 
-  _[, median_ci(i.aice - aice), 
-    by = .(forecast_time = update(forecast_time, year = 2001),
-           lag = as.numeric(as.Date(time) - forecast_time),
-           model)] |> 
-  _[, time := forecast_time + lag] |>
-  _[, time2 := update(time, year = 2001)] |>
-  _[, month := lubridate::month(forecast_time, label = TRUE)]
-
-hindcast_bias_extent +
-  hindcast_bias_extent_dif
-
+  scale_color_models + 
+  scale_fill_models +
+  guides(colour = "none", fill = "none") +
+  # scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"), 
+  #                    aesthetics = c("fill", "color"), guide = "none") +
+  facet_grid(measure ~ model, scales = "free_y", labeller = labeller(model = labels_models)) +
+  tag_facets(tag = "rc") +
+  coord_cartesian(clip = "off")
 ```
@@
 The difference between the initial conditions and the model preferred state can be attributed to the effect of data assimilation, which in +S2 is due solely to atmospheric and oceanic data assimilation.
-From June to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
+From May to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
 The rest of the year there is little if any difference between circles and triangles in +S2, indicating that the ocean and atmosphere data assimilation is not affecting sea ice and that this component of the model is virtually free-running.
@@
 ```{r extent-delta-compute}
+n <- 366*3
+w <- 11
+
 dif_hindcast <- hindcast_extent |> 
@@
   _[order(time)] |> 
-  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(model, forecast_time, member)] |> 
-  _[, median_ci(dif), by = .(model, time = update(time, year = 2000))]
-
-dif_hindcast <- rbind(
-  dif_hindcast |> 
-    copy() |> 
-    _[, time := update(time, year = 1999)],
-  dif_hindcast,
-  dif_hindcast |> 
-    copy() |> 
-    _[, time := update(time, year = 2001)]
-)
+  _[, .(aice = mean(aice)), by = .(model, forecast_time, time)] |> 
+  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), 
+    by = .(model, forecast_time)] |> 
+  _[, dif := frollmean(dif, n = w, align = "center"), by = .(model, forecast_time)] |> 
+  na.omit() |> 
+  _[, average(dif), by = .(model, month(forecast_time), time = update(time, year = 2000))]
 
@@
   _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
-  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
+  na.omit() |> 
+  _[, dif := frollmean(dif, n = w, align = "center"), by = .(dataset)] |> 
+  _[, average(dif), by = .(dataset, time = update(time, year = 2000))] |> 
   _[dataset == which_dataset]
 
-dif_extent <- rbind(
-  dif_extent |> 
-    copy() |> 
-    _[, time := update(time, year = 1999)],
-  dif_extent,
-  dif_extent |> 
-    copy() |> 
-    _[, time := update(time, year = 2001)]
-)
 ```
@@
 
-n <- 366*3
-w <- 11
 dif_extent |> 
   copy() |> 
-  _[, mid := frollmean(mid, n = w, align = "center"), by = dataset] |> 
-  ggplot(aes(time, mid)) +
+  ggplot(aes(time, estimate)) +
   geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
   geom_line(aes(color = dataset)) +
-  # geom_smooth(aes(group = dataset, color = dataset), 
-  #             method = "loess",
-  #             span = w/n, linewidth = 0.5, se = FALSE, n = n)  +
-  # geom_line() +
-
-  geom_line(data = dif_hindcast |> 
-              copy() |> 
-              _[, mid := frollmean(mid, n = w, align = "center"), by = model],
-            aes(color = model)) +
-  # geom_smooth(data = dif_hindcast,
-  #             method = "loess",
-  #             aes(color = model), span = w/n, linewidth = 0.5, se = FALSE, n = n) +
+  geom_line(data = dif_hindcast,
+            aes(color = model, group = interaction(model, month, month > month(time)))) +
   scale_x_datetime(NULL, date_breaks = "1 month", 
@@
 hindcast_clim <- lapply(setNames(months, months), \(month) {
-  here::here(glue::glue("data/derived/climatology/S2/{zeropad(month)}/em.nc")) |> 
+  lapply(c("S1", "S2"), \(model) {
+    here::here(glue::glue("data/derived/climatology/{model}/{zeropad(month)}/em.nc")) |> 
       cdo_selmonth(month) |> 
       cdo_execute() |> 
-    ReadNetCDF("aice")
+      ReadNetCDF("aice")  |> 
+      _[, model := model]
+  }) |> 
+    rbindlist()
 }) |> 
@@
   copy() |> 
-  _[, .(aice = mean(aice)), by = .(month(time), ygrid, xgrid, forecast_month)] |> 
+  _[, .(aice = mean(aice)), by = .(month(time), ygrid, xgrid, forecast_month, model)] |> 
   _[forecast_month == month]
@@
 ```{r fig-bias}
-#| fig-width: 9
+#| fig-width: 6
 #| fig-height: 6
 #| fig-cap: Mean difference between monthly sea ice concentrations at 0-month lead time +S2 forecast and observations. 
-
-ice_shelves_nsicd <- fread(here::here("data/derived/antarctic_ice_shelves_nsidc.csv"))
-
-hindcast_clim |> 
+#| fig-subcap:
+#|   - "+S1"
+#|   - "+S2" 
+#| layout-ncol: 1
+
+# ice_shelves_nsicd <- fread(here::here("data/derived/antarctic_ice_shelves_nsidc.csv"))
+
+s <- lapply(c("S1", "S2"), \(x) {
+  g <- hindcast_clim |>
+    _[model == x] |>
     # dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
@@
     geom_antarctica_fill +
-  facet_wrap(~ month, labeller = labeller(month = labels_month)) +
+    facet_wrap(~ month + model , labeller = labeller(month = labels_month)) +
     wide_legend +
     tag_facets()
+
+  print(g)
+
+})
+
 ```
@@
 ```{r fig-extent-anom}
-#| fig-cap: Monthly mean sea ice extent anomalies for +S1 and +S2 and observations (black) forecasted at selected lead times.
-The RMSE during the overlapping period (1990--2013) is shown on the top left.
+#| fig-cap: Monthly mean sea ice extent anomalies of the observation (black) and forecasts from +S1 (right column) and ACCESS-S2 (left column) at lead times of 0, 2, 4, and 6 months.
+The RMSE and correlation coefficient and their 95% confidence intervals during the overlapping period of ACCESS-S1 and S2 hindcasts (1990–2013) is shown on the top left and lower right corner of each panel.
 
@@
   _[, aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE),
-    by = .(month(time), dataset)]
+    by = .(month(time), dataset)] |>
+  _[]
 
@@
 
-hindcast_extent |> 
+hindast_extent_monthly <- hindcast_extent |> 
   copy() |> 
@@
     by = .(lag, model, time = update(time, year = 2000))] |>
-  _[, .(forecast = mean(aice)), by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
-  _[, lag := months_difference(time, forecast_time)] |> 
+  _[, .(forecast = mean(aice)),
+    by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
+  _[, lag := months_difference(time, forecast_time)]
+
+hindast_extent_monthly |> 
   _[, merge(.SD, monthly_extent, all = TRUE), by = .(model, lag)] |> 
@@
   geom_text(data = \(x) x[time %between% s1_range, 
-                          .(text = sqrt(mean((aice - forecast)^2, na.rm = TRUE))), 
+                          rmse(aice, forecast, signif = 2, scale = 1e12),
                           by = .(lag, model)],
-            aes(label = paste0(scales::number(text, scale = 1e-12), " M km²")), x = -Inf, y = Inf,
-            vjust = 1.2, hjust = -0.2) +
+            aes(label = text),
+            x = -Inf, y = Inf,
+            vjust = 1.2, hjust = 0) +
+
+  geom_text(data = \(x) x[time %between% s1_range, 
+                          correlate(aice, forecast),
+                          by = .(lag, model)],
+            aes(label = text),
+            vjust = 0, hjust = 1.2,
+            x = Inf, y = -Inf) +
+
   scale_y_continuous(NULL,
@@
   facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Month: ", x))) +
-  scale_color_models
+  scale_color_models +
+  scale_fill_models
 ```
@@
 @fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lead times. 
-Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at a lead time of 3 months.
-+S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.
+Compared with +S1, +S2 anomaly forecast is relatively poor (large RMSE) even for the first month (lead time 0), whereas +S1 has lower RMSE at 6 month lead time than +S2 has at zero lead time. 
++S2 shows much larger interannual variability than observations --particularly at short lead times-- with dramatic lows between 1995 and 2007, and highs between 2007 and 2015.
+
+Strangely, for +S2, RMSE improves with lead time, even though the correlation degrades with lead time. This is puzzling behaviour that goes contrary to what is usually seen in prediction models. 
+The explanation seems to be the mentioned increased interannual variability. 
+@fig-extent-sd shows monthly mean sea ice extent of the forecasts as a function of lead time compared with observations. 
++S1 standard deviation lies within the observed regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at 9 month lead time.
+
+```{r fig-exent-sd}
+#| fig-cap: Standard deviation and 95% interval of monthly mean sea ice extent of observations and forecast for +S1 and +S2 at different lead times.
+
+obs_sd <- monthly_extent[time %between% s1_range,
+                                   sd_ci(aice, na.rm = TRUE),
+                                   by = .(month(time), dataset)] |> 
+  _[, .(dataset, month, obs_sd = estimate)]
+
+hindcast_extent_sd <- hindast_extent_monthly[time %between% s1_range, 
+                       sd_ci(forecast, na.rm = TRUE), 
+                       by = .(model, lag, month(time))]
+
+hindcast_extent_sd |> 
+  _[obs_sd, on = .NATURAL] |> 
+  ggplot(aes(lag, estimate / obs_sd)) +
+  # geom_rect(data = obs_sd, inherit.aes = FALSE,
+  #            aes(ymin = low, ymax = high,  fill = dataset, xmin = -Inf, xmax = Inf),
+  #               alpha = 0.5) +
+  # geom_hline(data = obs_sd, aes(yintercept = estimate, color = dataset)) +
+  geom_ribbon(aes(ymin = low/obs_sd, ymax = high/obs_sd, fill = model), alpha = 0.5) +
+  geom_line(aes(color = model)) +
+
+  scale_y_continuous("Standardised sea ice extent standard deviation") +
+  scale_x_continuous("Lead time", breaks = c(0:9), expand = c(0, 0), 
+                     trans = scales::reverse_trans()) +
+  scale_color_models +
+  scale_fill_models +
+  facet_wrap(~ month, labeller = labeller(month = labels_month))
+
+```
+
+```{r forecast_example}
+mlag <- 0
+sample_date <- "2008-05-02"
+forecast_date <- as.Date(sample_date) 
+month(forecast_date) <- month(sample_date) - mlag
+mday(forecast_date) <- 1
+
+aice_forecast_example <- c(lapply(c("S1", "S2"), \(model) {
+  month_clim <- formatC(month(forecast_date), width = 2, flag = "0")
+
+  hindcast(forecast_date, model, 1) |> 
+    cdo_ydaysub(here::here(glue::glue("data/derived/climatology/{model}/{month_clim}/em.nc"))) |>
+    cdo_execute() |> 
+    ReadNetCDF("aice", subset = list(time = sample_date)) |> 
+    _[, model := model]
+  }),
+  list(CDR() |> 
+       anomalies() |>
+       ReadNetCDF("aice", subset = list(time = sample_date)) |> 
+       _[, model := "cdr"])) |> 
+  rbindlist()
+
+thickness_example <- c(S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/hi/daily/e01/di_hi_{format(forecast_date, '%Y%m%d')}_e01.nc",
+  S1 = "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/e01/di_ice_{format(forecast_date, '%Y%m%d')}_e01.nc") |> 
+  purrr::imap(\(f, model) {
+    f |> 
+      glue::glue() |> 
+      cdo_selname("hi") |> 
+      remap_cdr() |> 
+      cdo_execute(cache = TRUE) |> 
+      ReadNetCDF("hi", subset = list(time = sample_date))
+
+  }) |> 
+  rbindlist(idcol = "model")
+
+```
+
++S2 forecasts of sea-ice extent anomalies seem to align moderately well align with observations (leading to moderately high correlation) but their magnitude is overestimated (leading to large errors).
+This suggests that +S2 sea-ice is much more sensitive to atmospheric and oceanic forcing, perhaps due to different thickness. 
+@fig-forecast-example shows sea-ice concentration anomalies (top row) and sea-ice thickness and the difference between the two models (bottom row) for the `r format(as.Date(sample_date), "%e of %B %Y")` initialised one day prior; being that close to initialisation date, these are very approximately the initial conditions.
++S1 sea-ice concentrations anomalies are very close to observations as expected from the system assimilating these data.
++S2 sea-ice concentration anomalies, which are not assimilated, are not as close, but the general pattern is not terrible.
+The system simulates large positive anomalies in the Weddell and Ross seas and slight negative anomalies in the Amundsen and Bellingshausen seas. 
+The fact that +S2 can simulate this pattern without assimilating sea-ice data suggests that atmospheric and oceanic forcing was the dominant factor. 
+However, the magnitude of these anomalies is too big. 
+As shown in @fig-forecast-example bottom row, +S2's sea-ice was thinner than +S1's, even in regions of large positive se-ice concentration anomalies.
+
+```{r fig-forecast-example}
+#| fig.cap: +S1 and +S2 hindcasts for the 2 of May 2008.
+Top row shows sea-ice concentration anomalies forecasted by each system and the observations.
+Bottom row shows forecasted sea-ice thickness and the difference between +S1 and +S2.
+#| fig.height: 7
+#| fig.width: 5.5
+b <- seq(-0.5, 0.5, by = 0.1)
+b <- b[b!= 0]
+
+aice_forecast_example |> 
+  ggplot(aes(xgrid, ygrid)) +
+  geom_contour_fill(aes(z = aice, fill = ..level..),
+                    breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
+  scale_fill_divergent_discretised("Sea-ice concentration anomaly",
+                                   low = trans_pink,
+                                   high = trans_blue) +
+  geom_antarctica_fill + 
+  geomcoord_antarctica +
+  facet_wrap(~model, labeller = labeller(model = labels_models))  +
+  wide_legend  +
+
+thickness_example |> 
+  _[, model := factor(model, levels = c("S1", "S2", "Difference"))] |> 
+  ggplot(aes(xgrid, ygrid)) +
+  geom_contour_fill(aes(z = hi, fill = ..level..), 
+                    breaks = c(AnchorBreaks(exclude = 0, binwidth = 0.25)(c(0, 2)), Inf)) +
+  # as.discretised_scale(scale_fill_viridis_c)("Sea ice thickness", 
+  #                                            guide = guide_colorsteps(order = 9)) + 
+  scale_fill_divergent_discretised("Sea-ice thickness",
+                                   high = scales::muted("#5ab4ac"),
+                                   guide = guide_colorsteps(order = 9)) +
+  ggnewscale::new_scale_fill() +
+
+  geom_contour_fill(data = \(x) x |> 
+                      dcast(time + xgrid + ygrid ~ model, value.var = "hi") |> 
+                      _[, model := factor("Difference", levels = c("S1", "S2", "Difference"))], 
+                    aes(z = S1 - S2, fill = ..level..), breaks = c(-Inf, b, Inf)) +
+
+  scale_fill_divergent_discretised("Sea-ice thickness difference",
+                                   guide = guide_colorsteps(order = 0)) +
+  geom_antarctica_fill + 
+  geomcoord_antarctica +
+  wide_legend +
+  facet_wrap(~ model) +
+  plot_layout(ncol = 1)
+
+```
+
+The fact that +S1 and +S2 share the same model configuration and that the increased variability is more extreme at short lead times (@fig-exent-sd) suggests that this is an effect of the data assimilation.
+It is possible that sea-ice in the +S2 system is left in an unbalanced state after assimilating atmospheric and oceanic data but not sea-ice data, leading to large responses in the initial states which then subside at longer lead times when the model is balanced.
+
+```{r}
+giomas <- here::here("data/derived/giomas.csv") |> 
+  fread() |> 
+  _[, time := as.Date(time)] |> 
+  _[, .(thickness = mean(thickness)), by = month(time)]
+```
+
+```{r}
+
+mean_thickness <- function(concentration, thickness, cell_area = NULL) {
+  if (is.null(cell_area)) {
+    cell_area <- concentration |> 
+      rcdo::cdo_gridarea()
+  }
+
+  volume <- thickness |>
+    rcdo::cdo_mul(concentration) |>
+    rcdo::cdo_mul(cell_area) |> 
+    rcdo::cdo_fldsum()
+
+  area <- concentration |>
+    rcdo::cdo_mul(cell_area) |> 
+    rcdo::cdo_fldsum()
+
+  rcdo::cdo_div(volume, area) |>
+    rcdo::cdo_execute(options = "-L")
+}
+
+mean_thickness_access <- \(forecast_time, model) {
+  concentration <- hindcast(forecast_time, model = model, members = "em", variable = "aice")
+  thickness <- hindcast(forecast_time, model = model, members = "em", variable = "hi")
+
+  if (length(thickness) == 0) return(NA_character_)
+  if (length(concentration) == 0) return(NA_character_)
+
+  rcdo::cdo_cache_set(here::here("data/temp/cache/"))
+  cell_area <- thickness |> 
+      rcdo::cdo_gridarea()
+
+  mean_thickness(concentration, thickness, cell_area)
+}
+
+```
+
+```{r}
+
+get_dir_part <- function(path, n = 1) {
+  if (n == 1) return(basename(path))
+
+  return(get_dir_part(dirname(path), n = n - 1))
+}
+
+get_model <- function(path) {
+  get_dir_part(path, 8)
+}
+
+file <- "sithick_SImon_MCM-UA-1-0_historical_r1i1p1f1_gn_185001-201412.nc"
+dates <- function(path) {
+  file <- get_dir_part(path, 1)
+  l <- 6+1+6+3
+  n <- nchar(file)
+  s <- n-l+1
+
+  date_start <- substr(file, start = s, s + 5)
+  date_end <- substr(file, start = s+7, n-3)
+  list(start = lubridate::ym(date_start),
+       end = lubridate::ym(date_end))
+}
+
+thick <- "/g/data/oi10/replicas/CMIP6/CMIP/*/*/historical/r1i1p1f1/SImon/sithick/gn/*/*" |> 
+  Sys.glob()
+
+thick <- data.table(thickness = thick) |> 
+    _[, model := get_model(thickness)] |> 
+    _[, c("start", "end") := dates(thickness)] |> 
+    _[]
+
+conc <- "/g/data/oi10/replicas/CMIP6/CMIP/*/*/historical/r1i1p1f1/SImon/siconc/gn/*/*" |> 
+  Sys.glob()
+
+conc <- data.table(concentration = conc) |> 
+  _[, model := get_model(concentration)] |> 
+  _[, c("start", "end") := dates(concentration)]
+
+cmip <- merge(conc, thick, by = c("model", "start", "end"))
+
+```
+
+```{r}
+hindcast_thickness <- CJ(model = c("S2", "S1")) |>
+  _[, .(forecast_time = get_forecast_times(model)), by = model] |> 
+  _[mday(forecast_time) == 1] |> 
+  _[, thickness := unlist(mirai_map(data.frame(forecast_time, model), 
+                                    mean_thickness_access,
+                                    hindcast = hindcast, mean_thickness = mean_thickness)[.progress])] |> 
+  na.omit()
+
+```
+
+```{r}
+giomas_thickness <- "data/raw/giomas_thickness.nc" |> 
+  cdo_selindexbox(1, 360, 1, 71) |> 
+  cdo_setvrange(0, 500)
+
+giomas_concentration <- "data/raw/giomas_concentration.nc" |> 
+  cdo_selindexbox(1, 360, 1, 71) |> 
+  cdo_setvrange(0, 2)
+
+giomas_cellarea <- "data/raw/giomas_cell-area.nc" |> 
+  cdo_selindexbox(1, 360, 1, 71)
+
+giomas <- mean_thickness(giomas_concentration, giomas_thickness, giomas_cellarea) |> 
+  cdo_ymonmean() |> 
+  cdo_execute() |> 
+  ReadNetCDF(c(thickness = "heff")) |> 
+  _[, month := 1:12]
+
+```
+```{r}
+hindcast_mean_thickness <- hindcast_thickness |> 
+  _[, thickness |> cdo_monmean() |> cdo_execute(), by = .(model, forecast_time)] |>
+  _[, metR::ReadNetCDF(V1, c(thickness = "hi")), by = .(model, forecast_time)] |> 
+  _[, time := as.Date(update(time, day = 1))] |> 
+  _[, lag :=  months_difference(time, forecast_time)] 
+```
+
+```{r}
+cmip |> 
+  _[model %in% unique(model)[1:10]] |> 
+  _[, mean_thickness(cdo_sellonlatbox(concentration, 0, 360, -90, -40),
+                     cdo_sellonlatbox(thickness, 0, 360, -90, -40)), 
+            by = .(model, start, end)] |> 
+    _[, .(thickness = V1 |> cdo_mergetime() |> cdo_execute()), by = .(model)] |> 
+    _[, ReadNetCDF(thickness, "sithick")[, .(time, sithick)], by = .(model)] |> 
+    _[, .(sithick = mean(sithick)), by = .(model, month(time))] |> 
+    ggplot(aes(month, sithick)) +
+    geom_line(aes(group = model, color = "CMIP6 models")) +
+    geom_line(data = giomas, aes(y = thickness, color = "GIOMAS"), linewidth = 1.4) +
+    scale_color_manual(values = c("CMIP6 models" = "black", GIOMAS =  "#1a7e21")) +
+    scale_y_continuous("Mean sea-ice thickness") +
+    scale_x_continuous(NULL, breaks = 1:12, labels = labels_month)
+```
+
+```{r}
+hindcast_mean_thickness[time %between% s1_range,
+                          average(thickness), 
+                          keyby = .(lag, model, month(time))] |> 
+  ggplot(aes(lag, estimate)) +
+  geom_ribbon(aes(ymin = low, ymax = high, fill = model), alpha = 0.5) +
+  geom_line(aes(color = model)) +
+  # geom_smooth(method = "lm", aes(color = model)) +
+  scale_color_models +
+  scale_fill_models +
+  scale_y_continuous("Mean sea-ice thickness") +
+  scale_x_continuous("Lead time", breaks = c(0:9), expand = c(0, 0), 
+                     trans = scales::reverse_trans()) +
+  facet_wrap(~ month, labeller = labeller(month = labels_month))
+
+```
+
+```{r}
+hindcast_mean_thickness |> 
+  _[, .(thickness = mean(thickness)), by = .(model, lag, month(time))] |> 
+  _[lag %in% c(0)] |> 
+  ggplot(aes(month, thickness)) + 
+  geom_line(aes(color = model, group = interaction(model, lag))) +
+  geom_line(data = giomas) +
+  scale_color_models +
+  scale_y_continuous("Mean sea-ice thickness") +
+  scale_x_continuous(NULL, breaks = 1:12, labels = labels_month) 
+```
 
@@
   _[, lag := as.numeric(time - time_forecast)] |> 
+  _[version == "persistence", member := "0em"] |>
+  _[member == "0em"] |> 
   _[lag > 0] |> 
@@
 clim_std <- datasets |> 
+  _[1] |> 
   lapply(\(x) x |> 
-           anomalies() |> 
+           anomalies(climatology(x)) |> 
            cdo_sqr() |> 
@@
 
+```{r t_test}
+t_test <- function(x, y, ...) {
+
+  test <- t.test(x, y, ...)
+
+  list(low = test[["conf.int"]][1],
+       high = test[["conf.int"]][2], 
+       estimate = -diff(test$estimate), 
+       p.value = test[["p.value"]]
+  )
+}
+
+max_lag <- errors[version == "S1", max(lag)]
+
+dif <- errors |> 
+  na.omit() |> 
+  _[obs_dataset == which_dataset] |>
+  _[measure == "rmse"] |> 
+  _[version != "persistence"] |> 
+  _[lag <= max_lag] |> 
+  dcast(time + time_forecast + member + lag ~ version, value.var = "value") |> 
+  _[, t_test(na.omit(S1), na.omit(S2)), 
+    by = .(lag, month(time_forecast))] |> 
+  _[p.value > 0.01] |> 
+  _[, .SD[which.min(lag)], by = month] |> 
+  _[order(month), .(month, lag)]
+
+labels_month_significant <- setNames(paste0(month.abb, " (", dif$lag, ")"),
+                                     1:12)
+```
+
 ```{r fig-rmse}
 #| fig-cap: Median RMSE of sea ice concentration anomalies as a function of forecast lead time for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies (black) and climatology (gray).
+Only the first 120 days are shown.
 #| fig.height: 5
@@
 
-clim_std2 |> 
+errors2 <- clim_std2 |> 
   rbind(errors) |>
@@
   _[obs_dataset ==  which_dataset] |> 
-  _[, median_ci(value),
-    by = .(lag, version, measure, month(time_forecast), obs_dataset)] |>
-  ggplot(aes(lag, mid)) +
+  _[, average(value),
+    by = .(lag, version, measure, month(time_forecast), obs_dataset)]
+
+errors2 |> 
+  _[lag <= 120] |> 
+  ggplot(aes(lag, estimate)) +
 
@@
   #             alpha = 0.1) +
-  geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
+  geom_line(aes(color = version, group = interaction(obs_dataset, version))) +
 
@@
   scale_fill_models +
-  facet_wrap(~ month, labeller = labeller(month = labels_month)) 
+  facet_wrap(~ month, labeller = labeller(month = labels_month_significant)) +
+  coord_cartesian(clip = "off")
 ```
@@
 
+```{r lead-time-window}
+#| fig-cap: Minimum lead time at which each forecast's mean RMSE becomes lower than the 95% confidence interval of persistence forecast RMSE (black lines) and maximum lead time at which each forecast's mean RMSE remains lower than the 95% confidence interval of climatological forecast RMSE (gray lines).
+Green shading indicates the window where forecasts outperform both persistence (lead times longer than black line) and climatology (lead times shorter than gray line).
+Text labels show the date corresponding to the maximum lead time at which each forecast outperforms climatology.
+
+rbind(errors2 |> 
+  copy() |> 
+    _[obs_dataset ==  which_dataset] |> 
+  _[, base := low[version == "persistence"], by = .(lag, measure, month, obs_dataset)] |> 
+  _[version != "persistence"] |> 
+  _[version != "climatology"] |> 
+  _[estimate < base] |> 
+  _[, .SD[which.min(lag)], by = .(version, measure, month, obs_dataset)] |> 
+  _[, .(version, month, obs_dataset, lag, benchmark = "persistence")],
+
+errors2 |> 
+  copy() |> 
+  _[obs_dataset ==  which_dataset] |> 
+  _[, base := low[version == "climatology"], by = .(lag, measure, month, obs_dataset)] |> 
+  _[version != "climatology"] |> 
+  _[version != "persistence"] |> 
+  # _[lag > 2] |> 
+  _[, rle(estimate < base), by= .(version, measure, month, obs_dataset)] |> 
+  _[, .SD[1], by = .(version, measure, month, obs_dataset)] |> 
+  tidyr::complete(version, measure, month, obs_dataset, values, fill = list(lengths = 0)) |> 
+  as.data.table() |> 
+  _[values == TRUE] |> 
+  _[, .(version, month, obs_dataset, lag = lengths, benchmark = "climatology")] 
+) |> 
+  _[, time := make_date(2000, month, 1) + lag] |> 
+  # _[benchmark == "climatology" & version == "S1"] |> 
+  ggplot(aes(month, lag)) +
+  ggbraid::geom_braid(data = \(x) dcast(x, month + lag + version ~ benchmark, value.var = "lag"),
+                      alpha = 0.3,
+                      aes(ymin = persistence, ymax = climatology,
+                          fill = climatology > persistence)) +
+  geom_line(aes(color = benchmark)) +
+  # geom_point(aes(color = time), data = \(x) x[lag != 0]) +
+  geom_text(data = \(x) x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |> 
+              _[benchmark == "climatology"],
+            aes(label = format(time, "%d %b")),
+            vjust = -0.4) +
+  geom_point(data = \(x) x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |> 
+               _[benchmark == "climatology"],
+             aes(color = benchmark)) +
+  scale_color_models +
+  scale_y_continuous("Lead time (days)") +
+  scale_x_continuous("Forecast month", breaks = 1:12, labels = month.abb, minor_breaks = NULL) +
+  scale_fill_manual(guide = "none", values = c(`TRUE` = "seagreen", `FALSE` = "#FFFFFF00")) +
+  facet_wrap(~ version, ncol = 1, labeller = labeller(version = labels_models)) +
+  coord_cartesian(clip = "off") +
+  tag_facets()
+
+```
+
 ```{r topo_lonlat}
@@
 
+```{r}
+clim_std_lon <- datasets |> 
+  # _[1] |> 
+  lapply(\(x) x |> 
+           anomalies(climatology(x)) |> 
+           cdo_sqr() |> 
+           cdo_remapmean("r15x360") |>
+           cdo_mermean() |>
+           cdo_sqrt() |> 
+           cdo_execute() |> 
+           ReadNetCDF(c(climatology = "aice")) |> 
+           _[, let(lat = NULL, lon = lon + 12)]
+  ) |> 
+  rbindlist(idcol = "obs_dataset") |> 
+  _[, time := as.Date(time)] |> 
+  _[climatology == 0, climatology := NA] |> 
+  _[climatology < 0.05, climatology := NA]
+
+```
+
 ```{r rmse_lon_mean}
@@
   readRDS() |> 
-  _[version != "S1"] |> 
+  _[version == "persistence", member := "0em"] |> 
+  _[member == "0em"] |> 
+  _[, lon := lon + 12] |> 
+  _[, time := update(time, hour = 0)] |> 
+  _[, time := as.Date(time)]
+
+# base_rmse_lon <- clim_std_lon[rmse_lon_mean[version == "persistence"], on = .NATURAL] |> 
+#   _[, lag := as.numeric(as.Date(time) - time_forecast)] |> 
+#   _[, .(climatology = mean(climatology, na.rm = TRUE),
+#         value = mean(value, na.rm = TRUE)),
+#     by = .(lag, lon, obs_dataset, month = factor(month(time_forecast)))] |> 
+#   _[, base := climatology] |>
+#   _[, let(climatology = NULL, value = NULL)] |>
+#   _[]
+```
+
+```{r rmse_lon_mean2}
+# rmse_lon_mean <- rmse_lon_mean |> 
+#   copy() |> 
+#   # _[version != "S2"] |> 
+#   _[version != "persistence"] |> 
+#   _[member == "0em"] |> 
+#   _[base_rmse_lon, on = .NATURAL] |> 
+#   _[, lag := as.numeric(as.Date(time) - time_forecast)] |> 
+#   _[, .(dif = mean(value - base, na.rm = TRUE)),
+#     by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))]
+```
+
+```{r fig-rmse_lon, fig.height=7, fig.width=10}
+#| fig-cap: !expr glue::glue"Median difference between RMSE of +S2 forecasts and persistence forecast computed on {uniqueN(rmse_lon_mean$lon)} meridional slices {diff(unique(rmse_lon_mean$lon))[1]}º wide.
+Antarctica's coastline is shown at the bottom of each panel for reference.")
+#| fig-subcap: 
+#|  - ACCESS-S1
+#|  - ACCESS-S2
+
+binwidth <- 0.015
+gdata <- rmse_lon_mean |> 
+    _[obs_dataset == which_dataset] |> 
+    _[version != "persistence"] |> 
     _[, lag := as.numeric(as.Date(time) - time_forecast)] |> 
-  _[, base := value[version == "persistence"], 
-    by = .(time, lon, lag, obs_dataset)] |> 
-  _[version != "persistence"] |>
-  _[, .(dif = median(value - base, na.rm = TRUE)),
+    clim_std_lon[i = _, on = .NATURAL] |>
+    _[, .(dif = mean(value - climatology, na.rm = TRUE)),
       by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))] 
@@
 ```{r fig-rmse_lon, fig.height=7, fig.width=10}
-#| fig.cap: Median difference between RMSE of +S2 forecasts and persistence forecast computed on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]`º wide.
-Antarctica's coastline is shown at the bottom of each panel for reference.
-
-binwidth <- 0.015
-
-rmse_lon_mean |> 
-  _[obs_dataset == which_dataset] |> 
-  ggperiodic::periodic(lon = c(0, 360)) |> 
+lapply(c("S1", "S2"), \(x) {
+
+  g <- gdata |> 
+    _[version == x] |> 
+    _[, dif := Smooth2D(lag, lon, dif, method = smooth_svd(.1)), by = .(month, version)] |>
     ggplot(aes(lon, lag)) +
     geom_contour_fill(aes(z = dif, fill = after_stat(level)),
-                    breaks = AnchorBreaks(0, exclude = 0, binwidth = binwidth))  +
-  geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.4),
-                      breaks = AnchorBreaks(0, exclude = 0, binwidth = binwidth)) +
+                      breaks = AnchorBreaks(0, binwidth = binwidth))  +
+    geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.2),
+                        breaks = AnchorBreaks(0, binwidth = binwidth)) +
 
@@
     tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1))
+
+  print(g)
+}) -> sink
 ```
@@
 
-### Comparison with S1
-
-```{r t_test}
-t_test <- function(x, y, ...) {
-
-  test <- t.test(x, y, ...)
-
-  list(low = test[["conf.int"]][1],
-       high = test[["conf.int"]][2], 
-       estimate = -diff(test$estimate), 
-       p.value = test[["p.value"]]
-  )
-}
-
-max_lag <- errors[version == "S1", max(lag)]
-
-dif <- errors |> 
-  na.omit() |> 
+<!-- ```{r errors_ensemble} -->
+<!-- errors_ensemble <- readRDS(here::here("data/derived/rmse.Rds")) |>  -->
+<!--   _[version != "persistence"] |>  -->
+<!--   _[member != "0em"] |>  -->
+<!--   _[obs_dataset == which_dataset] |>  -->
+<!--   _[value < 0.01, value := NA] |>  -->
+<!--   _[value > 3e13, value := NA] |>  -->
+<!--   _[!is.na(value)] |>  -->
+<!--   _[, time := as.Date(time)] |>  -->
+<!--   _[, lag := as.numeric(time - time_forecast)] |>  -->
+<!--   _[lag > 0] |>  -->
+<!--   _[!(month(time_forecast) == 1 & member == "07" & version == "S1")] -->
+<!-- ``` -->
+
+<!-- ```{r fig-spread-errors} -->
+<!-- #| fig-cap: Total spread in RMSE, including ensemble spread and forecast spread, as a function of forecast lead time for ACCESS-S1 and ACCESS-S2 hindcasts.  -->
+
+<!-- errors_ensemble |>  -->
+<!--   _[, var(value, na.rm = TRUE),  -->
+<!--     by = .(lag, version, measure, obs_dataset, time_forecast))] |>  -->
+<!--   ggplot(aes(lag, V1)) + -->
+<!--   geom_line(aes(color = version)) + -->
+<!--   scale_color_models + -->
+<!--   scale_y_continuous("RSME ensemble variance") + -->
+<!--   facet_wrap(~ month, labeller = labeller(month = labels_month))  -->
+<!-- ``` -->
+
+```{r fig-initial-spread}
+#| fig-cap: Decomposition of forecast error spread at one-day lead time for ACCESS-S1 and ACCESS-S2 hindcasts across initialization months.
+The top panel shows the mean variance of RMSE errors across ensemble members, while the bottom panel shows the variance of the ensemble mean RMSE error.
+
+which_lags <- c(1, 3, 30)
+rbind(mean_variance = errors_ensemble |>
+        _[lag %in% which_lags] |> 
+        _[measure == "rmse"] |>
         _[obs_dataset == which_dataset] |>
-  _[measure == "iiee"] |> 
-  _[version != "persistence"] |> 
-  _[lag <= max_lag] |> 
-  dcast(time + time_forecast + member + lag ~ version, value.var = "value") |> 
-  _[, t_test(na.omit(S1), na.omit(S2)), 
-    by = .(lag, month(time_forecast))] |> 
-  _[p.value > 0.01] |> 
-  _[, .SD[which.min(lag)], by = month] |> 
-  _[order(month), .(month, lag)]
-
-labels_month_significant <- setNames(paste0(month.abb, " (", dif$lag, ")"),
-                                     1:12)
-```
+        _[, var(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
+        _[, .(value = mean(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
+
+      variance_of_mean = clim_std2 |> 
+        rbind(errors) |>
+        _[measure == "rmse"] |> 
+        _[lag %in% which_lags] |> 
+        _[measure == "rmse"] |>
+        # _[version != "persistence"] |> 
+        _[obs_dataset == which_dataset] |>
+        _[, .(value = var(value, na.rm = TRUE)),
+          by = .(version, 
+                 month(time_forecast),
+                 lag)],
+      idcol = "component") |> 
+  ggplot(aes(month, sqrt(value))) +
+  geom_line(aes(color = version), linewidth = 1) +
+  scale_y_continuous("Standard deviation") +
+  # scale_color_models +
+  scale_x_continuous("Month", breaks = 1:12, labels = month.abb, ) +
+  scale_color_models +
+  facet_grid(lag ~ component,
+             labeller = labeller(component = c(mean_variance = "Mean standard deviation of errors in ensemble members",
+                                               variance_of_mean = "Standard deviation of ensemble mean error."))) +
+  tag_facets(tag = "rc")
+```
+
+Also evident in @fig-initial-spread is the difference in the error spread at short lead times between +S2 and +S1.
+In all months +S1 has much narrower error spread at the first day of the forecasts.
+The error in the initial conditions is not only small, but also fairly constant.
+This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
+For +S2, this is true only between July and October.
+For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it is not constant.
 
@@
     by = .(lag, version, measure, obs_dataset, month(time_forecast))] |> 
-  ggplot(aes(lag, mid)) +
+  ggplot(aes(lag, estimate)) +
 
@@
   #              aes(x = lag, 
-  #                  ymin = mid - (high - low), ymax = mid + (high - low))) +
+  #                  ymin = estimate - (high - low), ymax = estimate + (high - low))) +
 
   # geom_point(data = \(x) x[dif, on = .NATURAL][version == "S1"],
-  # aes(lag, mid), size = 3, shape = 21) +
+  # aes(lag, estimate), size = 3, shape = 21) +
 

### 2025-10-06 R/datasets.R

@@ -220,57 +220,15 @@ S2_reanalysis <- function() {
     rcdo::cdo_execute(output = file, options = "-L")
 }
 
-hindcast <- function(forecast_times, model, members = 1:9) {
-  if (model == "S2") {
-    S2_hindcast(forecast_times, members)
-  } else {
-    S1_hindcast(forecast_times, members)
-  }
-}
-
-S2_hindcast <- function(forecast_times, members = 1:9) {
+hindcast <- function(forecast_times, model, members = 1:9, variable = "aice") {
+  members <- ifelse(members == "em", members, formatC(as.numeric(members), width = 2, flag = "0"))
+  
   data <- data.table::CJ(forecast_time = forecast_times, member = members)
   
-  raw_files <- here::here("data/derived/hindcast/S2/",
+  raw_files <- here::here("data/derived/hindcast", variable, model, 
                           data$forecast_time, 
                           paste0(formatC(data$member, width = 2, flag = "0"), ".nc"))
   raw_files[file.exists(raw_files)]
-}
-
-
-S1_hindcast <- function(forecast_times, members = 1:9) {
-  data <- data.table::CJ(forecast_time = forecast_times, member = members)
-  
-  raw_files <- paste0(
-    "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/e0",
-    data$member,
-    "/di_ice_",
-    format(as.Date(data$forecast_time), "%Y%m%d"),
-    "_e0",
-    data$member,
-    ".nc"
-  )
-  
-  exists <- file.exists(raw_files)
-  stopifnot(all(exists))
-  
-  files <- here::here("data/derived/hindcast/S1/", 
-                      data$forecast_time,
-                      paste0(formatC(data$member, width = 2, flag = "0"), ".nc"))
-  
-  furrr::future_map_chr(seq_along(files), \(f) {
-    if (file.exists(files[f])) {
-      return(files[f])
-    }
-    
-    dir.create(dirname(files[f]), showWarnings = FALSE, recursive = TRUE)
-    raw_files[f] |>
-      rcdo::cdo_selname("aice") |>
-      remap_cdr() |>
-      cdo_del29feb() |>
-      rcdo::cdo_execute(output = files[f], options = "-L")
-  })
-  files
 }
 
 OSI <- function() {

### 2025-10-06 R/functions.R

@@ -1,23 +1,33 @@
 median_ci <- function(x, ci = .95) {
-  quantile(na.omit(x),  probs = c(0.5 - ci/2, 0.5, 0.5 + ci/2), names = FALSE) |> 
-    setNames(c("low", "mid", "high")) |> 
+  
+  # list(mid = mean(x, na.rm = TRUE))
+  quantile(na.omit(x),  probs = c(0.5 - ci/2, 0.5, 0.5 + ci/2), names = FALSE) |>
+    setNames(c("low", "estimate", "high")) |>
     as.list()
 }
 
+average <- function(x) {
+  test <- t.test(x)
+  list(estimate = test$estimate,
+       low = test$conf.int[1],
+       high = test$conf.int[2])
+}
+
+
 on_gadi <- function() {
-    as.numeric(Sys.getenv("ON_GADI", unset = "0")) == 1
+  as.numeric(Sys.getenv("ON_GADI", unset = "0")) == 1
 }
 
 ncview <- function(file) {
-    system2("ncview", args = c(file), wait = FALSE)
+  system2("ncview", args = c(file), wait = FALSE)
 }
 
 get_forecast_times <- function(model) {
   
   dir <- switch(model,
-    S1 =  "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/",
-    S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/",
-    stop("Model needs to be S1 or S2")
+                S1 =  "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/",
+                S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/",
+                stop("Model needs to be S1 or S2")
   )
   
   dates <- list.files(file.path(dir, "e01")) |>
@@ -27,7 +37,8 @@ get_forecast_times <- function(model) {
     ) |> 
     _$date |> 
     lubridate::as_date()
-    
+  
+  
   dates
 }
 
@@ -37,49 +48,49 @@ compute_climatology <- function(dataset,
                                 runmean_number = 11,
                                 year = 2000,
                                 options = c("-L")) {
-    dates <- c(paste0(min(climatology), "-01-01"), paste0(max(climatology), "-12-31"))
-
-    noisy <- dataset |>
-        rcdo::cdo_seldate(
-            startdate = dates[1],
-            enddate = dates[2]
-        ) |>
[…]

### 2025-10-06 R/ggplot.R

@@ -25,8 +25,9 @@ ggplot2::theme_set(ggplot2::theme_minimal() +
                      ))
 wide_legend <- ggplot2::theme(legend.key.width = grid::unit(1, 'null'))
 
-trans_pink <- "#D07C8E"
-trans_blue <- "#0AA3CE"
+
+trans_pink <- scales::muted("#F7A8B8", l = 60, c = 100)
+trans_blue <- scales::muted("#55CDFC", l = 60, c = 100)
 
 
 

### 2025-10-06 draft/lag_correlation.R

@@ -1,0 +1,34 @@
+
+lags <- c(5, 10, 15)
+lags <- setNames(lags, lags)
+
+lag <- 30
+file <- anomalies(datasets$cdr)
+autocor <- function(file, lag) {
+  remove <- "timestep=1/{lag}/1" |> 
+    glue::glue()
+  
+  shift <- "-{lag}day" |> 
+    glue::glue()
+  
+  out <- file |> 
+    cdo_delete(timestep = remove) |>
+    cdo_timcor(file) |> 
+    cdo_execute(tempfile())
+  
+}
+
+
+corlags <- lapply(lags, \(lag) autocor(anomalies(datasets$cdr), lag))
+
+corlags |> 
+  setNames(lags) |> 
+  lapply(ReadNetCDF, c("aice", "pvalue")) |> 
+  rbindlist(idcol = "lag") |> 
+  _[aice <= 1] |> 
+  ggplot(aes(xgrid, ygrid)) +
+  geom_contour_fill(aes(z = aice)) +
+  # geom_point(data = \(x) x[pvalue < 0.5]) +
+  scale_fill_divergent() +
+  coord_equal() +
+  facet_wrap(~ lag)

### 2025-10-06 scripts/build_ensemble.R

@@ -1,7 +1,85 @@
+chunk <- as.numeric(Sys.getenv("CHUNK", 0))
+
 library(rcdo)
 source(here::here("R/datasets.R"))
+cdo_cache_set()
+future::plan("multicore", workers = 40)
 
-future::plan("multicore", workers = 24)
+process_ice <- function(file, shift, file_out, variable) {
+  file |> 
+    rcdo::cdo_selname(variable) |>
+    remap_cdr() |>
+    cdo_shifttime(shiftValue = paste0(shift, "day")) |> 
+    cdo_del29feb() |> 
+    cdo_execute(output = file_out, cache = TRUE, options = "-L")
+}
+
+
+get_file <- function(forecast_time, member, model, variable = c("aice", "hi")) {
+  
+  if (model == "S2") {
+    dir <- glue::glue("/g/data/ux62/access-s2/hindcast/raw_model/ice/{variable}/daily/e0")
+    prefix <- glue::glue("/di_{variable}_")
+  } else if (model == "S1") {
+    dir <- "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/e0"
+    prefix <- "/di_ice_"
+  }
+  
+  paste0(
+    dir,
+    member,
+    prefix,
+    format(as.Date(forecast_time), "%Y%m%d"),
+    "_e0",
+    member,
+    ".nc"
+  )  
+}
+
+
+
+
+build_ensemble <- function(forecasts, nominal_date, model, variable) {
+  folder <- here::here("data/derived/hindcast/", variable, model, nominal_date)
+  dir.create(folder, showWarnings = FALSE, recursive = TRUE)
+  
+  if (model == "S2") {
+    members <- 1:3 
+  } else {
+    members <- 1:9
+  }
+    
+  data <- data.table::CJ(original_forecast_date = forecasts, original_member = members) |> 
+    _[, file := get_file(original_forecast_date, original_member, model, variable = variable)] |> 
+    _[, member := 1:9] |> 
+    _[, forecast_date := as.Date(nominal_date)] |> 
+    _[, file_out := file.path(folder, paste0(formatC(member, width = 2, flag = "0"), ".nc"))] |>
+    _[, process_ice(file, as.numeric(forecast_date - original_forecast_date), file_out, variable),
+      by = .(forecast_date, member, original_forecast_date, original_member)]
+}
+
[…]

### 2025-10-06 scripts/build_ensemble_launch.R

@@ -1,0 +1,38 @@
+template <- r"(
+#!/bin/bash
+
+#PBS -P gb02
+#PBS -q normal
+#PBS -l ncpus=48
+#PBS -l mem=64GB
+#PBS -l walltime=12:00:00
+#PBS -l wd
+#PBS -l jobfs=400GB
+#PBS -o logs/
+#PBS -e logs/
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble{chunk}
+
+# Load module, always specify version number.
+module load R/4.3.1
+module load cdo/2.4.3
+
+# Must include `#PBS -l storage=scratch/ab12+gdata/yz98` if the job
+# needs access to `/scratch/ab12/` and `/g/data/yz98/`. Details on:
+# https://opus.nci.org.au/display/Help/PBS+Directives+Explained
+
+# Run R application
+export CHUNK={chunk}
+Rscript scripts/build_ensemble.R > logs/build_ensemble{chunk}.log
+)"
+
+
+chunks <- 1:10
+
+for (chunk in chunks) {
+  script <- glue::glue(here::here("scripts/pbs/build_ensemble{chunk}.sh"))
+  
+  writeLines(glue::glue(template), script)
+  system(glue::glue("qsub {script}"))
+  
+}

### 2025-10-06 scripts/ensemble_mean.R

@@ -1,0 +1,58 @@
+library(rcdo)
+library(data.table)
+library(metR)
+source(here::here("R/datasets.R"))
+
+future::plan("multicore", workers = 40)
+
+file <- get_forecast_times("S2") |> 
+  Filter(f = \(x) lubridate::day(x) == 1 & lubridate::month(x) == 1) |> 
+  _[2] |> 
+  hindcast(model = "S2", members = 1)
+
+ndays <- as.numeric(cdo_ntime(file) |> 
+                      cdo_execute())
+
+bytes_per_day <- file.size(file)/ndays
+
+compute_ensemble_mean <- function(forecast_time, model, variable) {
+  month <- data.table::month(forecast_time)
+  # the 7th member has some weird jumps and inconsistencies. 
+  if (model == "S1" && month == 1) {
+    members <- c(1:6, 8:9)
+  } else {
+    members <- 1:9
+  }
+  
+  files <- hindcast(forecast_time, model, members, variable)
+  
+  # Some forecast have less than the correct number of days. 
+  # Throw them out
+  size <- files |> 
+    file.size()
+  
+  ndays <- floor(size/bytes_per_day)
+  
+  out <- here::here("data/derived/hindcast", variable, model, forecast_time, "em.nc")
+  
+  files[ndays >= 215] |> 
+    cdo_ensmean() |> 
+    cdo_execute(out, options = "-L -O", cache = TRUE)
+}
+
+
+S2_times <- get_forecast_times("S2") |> 
+  Filter(f = \(x) mday(x) == 1) |> 
+  _[-1]
+
+
+S1_times <- get_forecast_times("S1") |> 
+  Filter(f = \(x) mday(x) == 1) 
+
+furrr::future_map(S2_times, compute_ensemble_mean, model = "S2", variable = "hi")
+furrr::future_map(S1_times, compute_ensemble_mean, model = "S1", variable = "hi")
+
+furrr::future_map(S2_times, compute_ensemble_mean, model = "S2", variable = "aice")
+furrr::future_map(S1_times, compute_ensemble_mean, model = "S1", variable = "aice")
+
+

### 2025-10-06 scripts/error_metrics.R

@@ -1,4 +1,6 @@
-logfile <- here::here("logs/error_metrics.log")
+chunk <- as.numeric(Sys.getenv("CHUNK", 0))
+
+logfile <- here::here(glue::glue("logs/error_metrics{chunk}.log"))
 log <- function(text) {
   messg <- paste0(format(lubridate::now()), ": ", text)
   if (interactive()) {
@@ -7,7 +9,6 @@ log <- function(text) {
     write(messg, logfile, append = TRUE)    
   }
 }
-on.exit(log("exiting"))
 writeLines("", logfile)
 log("Booting up")
 
@@ -157,6 +158,7 @@ compute_metrics <- function(forecast_time, member, version, obs_dataset) {
   tick <- lubridate::now()
   cdo_options_set(c("-L"))
   
+  log(glue::glue("Trying to get {forecast_time} forecast for {version} member {member}"))
   file <- hindcast(forecast_time, model = version, members = member)
 
   if (length(file) == 0) {
@@ -289,68 +291,47 @@ obs_dataset  <-  all_files[i, ]$obs_dataset
 version  <-  all_files[i, ]$version
 member  <-  all_files[i, ]$member
 
-future_pwalk(all_files, compute_metrics, .options = furrr_options(seed = NULL))
-
-
-read_measures <- function(dir) {
-  files <- list.files(dir, include.dirs = FALSE, recursive = TRUE, full.names = TRUE) 
+if (chunk != 0) {
+  log("Filtering..")
+  n <- floor(seq(1, nrow(all_files), length.out = 11))
   
-  dates <- paste0(here::here("data/derived"), "/(\\w*)/(\\w*)/(\\w*)/di_aice_(\\d{8})_e([\\dm]{2}).nc") |>
-    utils::strcapture(files,
-                      proto = list(
-                        measure = character(),
-                        version = character(),
-                        obs_dataset = character(),
-                        time_forecast = character(),
-                        member = character()
-                      ), perl = TRUE
-    ) |>
-    data.table::as.data.table() |>
-    _[, time_forecast := as.Date(time_forecast, format = "%Y%m%d")] |>
-    _[]
+  select <- seq(n[chunk], n[chunk+1])
   
-  # rmse_lon is 360 times larger than the other measures, so it 
-  # just doesnt fit into memory. :( 
[…]

### 2025-10-06 scripts/error_metrics_launch.R

@@ -1,0 +1,34 @@
+template <- r"(#!/bin/bash
+
+#PBS -P lo70
+#PBS -q normal
+#PBS -l ncpus=48
+#PBS -l mem=64GB
+#PBS -l walltime=3:00:00
+#PBS -l wd
+#PBS -l jobfs=100GB
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
+#PBS -N error_metrics{chunk}
+#PBS -o logs/
+#PBS -e logs/
+
+# Load module, always specify version number.
+module load R/4.3.1
+module load cdo/2.4.3
+
+# Run R application
+export PBS_WORKERS=35
+export CHUNK={chunk}
+Rscript scripts/error_metrics.R 
+)"
+
+
+chunks <- 1:10
+
+for (chunk in chunks) {
+  script <- glue::glue(here::here("scripts/pbs/error_metrics{chunk}.sh"))
+  
+  writeLines(glue::glue(template), script)
+  system(glue::glue("qsub {script}"))
+  
+}

### 2025-10-06 scripts/error_metrics_read.R

@@ -1,0 +1,92 @@
+logfile <- here::here(glue::glue("logs/error_metrics_read.log"))
+log <- function(text) {
+  messg <- paste0(format(lubridate::now()), ": ", text)
+  if (interactive()) {
+    message(messg)
+  } else {
+    write(messg, logfile, append = TRUE)    
+  }
+}
+writeLines("", logfile)
+log("Booting up")
+
+library(furrr)
+
+workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 2))
+plan_type <- "multicore"
+log(glue::glue("setting up {workers} workers in a {plan_type} plan"))
+
+plan(plan_type, workers = workers)
+
+
+rmse_dir <- here::here("data/derived", "rmse")
+
+rmse_lon_dir <- here::here("data/derived", "rmse_lon")
+
+iiee_dir <- here::here("data/derived", "iiee")
+
+
+read_measures <- function(dir) {
+  files <- list.files(dir, include.dirs = FALSE, recursive = TRUE, full.names = TRUE)
+  
+  dates <- paste0(here::here("data/derived"), "/(\\w*)/(\\w*)/(\\w*)/di_aice_(\\d{8})_e(\\d{2}|0em).nc") |>
+    utils::strcapture(files,
+                      proto = list(
+                        measure = character(),
+                        version = character(),
+                        obs_dataset = character(),
+                        time_forecast = character(),
+                        member = character()
+                      ), perl = TRUE
+    ) |>
+    data.table::as.data.table() |>
+    _[, time_forecast := as.Date(time_forecast, format = "%Y%m%d")] |>
+    _[]
+  
+  # rmse_lon is 360 times larger than the other measures, so it
+  # just doesnt fit into memory. :(
+  # if (dates$measure[1] == "rmse_lon") {
+  #   which_read <- dates[, which(obs_dataset == "cdr")]
+  #   files <- files[which_read]
+  #   dates <- dates[which_read]
+  # }
+  
+  future_map(seq_along(files), \(i) {
+    ncfile <- ncdf4::nc_open(files[[i]])
+    on.exit(ncdf4::nc_close(ncfile))
+    data <- try(metR::ReadNetCDF(ncfile, vars = c(value = "aice")), silent = TRUE)
+    
+    if (inherits(data, "try-error")) {
[…]

### 2025-10-06 scripts/giomas.R

@@ -1,0 +1,58 @@
+
+years <- 1981:2018
+
+download_giomas <- function(year) {
+  library(data.table)
+  gzip <- function(file) {
+    system(glue::glue("gzip -d {file}"))
+    gsub(".gz", "", file)
+  }
+
+  sh_subset <- list(j = 1:71)  # points in the southern hemisphere
+  
+  url <- glue::glue("http://pscfiles.apl.uw.edu/zhang/Global_seaice/heff.H{year}.nc.gz")
+  options(download.file.method="curl", download.file.extra="-k -L")
+  
+  thickness <- tempfile(fileext = ".nc.gz")
+  download.file(url, thickness)
+  thickness <- gzip(thickness) |> 
+    rcdo::cdo_settaxis(glue::glue("{year}-01-01,00:00:00,1mon")) |> 
+    rcdo::cdo_execute()
+  
+
+  url <- glue::glue("http://pscfiles.apl.uw.edu/zhang/Global_seaice/area.H{year}.nc.gz")
+  concentration <- tempfile(fileext = ".nc.gz")
+  download.file(url, concentration)
+  
+  concentration <- gzip(concentration) |> 
+    rcdo::cdo_settaxis(glue::glue("{year}-01-01,00:00:00,1mon")) |> 
+    rcdo::cdo_execute()
+
+  list(thickness = thickness, 
+      concentration = concentration,
+      year = year)
+}
+
+library(mirai)
+daemons(6)
+
+data <- mirai_map(years, download_giomas)[.progress] |> 
+  data.table::rbindlist() 
+
+
+data$thickness[1] |> 
+  rcdo::cdo_selname("lon_scaler,lat_scaler,dxt,dyt") |> 
+  rcdo::cdo_execute("data/raw/giomas_grid.nc", options = "-L")
+
+
+data$thickness |> 
+  lapply(\(x) rcdo::cdo_selname(x, "heff")) |> 
+  rcdo::cdo_mergetime() |> 
+  rcdo::cdo_execute("data/raw/giomas_thickness.nc", options = c("-L -O"))
+
+
+data$concentration |> 
+  lapply(\(x) rcdo::cdo_selname(x, "area")) |> 
+  rcdo::cdo_mergetime() |> 
+  rcdo::cdo_execute("data/raw/giomas_concentration.nc", options = c("-L -O"))
+  

### 2025-10-06 scripts/hindcast_climatology.R

@@ -3,7 +3,7 @@ library(data.table)
 library(metR)
 source(here::here("R/datasets.R"))
 
-future::plan("multicore", workers = 12)
+future::plan("multicore", workers = 20)
 
 file <- get_forecast_times("S2") |> 
   Filter(f = \(x) lubridate::day(x) == 1 & lubridate::month(x) == 1) |> 
@@ -24,26 +24,11 @@ compute_clim <- function(model, month) {
   dates <- get_forecast_times(model) |> 
     Filter(f = \(x) lubridate::day(x) == 1 & lubridate::month(x) == month & lubridate::year(x) %between% year_clim) 
   
-  # the 7th member has some weird jumps and inconsistencies. 
-  if (model == "S1" && month == 1) {
-    members <- c(1:6, 8:9)
-  } else {
-    members <- 1:9
-  }
-  forecasts <- hindcast(dates, model, 1)
-  
-  # Some forecast have less than the correct number of days. 
-  # Throw them out
-  size <- forecasts |> 
-    file.size()
-  
-  ndays <- floor(size/bytes_per_day)
-  
-  forecasts[ndays >= 215] |> 
+  hindcast(dates, model, members = "em") |> 
     cdo_mergetime() |> 
     cdo_ydaymean() |> 
     cdo_setyear(2000) |> 
-    cdo_execute(output = file, options = "-L")
+    cdo_execute(output = file, options = "-L -O", cache = TRUE)
 }
 
 

### 2025-10-06 scripts/thickness.R

@@ -1,0 +1,63 @@
+library(mirai)
+library(data.table)
+
+workers <- as.numeric(Sys.getenv("PBS_WORKERS", unset = 6))
+daemons(workers)
+
+source("R/datasets.R")
+source("R/functions.R")
+
+mean_thickness <- \(model, month, year) {
+  cdo_del29feb <- function(ifile, ofile = NULL) {
+    rcdo::cdo_operator("del29feb", params = NULL, 1, 1) |>
+      rcdo::cdo(
+        input = list(ifile),
+        params = NULL,
+        output = ofile
+      )
+  }
+  
+  month <- formatC(month, width = 2, flag = "0")
+  if (model == "S2") {
+    thickness <- "/g/data/ux62/access-s2/hindcast/raw_model/ice/hi/daily/e01/di_hi_{year}{month}01_e01.nc"
+  } else {
+    thickness <- "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/e01/di_ice_{year}{month}01_e01.nc"
+  }
+  thickness <- glue::glue(thickness)
+  
+  concentration <- hindcast(lubridate::make_date(year, month, 1), model = model, members = 1)
+  
+  if (!file.exists(thickness)) return(NA_character_)
+  
+  if (length(concentration) == 0) return(NA_character_)
+  
+  rcdo::cdo_cache_set(here::here("data/temp/cache/"))
+  
+  nsidc_grid <- here::here("data/raw/nsidc_grid.txt")
+  
+  volume <- thickness |> 
+    rcdo::cdo_selname("hi") |>
+    rcdo::cdo_remapbil(nsidc_grid) |> 
+    rcdo::cdo_mul(concentration) |> 
+    rcdo::cdo_fldint() |> 
+    cdo_del29feb() 
+  
+  area <- concentration |> 
+    rcdo::cdo_fldint() 
+  out <- glue::glue("data/derived/thickness/{model}/di_ice_{year}{month}01_e01.nc")
+  
+  rcdo::cdo_div(volume, area) |> 
+    rcdo::cdo_execute(options = "-L", output = out)
+  
+}
+
+
+forecast_thickness <- CJ(model = c("S2", "S1"),
+                         month = 1:12,
+                         year = 1981:2018) |>
+  _[, thickness := unlist(mirai_map(data.frame(month, year, model), mean_thickness,
+                                    hindcast = hindcast, 
[…]

### 2025-10-06 analysis/paper/access-ice.qmd
Forces sentence wrap

@@
 ---
+editor:
+  markdown:
+    wrap: sentence
 title: Evaluation of +S2 sea ice forecast and some other better title

### 2025-10-06 analysis/paper/access-ice.qmd
Updates introduction

@@
 
-Correctly modelling Antarctic sea ice is essential for understanding processes and improving climate projections to inform adaptation strategies. 
+Accurately modelling Antarctic sea ice is essential for understanding processes and improving climate projections to inform adaptation strategies.
 Accurate seasonal to sub-seasonal forecasts are also crucial for operation contingency planning in and around the Antarctic continent, including scientific missions, fisheries, and tourism [@desilva2020; @wagner2020].
-Getting sea-ice right might also help improve weather forecasts over and away from sea-ice regions [@rinke2006; @wang2024 @semmler2016].
-
-However, producing accurate Antarctic sea-ice forecasts has been challenging due to model biases, inherent large variability and complexity, and it has lagged behind Arctic sea-ice forecasts [@zampieri2019;  @gao2024a].
+Improvements in modelled sea-ice might also help improve weather forecasts over and away from sea-ice regions [@rinke2006; @wang2024 @semmler2016].
+
+However, producing accurate Antarctic sea-ice forecasts has been challenging due to model biases, inherent large variability and complexity, and they have lagged behind Arctic sea-ice forecasts[@zampieri2019; @gao2024a].
 Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023] and machine learning approaches [e.g.
@@
 
-Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea-ice initial conditions affect the quality of the forecast and in which timescales.
-Exploring seasonal predictions of Arctic sea-ice, @guemas2016 found that sea-ice initial conditions are only important in Autumn to predict Summer sea-ice, but the impact wasn't as dramatic when predicting Winter sea-ice.
-@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea-ice forecast initialised in July but not when initialised in January.
-
-For the Antarctic, @holland2013 studied the initial-value predictability of Antarctic sea-ice in the CCSESM3 model. 
+Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea-ice initial conditions affect the quality of the forecast and at what timescales.
+Exploring seasonal predictions of Arctic sea-ice, @guemas2016 found that sea-ice initial conditions are important in autumn to predict summer sea-ice, but the impact wasn’t as dramatic when predicting winter sea-ice.
+@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea-ice forecasts initialised in July but not when initialised in January.
+
+For the Antarctic, @holland2013 studied the initial-value predictability of Antarctic sea-ice in a perfect model study using the CCSM3 model.
 They found that sea-ice and ocean initial conditions provide predictive information to forecast sea-ice edge location several months in advance and that some predictability is retained for up to two years thanks to ocean heat content anomalies that are advected eastward.
-This is in contrast with @marchi2020, who ran perfect model experiments to argue that uncertainty in the atmosphere state and evolution is the main driver of uncertainty in Antarctic sea-ice extent prediction, with sea-ice and ocean initial conditions having lesser importance.
+This is in contrast with @marchi2020, who ran perfect model experiments to argue that uncertainty in the predicted atmospheric state and evolution is the main driver of uncertainty in Antarctic sea-ice extent prediction on seasonal timescales, with sea-ice and ocean initial conditions having lesser importance.
 More recently, @morioka2022 studied decadal forecasts of Antarctic sea-ice and found that initialising ocean and sea-ice improved the correlation between simulated and observed sea-ice concentration evolution in the Amundsen–Bellingshausen Sea.
-It is hard to compare these studies since they are based on forecasts initialised at different times of the year: @holland2013 ran 20 ensemble members initialised on the 1st of January of a particular year, @marchi2020 ran forecasts from the 1st of March and 1st of September and @morioka2022 run forecasts only from the 1st of March.
+It is hard to compare these studies since they are based on forecasts initialised at different times of the year and different frameworks: @holland2013 ran 20 ensemble members initialised on the 1st of January of a particular year, @marchi2020 ran forecasts from the 1st of March and 1st of September and @morioka2022 run forecasts only from the 1st of March.
 @marchi2020 also used a coupled ocean–sea-ice model instead of a fully coupled model like @holland2013 did.
-
-In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the Australian Community Climate and Earth System Simulator – Seasonal (ACCES-S) to its version two. 
-The change in version was focused on using ocean, sea-ice and land initial conditions generated by the BoM instead of depending on the UK’ Met Office. 
-Crucially, ACCESS-S2 does not assimilate sea-ice observations, and instead sea-ice is only affected by the ocean and atmospheric data assimilation via the coupled integration.
-
-Since model configuration is identical between the two systems, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year.
-This can be an opportunity to test the effect of sea-ice initial conditions on the forecast of sea-ice concentrations.
-
-In this study we compare sea-ice hindcasts produced by ACCESS-S1 and ACCESS-S2. 
+@morioka2022 used observed sea ice initial conditions and compared with observations, while @marchi2020 and @holland2013 are perfect model studies.
+
+In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the Australian Community Climate and Earth System Simulator – Seasonal (ACCES-S) from version S1 to S2.
+While the base model remained the same, the change in version was focused on using ocean, sea-ice and land initial conditions generated by the BoM instead of depending on the UK Met Office.
+Crucially, compared to ACCESS-S1, ACCESS-S2 does not assimilate sea-ice observations, so sea ice is only affected by the ocean and atmospheric data assimilation via the coupled integration.
+
+Since model configuration is identical between ACCESS-S1 and ACCESS-S2, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year, with the only difference being the initial conditions.
+This provides an opportunity to test the effect of sea-ice initial conditions on the forecast of sea-ice concentrations and the climate.
+
+In this study we compare sea-ice hindcasts produced by +S1 and +S2.
 We focus on seasonality of errors and biases and the effect of the data assimilation system. 

### 2025-10-06 analysis/paper/access-ice.qmd
Formatting

@@
 ```{r mirai_daemons}
-if (!mirai::daemons_set()) invisible(daemons(6))
+if (!mirai::daemons_set()) {
+  invisible(daemons(6))
+}
 ```
@@
 ```{r datasets}
-datasets <- list(cdr = CDR(),
-                 osi = OSI())
+datasets <- list(cdr = CDR(), osi = OSI())
 which_dataset <- "cdr"
@@
 ```{r range_max_lag}
-range_max_lag <- hindcast_extent[mday(time) == 2, .SD[which.max(lag)], 
-                                 by = .(forecast_time, model)] |> 
-  _[, .(max = max(lag),
-        min = min(lag)),
-    by = model]
+range_max_lag <- hindcast_extent[
+  mday(time) == 2,
+  .SD[which.max(lag)],
+  by = .(forecast_time, model)
+] |>
+  _[, .(max = max(lag), min = min(lag)), by = model]
 ```
@@
   tidyfast::dt_pivot_longer(cols = c(Forecast, Difference)) |>
-  _[, average(value), 
-    by = .(dataset, 
+  _[,
+    average(value),
+    by = .(
+      dataset,
       model,
@@
       forecast_time = update(forecast_time, year = 2001),
-           lag = as.numeric(as.Date(time) - forecast_time))] |> 
+      lag = as.numeric(as.Date(time) - forecast_time)
+    )
+  ] |>
   _[, time := forecast_time + lag] |>
@@
   ggplot(aes(time2, estimate)) +
-  geom_line(data = extent_climatology[dataset == which_dataset] |> 
+  geom_line(
+    data = extent_climatology[dataset == which_dataset] |>
       _[, measure := factor("Forecast", levels = c("Forecast", "Difference"))],
-            aes(x = as.Date(time), colour = dataset, 
-                group = dataset)) +
-
-  geom_hline(data = data.table(y = 0, measure = factor("Difference", levels = c("Forecast", "Difference"))), aes(yintercept = 0)) +
-  geom_line(aes(group = interaction(forecast_time, time != time2), colour = model)) +
-  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], 
-                           by = .(forecast_time, model, measure)], 
+    aes(x = as.Date(time), colour = dataset, group = dataset)
+  ) +
+
+  geom_hline(
+    data = data.table(
+      y = 0,
+      measure = factor("Difference", levels = c("Forecast", "Difference"))
+    ),
+    aes(yintercept = 0)
+  ) +
+  geom_line(aes(
+    group = interaction(forecast_time, time != time2),
+    colour = model
+  )) +
+  geom_point(
+    data = \(x) {
+      x[
+        mday(time) == 2,
+        .SD[which.max(lag)],
+        by = .(forecast_time, model, measure)
+      ]
+    },
     aes(fill = model),
-             shape = 24, size = 2.2) +
-  geom_point(data = \(x) x[lag == 1], 
-             aes(fill = model), 
-             shape = 21, size = 2) +
-
-  scale_x_date(NULL, date_breaks = "1 month", 
-               date_labels = "%b", expand = c(0, 0)) +
+    shape = 24,
+    size = 2.2
+  ) +
+  geom_point(data = \(x) x[lag == 1], aes(fill = model), shape = 21, size = 2) +
+
+  scale_x_date(
+    NULL,
+    date_breaks = "1 month",
+    date_labels = "%b",
+    expand = c(0, 0)
+  ) +
   scale_y_continuous(NULL, labels = labels_extent) +
@@
   #                    aesthetics = c("fill", "color"), guide = "none") +
-  facet_grid(measure ~ model, scales = "free_y", labeller = labeller(model = labels_models)) +
+  facet_grid(
+    measure ~ model,
+    scales = "free_y",
+    labeller = labeller(model = labels_models)
+  ) +
   tag_facets(tag = "rc") +
@@
 ```{r extent-delta-compute}
-n <- 366*3
+n <- 366 * 3
 w <- 11
@@
   _[, .(aice = mean(aice)), by = .(model, forecast_time, time)] |>
-  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), 
-    by = .(model, forecast_time)] |> 
-  _[, dif := frollmean(dif, n = w, align = "center"), by = .(model, forecast_time)] |> 
+  _[,
+    dif := c(NA, diff(aice) / diff(as.numeric(as.Date(time)))),
+    by = .(model, forecast_time)
+  ] |>
+  _[,
+    dif := frollmean(dif, n = w, align = "center"),
+    by = .(model, forecast_time)
+  ] |>
   na.omit() |>
-  _[, average(dif), by = .(model, month(forecast_time), time = update(time, year = 2000))]
+  _[,
+    average(dif),
+    by = .(model, month(forecast_time), time = update(time, year = 2000))
+  ]
 
@@
   _[order(time)] |>
-  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
+  _[,
+    dif := c(NA, diff(aice) / diff(as.numeric(as.Date(time)))),
+    by = .(dataset)
+  ] |>
   na.omit() |>
@@
   geom_line(aes(color = dataset)) +
-  geom_line(data = dif_hindcast,
-            aes(color = model, group = interaction(model, month, month > month(time)))) +
-  scale_x_datetime(NULL, date_breaks = "1 month", 
-                   date_labels = "%b", expand = c(0, 0)) +
-  scale_y_continuous(NULL, labels = \(x) labels_extent(x, units = "M km²/day")) +
+  geom_line(
+    data = dif_hindcast,
+    aes(color = model, group = interaction(model, month, month > month(time)))
+  ) +
+  scale_x_datetime(
+    NULL,
+    date_breaks = "1 month",
+    date_labels = "%b",
+    expand = c(0, 0)
+  ) +
+  scale_y_continuous(NULL, labels = \(x) {
+    labels_extent(x, units = "M km²/day")
+  }) +
   scale_color_models +
@@
 zeropad <- function(x, n = 2) formatC(x, width = 2, flag = "0")
-monnb <- function(d) { lt <- as.POSIXlt(as.Date(d, origin="1900-01-01")); 
-lt$year*12 + lt$mon } 
-mondf <- function(d1, d2) { monnb(d2) - monnb(d1) }
+monnb <- function(d) {
+  lt <- as.POSIXlt(as.Date(d, origin = "1900-01-01"))
+  lt$year * 12 + lt$mon
+}
+mondf <- function(d1, d2) {
+  monnb(d2) - monnb(d1)
+}
 
@@
   lapply(c("S1", "S2"), \(model) {
-    here::here(glue::glue("data/derived/climatology/{model}/{zeropad(month)}/em.nc")) |> 
+    here::here(glue::glue(
+      "data/derived/climatology/{model}/{zeropad(month)}/em.nc"
+    )) |>
       cdo_selmonth(month) |>
@@
   copy() |>
-  _[, .(aice = mean(aice)), by = .(month(time), ygrid, xgrid, forecast_month, model)] |> 
+  _[,
+    .(aice = mean(aice)),
+    by = .(month(time), ygrid, xgrid, forecast_month, model)
+  ] |>
   _[forecast_month == month]
@@
     ggplot(aes(xgrid, ygrid)) +
-    geom_contour_fill(aes(z = aice - obs, fill = after_stat(level)), 
-                      breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
-    scale_fill_divergent_discretised(paste0("Bias with respect to ",
-                                            labels_models[which_dataset]),
+    geom_contour_fill(
+      aes(z = aice - obs, fill = after_stat(level)),
+      breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)
+    ) +
+    scale_fill_divergent_discretised(
+      paste0("Bias with respect to ", labels_models[which_dataset]),
       low = trans_pink,
-                                     high = trans_blue) +
+      high = trans_blue
+    ) +
     # geom_polygon(data = ice_shelves_nsicd, aes(X, Y, group = group),
@@
     geom_antarctica_fill +
-    facet_wrap(~ month + model , labeller = labeller(month = labels_month)) +
+    facet_wrap(~ month + model, labeller = labeller(month = labels_month)) +
     wide_legend +
@@
   print(g)
-
 })
@@
 
-@fig-bias shows the difference of monthly mean sea ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest lead time
-From October to May, the model underestimates sea ice concentrations in most regions except for the inner Weddell Sea in April and May, where sea ice concentrations saturate to 1 both in the observations and forecasts.
+@fig-bias shows the difference of monthly mean sea ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest lead time From October to May, the model underestimates sea ice concentrations in most regions except for the inner Weddell Sea in April and May, where sea ice concentrations saturate to 1 both in the observations and forecasts.
 In winter, the differences are mostly on the sea ice edge with light positive bias in the African sector of East Antarctica and negative bias around the Indian Ocean sector which partially compensate, resulting in the near-zero extent bias seen in those months ([Fig.
@@
   _[aice == 0, aice := NA] |>
-  _[, .(aice = mean(aice)), by = .(dataset, time = as.Date(round_date(time, "month")))] |>
-  _[, aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE),
-    by = .(month(time), dataset)] |>
+  _[,
+    .(aice = mean(aice)),
+    by = .(dataset, time = as.Date(round_date(time, "month")))
+  ] |>
+  _[,
+    aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE),
+    by = .(month(time), dataset)
+  ] |>
   _[]
@@
   copy() |>
-  _[, aice := aice - mean(aice[year(time) %between% c(1981, 2011)]),
-    by = .(lag, model, time = update(time, year = 2000))] |>
-  _[, .(forecast = mean(aice)),
-    by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
+  _[,
+    aice := aice - mean(aice[year(time) %between% c(1981, 2011)]),
+    by = .(lag, model, time = update(time, year = 2000))
+  ] |>
+  _[,
+    .(forecast = mean(aice)),
+    by = .(model, forecast_time, time = as.Date(round_date(time, "month")))
+  ] |>
   _[, lag := months_difference(time, forecast_time)]
@@
   geom_line(aes(y = forecast, color = model)) +
-  geom_text(data = \(x) x[time %between% s1_range, 
+  geom_text(
+    data = \(x) {
+      x[
+        time %between% s1_range,
         rmse(aice, forecast, signif = 2, scale = 1e12),
-                          by = .(lag, model)],
+        by = .(lag, model)
+      ]
+    },
     aes(label = text),
-            x = -Inf, y = Inf,
-            vjust = 1.2, hjust = 0) +
-
-  geom_text(data = \(x) x[time %between% s1_range, 
-                          correlate(aice, forecast),
-                          by = .(lag, model)],
+    x = -Inf,
+    y = Inf,
+    vjust = 1.2,
+    hjust = 0
+  ) +
+
+  geom_text(
+    data = \(x) {
+      x[time %between% s1_range, correlate(aice, forecast), by = .(lag, model)]
+    },
     aes(label = text),
-            vjust = 0, hjust = 1.2,
-            x = Inf, y = -Inf) +
-
-  scale_y_continuous(NULL,
-                     labels = labels_extent) +
+    vjust = 0,
+    hjust = 1.2,
+    x = Inf,
+    y = -Inf
+  ) +
+
+  scale_y_continuous(NULL, labels = labels_extent) +
   scale_x_date(NULL, expand = c(0, 0)) +
-  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Month: ", x))) +
+  facet_grid(
+    lag ~ model,
+    labeller = labeller(lag = \(x) paste0("Month: ", x))
+  ) +
   scale_color_models +
@@
 
-Strangely, for +S2, RMSE improves with lead time, even though the correlation degrades with lead time. This is puzzling behaviour that goes contrary to what is usually seen in prediction models. 
+Strangely, for +S2, RMSE improves with lead time, even though the correlation degrades with lead time.
+This is puzzling behaviour that goes contrary to what is usually seen in prediction models.
 The explanation seems to be the mentioned increased interannual variability.
@@
 
-obs_sd <- monthly_extent[time %between% s1_range,
+obs_sd <- monthly_extent[
+  time %between% s1_range,
   sd_ci(aice, na.rm = TRUE),
-                                   by = .(month(time), dataset)] |> 
+  by = .(month(time), dataset)
+] |>
   _[, .(dataset, month, obs_sd = estimate)]
 
-hindcast_extent_sd <- hindast_extent_monthly[time %between% s1_range, 
+hindcast_extent_sd <- hindast_extent_monthly[
+  time %between% s1_range,
   sd_ci(forecast, na.rm = TRUE),
-                       by = .(model, lag, month(time))]
+  by = .(model, lag, month(time))
+]
 
@@
   # geom_hline(data = obs_sd, aes(yintercept = estimate, color = dataset)) +
-  geom_ribbon(aes(ymin = low/obs_sd, ymax = high/obs_sd, fill = model), alpha = 0.5) +
+  geom_ribbon(
+    aes(ymin = low / obs_sd, ymax = high / obs_sd, fill = model),
+    alpha = 0.5
+  ) +
   geom_line(aes(color = model)) +
@@
   scale_y_continuous("Standardised sea ice extent standard deviation") +
-  scale_x_continuous("Lead time", breaks = c(0:9), expand = c(0, 0), 
-                     trans = scales::reverse_trans()) +
+  scale_x_continuous(
+    "Lead time",
+    breaks = c(0:9),
+    expand = c(0, 0),
+    trans = scales::reverse_trans()
+  ) +
   scale_color_models +
   scale_fill_models +
-  facet_wrap(~ month, labeller = labeller(month = labels_month))
+  facet_wrap(~month, labeller = labeller(month = labels_month))
 
@@
 
-aice_forecast_example <- c(lapply(c("S1", "S2"), \(model) {
+aice_forecast_example <- c(
+  lapply(c("S1", "S2"), \(model) {
     month_clim <- formatC(month(forecast_date), width = 2, flag = "0")
@@
     hindcast(forecast_date, model, 1) |>
-    cdo_ydaysub(here::here(glue::glue("data/derived/climatology/{model}/{month_clim}/em.nc"))) |>
+      cdo_ydaysub(here::here(glue::glue(
+        "data/derived/climatology/{model}/{month_clim}/em.nc"
+      ))) |>
       cdo_execute() |>
@@
   }),
-  list(CDR() |> 
+  list(
+    CDR() |>
       anomalies() |>
       ReadNetCDF("aice", subset = list(time = sample_date)) |>
-       _[, model := "cdr"])) |> 
+      _[, model := "cdr"]
+  )
+) |>
   rbindlist()
 
-thickness_example <- c(S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/hi/daily/e01/di_hi_{format(forecast_date, '%Y%m%d')}_e01.nc",
-  S1 = "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/e01/di_ice_{format(forecast_date, '%Y%m%d')}_e01.nc") |> 
+thickness_example <- c(
+  S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/hi/daily/e01/di_hi_{format(forecast_date, '%Y%m%d')}_e01.nc",
+  S1 = "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/e01/di_ice_{format(forecast_date, '%Y%m%d')}_e01.nc"
+) |>
   purrr::imap(\(f, model) {
@@
       ReadNetCDF("hi", subset = list(time = sample_date))
-
   }) |>
@@
 b <- seq(-0.5, 0.5, by = 0.1)
-b <- b[b!= 0]
+b <- b[b != 0]
 
@@
   ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = aice, fill = ..level..),
-                    breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
-  scale_fill_divergent_discretised("Sea-ice concentration anomaly",
+  geom_contour_fill(
+    aes(z = aice, fill = ..level..),
+    breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)
+  ) +
+  scale_fill_divergent_discretised(
+    "Sea-ice concentration anomaly",
     low = trans_pink,
-                                   high = trans_blue) +
+    high = trans_blue
+  ) +
   geom_antarctica_fill +
@@
     ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = hi, fill = ..level..), 
-                    breaks = c(AnchorBreaks(exclude = 0, binwidth = 0.25)(c(0, 2)), Inf)) +
+  geom_contour_fill(
+    aes(z = hi, fill = ..level..),
+    breaks = c(AnchorBreaks(exclude = 0, binwidth = 0.25)(c(0, 2)), Inf)
+  ) +
   # as.discretised_scale(scale_fill_viridis_c)("Sea ice thickness",
   #                                            guide = guide_colorsteps(order = 9)) +
-  scale_fill_divergent_discretised("Sea-ice thickness",
+  scale_fill_divergent_discretised(
+    "Sea-ice thickness",
     high = scales::muted("#5ab4ac"),
-                                   guide = guide_colorsteps(order = 9)) +
+    guide = guide_colorsteps(order = 9)
+  ) +
   ggnewscale::new_scale_fill() +
 
-  geom_contour_fill(data = \(x) x |> 
+  geom_contour_fill(
+    data = \(x) {
+      x |>
         dcast(time + xgrid + ygrid ~ model, value.var = "hi") |>
-                      _[, model := factor("Difference", levels = c("S1", "S2", "Difference"))], 
-                    aes(z = S1 - S2, fill = ..level..), breaks = c(-Inf, b, Inf)) +
-
-  scale_fill_divergent_discretised("Sea-ice thickness difference",
-                                   guide = guide_colorsteps(order = 0)) +
+        _[, model := factor("Difference", levels = c("S1", "S2", "Difference"))]
+    },
+    aes(z = S1 - S2, fill = ..level..),
+    breaks = c(-Inf, b, Inf)
+  ) +
+
+  scale_fill_divergent_discretised(
+    "Sea-ice thickness difference",
+    guide = guide_colorsteps(order = 0)
+  ) +
   geom_antarctica_fill +
@@
   wide_legend +
-  facet_wrap(~ model) +
+  facet_wrap(~model) +
   plot_layout(ncol = 1)
@@
 mean_thickness_access <- \(forecast_time, model) {
-  concentration <- hindcast(forecast_time, model = model, members = "em", variable = "aice")
-  thickness <- hindcast(forecast_time, model = model, members = "em", variable = "hi")
-
-  if (length(thickness) == 0) return(NA_character_)
-  if (length(concentration) == 0) return(NA_character_)
+  concentration <- hindcast(
+    forecast_time,
+    model = model,
+    members = "em",
+    variable = "aice"
+  )
+  thickness <- hindcast(
+    forecast_time,
+    model = model,
+    members = "em",
+    variable = "hi"
+  )
+
+  if (length(thickness) == 0) {
+    return(NA_character_)
+  }
+  if (length(concentration) == 0) {
+    return(NA_character_)
+  }
 
@@
 get_dir_part <- function(path, n = 1) {
-  if (n == 1) return(basename(path))
+  if (n == 1) {
+    return(basename(path))
+  }
 
@@
   file <- get_dir_part(path, 1)
-  l <- 6+1+6+3
+  l <- 6 + 1 + 6 + 3
   n <- nchar(file)
-  s <- n-l+1
+  s <- n - l + 1
 
   date_start <- substr(file, start = s, s + 5)
-  date_end <- substr(file, start = s+7, n-3)
-  list(start = lubridate::ym(date_start),
-       end = lubridate::ym(date_end))
+  date_end <- substr(file, start = s + 7, n - 3)
+  list(start = lubridate::ym(date_start), end = lubridate::ym(date_end))
 }
@@
   _[mday(forecast_time) == 1] |>
-  _[, thickness := unlist(mirai_map(data.frame(forecast_time, model), 
+  _[,
+    thickness := unlist(mirai_map(
+      data.frame(forecast_time, model),
       mean_thickness_access,
-                                    hindcast = hindcast, mean_thickness = mean_thickness)[.progress])] |> 
+      hindcast = hindcast,
+      mean_thickness = mean_thickness
+    )[.progress])
+  ] |>
   na.omit()
@@
 
-giomas <- mean_thickness(giomas_concentration, giomas_thickness, giomas_cellarea) |> 
+giomas <- mean_thickness(
+  giomas_concentration,
+  giomas_thickness,
+  giomas_cellarea
+) |>
   cdo_ymonmean() |>
@@
 ```
+
 ```{r}
 hindcast_mean_thickness <- hindcast_thickness |>
-  _[, thickness |> cdo_monmean() |> cdo_execute(), by = .(model, forecast_time)] |>
-  _[, metR::ReadNetCDF(V1, c(thickness = "hi")), by = .(model, forecast_time)] |> 
+  _[,
+    thickness |> cdo_monmean() |> cdo_execute(),
+    by = .(model, forecast_time)
+  ] |>
+  _[,
+    metR::ReadNetCDF(V1, c(thickness = "hi")),
+    by = .(model, forecast_time)
+  ] |>
   _[, time := as.Date(update(time, day = 1))] |>
@@
   _[model %in% unique(model)[1:10]] |>
-  _[, mean_thickness(cdo_sellonlatbox(concentration, 0, 360, -90, -40),
-                     cdo_sellonlatbox(thickness, 0, 360, -90, -40)), 
-            by = .(model, start, end)] |> 
+  _[,
+    mean_thickness(
+      cdo_sellonlatbox(concentration, 0, 360, -90, -40),
+      cdo_sellonlatbox(thickness, 0, 360, -90, -40)
+    ),
+    by = .(model, start, end)
+  ] |>
   _[, .(thickness = V1 |> cdo_mergetime() |> cdo_execute()), by = .(model)] |>
@@
   geom_line(aes(group = model, color = "CMIP6 models")) +
-    geom_line(data = giomas, aes(y = thickness, color = "GIOMAS"), linewidth = 1.4) +
+  geom_line(
+    data = giomas,
+    aes(y = thickness, color = "GIOMAS"),
+    linewidth = 1.4
+  ) +
   scale_color_manual(values = c("CMIP6 models" = "black", GIOMAS = "#1a7e21")) +
@@
 ```{r}
-hindcast_mean_thickness[time %between% s1_range,
+hindcast_mean_thickness[
+  time %between% s1_range,
   average(thickness),
-                          keyby = .(lag, model, month(time))] |> 
+  keyby = .(lag, model, month(time))
+] |>
   ggplot(aes(lag, estimate)) +
@@
   scale_y_continuous("Mean sea-ice thickness") +
-  scale_x_continuous("Lead time", breaks = c(0:9), expand = c(0, 0), 
-                     trans = scales::reverse_trans()) +
-  facet_wrap(~ month, labeller = labeller(month = labels_month))
+  scale_x_continuous(
+    "Lead time",
+    breaks = c(0:9),
+    expand = c(0, 0),
+    trans = scales::reverse_trans()
+  ) +
+  facet_wrap(~month, labeller = labeller(month = labels_month))
 
@@
   _[1] |>
-  lapply(\(x) x |> 
+  lapply(\(x) {
+    x |>
       anomalies(climatology(x)) |>
@@
       _[, let(lat = NULL, lon = NULL)]
-  ) |> 
+  }) |>
   rbindlist(idcol = "obs_dataset") |>
@@
 clim_std2 <- clim_std[errors, on = .NATURAL] |>
-  _[, let(climatology = NULL,
-          value = climatology)] |> 
+  _[, let(climatology = NULL, value = climatology)] |>
   na.omit() |>
@@
 t_test <- function(x, y, ...) {
-
   test <- t.test(x, y, ...)
 
-  list(low = test[["conf.int"]][1],
+  list(
+    low = test[["conf.int"]][1],
     high = test[["conf.int"]][2],
@@
   dcast(time + time_forecast + member + lag ~ version, value.var = "value") |>
-  _[, t_test(na.omit(S1), na.omit(S2)), 
-    by = .(lag, month(time_forecast))] |> 
+  _[, t_test(na.omit(S1), na.omit(S2)), by = .(lag, month(time_forecast))] |>
   _[p.value > 0.01] |>
@@
 
-labels_month_significant <- setNames(paste0(month.abb, " (", dif$lag, ")"),
-                                     1:12)
+labels_month_significant <- setNames(
+  paste0(month.abb, " (", dif$lag, ")"),
+  1:12
+)
 ```
@@
   function(x) {
-    unique(x[, .(lag, month)])[, time := make_date(month = month, day = 1) + lag] |> 
-      _[mday(time) == day, .(lag, month, 
-                             month2 = lubridate::month(time, label = TRUE)
-      )] |> 
+    unique(x[, .(lag, month)])[,
+      time := make_date(month = month, day = 1) + lag
+    ] |>
+      _[
+        mday(time) == day,
+        .(lag, month, month2 = lubridate::month(time, label = TRUE))
+      ] |>
       _[, .SD[((seq_len(.N) + 1) %% step) == 0], by = month]
@@
   _[obs_dataset == which_dataset] |>
-  _[, average(value),
-    by = .(lag, version, measure, month(time_forecast), obs_dataset)]
+  _[,
+    average(value),
+    by = .(lag, version, measure, month(time_forecast), obs_dataset)
+  ]
 
@@
 
-  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
-             colour = "gray50", linewidth = 0.1) +
-
-  geom_text(data = month_from_lag(15, 1), aes(y = Inf, label = month2),
+  geom_vline(
+    data = month_from_lag(1, 1),
+    aes(xintercept = lag),
+    colour = "gray50",
+    linewidth = 0.1
+  ) +
+
+  geom_text(
+    data = month_from_lag(15, 1),
+    aes(y = Inf, label = month2),
     size = 2.5,
-            vjust = 1) +
-
-  geom_text(data = month_from_lag(1, 1), aes(y = -Inf, label = lag), 
+    vjust = 1
+  ) +
+
+  geom_text(
+    data = month_from_lag(1, 1),
+    aes(y = -Inf, label = lag),
     size = 2.5,
-            vjust = 0) +
+    vjust = 0
+  ) +
 
@@
   scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
-  labs(y = "RMSE", 
-       x = "Lead time (days)") +
+  labs(y = "RMSE", x = "Lead time (days)") +
   scale_color_models +
   scale_fill_models +
-  facet_wrap(~ month, labeller = labeller(month = labels_month_significant)) +
+  facet_wrap(~month, labeller = labeller(month = labels_month_significant)) +
   coord_cartesian(clip = "off")
@@
 
-rbind(errors2 |> 
+rbind(
+  errors2 |>
     copy() |>
     _[obs_dataset == which_dataset] |>
-  _[, base := low[version == "persistence"], by = .(lag, measure, month, obs_dataset)] |> 
+    _[,
+      base := low[version == "persistence"],
+      by = .(lag, measure, month, obs_dataset)
+    ] |>
     _[version != "persistence"] |>
@@
     _[obs_dataset == which_dataset] |>
-  _[, base := low[version == "climatology"], by = .(lag, measure, month, obs_dataset)] |> 
+    _[,
+      base := low[version == "climatology"],
+      by = .(lag, measure, month, obs_dataset)
+    ] |>
     _[version != "climatology"] |>
@@
     # _[lag > 2] |>
-  _[, rle(estimate < base), by= .(version, measure, month, obs_dataset)] |> 
+    _[, rle(estimate < base), by = .(version, measure, month, obs_dataset)] |>
     _[, .SD[1], by = .(version, measure, month, obs_dataset)] |>
-  tidyr::complete(version, measure, month, obs_dataset, values, fill = list(lengths = 0)) |> 
+    tidyr::complete(
+      version,
+      measure,
+      month,
+      obs_dataset,
+      values,
+      fill = list(lengths = 0)
+    ) |>
     as.data.table() |>
     _[values == TRUE] |>
-  _[, .(version, month, obs_dataset, lag = lengths, benchmark = "climatology")] 
+    _[, .(
+      version,
+      month,
+      obs_dataset,
+      lag = lengths,
+      benchmark = "climatology"
+    )]
 ) |>
@@
   ggplot(aes(month, lag)) +
-  ggbraid::geom_braid(data = \(x) dcast(x, month + lag + version ~ benchmark, value.var = "lag"),
+  ggbraid::geom_braid(
+    data = \(x) dcast(x, month + lag + version ~ benchmark, value.var = "lag"),
     alpha = 0.3,
-                      aes(ymin = persistence, ymax = climatology,
-                          fill = climatology > persistence)) +
+    aes(
+      ymin = persistence,
+      ymax = climatology,
+      fill = climatology > persistence
+    )
+  ) +
   geom_line(aes(color = benchmark)) +
   # geom_point(aes(color = time), data = \(x) x[lag != 0]) +
-  geom_text(data = \(x) x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |> 
-              _[benchmark == "climatology"],
+  geom_text(
+    data = \(x) {
+      x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |>
+        _[benchmark == "climatology"]
+    },
     aes(label = format(time, "%d %b")),
-            vjust = -0.4) +
-  geom_point(data = \(x) x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |> 
-               _[benchmark == "climatology"],
-             aes(color = benchmark)) +
+    vjust = -0.4
+  ) +
+  geom_point(
+    data = \(x) {
+      x[, .SD[which.max(lag)], by = .(version, month, obs_dataset)] |>
+        _[benchmark == "climatology"]
+    },
+    aes(color = benchmark)
+  ) +
   scale_color_models +
   scale_y_continuous("Lead time (days)") +
-  scale_x_continuous("Forecast month", breaks = 1:12, labels = month.abb, minor_breaks = NULL) +
-  scale_fill_manual(guide = "none", values = c(`TRUE` = "seagreen", `FALSE` = "#FFFFFF00")) +
-  facet_wrap(~ version, ncol = 1, labeller = labeller(version = labels_models)) +
+  scale_x_continuous(
+    "Forecast month",
+    breaks = 1:12,
+    labels = month.abb,
+    minor_breaks = NULL
+  ) +
+  scale_fill_manual(
+    guide = "none",
+    values = c(`TRUE` = "seagreen", `FALSE` = "#FFFFFF00")
+  ) +
+  facet_wrap(~version, ncol = 1, labeller = labeller(version = labels_models)) +
   coord_cartesian(clip = "off") +
@@
   # _[1] |>
-  lapply(\(x) x |> 
+  lapply(\(x) {
+    x |>
       anomalies(climatology(x)) |>
@@
       _[, let(lat = NULL, lon = lon + 12)]
-  ) |> 
+  }) |>
   rbindlist(idcol = "obs_dataset") |>
@@
   clim_std_lon[i = _, on = .NATURAL] |>
-    _[, .(dif = mean(value - climatology, na.rm = TRUE)),
-      by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))] 
+  _[,
+    .(dif = mean(value - climatology, na.rm = TRUE)),
+    by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))
+  ]
 ```
@@
 lapply(c("S1", "S2"), \(x) {
-
   g <- gdata |>
     _[version == x] |>
-    _[, dif := Smooth2D(lag, lon, dif, method = smooth_svd(.1)), by = .(month, version)] |>
+    _[,
+      dif := Smooth2D(lag, lon, dif, method = smooth_svd(.1)),
+      by = .(month, version)
+    ] |>
     ggplot(aes(lon, lag)) +
-    geom_contour_fill(aes(z = dif, fill = after_stat(level)),
-                      breaks = AnchorBreaks(0, binwidth = binwidth))  +
-    geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.2),
-                        breaks = AnchorBreaks(0, binwidth = binwidth)) +
-
-    geom_contour_fill(data = topo_lonlat,
-                      aes(lon, scales::rescale(lat, c(-60, 1)), 
-                          z = z, 
-                          fill = NULL),
+    geom_contour_fill(
+      aes(z = dif, fill = after_stat(level)),
+      breaks = AnchorBreaks(0, binwidth = binwidth)
+    ) +
+    geom_contour_tanaka(
+      aes(z = dif),
+      range = c(0.01, 0.2),
+      breaks = AnchorBreaks(0, binwidth = binwidth)
+    ) +
+
+    geom_contour_fill(
+      data = topo_lonlat,
+      aes(lon, scales::rescale(lat, c(-60, 1)), z = z, fill = NULL),
       fill = "#FAFAFA",
-                      breaks = c(0, Inf)) +
-
-    geom_contour2(data = topo_lonlat,
+      breaks = c(0, Inf)
+    ) +
+
+    geom_contour2(
+      data = topo_lonlat,
       aes(lon, scales::rescale(lat, c(-60, 1)), z = z),
-                  breaks = 0) +
-    geom_hline(data = month_from_lag(1, 1), aes(yintercept = lag),
-               colour = "gray50", linewidth = 0.1) +
-
-    geom_text(data = month_from_lag(15, 1), aes(x = 361, label = month2),
+      breaks = 0
+    ) +
+    geom_hline(
+      data = month_from_lag(1, 1),
+      aes(yintercept = lag),
+      colour = "gray50",
+      linewidth = 0.1
+    ) +
+
+    geom_text(
+      data = month_from_lag(15, 1),
+      aes(x = 361, label = month2),
       size = 2.5,
-              hjust = 0) +
-
-    geom_text(data = month_from_lag(1, 1), aes(x = -2, label = lag),
+      hjust = 0
+    ) +
+
+    geom_text(
+      data = month_from_lag(1, 1),
+      aes(x = -2, label = lag),
       size = 2.5,
-              hjust = 1) +
+      hjust = 1
+    ) +
     scale_fill_divergent_discretised(low = trans_blue, high = trans_pink) +
@@
     scale_y_continuous(breaks = NULL, expand = c(0, 0)) +
-    labs(fill = NULL, 
-         y = "Lead time (days)") +
+    labs(fill = NULL, y = "Lead time (days)") +
     wide_legend +
-    facet_wrap(~ month, labeller = labeller(month = labels_month)) +
-    theme(panel.background = element_blank(), panel.grid = element_blank(),
-          tagger.panel.tag.background = element_rect(colour = "white", fill = "white")) +
+    facet_wrap(~month, labeller = labeller(month = labels_month)) +
+    theme(
+      panel.background = element_blank(),
+      panel.grid = element_blank(),
+      tagger.panel.tag.background = element_rect(
+        colour = "white",
+        fill = "white"
+      )
+    ) +
     tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1))
@@
 <!-- ```{r errors_ensemble} -->
+
 <!-- errors_ensemble <- readRDS(here::here("data/derived/rmse.Rds")) |>  -->
+
 <!--   _[version != "persistence"] |>  -->
+
 <!--   _[member != "0em"] |>  -->
+
 <!--   _[obs_dataset == which_dataset] |>  -->
+
 <!--   _[value < 0.01, value := NA] |>  -->
+
 <!--   _[value > 3e13, value := NA] |>  -->
+
 <!--   _[!is.na(value)] |>  -->
+
 <!--   _[, time := as.Date(time)] |>  -->
+
 <!--   _[, lag := as.numeric(time - time_forecast)] |>  -->
+
 <!--   _[lag > 0] |>  -->
+
 <!--   _[!(month(time_forecast) == 1 & member == "07" & version == "S1")] -->
+
 <!-- ``` -->
@@
 <!-- ```{r fig-spread-errors} -->
+
 <!-- #| fig-cap: Total spread in RMSE, including ensemble spread and forecast spread, as a function of forecast lead time for ACCESS-S1 and ACCESS-S2 hindcasts.  -->
@@
 <!-- errors_ensemble |>  -->
+
 <!--   _[, var(value, na.rm = TRUE),  -->
+
 <!--     by = .(lag, version, measure, obs_dataset, time_forecast))] |>  -->
+
 <!--   ggplot(aes(lag, V1)) + -->
+
 <!--   geom_line(aes(color = version)) + -->
+
 <!--   scale_color_models + -->
+
 <!--   scale_y_continuous("RSME ensemble variance") + -->
+
 <!--   facet_wrap(~ month, labeller = labeller(month = labels_month))  -->
+
 <!-- ``` -->
@@
 which_lags <- c(1, 3, 30)
-rbind(mean_variance = errors_ensemble |>
+rbind(
+  mean_variance = errors_ensemble |>
     _[lag %in% which_lags] |>
@@
     _[, var(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |>
-        _[, .(value = mean(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
+    _[,
+      .(value = mean(V1, na.rm = TRUE)),
+      by = .(version, month(time_forecast), lag)
+    ],
 
@@
     _[obs_dataset == which_dataset] |>
-        _[, .(value = var(value, na.rm = TRUE)),
-          by = .(version, 
-                 month(time_forecast),
-                 lag)],
-      idcol = "component") |> 
+    _[,
+      .(value = var(value, na.rm = TRUE)),
+      by = .(version, month(time_forecast), lag)
+    ],
+  idcol = "component"
+) |>
   ggplot(aes(month, sqrt(value))) +
@@
   scale_color_models +
-  facet_grid(lag ~ component,
-             labeller = labeller(component = c(mean_variance = "Mean standard deviation of errors in ensemble members",
-                                               variance_of_mean = "Standard deviation of ensemble mean error."))) +
+  facet_grid(
+    lag ~ component,
+    labeller = labeller(
+      component = c(
+        mean_variance = "Mean standard deviation of errors in ensemble members",
+        variance_of_mean = "Standard deviation of ensemble mean error."
+      )
+    )
+  ) +
   tag_facets(tag = "rc")
@@
   # _[, month := lubridate::month(time_forecast, label = TRUE)]  |>
-  _[, median_ci(value), 
-    by = .(lag, version, measure, obs_dataset, month(time_forecast))] |> 
+  _[,
+    median_ci(value),
+    by = .(lag, version, measure, obs_dataset, month(time_forecast))
+  ] |>
   ggplot(aes(lag, estimate)) +
 
-  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
-             colour = "gray50", linewidth = 0.1) +
-
-  geom_text(data = month_from_lag(15, 2), aes(y = Inf, label = month2), 
+  geom_vline(
+    data = month_from_lag(1, 1),
+    aes(xintercept = lag),
+    colour = "gray50",
+    linewidth = 0.1
+  ) +
+
+  geom_text(
+    data = month_from_lag(15, 2),
+    aes(y = Inf, label = month2),
     size = 2.5,
-            vjust = 1) +
-
-  geom_text(data = month_from_lag(1, 2), aes(y = -Inf, label = lag), 
+    vjust = 1
+  ) +
+
+  geom_text(
+    data = month_from_lag(1, 2),
+    aes(y = -Inf, label = lag),
     size = 2.5,
-            vjust = 0) +
-
-  geom_ribbon(aes(ymin = low, 
-                  ymax = high, 
-                  color = version, 
-                  fill = version), alpha = 0.1) +
+    vjust = 0
+  ) +
+
+  geom_ribbon(
+    aes(ymin = low, ymax = high, color = version, fill = version),
+    alpha = 0.1
+  ) +
   geom_line(aes(color = version), linewidth = 1) +
@@
   scale_fill_models +
-  labs(y = NULL, 
-       x = "Lead time (days)") +
+  labs(y = NULL, x = "Lead time (days)") +
   #   scale_color_models +
   #   scale_fill_models +
-  facet_wrap(~ month, labeller = labeller(month = labels_month_significant)) 
+  facet_wrap(~month, labeller = labeller(month = labels_month_significant))
 ```
@@
 
-rbind(mean_variance = errors |>
+rbind(
+  mean_variance = errors |>
     _[measure == "iiee"] |>
@@
     _[, mean(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |>
-        _[, .(value = var(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
-      idcol = "component") |>
+    _[,
+      .(value = var(V1, na.rm = TRUE)),
+      by = .(version, month(time_forecast), lag)
+    ],
+  idcol = "component"
+) |>
 
@@
 
-  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
-             colour = "gray50", linewidth = 0.1) +
-
-  geom_text(data = month_from_lag(15, 2), aes(y = Inf, label = month2), 
+  geom_vline(
+    data = month_from_lag(1, 1),
+    aes(xintercept = lag),
+    colour = "gray50",
+    linewidth = 0.1
+  ) +
+
+  geom_text(
+    data = month_from_lag(15, 2),
+    aes(y = Inf, label = month2),
     size = 2.5,
-            vjust = 1) +
-
-  geom_text(data = month_from_lag(1, 2), aes(y = -Inf, label = lag), 
+    vjust = 1
+  ) +
+
+  geom_text(
+    data = month_from_lag(1, 2),
+    aes(y = -Inf, label = lag),
     size = 2.5,
-            vjust = 0) +
+    vjust = 0
+  ) +
 
   geom_line(aes(color = version), linewidth = 1) +
-  scale_y_continuous("Variance", 
-                     labels = labels_extent_2) +
+  scale_y_continuous("Variance", labels = labels_extent_2) +
   # scale_color_models +
@@
   scale_color_models +
-  ggh4x::facet_nested_wrap(month ~ component, ncol = 6, 
+  ggh4x::facet_nested_wrap(
+    month ~ component,
+    ncol = 6,
     remove_labels = "all",
-                           labeller = labeller(month = labels_month, 
-                                               component = c(mean_variance = "Mean variance",
-                                                             variance_of_mean = "Variance of mean"))) 
+    labeller = labeller(
+      month = labels_month,
+      component = c(
+        mean_variance = "Mean variance",
+        variance_of_mean = "Variance of mean"
+      )
+    )
+  )
 ```

### 2025-10-06 analysis/paper/access-ice.qmd
Updates data&methods

@@
 
-+S2 [@wedd2022] became operational in October 2021 by replacing the +S1 system [@hudson2017]..
-The model components of both +S2 and +S1 consist of the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0 [@best2011; @waters2017], Global Ocean 5.0 [@gurvan2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
-The atmosphere has a N216 horizontal resolution (\~60km in the mid-latitudes) with 85 vertical levels.
-The land model uses the same horizontal grid with 4 soil levels.
-The ocean component has a 1/4º resolution with 75 vertical levels.
-The sea ice component, based on CICE version 4.1, has the same resolution as the ocean and 5 sea ice categories as well as an open water category.
-
-Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for their hindcasts and from the Bureau's operational analysis for real-time forecasts.
-The main difference between the two are the ocean and sea ice initial conditions.
-+S1 ocean initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration.
-+S2, on the other hand, runs from initial conditions generated by the BoM data assimilation scheme described in @wedd2022.
-This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcasts and from the WMO Global Telecommunication System (GTS) and the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecasts.
-SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcasts and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecasts in areas where SSTs are over 0ºC.
-Relevant for this work, sea ice concentrations are not assimilated in +S2.
-
-The +S2 hindcast used in this study runs for the period 1981--2018.
-Each forecast consists of 9 ensemble members built from three consecutive 3-member forecasts initialised at the first of every month and the two previous days and run from 279 days.
-The +S1 hindcast is built in the same manner (albeit with 217-day runs) for the period 1990--2012.
-
-Anomalies for each hindcast are taken with respect to a climatology specific to each initalisation date for the period 1990--2012.
-This serves as a first-order correction of model drift.
+ACCESS-S2 [@wedd2022] is the Bureau of Meteorology's seasonal forecast system which became operational in October 2021, replacing the ACCESS-S1 system [@hudson2017].
+The model components of both ACCESS-S2 and ACCESS-S1 are identical with the same numbers of levels and resolution.
+They consist of the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], the Unified Model's Global Land 6.0 [@best2011; @waters2017], NEMO Global Ocean 5.0 [@gurvan2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
+The atmosphere has a N216 horizontal resolution (\~60 km in the mid-latitudes) with 85 vertical levels.
+The land model uses the same horizontal grid as the atmosphere with four soil levels.
+The ocean component has a nominal horizontal resolution of 1/4º with 75 vertical levels.
+The sea-ice component, based on CICE version 4.1, has the same resolution as the ocean component and five sea-ice thickness categories as well as an open water category.
+
+Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for their hindcasts.
+The main difference between the hindcasts of the two systems are the ocean and sea-ice initial conditions.
+ACCESS-S1 ocean initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea-ice concentration.
+ACCESS-S2, on the other hand, is initialised from ocean conditions generated by the BoM weakly coupled ensemble data assimilation scheme described in @wedd2022.
+This scheme uses an optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013].
+SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] in areas where SSTs are over 0ºC and Sea Surface Salinity is weakly nudged to the World Ocean Atlas 2013 climatology [@zweng2013].
+
+Of most relevance for this work, sea-ice concentrations are not assimilated in ACCESS-S2.
+Assimilation cycles are performed daily.
+The coupled model runs for 24 hours initialised from the previous cycle.
+Then the restart file fields of the ocean component are used as first guess in the data assimilation cycle and the innovations are used to build the next ocean initial conditions for the following cycle.
+The atmosphere fields from that daily integration are not used and instead the model atmosphere is initialised using ERA-Interim.
+The sea-ice initial conditions for the next cycle are the unaltered output of the previous daily integration.
+Then the cycle starts again and the coupled model runs for another 24 hours.
+During this integration the sea-ice component is affected by the ocean innovations and the new atmosphere initial conditions via the coupler.
+
+The ACCESS-S2 hindcast set used in this study runs for the period 1981–2018.
+Each hindcast consists of nine ensemble members built from three consecutive 3-member forecasts initialised at the first of every month and the two previous days and run for 279 days.
+The ACCESS-S1 hindcast set is built by perturbing the atmospheric fields only with a random field perturbation [@hudson2017] and runs for 217 days for the period 1990–2012.
+We analise the ensemble mean hindcasts unless otherwise specified.
+
+Anomalies for each hindcast set are taken with respect to their own climatology specific to each initialisation date and forecast lead time, for the period 1990–2012.
+This serves as a first-order correction of model bias and drift.
+
+Besides sea-ice concentration, we also analyse mean sea-ice thickness, which we compute as total sea-ice volume divided by total sea-ice area.
 
@@
 
-For verification we use satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.
-Different datasets derived using different algorithms provide different estimates, each with their own biases and uncertainties.
-@meier2019 estimated that the inter-product uncertainty of sea ice extent is of the order of 1 million $km^2$.
-As will be shown below, this spread is minimal compared with the typical errors in the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.
-
-We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset.
-It takes the maximum value of the NASA Team [@cavalieri1984] and NASA Bootstrap [@comiso2023] sea ice concentration products to reduce their low concentration bias [@meier2014; @meier2021].
+For verification we use satellite-derived sea-ice concentration, which estimates the proportion of each grid area that is covered with ice.
+Datasets derived using different algorithms and satellite platforms, each have their own biases and uncertainties.
+Estimates of inter-product uncertainty of sea-ice extent (SIE, defined here as the total region of the Southern Ocean with at least 15% sea-ice cover) is of the order of 0.5 million $km^2$ [@meier2019].
+As will be shown below, this spread is minimal compared with the typical errors in the +S2 and +S1 forecasts, so the overall conclusions of this study are independent of the verification dataset used.
+
+We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary sea-ice verification dataset.
+It takes the maximum value of the NASA Team [@cavalieri1984] and NASA Bootstrap [@comiso2023] sea-ice concentration products to reduce their low concentration bias [@meier2014; @meier2021].
 Both source algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
-The data has a resolution of 25 by 25 km and daily from 1978 onwards.
-
-The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [@OSISAF] is another satellite-derived sea ice concentration product.
+The data has a spatial resolution of 25 by 25 km and daily from 1978 onwards.
+
+The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility  [@OSISAF] based on the SSMIS sensor is another satellite-derived sea-ice concentration product.
 It is based on mostly the same sensors as the NOAA CDR but computed independently using different algorithms.
-Figures prepared with this dataset are provided in the supplementary material and don't differ significantly from the ones prepared using CDR.
+Figures prepared with this dataset are provided in the supplementary material and do not differ significantly from the ones prepared using CDR.
 
@@
 
-Sea-ice extent is defined as the area of the ocean covered with at least 15% ice.
-This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea ice conditions (cite -- probably something from NSIDC).
-
-Pan-Antarctic (net) sea ice extent serves as a rough hemispheric measure of the amount of sea ice, but it does not take into account the spatial distribution.
+Sea-ice extent is defined as the area of the ocean at least 15% covered by sea-ice. 
+This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea-ice concentrations [@cavalieri1991].
+
+Pan-Antarctic (net) sea-ice extent serves as a hemispheric measure of the amount of sea ice, but it does not take into account the spatial distribution. 
 A model could have a relatively accurate extent of the net ice but with different regional distributions. 
-To account for location errors, we computed the Root Mean Squared Error (RMSE) of sea ice concentration anomalies and the Integrated Ice Edge Error [IIEE; @goessling2016].
-
-We compute RMSE as the square root of the average squared differences between forecasted and observed sea ice concentration anomalies.
-We compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC CDRV4 southern hemisphere domain, and also a zonally-varying RMSE computed over twenty-four 15° longitude slices around Antarctica.
-
-The IIEE is defined as the area where the model miss-predicts sea ice concentration being above or below 15% ice.
-That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations.
-
-All error measures were computed on the NOAA/NSIDC CDRV4 domain grid and projection to which model output was bilinearly interpolated.
+To account for location errors, we computed the Root Mean Squared Error (RMSE) of grid-point sea-ice concentration anomalies.
+
+We compute RMSE as the square root of the area-averaged squared differences between grid-point forecasted and observed sea-ice concentration anomalies.
+We compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a zonally-varying RMSE computed over twenty-four 15° longitude slices around Antarctica.
+
+All error measures were computed on the NOAA/NSIDC CDRV4 domain grid, to which model output was bilinearly interpolated. 
+Note that the ACCESS CICE model grid has resolution between two and three times higher than NOAA/NSIDC CDRV4.
+
+Forecasts errors are also compared with hypothetical forecasts based on the persistence of anomalies. 
+The persistence forecast is generated by extending the observed anomalies in time.
 

### 2025-10-07 analysis/paper/access-ice.qmd
Modifies results

@@
   cache: true
+
 filters:
@@
 While the base model remained the same, the change in version was focused on using ocean, sea-ice and land initial conditions generated by the BoM instead of depending on the UK Met Office.
-Crucially, compared to ACCESS-S1, ACCESS-S2 does not assimilate sea-ice observations, so sea ice is only affected by the ocean and atmospheric data assimilation via the coupled integration.
-
-Since model configuration is identical between ACCESS-S1 and ACCESS-S2, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year, with the only difference being the initial conditions.
+Crucially, compared to +S1, +S2 does not assimilate sea-ice observations, so sea ice is only affected by the ocean and atmospheric data assimilation via the coupled integration.
+
+Since model configuration is identical between +S1 and +S2, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year, with the only difference being the initial conditions.
 This provides an opportunity to test the effect of sea-ice initial conditions on the forecast of sea-ice concentrations and the climate.
@@
 
-ACCESS-S2 [@wedd2022] is the Bureau of Meteorology's seasonal forecast system which became operational in October 2021, replacing the ACCESS-S1 system [@hudson2017].
-The model components of both ACCESS-S2 and ACCESS-S1 are identical with the same numbers of levels and resolution.
++S2 [@wedd2022] is the Bureau of Meteorology's seasonal forecast system which became operational in October 2021, replacing the +S1 system [@hudson2017].
+The model components of both +S2 and +S1 are identical with the same numbers of levels and resolution.
 They consist of the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], the Unified Model's Global Land 6.0 [@best2011; @waters2017], NEMO Global Ocean 5.0 [@gurvan2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
@@
 The main difference between the hindcasts of the two systems are the ocean and sea-ice initial conditions.
-ACCESS-S1 ocean initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea-ice concentration.
-ACCESS-S2, on the other hand, is initialised from ocean conditions generated by the BoM weakly coupled ensemble data assimilation scheme described in @wedd2022.
++S1 ocean initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea-ice concentration.
++S2, on the other hand, is initialised from ocean conditions generated by the BoM weakly coupled ensemble data assimilation scheme described in @wedd2022.
 This scheme uses an optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013].
@@
 
-Of most relevance for this work, sea-ice concentrations are not assimilated in ACCESS-S2.
+Of most relevance for this work, sea-ice concentrations are not assimilated in +S2.
 Assimilation cycles are performed daily.
@@
 
-The ACCESS-S2 hindcast set used in this study runs for the period 1981–2018.
+The +S2 hindcast set used in this study runs for the period 1981–2018.
 Each hindcast consists of nine ensemble members built from three consecutive 3-member forecasts initialised at the first of every month and the two previous days and run for 279 days.
-The ACCESS-S1 hindcast set is built by perturbing the atmospheric fields only with a random field perturbation [@hudson2017] and runs for 217 days for the period 1990–2012.
+The +S1 hindcast set is built by perturbing the atmospheric fields only with a random field perturbation [@hudson2017] and runs for 217 days for the period 1990–2012.
 We analise the ensemble mean hindcasts unless otherwise specified.
@@
 
-Forecasts errors are also compared with hypothetical forecasts based on the persistence of anomalies. 
+Forecasts errors are also compared with hypothetical forecasts based on the persistence of anomalies and on climatology. 
 The persistence forecast is generated by extending the observed anomalies in time.
+
+As a measure of forecast improvement over the hypothetical forecast, we use the skill score [@murphy1985], defined as
+
+$$
+S = 1 - \frac{RMSE_{f}}{RMSE_{r}}
+$$
+
+Where $RMSE_{f}$ is the RMSE of the forecast, $RMSE_{r}$ is the RMSE of the reference forecast. 
+Negative skill score indicates that the forecast is worse than the reference forecast while positive values indicate an improvement. 
+A perfect forecast, would have zero RMSE and thus a skill score of 1.
 
@@
 ```{r fig-hindcast-extent}
-#| fig-cap: Pan-Antarctic median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month with the median sea ice extent of +CDR in the left column and the median difference in the right column.
-The +CDR climatology is computed in the period corresponding to each hindcast.
-Circles represent the initial conditions at the first of every month and triangles represent the median value at the first of every month forecasted with the largest possible lead time.
-Colours indicate the initialisation month of the forecast.
+#| fig-cap: "Row a: Pan-Antarctic mean sea-ice extent for all hindcasts initialised on the first of each calendar month for +S1 (column 1; green) and +S1 (column 2; purple).
+Mean sea-ice extent from NOAA/NSIDC CDRV4 in each corresponding hindcast period is shown in black.
+Row b: Mean differences between the forecast and the observed values.
+Circles represent the initial conditions at the start of forecasts (i.e., the first of every month), and triangles represent the mean values at the end of forecasts (i.e., the longest possible lead time)."
 
@@
 
-@fig-hindcast-extent a shows the median sea ice extent (left column) and median difference with respect of +CDR for the +S2 and +S1 hindcasts (right column).
-Median extent of initial conditions at the first of every month are indicated with circles, while median extent at the same date but forecasted with the largest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
-At this large lead time, the the information of the initial conditions is essentially lost and the forecast reverts to each model's preferred equilibrium state.
-
-+S2 initial conditions (circles) show an overall negative bias, especially in the late summer-early autumn, while +S1 initial conditions are very close to observations.
-Both models' equilibrium state (triangles) show a negative bias of sea ice extent, particularly in the late-autumn and winter months.
-This is due primarily to faster melt during a longer melt season between January and March and slower growth during March and April.
-This is then partially balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
-Many sea ice models exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
-It is not surprising that both forecasting systems converge to a similar climatology because they share the same model formulation.
-
-The difference between the initial conditions and the model preferred state can be attributed to the effect of data assimilation, which in +S2 is due solely to atmospheric and oceanic data assimilation.
+@fig-hindcast-extent shows the mean sea-ice extent of the +S2 and +S1 hindcasts (row a) and their differences from the mean sea-ice extent of NOAA/NSIDC CDRV4 (row b).
+Mean extent at the first of every month is indicated with circles for the initial conditions and with triangles for the longest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
+At this long lead time, the information of the initial conditions is essentially lost and the forecast reverts to each model's preferred equilibrium state.
+
++S2 initial conditions (circles; @fig-hindcast-extent column 1) show an overall negative bias, especially in the late summer-early autumn, while +S1 initial conditions (circles in @fig-hindcast-extent column 2) are very close to observations, as expected from the assimilation of sea-ice observations to produce the initial conditions of +S1.
+Both systems' equilibrium states (triangles) show negative biases of sea-ice extent, particularly in the growth phase of late-autumn and winter months.
+This is due primarily to the melt season being longer than in observations and with faster melt between January and March and the growing seasons being shorter with slower growth during March and April.
+This is then followed by faster growth between May and July (@fig-mean-growth).
+Many sea-ice models exhibit this systematic underestimation during the sea-ice minimum and early freezing season [@massonnet2023], which could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
+It is also not surprising that both forecasting systems converge to a similar equilibrium state because they share the same model formulation.
+
+The difference between the initial conditions (circles) and the model equilibrium state (triangles) can be mostly attributed to the effect of data assimilation, which in +S2 is due solely to atmospheric and oceanic data assimilation.
 From May to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
-The rest of the year there is little if any difference between circles and triangles in +S2, indicating that the ocean and atmosphere data assimilation is not affecting sea ice and that this component of the model is virtually free-running.
+During these months, +S1 can overestimate the sea-ice extent at short lead time.
+For the rest of the year circles are overlaid with triangles in +S2, indicating that the ocean and atmosphere data assimilation is not affecting sea ice and that this component of the model is virtually free-running.
 
@@
 ```{r fig-mean-growth}
-#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations computed as the median daily differences in sea ice extent for each date and all lead times.
+#| fig-cap:  Mean daily sea-ice extent growth (10^6 km^2/day) in +S1 (green) and +S2 (purple) hindcasts and observations (black), computed as the mean daily differences in sea-ice extent between each date and the next for each forecast month.
 Values are smoothed with a 11-day running mean.
@@
 #| fig-height: 6
-#| fig-cap: Mean difference between monthly sea ice concentrations at 0-month lead time +S2 forecast and observations.
+#| fig-cap: Ensemble mean difference between monthly sea-ice concentration of ACCESS-S2 ensemble mean forecast at 0-month lead time (monthly mean values forecasted from the forecast initialised at the first of the month) and observations .
 #| fig-subcap:
@@
       aes(z = aice - obs, fill = after_stat(level)),
-      breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)
+      breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)(c(-0.999, 0.999))
     ) +
@@
     geom_antarctica_fill +
-    facet_wrap(~ month + model, labeller = labeller(month = labels_month)) +
+    facet_wrap(~month, labeller = labeller(month = labels_month)) +
     wide_legend +
@@
 
-@fig-bias shows the difference of monthly mean sea ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest lead time From October to May, the model underestimates sea ice concentrations in most regions except for the inner Weddell Sea in April and May, where sea ice concentrations saturate to 1 both in the observations and forecasts.
-In winter, the differences are mostly on the sea ice edge with light positive bias in the African sector of East Antarctica and negative bias around the Indian Ocean sector which partially compensate, resulting in the near-zero extent bias seen in those months ([Fig.
-@fig-hindcast-extent]).
+To further understand the bias in +S2, @fig-bias-1 shows spatial patterns of the differences of monthly mean sea-ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest monthly lead time.
+From October to May, the model underestimates sea-ice concentrations in most regions except for the inner Weddell Sea in April and May, where sea-ice concentrations saturate to 1 both in the observations and forecasts.
+In winter, the differences are limited to a narrow band around the sea-ice edge with slight positive biases in the African sector of East Antarctica and negative biases around the Indian Ocean sector which partially compensate, resulting in the near-zero extent bias seen in those months (@fig-hindcast-extent).
+
++S1 has a comparatively smaller overall bias (@fig-bias-2). 
+The largest values are found between April and June, when the faster growth resutls in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen seas.
 
@@
 ```{r fig-extent-anom}
-#| fig-cap: Monthly mean sea ice extent anomalies of the observation (black) and forecasts from +S1 (right column) and ACCESS-S2 (left column) at lead times of 0, 2, 4, and 6 months.
-The RMSE and correlation coefficient and their 95% confidence intervals during the overlapping period of ACCESS-S1 and S2 hindcasts (1990–2013) is shown on the top left and lower right corner of each panel.
+#| fig-cap: Monthly mean sea-ice extent anomalies of the observations (black) and forecasts from +S1 (right column; purple) and +S2(left column; green) at lead times of 0, 2, 4, and 6 months.
+The RMSE and correlation during the overlapping period of +S1 and +S2 hindcasts(1990–2013) is shown on the top left and bottom right of each panel respectively.
 
@@
     },
-    aes(label = text),
+    aes(label = paste0("RMSE: ", text)),
     x = -Inf,
@@
     },
-    aes(label = text),
+    aes(label = paste0("cor: ", text)),
     vjust = 0,
@@
 
-@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lead times.
-Compared with +S1, +S2 anomaly forecast is relatively poor (large RMSE) even for the first month (lead time 0), whereas +S1 has lower RMSE at 6 month lead time than +S2 has at zero lead time.
-+S2 shows much larger interannual variability than observations --particularly at short lead times-- with dramatic lows between 1995 and 2007, and highs between 2007 and 2015.
-
-Strangely, for +S2, RMSE improves with lead time, even though the correlation degrades with lead time.
+@fig-extent-anom shows monthly sea-ice extent anomalies forecasted at selected lead times.
+Compared with +S1, +S2 anomaly forecasts are relatively poor (large RMSE) even for the first month (lead time 0), whereas +S1 forecasts stay relatively skilful even at a lead time of three months.
++S2 shows much larger interannual variability than observations, with dramatic lows between 1995 and 2007, and highs between 2007 and 2015.
+
+Unexpectedly, for +S2, RMSE improves with lead time, even though the correlation degrades with lead time.
 This is puzzling behaviour that goes contrary to what is usually seen in prediction models.
 The explanation seems to be the mentioned increased interannual variability.
-@fig-extent-sd shows monthly mean sea ice extent of the forecasts as a function of lead time compared with observations.
-+S1 standard deviation lies within the observed regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at 9 month lead time.
+@fig-extent-sd shows monthly mean sea-ice extent of the forecasts as a function of lead time compared with observations.
++S1 standard deviation lies within the observed standard deviation regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at 9 month lead time for most months.
 
 ```{r fig-exent-sd}
-#| fig-cap: Standard deviation and 95% interval of monthly mean sea ice extent of observations and forecast for +S1 and +S2 at different lead times.
+#| fig-cap: Interannual standard deviation with 95% confidence interval of monthly mean sea-ice extent forecasted for each month divided by that month’s sea-ice extent observation standard deviation.
++S1 and +S2 at different lead times.
+Each panel indicates the target month.
+Note the reverse horizontal axis.
 
@@
 
-+S2 forecasts of sea-ice extent anomalies seem to align moderately well align with observations (leading to moderately high correlation) but their magnitude is overestimated (leading to large errors).
-This suggests that +S2 sea-ice is much more sensitive to atmospheric and oceanic forcing, perhaps due to different thickness.
-@fig-forecast-example shows sea-ice concentration anomalies (top row) and sea-ice thickness and the difference between the two models (bottom row) for the `r format(as.Date(sample_date), "%e of %B %Y")` initialised one day prior; being that close to initialisation date, these are very approximately the initial conditions.
++S2 forecasts of sea-ice extent anomalies seem to align moderately well with observations (leading to moderately high correlation) but their magnitude is overestimated (leading to large errors).
+This could be caused by +S2 sea ice being much more sensitive to atmospheric and oceanic forcing, perhaps due to lower thickness.
+
+As an example, @fig-forecast-exampleshows sea-ice concentration anomalies (top row) and sea-ice thickness and the difference between the two models (bottom row) for `r format(as.Date(sample_date), "%e %B %Y")` initialised one day prior; being that close to initialisation date, these are very approximately the initial conditions.
 +S1 sea-ice concentrations anomalies are very close to observations as expected from the system assimilating these data. 
-+S2 sea-ice concentration anomalies, which are not assimilated, are not as close, but the general pattern is not terrible.
++S2 sea-ice concentration anomalies, which are not assimilated, are not as close, but the large-scale pattern is aligned with observations.
 The system simulates large positive anomalies in the Weddell and Ross seas and slight negative anomalies in the Amundsen and Bellingshausen seas.
 The fact that +S2 can simulate this pattern without assimilating sea-ice data suggests that atmospheric and oceanic forcing was the dominant factor.
-However, the magnitude of these anomalies is too big.
-As shown in @fig-forecast-example bottom row, +S2's sea-ice was thinner than +S1's, even in regions of large positive se-ice concentration anomalies.
+However, the magnitude of the sea-ice anomalies is too big. It is plausible that this is due to the thinner ice simulated by +S2 (bottom row).
 
 ```{r fig-forecast-example}
-#| fig.cap: +S1 and +S2 hindcasts for the 2 of May 2008.
+#| fig.cap: +S1 and +S2 hindcasts for 2 May 2008 at one day lead time.
 Top row shows sea-ice concentration anomalies forecasted by each system and the observations.
@@
   plot_layout(ncol = 1)
-
-```
-
-The fact that +S1 and +S2 share the same model configuration and that the increased variability is more extreme at short lead times (@fig-exent-sd) suggests that this is an effect of the data assimilation.
-It is possible that sea-ice in the +S2 system is left in an unbalanced state after assimilating atmospheric and oceanic data but not sea-ice data, leading to large responses in the initial states which then subside at longer lead times when the model is balanced.
-
-```{r}
-giomas <- here::here("data/derived/giomas.csv") |>
-  fread() |>
-  _[, time := as.Date(time)] |>
-  _[, .(thickness = mean(thickness)), by = month(time)]
-```
-
-```{r}
+```
+
+```{r mean_thickness_funs}
 
@@
 }
-
 mean_thickness_access <- \(forecast_time, model) {
@@
 }
-
-```
-
-```{r}
-
-get_dir_part <- function(path, n = 1) {
-  if (n == 1) {
-    return(basename(path))
-  }
-
-  return(get_dir_part(dirname(path), n = n - 1))
-}
-
-get_model <- function(path) {
-  get_dir_part(path, 8)
-}
-
-file <- "sithick_SImon_MCM-UA-1-0_historical_r1i1p1f1_gn_185001-201412.nc"
-dates <- function(path) {
-  file <- get_dir_part(path, 1)
-  l <- 6 + 1 + 6 + 3
-  n <- nchar(file)
-  s <- n - l + 1
-
-  date_start <- substr(file, start = s, s + 5)
-  date_end <- substr(file, start = s + 7, n - 3)
-  list(start = lubridate::ym(date_start), end = lubridate::ym(date_end))
-}
-
-thick <- "/g/data/oi10/replicas/CMIP6/CMIP/*/*/historical/r1i1p1f1/SImon/sithick/gn/*/*" |>
-  Sys.glob()
-
-thick <- data.table(thickness = thick) |>
-  _[, model := get_model(thickness)] |>
-  _[, c("start", "end") := dates(thickness)] |>
-  _[]
-
-conc <- "/g/data/oi10/replicas/CMIP6/CMIP/*/*/historical/r1i1p1f1/SImon/siconc/gn/*/*" |>
-  Sys.glob()
-
-conc <- data.table(concentration = conc) |>
-  _[, model := get_model(concentration)] |>
-  _[, c("start", "end") := dates(concentration)]
-
-cmip <- merge(conc, thick, by = c("model", "start", "end"))
-
-```
-
-```{r}
+```
+
+```{r hindcast_thickness}
 hindcast_thickness <- CJ(model = c("S2", "S1")) |>
@@
 
-```{r}
-giomas_thickness <- "data/raw/giomas_thickness.nc" |>
-  cdo_selindexbox(1, 360, 1, 71) |>
-  cdo_setvrange(0, 500)
-
-giomas_concentration <- "data/raw/giomas_concentration.nc" |>
-  cdo_selindexbox(1, 360, 1, 71) |>
-  cdo_setvrange(0, 2)
-
-giomas_cellarea <- "data/raw/giomas_cell-area.nc" |>
-  cdo_selindexbox(1, 360, 1, 71)
-
-giomas <- mean_thickness(
-  giomas_concentration,
-  giomas_thickness,
-  giomas_cellarea
-) |>
-  cdo_ymonmean() |>
-  cdo_execute() |>
-  ReadNetCDF(c(thickness = "heff")) |>
-  _[, month := 1:12]
-
-```
-
-```{r}
+```{r hindcast_mean_thickness}
 hindcast_mean_thickness <- hindcast_thickness |>
@@
 ```{r}
-cmip |>
-  _[model %in% unique(model)[1:10]] |>
-  _[,
-    mean_thickness(
-      cdo_sellonlatbox(concentration, 0, 360, -90, -40),
-      cdo_sellonlatbox(thickness, 0, 360, -90, -40)
-    ),
-    by = .(model, start, end)
-  ] |>
-  _[, .(thickness = V1 |> cdo_mergetime() |> cdo_execute()), by = .(model)] |>
-  _[, ReadNetCDF(thickness, "sithick")[, .(time, sithick)], by = .(model)] |>
-  _[, .(sithick = mean(sithick)), by = .(model, month(time))] |>
-  ggplot(aes(month, sithick)) +
-  geom_line(aes(group = model, color = "CMIP6 models")) +
-  geom_line(
-    data = giomas,
-    aes(y = thickness, color = "GIOMAS"),
-    linewidth = 1.4
-  ) +
-  scale_color_manual(values = c("CMIP6 models" = "black", GIOMAS = "#1a7e21")) +
-  scale_y_continuous("Mean sea-ice thickness") +
-  scale_x_continuous(NULL, breaks = 1:12, labels = labels_month)
-```
-
-```{r}
+#| label: fig-mean-thickness
+#| fig-cap: Mean and 95% interval of monthly mean sea-ice thickness for +S1 and +S2 at different lead times. Each panel indicates the target month. Note the reverse horizontal axis.
 hindcast_mean_thickness[
@@
   facet_wrap(~month, labeller = labeller(month = labels_month))
-
-```
-
-```{r}
-hindcast_mean_thickness |>
-  _[, .(thickness = mean(thickness)), by = .(model, lag, month(time))] |>
-  _[lag %in% c(0)] |>
-  ggplot(aes(month, thickness)) +
-  geom_line(aes(color = model, group = interaction(model, lag))) +
-  geom_line(data = giomas) +
-  scale_color_models +
-  scale_y_continuous("Mean sea-ice thickness") +
-  scale_x_continuous(NULL, breaks = 1:12, labels = labels_month)
-```
-
-To assess +S2 forecasts quantitatively, we compute error measures for all hindcasts started on the 1st of every month.
+```
+
+Extending beyond the one case in @fig-forecast-example, @fig-mean-thickness shows monthly mean sea-ice thickness as a function of lead time for +S1 and +S2.
+Supporting the idea that thinner ice is what causes the increased extent variability in +S2, this system  simulates thinner sea-ice compared to +S1 overall at almost all lead times and in all months except for summer at short lead times (Dec-Jan, 0-1 months; Feb-Mar, 0-2 months).
+However, in both systems, forecasted sea-ice is thicker at shorter lead times and then decreases, particularly in the summer months.
+If thinner ice were a sufficient cause of increased variability, then we would expect variability to increase with lead time in both forecasting systems.
+
+The fact that +S1 and +S2 share the same model configuration and that the increased variability is more extreme at short lead times ([Fig.
+@fig-exent-sd]) suggests that the data assimilation procedure is partly to blame.
+It is possible that sea-ice in the +S2 system is left in an unbalanced state after assimilating atmospheric and oceanic data but not sea-ice data, leading to large responses that are amplified by the thin ice in the initial states which then subside at longer lead times when the model is balanced.
+
+To assess +S2 forecasts in more detail, we compute error measures for all hindcasts started on the 1st of every month.
 
@@
 ```{r fig-rmse}
-#| fig-cap: Median RMSE of sea ice concentration anomalies as a function of forecast lead time for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies (black) and climatology (gray).
+#| fig-cap: Mean RMSE of sea-ice concentration anomalies as a function of forecast lead time for all forecasts initialised on the first of each month compared with a reference forecast of persistence of anomalies (black) and climatology (gray).
 Only the first 120 days are shown.
@@
 
-@fig-rmse shows the median RMSE of sea ice concentration anomalies for +S2 and +S1 hindcasts compared with a benchmark of persistence and climatology.
-Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lead times, but that the persistence forecast errors would grow faster and eventually surpass the model forecast, after when the model is statistically useful.
-Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model does not have skill at any lead time and in any month.
-The only exception is around February, where the model has lower RMSE than the persistence forecast at virtually every lead time.
-
-```{r lead-time-window}
-#| fig-cap: Minimum lead time at which each forecast's mean RMSE becomes lower than the 95% confidence interval of persistence forecast RMSE (black lines) and maximum lead time at which each forecast's mean RMSE remains lower than the 95% confidence interval of climatological forecast RMSE (gray lines).
+@fig-rmse shows the mean RMSE of sea-ice concentration anomalies for +S1 and +S2 hindcasts compared against persistence and climatological forecasts used as a benchmark.
+Due to errors in the initial conditions, it is expected that persistence forecasts would be better than the model forecasts at very short lead times, but that the persistence forecast errors would grow faster and may eventually surpass the model forecast errors.
+The black line shows that the persistence forecast error indeed grows rapidly and reaches its maximum in about 30 days for most months except for February, when it grows much slower.
+The +S1 forecast errors grow slower than persistence forecast errors and remain lower after less than 10 days on average.
+The +S2 forecast error starts high in all months and is lower than the persistence forecast error after more than 15 days in most months except for forecast initialised in February, when it takes 80 days.
+
+At longer lead times, it is more appropriate to compare errors with the climatological forecast error. 
+The lead time at which +S1 forecast error is higher than the climatological forecast error varies between more than 60 and less than 20 days depending on forecast initialisation month with the minimum in June.
++S2 forecasts never have lower error than climatology, on the other hand, except marginally in October forecasts.
+
+@fig-lead-time-window summarises the lead time window in which each hindcast is better than both the persistence forecast and the climatological forecast as a function of forecast month. 
++S1 forecasts have a wider lead time window in the Summer than the other seasons and is not better than both benchmarks at forecasting June sea-ice concentration anomalies. 
+Forecasts initialised in May and June are particularly poor, and July cannot be forecasted better than the benchmarks. 
+This is consistent with the mid-winter loss of predictability observed by @libera2022, who attributed it to deep warm water entraining into the mixed layer.
+
+```{r fig-lead-time-window}
+#| fig-cap: Minimum lead time at which each forecast's mean RMSE becomes larger than the lower bound of the 95% confidence interval of persistence forecast RMSE (black lines) and maximum lead time at which each forecast's mean RMSE remains lower than the lower bound of the 95% confidence interval of climatological forecast RMSE (gray lines).
 Green shading indicates the window where forecasts outperform both persistence (lead times longer than black line) and climatology (lead times shorter than gray line).
@@
 
-```{r fig-rmse_lon, fig.height=7, fig.width=10}
-#| fig-cap: !expr glue::glue"Median difference between RMSE of +S2 forecasts and persistence forecast computed on {uniqueN(rmse_lon_mean$lon)} meridional slices {diff(unique(rmse_lon_mean$lon))[1]}º wide.
-Antarctica's coastline is shown at the bottom of each panel for reference.")
-#| fig-subcap:
-#|  - ACCESS-S1
-#|  - ACCESS-S2
-
-binwidth <- 0.015
+```{r rmse_lon-gdata}
+
 gdata <- rmse_lon_mean |>
@@
   _[,
-    .(dif = mean(value - climatology, na.rm = TRUE)),
+    .(
+      dif = mean(value - climatology, na.rm = TRUE),
+      skill = 1 - (mean(value, na.rm = TRUE) / mean(climatology, na.rm = TRUE))
+    ),
     by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))
@@
 
-```{r fig-rmse_lon, fig.height=7, fig.width=10}
+```{r, fig.height=7, fig.width=10}
+#| label: fig-rmse_lon
+#| fig-cap: !expr glue::glue("RMSE skill score of model forecasts with climatological forecast as reference computed on {uniqueN(rmse_lon_mean$lon)} meridional slices {diff(unique(rmse_lon_mean$lon))[1]}º wide as a function of lead time and longitude.
+Antarctica’s coastline is shown at the bottom of each panel for reference.")
+#| fig-subcap:
+#|  - +S1
+#|  - +S2
+binwidth <- 0.1
 lapply(c("S1", "S2"), \(x) {
@@
     _[version == x] |>
-    _[,
-      dif := Smooth2D(lag, lon, dif, method = smooth_svd(.1)),
-      by = .(month, version)
-    ] |>
+    # _[,
+    #   skill := Smooth2D(lag, lon, skill, method = smooth_svd(0.05)),
+    #   by = .(month, version)
+    # ] |>
     ggplot(aes(lon, lag)) +
     geom_contour_fill(
-      aes(z = dif, fill = after_stat(level)),
+      aes(z = skill, fill = after_stat(level)),
       breaks = AnchorBreaks(0, binwidth = binwidth)
@@
     geom_contour_tanaka(
-      aes(z = dif),
+      aes(z = skill),
       range = c(0.01, 0.2),
@@
 
-Even though +S2 is no better than persistence at forecasting pan-Antarctic sea ice concentration anomalies, the quality of the forecast might vary between regions.
-To analyse the spatial distribution of the model error, we computed sea ice concentration anomalies RMSE on `r uniqueN(rmse_lon_mean$lon)` meridional slices `r diff(unique(rmse_lon_mean$lon))[1]`º wide.
-The median difference between the forecast and persistence RMSE is shown on @fig-rmse_lon, where negative values indicate that the model has lower RMSE than the benchmark.
-
-The skill shown by +S2 at February-March forecasts is observed
-
-by a band of negative values across almost all longitudes.
-The only region where February-March sea ice is not well forecasted is east of the Antarctic Peninsula, which is the only region with consistent summer sea ice.
-This suggests that the positive skill around the summer minimum comes from the model correctly forecasting no ice where no ice is ever found.
-However, there are other regions of skilful forecasts.
-
-Forecasts initialised on January 1st ([Fig.
-@fig-rmse_lon] panel a) show RMSE values below persistence at lead times larger than 180 days (corresponding to July through September) in the Ross and Weddell Sea.
-These two regions are also relatively well forecasted from other months.
-Notably, May forecasts of Weddell sea ice concentration anomalies ([Fig. @fig-rmse_lon] panel e) are skilful after about 20 days and remain so until December.
-The Weddell Sea is also the region with the maximum error --around June regardless of initialisation date.
-
-<!-- ```{r errors_ensemble} -->
-
-<!-- errors_ensemble <- readRDS(here::here("data/derived/rmse.Rds")) |>  -->
-
-<!--   _[version != "persistence"] |>  -->
-
-<!--   _[member != "0em"] |>  -->
-
-<!--   _[obs_dataset == which_dataset] |>  -->
-
-<!--   _[value < 0.01, value := NA] |>  -->
-
-<!--   _[value > 3e13, value := NA] |>  -->
-
-<!--   _[!is.na(value)] |>  -->
-
-<!--   _[, time := as.Date(time)] |>  -->
-
-<!--   _[, lag := as.numeric(time - time_forecast)] |>  -->
-
-<!--   _[lag > 0] |>  -->
-
-<!--   _[!(month(time_forecast) == 1 & member == "07" & version == "S1")] -->
-
-<!-- ``` -->
-
-<!-- ```{r fig-spread-errors} -->
-
-<!-- #| fig-cap: Total spread in RMSE, including ensemble spread and forecast spread, as a function of forecast lead time for ACCESS-S1 and ACCESS-S2 hindcasts.  -->
-
-<!-- errors_ensemble |>  -->
-
-<!--   _[, var(value, na.rm = TRUE),  -->
-
-<!--     by = .(lag, version, measure, obs_dataset, time_forecast))] |>  -->
-
-<!--   ggplot(aes(lag, V1)) + -->
-
-<!--   geom_line(aes(color = version)) + -->
-
-<!--   scale_color_models + -->
-
-<!--   scale_y_continuous("RSME ensemble variance") + -->
-
-<!--   facet_wrap(~ month, labeller = labeller(month = labels_month))  -->
-
-<!-- ``` -->
+To analyse the spatial distribution of the model error, we computed the RMSE of zonal mean sea-ice concentration anomalies on `r uniqueN(rmse_lon_mean$lon)` slices of `diff(unique(rmse_lon_mean$lon))[1]`° longitude span for each forecasting system.
+We control for some areas being naturally easier to forecast than others by computing the RMSE skill score with the climatological forecast RMSE as reference.
+
+For +S1 forecasts (@fig-rmse_lon-1), skill tends to be lower off the coast of Eastern Antarctica even at short lead times; for instance, the skill score for forecasts initialised in May and June are negative between 0° and 120°E even at almost zero lead time.
+In western Antarctica there is a hint of easterly-propagating skill in forecasts initialised in February and March. 
+This is consistent with @holland2013 findings that memory of sea-ice anomalies are stored in ocean heat content anomalies that are transported east by the Antarctic Circumpolar Current.
+
++S2 forecasts  (@fig-rmse_lon-2) also have lower skill over Eastern Antarctica. 
+From July to December even though the pan-Antarctic average skill is negative at all lead times ([Fig. @fig-lead-time-window]), it is positive for up to a month in Western Antarctica. 
+Since oceanic and atmospheric forcing is the only source of information, this suggests that sea-ice in this region is particularly sensitive to oceanic and atmospheric forcing and suggests a role of the Pacific-South American mode and the Amundsen Sea Low to shape sea-ice concentration anomalies.
+The fact that this is evident in the months in which ENSO teleconnections are more important for atmospheric circulation also suggest the influence of tropical Pacific variability.
+February and March are the only two months that can be forecasted with marginally positive skill in large regions.
+
+```{r errors_ensemble} 
+errors_ensemble <- readRDS(here::here("data/derived/rmse.Rds")) |>
+  _[version != "persistence"] |>
+  _[member != "0em"] |>
+  _[obs_dataset == which_dataset] |>
+  _[value < 0.01, value := NA] |>
+  _[value > 3e13, value := NA] |>
+  _[!is.na(value)] |>
+  _[, time := as.Date(time)] |>
+  _[, lag := as.numeric(time - time_forecast)] |>
+  _[lag > 0] |>
+  _[!(month(time_forecast) == 1 & member == "07" & version == "S1")]
+```
 
 ```{r fig-initial-spread}
-#| fig-cap: Decomposition of forecast error spread at one-day lead time for ACCESS-S1 and ACCESS-S2 hindcasts across initialization months.
-The top panel shows the mean variance of RMSE errors across ensemble members, while the bottom panel shows the variance of the ensemble mean RMSE error.
+#| fig-cap: Decomposition of forecast error spread at 1, 5 and 30 days lead time for +S1 and +S2 hindcasts across initialization months.
+The left column shows the mean standard deviation of RMSE errors across ensemble members, while the right column shows the standard deviation of the ensemble mean RMSE error and the spread of the persistence and climatology forecasts errors.
 
@@
 
-Also evident in @fig-initial-spread is the difference in the error spread at short lead times between +S2 and +S1.
-In all months +S1 has much narrower error spread at the first day of the forecasts.
-The error in the initial conditions is not only small, but also fairly constant.
-This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
-For +S2, this is true only between July and October.
-For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it is not constant.
-
-```{r fig-iiee}
-#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lead time for all forecast initialised on the first of each month for +S1 and +S2 hindcasts.
-For each month, the number in parenthesis indicates the minimum lead time at which the mean error of each model is not statistically significant at a 1% level using a two-sided t-test.
-
-errors |>
-  _[measure == "iiee"] |>
-  _[obs_dataset == which_dataset] |>
-  # _[, month := lubridate::month(time_forecast, label = TRUE)]  |>
-  _[,
-    median_ci(value),
-    by = .(lag, version, measure, obs_dataset, month(time_forecast))
-  ] |>
-  ggplot(aes(lag, estimate)) +
-
-  geom_vline(
-    data = month_from_lag(1, 1),
-    aes(xintercept = lag),
-    colour = "gray50",
-    linewidth = 0.1
-  ) +
-
-  geom_text(
-    data = month_from_lag(15, 2),
-    aes(y = Inf, label = month2),
-    size = 2.5,
-    vjust = 1
-  ) +
-
-  geom_text(
-    data = month_from_lag(1, 2),
-    aes(y = -Inf, label = lag),
-    size = 2.5,
-    vjust = 0
-  ) +
-
-  geom_ribbon(
-    aes(ymin = low, ymax = high, color = version, fill = version),
-    alpha = 0.1
-  ) +
-  geom_line(aes(color = version), linewidth = 1) +
-
-  # geom_errorbar(data = \(x) x[dif, on = .NATURAL][version == "S1"],
-  #              aes(x = lag,
-  #                  ymin = estimate - (high - low), ymax = estimate + (high - low))) +
-
-  # geom_point(data = \(x) x[dif, on = .NATURAL][version == "S1"],
-  # aes(lag, estimate), size = 3, shape = 21) +
-
-  scale_y_continuous(labels = labels_extent) +
-  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
-  scale_color_models +
-  scale_fill_models +
-  labs(y = NULL, x = "Lead time (days)") +
-  #   scale_color_models +
-  #   scale_fill_models +
-  facet_wrap(~month, labeller = labeller(month = labels_month_significant))
-```
-
-To compare +S2 with +S1, we computed the IIEE for both models.
-This error measure is shown in @fig-iiee for forecasts initialised at the first of every month at all lead times.
-+S1 has lower error at lead times up to 60 days at all months, with the errors converging to +S2 as the forecast goes on.
-The time to convergence depends on the month and it can be as short as two weeks June and July to as long as 160 days for forecasts initialised in February.
-Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the memory of sea ice to initial conditions --at least from October to March when the data assimilated from the other components has little to no influence on sea ice.
-
-Also evident in @fig-iiee is the difference in the error spread at short lead times between +S2 and +S1.
-In all months +S1 has much narrower error spread at the first day of the forecasts.
-The error in the initial conditions is not only small, but also fairly constant.
-This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
-For +S2, this is true only between July and October.
-For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it is not constant.
-
-```{r fig-iiee-variance}
-#| fig-cap: Mean spread of IIEE at different lead times for different models.
-#| fig.height: 6
-#| fig.width: 8
-labels_extent_2 <- function(x) {
-  m <- which.max(x)
-  x <- scales::label_number(scale = (1e-12)^2)(x)
-  x[m] <- paste0(x[m], "\nM km⁴")
-  x
-}
-
-rbind(
-  mean_variance = errors |>
-    _[measure == "iiee"] |>
-    _[obs_dataset == which_dataset] |>
-    _[, var(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |>
-    _[,
-      .(value = mean(V1, na.rm = TRUE)),
-      by = .(version, month(time_forecast), lag)
-    ],
-
-  variance_of_mean = errors |>
-    _[measure == "iiee"] |>
-    _[obs_dataset == which_dataset] |>
-    _[, mean(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |>
-    _[,
-      .(value = var(V1, na.rm = TRUE)),
-      by = .(version, month(time_forecast), lag)
-    ],
-  idcol = "component"
-) |>
-
-  ggplot(aes(lag, value)) +
-
-  geom_vline(
-    data = month_from_lag(1, 1),
-    aes(xintercept = lag),
-    colour = "gray50",
-    linewidth = 0.1
-  ) +
-
-  geom_text(
-    data = month_from_lag(15, 2),
-    aes(y = Inf, label = month2),
-    size = 2.5,
-    vjust = 1
-  ) +
-
-  geom_text(
-    data = month_from_lag(1, 2),
-    aes(y = -Inf, label = lag),
-    size = 2.5,
-    vjust = 0
-  ) +
-
-  geom_line(aes(color = version), linewidth = 1) +
-  scale_y_continuous("Variance", labels = labels_extent_2) +
-  # scale_color_models +
-  scale_x_continuous("Lead time (days)", breaks = NULL) +
-  scale_color_models +
-  ggh4x::facet_nested_wrap(
-    month ~ component,
-    ncol = 6,
-    remove_labels = "all",
-    labeller = labeller(
-      month = labels_month,
-      component = c(
-        mean_variance = "Mean variance",
-        variance_of_mean = "Variance of mean"
-      )
-    )
-  )
-```
-
-The large initial error spread could be due either to large spread of ensemble members or to large spread of individual forecasts.
-@fig-iiee-variance splits the IIEE variance for each lead time into the mean variance of each individual forecast and the variance of the mean error of each individual forecast, which adds up to the total variance.
-The average variance of each forecast is almost identical between the two forecast systems in all months.
-This shows that the ensemble spread of individual forecasts evolves identically, which, again, is not unexpected because both systems share the same model formulation.
-This also shows that the perturbation scheme in +S2 is comparable to the one in +S1.
-
-On the other hand, the spread of the mean error is always larger in +S2 than +S1.
-The difference is particularly large at short lead times in some months, which coincide with the ones in which the data assimilation scheme is not influencing sea ice initial conditions.
+In Figure 5 the mean error was shown. 
+@fig-initial-spread column 1 shows the mean standard deviation of errors among ensemble members at various lead times.
+At one day lead time ([Fig. @fig-initial-spread] a.1) +S2 has a slightly larger spread than +S1 due to the way that ensemble members are generated.
++S1 ensemble members are generated by adding random field perturbations to the atmosphere only, which then are transferred to the other components via the coupled simulation [@hudson2017]. 
+With this scheme, ensemble members are all but guaranteed to be underdispersed in the ocean and sea-ice components. The time-lag ensemble used for +S2 ensures greater spread. 
+This difference is gone after about just two days, and both systems have a comparable spread in ensemble member error afterwards ([Fig. @fig-initial-spread] b1 and c1).
+
+@fig-initial-spread column 2, on the other hand, shows the standard deviation of ensemble mean error of each hindcast and the persistence forecast. 
+At one day lead time, +S2 ensemble mean error standard deviation is much larger than +S1’s, which in turn is comparable to the persistence forecast error standard deviation. 
+At longer lead times, the spread of +S1 and persistence forecast standard deviation increases to eventually be comparable to +S2 and the standard deviation in climatological forecast errors. 
++S2 error standard deviation is fairly independent of lead time and similar to the climatological forecast error standard deviation at all lead times.
 

### 2025-10-07 R/ggplot.R
Hardcodes sic_projection (quarto was failing for some reason otherwise)

@@ -4,9 +4,9 @@ element_text_first_line <- function(...) {
   el
 }
 
-# Thanks to the wizard Teun van den Brand 
+# Thanks to the wizard Teun van den Brand
 # https://fosstodon.org/@teunbrand/113248074175175782
-element_grob.element_text_first_line <- 
+element_grob.element_text_first_line <-
   function(element, label, x = NULL, y = NULL, ...) {
     twolines <- grepl("\n", label)
     y <- y - grid::unit(twolines * 0.5, "lines")
@@ -14,21 +14,23 @@ element_grob.element_text_first_line <-
   }
 
 
-ggplot2::theme_set(ggplot2::theme_minimal() +
-                     ggplot2::theme(panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
-                                    legend.position = "bottom",
-                                    legend.title.position = "top", 
-                                    legend.title = ggplot2::element_text(hjust = 0.5),
-                                    legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
-                                    legend.key.height = grid::unit(0.75, "lines"),
-                                    axis.text.y = element_text_first_line()
-                     ))
+ggplot2::theme_set(
+  ggplot2::theme_minimal() +
+    ggplot2::theme(
+      panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
+      legend.position = "bottom",
+      legend.title.position = "top",
+      legend.title = ggplot2::element_text(hjust = 0.5),
+      legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
+      legend.key.height = grid::unit(0.75, "lines"),
+      axis.text.y = element_text_first_line()
+    )
+)
 wide_legend <- ggplot2::theme(legend.key.width = grid::unit(1, 'null'))
 
 
 trans_pink <- scales::muted("#F7A8B8", l = 60, c = 100)
 trans_blue <- scales::muted("#55CDFC", l = 60, c = 100)
-
 
 
 colours_models <- c(
@@ -53,14 +55,16 @@ labels_models <- c(
   climatology = "Climatology"
 )
 
-scale_color_models <- ggplot2::scale_color_manual(NULL,
-                                                  values = colours_models,
[…]

### 2025-10-07 analysis/paper/access-ice.qmd
Hardcodes sic_projection (quarto was failing for some reason otherwise)
Updates conclusions

@@
 
-```{r mirai_daemons}
+```{r mirai_daemons, cache = FALSE}
 if (!mirai::daemons_set()) {
@@
 For +S1 forecasts (@fig-rmse_lon-1), skill tends to be lower off the coast of Eastern Antarctica even at short lead times; for instance, the skill score for forecasts initialised in May and June are negative between 0° and 120°E even at almost zero lead time.
+This mirrors @libera2022 findings of a “winter predictability barrier”, although they focus on the Weddell sea and here we show that the effect seems to be stronger more to the East. 
 In western Antarctica there is a hint of easterly-propagating skill in forecasts initialised in February and March. 
@@
 
-Sea ice forecasts from the +S2 model show a significant low extent bias, particularly during late summer and early autumn.
+Sea-ice forecasts from the +S2 system show a significant low extent bias, particularly during late summer and early autumn. 
 This bias is attributed to a faster and longer melt season between January and March, and slower growth between March and April. 
-This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems either with the model's thermodynamic representation or with short wave radiation forcing, as shown in other climate models [@zampieri2019; @roach2020].
-Even though +S2 shares the same model formulation with +S1, the latter does not suffer from this bias.
-This is due to the assimilation of sea ice concentrations into the initial conditions, which successfully corrects for the negative bias in the model.
-Our analysis suggests that the data assimilation system in +S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea ice component runs virtually free, reverting to its biased equilibrium state.
+This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems either with the model’s thermodynamic representation or with short wave radiation forcing, as shown in other climate models [@zampieri2019; @roach2020].
+Even though +S2 shares the same model formulation with +S1, the latter does not suffer from this bias, indicating that assimilating sea-ice concentrations successfully corrects for the negative bias that exists in the free-running model.
+
+Ensemble spread grows quickly even when perturbations are only implemented in the atmosphere component (in +S1), indicating that sea-ice is indeed responding quickly to atmospheric perturbations. 
+However, our analysis suggests that the atmosphere and ocean data assimilation implemented in +S2 is only effectively influencing sea-ice initial conditions from June to October, while the rest of the year, the sea-ice component runs virtually free, reverting to its biased equilibrium state.
 
@@
 
-Based on the observation that +S2 sea ice initial conditions are essentially not initialised, comparing its forecasts with +S1's allows us to estimate the time-scale for which initial conditions are important.
-This highlights February initial conditions as crucial for determining sea ice evolution at least up to late June.
-Arctic sea ice forecasts also show greater sensitivity to initial conditions in boreal summer compared with boreal winter [@day2014; @bunzel2016], so a similar mechanism might be playing a role.
-
-Although forecasts are not skilful for forecasting pan-Antarctic sea ice concentrations, there are some areas where the model does show skill.
-The Weddell Sea, and the Ross Sea to a lesser extent are particularly well forecasted between June and November for forecasts initialised from January to June.
-These regions have been recognised as a region of high predictability thanks to persistent and eastwardly advected upper ocean heat content anomalies [@bushuk2021].
-Sea surface temperatures are assimilated by +S2 only in areas with temperatures greater than 0°C and errors don't show a clear easterly-propagating signal, so it is not clear if +S2 is leveraging the same source of predictability..
+Although +S1 only assimilates sea-ice concentration, it is clear that sea-ice thickness is also affected. 
++S1 simulates significantly thicker ice than +S2 and in both systems sea-ice is thicker at shorter lead times than at longer lead times. 
+Both the explicit data assimilation in +S2 and the effects of atmospheric and oceanic data assimilation in +S1 might be nudging simulated sea ice to be thicker than the model equilimbrium state.  
+We believe that the thinner sea-ice in +S2 is a partial cause for the large sea-ice extent variance, but other mechanisms, such as unbalanced initial conditions might be important.
+
+Given that +S2 sea-ice extent is essentially not initialised by the observations, comparing its forecasts with those of +S1 allows us to estimate the time-scale for which initial conditions are important.
+We find that initial conditions affect Antarctic sea-ice forecasts in ther order of a few months, but that effect is seasonally dependent. 
+January to April initial conditions improve forecasts for up to 3 months with, February initial conditions in particular are shown to be crucial for determining sea-ice evolution at least up to May. 
+Arctic sea-ice forecasts also show greater sensitivity to initial conditions in boreal Summer, compared with boreal Winter [@day2014; @bunzel2016], suggesting a similar mechanism might be playing a role.
+
+Forecasts initialised in winter have very little skill and +S1 and +S2 forecast errors are not statistically different after just 2 weeks. 
+This is consistent with @libera2022's finding of a “winter predictability barrier” in the Weddell sea, although they describe the barrier as sharp loss of predictability in July, and here we find a gradual reduction in skill compared with climatology around June.
+This difference might be due to our use of pan-Antarctic RMSE, since our regional analysis indicates that the degraded skill is most dramatic in the King Haakon sea.
+
+These findings have important implications for both operational forecasting, model development and predictability studies.
+For operational centers, our results suggest that efforts to improve sea-ice data assimilation should prioritize the summer and autumn months when initial conditions have the greatest impact on forecast skill.
+Additionally, the substantial bias in +S2 highlights the need for improved model physics, particularly in the representation of sea-ice thermodynamics and radiation processes. 
+Crucially, our results suggest dramatic seasonal variations in sea-ice predictability. 
+Future studies should therefore use initial conditions through the whole year rather than focusing on a limited number of initialisation dates.
 

### 2025-10-07 analysis/paper/access-ice.qmd
Uses Nature template

@@
 ---
-editor:
-  markdown:
-    wrap: sentence
-title: Evaluation of +S2 sea ice forecast and some other better title
-pdf-engine: pdflatex
+title: Quarto Template for Springer Nature
+format:
+  nature-pdf:
+    # journal.cite-style is included in the tex file but ignored by pandoc if 
+    # cite-method is not `natbib`.
     journal:
-  name: tc
+      cite-style: sn-nature
+    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
+    # to use the bst styles from the upstream template.
+    cite-method: natbib
+    keep-tex: true
+    equal-margins: false
+    classoption: referee
 author:
-  - given_name: Elio 
-    surname: Campitelli
-    affiliation: "1,2"
-    email: elio.campitelli@monash.edu
+  - name: Author One
+    affiliations:
+      - name: Harvard University
+        id: 1
+        department: Department of Government
+        address: 1737 Cambridge Street
+        city: Cambridge
+        state: MA
+        postal-code: 02138
+      - name: Harvard University
+        id: 2
+        department: Department of Statistics
+        address: 1 Oxford Street
+        city: Cambridge
+        state: MA
+        postal-code: 02138
+    attributes:
+      equal-contributor: true
+  - name: Author Two
+    affiliations:
+      - name: Yale University
+        id: 3
+        department: Department of Political Science
+        address: 115 Prospect Street
+        city: New Haven
+        state: CT
+        postal-code: 06511
+    email: corresponding@email.com
+    attributes:
+      equal-contributor: true
       corresponding: true
-  - given_name: Ariaan 
-    surname: Purich
-    affiliation: "1,2"
-  - given_name: Julie
-    surname: Arblaster
-    affiliation: "1,2"
-  - given_name: Eun-Pa
-    surname: Lim
-    affiliation: "3"
-  - given_name: Matthew
-    surname: Wheeler
-    affiliation: "3"
-  - given_name: Phillip
-    surname: Reid
-    affiliation: "3"
-
-affiliation:
-  - code: 1
-    address: School of Earth, Atmosphere and Environment, Monash University, Australia
-  - code: 2
-    address: Securing Antarctica’s Environmental Future, Monash University, Australia
-  - code: 3
-    address: Bureau of Meteorology
-
-abstract: The abstract goes here.
-bibliography: references.bib
-running:
-  title: +S2 sea ice forecast
-  author: Campitelli et al.
-# This section is mandatory even if you declare that no competing interests are present.
-competinginterests: The authors declare no competing interests.
-# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
-# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
-copyrightstatement: The author's copyright for this publication is transferred to institution/company. 
-### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
-### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
-### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
-availability:
-  #code: |
-  #  use this to add a statement when having only software code available
-  #data: |
-  #  use this to add a statement when having only data sets available
-  codedata: use this to add a statement when having data sets and software code available
-  sample: use this section when having geoscientific samples available
-  videosupplement: use this section when having video supplements available
-  authorcontribution: Contributions
-  disclaimer: We like Copernicus.
-  acknowledgements: Acknowledgements will gohere
-  appendix: |
-    \section{Figures and tables in appendices}
-    Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
-    \subsection{Option 1}
-    If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
-    They will be correctly named automatically.
-    \subsection{Option 2}
-    If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
-
-    To rename them correctly to A1, A2, etc., please add the following commands in front of them:
-    `\appendixfigures` needs to be added in front of appendix figures
-    `\appendixtables` needs to be added in front of appendix tables
-
-    Please add `\clearpage` between each table and/or figure.
-
-format: pdf
+  - name: Author Three
+    affiliations:
+      - ref: 1
+abstract: |
+  The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
+Authors are advised to check the author instructions for the journal they are submitting to for word limits and if structural elements like subheadings, citations, or equations are permitted.
+keywords: [template, demo]
+bibliography: bibliography.bib
 
@@
   +CDR: NOAA/NSIDC CDRV4
+
 ---

### 2025-10-08 analysis/paper/access-ice.qmd
Reading for final version

@@
 ---
-title: Quarto Template for Springer Nature
+title: Evaluating Seasonal Predictability of Antarctic Sea Ice and the importance of initial conditions with fully coupled models
 format:
@@
 author:
-  - name: Author One
+  - name: Elio Campitelli
     affiliations:
-      - name: Harvard University
+      - name: Monash University, Australia,
         id: 1
-        department: Department of Government
-        address: 1737 Cambridge Street
-        city: Cambridge
-        state: MA
-        postal-code: 02138
-      - name: Harvard University
+        department: School of Earth, Atmosphere and Environment
+        # address: 1737 Cambridge Street
+        # city: Cambridge
+        # state: MA
+        # postal-code: 02138
+      - name: Monash University, Australia
         id: 2
-        department: Department of Statistics
-        address: 1 Oxford Street
-        city: Cambridge
-        state: MA
-        postal-code: 02138
+        department: Securing Antarctica’s Environmental Future
+        # address: 1 Oxford Street
+        # city: Cambridge
+        # state: MA
+        # postal-code: 02138
+    email: elio.campitelli@monash.edu
     attributes:
-      equal-contributor: true
-  - name: Author Two
-    affiliations:
-      - name: Yale University
-        id: 3
-        department: Department of Political Science
-        address: 115 Prospect Street
-        city: New Haven
-        state: CT
-        postal-code: 06511
-    email: corresponding@email.com
-    attributes:
-      equal-contributor: true
       corresponding: true
-  - name: Author Three
+  - name: Ariaan Purich
     affiliations:
       - ref: 1
-abstract: |
-  The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications.
-Authors are advised to check the author instructions for the journal they are submitting to for word limits and if structural elements like subheadings, citations, or equations are permitted.
-keywords: [template, demo]
-bibliography: bibliography.bib
+      - ref: 2
+  - name: Julie Arblaster
+    affiliations:
+      - ref: 1
+      - ref: 2
+  - name: Eun-Pa Lim
+    affiliations:
+      - department: Research, Bureau of Meteorology
+        id: 3
+  - name: Mathew Wheeler
+    affiliations:
+      - ref: 3      
+  - name: Phillip Reid
+    affiliations:
+      - ref: 3      
+keywords: [sea ice, predictabilty, initial conditions, forecasting]
+bibliography: references.bib
 
@@
 filters:
-  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
+  - _extensions/ute/search-replace/search-replace.lua
+  - _extensions/pandoc-ext/abstract-section/abstract-section.lua
   # - ../../_extensions/mloubout/critic-markup/critic-markup.lua
-  - ../../_extensions/ute/search-replace/search-replace.lua
+
 search-replace:
@@
 
-balbalbal
+Accurate Antarctic sea-ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
+We investigate the role of initial conditions on Antarctic sea-ice forecasts by comparing hindcasts from two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system: +S1 and +S2.
+They use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations. 
+Our analysis reveals that both systems experience an extended melt season and delayed growth phase. 
+This leads to a significant negative sea-ice extent bias which is corrected only in +S1 by the data assimilation system. 
+The impact of the initial conditions on forecast errors varies dramatically by season: summer and autumn initial conditions (January-April) provide predictive skill for up to three months, with February initial conditions being particularly crucial.
+In contrast, winter forecasts off the two systems are statistically indistinguishable after only two weeks. 
+Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over Eastern Antarctica, where evern +S1 shows negative skill. 
+These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritize the summer-autumn period when initial conditions have maximum impact on forecast skill.
 
@@
 
-```{r fig-exent-sd}
+```{r fig-extent-sd}
 #| fig-cap: Interannual standard deviation with 95% confidence interval of monthly mean sea-ice extent forecasted for each month divided by that month’s sea-ice extent observation standard deviation.
@@
 
-As an example, @fig-forecast-exampleshows sea-ice concentration anomalies (top row) and sea-ice thickness and the difference between the two models (bottom row) for `r format(as.Date(sample_date), "%e %B %Y")` initialised one day prior; being that close to initialisation date, these are very approximately the initial conditions.
+As an example, @fig-forecast-example shows sea-ice concentration anomalies (top row) and sea-ice thickness and the difference between the two models (bottom row) for `r format(as.Date(sample_date), "%e %B %Y")` initialised one day prior; being that close to initialisation date, these are very approximately the initial conditions.
 +S1 sea-ice concentrations anomalies are very close to observations as expected from the system assimilating these data. 
@@
 The fact that +S1 and +S2 share the same model configuration and that the increased variability is more extreme at short lead times ([Fig.
-@fig-exent-sd]) suggests that the data assimilation procedure is partly to blame.
+@fig-extent-sd]) suggests that the data assimilation procedure is partly to blame.
 It is possible that sea-ice in the +S2 system is left in an unbalanced state after assimilating atmospheric and oceanic data but not sea-ice data, leading to large responses that are amplified by the thin ice in the initial states which then subside at longer lead times when the model is balanced.
@@
 Only the first 120 days are shown.
+In parenthersis, the shortest time at which +S1 and +S2 mean RMSE is not statistically different at the 99% confidence level.
 #| fig.height: 5
@@
 @fig-lead-time-window summarises the lead time window in which each hindcast is better than both the persistence forecast and the climatological forecast as a function of forecast month. 
-+S1 forecasts have a wider lead time window in the Summer than the other seasons and is not better than both benchmarks at forecasting June sea-ice concentration anomalies. 
++S1 forecasts have a wider lead time window in the summer than the other seasons and is not better than both benchmarks at forecasting June sea-ice concentration anomalies. 
 Forecasts initialised in May and June are particularly poor, and July cannot be forecasted better than the benchmarks. 
@@
 January to April initial conditions improve forecasts for up to 3 months with, February initial conditions in particular are shown to be crucial for determining sea-ice evolution at least up to May. 
-Arctic sea-ice forecasts also show greater sensitivity to initial conditions in boreal Summer, compared with boreal Winter [@day2014; @bunzel2016], suggesting a similar mechanism might be playing a role.
+Arctic sea-ice forecasts also show greater sensitivity to initial conditions in boreal summer, compared with boreal winter [@day2014; @bunzel2016], suggesting a similar mechanism might be playing a role.
 

### 2025-10-08 analysis/paper/access-ice.qmd
Formats figures

@@
     journal:
-      cite-style: sn-nature
+      cite-style: sn-mathphys-num
     # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
@@
     classoption: referee
+    fig-width: 5  
 author:
@@
   +CDR: NOAA/NSIDC CDRV4
-
 ---
@@
 Accurate seasonal to sub-seasonal forecasts are also crucial for operation contingency planning in and around the Antarctic continent, including scientific missions, fisheries, and tourism [@desilva2020; @wagner2020].
-Improvements in modelled sea-ice might also help improve weather forecasts over and away from sea-ice regions [@rinke2006; @wang2024 @semmler2016].
+Improvements in modelled sea-ice might also help improve weather forecasts over and away from sea-ice regions [@rinke2006; @wang2024; @semmler2016].
 
@@
 Circles represent the initial conditions at the start of forecasts (i.e., the first of every month), and triangles represent the mean values at the end of forecasts (i.e., the longest possible lead time)."
+#| fig-height: 4
 
@@
     NULL,
-    date_breaks = "1 month",
+    date_breaks = "2 month",
     date_labels = "%b",
@@
 Values are smoothed with a 11-day running mean.
+#| fig-height: 4
 
@@
 
-```{r fig-bias}
-#| fig-width: 6
-#| fig-height: 6
-#| fig-cap: Ensemble mean difference between monthly sea-ice concentration of ACCESS-S2 ensemble mean forecast at 0-month lead time (monthly mean values forecasted from the forecast initialised at the first of the month) and observations .
-#| fig-subcap:
-#|   - "+S1"
-#|   - "+S2"
-#| layout-ncol: 1
-
-# ice_shelves_nsicd <- fread(here::here("data/derived/antarctic_ice_shelves_nsidc.csv"))
-
-s <- lapply(c("S1", "S2"), \(x) {
-  g <- hindcast_clim |>
+```{r}
+#| label: plot_bias
+plot_bias <- lapply(c("S2", "S1"), \(x) {
+  hindcast_clim |>
     _[model == x] |>
@@
     tag_facets()
-
-  print(g)
 })
 
+```
+
+```{r fig-bias-1}
+#| fig-cap: Ensemble mean difference between monthly sea-ice concentration of +S2 ensemble mean forecast at 0-month lead time (monthly mean values forecasted from the forecast initialised at the first of the month) and observations.
+#| fig-height: 6
+
+plot_bias[[1]]
+```
+
+```{r fig-bias-2}
+#| fig-cap: Same as @fig-bias-1 but for +S1
+#| fig-height: 6
+
+plot_bias[[2]]
 ```
@@
 ```{r fig-extent-anom}
-#| fig-cap: Monthly mean sea-ice extent anomalies of the observations (black) and forecasts from +S1 (right column; purple) and +S2(left column; green) at lead times of 0, 2, 4, and 6 months.
-The RMSE and correlation during the overlapping period of +S1 and +S2 hindcasts(1990–2013) is shown on the top left and bottom right of each panel respectively.
+#| fig-cap: Monthly mean sea-ice extent anomalies of the observations (black) and forecasts from +S1 (right column; purple) and +S2 (left column; green) at lead times of 0, 2, 4, and 6 months.
+The RMSE and correlation during the overlapping period of +S1 and +S2 hindcasts (1990–2013) is shown on the top left and bottom left of each panel respectively.
+#| fig-height: 6
 
@@
     y = Inf,
+    size = 3,
     vjust = 1.2,
@@
     aes(label = paste0("cor: ", text)),
-    vjust = 0,
-    hjust = 1.2,
-    x = Inf,
+    vjust = -0.8,
+    hjust = 0,
+    size = 3,
+    x = -Inf,
     y = -Inf
@@
 Note the reverse horizontal axis.
+#| fig-height: 5
 
@@
 Bottom row shows forecasted sea-ice thickness and the difference between +S1 and +S2.
-#| fig.height: 7
-#| fig.width: 5.5
+#| fig-height: 6
+
 b <- seq(-0.5, 0.5, by = 0.1)
@@
 #| fig-cap: Mean and 95% interval of monthly mean sea-ice thickness for +S1 and +S2 at different lead times. Each panel indicates the target month. Note the reverse horizontal axis.
+#| fig-height: 5
+
 hindcast_mean_thickness[
@@
 In parenthersis, the shortest time at which +S1 and +S2 mean RMSE is not statistically different at the 99% confidence level.
-#| fig.height: 5
-#| fig.width: 9
+#| fig-height: 5
 
@@
   scale_fill_models +
-  facet_wrap(~month, labeller = labeller(month = labels_month_significant)) +
+  facet_wrap(
+    ~month,
+    labeller = labeller(month = labels_month_significant),
+    ncol = 3
+  ) +
   coord_cartesian(clip = "off")
@@
 Text labels show the date corresponding to the maximum lead time at which each forecast outperforms climatology.
+#| fig-height: 4
 
@@
   # geom_point(aes(color = time), data = \(x) x[lag != 0]) +
-  geom_text(
+  ggrepel::geom_text_repel(
     data = \(x) {
@@
 ```{r, fig.height=7, fig.width=10}
-#| label: fig-rmse_lon
-#| fig-cap: !expr glue::glue("RMSE skill score of model forecasts with climatological forecast as reference computed on {uniqueN(rmse_lon_mean$lon)} meridional slices {diff(unique(rmse_lon_mean$lon))[1]}º wide as a function of lead time and longitude.
-Antarctica’s coastline is shown at the bottom of each panel for reference.")
-#| fig-subcap:
-#|  - +S1
-#|  - +S2
+#| label: rmse_lon_plots
 binwidth <- 0.1
-lapply(c("S1", "S2"), \(x) {
+rmse_lon_plots <- lapply(c("S1", "S2"), \(x) {
   g <- gdata |>
@@
       colour = "gray50",
-      linewidth = 0.1
+      alpha = 0.5,
+      linewidth = 0.2
     ) +
@@
     wide_legend +
-    facet_wrap(~month, labeller = labeller(month = labels_month)) +
+    facet_wrap(~month, labeller = labeller(month = labels_month), ncol = 3) +
     theme(
@@
     tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1))
-
-  print(g)
-}) -> sink
+})
+```
+
+```{r}
+#| label: fig-rmse_lon-1
+#| fig-cap: !expr glue::glue("RMSE skill score of +S1 forecasts with climatological forecast as reference computed on {uniqueN(rmse_lon_mean$lon)} meridional slices {diff(unique(rmse_lon_mean$lon))[1]}º wide as a function of lead time and longitude.
+Antarctica’s coastline is shown at the bottom of each panel for reference.")
+#| fig-height: 7
+rmse_lon_plots[[1]]
+
+```
+
+```{r}
+#| label: fig-rmse_lon-2
+#| fig-cap: Same as @fig-rmse_lon-1 but for +S2
+#| fig-height: 7
+rmse_lon_plots[[2]]
 ```
@@
 The left column shows the mean standard deviation of RMSE errors across ensemble members, while the right column shows the standard deviation of the ensemble mean RMSE error and the spread of the persistence and climatology forecasts errors.
-
+#| fig-height: 4
 which_lags <- c(1, 3, 30)
@@
   # scale_color_models +
-  scale_x_continuous("Month", breaks = 1:12, labels = month.abb, ) +
+  scale_x_continuous("Month", breaks = seq(1, 12, by = 2), labels = \(x) {
+    month.abb[x]
+  }) +
   scale_color_models +
@@
       component = c(
-        mean_variance = "Mean standard deviation of errors in ensemble members",
-        variance_of_mean = "Standard deviation of ensemble mean error."
+        mean_variance = "Mean standard deviation\nof errors in ensemble members",
+        variance_of_mean = "Standard deviation\nof ensemble mean error."
       )

### 2025-10-08 R/ggplot.R
Adds scripts

@@ -84,18 +84,18 @@ topo <- here::here("data/raw/ETOPO.nc") |>
   metR::ReadNetCDF("z") |>
   _[, .(x = xgrid, y = ygrid, z)]
 
-# get_proj <- function(x) {
-#   x <- ncdf4::nc_open(x)
-#   on.exit(ncdf4::nc_close(x))
-#   x |>
-#     ncdf4::ncatt_get(varid = "crs") |>
-#     _[["proj_params"]]
-# }
+get_proj <- function(x) {
+  x <- ncdf4::nc_open(x)
+  on.exit(ncdf4::nc_close(x))
+  x |>
+    ncdf4::ncatt_get(varid = "crs") |>
+    _[["proj_params"]]
+}
 
-# sic_projection <- CDR() |>
-#   get_proj()
+sic_projection <- CDR() |>
+  get_proj()
 
-sic_projection <- "+proj=stere +lat_0=-90 +lat_ts=-70 +lon_0=0 +k=1 +x_0=0 +y_0=0 +a=6378273 +b=6356889.449 +units=m +no_defs"
+# sic_projection <- "+proj=stere +lat_0=-90 +lat_ts=-70 +lon_0=0 +k=1 +x_0=0 +y_0=0 +a=6378273 +b=6356889.449 +units=m +no_defs"
 
 contour_antarctica <- ggplot2::StatContour$compute_group(topo, breaks = 0)
 geom_antarctica_path <- ggplot2::geom_path(
@@ -114,13 +114,12 @@ geom_antarctica_fill <- ggplot2::geom_polygon(
 )
 
 geomcoord_antarctica <- list(
-  NULL,
+  ggplot2::coord_equal(),
   # ggplot2::coord_sf(
   #   crs = sic_projection,
   #   lims_method = "box",
   #   label_axes = "----"
   # ),
   ggplot2::scale_x_continuous(name = NULL, expand = c(0, 0), labels = NULL),
-  ggplot2::scale_y_continuous(name = NULL, expand = c(0, 0), labels = NULL),
-  geom_antarctica_path
+  ggplot2::scale_y_continuous(name = NULL, expand = c(0, 0), labels = NULL)
 )

### 2025-10-08 scripts/pbs/build_ensemble.sh

@@ -4,12 +4,12 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=03:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
 #PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
 #PBS -N build_ensemble
 
 # Load module, always specify version number.

### 2025-10-08 scripts/pbs/build_ensemble1.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble1
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=1
+Rscript scripts/build_ensemble.R > logs/build_ensemble1.log

### 2025-10-08 scripts/pbs/build_ensemble10.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble10
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=10
+Rscript scripts/build_ensemble.R > logs/build_ensemble10.log

### 2025-10-08 scripts/pbs/build_ensemble2.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble2
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=2
+Rscript scripts/build_ensemble.R > logs/build_ensemble2.log

### 2025-10-08 scripts/pbs/build_ensemble3.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble3
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=3
+Rscript scripts/build_ensemble.R > logs/build_ensemble3.log

### 2025-10-08 scripts/pbs/build_ensemble4.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble4
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=4
+Rscript scripts/build_ensemble.R > logs/build_ensemble4.log

### 2025-10-08 scripts/pbs/build_ensemble5.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble5
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=5
+Rscript scripts/build_ensemble.R > logs/build_ensemble5.log

### 2025-10-08 scripts/pbs/build_ensemble6.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble6
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=6
+Rscript scripts/build_ensemble.R > logs/build_ensemble6.log

### 2025-10-08 scripts/pbs/build_ensemble7.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble7
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=7
+Rscript scripts/build_ensemble.R > logs/build_ensemble7.log

### 2025-10-08 scripts/pbs/build_ensemble8.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble8
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=8
+Rscript scripts/build_ensemble.R > logs/build_ensemble8.log

### 2025-10-08 scripts/pbs/build_ensemble9.sh

@@ -1,16 +1,16 @@
 #!/bin/bash
 
-#PBS -P lo70
+#PBS -P gb02
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=08:00:00
+#PBS -l walltime=12:00:00
 #PBS -l wd
-#PBS -l jobfs=20GB
+#PBS -l jobfs=400GB
 #PBS -o logs/
 #PBS -e logs/
-#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo70+scratch/lo70
+#PBS -N build_ensemble9
 
 # Load module, always specify version number.
 module load R/4.3.1
@@ -21,5 +21,5 @@ module load cdo/2.4.3
 # https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
-Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log
+export CHUNK=9
+Rscript scripts/build_ensemble.R > logs/build_ensemble9.log

### 2025-10-08 scripts/pbs/ensemble_mean.sh

@@ -10,16 +10,13 @@
 #PBS -o logs/
 #PBS -e logs/
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N hindcast_climatolgy
+#PBS -N ensemble_mean
 
 # Load module, always specify version number.
 module load R/4.3.1
 module load cdo/2.4.3
 
-# Must include `#PBS -l storage=scratch/ab12+gdata/yz98` if the job
-# needs access to `/scratch/ab12/` and `/g/data/yz98/`. Details on:
-# https://opus.nci.org.au/display/Help/PBS+Directives+Explained
 
 # Run R application
-SCRIPT=hindcast_climatology.R
+SCRIPT=ensemble_mean.R
 Rscript scripts/${SCRIPT} > logs/${SCRIPT}.log

### 2025-10-08 scripts/pbs/error_metrics1.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics1
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=1
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics10.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics10
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=10
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics2.sh

@@ -4,7 +4,7 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07

### 2025-10-08 scripts/pbs/error_metrics3.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics3
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=3
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics4.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics4
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=4
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics5.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics5
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=5
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics6.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics6
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=6
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics7.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics7
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=7
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics8.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics8
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=8
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics9.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics9
 #PBS -o logs/
 #PBS -e logs/
 
@@ -18,5 +18,5 @@ module load cdo/2.4.3
 
 # Run R application
 export PBS_WORKERS=35
-export CHUNK=2
+export CHUNK=9
 Rscript scripts/error_metrics.R 

### 2025-10-08 scripts/pbs/error_metrics_read.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=03:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N error_metrics_read
 #PBS -o logs/
 #PBS -e logs/
 
@@ -17,6 +17,5 @@ module load R/4.3.1
 module load cdo/2.4.3
 
 # Run R application
-export PBS_WORKERS=35
-export CHUNK=2
-Rscript scripts/error_metrics.R 
+export PBS_WORKERS=20
+Rscript scripts/error_metrics_read.R 

### 2025-10-08 scripts/pbs/thickness.sh

@@ -4,11 +4,11 @@
 #PBS -q normal
 #PBS -l ncpus=48
 #PBS -l mem=64GB
-#PBS -l walltime=20:00:00
+#PBS -l walltime=3:00:00
 #PBS -l wd
 #PBS -l jobfs=100GB
 #PBS -l storage=gdata/ux62+scratch/k10+gdata/k10+gdata/ub7+gdata/rt52+gdata/dx2+gdata/lo70+gdata/hh5+gdata/lo07
-#PBS -N error_metrics2
+#PBS -N thickness
 #PBS -o logs/
 #PBS -e logs/
 
@@ -17,6 +17,5 @@ module load R/4.3.1
 module load cdo/2.4.3
 
 # Run R application
-export PBS_WORKERS=35
-export CHUNK=2
-Rscript scripts/error_metrics.R 
+export PBS_WORKERS=40
+Rscript scripts/thickness.R 

### 2025-10-08 analysis/paper/access-ice.tex
Redners paper

@@
+% Options for packages loaded elsewhere
 % Options for packages loaded elsewhere
@@
 %
-\documentclass[tc, manuscript]{copernicus}
+\documentclass[
+  sn-mathphys-num,
+  referee]{sn-jnl}
 
+\usepackage{xcolor}
 \usepackage{amsmath,amssymb}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
@@
 \makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
@@
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 
-\usepackage[inline]{trackchanges}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+%%%% Standard Packages
+
+\usepackage{graphicx}%
+\usepackage{multirow}%
+\usepackage{amsmath,amssymb,amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithm}%
+\usepackage{algorithmicx}%
+\usepackage{algpseudocode}%
+\usepackage{listings}%
+
+%%%%
+
+\raggedbottom
 \makeatletter
@@
 \makeatother
-
 \usepackage{bookmark}
-
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
-  pdftitle={Evaluation of +S2 sea ice forecast and some other better title},
+  pdftitle={Evaluating Seasonal Predictability of Antarctic Sea Ice and the importance of initial conditions with fully coupled models},
+  pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Mathew Wheeler; Phillip Reid},
+  pdfkeywords={sea ice, predictabilty, initial conditions, forecasting},
   colorlinks=true,
@@
   citecolor={Blue},
-  urlcolor={red},
+  urlcolor={Blue},
   pdfcreator={LaTeX via pandoc}}
 
-\title{Evaluation of +S2 sea ice forecast and some other better title}
+\title[Evaluating Seasonal Predictability of Antarctic Sea Ice and the
+importance of initial conditions with fully coupled models]{Evaluating
+Seasonal Predictability of Antarctic Sea Ice and the importance of
+initial conditions with fully coupled models}
 
-\Author[1,2]{Elio}{Campitelli}
-\Author[1,2]{Ariaan}{Purich}
-\Author[1,2]{Julie}{Arblaster}
-\Author[3]{Eun-Pa}{Lim}
-\Author[3]{Matthew}{Wheeler}
-\Author[3]{Phillip}{Reid}
+% author setup
+\author*[1,2]{\fnm{Elio} \sur{Campitelli}}\email{elio.campitelli@monash.edu}\author[1,2]{\fnm{Ariaan} \sur{Purich}}\author[1,2]{\fnm{Julie} \sur{Arblaster}}\author[3]{\fnm{Eun-Pa} \sur{Lim}}\author[3]{\fnm{Mathew} \sur{Wheeler}}\author[3]{\fnm{Phillip} \sur{Reid}}
+% affil setup
+\affil[1]{\orgdiv{School of Earth, Atmosphere and
+Environment}, \orgname{Monash University, Australia,}}
+\affil[2]{\orgdiv{Securing Antarctica's Environmental
+Future}, \orgname{Monash University, Australia}}
+\affil[3]{\orgdiv{Research, Bureau of Meteorology}}
 
-\affil[1]{School of Earth, Atmosphere and Environment, Monash
-University, Australia}
-\affil[2]{Securing Antarctica's Environmental Future, Monash University,
-Australia}
-\affil[3]{Bureau of Meteorology}
+% abstract
 
-\runningtitle{+S2 sea ice forecast}
+\abstract{Accurate Antarctic sea-ice forecasts are crucial for climate
+monitoring and operational planning, yet they remain challenging due to
+model biases and complex ice-ocean-atmosphere interactions. We
+investigate the role of initial conditions on Antarctic sea-ice
+forecasts by comparing hindcasts from two versions of the Australian
+Bureau of Meteorology's ACCESS seasonal forecast system: ACCESS-S1 and
+ACCESS-S2. They use identical model configuration and differ only in
+their initial conditions; primarily in that ACCESS-S2 does not
+assimilate sea-ice observations. Our analysis reveals that both systems
+experience an extended melt season and delayed growth phase. This leads
+to a significant negative sea-ice extent bias which is corrected only in
+ACCESS-S1 by the data assimilation system. The impact of the initial
+conditions on forecast errors varies dramatically by season: summer and
+autumn initial conditions (January-April) provide predictive skill for
+up to three months, with February initial conditions being particularly
+crucial. In contrast, winter forecasts off the two systems are
+statistically indistinguishable after only two weeks. Regional analysis
+of forecast skill suggests that this winter predictability barrier is
+most dramatic over Eastern Antarctica, where evern ACCESS-S1 shows
+negative skill. These findings highlight the critical importance of
+comprehensive year-round sampling in predictability studies and suggest
+that operational sea-ice data assimilation efforts should prioritize the
+summer-autumn period when initial conditions have maximum impact on
+forecast skill.}
 
-\runningauthor{Campitelli et al.}
-
-\correspondence{Elio\ Campitelli\ (elio.campitelli@monash.edu)}
+% keywords
+\keywords{sea ice,  predictabilty,  initial conditions,  forecasting}
 
 \begin{document}
-
-\received{}
-\pubdiscuss{} %% only important for two-stage journals
-\revised{}
-\accepted{}
-\published{}
-
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
-
-\firstpage{1}
 \maketitle
-\begin{abstract}
-The abstract goes here. It can also be on \emph{multiple lines}.
-\end{abstract}
-\copyrightstatement{The author's copyright for this publication is
-transferred to institution/company.}
 
@@
 
-Unlike Arctic sea ice, which has been steadily retreating at least since
-the start of satellite records in the early 80s, Antarctic sea ice has
-experienced a slightly increasing trend up to 2015, which models
-systematically failed to reproduce and puzzled researchers
-\citep{hobbs2016}. Then, in 2016 Antarctic sea ice extent dropped
-precipitously \citep{turner2017a} and has been at low and record low
-levels since \citep{purich2023, hobs2024, gilbert2024, josey2024}, even
-more highlighting our lack of understanding of Antarctic sea ice
-variability and long-term change \citep{espinosa2024}.
+Accurately modelling Antarctic sea ice is essential for understanding
+processes and improving climate projections to inform adaptation
+strategies. Accurate seasonal to sub-seasonal forecasts are also crucial
+for operation contingency planning in and around the Antarctic
+continent, including scientific missions, fisheries, and tourism
+\citep{desilva2020, wagner2020}. Improvements in modelled sea-ice might
+also help improve weather forecasts over and away from sea-ice regions
+\citep{rinke2006, wang2024, semmler2016}.
 
-Antarctic sea ice covers the polar ocean, yet the potential impacts of
-its variability extend far into the mid and lower latitudes, affecting
-the circulation of both the ocean and the atmosphere, ocean carbon
-uptake and biological processes \citep[\citet{england2020},
-\citet{swadling2023}]{fogwill2020}. In coupled climate model
-experiments, a reduction in sea ice extent has been shown to reduce the
-meridional temperature gradient, in turn reducing the eddy-driven jet
-strength and shifting it equatorward \citetext{\citealp[
-\citet{england2020}]{england2018}; \citealp{ayres2019}; \citealp{ayres2022}}
-and vice versa for an increase in sea ice
-\citep{raphael2003a, raphael2011}, which is a crucial control on
-extratropical weather. In the Northern Hemisphere, it has been
-hypothesised that the reduction in Arctic sea ice has led to a weaker,
-wavier jet, increasing the frequency of extreme events
-\citep{barnes2015, smith2022b}. There is some evidence of this effect in
-the Southern Hemisphere. For example, model simulations show a weakened
-vortex in the middle atmosphere, increased vertical wave flux and
-greater zonal wave 1 amplitude in austral spring following lower than
-average sea ice extent at in early winter
-\citep[\citet{song2025}]{rea2024}. But models different models disagree
-on the exact phase of the zonal wave response, and observations do not
-show a clear signal \citep{rea2024}. The magnitude of the eddy-driven
-jet response to sea ice loss has been shown to be model dependent, and
-is sensitive to its climatological location \citep{ayres2019}. Further,
-sea ice induced changes in the jet have been found to be relatively
-small compared with the jet-response to the direct effect of CO2
-increase and global sea surface temperature (SST) warming
-\citep{ayres2022}.
+However, producing accurate Antarctic sea-ice forecasts has been
+challenging due to model biases, inherent large variability and
+complexity, and they have lagged behind Arctic sea-ice
+forecasts\citep{zampieri2019, gao2024a}. Dynamical seasonal forecasts of
+summer Antarctic sea ice have been shown to perform worse than
+relatively simple statistical methods \citep{massonnet2023} and machine
+learning approaches \citep[e.g.][]{dong2024, lin2025}, which also
+underscores the need for better understanding of sea-ice dynamics, and
+drivers of its variability.
 
-Correctly modelling Antarctic sea is not only necessary for process
-understanding and climate projections to inform adaptation strategies
-but also for accurate seasonal to sub-seasonal forecasts, which are
-crucial for operations in and around the Antarctic continent such as
-scientific missions, fisheries and tourism \citep[
-\citet{wagner2020}]{desilva2020}. However, producing accurate Antarctic
-sea ice forecasts has been challenging due to model biases, inherent
-large variability and complexity, and it has lagged behind Arctic sea
-ice forecasts \citep{zampieri2019}. Dynamical seasonal forecasts of
-summer Antarctic sea ice have been shown to perform worse than
-relatively simple statistical methods \citep{massonnet2023}, which also
-underscores the need for better understanding of sea ice dynamics, and
-drivers.
+Good initial conditions are generally required for a good forecast,
+however, it is not entirely known to what extent accurate sea-ice
+initial conditions affect the quality of the forecast and at what
+timescales. Exploring seasonal predictions of Arctic sea-ice,
+\citet{guemas2016} found that sea-ice initial conditions are important
+in autumn to predict summer sea-ice, but the impact wasn't as dramatic
+when predicting winter sea-ice. \citet{day2014} also found
+seasonally-varying differences in the effect of initialisation, noting
+that accurate Arctic sea-ice thickness lead to improved sea-ice
+forecasts initialised in July but not when initialised in January.
 
-In this work we evaluate sea ice forecasts produced by the Australian
-Community Climate and Earth System Simulator -- seasonal version2
-(ACCESS-S2), which is the Australian Bureau of Meteorology (BoM)'s
-current operational subseasonal-to-seasonal prediction system, and
-compare its forecast skill with that of the former system ACCESS-S1.
-This evaluation will inform future work with the prediction system as a
-research tool and explore the potential of using its sea-ice forecasts
-for decision-making.
+For the Antarctic, \citet{holland2013} studied the initial-value
+predictability of Antarctic sea-ice in a perfect model study using the
+CCSM3 model. They found that sea-ice and ocean initial conditions
+provide predictive information to forecast sea-ice edge location several
+months in advance and that some predictability is retained for up to two
+years thanks to ocean heat content anomalies that are advected eastward.
+This is in contrast with \citet{marchi2020}, who ran perfect model
+experiments to argue that uncertainty in the predicted atmospheric state
+and evolution is the main driver of uncertainty in Antarctic sea-ice
+extent prediction on seasonal timescales, with sea-ice and ocean initial
+conditions having lesser importance. More recently, \citet{morioka2022}
+studied decadal forecasts of Antarctic sea-ice and found that
+initialising ocean and sea-ice improved the correlation between
+simulated and observed sea-ice concentration evolution in the
+Amundsen--Bellingshausen Sea. It is hard to compare these studies since
+they are based on forecasts initialised at different times of the year
+and different frameworks: \citet{holland2013} ran 20 ensemble members
+initialised on the 1st of January of a particular year,
+\citet{marchi2020} ran forecasts from the 1st of March and 1st of
+September and \citet{morioka2022} run forecasts only from the 1st of
+March. \citet{marchi2020} also used a coupled ocean--sea-ice model
+instead of a fully coupled model like \citet{holland2013} did.
+\citet{morioka2022} used observed sea ice initial conditions and
+compared with observations, while \citet{marchi2020} and
+\citet{holland2013} are perfect model studies.
+
+In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the
+Australian Community Climate and Earth System Simulator -- Seasonal
+(ACCES-S) from version S1 to S2. While the base model remained the same,
+the change in version was focused on using ocean, sea-ice and land
+initial conditions generated by the BoM instead of depending on the UK
+Met Office. Crucially, compared to ACCESS-S1, ACCESS-S2 does not
+assimilate sea-ice observations, so sea ice is only affected by the
+ocean and atmospheric data assimilation via the coupled integration.
+
+Since model configuration is identical between ACCESS-S1 and ACCESS-S2,
+they form a sort of ``natural experiment'' where the same forecasting
+model was run over a long period of time with multiple ensemble
+forecasts initialised throughout the year, with the only difference
+being the initial conditions. This provides an opportunity to test the
+effect of sea-ice initial conditions on the forecast of sea-ice
+concentrations and the climate.
+
+In this study we compare sea-ice hindcasts produced by ACCESS-S1 and
+ACCESS-S2. We focus on seasonality of errors and biases and the effect
+of the data assimilation system. This comparison will also inform future
+work with the prediction system as a research tool to better understand
+the dynamics and variability of the Antarctic sea ice and its impacts on
+the climate system as well as to explore the potential of using its
+sea-ice forecasts for decision-making. The work will also serve as a
+benchmark for future prediction systems to attempt to improve upon.
 
@@
 
-ACCESS-S2 \citep{wedd2022} became operational in October 2021 by
-replacing the ACCESS-S1 system \citep{hudson2017}.. The model components
-of both ACCESS-S2 and ACCESS-S1 consist of the Global Atmosphere 6.0
-(GA6) \citep{williams2015, waters2017}, Global Land 6.0
-\citep{best2011, waters2017}, Global Ocean 5.0
-\citep{madec2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
+ACCESS-S2 \citep{wedd2022} is the Bureau of Meteorology's seasonal
+forecast system which became operational in October 2021, replacing the
+ACCESS-S1 system \citep{hudson2017}. The model components of both
+ACCESS-S2 and ACCESS-S1 are identical with the same numbers of levels
+and resolution. They consist of the Global Atmosphere 6.0 (GA6)
+\citep{williams2015, waters2017}, the Unified Model's Global Land 6.0
+\citep{best2011, waters2017}, NEMO Global Ocean 5.0
+\citep{gurvan2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
 \citet{rae2015}{]}. The atmosphere has a N216 horizontal resolution
-(\textasciitilde60km in the mid-latitudes) with 85 vertical levels. The
-land model uses the same horizontal grid with 4 soil levels. The ocean
-component has a 1/4º resolution with 75 vertical levels. The sea ice
-component, based on CICE version 4.1, has the same resolution as the
-ocean and 5 sea ice categories as well as an open water category.
+(\textasciitilde60 km in the mid-latitudes) with 85 vertical levels. The
+land model uses the same horizontal grid as the atmosphere with four
+soil levels. The ocean component has a nominal horizontal resolution of
+1/4º with 75 vertical levels. The sea-ice component, based on CICE
+version 4.1, has the same resolution as the ocean component and five
+sea-ice thickness categories as well as an open water category.
 
 Both systems take atmospheric initial conditions derived from
-ERA-interim \citep{dee2011} for their hindcasts and from the Bureau's
-operational analysis for real-time forecasts. The main difference
-between the two are the ocean and sea ice initial conditions. ACCESS-S1
-ocean initial conditions come from the Met Office FOAM system, which
-uses a multivariate, incremental three-dimensional variational (3D-Var),
-first-guess-at-appropriate-time (FGAT) data assimilation scheme
-\citep{waters2015} and assimilates sea surface temperature (SST), sea
-surface height (SSH), in situ temperature and salinity profiles, and sea
-ice concentration. ACCESS-S2, on the other hand, runs from initial
-conditions generated by the BoM data assimilation scheme described in
-\citet{wedd2022}. This scheme is a weakly coupled ensemble optimal
-interpolation method and assimilates temperature and salinity profiles
-from EN4 \citep{good2013} for the hindcasts and from the WMO Global
-Telecommunication System (GTS) and the Coriolis and USGODAE Global Data
-Assembly Centers (GDACs) for the real-time forecasts. SSTs are nudged to
-Reynolds OISSTv2.1 \citep{reynolds2007} for the hindcasts and to the
-Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs
-2008) for the real-time forecasts in areas where SSTs are over 0ºC.
-Relevant for this work, sea ice concentrations are not assimilated in
-ACCESS-S2.
+ERA-interim \citep{dee2011} for their hindcasts. The main difference
+between the hindcasts of the two systems are the ocean and sea-ice
+initial conditions. ACCESS-S1 ocean initial conditions come from the Met
+Office FOAM system, which uses a multivariate, incremental
+three-dimensional variational (3D-Var), first-guess-at-appropriate-time
+(FGAT) data assimilation scheme \citep{waters2015} and assimilates sea
+surface temperature (SST), sea surface height (SSH), in situ temperature
+and salinity profiles, and sea-ice concentration. ACCESS-S2, on the
+other hand, is initialised from ocean conditions generated by the BoM
+weakly coupled ensemble data assimilation scheme described in
+\citet{wedd2022}. This scheme uses an optimal interpolation method and
+assimilates temperature and salinity profiles from EN4 \citep{good2013}.
+SSTs are nudged to Reynolds OISSTv2.1 \citep{reynolds2007} in areas
+where SSTs are over 0ºC and Sea Surface Salinity is weakly nudged to the
+World Ocean Atlas 2013 climatology \citep{zweng2013}.
 
-The ACCESS-S2 hindcast used in this study runs for the period
-1981--2018. Each forecast consists of 9 ensemble members built from
+Of most relevance for this work, sea-ice concentrations are not
+assimilated in ACCESS-S2. Assimilation cycles are performed daily. The
+coupled model runs for 24 hours initialised from the previous cycle.
+Then the restart file fields of the ocean component are used as first
+guess in the data assimilation cycle and the innovations are used to
+build the next ocean initial conditions for the following cycle. The
+atmosphere fields from that daily integration are not used and instead
+the model atmosphere is initialised using ERA-Interim. The sea-ice
+initial conditions for the next cycle are the unaltered output of the
+previous daily integration. Then the cycle starts again and the coupled
+model runs for another 24 hours. During this integration the sea-ice
+component is affected by the ocean innovations and the new atmosphere
+initial conditions via the coupler.
+
+The ACCESS-S2 hindcast set used in this study runs for the period
+1981--2018. Each hindcast consists of nine ensemble members built from
 three consecutive 3-member forecasts initialised at the first of every
-month and the two previous days and run from 279 days. The ACCESS-S1
-hindcast is built in the same manner (albeit with 217-day runs) for the
-period 1990--2012.
+month and the two previous days and run for 279 days. The ACCESS-S1
+hindcast set is built by perturbing the atmospheric fields only with a
+random field perturbation \citep{hudson2017} and runs for 217 days for
+the period 1990--2012. We analise the ensemble mean hindcasts unless
+otherwise specified.
 
-Anomalies for each hindcast are be taken with respect to a
-forecast-dependent climatology for the period 1990--2012. This serves as
-a first-order correction of model drift.
+Anomalies for each hindcast set are taken with respect to their own
+climatology specific to each initialisation date and forecast lead time,
+for the period 1990--2012. This serves as a first-order correction of
+model bias and drift.
+
+Besides sea-ice concentration, we also analyse mean sea-ice thickness,
+which we compute as total sea-ice volume divided by total sea-ice area.
 
@@
 
-For verification we use satellite-derived estimates of sea ice
-concentration, which estimates the proportion of each grid area that is
-covered with ice. Different datasets derived using different algorithms
-provide different estimates, each with their own biases and
-uncertainties. \citet{meier2019} estimated that the inter-product
-uncertainty of sea ice extent is of the order of 1 million \(km^2\). As
-will be shown below, this spread is minimal compared with the typical
-errors in the ACCESS-S2 and ACCESS-S1 forecasts, so the conclusions are
-independent of the dataset used.
+For verification we use satellite-derived sea-ice concentration, which
+estimates the proportion of each grid area that is covered with ice.
+Datasets derived using different algorithms and satellite platforms,
+each have their own biases and uncertainties. Estimates of inter-product
+uncertainty of sea-ice extent (SIE, defined here as the total region of
+the Southern Ocean with at least 15\% sea-ice cover) is of the order of
+0.5 million \(km^2\) \citep{meier2019}. As will be shown below, this
+spread is minimal compared with the typical errors in the ACCESS-S2 and
+ACCESS-S1 forecasts, so the overall conclusions of this study are
+independent of the verification dataset used.
 
 We use NOAA/NSIDC's Climate Data Record V4 {[}CDR; \citet{meier2014}{]}
-as the primary verification dataset. It takes the maximum value of the
-NASA Team \citep{cavalieri1984} and NASA Bootstrap \citep{comiso2023}
-sea ice concentration products to reduce their low concentration bias
-\citep{meier2014, meier2021}. Both source algorithms use data from the
-Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7
-satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors
-on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and
--F13 satellites. The data has a resolution of 25 by 25 km and daily from
-1978 onwards.
+as the primary sea-ice verification dataset. It takes the maximum value
+of the NASA Team \citep{cavalieri1984} and NASA Bootstrap
+\citep{comiso2023} sea-ice concentration products to reduce their low
+concentration bias \citep{meier2014, meier2021}. Both source algorithms
+use data from the Scanning Multichannel Microwave Radiometer (SMMR) on
+the Nimbus-7 satellite and from the Special Sensor Microwave/Imager
+(SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP)
+-F8, -F11, and -F13 satellites. The data has a spatial resolution of 25
+by 25 km and daily from 1978 onwards.
 
@@
 Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility
-(OSI SAF) \citep{OSISAF} is another satellite-derived sea ice
-concentration product. It is based on mostly the same sensors as the
-NOAA CDR but computed independently using different algorithms. Figures
-prepared with this dataset are provided in the supplementary material
-and don't differ significantly from the ones prepared using CDR.
+\citep{OSISAF} based on the SSMIS sensor is another satellite-derived
+sea-ice concentration product. It is based on mostly the same sensors as
+the NOAA CDR but computed independently using different algorithms.
+Figures prepared with this dataset are provided in the supplementary
+material and do not differ significantly from the ones prepared using
+CDR.
 
@@
 
-Sea-ice extent is defined as the area of ocean covered with at least
-15\% ice. This threshold is motivated by the limitations in satellite
-retrieval, which is increasingly unreliable for lower sea ice conditions
-(cite -- probably something from NSIDC).
+Sea-ice extent is defined as the area of the ocean at least 15\% covered
+by sea-ice. This threshold is motivated by the limitations in satellite
+retrieval, which is increasingly unreliable for lower sea-ice
+concentrations \citep{cavalieri1991}.
 
-Pan-Antarctic (net) sea ice extent serves as a rough hemispheric measure
-of the amount of sea ice, but it does not take into account the spatial
-distribution. A model could have relatively accurate extent of the net
+Pan-Antarctic (net) sea-ice extent serves as a hemispheric measure of
+the amount of sea ice, but it does not take into account the spatial
+distribution. A model could have a relatively accurate extent of the net
 ice but with different regional distributions. To account for location
-errors, we computed the Root Mean Squared Error (RMSE) of sea ice
-concentration anomalies and the Integrated Ice Edge Error {[}IIEE;
-\citet{goessling2016}{]}.
+errors, we computed the Root Mean Squared Error (RMSE) of grid-point
+sea-ice concentration anomalies.
 
-We compute RMSE as the square root of the average squared differences
-between forecasted and observed sea ice concentration anomalies. We
-compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC
-CDRV4 southern hemisphere domain, and also a zonally-varying RMSE
-computed over twenty-four 15° longitude slices around Antarctica.
+We compute RMSE as the square root of the area-averaged squared
+differences between grid-point forecasted and observed sea-ice
+concentration anomalies. We compute a pan-Antarctic RMSE by averaging
+over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a
+zonally-varying RMSE computed over twenty-four 15° longitude slices
+around Antarctica.
 
-The IIEE is defined as the area where the model miss-predicts sea ice
-concentration being above or below 15\% ice. That is, dichotomise sea
-ice concentration into areas with more and less than 15\% sea ice both
-in the forecast and observations.
+All error measures were computed on the NOAA/NSIDC CDRV4 domain grid, to
+which model output was bilinearly interpolated. Note that the ACCESS
+CICE model grid has resolution between two and three times higher than
+NOAA/NSIDC CDRV4.
 
-All error measures were computed on the NOAA/NSIDC CDRV4 domain grid and
-projection to which model output was bilinearly interpolated.
+Forecasts errors are also compared with hypothetical forecasts based on
+the persistence of anomalies and on climatology. The persistence
+forecast is generated by extending the observed anomalies in time.
+
+As a measure of forecast improvement over the hypothetical forecast, we
+use the skill score \citep{murphy1985}, defined as
+
+\[
+S = 1 - \frac{RMSE_{f}}{RMSE_{r}}
+\]
+
+Where \(RMSE_{f}\) is the RMSE of the forecast, \(RMSE_{r}\) is the RMSE
+of the reference forecast. Negative skill score indicates that the
+forecast is worse than the reference forecast while positive values
+indicate an improvement. A perfect forecast, would have zero RMSE and
+thus a skill score of 1.
 
@@
 
-\caption{\label{fig-hindcast-extent}Pan-antarctic median sea ice extent
-for al hindcasts initialised the first of the month for ACCESS-S2 and
-ACCESS-S1 in colours representing the start month with the median sea
-ice extent of NOAA/NSIDC CDRV4 in the left column and the median
-difference in the right column. The NOAA/NSIDC CDRV4 climatology is
-computed in the period corresponding to each hindcast. Circles represent
-the initial conditions at the first of every month and triangles
-represent the median value at the first of every month forecasted with
-the largest possible lead time. Colours indicate the initialisation
-month of the forecast.}
+\caption{\label{fig-hindcast-extent}Row a: Pan-Antarctic mean sea-ice
+extent for all hindcasts initialised on the first of each calendar month
+for ACCESS-S1 (column 1; green) and ACCESS-S1 (column 2; purple). Mean
+sea-ice extent from NOAA/NSIDC CDRV4 in each corresponding hindcast
+period is shown in black. Row b: Mean differences between the forecast
+and the observed values. Circles represent the initial conditions at the
+start of forecasts (i.e., the first of every month), and triangles
+represent the mean values at the end of forecasts (i.e., the longest
+possible lead time).}
 
@@
 
-Figure~\ref{fig-hindcast-extent} a shows the median sea ice extent (left
-column) and median difference with respect of NOAA/NSIDC CDRV4 for the
-ACCESS-S2 and ACCESS-S1 hindcasts (right column). Median extent of
-initial conditions at the first of every month are indicated with
-circles, while median extent at the same date but forecasted with the
-largest lead time possible for each model (between 274 and 277 days for
-ACCESS-S2 and between 213 and 216 days for ACCESS-S1). At this large
-lead time, the the information of the initial conditions is essentially
-lost and the forecast reverts to each model's preferred equilibrium
-state.
+Figure~\ref{fig-hindcast-extent} shows the mean sea-ice extent of the
+ACCESS-S2 and ACCESS-S1 hindcasts (row a) and their differences from the
+mean sea-ice extent of NOAA/NSIDC CDRV4 (row b). Mean extent at the
+first of every month is indicated with circles for the initial
+conditions and with triangles for the longest lead time possible for
+each model (between 274 and 277 days for ACCESS-S2 and between 213 and
+216 days for ACCESS-S1). At this long lead time, the information of the
+initial conditions is essentially lost and the forecast reverts to each
+model's preferred equilibrium state.
 
-ACCESS-S2 initial conditions (circles) show an overall negative bias,
-especially in the late summer-early autumn, while ACCESS-S1 initial
-conditions are very close to observations. Both models' equilibrium
-state (triangles) show a negative bias of sea ice extent, particularly
-in the late-autumn and winter months. This is due primarily to faster
-melt during a longer melt season between January and March and slower
-growth during March and April. This is then partially balanced with
-faster growth between May and July (Fig.~\ref{fig-mean-growth}). Many
-sea ice models exhibit this systematic underestimation during the sea
-ice minimum and early freezing season \citep{massonnet2023} and could
-indicate problems in the representation of thermodynamics in the model
-\citep{zampieri2019}. It is not surprising that both forecasting systems
-converge to a similar climatology because they share the same model
-formulation.
+ACCESS-S2 initial conditions (circles; Figure~\ref{fig-hindcast-extent}
+column 1) show an overall negative bias, especially in the late
+summer-early autumn, while ACCESS-S1 initial conditions (circles in
+Figure~\ref{fig-hindcast-extent} column 2) are very close to
+observations, as expected from the assimilation of sea-ice observations
+to produce the initial conditions of ACCESS-S1. Both systems'
+equilibrium states (triangles) show negative biases of sea-ice extent,
+particularly in the growth phase of late-autumn and winter months. This
+is due primarily to the melt season being longer than in observations
+and with faster melt between January and March and the growing seasons
+being shorter with slower growth during March and April. This is then
+followed by faster growth between May and July
+(Figure~\ref{fig-mean-growth}). Many sea-ice models exhibit this
+systematic underestimation during the sea-ice minimum and early freezing
+season \citep{massonnet2023}, which could indicate problems in the
+representation of thermodynamics in the model \citep{zampieri2019}. It
+is also not surprising that both forecasting systems converge to a
+similar equilibrium state because they share the same model formulation.
 
-The difference between the initial conditions and the model preferred
-state can be attributed to the effect of data assimilation, which in
-ACCESS-S2 is due solely to atmospheric and oceanic data assimilation.
-From June to October, in ACCESS-S2 circles are closer to observations
-than to the triangles, indicating that the information from the ocean
-and atmosphere data assimilation is affecting sea ice and improving the
-initial conditions. The rest of the year there is little if any
-difference between circles and triangles in ACCESS-S2, indicating that
-the ocean and atmosphere data assimilation is not affecting seas ice and
-that this component of the model is virtually free-running.
+The difference between the initial conditions (circles) and the model
+equilibrium state (triangles) can be mostly attributed to the effect of
+data assimilation, which in ACCESS-S2 is due solely to atmospheric and
+oceanic data assimilation. From May to October, in ACCESS-S2 circles are
+closer to observations than to the triangles, indicating that the
+information from the ocean and atmosphere data assimilation is affecting
+sea ice and improving the initial conditions. During these months,
+ACCESS-S1 can overestimate the sea-ice extent at short lead time. For
+the rest of the year circles are overlaid with triangles in ACCESS-S2,
+indicating that the ocean and atmosphere data assimilation is not
+affecting sea ice and that this component of the model is virtually
+free-running.
 
@@
 
-\caption{\label{fig-mean-growth}Median daily sea ice extent growth of
-ACCESS-S1 and ACCESS-S2 hindcasts and observations computed as the
-median daily differences in sea ice extent for each date and all lead
-times. Values are smoothed with a 11-day running mean.}
+\caption{\label{fig-mean-growth}Mean daily sea-ice extent growth
+(10\^{}6 km\^{}2/day) in ACCESS-S1 (green) and ACCESS-S2 (purple)
+hindcasts and observations (black), computed as the mean daily
+differences in sea-ice extent between each date and the next for each
+forecast month. Values are smoothed with a 11-day running mean.}
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1-1.pdf}}
 
@@
 
-\caption{\label{fig-bias}Mean difffernece between monthly sea ice
-concentrations at 0-month lead time ACCESS-S2 forecast and
-observations.}
+\caption{\label{fig-bias-1}Ensemble mean difference between monthly
+sea-ice concentration of ACCESS-S2 ensemble mean forecast at 0-month
+lead time (monthly mean values forecasted from the forecast initialised
+at the first of the month) and observations.}
 
@@
 
-Figure~\ref{fig-bias} shows the difference of monthly mean sea ice
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-2-1.pdf}}
+
+}
+
+\caption{\label{fig-bias-2}Same as Figure~\ref{fig-bias-1} but for
+ACCESS-S1}
+
+\end{figure}%
+
+To further understand the bias in ACCESS-S2, Figure~\ref{fig-bias-1}
+shows spatial patterns of the differences of monthly mean sea-ice
 concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
-shortest lead time From October to May, the model underestimates sea ice
-concentrations in most regions except for the inner Weddell Sea in April
-and May, where sea ice concentrations saturate to 1 both in the
-observations and forecasts. In winter, the differences are mostly on the
-sea ice edge with light positive bias in the African sector of East
-Antarctica and negative bias around the Indian Ocean sector which
-partially compensate, resulting in the near-zero extent bias seen in
-those months (Fig.~\ref{fig-hindcast-extent}).
+shortest monthly lead time. From October to May, the model
+underestimates sea-ice concentrations in most regions except for the
+inner Weddell Sea in April and May, where sea-ice concentrations
+saturate to 1 both in the observations and forecasts. In winter, the
+differences are limited to a narrow band around the sea-ice edge with
+slight positive biases in the African sector of East Antarctica and
+negative biases around the Indian Ocean sector which partially
+compensate, resulting in the near-zero extent bias seen in those months
+(Figure~\ref{fig-hindcast-extent}).
+
+ACCESS-S1 has a comparatively smaller overall bias
+(Figure~\ref{fig-bias-2}). The largest values are found between April
+and June, when the faster growth resutls in large positive bias along
+the sea-ice edge, and in January, when the faster melt leads to large
+negative bias in the Weddell and Admunsen seas.
 
@@
 
-\caption{\label{fig-extent-anom}Monthly mean sea ice extent anomalies
-for ACCESS-S1 and ACCESS-S2 and observations (black) forecasted at
-selected lead times. The RMSE during the overlapping period (1990--2013)
-is shown on the top left.}
+\caption{\label{fig-extent-anom}Monthly mean sea-ice extent anomalies of
+the observations (black) and forecasts from ACCESS-S1 (right column;
+purple) and ACCESS-S2 (left column; green) at lead times of 0, 2, 4, and
+6 months. The RMSE and correlation during the overlapping period of
+ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) is shown on the top left
+and bottom left of each panel respectively.}
 
@@
 
-Figure~\ref{fig-extent-anom} shows monthly sea ice extent anomalies
+Figure~\ref{fig-extent-anom} shows monthly sea-ice extent anomalies
 forecasted at selected lead times. Compared with ACCESS-S1, ACCESS-S2
-anomaly forecast is relatively poor even in the first month, which stays
-relatively skilful even at a lead time of 3 months. ACCESS-S2 shows much
-bigger variability than observations, with dramatic lows between 1995
-and 2007 and highs between 2007 and 2015.
+anomaly forecasts are relatively poor (large RMSE) even for the first
+month (lead time 0), whereas ACCESS-S1 forecasts stay relatively skilful
+even at a lead time of three months. ACCESS-S2 shows much larger
+interannual variability than observations, with dramatic lows between
+1995 and 2007, and highs between 2007 and 2015.
 
-To assess ACCESS-S2 forecasts quantitatively, we compute error measures
+Unexpectedly, for ACCESS-S2, RMSE improves with lead time, even though
+the correlation degrades with lead time. This is puzzling behaviour that
+goes contrary to what is usually seen in prediction models. The
+explanation seems to be the mentioned increased interannual variability.
+Figure~\ref{fig-extent-sd} shows monthly mean sea-ice extent of the
+forecasts as a function of lead time compared with observations.
+ACCESS-S1 standard deviation lies within the observed standard deviation
+regardless of lead time, while ACCESS-S2 standard deviation is more than
+twice that of observations at zero lead time and only approaches the
+observed value at 9 month lead time for most months.
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-sd-1.pdf}}
+
+}
+
+\caption{\label{fig-extent-sd}Interannual standard deviation with 95\%
+confidence interval of monthly mean sea-ice extent forecasted for each
+month divided by that month's sea-ice extent observation standard
+deviation. ACCESS-S1 and ACCESS-S2 at different lead times. Each panel
+indicates the target month. Note the reverse horizontal axis.}
+
+\end{figure}%
+
+ACCESS-S2 forecasts of sea-ice extent anomalies seem to align moderately
+well with observations (leading to moderately high correlation) but
+their magnitude is overestimated (leading to large errors). This could
+be caused by ACCESS-S2 sea ice being much more sensitive to atmospheric
+and oceanic forcing, perhaps due to lower thickness.
+
+As an example, Figure~\ref{fig-forecast-example} shows sea-ice
+concentration anomalies (top row) and sea-ice thickness and the
+difference between the two models (bottom row) for 2 May 2008
+initialised one day prior; being that close to initialisation date,
+these are very approximately the initial conditions. ACCESS-S1 sea-ice
+concentrations anomalies are very close to observations as expected from
+the system assimilating these data. ACCESS-S2 sea-ice concentration
+anomalies, which are not assimilated, are not as close, but the
+large-scale pattern is aligned with observations. The system simulates
+large positive anomalies in the Weddell and Ross seas and slight
+negative anomalies in the Amundsen and Bellingshausen seas. The fact
+that ACCESS-S2 can simulate this pattern without assimilating sea-ice
+data suggests that atmospheric and oceanic forcing was the dominant
+factor. However, the magnitude of the sea-ice anomalies is too big. It
+is plausible that this is due to the thinner ice simulated by ACCESS-S2
+(bottom row).
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-forecast-example-1.pdf}}
+
+}
+
+\caption{\label{fig-forecast-example}ACCESS-S1 and ACCESS-S2 hindcasts
+for 2 May 2008 at one day lead time. Top row shows sea-ice concentration
+anomalies forecasted by each system and the observations. Bottom row
+shows forecasted sea-ice thickness and the difference between ACCESS-S1
+and ACCESS-S2.}
+
+\end{figure}%
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-thickness-1.pdf}}
+
+}
+
+\caption{\label{fig-mean-thickness}Mean and 95\% interval of monthly
+mean sea-ice thickness for ACCESS-S1 and ACCESS-S2 at different lead
+times. Each panel indicates the target month. Note the reverse
+horizontal axis.}
+
+\end{figure}%
+
+Extending beyond the one case in Figure~\ref{fig-forecast-example},
+Figure~\ref{fig-mean-thickness} shows monthly mean sea-ice thickness as
+a function of lead time for ACCESS-S1 and ACCESS-S2. Supporting the idea
+that thinner ice is what causes the increased extent variability in
+ACCESS-S2, this system simulates thinner sea-ice compared to ACCESS-S1
+overall at almost all lead times and in all months except for summer at
+short lead times (Dec-Jan, 0-1 months; Feb-Mar, 0-2 months). However, in
+both systems, forecasted sea-ice is thicker at shorter lead times and
+then decreases, particularly in the summer months. If thinner ice were a
+sufficient cause of increased variability, then we would expect
+variability to increase with lead time in both forecasting systems.
+
+The fact that ACCESS-S1 and ACCESS-S2 share the same model configuration
+and that the increased variability is more extreme at short lead times
+(Fig.~\ref{fig-extent-sd}) suggests that the data assimilation procedure
+is partly to blame. It is possible that sea-ice in the ACCESS-S2 system
+is left in an unbalanced state after assimilating atmospheric and
+oceanic data but not sea-ice data, leading to large responses that are
+amplified by the thin ice in the initial states which then subside at
+longer lead times when the model is balanced.
+
+To assess ACCESS-S2 forecasts in more detail, we compute error measures
 for all hindcasts started on the 1st of every month.
@@
 
-\caption{\label{fig-rmse}Median RMSE of sea ice concentration anomalies
-as a function of forecast lead time for all forecast initialised on the
+\caption{\label{fig-rmse}Mean RMSE of sea-ice concentration anomalies as
+a function of forecast lead time for all forecasts initialised on the
 first of each month compared with a reference forecast of persistence of
-anomalies (black) and climatology (gray).}
+anomalies (black) and climatology (gray). Only the first 120 days are
+shown. In parenthersis, the shortest time at which ACCESS-S1 and
+ACCESS-S2 mean RMSE is not statistically different at the 99\%
+confidence level.}
 
@@
 
-Figure~\ref{fig-rmse} shows the median RMSE of sea ice concentration
-anomalies for ACCESS-S2 and ACCESS-S1 hindcasts compared with a
-benchmark of persistence and climatology. Due to errors in the initial
-conditions, it is expected that a persistence forecast would be better
-than the model forecast at very short lead times, but that the
-persistence forecast errors would grow faster and eventually surpass the
-model forecast, after when the model is statistically useful. Here the
-persistence errors are almost always lower than the ACCESS-S2 forecast,
-indicating that the model does not have skill at any lead time and in
-any month. The only exception is around February, where the model has
-lower RMSE than the persistence forecast at virtually every lead time.
+Figure~\ref{fig-rmse} shows the mean RMSE of sea-ice concentration
+anomalies for ACCESS-S1 and ACCESS-S2 hindcasts compared against
+persistence and climatological forecasts used as a benchmark. Due to
+errors in the initial conditions, it is expected that persistence
+forecasts would be better than the model forecasts at very short lead
+times, but that the persistence forecast errors would grow faster and
+may eventually surpass the model forecast errors. The black line shows
+that the persistence forecast error indeed grows rapidly and reaches its
+maximum in about 30 days for most months except for February, when it
+grows much slower. The ACCESS-S1 forecast errors grow slower than
+persistence forecast errors and remain lower after less than 10 days on
+average. The ACCESS-S2 forecast error starts high in all months and is
+lower than the persistence forecast error after more than 15 days in
+most months except for forecast initialised in February, when it takes
+80 days.
+
+At longer lead times, it is more appropriate to compare errors with the
+climatological forecast error. The lead time at which ACCESS-S1 forecast
+error is higher than the climatological forecast error varies between
+more than 60 and less than 20 days depending on forecast initialisation
+month with the minimum in June. ACCESS-S2 forecasts never have lower
+error than climatology, on the other hand, except marginally in October
+forecasts.
+
+Figure~\ref{fig-lead-time-window} summarises the lead time window in
+which each hindcast is better than both the persistence forecast and the
+climatological forecast as a function of forecast month. ACCESS-S1
+forecasts have a wider lead time window in the summer than the other
+seasons and is not better than both benchmarks at forecasting June
+sea-ice concentration anomalies. Forecasts initialised in May and June
+are particularly poor, and July cannot be forecasted better than the
+benchmarks. This is consistent with the mid-winter loss of
+predictability observed by \citet{libera2022}, who attributed it to deep
+warm water entraining into the mixed layer.
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-lead-time-window-1.pdf}}
 
@@
 
-\caption{\label{fig-rmse_lon}Median difference between RMSE of ACCESS-S2
-forecasts and persistence forecast computed on
-\texttt{r\ uniqueN(rmse\_lon\_mean\$lon)} meridional slices
-\texttt{r\ diff(unique(rmse\_lon\_mean\$lon)){[}1{]}}º wide.
-Antarctica's coastline is shown at the bottom of each panel for
-reference.}
+\caption{\label{fig-lead-time-window}Minimum lead time at which each
+forecast's mean RMSE becomes larger than the lower bound of the 95\%
+confidence interval of persistence forecast RMSE (black lines) and
+maximum lead time at which each forecast's mean RMSE remains lower than
+the lower bound of the 95\% confidence interval of climatological
+forecast RMSE (gray lines). Green shading indicates the window where
+forecasts outperform both persistence (lead times longer than black
+line) and climatology (lead times shorter than gray line). Text labels
+show the date corresponding to the maximum lead time at which each
+forecast outperforms climatology.}
 
 \end{figure}%
-
-Even though ACCESS-S2 is no better than persistence at forecasting
-pan-Antarctic sea ice concentration anomalies, the quality of the
-forecast might vary between regions. To analyse the spatial distribution
-of the model error, we computed sea ice concentration anomalies RMSE on
-15 meridional slices 24º wide. The median difference between the
-forecast and persistence RMSE is shown on Figure~\ref{fig-rmse_lon},
-where negative values indicate that the model has lower RMSE than the
-benchmark.
-
-The skill shown by ACCESS-S2 at February-March forecasts is observed
-
-by a band of negative values across almost all longitudes. The only
-region where February-March sea ice is not well forecasted is east of
-the Antarctic Peninsula, which is the only region with consistent summer
-sea ice. This suggests that the positive skill around the summer minimum
-comes from the model correctly forecasting no ice where no ice is ever
-found. However, there are other regions of skilful forecasts.
-
-Forecasts initialised on January 1st (Fig.~\ref{fig-rmse_lon} panel a)
-show RMSE values below persistence at lead times larger than 180 days
-(corresponding to July through September) in the Ross and Weddell Sea.
-These two regions are also relatively well forecasted from other months.
-Notably, May forecasts of Weddell sea ice concentration anomalies
-(Fig.~\ref{fig-rmse_lon} panel e) are skilful after about 20 days and
-remain so until December. The Weddell Sea is also the region with the
-maximum error --around June regardless of initialisation date.
-
-\subsubsection{Comparison with S1}\label{comparison-with-s1}
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-iiee-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-1.pdf}}
 
@@
 
-\caption{\label{fig-iiee}Median and 95\% coverage of Integrated Ice Edge
-Error as a function of forecast lead time for all forecast initialised
-on the first of each month for ACCESS-S1 and ACCESS-S2 hindcasts. For
-each month, the number in parenthesis indicates the minimum lead time at
-which the the mean error of each model is not statistically significant
-at a 1\% level using a two-sisded t-test.}
+\caption{\label{fig-rmse_lon-1}RMSE skill score of ACCESS-S1 forecasts
+with climatological forecast as reference computed on 15 meridional
+slices 24º wide as a function of lead time and longitude. Antarctica's
+coastline is shown at the bottom of each panel for reference.}
 
 \end{figure}%
-
-To compare ACCESS-S2 with ACCESS-S1, we computed the IIEE for both
-models. This error measure is shown in Figure~\ref{fig-iiee} for
-forecasts initialised at the first of every month at all lead times.
-ACCESS-S1 has lower error at lead times up to 60 days at all months,
-with the errors converging to ACCESS-S2 as the forecast goes on. The
-time to convergence depends on the month and it can be as short as two
-weeks June and July to as long as 160 days for forecasts initialised in
-February. Since the only difference between these forecasts are the
-initial conditions, this timescale is an indication of the memory of sea
-ice to initial conditions --at least from October to March when the data
-assimilated form the other components has little to no influence on sea
-ice.
-
-Also evident in Figure~\ref{fig-iiee} is the difference in the error
-spread at short lead times between ACCESS-S2 and ACCESS-S1. In all
-months ACCESS-S1 has much narrower error spread at the first day of the
-forecasts. The error in the initial conditions is not only small, but
-also fairly constant. This spread then grows towards a climatological
-spread as errors accumulate differently in different forecasts. For
-ACCESS-S2, this is true only between July and October. For all other
-months, the error spread is more or less stable throughout the forecast
-window, indicating that not only the initial error is high, but it is
-not constant.
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-iiee-variance-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-1.pdf}}
 
@@
 
-\caption{\label{fig-iiee-variance}Mean spread of IIEE at different lead
-times for different models.}
+\caption{\label{fig-rmse_lon-2}Same as Figure~\ref{fig-rmse_lon-1} but
+for ACCESS-S2}
 
@@
 
-The large initial error spread could be due either to large spread of
-ensemble members or to large spread of individual forecasts.
-Figure~\ref{fig-iiee-variance} splits the IIEE variance for each lead
-time into the mean variance of each individual forecast and the variance
-of the mean error of each individual forecast, which adds up to the
-total variance. The average variance of each forecast is almost
-identical between the two forecast systems in all months. This shows
-that the ensemble spread of individual forecasts evolves identically,
-which, again, is not unexpected because both systems share the same
-model formulation. This also shows that the perturbation scheme in
-ACCESS-S2 is comparable to the one in ACCESS-S1.
+To analyse the spatial distribution of the model error, we computed the
+RMSE of zonal mean sea-ice concentration anomalies on 15 slices of
+\texttt{diff(unique(rmse\_lon\_mean\$lon)){[}1{]}}° longitude span for
+each forecasting system. We control for some areas being naturally
+easier to forecast than others by computing the RMSE skill score with
+the climatological forecast RMSE as reference.
 
-On the other hand, the spread of the mean error is always larger in
-ACCESS-S2 than ACCESS-S1. The difference is particularly large at short
-lead times in some months, which coincide with the ones in which the
-data assimilation scheme is not influencing sea ice initial conditions.
+For ACCESS-S1 forecasts (Figure~\ref{fig-rmse_lon-1}), skill tends to be
+lower off the coast of Eastern Antarctica even at short lead times; for
+instance, the skill score for forecasts initialised in May and June are
+negative between 0° and 120°E even at almost zero lead time. This
+mirrors \citet{libera2022} findings of a ``winter predictability
+barrier'', although they focus on the Weddell sea and here we show that
+the effect seems to be stronger more to the East. In western Antarctica
+there is a hint of easterly-propagating skill in forecasts initialised
+in February and March. This is consistent with \citet{holland2013}
+findings that memory of sea-ice anomalies are stored in ocean heat
+content anomalies that are transported east by the Antarctic Circumpolar
+Current.
+
+ACCESS-S2 forecasts (Figure~\ref{fig-rmse_lon-2}) also have lower skill
+over Eastern Antarctica. From July to December even though the
+pan-Antarctic average skill is negative at all lead times
+(Fig.~\ref{fig-lead-time-window}), it is positive for up to a month in
+Western Antarctica. Since oceanic and atmospheric forcing is the only
+source of information, this suggests that sea-ice in this region is
+particularly sensitive to oceanic and atmospheric forcing and suggests a
+role of the Pacific-South American mode and the Amundsen Sea Low to
+shape sea-ice concentration anomalies. The fact that this is evident in
+the months in which ENSO teleconnections are more important for
+atmospheric circulation also suggest the influence of tropical Pacific
+variability. February and March are the only two months that can be
+forecasted with marginally positive skill in large regions.
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-initial-spread-1.pdf}}
+
+}
+
+\caption{\label{fig-initial-spread}Decomposition of forecast error
+spread at 1, 5 and 30 days lead time for ACCESS-S1 and ACCESS-S2
+hindcasts across initialization months. The left column shows the mean
+standard deviation of RMSE errors across ensemble members, while the
+right column shows the standard deviation of the ensemble mean RMSE
+error and the spread of the persistence and climatology forecasts
+errors.}
+
+\end{figure}%
+
+In Figure 5 the mean error was shown. Figure~\ref{fig-initial-spread}
+column 1 shows the mean standard deviation of errors among ensemble
+members at various lead times. At one day lead time
+(Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a slightly larger
+spread than ACCESS-S1 due to the way that ensemble members are
+generated. ACCESS-S1 ensemble members are generated by adding random
+field perturbations to the atmosphere only, which then are transferred
+to the other components via the coupled simulation \citep{hudson2017}.
+With this scheme, ensemble members are all but guaranteed to be
+underdispersed in the ocean and sea-ice components. The time-lag
+ensemble used for ACCESS-S2 ensures greater spread. This difference is
+gone after about just two days, and both systems have a comparable
+spread in ensemble member error afterwards
+(Fig.~\ref{fig-initial-spread} b1 and c1).
+
+Figure~\ref{fig-initial-spread} column 2, on the other hand, shows the
+standard deviation of ensemble mean error of each hindcast and the
+persistence forecast. At one day lead time, ACCESS-S2 ensemble mean
+error standard deviation is much larger than ACCESS-S1's, which in turn
+is comparable to the persistence forecast error standard deviation. At
+longer lead times, the spread of ACCESS-S1 and persistence forecast
+standard deviation increases to eventually be comparable to ACCESS-S2
+and the standard deviation in climatological forecast errors. ACCESS-S2
+error standard deviation is fairly independent of lead time and similar
+to the climatological forecast error standard deviation at all lead
+times.
 
@@
 
-Sea ice forecasts from the ACCESS-S2 model show a significant low extent
-bias, particularly during late summer and early autumn. This bias is
-attributed to a faster and longer melt season between January and March,
-and slower growth between March and April. This underestimation during
-the minimum and early freezing season is a common issue in many
+Sea-ice forecasts from the ACCESS-S2 system show a significant low
+extent bias, particularly during late summer and early autumn. This bias
+is attributed to a faster and longer melt season between January and
+March, and slower growth between March and April. This underestimation
+during the minimum and early freezing season is a common issue in many
 seasonal-to-subseasonal (S2S) systems, suggesting potential problems
@@
 model formulation with ACCESS-S1, the latter does not suffer from this
-bias. This is due to the assimilation of sea ice concentrations into the
-initial conditions, which successful corrects for the negative bias in
-the model. Our analysis suggests that the data assimilation system in
-ACCESS-S2 is only effectively influencing sea ice initial conditions
-from June to October, while the rest of the year, the sea ice component
-runs virtually free, reverting to its biased equilibrium state.
+bias, indicating that assimilating sea-ice concentrations successfully
+corrects for the negative bias that exists in the free-running model.
+
+Ensemble spread grows quickly even when perturbations are only
+implemented in the atmosphere component (in ACCESS-S1), indicating that
+sea-ice is indeed responding quickly to atmospheric perturbations.
+However, our analysis suggests that the atmosphere and ocean data
+assimilation implemented in ACCESS-S2 is only effectively influencing
+sea-ice initial conditions from June to October, while the rest of the
+year, the sea-ice component runs virtually free, reverting to its biased
+equilibrium state.
 
@@
 
-Based on the observation that ACCESS-S2 sea ice initial conditions are
-essentially not initialised, comparing its forecasts with ACCESS-S1's
-allows us to estimate the time-scale for which initial conditions are
-important. This highlights February initial conditions as a crucial for
-determining sea ice evolution at least up to late June. Arctic sea ice
-forecast also show greater sensitivity to initial conditions in boreal
-summer compared with boreal winter \citep{day2014, bunzel2016}, so
-similar mechanism might be playing a role.
+Although ACCESS-S1 only assimilates sea-ice concentration, it is clear
+that sea-ice thickness is also affected. ACCESS-S1 simulates
+significantly thicker ice than ACCESS-S2 and in both systems sea-ice is
+thicker at shorter lead times than at longer lead times. Both the
+explicit data assimilation in ACCESS-S2 and the effects of atmospheric
+and oceanic data assimilation in ACCESS-S1 might be nudging simulated
+sea ice to be thicker than the model equilimbrium state.\\
+We believe that the thinner sea-ice in ACCESS-S2 is a partial cause for
+the large sea-ice extent variance, but other mechanisms, such as
+unbalanced initial conditions might be important.
 
-Although forecasts are not skilful for forecasting pan-Antarctic sea ice
-concentrations, there are some areas where the model does show skill.
-The Weddell Sea, and the Ross Sea to a lesser extent are particularly
-well forecasted between June and November for forecasts initialised from
-January to June. These regions have been recognised as a regions of high
-predictability thanks to persistent and eastwardly advected upper ocean
-heat content anomalies \citep{bushuk2021}. Sea surface temperatures are
-assimilated by ACCESS-S2 only in areas with temperatures greater than
-0°C and errors don't show a clear easterly-propagating signal, so it is
-not clear if ACCESS-S2 is leveraging the same source of predictability..
+Given that ACCESS-S2 sea-ice extent is essentially not initialised by
+the observations, comparing its forecasts with those of ACCESS-S1 allows
+us to estimate the time-scale for which initial conditions are
+important. We find that initial conditions affect Antarctic sea-ice
+forecasts in ther order of a few months, but that effect is seasonally
+dependent. January to April initial conditions improve forecasts for up
+to 3 months with, February initial conditions in particular are shown to
+be crucial for determining sea-ice evolution at least up to May. Arctic
+sea-ice forecasts also show greater sensitivity to initial conditions in
+boreal summer, compared with boreal winter \citep{day2014, bunzel2016},
+suggesting a similar mechanism might be playing a role.
+
+Forecasts initialised in winter have very little skill and ACCESS-S1 and
+ACCESS-S2 forecast errors are not statistically different after just 2
+weeks. This is consistent with \citet{libera2022}'s finding of a
+``winter predictability barrier'' in the Weddell sea, although they
+describe the barrier as sharp loss of predictability in July, and here
+we find a gradual reduction in skill compared with climatology around
+June. This difference might be due to our use of pan-Antarctic RMSE,
+since our regional analysis indicates that the degraded skill is most
+dramatic in the King Haakon sea.
+
+These findings have important implications for both operational
+forecasting, model development and predictability studies. For
+operational centers, our results suggest that efforts to improve sea-ice
+data assimilation should prioritize the summer and autumn months when
+initial conditions have the greatest impact on forecast skill.
+Additionally, the substantial bias in ACCESS-S2 highlights the need for
+improved model physics, particularly in the representation of sea-ice
+thermodynamics and radiation processes. Crucially, our results suggest
+dramatic seasonal variations in sea-ice predictability. Future studies
+should therefore use initial conditions through the whole year rather
+than focusing on a limited number of initialisation dates.
 
@@
 
-\section{Suplementary figures}\label{suplementary-figures}
-
-\codedataavailability{use this to add a statement when having data sets
-and software code
-available} %% use this section when having data sets and software code available
-
-\sampleavailability{use this section when having geoscientific samples
-available} %% use this section when having geoscientific samples available
-
-\videosupplement{use this section when having video supplements
-available} %% use this section when having geoscientific samples available
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{Figures and tables in appendices}
-
-Regarding figures and tables in appendices, the following two options
-are possible depending on your general handling of figures and tables in
-the manuscript environment:
-
-\subsection{Option 1}
-
-If you sorted all figures and tables into the sections of the text,
-please also sort the appendix figures and appendix tables into the
-respective appendix sections. They will be correctly named
-automatically.
-
-\subsection{Option 2}
-
-If you put all figures after the reference list, please insert appendix
-tables and figures after the normal tables and figures.
-
-To rename them correctly to A1, A2, etc., please add the following
-commands in front of them: \texttt{\textbackslash{}appendixfigures}
-needs to be added in front of appendix figures
-\texttt{\textbackslash{}appendixtables} needs to be added in front of
-appendix tables
-
-Please add \texttt{\textbackslash{}clearpage} between each table and/or
-figure.
-\noappendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{Contributions} %% optional section
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing
-interests.} %% this section is mandatory even if you declare that no competing interests are present
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\disclaimer{We like Copernicus.} %% optional section
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgements}
-Acknowledgements will gohere
-\end{acknowledgements}
-
-\bibliographystyle{copernicus}
+\section{Supplementary figures}\label{supplementary-figures}
 

### 2025-10-09 analysis/paper/access-ice.qmd
Polishes rmse_lon plots

@@
         id: 3
-  - name: Mathew Wheeler
+  - name: Matthew Wheeler
     affiliations:
@@
 #| label: rmse_lon_plots
+range <- range(gdata$skill)
 binwidth <- 0.1
+breaks <- AnchorBreaks(0, binwidth = binwidth)(range)
 rmse_lon_plots <- lapply(c("S1", "S2"), \(x) {
-  g <- gdata |>
+  gdata |>
     _[version == x] |>
@@
       aes(z = skill, fill = after_stat(level)),
-      breaks = AnchorBreaks(0, binwidth = binwidth)
+      breaks = breaks
     ) +
@@
       aes(z = skill),
-      range = c(0.01, 0.2),
-      breaks = AnchorBreaks(0, binwidth = binwidth)
+      range = c(0.005, 0.05),
+      breaks = breaks
     ) +
@@
       aes(lon, scales::rescale(lat, c(-60, 1)), z = z),
+      linewidth = 0.2,
       breaks = 0
     ) +
-    geom_hline(
+
+    geom_segment(
       data = month_from_lag(1, 1),
-      aes(yintercept = lag),
+      aes(y = lag, yend = lag, x = 7.5, xend = 360 - 7.5),
       colour = "gray50",
@@
       data = month_from_lag(15, 1),
-      aes(x = 361, label = month2),
-      size = 2.5,
-      hjust = 0
+      aes(x = 360, label = month2),
+      size = 1.5,
+      hjust = 0.2
     ) +
@@
       data = month_from_lag(1, 1),
-      aes(x = -2, label = lag),
-      size = 2.5,
-      hjust = 1
+      aes(x = 0, label = lag),
+      size = 1.5,
+      hjust = 0.8
     ) +
     scale_fill_divergent_discretised(low = trans_blue, high = trans_pink) +
-    scale_x_longitude(ticks = 60, expand = c(0.08, 0)) +
+    scale_x_longitude(ticks = 90, expand = c(0.08, 0)) +
 
@@
     ) +
-    tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1))
+    tag_facets(position = list(x = 0.08, y = 0, hjust = 0, vjust = 0))
 })
@@
 rmse_lon_plots[[2]]
+
 ```

### 2025-10-09 analysis/paper/access-ice.qmd
Adds OSI supplementary figures

@@
   _[year(time) < 2010 & anom < -2.5e12, aice := NA] |>
-  _[, anom := NULL]
+  _[, anom := NULL] |>
+  _[, time := update(time, hour = 0)]
 ```
@@
 hindcast_extent[, .(model, forecast_time, member, time, Forecast = aice)] |>
-  extent_daily[i = _, on = c("time")] |>
+  extent_daily[i = _, on = c("time"), allow.cartesian = TRUE] |>
   _[, Difference := Forecast - aice] |>
@@
   _[, measure := factor(measure, levels = c("Forecast", "Difference"))] |>
-
+  _[dataset == which_dataset] |>
   ggplot(aes(time2, estimate)) +
@@
   _[, dif := frollmean(dif, n = w, align = "center"), by = .(dataset)] |>
-  _[, average(dif), by = .(dataset, time = update(time, year = 2000))] |>
-  _[dataset == which_dataset]
+  _[, average(dif), by = .(dataset, time = update(time, year = 2000))]
 
@@
 dif_extent |>
-  copy() |>
+  _[dataset == which_dataset] |>
   ggplot(aes(time, estimate)) +
@@
 
-obs_clim <- CDR() |>
+obs_clim <- lapply(datasets, \(x) {
+  x |>
     climatology() |>
@@
     cdo_execute() |>
-  ReadNetCDF(c(obs = "aice")) |>
+    ReadNetCDF(c(obs = "aice"))
+}) |>
+  rbindlist(idcol = "dataset") |>
   _[, month := month(time)]
@@
     _[model == x] |>
+    _[dataset == which_dataset] |>
     # dcast(time + xgrid + ygrid ~ model, value.var = "aice") |>
@@
   _[obs_sd, on = .NATURAL] |>
+  _[dataset == which_dataset] |>
   ggplot(aes(lag, estimate / obs_sd)) +
@@
   _[!(month(time_forecast) == 1 & member == "07" & version == "S1")] # This member is wrong
-
 ```
@@
 clim_std <- datasets |>
-  _[1] |>
   lapply(\(x) {
@@
   _[climatology < 0.05, climatology := NA]
-
 ```
@@
   _[, version := "climatology"]
+
 ```
@@
   na.omit() |>
-  _[obs_dataset == which_dataset] |>
+  # _[obs_dataset == which_dataset] |>
   _[measure == "rmse"] |>
@@
   _[lag <= max_lag] |>
-  dcast(time + time_forecast + member + lag ~ version, value.var = "value") |>
-  _[, t_test(na.omit(S1), na.omit(S2)), by = .(lag, month(time_forecast))] |>
+  dcast(
+    obs_dataset + time + time_forecast + member + lag ~ version,
+    value.var = "value"
+  ) |>
+  _[,
+    t_test(na.omit(S1), na.omit(S2)),
+    by = .(obs_dataset, lag, month(time_forecast))
+  ] |>
   _[p.value > 0.01] |>
-  _[, .SD[which.min(lag)], by = month] |>
-  _[order(month), .(month, lag)]
-
-labels_month_significant <- setNames(
-  paste0(month.abb, " (", dif$lag, ")"),
-  1:12
-)
+  _[, .SD[which.min(lag)], by = .(obs_dataset, month)] |>
+  _[order(month), .(month, lag, obs_dataset)]
+
 ```
@@
 
+labels_month_significant <- setNames(
+  paste0(month.abb, " (", dif[obs_dataset == which_dataset]$lag, ")"),
+  1:12
+)
+
 errors2 <- clim_std2 |>
@@
   # merge(clim_std) |>
-  _[obs_dataset == which_dataset] |>
   _[,
@@
 errors2 |>
+  _[obs_dataset == which_dataset] |>
   _[lag <= 120] |>
@@
 clim_std_lon <- datasets |>
-  # _[1] |>
   lapply(\(x) {
@@
   _[, time := as.Date(time)]
-
-# base_rmse_lon <- clim_std_lon[rmse_lon_mean[version == "persistence"], on = .NATURAL] |>
-#   _[, lag := as.numeric(as.Date(time) - time_forecast)] |>
-#   _[, .(climatology = mean(climatology, na.rm = TRUE),
-#         value = mean(value, na.rm = TRUE)),
-#     by = .(lag, lon, obs_dataset, month = factor(month(time_forecast)))] |>
-#   _[, base := climatology] |>
-#   _[, let(climatology = NULL, value = NULL)] |>
-#   _[]
-```
-
-```{r rmse_lon_mean2}
-# rmse_lon_mean <- rmse_lon_mean |>
-#   copy() |>
-#   # _[version != "S2"] |>
-#   _[version != "persistence"] |>
-#   _[member == "0em"] |>
-#   _[base_rmse_lon, on = .NATURAL] |>
-#   _[, lag := as.numeric(as.Date(time) - time_forecast)] |>
-#   _[, .(dif = mean(value - base, na.rm = TRUE)),
-#     by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))]
 ```
@@
 
-gdata <- rmse_lon_mean |>
-  _[obs_dataset == which_dataset] |>
+rmse_lon_mean_gdata <- rmse_lon_mean |>
   _[version != "persistence"] |>
@@
 #| label: rmse_lon_plots
-range <- range(gdata$skill)
+
+rmse_lon_mean_gdata2 <- rmse_lon_mean_gdata[obs_dataset == which_dataset]
+
+range <- range(rmse_lon_mean_gdata2$skill)
 binwidth <- 0.1
@@
 rmse_lon_plots <- lapply(c("S1", "S2"), \(x) {
-  gdata |>
+  rmse_lon_mean_gdata2 |>
     _[version == x] |>
@@
 })
+
 ```
@@
 rmse_lon_plots[[1]]
-
 ```
@@
 rmse_lon_plots[[2]]
-
 ```
@@
   _[member != "0em"] |>
-  _[obs_dataset == which_dataset] |>
   _[value < 0.01, value := NA] |>
@@
   _[!(month(time_forecast) == 1 & member == "07" & version == "S1")]
+
 ```
@@
   tag_facets(tag = "rc")
+
 ```
@@
 
-<!-- ## ERA5 -->
-
-<!-- ```{r dataset_era} -->
-
-<!-- which_dataset <- "era5" -->
-
-<!-- ``` -->
-
-<!-- ```{r fig-hindcast-extent-era5} -->
-
-<!-- which_dataset <- "era5" -->
-
-<!-- <<fig-hindcast-extent>> -->
-
-<!-- ``` -->
-
-<!-- ```{r fig-hindcast-extent-bt} -->
-
-<!-- which_dataset <- "bt" -->
-
-<!-- <<fig-hindcast-extent>> -->
-
-<!-- ``` -->
-
-<!-- ```{r fig-mean-growth} -->
-
-<!-- <<fig-mean-growth>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<fig-bias>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<<fig-extent-anom>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<fig-rmse>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<fig-iiee>> -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- <<fig-iiee-variance>> -->
-
-<!-- ``` -->
+## ERA5
+
+```{r dataset_era}
+which_dataset <- "osi"
+```
+
+```{r fig-hindcast-extent-osi, ref.label=I('fig-hindcast-extent')}
+
+```
+
+```{r, fig-mean-growth-osi, ref.label=I("fig-mean-growth")}
+
+```
+
+```{r, plot_bias-osi, ref.label=I("plot_bias")}
+
+```
+
+```{r, fig-bias-1-osi, ref.label=I("fig-bias-1")}
+
+```
+
+```{r, fig-bias-2-osi, ref.label=I("fig-bias-2")}
+
+```
+
+```{r, fig-extent-anom-osi, ref.label=I("fig-extent-anom")}
+
+```
+
+```{r, fig-extent-sd-osi, ref.label=I("fig-extent-sd")}
+```
+
+```{r, fig-rmse-osi, ref.label=I("fig-rmse")}
+
+```
+
+```{r, fig-lead-time-window-osi, ref.label=I("fig-lead-time-window")}
+
+```
+
+```{r, rmse_lon_plots-osi, ref.label=I("rmse_lon_plots")}
+
+```
+
+```{r, fig-rmse_lon-1-osi, ref.label=I("fig-rmse_lon-1")}
+
+```
+
+```{r, fig-rmse_lon-2-osi, ref.label=I("fig-rmse_lon-2")}
+
+```
+
+```{r, fig-initial-spread-osi, ref.label=I("fig-initial-spread")}
+
+```

### 2025-10-09 analysis/paper/access-ice.qmd
Figures stay in their sections

@@
     fig-width: 5  
+    fig-pos: "!htb"
+header-includes: | 
+   \usepackage[section]{placeins}
 author:
@@
 #| fig-cap: "Row a: Pan-Antarctic mean sea-ice extent for all hindcasts initialised on the first of each calendar month for +S1 (column 1; green) and +S1 (column 2; purple).
-Mean sea-ice extent from NOAA/NSIDC CDRV4 in each corresponding hindcast period is shown in black.
+Observed mean sea-ice extent in each corresponding hindcast period is shown in black.
 Row b: Mean differences between the forecast and the observed values.
@@
 The largest values are found between April and June, when the faster growth resutls in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen seas.
+
+\FloatBarrier
 
@@
 
+\FloatBarrier
+
 ## Conclusions
@@
 
-------------------------------------------------------------------------
-
 # References
@@
 
-# Supplementary figures
-
-## ERA5
+\appendix  
+\renewcommand{\thefigure}{A\arabic{figure}}
+\setcounter{figure}{0}
+
+# Supplementary figures {.appendix}
+
+The following are the same figures from the main paper but using the OSI dataset instead of CDR.
 

### 2025-10-09 analysis/paper/access-ice.tex
Render

@@
 \raggedbottom
+\usepackage[section]{placeins}
 \makeatletter
@@
   pdftitle={Evaluating Seasonal Predictability of Antarctic Sea Ice and the importance of initial conditions with fully coupled models},
-  pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Mathew Wheeler; Phillip Reid},
+  pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Matthew Wheeler; Phillip Reid},
   pdfkeywords={sea ice, predictabilty, initial conditions, forecasting},
@@
 % author setup
-\author*[1,2]{\fnm{Elio} \sur{Campitelli}}\email{elio.campitelli@monash.edu}\author[1,2]{\fnm{Ariaan} \sur{Purich}}\author[1,2]{\fnm{Julie} \sur{Arblaster}}\author[3]{\fnm{Eun-Pa} \sur{Lim}}\author[3]{\fnm{Mathew} \sur{Wheeler}}\author[3]{\fnm{Phillip} \sur{Reid}}
+\author*[1,2]{\fnm{Elio} \sur{Campitelli}}\email{elio.campitelli@monash.edu}\author[1,2]{\fnm{Ariaan} \sur{Purich}}\author[1,2]{\fnm{Julie} \sur{Arblaster}}\author[3]{\fnm{Eun-Pa} \sur{Lim}}\author[3]{\fnm{Matthew} \sur{Wheeler}}\author[3]{\fnm{Phillip} \sur{Reid}}
 % affil setup
@@
 
-\begin{figure}
+\begin{figure}[!htb]
 
@@
 extent for all hindcasts initialised on the first of each calendar month
-for ACCESS-S1 (column 1; green) and ACCESS-S1 (column 2; purple). Mean
-sea-ice extent from NOAA/NSIDC CDRV4 in each corresponding hindcast
-period is shown in black. Row b: Mean differences between the forecast
-and the observed values. Circles represent the initial conditions at the
-start of forecasts (i.e., the first of every month), and triangles
-represent the mean values at the end of forecasts (i.e., the longest
-possible lead time).}
+for ACCESS-S1 (column 1; green) and ACCESS-S1 (column 2; purple).
+Observed mean sea-ice extent in each corresponding hindcast period is
+shown in black. Row b: Mean differences between the forecast and the
+observed values. Circles represent the initial conditions at the start
+of forecasts (i.e., the first of every month), and triangles represent
+the mean values at the end of forecasts (i.e., the longest possible lead
+time).}
 
@@
 
+\FloatBarrier
+
 \subsection{RMSE}\label{rmse}
 
-\begin{figure}
+\begin{figure}[!htb]
 
@@
 
+\FloatBarrier
+
 \subsection{Conclusions}\label{conclusions}
@@
 
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
-
 \section{References}\label{references}
@@
 
+\appendix  
+\renewcommand{\thefigure}{A\arabic{figure}}
+\setcounter{figure}{0}
+
 \section{Supplementary figures}\label{supplementary-figures}
 
+The following are the same figures from the main paper but using the OSI
+dataset instead of CDR.
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-hindcast-extent-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-hindcast-extent-osi}Row a: Pan-Antarctic mean
+sea-ice extent for all hindcasts initialised on the first of each
+calendar month for ACCESS-S1 (column 1; green) and ACCESS-S1 (column 2;
+purple). Observed mean sea-ice extent in each corresponding hindcast
+period is shown in black. Row b: Mean differences between the forecast
+and the observed values. Circles represent the initial conditions at the
+start of forecasts (i.e., the first of every month), and triangles
+represent the mean values at the end of forecasts (i.e., the longest
+possible lead time).}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-growth-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-mean-growth-osi}Mean daily sea-ice extent growth
+(10\^{}6 km\^{}2/day) in ACCESS-S1 (green) and ACCESS-S2 (purple)
+hindcasts and observations (black), computed as the mean daily
+differences in sea-ice extent between each date and the next for each
+forecast month. Values are smoothed with a 11-day running mean.}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-bias-1-osi}Ensemble mean difference between monthly
+sea-ice concentration of ACCESS-S2 ensemble mean forecast at 0-month
+lead time (monthly mean values forecasted from the forecast initialised
+at the first of the month) and observations.}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-2-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-bias-2-osi}Same as Figure~\ref{fig-bias-1} but for
+ACCESS-S1}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-anom-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-extent-anom-osi}Monthly mean sea-ice extent
+anomalies of the observations (black) and forecasts from ACCESS-S1
+(right column; purple) and ACCESS-S2 (left column; green) at lead times
+of 0, 2, 4, and 6 months. The RMSE and correlation during the
+overlapping period of ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) is
+shown on the top left and bottom left of each panel respectively.}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-sd-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-extent-sd-osi}Interannual standard deviation with
+95\% confidence interval of monthly mean sea-ice extent forecasted for
+each month divided by that month's sea-ice extent observation standard
+deviation. ACCESS-S1 and ACCESS-S2 at different lead times. Each panel
+indicates the target month. Note the reverse horizontal axis.}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-rmse-osi}Mean RMSE of sea-ice concentration
+anomalies as a function of forecast lead time for all forecasts
+initialised on the first of each month compared with a reference
+forecast of persistence of anomalies (black) and climatology (gray).
+Only the first 120 days are shown. In parenthersis, the shortest time at
+which ACCESS-S1 and ACCESS-S2 mean RMSE is not statistically different
+at the 99\% confidence level.}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-lead-time-window-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-lead-time-window-osi}Minimum lead time at which each
+forecast's mean RMSE becomes larger than the lower bound of the 95\%
+confidence interval of persistence forecast RMSE (black lines) and
+maximum lead time at which each forecast's mean RMSE remains lower than
+the lower bound of the 95\% confidence interval of climatological
+forecast RMSE (gray lines). Green shading indicates the window where
+forecasts outperform both persistence (lead times longer than black
+line) and climatology (lead times shorter than gray line). Text labels
+show the date corresponding to the maximum lead time at which each
+forecast outperforms climatology.}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-rmse_lon-1-osi}RMSE skill score of ACCESS-S1
+forecasts with climatological forecast as reference computed on 15
+meridional slices 24º wide as a function of lead time and longitude.
+Antarctica's coastline is shown at the bottom of each panel for
+reference.}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-rmse_lon-2-osi}Same as Figure~\ref{fig-rmse_lon-1}
+but for ACCESS-S2}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-initial-spread-osi-1.pdf}}
+
+}
+
+\caption{\label{fig-initial-spread-osi}Decomposition of forecast error
+spread at 1, 5 and 30 days lead time for ACCESS-S1 and ACCESS-S2
+hindcasts across initialization months. The left column shows the mean
+standard deviation of RMSE errors across ensemble members, while the
+right column shows the standard deviation of the ensemble mean RMSE
+error and the spread of the persistence and climatology forecasts
+errors.}
+
+\end{figure}%
+
 \end{document}

### 2025-10-09 analysis/paper/access-ice.qmd
Final polish after printing

@@
 ---
-title: Evaluating Seasonal Predictability of Antarctic Sea Ice and the importance of initial conditions with fully coupled models
+title: Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Model
 format:
@@
       - ref: 3      
-keywords: [sea ice, predictabilty, initial conditions, forecasting]
+keywords: [sea ice, predictability, initial conditions, forecasting]
 bibliography: references.bib
@@
 
-Accurate Antarctic sea-ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
-We investigate the role of initial conditions on Antarctic sea-ice forecasts by comparing hindcasts from two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system: +S1 and +S2.
-They use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations. 
+Accurate Antarctic sea ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
+The two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system, +S1 and +S2, use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations.
+This makes for a good "natural experiment" to analyse the role of initial conditions on Antarctic sea ice forecasts using a long run and a moderately large ensemble with a fully coupled model.
 Our analysis reveals that both systems experience an extended melt season and delayed growth phase. 
@@
 The impact of the initial conditions on forecast errors varies dramatically by season: summer and autumn initial conditions (January-April) provide predictive skill for up to three months, with February initial conditions being particularly crucial.
-In contrast, winter forecasts off the two systems are statistically indistinguishable after only two weeks. 
-Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over Eastern Antarctica, where evern +S1 shows negative skill. 
+In contrast, winter forecasts of the two systems are statistically indistinguishable after only two weeks. 
+Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over Eastern Antarctica, where even +S1 shows negative skill. 
 These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritize the summer-autumn period when initial conditions have maximum impact on forecast skill.
@@
 
-However, producing accurate Antarctic sea-ice forecasts has been challenging due to model biases, inherent large variability and complexity, and they have lagged behind Arctic sea-ice forecasts[@zampieri2019; @gao2024a].
-Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023] and machine learning approaches [e.g.
-@dong2024; @lin2025], which also underscores the need for better understanding of sea-ice dynamics, and drivers of its variability.
-
-Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea-ice initial conditions affect the quality of the forecast and at what timescales.
-Exploring seasonal predictions of Arctic sea-ice, @guemas2016 found that sea-ice initial conditions are important in autumn to predict summer sea-ice, but the impact wasn’t as dramatic when predicting winter sea-ice.
-@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea-ice forecasts initialised in July but not when initialised in January.
-
-For the Antarctic, @holland2013 studied the initial-value predictability of Antarctic sea-ice in a perfect model study using the CCSM3 model.
+However, Antarctic sea ice forecasts have lagged behind Arctic sea ice forecasts due to model biases, inherent large variability and complexity [@zampieri2019; @gao2024a].
+Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simpler statistical methods [@massonnet2023] and machine learning approaches (e.g.
+@dong2024, @lin2025), which also underscores the need for better understanding of sea-ice dynamics, and drivers of its variability.
+
+Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea ice initial conditions affect the quality of the forecast and at what timescales.
+Exploring seasonal predictions of Arctic sea ice, @guemas2016 found that sea ice initial conditions are important in autumn to predict summer sea ice, but the impact wasn’t as dramatic when predicting winter sea ice.
+@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea ice forecasts initialised in July but not when initialised in January.
+
+For the Antarctic, @holland2013 studied the initial-value predictability of Antarctic sea ice in a perfect model study using the CCSM3 model.
 They found that sea-ice and ocean initial conditions provide predictive information to forecast sea-ice edge location several months in advance and that some predictability is retained for up to two years thanks to ocean heat content anomalies that are advected eastward.
 This is in contrast with @marchi2020, who ran perfect model experiments to argue that uncertainty in the predicted atmospheric state and evolution is the main driver of uncertainty in Antarctic sea-ice extent prediction on seasonal timescales, with sea-ice and ocean initial conditions having lesser importance.
-More recently, @morioka2022 studied decadal forecasts of Antarctic sea-ice and found that initialising ocean and sea-ice improved the correlation between simulated and observed sea-ice concentration evolution in the Amundsen–Bellingshausen Sea.
+More recently, @morioka2022 studied decadal forecasts of Antarctic sea ice and found that initialising ocean and sea ice improved the correlation between simulated and observed sea-ice concentration evolution in the Amundsen–Bellingshausen Sea.
 It is hard to compare these studies since they are based on forecasts initialised at different times of the year and different frameworks: @holland2013 ran 20 ensemble members initialised on the 1st of January of a particular year, @marchi2020 ran forecasts from the 1st of March and 1st of September and @morioka2022 run forecasts only from the 1st of March.
@@
 
-In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the Australian Community Climate and Earth System Simulator – Seasonal (ACCES-S) from version S1 to S2.
+In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the Australian Community Climate and Earth System Simulator – Seasonal (ACCESS-S) from version S1 to S2.
 While the base model remained the same, the change in version was focused on using ocean, sea-ice and land initial conditions generated by the BoM instead of depending on the UK Met Office.
@@
 Since model configuration is identical between +S1 and +S2, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year, with the only difference being the initial conditions.
-This provides an opportunity to test the effect of sea-ice initial conditions on the forecast of sea-ice concentrations and the climate.
-
-In this study we compare sea-ice hindcasts produced by +S1 and +S2.
+This provides an opportunity to test the effect of sea ice initial conditions on the forecast of sea-ice concentrations and the climate.
+
+In this study we compare sea ice hindcasts produced by +S1 and +S2.
 We focus on seasonality of errors and biases and the effect of the data assimilation system.
-This comparison will also inform future work with the prediction system as a research tool to better understand the dynamics and variability of the Antarctic sea ice and its impacts on the climate system as well as to explore the potential of using its sea-ice forecasts for decision-making.
+This comparison will also inform future work with the prediction system as a research tool to better understand the dynamics and variability of the Antarctic sea ice and its impacts on the climate system as well as to explore the potential of using its sea ice forecasts for decision-making.
 The work will also serve as a benchmark for future prediction systems to attempt to improve upon.
@@
 Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for their hindcasts.
-The main difference between the hindcasts of the two systems are the ocean and sea-ice initial conditions.
-+S1 ocean initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea-ice concentration.
+The main difference between the hindcasts of the two systems are the ocean and sea ice initial conditions.
++S1's ocean and sea ice initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and satellite observations of sea-ice concentration using the EUMETSAT OSISAF product described in the next section.
 +S2, on the other hand, is initialised from ocean conditions generated by the BoM weakly coupled ensemble data assimilation scheme described in @wedd2022.
@@
 The atmosphere fields from that daily integration are not used and instead the model atmosphere is initialised using ERA-Interim.
-The sea-ice initial conditions for the next cycle are the unaltered output of the previous daily integration.
+The sea ice initial conditions for the next cycle are the unaltered output of the previous daily integration.
 Then the cycle starts again and the coupled model runs for another 24 hours.
@@
 
+The +S1 hindcast set is made up of nine members created by perturbing the atmospheric fields only with a random field perturbation [@hudson2017] and runs for 217 days for the period 1990–2012 initialised at the first of every month.
 The +S2 hindcast set used in this study runs for the period 1981–2018.
-Each hindcast consists of nine ensemble members built from three consecutive 3-member forecasts initialised at the first of every month and the two previous days and run for 279 days.
-The +S1 hindcast set is built by perturbing the atmospheric fields only with a random field perturbation [@hudson2017] and runs for 217 days for the period 1990–2012.
-We analise the ensemble mean hindcasts unless otherwise specified.
+Each hindcast consists of nine ensemble members built from three consecutive 3-member forecasts created in the same manner as +S1 members initialised at the first of every month and the two previous days and run for 279 days.
+We analyse the ensemble mean hindcasts unless otherwise specified.
 
@@
 This serves as a first-order correction of model bias and drift.
+For monthly means, we define "0 lead time months" as the monthly mean forecast of the same month of initialisation.
 
@@
 We compute RMSE as the square root of the area-averaged squared differences between grid-point forecasted and observed sea-ice concentration anomalies.
-We compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a zonally-varying RMSE computed over twenty-four 15° longitude slices around Antarctica.
+We compute a pan-Antarctic RMSE by averaging over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a zonally-varying RMSE computed over 15 longitude slices 24° wide around Antarctica.
 
@@
 
-@fig-hindcast-extent shows the mean sea-ice extent of the +S2 and +S1 hindcasts (row a) and their differences from the mean sea-ice extent of NOAA/NSIDC CDRV4 (row b).
+@fig-hindcast-extent shows mean sea-ice extent of the +S1 and +S2 hindcasts (row a) and their differences from mean sea-ice extent of +CDR (row b).
 Mean extent at the first of every month is indicated with circles for the initial conditions and with triangles for the longest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
-At this long lead time, the information of the initial conditions is essentially lost and the forecast reverts to each model's preferred equilibrium state.
-
-+S2 initial conditions (circles; @fig-hindcast-extent column 1) show an overall negative bias, especially in the late summer-early autumn, while +S1 initial conditions (circles in @fig-hindcast-extent column 2) are very close to observations, as expected from the assimilation of sea-ice observations to produce the initial conditions of +S1.
+At this long lead time, information of the initial conditions is essentially lost and the forecast reverts to each model's preferred equilibrium state.
+
++S2 initial conditions (circles in [Fig.
+@fig-hindcast-extent] column 2) show an overall negative bias, especially in the late summer-early autumn, while +S1 initial conditions (circles in [Fig.
+@fig-hindcast-extent] column 1) are very close to observations, as expected from the assimilation of sea-ice observations to produce the initial conditions of +S1.
 Both systems' equilibrium states (triangles) show negative biases of sea-ice extent, particularly in the growth phase of late-autumn and winter months.
@@
 
-The difference between the initial conditions (circles) and the model equilibrium state (triangles) can be mostly attributed to the effect of data assimilation, which in +S2 is due solely to atmospheric and oceanic data assimilation.
-From May to October, in +S2 circles are closer to observations than to the triangles, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
+The difference between the initial conditions (circles) and the model equilibrium state (triangles) can be mostly attributed to the effect of data assimilation, which in +S2 is due solely to the coupling of sea-ice with the atmosphere and the ocean.
+From May to October, in +S2 circles are closer to observations than the triangles are, indicating that the information from the ocean and atmosphere data assimilation is affecting sea ice and improving the initial conditions.
 During these months, +S1 can overestimate the sea-ice extent at short lead time.
@@
 ```{r fig-mean-growth}
-#| fig-cap:  Mean daily sea-ice extent growth (10^6 km^2/day) in +S1 (green) and +S2 (purple) hindcasts and observations (black), computed as the mean daily differences in sea-ice extent between each date and the next for each forecast month.
+#| fig-cap:  Mean daily sea-ice extent growth ($10^6 km^2/day$) in +S1 (green) and +S2 (purple) hindcasts and observations (black), computed as the mean daily differences in sea-ice extent between each date and the next for each forecast month.
 Values are smoothed with a 11-day running mean.
@@
 +S1 has a comparatively smaller overall bias (@fig-bias-2). 
-The largest values are found between April and June, when the faster growth resutls in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen seas.
+The largest values are found between April and June, when the faster growth results in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen seas.
 
@@
   ggplot(aes(time)) +
-  geom_line(aes(y = aice, group = dataset, color = dataset)) +
-  geom_line(aes(y = forecast, color = model)) +
+  geom_line(aes(y = aice, group = dataset, color = dataset), linewidth = 0.2) +
+  geom_line(aes(y = forecast, color = model), linewidth = 0.2) +
   geom_text(
@@
     y = Inf,
-    size = 3,
-    vjust = 1.2,
+    size = 2.5,
+    vjust = 2.5,
     hjust = 0
@@
     aes(label = paste0("cor: ", text)),
-    vjust = -0.8,
+    vjust = -1.5,
     hjust = 0,
-    size = 3,
+    size = 2.5,
     x = -Inf,
@@
 The explanation seems to be the mentioned increased interannual variability.
-@fig-extent-sd shows monthly mean sea-ice extent of the forecasts as a function of lead time compared with observations.
+@fig-extent-sd shows the interannual standard deviation of monthly sea-ice extent of the forecasts as a function of lead time compared with observations.
 +S1 standard deviation lies within the observed standard deviation regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at 9 month lead time for most months.
@@
     aes(label = format(time, "%d %b")),
+    size = 2,
     vjust = -0.4
@@
 
-To analyse the spatial distribution of the model error, we computed the RMSE of zonal mean sea-ice concentration anomalies on `r uniqueN(rmse_lon_mean$lon)` slices of `diff(unique(rmse_lon_mean$lon))[1]`° longitude span for each forecasting system.
+To analyse the spatial distribution of the model error, we computed the RMSE of zonal mean sea-ice concentration anomalies on `r uniqueN(rmse_lon_mean$lon)` slices of `r diff(unique(rmse_lon_mean$lon))[1]`° longitude span for each forecasting system.
 We control for some areas being naturally easier to forecast than others by computing the RMSE skill score with the climatological forecast RMSE as reference.
@@
   ggplot(aes(month, sqrt(value))) +
-  geom_line(aes(color = version), linewidth = 1) +
+  geom_line(aes(color = version), linewidth = 0.5) +
   scale_y_continuous("Standard deviation") +
@@
 
-In Figure 5 the mean error was shown. 
+In @fig-rmse the mean error was shown. 
 @fig-initial-spread column 1 shows the mean standard deviation of errors among ensemble members at various lead times.
@@
 
-Sea-ice forecasts from the +S2 system show a significant low extent bias, particularly during late summer and early autumn. 
+Sea ice forecasts from the +S2 system show a significant low extent bias, particularly during late summer and early autumn. 
 This bias is attributed to a faster and longer melt season between January and March, and slower growth between March and April. 
@@
 Ensemble spread grows quickly even when perturbations are only implemented in the atmosphere component (in +S1), indicating that sea-ice is indeed responding quickly to atmospheric perturbations. 
-However, our analysis suggests that the atmosphere and ocean data assimilation implemented in +S2 is only effectively influencing sea-ice initial conditions from June to October, while the rest of the year, the sea-ice component runs virtually free, reverting to its biased equilibrium state.
+However, our analysis suggests that the atmosphere and ocean data assimilation implemented in +S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea-ice component runs virtually free, reverting to its biased equilibrium state.
 
@@
 Both the explicit data assimilation in +S2 and the effects of atmospheric and oceanic data assimilation in +S1 might be nudging simulated sea ice to be thicker than the model equilimbrium state.  
-We believe that the thinner sea-ice in +S2 is a partial cause for the large sea-ice extent variance, but other mechanisms, such as unbalanced initial conditions might be important.
+We believe that the thinner sea ice in +S2 is a partial cause for the large sea-ice extent variance, but other mechanisms, such as unbalanced initial conditions might be important.
 
 Given that +S2 sea-ice extent is essentially not initialised by the observations, comparing its forecasts with those of +S1 allows us to estimate the time-scale for which initial conditions are important.
-We find that initial conditions affect Antarctic sea-ice forecasts in ther order of a few months, but that effect is seasonally dependent. 
+We find that initial conditions affect Antarctic sea ice forecasts in the order of a few months, but that effect is seasonally dependent. 
 January to April initial conditions improve forecasts for up to 3 months with, February initial conditions in particular are shown to be crucial for determining sea-ice evolution at least up to May. 
-Arctic sea-ice forecasts also show greater sensitivity to initial conditions in boreal summer, compared with boreal winter [@day2014; @bunzel2016], suggesting a similar mechanism might be playing a role.
+Arctic sea ice forecasts also show greater sensitivity to initial conditions in boreal summer, compared with boreal winter [@day2014; @bunzel2016], suggesting a similar mechanism might be playing a role.
 
@@
 
+\clearpage
+
 \appendix  

### 2025-10-09 R/ggplot.R
Makes plot font a bit smaller

@@ -15,7 +15,7 @@ element_grob.element_text_first_line <-
 
 
 ggplot2::theme_set(
-  ggplot2::theme_minimal() +
+  ggplot2::theme_minimal(base_size = 9) +
     ggplot2::theme(
       panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
       legend.position = "bottom",

### 2025-10-09 R/functions.R
Format functions.R and also uses em dash instead of --

@@ -1,16 +1,21 @@
 median_ci <- function(x, ci = .95) {
-  
   # list(mid = mean(x, na.rm = TRUE))
-  quantile(na.omit(x),  probs = c(0.5 - ci/2, 0.5, 0.5 + ci/2), names = FALSE) |>
+  quantile(
+    na.omit(x),
+    probs = c(0.5 - ci / 2, 0.5, 0.5 + ci / 2),
+    names = FALSE
+  ) |>
     setNames(c("low", "estimate", "high")) |>
     as.list()
 }
 
 average <- function(x) {
   test <- t.test(x)
-  list(estimate = test$estimate,
-       low = test$conf.int[1],
-       high = test$conf.int[2])
+  list(
+    estimate = test$estimate,
+    low = test$conf.int[1],
+    high = test$conf.int[2]
+  )
 }
 
 
@@ -23,33 +28,37 @@ ncview <- function(file) {
 }
 
 get_forecast_times <- function(model) {
-  
-  dir <- switch(model,
-                S1 =  "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/",
-                S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/",
-                stop("Model needs to be S1 or S2")
+  dir <- switch(
+    model,
+    S1 = "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/",
+    S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/aice/daily/",
+    stop("Model needs to be S1 or S2")
   )
-  
+
   dates <- list.files(file.path(dir, "e01")) |>
     strcapture(
       pattern = "di_a?ice_(\\d{8})_e01.nc",
       proto = list(date = character(1))
-    ) |> 
-    _$date |> 
+    ) |>
+    _$date |>
     lubridate::as_date()
-  
-  
+
   dates
 }
 
-compute_climatology <- function(dataset,
-                                output,
-                                climatology = 1981:2011,
-                                runmean_number = 11,
-                                year = 2000,
-                                options = c("-L")) {
-  dates <- c(paste0(min(climatology), "-01-01"), paste0(max(climatology), "-12-31"))
-  
+compute_climatology <- function(
+  dataset,
+  output,
+  climatology = 1981:2011,
[…]

### 2025-10-09 analysis/paper/access-ice.tex
Renders

@@
 \hypersetup{
-  pdftitle={Evaluating Seasonal Predictability of Antarctic Sea Ice and the importance of initial conditions with fully coupled models},
+  pdftitle={Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Model},
   pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Matthew Wheeler; Phillip Reid},
-  pdfkeywords={sea ice, predictabilty, initial conditions, forecasting},
+  pdfkeywords={sea ice, predictability, initial
+conditions, forecasting},
   colorlinks=true,
@@
 
-\title[Evaluating Seasonal Predictability of Antarctic Sea Ice and the
-importance of initial conditions with fully coupled models]{Evaluating
-Seasonal Predictability of Antarctic Sea Ice and the importance of
-initial conditions with fully coupled models}
+\title[Evaluating the Importance of Initial Conditions for Antarctic Sea
+Ice Seasonal Predictability with a Fully Coupled Model]{Evaluating the
+Importance of Initial Conditions for Antarctic Sea Ice Seasonal
+Predictability with a Fully Coupled Model}
 
@@
 
-\abstract{Accurate Antarctic sea-ice forecasts are crucial for climate
+\abstract{Accurate Antarctic sea ice forecasts are crucial for climate
 monitoring and operational planning, yet they remain challenging due to
-model biases and complex ice-ocean-atmosphere interactions. We
-investigate the role of initial conditions on Antarctic sea-ice
-forecasts by comparing hindcasts from two versions of the Australian
-Bureau of Meteorology's ACCESS seasonal forecast system: ACCESS-S1 and
-ACCESS-S2. They use identical model configuration and differ only in
-their initial conditions; primarily in that ACCESS-S2 does not
-assimilate sea-ice observations. Our analysis reveals that both systems
-experience an extended melt season and delayed growth phase. This leads
-to a significant negative sea-ice extent bias which is corrected only in
-ACCESS-S1 by the data assimilation system. The impact of the initial
-conditions on forecast errors varies dramatically by season: summer and
-autumn initial conditions (January-April) provide predictive skill for
-up to three months, with February initial conditions being particularly
-crucial. In contrast, winter forecasts off the two systems are
-statistically indistinguishable after only two weeks. Regional analysis
-of forecast skill suggests that this winter predictability barrier is
-most dramatic over Eastern Antarctica, where evern ACCESS-S1 shows
-negative skill. These findings highlight the critical importance of
-comprehensive year-round sampling in predictability studies and suggest
-that operational sea-ice data assimilation efforts should prioritize the
-summer-autumn period when initial conditions have maximum impact on
-forecast skill.}
+model biases and complex ice-ocean-atmosphere interactions. The two
+versions of the Australian Bureau of Meteorology's ACCESS seasonal
+forecast system, ACCESS-S1 and ACCESS-S2, use identical model
+configuration and differ only in their initial conditions; primarily in
+that ACCESS-S2 does not assimilate sea-ice observations. This makes for
+a good ``natural experiment'' to analyse the role of initial conditions
+on Antarctic sea ice forecasts using a long run and a moderately large
+ensemble with a fully coupled model. Our analysis reveals that both
+systems experience an extended melt season and delayed growth phase.
+This leads to a significant negative sea-ice extent bias which is
+corrected only in ACCESS-S1 by the data assimilation system. The impact
+of the initial conditions on forecast errors varies dramatically by
+season: summer and autumn initial conditions (January-April) provide
+predictive skill for up to three months, with February initial
+conditions being particularly crucial. In contrast, winter forecasts of
+the two systems are statistically indistinguishable after only two
+weeks. Regional analysis of forecast skill suggests that this winter
+predictability barrier is most dramatic over Eastern Antarctica, where
+even ACCESS-S1 shows negative skill. These findings highlight the
+critical importance of comprehensive year-round sampling in
+predictability studies and suggest that operational sea-ice data
+assimilation efforts should prioritize the summer-autumn period when
+initial conditions have maximum impact on forecast skill.}
 
 % keywords
-\keywords{sea ice,  predictabilty,  initial conditions,  forecasting}
+\keywords{sea ice,  predictability,  initial conditions,  forecasting}
 
@@
 
-However, producing accurate Antarctic sea-ice forecasts has been
-challenging due to model biases, inherent large variability and
-complexity, and they have lagged behind Arctic sea-ice
-forecasts\citep{zampieri2019, gao2024a}. Dynamical seasonal forecasts of
-summer Antarctic sea ice have been shown to perform worse than
-relatively simple statistical methods \citep{massonnet2023} and machine
-learning approaches \citep[e.g.][]{dong2024, lin2025}, which also
+However, Antarctic sea ice forecasts have lagged behind Arctic sea ice
+forecasts due to model biases, inherent large variability and complexity
+\citep{zampieri2019, gao2024a}. Dynamical seasonal forecasts of summer
+Antarctic sea ice have been shown to perform worse than relatively
+simpler statistical methods \citep{massonnet2023} and machine learning
+approaches (e.g. \citet{dong2024}, \citet{lin2025}), which also
 underscores the need for better understanding of sea-ice dynamics, and
@@
 Good initial conditions are generally required for a good forecast,
-however, it is not entirely known to what extent accurate sea-ice
+however, it is not entirely known to what extent accurate sea ice
 initial conditions affect the quality of the forecast and at what
-timescales. Exploring seasonal predictions of Arctic sea-ice,
-\citet{guemas2016} found that sea-ice initial conditions are important
-in autumn to predict summer sea-ice, but the impact wasn't as dramatic
-when predicting winter sea-ice. \citet{day2014} also found
+timescales. Exploring seasonal predictions of Arctic sea ice,
+\citet{guemas2016} found that sea ice initial conditions are important
+in autumn to predict summer sea ice, but the impact wasn't as dramatic
+when predicting winter sea ice. \citet{day2014} also found
 seasonally-varying differences in the effect of initialisation, noting
-that accurate Arctic sea-ice thickness lead to improved sea-ice
+that accurate Arctic sea-ice thickness lead to improved sea ice
 forecasts initialised in July but not when initialised in January.
@@
 For the Antarctic, \citet{holland2013} studied the initial-value
-predictability of Antarctic sea-ice in a perfect model study using the
+predictability of Antarctic sea ice in a perfect model study using the
 CCSM3 model. They found that sea-ice and ocean initial conditions
@@
 conditions having lesser importance. More recently, \citet{morioka2022}
-studied decadal forecasts of Antarctic sea-ice and found that
-initialising ocean and sea-ice improved the correlation between
+studied decadal forecasts of Antarctic sea ice and found that
+initialising ocean and sea ice improved the correlation between
 simulated and observed sea-ice concentration evolution in the
@@
 Australian Community Climate and Earth System Simulator -- Seasonal
-(ACCES-S) from version S1 to S2. While the base model remained the same,
-the change in version was focused on using ocean, sea-ice and land
+(ACCESS-S) from version S1 to S2. While the base model remained the
+same, the change in version was focused on using ocean, sea-ice and land
 initial conditions generated by the BoM instead of depending on the UK
@@
 being the initial conditions. This provides an opportunity to test the
-effect of sea-ice initial conditions on the forecast of sea-ice
+effect of sea ice initial conditions on the forecast of sea-ice
 concentrations and the climate.
 
-In this study we compare sea-ice hindcasts produced by ACCESS-S1 and
+In this study we compare sea ice hindcasts produced by ACCESS-S1 and
 ACCESS-S2. We focus on seasonality of errors and biases and the effect
@@
 the dynamics and variability of the Antarctic sea ice and its impacts on
-the climate system as well as to explore the potential of using its
-sea-ice forecasts for decision-making. The work will also serve as a
+the climate system as well as to explore the potential of using its sea
+ice forecasts for decision-making. The work will also serve as a
 benchmark for future prediction systems to attempt to improve upon.
@@
 ERA-interim \citep{dee2011} for their hindcasts. The main difference
-between the hindcasts of the two systems are the ocean and sea-ice
-initial conditions. ACCESS-S1 ocean initial conditions come from the Met
-Office FOAM system, which uses a multivariate, incremental
-three-dimensional variational (3D-Var), first-guess-at-appropriate-time
-(FGAT) data assimilation scheme \citep{waters2015} and assimilates sea
-surface temperature (SST), sea surface height (SSH), in situ temperature
-and salinity profiles, and sea-ice concentration. ACCESS-S2, on the
-other hand, is initialised from ocean conditions generated by the BoM
-weakly coupled ensemble data assimilation scheme described in
-\citet{wedd2022}. This scheme uses an optimal interpolation method and
-assimilates temperature and salinity profiles from EN4 \citep{good2013}.
-SSTs are nudged to Reynolds OISSTv2.1 \citep{reynolds2007} in areas
-where SSTs are over 0ºC and Sea Surface Salinity is weakly nudged to the
-World Ocean Atlas 2013 climatology \citep{zweng2013}.
+between the hindcasts of the two systems are the ocean and sea ice
+initial conditions. ACCESS-S1's ocean and sea ice initial conditions
+come from the Met Office FOAM system, which uses a multivariate,
+incremental three-dimensional variational (3D-Var),
+first-guess-at-appropriate-time (FGAT) data assimilation scheme
+\citep{waters2015} and assimilates sea surface temperature (SST), sea
+surface height (SSH), in situ temperature and salinity profiles, and
+satellite observations of sea-ice concentration using the EUMETSAT
+OSISAF product described in the next section. ACCESS-S2, on the other
+hand, is initialised from ocean conditions generated by the BoM weakly
+coupled ensemble data assimilation scheme described in \citet{wedd2022}.
+This scheme uses an optimal interpolation method and assimilates
+temperature and salinity profiles from EN4 \citep{good2013}. SSTs are
+nudged to Reynolds OISSTv2.1 \citep{reynolds2007} in areas where SSTs
+are over 0ºC and Sea Surface Salinity is weakly nudged to the World
+Ocean Atlas 2013 climatology \citep{zweng2013}.
 
@@
 atmosphere fields from that daily integration are not used and instead
-the model atmosphere is initialised using ERA-Interim. The sea-ice
+the model atmosphere is initialised using ERA-Interim. The sea ice
 initial conditions for the next cycle are the unaltered output of the
@@
 
-The ACCESS-S2 hindcast set used in this study runs for the period
-1981--2018. Each hindcast consists of nine ensemble members built from
-three consecutive 3-member forecasts initialised at the first of every
-month and the two previous days and run for 279 days. The ACCESS-S1
-hindcast set is built by perturbing the atmospheric fields only with a
-random field perturbation \citep{hudson2017} and runs for 217 days for
-the period 1990--2012. We analise the ensemble mean hindcasts unless
-otherwise specified.
+The ACCESS-S1 hindcast set is made up of nine members created by
+perturbing the atmospheric fields only with a random field perturbation
+\citep{hudson2017} and runs for 217 days for the period 1990--2012
+initialised at the first of every month. The ACCESS-S2 hindcast set used
+in this study runs for the period 1981--2018. Each hindcast consists of
+nine ensemble members built from three consecutive 3-member forecasts
+created in the same manner as ACCESS-S1 members initialised at the first
+of every month and the two previous days and run for 279 days. We
+analyse the ensemble mean hindcasts unless otherwise specified.
 
@@
 for the period 1990--2012. This serves as a first-order correction of
-model bias and drift.
+model bias and drift. For monthly means, we define ``0 lead time
+months'' as the monthly mean forecast of the same month of
+initialisation.
 
@@
 over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a
-zonally-varying RMSE computed over twenty-four 15° longitude slices
-around Antarctica.
+zonally-varying RMSE computed over 15 longitude slices 24° wide around
+Antarctica.
 
@@
 
-Figure~\ref{fig-hindcast-extent} shows the mean sea-ice extent of the
-ACCESS-S2 and ACCESS-S1 hindcasts (row a) and their differences from the
+Figure~\ref{fig-hindcast-extent} shows mean sea-ice extent of the
+ACCESS-S1 and ACCESS-S2 hindcasts (row a) and their differences from
 mean sea-ice extent of NOAA/NSIDC CDRV4 (row b). Mean extent at the
@@
 each model (between 274 and 277 days for ACCESS-S2 and between 213 and
-216 days for ACCESS-S1). At this long lead time, the information of the
+216 days for ACCESS-S1). At this long lead time, information of the
 initial conditions is essentially lost and the forecast reverts to each
@@
 
-ACCESS-S2 initial conditions (circles; Figure~\ref{fig-hindcast-extent}
-column 1) show an overall negative bias, especially in the late
+ACCESS-S2 initial conditions (circles in Fig.~\ref{fig-hindcast-extent}
+column 2) show an overall negative bias, especially in the late
 summer-early autumn, while ACCESS-S1 initial conditions (circles in
-Figure~\ref{fig-hindcast-extent} column 2) are very close to
-observations, as expected from the assimilation of sea-ice observations
-to produce the initial conditions of ACCESS-S1. Both systems'
-equilibrium states (triangles) show negative biases of sea-ice extent,
-particularly in the growth phase of late-autumn and winter months. This
-is due primarily to the melt season being longer than in observations
-and with faster melt between January and March and the growing seasons
-being shorter with slower growth during March and April. This is then
-followed by faster growth between May and July
-(Figure~\ref{fig-mean-growth}). Many sea-ice models exhibit this
-systematic underestimation during the sea-ice minimum and early freezing
-season \citep{massonnet2023}, which could indicate problems in the
-representation of thermodynamics in the model \citep{zampieri2019}. It
-is also not surprising that both forecasting systems converge to a
-similar equilibrium state because they share the same model formulation.
+Fig.~\ref{fig-hindcast-extent} column 1) are very close to observations,
+as expected from the assimilation of sea-ice observations to produce the
+initial conditions of ACCESS-S1. Both systems' equilibrium states
+(triangles) show negative biases of sea-ice extent, particularly in the
+growth phase of late-autumn and winter months. This is due primarily to
+the melt season being longer than in observations and with faster melt
+between January and March and the growing seasons being shorter with
+slower growth during March and April. This is then followed by faster
+growth between May and July (Figure~\ref{fig-mean-growth}). Many sea-ice
+models exhibit this systematic underestimation during the sea-ice
+minimum and early freezing season \citep{massonnet2023}, which could
+indicate problems in the representation of thermodynamics in the model
+\citep{zampieri2019}. It is also not surprising that both forecasting
+systems converge to a similar equilibrium state because they share the
+same model formulation.
 
@@
 equilibrium state (triangles) can be mostly attributed to the effect of
-data assimilation, which in ACCESS-S2 is due solely to atmospheric and
-oceanic data assimilation. From May to October, in ACCESS-S2 circles are
-closer to observations than to the triangles, indicating that the
-information from the ocean and atmosphere data assimilation is affecting
-sea ice and improving the initial conditions. During these months,
-ACCESS-S1 can overestimate the sea-ice extent at short lead time. For
-the rest of the year circles are overlaid with triangles in ACCESS-S2,
-indicating that the ocean and atmosphere data assimilation is not
-affecting sea ice and that this component of the model is virtually
-free-running.
+data assimilation, which in ACCESS-S2 is due solely to the coupling of
+sea-ice with the atmosphere and the ocean. From May to October, in
+ACCESS-S2 circles are closer to observations than the triangles are,
+indicating that the information from the ocean and atmosphere data
+assimilation is affecting sea ice and improving the initial conditions.
+During these months, ACCESS-S1 can overestimate the sea-ice extent at
+short lead time. For the rest of the year circles are overlaid with
+triangles in ACCESS-S2, indicating that the ocean and atmosphere data
+assimilation is not affecting sea ice and that this component of the
+model is virtually free-running.
 
@@
 \caption{\label{fig-mean-growth}Mean daily sea-ice extent growth
-(10\^{}6 km\^{}2/day) in ACCESS-S1 (green) and ACCESS-S2 (purple)
+(\(10^6 km^2/day\)) in ACCESS-S1 (green) and ACCESS-S2 (purple)
 hindcasts and observations (black), computed as the mean daily
@@
 (Figure~\ref{fig-bias-2}). The largest values are found between April
-and June, when the faster growth resutls in large positive bias along
+and June, when the faster growth results in large positive bias along
 the sea-ice edge, and in January, when the faster melt leads to large
@@
 explanation seems to be the mentioned increased interannual variability.
-Figure~\ref{fig-extent-sd} shows monthly mean sea-ice extent of the
-forecasts as a function of lead time compared with observations.
-ACCESS-S1 standard deviation lies within the observed standard deviation
-regardless of lead time, while ACCESS-S2 standard deviation is more than
-twice that of observations at zero lead time and only approaches the
-observed value at 9 month lead time for most months.
+Figure~\ref{fig-extent-sd} shows the interannual standard deviation of
+monthly sea-ice extent of the forecasts as a function of lead time
+compared with observations. ACCESS-S1 standard deviation lies within the
+observed standard deviation regardless of lead time, while ACCESS-S2
+standard deviation is more than twice that of observations at zero lead
+time and only approaches the observed value at 9 month lead time for
+most months.
 
@@
 To analyse the spatial distribution of the model error, we computed the
-RMSE of zonal mean sea-ice concentration anomalies on 15 slices of
-\texttt{diff(unique(rmse\_lon\_mean\$lon)){[}1{]}}° longitude span for
-each forecasting system. We control for some areas being naturally
-easier to forecast than others by computing the RMSE skill score with
-the climatological forecast RMSE as reference.
+RMSE of zonal mean sea-ice concentration anomalies on 15 slices of 24°
+longitude span for each forecasting system. We control for some areas
+being naturally easier to forecast than others by computing the RMSE
+skill score with the climatological forecast RMSE as reference.
 
@@
 
-In Figure 5 the mean error was shown. Figure~\ref{fig-initial-spread}
-column 1 shows the mean standard deviation of errors among ensemble
-members at various lead times. At one day lead time
-(Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a slightly larger
-spread than ACCESS-S1 due to the way that ensemble members are
-generated. ACCESS-S1 ensemble members are generated by adding random
-field perturbations to the atmosphere only, which then are transferred
-to the other components via the coupled simulation \citep{hudson2017}.
-With this scheme, ensemble members are all but guaranteed to be
-underdispersed in the ocean and sea-ice components. The time-lag
-ensemble used for ACCESS-S2 ensures greater spread. This difference is
-gone after about just two days, and both systems have a comparable
-spread in ensemble member error afterwards
+In Figure~\ref{fig-rmse} the mean error was shown.
+Figure~\ref{fig-initial-spread} column 1 shows the mean standard
+deviation of errors among ensemble members at various lead times. At one
+day lead time (Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a
+slightly larger spread than ACCESS-S1 due to the way that ensemble
+members are generated. ACCESS-S1 ensemble members are generated by
+adding random field perturbations to the atmosphere only, which then are
+transferred to the other components via the coupled simulation
+\citep{hudson2017}. With this scheme, ensemble members are all but
+guaranteed to be underdispersed in the ocean and sea-ice components. The
+time-lag ensemble used for ACCESS-S2 ensures greater spread. This
+difference is gone after about just two days, and both systems have a
+comparable spread in ensemble member error afterwards
 (Fig.~\ref{fig-initial-spread} b1 and c1).
@@
 
-Sea-ice forecasts from the ACCESS-S2 system show a significant low
+Sea ice forecasts from the ACCESS-S2 system show a significant low
 extent bias, particularly during late summer and early autumn. This bias
@@
 assimilation implemented in ACCESS-S2 is only effectively influencing
-sea-ice initial conditions from June to October, while the rest of the
+sea ice initial conditions from June to October, while the rest of the
 year, the sea-ice component runs virtually free, reverting to its biased
@@
 sea ice to be thicker than the model equilimbrium state.\\
-We believe that the thinner sea-ice in ACCESS-S2 is a partial cause for
+We believe that the thinner sea ice in ACCESS-S2 is a partial cause for
 the large sea-ice extent variance, but other mechanisms, such as
@@
 us to estimate the time-scale for which initial conditions are
-important. We find that initial conditions affect Antarctic sea-ice
-forecasts in ther order of a few months, but that effect is seasonally
+important. We find that initial conditions affect Antarctic sea ice
+forecasts in the order of a few months, but that effect is seasonally
 dependent. January to April initial conditions improve forecasts for up
@@
 be crucial for determining sea-ice evolution at least up to May. Arctic
-sea-ice forecasts also show greater sensitivity to initial conditions in
+sea ice forecasts also show greater sensitivity to initial conditions in
 boreal summer, compared with boreal winter \citep{day2014, bunzel2016},
@@
 \caption{\label{fig-mean-growth-osi}Mean daily sea-ice extent growth
-(10\^{}6 km\^{}2/day) in ACCESS-S1 (green) and ACCESS-S2 (purple)
+(\(10^6 km^2/day\)) in ACCESS-S1 (green) and ACCESS-S2 (purple)
 hindcasts and observations (black), computed as the mean daily

### 2025-10-09 analysis/paper/access-ice.qmd
Adds acknowledgments and code + data (placeholder)

@@
 
+# Acknowledgments
+
+This work was supported by ARC SRIEAS Grant SR200100005 Securing Antarctica’s Environmental Future.
+
+# Code and data availability
+
+The underlying code for this study is available on github and can be accessed via this link https://github.com/eliocamp/access-s2_ice-eval.
+
+Raw data of +S1 and +S2 forecast are not available due to size. 
+Derived datasets required to reproduce the results are available in this zenodo repository: XXXXX
+
 # References

### 2025-10-09 analysis/paper/access-ice.tex
Adds acknowledgments and code + data (placeholder)

@@
 
+\section{Acknowledgments}\label{acknowledgments}
+
+This work was supported by ARC SRIEAS Grant SR200100005 Securing
+Antarctica's Environmental Future.
+
+\section{Code and data availability}\label{code-and-data-availability}
+
+The underlying code for this study is available on github and can be
+accessed via this link https://github.com/eliocamp/access-s2\_ice-eval.
+
+Raw data of ACCESS-S1 and ACCESS-S2 forecast are not available due to
+size. Derived datasets required to reproduce the results are available
+in this zenodo repository: XXXXX
+
 \section{References}\label{references}

### 2025-10-23 analysis/paper/access-ice.qmd
Adds zenodo data management

@@
   message: false
-  cache: true
+  cache: refresh
 
@@
 cdo_cache_set(here::here("data/temp/cache"))
+```
+
+```{r zenodo, cache = FALSE}
+source(here::here("R/zenodo.R"))
+zenodo_init()
+zenodo_download_data()
 ```
@@
 ```{r hindcast_extent}
-hindcast_extent <- fread(here::here("data/derived/hindcast_extent.csv")) |>
+hindcast_extent <- here::here("data/derived/hindcast_extent.csv") |>
+  zenodo() |>
+  fread() |>
   _[, model := relevel(factor(model), "S2")] |>
@@
 ```{r hindcast_clim}
-zeropad <- function(x, n = 2) formatC(x, width = 2, flag = "0")
+
+hindcast_clim <- hindcast_0lag_clim() |>
+  _[, ReadNetCDF(file, "aice"), by = .(model)] |>
+  _[, month := month(time)] |>
+  _[, time := NULL]
+
 monnb <- function(d) {
@@
 }
-
-months <- 1:12
-
-hindcast_clim <- lapply(setNames(months, months), \(month) {
-  lapply(c("S1", "S2"), \(model) {
-    here::here(glue::glue(
-      "data/derived/climatology/{model}/{zeropad(month)}/em.nc"
-    )) |>
-      cdo_selmonth(month) |>
-      cdo_execute() |>
-      ReadNetCDF("aice") |>
-      _[, model := model]
-  }) |>
-    rbindlist()
-}) |>
-  rbindlist(idcol = "forecast_month") |>
-  _[, forecast_month := as.numeric(forecast_month)]
-
-hindcast_clim <- hindcast_clim |>
-  copy() |>
-  _[,
-    .(aice = mean(aice)),
-    by = .(month(time), ygrid, xgrid, forecast_month, model)
-  ] |>
-  _[forecast_month == month]
 
@@
     month_clim <- formatC(month(forecast_date), width = 2, flag = "0")
-
+    clim <- here::here(glue::glue(
+      "data/derived/climatology/{model}/{month_clim}/em.nc"
+    )) |>
+      zenodo()
     hindcast(forecast_date, model, 1) |>
-      cdo_ydaysub(here::here(glue::glue(
-        "data/derived/climatology/{model}/{month_clim}/em.nc"
-      ))) |>
+      cdo_ydaysub(clim) |>
       cdo_execute() |>
@@
 ) |>
+  zenodo() |>
   purrr::imap(\(f, model) {
@@
 
-mean_thickness <- function(concentration, thickness, cell_area = NULL) {
-  if (is.null(cell_area)) {
-    cell_area <- concentration |>
-      rcdo::cdo_gridarea()
-  }
-
-  volume <- thickness |>
-    rcdo::cdo_mul(concentration) |>
-    rcdo::cdo_mul(cell_area) |>
-    rcdo::cdo_fldsum()
-
-  area <- concentration |>
-    rcdo::cdo_mul(cell_area) |>
-    rcdo::cdo_fldsum()
-
-  rcdo::cdo_div(volume, area) |>
-    rcdo::cdo_execute(options = "-L")
-}
-mean_thickness_access <- \(forecast_time, model) {
-  concentration <- hindcast(
-    forecast_time,
-    model = model,
-    members = "em",
-    variable = "aice"
-  )
-  thickness <- hindcast(
-    forecast_time,
-    model = model,
-    members = "em",
-    variable = "hi"
-  )
-
-  if (length(thickness) == 0) {
-    return(NA_character_)
-  }
-  if (length(concentration) == 0) {
-    return(NA_character_)
-  }
-
-  rcdo::cdo_cache_set(here::here("data/temp/cache/"))
-  cell_area <- thickness |>
-    rcdo::cdo_gridarea()
-
-  mean_thickness(concentration, thickness, cell_area)
-}
+# mean_thickness <- function(concentration, thickness, cell_area = NULL) {
+#   rcdo::cdo_cache_set(here::here("data/temp/cache/"))
+#   if (is.null(cell_area)) {
+#     cell_area <- thickness |>
+#       rcdo::cdo_gridarea()
+#   }
+
+#   volume <- thickness |>
+#     rcdo::cdo_mul(concentration) |>
+#     rcdo::cdo_mul(cell_area) |>
+#     rcdo::cdo_fldsum()
+
+#   area <- concentration |>
+#     rcdo::cdo_mul(cell_area) |>
+#     rcdo::cdo_fldsum()
+
+#   rcdo::cdo_div(volume, area) |>
+#     rcdo::cdo_execute(options = "-L")
+# }
+
+# mean_thickness_access <- \(forecast_time, model) {
+#   concentration <- hindcast(
+#     forecast_time,
+#     model = model,
+#     members = "em",
+#     variable = "aice"
+#   )
+#   thickness <- hindcast(
+#     forecast_time,
+#     model = model,
+#     members = "em",
+#     variable = "hi"
+#   )
+
+#   if (length(thickness) == 0) {
+#     return(NA_character_)
+#   }
+#   if (length(concentration) == 0) {
+#     return(NA_character_)
+#   }
+
+#   cell_area <- thickness |>
+#     rcdo::cdo_gridarea()
+
+#   mean_thickness(concentration, thickness, cell_area)
+# }
 ```
@@
 ```{r hindcast_thickness}
-hindcast_thickness <- CJ(model = c("S2", "S1")) |>
-  _[, .(forecast_time = get_forecast_times(model)), by = model] |>
-  _[mday(forecast_time) == 1] |>
-  _[,
-    thickness := unlist(mirai_map(
-      data.frame(forecast_time, model),
-      mean_thickness_access,
-      hindcast = hindcast,
-      mean_thickness = mean_thickness
-    )[.progress])
-  ] |>
-  na.omit()
+# hindcast_thickness <- CJ(model = c("S2", "S1")) |>
+#   _[, .(forecast_time = get_forecast_times(model)), by = model] |>
+#   _[mday(forecast_time) == 1] |>
+#   _[,
+#     concentration := hindcast(
+#       forecast_time,
+#       model,
+#       members = "em",
+#       variable = "aice"
+#     ) |>
+#       zenodo(),
+#     by = .(model)
+#   ] |>
+#   _[mday(forecast_time) == 1] |>
+#   _[,
+#     thickness := hindcast(
+#       forecast_time,
+#       model,
+#       members = "em",
+#       variable = "hi"
+#     ) |>
+#       zenodo(),
+#     by = .(model)
+#   ] |>
+#   na.omit() |>
+#   _[,
+#     thickness := unlist(mirai_map(
+#       data.frame(concentration, thickness),
+#       mean_thickness
+#     )[.progress])
+#   ] |>
+#   na.omit()
 
@@
 ```{r hindcast_mean_thickness}
-hindcast_mean_thickness <- hindcast_thickness |>
-  _[,
-    thickness |> cdo_monmean() |> cdo_execute(),
-    by = .(model, forecast_time)
-  ] |>
-  _[,
-    metR::ReadNetCDF(V1, c(thickness = "hi")),
+hindcast_mean_thickness <- hindcast_mean_thickness() |>
+  _[,
+    ReadNetCDF(mean_thickness, c(thickness = "hi")),
     by = .(model, forecast_time)
@@
 ```{r errors}
-errors <- rbind(
-  readRDS(here::here("data/derived/rmse.Rds")),
-  readRDS(here::here("data/derived/iiee.Rds"))
-) |>
+errors <- here::here("data/derived/rmse.Rds") |>
+  zenodo() |>
+  readRDS() |>
   _[value < 0.01, value := NA] |>
@@
 topo_lonlat <- here::here("data/raw/ETOPO.nc") |>
+  zenodo() |>
   ReadNetCDF() |>
@@
 rmse_lon_mean <- here::here("data/derived/rmse_lon.Rds") |>
+  zenodo() |>
   readRDS() |>
@@
 ```{r errors_ensemble} 
-errors_ensemble <- readRDS(here::here("data/derived/rmse.Rds")) |>
+errors_ensemble <- here::here("data/derived/rmse.Rds") |>
+  zenodo() |>
+  readRDS() |>
   _[version != "persistence"] |>

### 2025-10-23 R/datasets.R

@@ -1,4 +1,5 @@
 source(here::here("R/functions.R"))
+source(here::here("R/zenodo.R"))
 
 cdo_apply <- function(ifile, op, ofile = NULL) {
   rcdo::cdo_operator("apply", params = c("op"), Inf, 1) |>
@@ -10,17 +11,18 @@ cdo_apply <- function(ifile, op, ofile = NULL) {
 }
 
 CDR_grid <- function() {
-  here::here("data/raw/nsidc_grid.txt")
+  here::here("data/raw/nsidc_grid.txt") |>
+    zenodo()
 }
 
 CDR <- function() {
   file <- here::here("data/raw/cdr.nc")
   years <- 1981:2023
-  
+
   if (file.exists(file)) {
     return(file)
   }
-  
+
   urls <- paste0(
     "https://polarwatch.noaa.gov/erddap/griddap/nsidcG02202v4sh1day.nc?cdr_seaice_conc[(",
     years,
@@ -28,17 +30,17 @@ CDR <- function() {
     years,
     "-12-31T00:00:00Z)][(4350000.0):1:(-3950000.0)][(-3950000.0):1:(3950000.0)]"
   )
-  
+
   files <- replicate(length(urls), tempfile())
   res <- curl::multi_download(urls, files, resume = TRUE)
-  
+
   files |>
     rcdo::cdo_mergetime() |>
     cdo_chname("cdr_seaice_conc", "aice") |>
     cdo_del29feb() |>
     rcdo::cdo_setgrid(CDR_grid()) |>
     rcdo::cdo_execute(output = file, options = c("-L", "-O"))
-  
+
   remove(files)
   return(file)
 }
@@ -46,11 +48,11 @@ CDR <- function() {
 CDR_BT <- function() {
   file <- here::here("data/raw/cdr_bt.nc")
   years <- 1981:2024
-  
+
   if (file.exists(file)) {
     return(file)
   }
-  
+
   urls <- paste0(
     "https://polarwatch.noaa.gov/erddap/griddap/nsidcG02202v4sh1day.nc?nsidc_bt_seaice_conc[(",
     years,
@@ -58,17 +60,17 @@ CDR_BT <- function() {
     years,
     "-12-31T00:00:00Z)][(4350000.0):1:(-3950000.0)][(-3950000.0):1:(3950000.0)]"
   )
-  
+
   files <- replicate(length(urls), tempfile())
[…]

### 2025-10-23 R/zenodo.R

@@ -1,0 +1,262 @@
+# Minimal Zenodo helpers for a single fixed deposition
+# Uses httr2 for HTTP requests. Keep this file small and focused.
+
+# Hard-coded deposition ID for this project. Replace with the real ID.
+ZENODO_DEPOSITION_ID <- 386004
+
+# Internal: read the token from environment and validate
+.zenodo_get_token <- function() {
+  token <- Sys.getenv("ZENODO_TOKEN", unset = "")
+  if (identical(token, "") || is.na(token)) {
+    stop(
+      "ZENODO_TOKEN environment variable is not set or empty. Set it to your Zenodo API token."
+    )
+  }
+  token
+}
+
+# Internal: choose base API URL. If ZENODO_SANDBOX is set (non-empty), use the sandbox API.
+.zenodo_base_url <- function() {
+  if (nzchar(Sys.getenv("ZENODO_SANDBOX", ""))) {
+    return("https://sandbox.zenodo.org/api")
+  }
+  "https://zenodo.org/api"
+}
+
+# List files attached to the fixed deposition
+# Returns a list (as returned by the API) or throws on error
+zenodo_list_files <- function(deposition_id = ZENODO_DEPOSITION_ID) {
+  token <- .zenodo_get_token()
+  url <- paste0(.zenodo_base_url(), "/deposit/depositions/", deposition_id)
+
+  resp <- httr2::request(url) |>
+    httr2::req_auth_bearer_token(token) |>
+    httr2::req_perform()
+
+  if (httr2::resp_status(resp) >= 400) {
+    stop(sprintf(
+      "Zenodo API returned HTTP %s: %s",
+      httr2::resp_status(resp),
+      httr2::resp_body_json(resp, simplifyVector = TRUE)$message
+    ))
+  }
+
+  body <- httr2::resp_body_json(resp, simplifyVector = FALSE)
+  # body$files is typically a list with elements containing 'filename', 'id', 'links', etc.
+  body$files
+}
+
+
+get_bucket_for_deposition <- function(deposition_id = ZENODO_DEPOSITION_ID) {
+  token <- .zenodo_get_token()
+  url <- paste0(.zenodo_base_url(), "/deposit/depositions/", deposition_id)
+
+  resp <- httr2::request(url) |>
+    httr2::req_auth_bearer_token(token) |>
+    httr2::req_perform()
+
+  if (httr2::resp_status(resp) >= 400) {
+    stop("Failed to fetch deposition: HTTP ", httr2::resp_status(resp))
+  }
+
[…]

### 2025-10-23 analysis/paper/access-ice.qmd
Adds descriptions to zenodo list

@@
 ```{r datasets}
-datasets <- list(cdr = CDR(), osi = OSI())
+datasets <- list(
+  cdr = CDR() |>
+    zenodo("NSIDC CDR sea-ice concentration."),
+  osi = OSI() |>
+    zenodo("OSI sea-ice concentration.")
+)
 which_dataset <- "cdr"
@@
 hindcast_extent <- here::here("data/derived/hindcast_extent.csv") |>
-  zenodo() |>
+  zenodo("Sea-ice extent from hindcast.") |>
   fread() |>
@@
     )) |>
-      zenodo()
+      zenodo("{model} sea-ice concentration climatology for {month_clim}")
+
     hindcast(forecast_date, model, 1) |>
+      zenodo("{model} hindcast for initialised on {forecast_date}") |>
       cdo_ydaysub(clim) |>
@@
 thickness_example <- c(
-  S2 = "/g/data/ux62/access-s2/hindcast/raw_model/ice/hi/daily/e01/di_hi_{format(forecast_date, '%Y%m%d')}_e01.nc",
-  S1 = "/g/data/ub7/access-s1/hc/raw_model/unchecked/ice/ice/daily/e01/di_ice_{format(forecast_date, '%Y%m%d')}_e01.nc"
+  S2 = hindcast(forecast_date, "S2", members = 1, variable = "hi") |>
+    zenodo(
+      "S2 sea-ice thickness hindcast for {forecast_date} (1st ensemble member)"
+    ),
+  S1 = hindcast(forecast_date, "S1", members = 1, variable = "hi") |>
+    zenodo(
+      "S1 sea-ice thickness hindcast for {forecast_date} (1st ensemble member)"
+    )
 ) |>
-  zenodo() |>
   purrr::imap(\(f, model) {
@@
   plot_layout(ncol = 1)
-```
-
-```{r mean_thickness_funs}
-
-# mean_thickness <- function(concentration, thickness, cell_area = NULL) {
-#   rcdo::cdo_cache_set(here::here("data/temp/cache/"))
-#   if (is.null(cell_area)) {
-#     cell_area <- thickness |>
-#       rcdo::cdo_gridarea()
-#   }
-
-#   volume <- thickness |>
-#     rcdo::cdo_mul(concentration) |>
-#     rcdo::cdo_mul(cell_area) |>
-#     rcdo::cdo_fldsum()
-
-#   area <- concentration |>
-#     rcdo::cdo_mul(cell_area) |>
-#     rcdo::cdo_fldsum()
-
-#   rcdo::cdo_div(volume, area) |>
-#     rcdo::cdo_execute(options = "-L")
-# }
-
-# mean_thickness_access <- \(forecast_time, model) {
-#   concentration <- hindcast(
-#     forecast_time,
-#     model = model,
-#     members = "em",
-#     variable = "aice"
-#   )
-#   thickness <- hindcast(
-#     forecast_time,
-#     model = model,
-#     members = "em",
-#     variable = "hi"
-#   )
-
-#   if (length(thickness) == 0) {
-#     return(NA_character_)
-#   }
-#   if (length(concentration) == 0) {
-#     return(NA_character_)
-#   }
-
-#   cell_area <- thickness |>
-#     rcdo::cdo_gridarea()
-
-#   mean_thickness(concentration, thickness, cell_area)
-# }
-```
-
-```{r hindcast_thickness}
-# hindcast_thickness <- CJ(model = c("S2", "S1")) |>
-#   _[, .(forecast_time = get_forecast_times(model)), by = model] |>
-#   _[mday(forecast_time) == 1] |>
-#   _[,
-#     concentration := hindcast(
-#       forecast_time,
-#       model,
-#       members = "em",
-#       variable = "aice"
-#     ) |>
-#       zenodo(),
-#     by = .(model)
-#   ] |>
-#   _[mday(forecast_time) == 1] |>
-#   _[,
-#     thickness := hindcast(
-#       forecast_time,
-#       model,
-#       members = "em",
-#       variable = "hi"
-#     ) |>
-#       zenodo(),
-#     by = .(model)
-#   ] |>
-#   na.omit() |>
-#   _[,
-#     thickness := unlist(mirai_map(
-#       data.frame(concentration, thickness),
-#       mean_thickness
-#     )[.progress])
-#   ] |>
-#   na.omit()
-
 ```
@@
 errors <- here::here("data/derived/rmse.Rds") |>
-  zenodo() |>
+  zenodo("Hindcast RMSE of sea-ice concentration anomalies") |>
   readRDS() |>
@@
 topo_lonlat <- here::here("data/raw/ETOPO.nc") |>
-  zenodo() |>
+  zenodo("ETOPO topography") |>
   ReadNetCDF() |>
@@
 
-```{r}
+```{r clim_std_lon} 
 clim_std_lon <- datasets |>
@@
 rmse_lon_mean <- here::here("data/derived/rmse_lon.Rds") |>
-  zenodo() |>
+  zenodo("Hindcast RMSE in latitude slices") |>
   readRDS() |>

### 2025-10-23 R/datasets.R

@@ -12,7 +12,7 @@ cdo_apply <- function(ifile, op, ofile = NULL) {
 
 CDR_grid <- function() {
   here::here("data/raw/nsidc_grid.txt") |>
-    zenodo()
+    zenodo("NSDIC CDR grid description")
 }
 
 CDR <- function() {
@@ -416,7 +416,12 @@ hindcast_0lag_clim <- function(models = c("S1", "S2")) {
         rcdo::cdo_execute(output = file, options = "-O")
     }
 
-    return(list(file = zenodo(file)))
+    return(list(
+      file = zenodo(
+        file,
+        "{model} hindcast monthly sea-ice concentration climatology for the first month of forecast."
+      )
+    ))
   }) |>
     data.table::rbindlist(idcol = "model")
 }
@@ -495,6 +500,14 @@ hindcast_mean_thickness <- function() {
       na.omit()
   }
 
-  zenodo(thick$mean_thickness)
+  thick[,
+    mean_thickness := zenodo(
+      mean_thickness,
+      glue::glue_data(
+        .SD,
+        "{model} hindcast monthly spatial average sea-ice thickness for {forecast_time}"
+      )
+    )
+  ]
   thick
 }

### 2025-10-23 R/zenodo.R

@@ -132,28 +132,31 @@ zenodo_checksum_matches <- function(files) {
   ))
 }
 
-zenodo_upload_data <- function() {
-  files <- readLines(zenodo_files()) |>
-    Filter(f = nzchar)
 
-  if (length(files) == 0) {
+relative_path <- function(path) {
+  root <- here::here("")
+  gsub(root, "", x = path)
+}
+
+zenodo_upload_data <- function() {
+  files <- read.csv(zenodo_files())
+
+  files <- files[!duplicated(files$file)]
+
+  if (nrow(files) == 0) {
     stop("No files in zenodo_files.txt")
   }
 
-  checksum <- zenodo_checksum_matches(files)
+  checksum <- zenodo_checksum_matches(files$files)
 
   if (checksum$matches) {
     message("Remote and local checksum match. Nothing to upload.")
     return(invisible(NULL))
   }
 
-  files <- c(files, zenodo_files())
-  # Use relative paths inside zip
-  root <- here::here("")
-  files <- gsub(root, "", x = files)
-
   zipfile <- file.path(tempdir(), "data.zip")
-  zip(zipfile, files)
+  zip(zipfile, files$file)
+
   zenodo_upload_file(zipfile)
 
   checksum_file <- here::here("data/checksum")
@@ -223,11 +226,11 @@ zenodo_download_data <- function() {
   file_list <- zenodo_files()
 
   if (file.exists(file_list)) {
-    needed_files <- readLines(file_list) |>
+    needed_files <- read.csv(file_list)$file |>
       Filter(f = nzchar)
 
     if (all(file.exists(needed_files))) {
-      checksum <- zenodo_checksum_matches(files)
+      checksum <- zenodo_checksum_matches(needed_files)
       if (checksum$matches) return(invisible(NULL))
     }
   }
@@ -242,21 +245,35 @@ zenodo_download_data <- function() {
 on_gadi <- function() Sys.getenv("ON_GADI", unset = "null") != "null"
 
 
-zenodo <- function(files) {
-  zenodo_one <- function(file) {
-    # On gadi, we mark the file as used.
-    if (on_gadi()) {
-      if (file.exists(file)) {
-        write(file, file = zenodo_files(), append = TRUE)
-      }
-    }
-    return(file)
+zenodo <- function(files, description = rep("", length(files))) {
+  if (on_gadi()) {
+    env <- parent.frame()
+    description <- vapply(
+      description,
+      \(x) glue::glue(x, .envir = env),
+      character(1)
[…]

### 2025-10-23 R/zenodo.R
Fixes upload data

@@ -141,7 +141,7 @@ relative_path <- function(path) {
 zenodo_upload_data <- function() {
   files <- read.csv(zenodo_files())
 
-  files <- files[!duplicated(files$file)]
+  files <- files[!duplicated(files$file), ]
 
   if (nrow(files) == 0) {
     stop("No files in zenodo_files.txt")
@@ -154,11 +154,14 @@ zenodo_upload_data <- function() {
     return(invisible(NULL))
   }
 
+  message("Creating zip file")
   zipfile <- file.path(tempdir(), "data.zip")
   zip(zipfile, files$file)
 
+  message("Uploading zip file")
   zenodo_upload_file(zipfile)
 
+  message("Uploading checksum file")
   checksum_file <- here::here("data/checksum")
   writeLines(checksum$local, checksum_file)
   zenodo_upload_file(checksum_file)

### 2025-10-24 R/zenodo.R
Manages download and upload better

@@ -1,51 +1,29 @@
-# Minimal Zenodo helpers for a single fixed deposition
-# Uses httr2 for HTTP requests. Keep this file small and focused.
-
-# Hard-coded deposition ID for this project. Replace with the real ID.
-ZENODO_DEPOSITION_ID <- 386004
+# Hard-coded deposition ID for this project.
+ZENODO_DEPOSITION_ID <- 391075
 
 # Internal: read the token from environment and validate
 .zenodo_get_token <- function() {
   token <- Sys.getenv("ZENODO_TOKEN", unset = "")
   if (identical(token, "") || is.na(token)) {
-    stop(
-      "ZENODO_TOKEN environment variable is not set or empty. Set it to your Zenodo API token."
-    )
+    token <- NULL
   }
   token
 }
 
 # Internal: choose base API URL. If ZENODO_SANDBOX is set (non-empty), use the sandbox API.
-.zenodo_base_url <- function() {
+.zenodo_base_url <- function(api = TRUE) {
   if (nzchar(Sys.getenv("ZENODO_SANDBOX", ""))) {
-    return("https://sandbox.zenodo.org/api")
-  }
-  "https://zenodo.org/api"
-}
-
-# List files attached to the fixed deposition
-# Returns a list (as returned by the API) or throws on error
-zenodo_list_files <- function(deposition_id = ZENODO_DEPOSITION_ID) {
-  token <- .zenodo_get_token()
-  url <- paste0(.zenodo_base_url(), "/deposit/depositions/", deposition_id)
-
-  resp <- httr2::request(url) |>
-    httr2::req_auth_bearer_token(token) |>
-    httr2::req_perform()
-
-  if (httr2::resp_status(resp) >= 400) {
-    stop(sprintf(
-      "Zenodo API returned HTTP %s: %s",
-      httr2::resp_status(resp),
-      httr2::resp_body_json(resp, simplifyVector = TRUE)$message
-    ))
+    url <- "https://sandbox.zenodo.org/"
+  } else {
+    url <- "https://zenodo.org"
   }
 
-  body <- httr2::resp_body_json(resp, simplifyVector = FALSE)
-  # body$files is typically a list with elements containing 'filename', 'id', 'links', etc.
-  body$files
-}
+  if (api) {
+    url <- paste0(url, "api")
+  }
 
+  return(url)
+}
 
 get_bucket_for_deposition <- function(deposition_id = ZENODO_DEPOSITION_ID) {
   token <- .zenodo_get_token()
[…]

### 2025-10-24 R/functions.R
Default climatology works now

@@ -111,14 +111,14 @@ climatology <- function(dataset) {
 }
 
 
-anomalies <- function(dataset, climatology = climatology(dataset)) {
+anomalies <- function(dataset, clim = climatology(dataset)) {
   file <- here::here("data/derived/anomalies", name_from_dataset(dataset))
   dir.create(dirname(file), FALSE, TRUE)
 
   if (file.exists(file)) {
     return(file)
   }
-  cdo_ydaysub(cdo_del29feb(dataset), cdo_del29feb(climatology)) |>
+  cdo_ydaysub(cdo_del29feb(dataset), cdo_del29feb(clim)) |>
     cdo_execute(output = file, options = "-L")
 }
 

### 2025-10-24 R/datasets.R
Removes nonexistent hidncast time in thickness

@@ -452,6 +452,11 @@ hindcast_mean_thickness <- function() {
   thick <- data.table::CJ(model = c("S2", "S1")) |>
     _[, .(forecast_time = get_forecast_times(model)), by = model] |>
     _[data.table::mday(forecast_time) == 1] |>
+    _[
+      model == "S2" & forecast_time == as.IDate("1981-01-01"),
+      forecast_time := NA
+    ] |> # first time doesn't have mnembers
+    na.omit() |>
     _[,
       mean_thickness := here::here(glue::glue_data(
         .SD,
@@ -509,5 +514,5 @@ hindcast_mean_thickness <- function() {
       )
     )
   ]
-  thick
+  thick[]
 }

### 2025-10-24 draft/lag_correlation.R
Removes lag correlation file

@@ -1,34 +1,0 @@
-
-lags <- c(5, 10, 15)
-lags <- setNames(lags, lags)
-
-lag <- 30
-file <- anomalies(datasets$cdr)
-autocor <- function(file, lag) {
-  remove <- "timestep=1/{lag}/1" |> 
-    glue::glue()
-  
-  shift <- "-{lag}day" |> 
-    glue::glue()
-  
-  out <- file |> 
-    cdo_delete(timestep = remove) |>
-    cdo_timcor(file) |> 
-    cdo_execute(tempfile())
-  
-}
-
-
-corlags <- lapply(lags, \(lag) autocor(anomalies(datasets$cdr), lag))
-
-corlags |> 
-  setNames(lags) |> 
-  lapply(ReadNetCDF, c("aice", "pvalue")) |> 
-  rbindlist(idcol = "lag") |> 
-  _[aice <= 1] |> 
-  ggplot(aes(xgrid, ygrid)) +
-  geom_contour_fill(aes(z = aice)) +
-  # geom_point(data = \(x) x[pvalue < 0.5]) +
-  scale_fill_divergent() +
-  coord_equal() +
-  facet_wrap(~ lag)

### 2025-10-24 analysis/paper/access-ice.qmd
Init needds to come after download.
Adds software citations

@@
   message: false
-  cache: refresh
+  cache: true
 
@@
 source(here::here("R/zenodo.R"))
+zenodo_download_data()
 zenodo_init()
-zenodo_download_data()
 ```
@@
 A perfect forecast, would have zero RMSE and thus a skill score of 1.
+
+## Computational procedures
+
+We performed all analysis in this paper using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020b] packages.
+Significant processing was performed using the CDO command line operators [@schulzweida2023].
+All graphics are made using ggplot2 [@wickham2009].
+The paper was rendered using knitr and quarto [@xie2015; @allaire2022].
 

### 2025-10-24 analysis/paper/access-ice.tex
Adds software citations

@@
 
+\subsection{Computational procedures}\label{computational-procedures}
+
+We performed all analysis in this paper using the R programming language
+\citep{rcoreteam2020}, using data.table \citep{dowle2020} and metR
+\citep{campitelli2020b} packages. Significant processing was performed
+using the CDO command line operators \citep{schulzweida2023}. All
+graphics are made using ggplot2 \citep{wickham2009}. The paper was
+rendered using knitr and quarto \citep{xie2015, allaire2022}.
+
 \section{Results and discussion}\label{results-and-discussion}

### 2025-10-30 analysis/paper/access-ice.qmd
Adds Ariaan's comments

@@
 ---
-title: Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Model
+title: Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Forecast Model
 format:
@@
     equal-margins: false
-    classoption: referee
+    classoption: referee,lineno
     fig-width: 5  
@@
       - ref: 3      
-keywords: [sea ice, predictability, initial conditions, forecasting]
+keywords: [sea ice, seasonal predictability, initial conditions, forecasting]
 bibliography: references.bib
@@
 
-Accurate Antarctic sea ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
-The two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system, +S1 and +S2, use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations.
-This makes for a good "natural experiment" to analyse the role of initial conditions on Antarctic sea ice forecasts using a long run and a moderately large ensemble with a fully coupled model.
-Our analysis reveals that both systems experience an extended melt season and delayed growth phase. 
-This leads to a significant negative sea-ice extent bias which is corrected only in +S1 by the data assimilation system. 
-The impact of the initial conditions on forecast errors varies dramatically by season: summer and autumn initial conditions (January-April) provide predictive skill for up to three months, with February initial conditions being particularly crucial.
+Accurate Antarctic sea-ice forecasts are crucial for climate monitoring and operational planning, yet they remain challenging due to model biases and complex ice-ocean-atmosphere interactions. 
+The two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system, +S1 and +S2, use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations, whereas +S1 does.
+This provides an opportunistic experiment to assess the role of initial conditions on Antarctic sea-ice forecasts using more than 20 years of fully coupled simulations with two 9-member ensembles.
+Our analysis reveals that both systems experience an extended melt season and delayed growth phase compared with observations. 
+This leads to a significant negative sea-ice extent bias, which is corrected only in +S1 by the data assimilation system. 
+The impact of the differing initial conditions on forecast errors varies dramatically by season: summer and autumn initial conditions (January-April) provide predictive skill for up to three months, with February initial conditions being particularly crucial.
 In contrast, winter forecasts of the two systems are statistically indistinguishable after only two weeks. 
-Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over Eastern Antarctica, where even +S1 shows negative skill. 
+Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over East Antarctica, where even +S1 shows negative skill. 
 These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritize the summer-autumn period when initial conditions have maximum impact on forecast skill.
@@
 
-However, Antarctic sea ice forecasts have lagged behind Arctic sea ice forecasts due to model biases, inherent large variability and complexity [@zampieri2019; @gao2024a].
+However, progress in Antarctic sea-ice forecasting system has lagged behind Arctic sea-ice forecasts due to model biases, inherent large variability and complexity [@zampieri2019; @gao2024a].
 Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simpler statistical methods [@massonnet2023] and machine learning approaches (e.g.
@@
 
-Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea ice initial conditions affect the quality of the forecast and at what timescales.
-Exploring seasonal predictions of Arctic sea ice, @guemas2016 found that sea ice initial conditions are important in autumn to predict summer sea ice, but the impact wasn’t as dramatic when predicting winter sea ice.
-@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea ice forecasts initialised in July but not when initialised in January.
+Good initial conditions are generally required for a good forecast, however, it is not entirely known to what extent accurate sea-ice initial conditions affect the quality of the forecast and at what timescales.
+Exploring seasonal predictions of Arctic sea ice, @guemas2016 found that sea-ice initial conditions are important in autumn to predict summer sea ice, but the impact wasn’t as dramatic when predicting winter sea ice.
+@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea-ice forecasts initialised in July but not when initialised in January.
 
@@
 @marchi2020 also used a coupled ocean–sea-ice model instead of a fully coupled model like @holland2013 did.
-@morioka2022 used observed sea ice initial conditions and compared with observations, while @marchi2020 and @holland2013 are perfect model studies.
+@morioka2022 used observed sea-ice initial conditions and compared with observations, while @marchi2020 and @holland2013 are perfect model studies.
 
@@
 
-Since model configuration is identical between +S1 and +S2, they form a sort of “natural experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year, with the only difference being the initial conditions.
-This provides an opportunity to test the effect of sea ice initial conditions on the forecast of sea-ice concentrations and the climate.
-
-In this study we compare sea ice hindcasts produced by +S1 and +S2.
+Since model configuration is identical between +S1 and +S2, they form a sort of "opportunistic experiment” where the same forecasting model was run over a long period of time with multiple ensemble forecasts initialised throughout the year, with the only difference being the initial conditions.
+This provides an opportunity to test the effect of sea-ice initial conditions on the forecast of sea-ice concentrations and the climate.
+
+In this study we compare sea-ice hindcasts produced by +S1 and +S2.
 We focus on seasonality of errors and biases and the effect of the data assimilation system.
-This comparison will also inform future work with the prediction system as a research tool to better understand the dynamics and variability of the Antarctic sea ice and its impacts on the climate system as well as to explore the potential of using its sea ice forecasts for decision-making.
+This comparison will inform future work with the prediction system as a research tool to better understand the dynamics and variability of the Antarctic sea ice and its impacts on the climate system as well as to explore the potential of using its sea-ice forecasts for decision-making.
 The work will also serve as a benchmark for future prediction systems to attempt to improve upon.
@@
 Both systems take atmospheric initial conditions derived from ERA-interim [@dee2011] for their hindcasts.
-The main difference between the hindcasts of the two systems are the ocean and sea ice initial conditions.
-+S1's ocean and sea ice initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and satellite observations of sea-ice concentration using the EUMETSAT OSISAF product described in the next section.
+The main difference between the hindcasts of the two systems are the ocean and sea-ice initial conditions.
++S1's ocean and sea-ice initial conditions come from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) data assimilation scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and satellite observations of sea-ice concentration using the EUMETSAT OSISAF product described in the next section.
 +S2, on the other hand, is initialised from ocean conditions generated by the BoM weakly coupled ensemble data assimilation scheme described in @wedd2022.
@@
 The atmosphere fields from that daily integration are not used and instead the model atmosphere is initialised using ERA-Interim.
-The sea ice initial conditions for the next cycle are the unaltered output of the previous daily integration.
+The sea-ice initial conditions for the next cycle are the unaltered output of the previous daily integration.
 Then the cycle starts again and the coupled model runs for another 24 hours.
@@
 The +S2 hindcast set used in this study runs for the period 1981–2018.
-Each hindcast consists of nine ensemble members built from three consecutive 3-member forecasts created in the same manner as +S1 members initialised at the first of every month and the two previous days and run for 279 days.
+Each hindcast consists of nine ensemble members built from three consecutive three-member forecasts created in the same manner as +S1 members initialised at the first of every month and the two previous days and run for 279 days.
 We analyse the ensemble mean hindcasts unless otherwise specified.
@@
 )
+labs_dataset <- c(cdr = "CDR", osi = "OSI")
 which_dataset <- "cdr"
@@
 Both source algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
-The data has a spatial resolution of 25 by 25 km and daily from 1978 onwards.
-
-The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility [@OSISAF] based on the SSMIS sensor is another satellite-derived sea-ice concentration product.
+The data has a spatial resolution of 25 by 25 km and daily from November 1978 onwards.
+
+The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility [OSI; @OSISAF] based on the SSMIS sensor is another satellite-derived sea-ice concentration product.
 It is based on mostly the same sensors as the NOAA CDR but computed independently using different algorithms. 
@@
 For evaluation purposes, we use a series of measures. 
-
 Sea-ice extent is defined as the area of the ocean at least 15% covered by sea-ice. 
@@
 ```{r fig-bias-1}
-#| fig-cap: Ensemble mean difference between monthly sea-ice concentration of +S2 ensemble mean forecast at 0-month lead time (monthly mean values forecasted from the forecast initialised at the first of the month) and observations.
+#| fig-cap: !expr glue::glue("Ensemble mean difference between monthly sea-ice concentration of +S2 ensemble mean forecast at 0-month lead time (monthly mean values forecasted from the forecast initialised at the first of the month) and observations ({labs_dataset[which_dataset]})").
 #| fig-height: 6
@@
 ```{r fig-bias-2}
-#| fig-cap: Same as @fig-bias-1 but for +S1
+#| fig-cap: Same as @fig-bias-1 but for +S1.
 #| fig-height: 6
@@
 +S1 has a comparatively smaller overall bias (@fig-bias-2). 
-The largest values are found between April and June, when the faster growth results in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen seas.
+The largest values are found between April and June, when the faster growth results in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen Seas.
 
@@
 #| fig-cap: Monthly mean sea-ice extent anomalies of the observations (black) and forecasts from +S1 (right column; purple) and +S2 (left column; green) at lead times of 0, 2, 4, and 6 months.
-The RMSE and correlation during the overlapping period of +S1 and +S2 hindcasts (1990–2013) is shown on the top left and bottom left of each panel respectively.
+The RMSE and correlation during the overlapping period of +S1 and +S2 hindcasts (1990–2013) are shown on the top left and bottom left of each panel respectively.
 #| fig-height: 6
@@
 @fig-extent-sd shows the interannual standard deviation of monthly sea-ice extent of the forecasts as a function of lead time compared with observations.
-+S1 standard deviation lies within the observed standard deviation regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at 9 month lead time for most months.
++S1 standard deviation lies within the observed standard deviation regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at nine month lead time for most months.
 
@@
 
-  scale_y_continuous("Standardised sea ice extent standard deviation") +
+  scale_y_continuous("Standardised sea-ice extent standard deviation") +
   scale_x_continuous(
@@
 +S2 sea-ice concentration anomalies, which are not assimilated, are not as close, but the large-scale pattern is aligned with observations.
-The system simulates large positive anomalies in the Weddell and Ross seas and slight negative anomalies in the Amundsen and Bellingshausen seas.
-The fact that +S2 can simulate this pattern without assimilating sea-ice data suggests that atmospheric and oceanic forcing was the dominant factor.
+The system simulates large positive anomalies in the Weddell and Ross Seas and slight negative anomalies in the Amundsen and Bellingshausen Seas.
+The fact that +S2 can simulate this pattern without assimilating sea-ice data suggests that atmospheric and oceanic forcing were the dominant drivers.
 However, the magnitude of the sea-ice anomalies is too big. It is plausible that this is due to the thinner ice simulated by +S2 (bottom row).
@@
 The fact that +S1 and +S2 share the same model configuration and that the increased variability is more extreme at short lead times ([Fig.
-@fig-extent-sd]) suggests that the data assimilation procedure is partly to blame.
+@fig-extent-sd]) suggests that the data assimilation procedure is partly responsible.
 It is possible that sea-ice in the +S2 system is left in an unbalanced state after assimilating atmospheric and oceanic data but not sea-ice data, leading to large responses that are amplified by the thin ice in the initial states which then subside at longer lead times when the model is balanced.
@@
 #| label: fig-rmse_lon-2
-#| fig-cap: Same as @fig-rmse_lon-1 but for +S2
+#| fig-cap: Same as @fig-rmse_lon-1 but for +S2.
 #| fig-height: 7
@@
 For +S1 forecasts (@fig-rmse_lon-1), skill tends to be lower off the coast of Eastern Antarctica even at short lead times; for instance, the skill score for forecasts initialised in May and June are negative between 0° and 120°E even at almost zero lead time.
-This mirrors @libera2022 findings of a “winter predictability barrier”, although they focus on the Weddell sea and here we show that the effect seems to be stronger more to the East. 
-In western Antarctica there is a hint of easterly-propagating skill in forecasts initialised in February and March. 
+This mirrors @libera2022 findings of a “winter predictability barrier”, although they focus on the Weddell Sea and here we show that the effect seems to be stronger more to the east. 
+In West Antarctica there is a hint of easterly-propagating skill in forecasts initialised in February and March. 
 This is consistent with @holland2013 findings that memory of sea-ice anomalies are stored in ocean heat content anomalies that are transported east by the Antarctic Circumpolar Current.
 
-+S2 forecasts  (@fig-rmse_lon-2) also have lower skill over Eastern Antarctica. 
-From July to December even though the pan-Antarctic average skill is negative at all lead times ([Fig. @fig-lead-time-window]), it is positive for up to a month in Western Antarctica. 
++S2 forecasts  (@fig-rmse_lon-2) also have lower skill over East Antarctica. 
+From July to December even though the pan-Antarctic average skill is negative at all lead times ([Fig. @fig-lead-time-window]), it is positive for up to a month in West Antarctica. 
 Since oceanic and atmospheric forcing is the only source of information, this suggests that sea-ice in this region is particularly sensitive to oceanic and atmospheric forcing and suggests a role of the Pacific-South American mode and the Amundsen Sea Low to shape sea-ice concentration anomalies.
-The fact that this is evident in the months in which ENSO teleconnections are more important for atmospheric circulation also suggest the influence of tropical Pacific variability.
+The fact that this is evident in the months in which El Niño--Southern Oscillation teleconnections are more important for atmospheric circulation also suggest the influence of tropical Pacific variability.
 February and March are the only two months that can be forecasted with marginally positive skill in large regions.
@@
 
-Sea ice forecasts from the +S2 system show a significant low extent bias, particularly during late summer and early autumn. 
+Sea-ice forecasts from the +S2 system show a significant low extent bias, particularly during late summer and early autumn. 
 This bias is attributed to a faster and longer melt season between January and March, and slower growth between March and April. 
 This underestimation during the minimum and early freezing season is a common issue in many seasonal-to-subseasonal (S2S) systems, suggesting potential problems either with the model’s thermodynamic representation or with short wave radiation forcing, as shown in other climate models [@zampieri2019; @roach2020].
-Even though +S2 shares the same model formulation with +S1, the latter does not suffer from this bias, indicating that assimilating sea-ice concentrations successfully corrects for the negative bias that exists in the free-running model.
-
-Ensemble spread grows quickly even when perturbations are only implemented in the atmosphere component (in +S1), indicating that sea-ice is indeed responding quickly to atmospheric perturbations. 
-However, our analysis suggests that the atmosphere and ocean data assimilation implemented in +S2 is only effectively influencing sea ice initial conditions from June to October, while the rest of the year, the sea-ice component runs virtually free, reverting to its biased equilibrium state.
+Even though +S2 shares the same model components as +S1, the latter does not suffer from this bias, indicating that assimilating sea-ice concentrations successfully corrects for the negative bias that exists in the free-running model.
+
+Ensemble spread grows quickly even when perturbations are only implemented in the atmosphere component (in +S1), indicating that sea ice is indeed responding quickly to atmospheric perturbations. 
+However, our analysis suggests that the atmosphere and ocean data assimilation implemented in +S2 is only effectively influencing sea-ice initial conditions from June to October, while the rest of the year, the sea-ice component runs virtually free, reverting to its biased equilibrium state.
 
@@
 
-Although +S1 only assimilates sea-ice concentration, it is clear that sea-ice thickness is also affected. 
+Although +S1 only assimilates sea-ice concentration, it is clear that sea-ice thickness is also affected through the assimilation process. 
 +S1 simulates significantly thicker ice than +S2 and in both systems sea-ice is thicker at shorter lead times than at longer lead times. 
 Both the explicit data assimilation in +S2 and the effects of atmospheric and oceanic data assimilation in +S1 might be nudging simulated sea ice to be thicker than the model equilimbrium state.
-We believe that the thinner sea ice in +S2 is a partial cause for the large sea-ice extent variance, but other mechanisms, such as unbalanced initial conditions might be important.
-
-Given that +S2 sea-ice extent is essentially not initialised by the observations, comparing its forecasts with those of +S1 allows us to estimate the time-scale for which initial conditions are important.
-We find that initial conditions affect Antarctic sea ice forecasts in the order of a few months, but that effect is seasonally dependent. 
-January to April initial conditions improve forecasts for up to 3 months with, February initial conditions in particular are shown to be crucial for determining sea-ice evolution at least up to May. 
-Arctic sea ice forecasts also show greater sensitivity to initial conditions in boreal summer, compared with boreal winter [@day2014; @bunzel2016], suggesting a similar mechanism might be playing a role.
-
-Forecasts initialised in winter have very little skill and +S1 and +S2 forecast errors are not statistically different after just 2 weeks. 
-This is consistent with @libera2022's finding of a “winter predictability barrier” in the Weddell sea, although they describe the barrier as sharp loss of predictability in July, and here we find a gradual reduction in skill compared with climatology around June.
+We suggest that the thinner sea ice in +S2 contributes to the large sea-ice extent variance, but other mechanisms, such as unbalanced initial conditions might also be important.
+
+Given that +S2 sea-ice extent is not directly initialised by sea-ice observations, comparing its forecasts with those of +S1 allows us to estimate the time-scale over which initial conditions are important.
+We find that initial conditions affect Antarctic sea-ice forecasts in the order of a few months, but that effect is seasonally dependent. 
+January to April initial conditions improve forecasts for up to three months with, February initial conditions in particular are shown to be crucial for determining sea-ice evolution at least up to May.
+Arctic sea-ice forecasts also show greater sensitivity to initial conditions in boreal summer, compared with boreal winter [@day2014; @bunzel2016], suggesting a similar mechanism might be playing a role.
+
+Forecasts initialised in winter have very little skill and +S1 and +S2 forecast errors are not statistically different after just two weeks. 
+This is consistent with the findings of @libera2022's finding of a “winter predictability barrier” in the Weddell Sea, although they describe the barrier as sharp loss of predictability in July, and here we find a gradual reduction in skill compared with climatology around June.
 This difference might be due to our use of pan-Antarctic RMSE, since our regional analysis indicates that the degraded skill is most dramatic in the King Haakon sea.

### 2025-10-30 analysis/paper/access-ice.qmd
Adds Matt's comments

@@
     affiliations:
-      - department: Research, Bureau of Meteorology
+      - department: Research, Bureau of Meteorology, Australia
         id: 3
-  - name: Matthew Wheeler
+  - name: Matthew C. Wheeler
     affiliations:
@@
 The two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system, +S1 and +S2, use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations, whereas +S1 does.
-This provides an opportunistic experiment to assess the role of initial conditions on Antarctic sea-ice forecasts using more than 20 years of fully coupled simulations with two 9-member ensembles.
+This provides an convenient opportunistic experiment to assess the role of initial conditions on Antarctic sea-ice forecasts using more than 20 years of fully coupled simulations with two 9-member ensembles.
 Our analysis reveals that both systems experience an extended melt season and delayed growth phase compared with observations. 
@@
 
-However, progress in Antarctic sea-ice forecasting system has lagged behind Arctic sea-ice forecasts due to model biases, inherent large variability and complexity [@zampieri2019; @gao2024a].
+However, progress in Antarctic sea-ice forecasting system has lagged behind Arctic sea-ice forecasts due to model biases, and inherent large variability and complexity [@zampieri2019; @gao2024a].
 Dynamical seasonal forecasts of summer Antarctic sea ice have been shown to perform worse than relatively simpler statistical methods [@massonnet2023] and machine learning approaches (e.g.
-@dong2024, @lin2025), which also underscores the need for better understanding of sea-ice dynamics, and drivers of its variability.
+@dong2024, @lin2025), which also underscores the need for better understanding and physical modelling of sea-ice dynamics, and drivers of its variability.
 
@@
 Exploring seasonal predictions of Arctic sea ice, @guemas2016 found that sea-ice initial conditions are important in autumn to predict summer sea ice, but the impact wasn’t as dramatic when predicting winter sea ice.
-@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness lead to improved sea-ice forecasts initialised in July but not when initialised in January.
+@day2014 also found seasonally-varying differences in the effect of initialisation, noting that accurate Arctic sea-ice thickness leads to improved sea-ice forecasts initialised in July but not when initialised in January.
 
@@
 More recently, @morioka2022 studied decadal forecasts of Antarctic sea ice and found that initialising ocean and sea ice improved the correlation between simulated and observed sea-ice concentration evolution in the Amundsen–Bellingshausen Sea.
-It is hard to compare these studies since they are based on forecasts initialised at different times of the year and different frameworks: @holland2013 ran 20 ensemble members initialised on the 1st of January of a particular year, @marchi2020 ran forecasts from the 1st of March and 1st of September and @morioka2022 run forecasts only from the 1st of March.
+It is hard to compare these studies since they are based on forecasts initialised at different times of the year and different frameworks: @holland2013 ran 20 ensemble members initialised on the 1st of January of a particular year, @marchi2020 ran forecasts from the 1st of March and 1st of September, and @morioka2022 ran forecasts only from the 1st of March.
 @marchi2020 also used a coupled ocean–sea-ice model instead of a fully coupled model like @holland2013 did.
-@morioka2022 used observed sea-ice initial conditions and compared with observations, while @marchi2020 and @holland2013 are perfect model studies.
+@morioka2022 used observed sea-ice initial conditions and compared with observations, while @marchi2020 and @holland2013 were perfect model studies.
 

### 2025-10-30 analysis/paper/access-ice.qmd
Fixes caption

@@
 ```{r fig-bias-1}
-#| fig-cap: !expr glue::glue("Ensemble mean difference between monthly sea-ice concentration of +S2 ensemble mean forecast at 0-month lead time (monthly mean values forecasted from the forecast initialised at the first of the month) and observations ({labs_dataset[which_dataset]})").
+#| fig-cap: !expr glue::glue("Ensemble mean difference between monthly sea-ice concentration of +S2 ensemble mean forecast at 0-month lead time (monthly mean values forecasted from the forecast initialised at the first of the month) and observations ({labs_dataset[which_dataset]}).")
 #| fig-height: 6

### 2025-10-30 analysis/paper/access-ice.tex
Renders

@@
   sn-mathphys-num,
-  referee]{sn-jnl}
+  referee,lineno]{sn-jnl}
 
@@
 \hypersetup{
-  pdftitle={Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Model},
-  pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Matthew Wheeler; Phillip Reid},
-  pdfkeywords={sea ice, predictability, initial
+  pdftitle={Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Forecast Model},
+  pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Matthew C. Wheeler; Phillip Reid},
+  pdfkeywords={sea ice, seasonal predictability, initial
 conditions, forecasting},
@@
 \title[Evaluating the Importance of Initial Conditions for Antarctic Sea
-Ice Seasonal Predictability with a Fully Coupled Model]{Evaluating the
-Importance of Initial Conditions for Antarctic Sea Ice Seasonal
-Predictability with a Fully Coupled Model}
+Ice Seasonal Predictability with a Fully Coupled Forecast
+Model]{Evaluating the Importance of Initial Conditions for Antarctic Sea
+Ice Seasonal Predictability with a Fully Coupled Forecast Model}
 
 % author setup
-\author*[1,2]{\fnm{Elio} \sur{Campitelli}}\email{elio.campitelli@monash.edu}\author[1,2]{\fnm{Ariaan} \sur{Purich}}\author[1,2]{\fnm{Julie} \sur{Arblaster}}\author[3]{\fnm{Eun-Pa} \sur{Lim}}\author[3]{\fnm{Matthew} \sur{Wheeler}}\author[3]{\fnm{Phillip} \sur{Reid}}
+\author*[1,2]{\fnm{Elio} \sur{Campitelli}}\email{elio.campitelli@monash.edu}\author[1,2]{\fnm{Ariaan} \sur{Purich}}\author[1,2]{\fnm{Julie} \sur{Arblaster}}\author[3]{\fnm{Eun-Pa} \sur{Lim}}\author[3]{\fnm{Matthew C.} \sur{Wheeler}}\author[3]{\fnm{Phillip} \sur{Reid}}
 % affil setup
@@
 Future}, \orgname{Monash University, Australia}}
-\affil[3]{\orgdiv{Research, Bureau of Meteorology}}
+\affil[3]{\orgdiv{Research, Bureau of Meteorology, Australia}}
 
@@
 
-\abstract{Accurate Antarctic sea ice forecasts are crucial for climate
+\abstract{Accurate Antarctic sea-ice forecasts are crucial for climate
 monitoring and operational planning, yet they remain challenging due to
@@
 configuration and differ only in their initial conditions; primarily in
-that ACCESS-S2 does not assimilate sea-ice observations. This makes for
-a good ``natural experiment'' to analyse the role of initial conditions
-on Antarctic sea ice forecasts using a long run and a moderately large
-ensemble with a fully coupled model. Our analysis reveals that both
-systems experience an extended melt season and delayed growth phase.
-This leads to a significant negative sea-ice extent bias which is
-corrected only in ACCESS-S1 by the data assimilation system. The impact
-of the initial conditions on forecast errors varies dramatically by
-season: summer and autumn initial conditions (January-April) provide
-predictive skill for up to three months, with February initial
-conditions being particularly crucial. In contrast, winter forecasts of
-the two systems are statistically indistinguishable after only two
-weeks. Regional analysis of forecast skill suggests that this winter
-predictability barrier is most dramatic over Eastern Antarctica, where
-even ACCESS-S1 shows negative skill. These findings highlight the
-critical importance of comprehensive year-round sampling in
-predictability studies and suggest that operational sea-ice data
-assimilation efforts should prioritize the summer-autumn period when
-initial conditions have maximum impact on forecast skill.}
+that ACCESS-S2 does not assimilate sea-ice observations, whereas
+ACCESS-S1 does.\\
+This provides an convenient opportunistic experiment to assess the role
+of initial conditions on Antarctic sea-ice forecasts using more than 20
+years of fully coupled simulations with two 9-member ensembles. Our
+analysis reveals that both systems experience an extended melt season
+and delayed growth phase compared with observations. This leads to a
+significant negative sea-ice extent bias, which is corrected only in
+ACCESS-S1 by the data assimilation system. The impact of the differing
+initial conditions on forecast errors varies dramatically by season:
+summer and autumn initial conditions (January-April) provide predictive
+skill for up to three months, with February initial conditions being
+particularly crucial. In contrast, winter forecasts of the two systems
+are statistically indistinguishable after only two weeks. Regional
+analysis of forecast skill suggests that this winter predictability
+barrier is most dramatic over East Antarctica, where even ACCESS-S1
+shows negative skill. These findings highlight the critical importance
+of comprehensive year-round sampling in predictability studies and
+suggest that operational sea-ice data assimilation efforts should
+prioritize the summer-autumn period when initial conditions have maximum
+impact on forecast skill.}
 
 % keywords
-\keywords{sea ice,  predictability,  initial conditions,  forecasting}
+\keywords{sea ice,  seasonal predictability,  initial
+conditions,  forecasting}
 
@@
 
-However, Antarctic sea ice forecasts have lagged behind Arctic sea ice
-forecasts due to model biases, inherent large variability and complexity
-\citep{zampieri2019, gao2024a}. Dynamical seasonal forecasts of summer
-Antarctic sea ice have been shown to perform worse than relatively
-simpler statistical methods \citep{massonnet2023} and machine learning
-approaches (e.g. \citet{dong2024}, \citet{lin2025}), which also
-underscores the need for better understanding of sea-ice dynamics, and
+However, progress in Antarctic sea-ice forecasting system has lagged
+behind Arctic sea-ice forecasts due to model biases, and inherent large
+variability and complexity \citep{zampieri2019, gao2024a}. Dynamical
+seasonal forecasts of summer Antarctic sea ice have been shown to
+perform worse than relatively simpler statistical methods
+\citep{massonnet2023} and machine learning approaches (e.g.
+\citet{dong2024}, \citet{lin2025}), which also underscores the need for
+better understanding and physical modelling of sea-ice dynamics, and
 drivers of its variability.
@@
 Good initial conditions are generally required for a good forecast,
-however, it is not entirely known to what extent accurate sea ice
+however, it is not entirely known to what extent accurate sea-ice
 initial conditions affect the quality of the forecast and at what
 timescales. Exploring seasonal predictions of Arctic sea ice,
-\citet{guemas2016} found that sea ice initial conditions are important
+\citet{guemas2016} found that sea-ice initial conditions are important
 in autumn to predict summer sea ice, but the impact wasn't as dramatic
@@
 seasonally-varying differences in the effect of initialisation, noting
-that accurate Arctic sea-ice thickness lead to improved sea ice
+that accurate Arctic sea-ice thickness leads to improved sea-ice
 forecasts initialised in July but not when initialised in January.
@@
 \citet{marchi2020} ran forecasts from the 1st of March and 1st of
-September and \citet{morioka2022} run forecasts only from the 1st of
+September, and \citet{morioka2022} ran forecasts only from the 1st of
 March. \citet{marchi2020} also used a coupled ocean--sea-ice model
 instead of a fully coupled model like \citet{holland2013} did.
-\citet{morioka2022} used observed sea ice initial conditions and
+\citet{morioka2022} used observed sea-ice initial conditions and
 compared with observations, while \citet{marchi2020} and
-\citet{holland2013} are perfect model studies.
+\citet{holland2013} were perfect model studies.
 
@@
 Since model configuration is identical between ACCESS-S1 and ACCESS-S2,
-they form a sort of ``natural experiment'' where the same forecasting
-model was run over a long period of time with multiple ensemble
-forecasts initialised throughout the year, with the only difference
-being the initial conditions. This provides an opportunity to test the
-effect of sea ice initial conditions on the forecast of sea-ice
+they form a sort of ``opportunistic experiment'' where the same
+forecasting model was run over a long period of time with multiple
+ensemble forecasts initialised throughout the year, with the only
+difference being the initial conditions. This provides an opportunity to
+test the effect of sea-ice initial conditions on the forecast of sea-ice
 concentrations and the climate.
 
-In this study we compare sea ice hindcasts produced by ACCESS-S1 and
+In this study we compare sea-ice hindcasts produced by ACCESS-S1 and
 ACCESS-S2. We focus on seasonality of errors and biases and the effect
-of the data assimilation system. This comparison will also inform future
-work with the prediction system as a research tool to better understand
-the dynamics and variability of the Antarctic sea ice and its impacts on
-the climate system as well as to explore the potential of using its sea
-ice forecasts for decision-making. The work will also serve as a
-benchmark for future prediction systems to attempt to improve upon.
+of the data assimilation system. This comparison will inform future work
+with the prediction system as a research tool to better understand the
+dynamics and variability of the Antarctic sea ice and its impacts on the
+climate system as well as to explore the potential of using its sea-ice
+forecasts for decision-making. The work will also serve as a benchmark
+for future prediction systems to attempt to improve upon.
 
@@
 ERA-interim \citep{dee2011} for their hindcasts. The main difference
-between the hindcasts of the two systems are the ocean and sea ice
-initial conditions. ACCESS-S1's ocean and sea ice initial conditions
+between the hindcasts of the two systems are the ocean and sea-ice
+initial conditions. ACCESS-S1's ocean and sea-ice initial conditions
 come from the Met Office FOAM system, which uses a multivariate,
@@
 atmosphere fields from that daily integration are not used and instead
-the model atmosphere is initialised using ERA-Interim. The sea ice
+the model atmosphere is initialised using ERA-Interim. The sea-ice
 initial conditions for the next cycle are the unaltered output of the
@@
 in this study runs for the period 1981--2018. Each hindcast consists of
-nine ensemble members built from three consecutive 3-member forecasts
-created in the same manner as ACCESS-S1 members initialised at the first
-of every month and the two previous days and run for 279 days. We
-analyse the ensemble mean hindcasts unless otherwise specified.
+nine ensemble members built from three consecutive three-member
+forecasts created in the same manner as ACCESS-S1 members initialised at
+the first of every month and the two previous days and run for 279 days.
+We analyse the ensemble mean hindcasts unless otherwise specified.
 
@@
 -F8, -F11, and -F13 satellites. The data has a spatial resolution of 25
-by 25 km and daily from 1978 onwards.
+by 25 km and daily from November 1978 onwards.
 
@@
 Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility
-\citep{OSISAF} based on the SSMIS sensor is another satellite-derived
-sea-ice concentration product. It is based on mostly the same sensors as
-the NOAA CDR but computed independently using different algorithms.
-Figures prepared with this dataset are provided in the supplementary
-material and do not differ significantly from the ones prepared using
-CDR.
+{[}OSI; \citet{OSISAF}{]} based on the SSMIS sensor is another
+satellite-derived sea-ice concentration product. It is based on mostly
+the same sensors as the NOAA CDR but computed independently using
+different algorithms. Figures prepared with this dataset are provided in
+the supplementary material and do not differ significantly from the ones
+prepared using CDR.
 
@@
 
-For evaluation purposes, we use a series of measures.
-
-Sea-ice extent is defined as the area of the ocean at least 15\% covered
-by sea-ice. This threshold is motivated by the limitations in satellite
-retrieval, which is increasingly unreliable for lower sea-ice
-concentrations \citep{cavalieri1991}.
+For evaluation purposes, we use a series of measures. Sea-ice extent is
+defined as the area of the ocean at least 15\% covered by sea-ice. This
+threshold is motivated by the limitations in satellite retrieval, which
+is increasingly unreliable for lower sea-ice concentrations
+\citep{cavalieri1991}.
 
@@
 lead time (monthly mean values forecasted from the forecast initialised
-at the first of the month) and observations.}
+at the first of the month) and observations (CDR).}
 
@@
 \caption{\label{fig-bias-2}Same as Figure~\ref{fig-bias-1} but for
-ACCESS-S1}
+ACCESS-S1.}
 
@@
 the sea-ice edge, and in January, when the faster melt leads to large
-negative bias in the Weddell and Admunsen seas.
+negative bias in the Weddell and Admunsen Seas.
 
@@
 6 months. The RMSE and correlation during the overlapping period of
-ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) is shown on the top left
+ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) are shown on the top left
 and bottom left of each panel respectively.}
@@
 standard deviation is more than twice that of observations at zero lead
-time and only approaches the observed value at 9 month lead time for
+time and only approaches the observed value at nine month lead time for
 most months.
@@
 large-scale pattern is aligned with observations. The system simulates
-large positive anomalies in the Weddell and Ross seas and slight
-negative anomalies in the Amundsen and Bellingshausen seas. The fact
+large positive anomalies in the Weddell and Ross Seas and slight
+negative anomalies in the Amundsen and Bellingshausen Seas. The fact
 that ACCESS-S2 can simulate this pattern without assimilating sea-ice
-data suggests that atmospheric and oceanic forcing was the dominant
-factor. However, the magnitude of the sea-ice anomalies is too big. It
+data suggests that atmospheric and oceanic forcing were the dominant
+drivers. However, the magnitude of the sea-ice anomalies is too big. It
 is plausible that this is due to the thinner ice simulated by ACCESS-S2
@@
 (Fig.~\ref{fig-extent-sd}) suggests that the data assimilation procedure
-is partly to blame. It is possible that sea-ice in the ACCESS-S2 system
-is left in an unbalanced state after assimilating atmospheric and
+is partly responsible. It is possible that sea-ice in the ACCESS-S2
+system is left in an unbalanced state after assimilating atmospheric and
 oceanic data but not sea-ice data, leading to large responses that are
@@
 \caption{\label{fig-rmse_lon-2}Same as Figure~\ref{fig-rmse_lon-1} but
-for ACCESS-S2}
+for ACCESS-S2.}
 
@@
 mirrors \citet{libera2022} findings of a ``winter predictability
-barrier'', although they focus on the Weddell sea and here we show that
-the effect seems to be stronger more to the East. In western Antarctica
+barrier'', although they focus on the Weddell Sea and here we show that
+the effect seems to be stronger more to the east. In West Antarctica
 there is a hint of easterly-propagating skill in forecasts initialised
@@
 ACCESS-S2 forecasts (Figure~\ref{fig-rmse_lon-2}) also have lower skill
-over Eastern Antarctica. From July to December even though the
+over East Antarctica. From July to December even though the
 pan-Antarctic average skill is negative at all lead times
 (Fig.~\ref{fig-lead-time-window}), it is positive for up to a month in
-Western Antarctica. Since oceanic and atmospheric forcing is the only
+West Antarctica. Since oceanic and atmospheric forcing is the only
 source of information, this suggests that sea-ice in this region is
@@
 shape sea-ice concentration anomalies. The fact that this is evident in
-the months in which ENSO teleconnections are more important for
-atmospheric circulation also suggest the influence of tropical Pacific
-variability. February and March are the only two months that can be
-forecasted with marginally positive skill in large regions.
+the months in which El Niño--Southern Oscillation teleconnections are
+more important for atmospheric circulation also suggest the influence of
+tropical Pacific variability. February and March are the only two months
+that can be forecasted with marginally positive skill in large regions.
 
@@
 
-Sea ice forecasts from the ACCESS-S2 system show a significant low
+Sea-ice forecasts from the ACCESS-S2 system show a significant low
 extent bias, particularly during late summer and early autumn. This bias
@@
 \citep{zampieri2019, roach2020}. Even though ACCESS-S2 shares the same
-model formulation with ACCESS-S1, the latter does not suffer from this
+model components as ACCESS-S1, the latter does not suffer from this
 bias, indicating that assimilating sea-ice concentrations successfully
@@
 implemented in the atmosphere component (in ACCESS-S1), indicating that
-sea-ice is indeed responding quickly to atmospheric perturbations.
+sea ice is indeed responding quickly to atmospheric perturbations.
 However, our analysis suggests that the atmosphere and ocean data
 assimilation implemented in ACCESS-S2 is only effectively influencing
-sea ice initial conditions from June to October, while the rest of the
+sea-ice initial conditions from June to October, while the rest of the
 year, the sea-ice component runs virtually free, reverting to its biased
@@
 Although ACCESS-S1 only assimilates sea-ice concentration, it is clear
-that sea-ice thickness is also affected. ACCESS-S1 simulates
-significantly thicker ice than ACCESS-S2 and in both systems sea-ice is
-thicker at shorter lead times than at longer lead times. Both the
-explicit data assimilation in ACCESS-S2 and the effects of atmospheric
-and oceanic data assimilation in ACCESS-S1 might be nudging simulated
-sea ice to be thicker than the model equilimbrium state.\\
-We believe that the thinner sea ice in ACCESS-S2 is a partial cause for
-the large sea-ice extent variance, but other mechanisms, such as
-unbalanced initial conditions might be important.
-
-Given that ACCESS-S2 sea-ice extent is essentially not initialised by
-the observations, comparing its forecasts with those of ACCESS-S1 allows
-us to estimate the time-scale for which initial conditions are
-important. We find that initial conditions affect Antarctic sea ice
+that sea-ice thickness is also affected through the assimilation
+process. ACCESS-S1 simulates significantly thicker ice than ACCESS-S2
+and in both systems sea-ice is thicker at shorter lead times than at
+longer lead times. Both the explicit data assimilation in ACCESS-S2 and
+the effects of atmospheric and oceanic data assimilation in ACCESS-S1
+might be nudging simulated sea ice to be thicker than the model
+equilimbrium state. We suggest that the thinner sea ice in ACCESS-S2
+contributes to the large sea-ice extent variance, but other mechanisms,
+such as unbalanced initial conditions might also be important.
+
+Given that ACCESS-S2 sea-ice extent is not directly initialised by
+sea-ice observations, comparing its forecasts with those of ACCESS-S1
+allows us to estimate the time-scale over which initial conditions are
+important. We find that initial conditions affect Antarctic sea-ice
 forecasts in the order of a few months, but that effect is seasonally
 dependent. January to April initial conditions improve forecasts for up
-to 3 months with, February initial conditions in particular are shown to
-be crucial for determining sea-ice evolution at least up to May. Arctic
-sea ice forecasts also show greater sensitivity to initial conditions in
-boreal summer, compared with boreal winter \citep{day2014, bunzel2016},
-suggesting a similar mechanism might be playing a role.
+to three months with, February initial conditions in particular are
+shown to be crucial for determining sea-ice evolution at least up to
+May. Arctic sea-ice forecasts also show greater sensitivity to initial
+conditions in boreal summer, compared with boreal winter
+\citep{day2014, bunzel2016}, suggesting a similar mechanism might be
+playing a role.
 
 Forecasts initialised in winter have very little skill and ACCESS-S1 and
-ACCESS-S2 forecast errors are not statistically different after just 2
-weeks. This is consistent with \citet{libera2022}'s finding of a
-``winter predictability barrier'' in the Weddell sea, although they
-describe the barrier as sharp loss of predictability in July, and here
-we find a gradual reduction in skill compared with climatology around
-June. This difference might be due to our use of pan-Antarctic RMSE,
-since our regional analysis indicates that the degraded skill is most
-dramatic in the King Haakon sea.
+ACCESS-S2 forecast errors are not statistically different after just two
+weeks. This is consistent with the findings of \citet{libera2022}'s
+finding of a ``winter predictability barrier'' in the Weddell Sea,
+although they describe the barrier as sharp loss of predictability in
+July, and here we find a gradual reduction in skill compared with
+climatology around June. This difference might be due to our use of
+pan-Antarctic RMSE, since our regional analysis indicates that the
+degraded skill is most dramatic in the King Haakon sea.
 
@@
 lead time (monthly mean values forecasted from the forecast initialised
-at the first of the month) and observations.}
+at the first of the month) and observations (OSI).}
 
@@
 \caption{\label{fig-bias-2-osi}Same as Figure~\ref{fig-bias-1} but for
-ACCESS-S1}
+ACCESS-S1.}
 
@@
 of 0, 2, 4, and 6 months. The RMSE and correlation during the
-overlapping period of ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) is
+overlapping period of ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) are
 shown on the top left and bottom left of each panel respectively.}
@@
 \caption{\label{fig-rmse_lon-2-osi}Same as Figure~\ref{fig-rmse_lon-1}
-but for ACCESS-S2}
+but for ACCESS-S2.}
 

### 2025-10-30 R/zenodo.R
Changes zenodo deposition id

@@ -1,5 +1,5 @@
 # Hard-coded deposition ID for this project.
-ZENODO_DEPOSITION_ID <- 391075
+ZENODO_DEPOSITION_ID <- 17479538
 
 # Internal: read the token from environment and validate
 .zenodo_get_token <- function() {

### 2025-11-10 analysis/paper/access-ice.qmd
Adds data repository
Changes affiliation

@@
     affiliations:
-      - name: Monash University, Australia,
+      - name: Monash University, Kulin Nations, Clayton, Victoria, Australia.
         id: 1
@@
         # postal-code: 02138
-      - name: Monash University, Australia
+      - name: Clayton, Kulin Nations, Victoria, Australia
         id: 2
-        department: Securing Antarctica’s Environmental Future
+        department: SARC Special Research Initiative for Securing Antarctica’s Environmental Future
         # address: 1 Oxford Street
@@
 Raw data of +S1 and +S2 forecast are not available due to size. 
-Derived datasets required to reproduce the results are available in this zenodo repository: XXXXX
+Derived datasets required to reproduce the results are available in this zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
 

### 2025-11-10 analysis/paper/access-ice.qmd
Adds Bethan comments

@@
 ```{r fig-hindcast-extent}
-#| fig-cap: "Row a: Pan-Antarctic mean sea-ice extent for all hindcasts initialised on the first of each calendar month for +S1 (column 1; green) and +S1 (column 2; purple).
+#| fig-cap: "Row a: Pan-Antarctic mean sea-ice extent for all hindcasts initialised on the first of each calendar month for +S1 (column 1; green) and +S2 (column 2; purple).
 Observed mean sea-ice extent in each corresponding hindcast period is shown in black.
@@
   scale_fill_models +
-  guides(colour = "none", fill = "none") +
+  guides(fill = "none") +
   # scale_color_manual(values = cetcolor::cet_pal(12, name = "c2"),

### 2025-11-19 analysis/paper/access-ice.qmd
Adds Xiaobing comments

@@
 ---
-title: Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Forecast Model
+title: The Importance of Initial Conditions in Seasonal Predictions of Antarctic Sea Ice
 format:
@@
     fig-pos: "!htb"
+    fig-dpi: 300
 header-includes: | 
@@
 The +S2 hindcast set used in this study runs for the period 1981–2018.
-Each hindcast consists of nine ensemble members built from three consecutive three-member forecasts created in the same manner as +S1 members initialised at the first of every month and the two previous days and run for 279 days.
+Ensemble members are created in the same manner as +S1 members, however, due to computing cost limitations, only three members per forecast initialisation date were run for 279 days. 
+Bigger ensambles were generated aggregating several three-member ensembles initialised on successive days [@wedd2022]. 
+Here, we build a nine-member time-lagged ensemble from three consecutive three-member forecasts initialised at the first of every month and the two previous days and run for 279 days. 
 We analyse the ensemble mean hindcasts unless otherwise specified.
@@
 Forecasts errors are also compared with hypothetical forecasts based on the persistence of anomalies and on climatology. 
-The persistence forecast is generated by extending the observed anomalies in time.
+The persistence forecast is generated by extending the observed sea-ice concentration anomalies the day of the forecast initialisation and comparing it with the actual anomalies observed. 
+The climatological forecast error is computed as the standard deviation of daily anomalies.
 
@@
 ```{r fig-hindcast-extent}
-#| fig-cap: "Row a: Pan-Antarctic mean sea-ice extent for all hindcasts initialised on the first of each calendar month for +S1 (column 1; green) and +S2 (column 2; purple).
+#| fig-cap: "Row a: Pan-Antarctic daily mean sea-ice extent for all hindcasts initialised on the first of each calendar month for +S1 (column 1; green) and +S2 (column 2; purple).
 Observed mean sea-ice extent in each corresponding hindcast period is shown in black.
 Row b: Mean differences between the forecast and the observed values.
-Circles represent the initial conditions at the start of forecasts (i.e., the first of every month), and triangles represent the mean values at the end of forecasts (i.e., the longest possible lead time)."
+Circles represent the initial conditions at the start of forecasts (i.e., the first of every month), and triangles represent the mean values at the lead time corresponding to the maximum lead time in S1 (between 213 and 216 days, depending on the month)"
 #| fig-height: 4
@@
   )) +
+  # geom_point(
+  #   data = \(x) {
+  #     x[
+  #       mday(time) == 2,
+  #       .SD[which.max(lag)],
+  #       by = .(forecast_time, model, measure)
+  #     ]
+  #   },
+  #   aes(fill = model),
+  #   shape = 24,
+  #   size = 2.2
+  # ) +
+
   geom_point(
@@
         mday(time) == 2,
-        .SD[which.max(lag)],
-        by = .(forecast_time, model, measure)
+        .SD[lag == max(lag[model == "S1"])],
+        by = .(forecast_time, measure)
       ]
@@
   ) +
+
   geom_point(data = \(x) x[lag == 1], aes(fill = model), shape = 21, size = 2) +
@@
 Mean extent at the first of every month is indicated with circles for the initial conditions and with triangles for the longest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
-At this long lead time, information of the initial conditions is essentially lost and the forecast reverts to each model's preferred equilibrium state.
+At this long lead time, information of the initial conditions is essentially lost and the forecast reverts close to each model's preferred equilibrium state.
 
@@
 
-```{r forecast_example}
+```{r plot_forecast_example}
+
+plot_forecast_example <- function(sample_date) {
   mlag <- 0
-sample_date <- "2008-05-02"
   forecast_date <- as.Date(sample_date)
@@
 
-```
-
-+S2 forecasts of sea-ice extent anomalies seem to align moderately well with observations (leading to moderately high correlation) but their magnitude is overestimated (leading to large errors).
-This could be caused by +S2 sea ice being much more sensitive to atmospheric and oceanic forcing, perhaps due to lower thickness.
-
-As an example, @fig-forecast-example shows sea-ice concentration anomalies (top row) and sea-ice thickness and the difference between the two models (bottom row) for `r format(as.Date(sample_date), "%e %B %Y")` initialised one day prior; being that close to initialisation date, these are very approximately the initial conditions.
-+S1 sea-ice concentrations anomalies are very close to observations as expected from the system assimilating these data. 
-+S2 sea-ice concentration anomalies, which are not assimilated, are not as close, but the large-scale pattern is aligned with observations.
-The system simulates large positive anomalies in the Weddell and Ross Seas and slight negative anomalies in the Amundsen and Bellingshausen Seas.
-The fact that +S2 can simulate this pattern without assimilating sea-ice data suggests that atmospheric and oceanic forcing were the dominant drivers.
-However, the magnitude of the sea-ice anomalies is too big. It is plausible that this is due to the thinner ice simulated by +S2 (bottom row).
-
-```{r fig-forecast-example}
-#| fig.cap: +S1 and +S2 hindcasts for 2 May 2008 at one day lead time.
-Top row shows sea-ice concentration anomalies forecasted by each system and the observations.
-Bottom row shows forecasted sea-ice thickness and the difference between +S1 and +S2.
-#| fig-height: 6
-
   b <- seq(-0.5, 0.5, by = 0.1)
@@
   aice_forecast_example |>
+    _[, model := factor(model, labels = c("S1", "S2", which_dataset))] |>
     ggplot(aes(xgrid, ygrid)) +
@@
           dcast(time + xgrid + ygrid ~ model, value.var = "hi") |>
-        _[, model := factor("Difference", levels = c("S1", "S2", "Difference"))]
+          _[,
+            model := factor("Difference", levels = c("S1", "S2", "Difference"))
+          ]
       },
@@
     wide_legend +
-  facet_wrap(~model) +
+    facet_wrap(~model, labeller = labeller(model = labels_models)) +
     plot_layout(ncol = 1)
+}
+
+```
+
++S2 forecasts of sea-ice extent anomalies seem to align moderately well with observations (leading to moderately high correlation) but their magnitude is overestimated (leading to large errors).
+This could be caused by +S2 sea ice being much more sensitive to atmospheric and oceanic forcing, perhaps due to lower thickness.
+
+As an example, @fig-forecast-example shows sea-ice concentration anomalies (top row) and sea-ice thickness and the difference between the two models (bottom row) for `r format(as.Date("2008-05-02"), "%e %B %Y")` initialised one day prior; being that close to initialisation date, these are very approximately the initial conditions.
++S1 sea-ice concentrations anomalies are very close to observations as expected from the system assimilating these data. 
++S2 sea-ice concentration anomalies, which are not assimilated, are not as close, but the large-scale pattern is aligned with observations.
+The system simulates large positive anomalies in the Weddell and Ross Seas and slight negative anomalies in the Amundsen and Bellingshausen Seas.
+The fact that +S2 can simulate this pattern without assimilating sea-ice data suggests that atmospheric and oceanic forcing were the dominant drivers.
+However, the magnitude of the sea-ice anomalies is too big. It is plausible that this is due to the thinner ice simulated by +S2 (bottom row).
+
+```{r fig-forecast-example}
+#| fig.cap: +S1 and +S2 hindcasts for 2 May 2008 at one day lead time.
+Top row shows sea-ice concentration anomalies forecasted by each system and the observations.
+Bottom row shows forecasted sea-ice thickness and the difference between +S1 and +S2.
+#| fig-height: 6
+plot_forecast_example(as.Date("2008-05-02"))
 ```
@@
 It is possible that sea-ice in the +S2 system is left in an unbalanced state after assimilating atmospheric and oceanic data but not sea-ice data, leading to large responses that are amplified by the thin ice in the initial states which then subside at longer lead times when the model is balanced.
-
-To assess +S2 forecasts in more detail, we compute error measures for all hindcasts started on the 1st of every month.
 
@@
 
+To assess +S2 forecasts in more detail, we compute error measures for all hindcasts started on the 1st of every month.
 @fig-rmse shows the mean RMSE of sea-ice concentration anomalies for +S1 and +S2 hindcasts compared against persistence and climatological forecasts used as a benchmark.
@@
 breaks <- AnchorBreaks(0, binwidth = binwidth)(range)
-rmse_lon_plots <- lapply(c("S1", "S2"), \(x) {
+rmse_lon_plots <- lapply(c("S1", "S2", "diff"), \(x) {
+  if (x == "diff") {
+    rmse_lon_mean_gdata2 <- rmse_lon_mean_gdata2 |>
+      dcast(lag + lon + obs_dataset + month ~ version, value.var = "skill") |>
+      _[, skill := S1 - S2]
+    breaks <- NULL
+  } else {
+    rmse_lon_mean_gdata2 <- rmse_lon_mean_gdata2 |>
+      _[version == x]
+  }
+
   rmse_lon_mean_gdata2 |>
-    _[version == x] |>
-    # _[,
-    #   skill := Smooth2D(lag, lon, skill, method = smooth_svd(0.05)),
-    #   by = .(month, version)
-    # ] |>
     ggplot(aes(lon, lag)) +
@@
 #| fig-height: 7
+#| dev: png
 rmse_lon_plots[[1]]
@@
 #| fig-height: 7
+#| dev: png
 rmse_lon_plots[[2]]
@@
 February and March are the only two months that can be forecasted with marginally positive skill in large regions.
+
+```{r}
+#| label: fig-rmse_lon-3
+#| fig-cap: Same as @fig-rmse_lon-1 but for the difference between +S1 and +S2.
+#| fig-height: 7
+#| dev: png
+rmse_lon_plots[[3]]
+```
+
+Finally, @fig-rmse_lon-3 shows the difference in skill between +S1 and +S2. 
+Large differences in skill indicate areas and months that are most affected by the data assimilation present in +S1.
+Between January and March, which are the months in which +S1 is the most skillfuil (@fig-lead-time-window), most of the improvement compared with +S2 is present in the Ross and Weddell Sea. 
+In April and May, the improvement seems more homogeneous.
 
@@
 However, our analysis suggests that the atmosphere and ocean data assimilation implemented in +S2 is only effectively influencing sea-ice initial conditions from June to October, while the rest of the year, the sea-ice component runs virtually free, reverting to its biased equilibrium state.
+@zhou2022 had previously evaluated sea-ice forecasts in +S2 and also highlighted the poor performance of this forecasting system attributed to the lack of good initial conditions.
 
@@
 
+We thank the internal reviewers Bethan White and Xiaobing Zhou for their comments and feedback.
+
+This work benefited by earlier unpublished work by Laura Davies, Phil Reid, Andrew G. Marshall.
+
+This research was undertaken with the assistance of resources from the National Computational Infrastructure (NCI Australia), an NCRIS enabled capability supported by the Australian Government.
+
 This work was supported by ARC SRIEAS Grant SR200100005 Securing Antarctica’s Environmental Future.
@@
 
+```{r, fig-rmse_lon-3-osi, ref.label=I("fig-rmse_lon-3")}
+
+```
+
 ```{r, fig-initial-spread-osi, ref.label=I("fig-initial-spread")}

### 2025-11-19 analysis/paper/access-ice.qmd
Fix typos

@@
 The two versions of the Australian Bureau of Meteorology's ACCESS seasonal forecast system, +S1 and +S2, use identical model configuration and differ only in their initial conditions; primarily in that +S2 does not assimilate sea-ice observations, whereas +S1 does.
-This provides an convenient opportunistic experiment to assess the role of initial conditions on Antarctic sea-ice forecasts using more than 20 years of fully coupled simulations with two 9-member ensembles.
+This provides a convenient opportunistic experiment to assess the role of initial conditions on Antarctic sea-ice forecasts using more than 20 years of fully coupled simulations with two 9-member ensembles.
 Our analysis reveals that both systems experience an extended melt season and delayed growth phase compared with observations. 
@@
 The land model uses the same horizontal grid as the atmosphere with four soil levels.
-The ocean component has a nominal horizontal resolution of 1/4º with 75 vertical levels.
+The ocean component has a nominal horizontal resolution of 1/4° with 75 vertical levels.
 The sea-ice component, based on CICE version 4.1, has the same resolution as the ocean component and five sea-ice thickness categories as well as an open water category.
@@
 This scheme uses an optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013].
-SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] in areas where SSTs are over 0ºC and Sea Surface Salinity is weakly nudged to the World Ocean Atlas 2013 climatology [@zweng2013].
+SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] in areas where SSTs are over 0°C and Sea Surface Salinity is weakly nudged to the World Ocean Atlas 2013 climatology [@zweng2013].
 
@@
 Ensemble members are created in the same manner as +S1 members, however, due to computing cost limitations, only three members per forecast initialisation date were run for 279 days. 
-Bigger ensambles were generated aggregating several three-member ensembles initialised on successive days [@wedd2022]. 
+Bigger ensembles were generated by aggregating several three-member ensembles initialised on successive days [@wedd2022]. 
 Here, we build a nine-member time-lagged ensemble from three consecutive three-member forecasts initialised at the first of every month and the two previous days and run for 279 days. 
@@
 Datasets derived using different algorithms and satellite platforms, each have their own biases and uncertainties.
-Estimates of inter-product uncertainty of sea-ice extent (SIE, defined here as the total region of the Southern Ocean with at least 15% sea-ice cover) is of the order of 0.5 million $km^2$ [@meier2019].
+Estimates of inter-product uncertainty of sea-ice extent (SIE, defined here as the total region of the Southern Ocean with at least 15% sea-ice cover) are of the order of 0.5 million $km^2$ [@meier2019].
 As will be shown below, this spread is minimal compared with the typical errors in the +S2 and +S1 forecasts, so the overall conclusions of this study are independent of the verification dataset used.
@@
 Both source algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites.
-The data has a spatial resolution of 25 by 25 km and daily from November 1978 onwards.
+The data have a spatial resolution of 25 by 25 km and daily from November 1978 onwards.
 
@@
 
-Forecasts errors are also compared with hypothetical forecasts based on the persistence of anomalies and on climatology. 
+Forecast errors are also compared with hypothetical forecasts based on the persistence of anomalies and on climatology. 
 The persistence forecast is generated by extending the observed sea-ice concentration anomalies the day of the forecast initialisation and comparing it with the actual anomalies observed. 
@@
 Negative skill score indicates that the forecast is worse than the reference forecast while positive values indicate an improvement. 
-A perfect forecast, would have zero RMSE and thus a skill score of 1.
+A perfect forecast would have zero RMSE and thus a skill score of 1.
 
@@
 
-We performed all analysis in this paper using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020b] packages.
+We performed all analyses in this paper using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020b] packages.
 Significant processing was performed using the CDO command line operators [@schulzweida2023].
 All graphics are made using ggplot2 [@wickham2009].
-The paper was rendered using knitr and quarto [@xie2015; @allaire2022].
+The paper was rendered using knitr and Quarto [@xie2015; @allaire2022].
 
@@
 +S1 has a comparatively smaller overall bias (@fig-bias-2). 
-The largest values are found between April and June, when the faster growth results in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Admunsen Seas.
+The largest values are found between April and June, when the faster growth results in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Amundsen Seas.
 
@@
 Only the first 120 days are shown.
-In parenthersis, the shortest time at which +S1 and +S2 mean RMSE is not statistically different at the 99% confidence level.
+In parentheses, the shortest time at which +S1 and +S2 mean RMSE is not statistically different at the 99% confidence level.
 #| fig-height: 5
@@
 #| label: fig-rmse_lon-1
-#| fig-cap: !expr glue::glue("RMSE skill score of +S1 forecasts with climatological forecast as reference computed on {uniqueN(rmse_lon_mean$lon)} meridional slices {diff(unique(rmse_lon_mean$lon))[1]}º wide as a function of lead time and longitude.
+#| fig-cap: !expr glue::glue("RMSE skill score of +S1 forecasts with climatological forecast as reference computed on {uniqueN(rmse_lon_mean$lon)} meridional slices {diff(unique(rmse_lon_mean$lon))[1]}° wide as a function of lead time and longitude.
 Antarctica’s coastline is shown at the bottom of each panel for reference.")
@@
 Since oceanic and atmospheric forcing is the only source of information, this suggests that sea-ice in this region is particularly sensitive to oceanic and atmospheric forcing and suggests a role of the Pacific-South American mode and the Amundsen Sea Low to shape sea-ice concentration anomalies.
-The fact that this is evident in the months in which El Niño--Southern Oscillation teleconnections are more important for atmospheric circulation also suggest the influence of tropical Pacific variability.
+The fact that this is evident in the months in which El Niño--Southern Oscillation teleconnections are more important for atmospheric circulation also suggests the influence of tropical Pacific variability.
 February and March are the only two months that can be forecasted with marginally positive skill in large regions.
@@
 Large differences in skill indicate areas and months that are most affected by the data assimilation present in +S1.
-Between January and March, which are the months in which +S1 is the most skillfuil (@fig-lead-time-window), most of the improvement compared with +S2 is present in the Ross and Weddell Sea. 
+Between January and March, which are the months in which +S1 is the most skilful (@fig-lead-time-window), most of the improvement compared with +S2 is present in the Ross and Weddell Sea. 
 In April and May, the improvement seems more homogeneous.
@@
 +S1 simulates significantly thicker ice than +S2 and in both systems sea-ice is thicker at shorter lead times than at longer lead times. 
-Both the explicit data assimilation in +S2 and the effects of atmospheric and oceanic data assimilation in +S1 might be nudging simulated sea ice to be thicker than the model equilimbrium state.
+Both the explicit data assimilation in +S2 and the effects of atmospheric and oceanic data assimilation in +S1 might be nudging simulated sea ice to be thicker than the model equilibrium state.
 We suggest that the thinner sea ice in +S2 contributes to the large sea-ice extent variance, but other mechanisms, such as unbalanced initial conditions might also be important.
@@
 We find that initial conditions affect Antarctic sea-ice forecasts in the order of a few months, but that effect is seasonally dependent. 
-January to April initial conditions improve forecasts for up to three months with, February initial conditions in particular are shown to be crucial for determining sea-ice evolution at least up to May.
+January to April initial conditions improve forecasts for up to three months.
+February initial conditions in particular are shown to be crucial for determining sea-ice evolution at least up to May. 
 Arctic sea-ice forecasts also show greater sensitivity to initial conditions in boreal summer, compared with boreal winter [@day2014; @bunzel2016], suggesting a similar mechanism might be playing a role.
 
-Forecasts initialised in winter have very little skill and +S1 and +S2 forecast errors are not statistically different after just two weeks. 
-This is consistent with the findings of @libera2022's finding of a “winter predictability barrier” in the Weddell Sea, although they describe the barrier as sharp loss of predictability in July, and here we find a gradual reduction in skill compared with climatology around June.
-This difference might be due to our use of pan-Antarctic RMSE, since our regional analysis indicates that the degraded skill is most dramatic in the King Haakon sea.
+Forecasts initialised in winter have very little skill and +S1 and +S2 forecast errors are statistically indistinguishable after just two weeks. 
+This is consistent with @libera2022's finding of a “winter predictability barrier”  in the Weddell Sea, although they describe the barrier as a sharp loss of predictability in July, and here we find a gradual reduction in skill compared with climatology around June.
+This difference might be due to our use of pan-Antarctic RMSE, since our regional analysis indicates that the degraded skill is most dramatic in the King Haakon Sea.
 
@@
 
-This work benefited by earlier unpublished work by Laura Davies, Phil Reid, Andrew G. Marshall.
+This work benefited from earlier unpublished work by Laura Davies, Phil Reid, Andrew G. Marshall.
 
@@
 
-The underlying code for this study is available on github and can be accessed via this link https://github.com/eliocamp/access-s2_ice-eval.
-
-Raw data of +S1 and +S2 forecast are not available due to size. 
-Derived datasets required to reproduce the results are available in this zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
+The underlying code for this study is available on GitHub and can be accessed via this link https://github.com/eliocamp/access-s2_ice-eval.
+
+Raw data of +S1 and +S2 forecasts are not available due to size. 
+Derived datasets required to reproduce the results are available in this Zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
 

### 2025-11-19 analysis/paper/access-ice.tex
Render

@@
 \hypersetup{
-  pdftitle={Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Forecast Model},
+  pdftitle={The Importance of Initial Conditions in Seasonal Predictions of Antarctic Sea Ice},
   pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Matthew C. Wheeler; Phillip Reid},
@@
 
-\title[Evaluating the Importance of Initial Conditions for Antarctic Sea
-Ice Seasonal Predictability with a Fully Coupled Forecast
-Model]{Evaluating the Importance of Initial Conditions for Antarctic Sea
-Ice Seasonal Predictability with a Fully Coupled Forecast Model}
+\title[The Importance of Initial Conditions in Seasonal Predictions of
+Antarctic Sea Ice]{The Importance of Initial Conditions in Seasonal
+Predictions of Antarctic Sea Ice}
 
@@
 \affil[1]{\orgdiv{School of Earth, Atmosphere and
-Environment}, \orgname{Monash University, Australia,}}
-\affil[2]{\orgdiv{Securing Antarctica's Environmental
-Future}, \orgname{Monash University, Australia}}
+Environment}, \orgname{Monash University, Kulin Nations, Clayton,
+Victoria, Australia.}}
+\affil[2]{\orgdiv{SARC Special Research Initiative for Securing
+Antarctica's Environmental Future}, \orgname{Clayton, Kulin Nations,
+Victoria, Australia}}
 \affil[3]{\orgdiv{Research, Bureau of Meteorology, Australia}}
@@
 ACCESS-S1 does.\\
-This provides an convenient opportunistic experiment to assess the role
+This provides a convenient opportunistic experiment to assess the role
 of initial conditions on Antarctic sea-ice forecasts using more than 20
@@
 soil levels. The ocean component has a nominal horizontal resolution of
-1/4º with 75 vertical levels. The sea-ice component, based on CICE
+1/4° with 75 vertical levels. The sea-ice component, based on CICE
 version 4.1, has the same resolution as the ocean component and five
@@
 nudged to Reynolds OISSTv2.1 \citep{reynolds2007} in areas where SSTs
-are over 0ºC and Sea Surface Salinity is weakly nudged to the World
+are over 0°C and Sea Surface Salinity is weakly nudged to the World
 Ocean Atlas 2013 climatology \citep{zweng2013}.
@@
 initialised at the first of every month. The ACCESS-S2 hindcast set used
-in this study runs for the period 1981--2018. Each hindcast consists of
-nine ensemble members built from three consecutive three-member
-forecasts created in the same manner as ACCESS-S1 members initialised at
-the first of every month and the two previous days and run for 279 days.
-We analyse the ensemble mean hindcasts unless otherwise specified.
+in this study runs for the period 1981--2018. Ensemble members are
+created in the same manner as ACCESS-S1 members, however, due to
+computing cost limitations, only three members per forecast
+initialisation date were run for 279 days. Bigger ensembles were
+generated by aggregating several three-member ensembles initialised on
+successive days \citep{wedd2022}. Here, we build a nine-member
+time-lagged ensemble from three consecutive three-member forecasts
+initialised at the first of every month and the two previous days and
+run for 279 days. We analyse the ensemble mean hindcasts unless
+otherwise specified.
 
@@
 uncertainty of sea-ice extent (SIE, defined here as the total region of
-the Southern Ocean with at least 15\% sea-ice cover) is of the order of
+the Southern Ocean with at least 15\% sea-ice cover) are of the order of
 0.5 million \(km^2\) \citep{meier2019}. As will be shown below, this
@@
 (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP)
--F8, -F11, and -F13 satellites. The data has a spatial resolution of 25
+-F8, -F11, and -F13 satellites. The data have a spatial resolution of 25
 by 25 km and daily from November 1978 onwards.
@@
 
-Forecasts errors are also compared with hypothetical forecasts based on
+Forecast errors are also compared with hypothetical forecasts based on
 the persistence of anomalies and on climatology. The persistence
-forecast is generated by extending the observed anomalies in time.
+forecast is generated by extending the observed sea-ice concentration
+anomalies the day of the forecast initialisation and comparing it with
+the actual anomalies observed. The climatological forecast error is
+computed as the standard deviation of daily anomalies.
 
@@
 forecast is worse than the reference forecast while positive values
-indicate an improvement. A perfect forecast, would have zero RMSE and
+indicate an improvement. A perfect forecast would have zero RMSE and
 thus a skill score of 1.
@@
 
-We performed all analysis in this paper using the R programming language
+We performed all analyses in this paper using the R programming language
 \citep{rcoreteam2020}, using data.table \citep{dowle2020} and metR
@@
 graphics are made using ggplot2 \citep{wickham2009}. The paper was
-rendered using knitr and quarto \citep{xie2015, allaire2022}.
+rendered using knitr and Quarto \citep{xie2015, allaire2022}.
 
@@
 
-\caption{\label{fig-hindcast-extent}Row a: Pan-Antarctic mean sea-ice
-extent for all hindcasts initialised on the first of each calendar month
-for ACCESS-S1 (column 1; green) and ACCESS-S1 (column 2; purple).
-Observed mean sea-ice extent in each corresponding hindcast period is
-shown in black. Row b: Mean differences between the forecast and the
-observed values. Circles represent the initial conditions at the start
-of forecasts (i.e., the first of every month), and triangles represent
-the mean values at the end of forecasts (i.e., the longest possible lead
-time).}
+\caption{\label{fig-hindcast-extent}Row a: Pan-Antarctic daily mean
+sea-ice extent for all hindcasts initialised on the first of each
+calendar month for ACCESS-S1 (column 1; green) and ACCESS-S2 (column 2;
+purple). Observed mean sea-ice extent in each corresponding hindcast
+period is shown in black. Row b: Mean differences between the forecast
+and the observed values. Circles represent the initial conditions at the
+start of forecasts (i.e., the first of every month), and triangles
+represent the mean values at the lead time corresponding to the maximum
+lead time in S1 (between 213 and 216 days, depending on the month)}
 
@@
 216 days for ACCESS-S1). At this long lead time, information of the
-initial conditions is essentially lost and the forecast reverts to each
-model's preferred equilibrium state.
+initial conditions is essentially lost and the forecast reverts close to
+each model's preferred equilibrium state.
 
@@
 the sea-ice edge, and in January, when the faster melt leads to large
-negative bias in the Weddell and Admunsen Seas.
+negative bias in the Weddell and Amundsen Seas.
 
@@
 
-To assess ACCESS-S2 forecasts in more detail, we compute error measures
-for all hindcasts started on the 1st of every month.
-
 \begin{figure}[!htb]
@@
 anomalies (black) and climatology (gray). Only the first 120 days are
-shown. In parenthersis, the shortest time at which ACCESS-S1 and
+shown. In parentheses, the shortest time at which ACCESS-S1 and
 ACCESS-S2 mean RMSE is not statistically different at the 99\%
@@
 
+To assess ACCESS-S2 forecasts in more detail, we compute error measures
+for all hindcasts started on the 1st of every month.
 Figure~\ref{fig-rmse} shows the mean RMSE of sea-ice concentration
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-1.png}}
 
@@
 with climatological forecast as reference computed on 15 meridional
-slices 24º wide as a function of lead time and longitude. Antarctica's
+slices 24° wide as a function of lead time and longitude. Antarctica's
 coastline is shown at the bottom of each panel for reference.}
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-1.png}}
 
@@
 the months in which El Niño--Southern Oscillation teleconnections are
-more important for atmospheric circulation also suggest the influence of
-tropical Pacific variability. February and March are the only two months
-that can be forecasted with marginally positive skill in large regions.
+more important for atmospheric circulation also suggests the influence
+of tropical Pacific variability. February and March are the only two
+months that can be forecasted with marginally positive skill in large
+regions.
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-3-1.png}}
+
+}
+
+\caption{\label{fig-rmse_lon-3}Same as Figure~\ref{fig-rmse_lon-1} but
+for the difference between ACCESS-S1 and ACCESS-S2.}
+
+\end{figure}%
+
+Finally, Figure~\ref{fig-rmse_lon-3} shows the difference in skill
+between ACCESS-S1 and ACCESS-S2. Large differences in skill indicate
+areas and months that are most affected by the data assimilation present
+in ACCESS-S1. Between January and March, which are the months in which
+ACCESS-S1 is the most skilful (Figure~\ref{fig-lead-time-window}), most
+of the improvement compared with ACCESS-S2 is present in the Ross and
+Weddell Sea. In April and May, the improvement seems more homogeneous.
 
@@
 year, the sea-ice component runs virtually free, reverting to its biased
-equilibrium state.
+equilibrium state. \citet{zhou2022} had previously evaluated sea-ice
+forecasts in ACCESS-S2 and also highlighted the poor performance of this
+forecasting system attributed to the lack of good initial conditions.
 
@@
 might be nudging simulated sea ice to be thicker than the model
-equilimbrium state. We suggest that the thinner sea ice in ACCESS-S2
+equilibrium state. We suggest that the thinner sea ice in ACCESS-S2
 contributes to the large sea-ice extent variance, but other mechanisms,
@@
 dependent. January to April initial conditions improve forecasts for up
-to three months with, February initial conditions in particular are
-shown to be crucial for determining sea-ice evolution at least up to
-May. Arctic sea-ice forecasts also show greater sensitivity to initial
-conditions in boreal summer, compared with boreal winter
-\citep{day2014, bunzel2016}, suggesting a similar mechanism might be
-playing a role.
+to three months. February initial conditions in particular are shown to
+be crucial for determining sea-ice evolution at least up to May. Arctic
+sea-ice forecasts also show greater sensitivity to initial conditions in
+boreal summer, compared with boreal winter \citep{day2014, bunzel2016},
+suggesting a similar mechanism might be playing a role.
 
 Forecasts initialised in winter have very little skill and ACCESS-S1 and
-ACCESS-S2 forecast errors are not statistically different after just two
-weeks. This is consistent with the findings of \citet{libera2022}'s
-finding of a ``winter predictability barrier'' in the Weddell Sea,
-although they describe the barrier as sharp loss of predictability in
-July, and here we find a gradual reduction in skill compared with
-climatology around June. This difference might be due to our use of
-pan-Antarctic RMSE, since our regional analysis indicates that the
-degraded skill is most dramatic in the King Haakon sea.
+ACCESS-S2 forecast errors are statistically indistinguishable after just
+two weeks. This is consistent with \citet{libera2022}'s finding of a
+``winter predictability barrier'' in the Weddell Sea, although they
+describe the barrier as a sharp loss of predictability in July, and here
+we find a gradual reduction in skill compared with climatology around
+June. This difference might be due to our use of pan-Antarctic RMSE,
+since our regional analysis indicates that the degraded skill is most
+dramatic in the King Haakon Sea.
 
@@
 
+We thank the internal reviewers Bethan White and Xiaobing Zhou for their
+comments and feedback.
+
+This work benefited from earlier unpublished work by Laura Davies, Phil
+Reid, Andrew G. Marshall.
+
+This research was undertaken with the assistance of resources from the
+National Computational Infrastructure (NCI Australia), an NCRIS enabled
+capability supported by the Australian Government.
+
 This work was supported by ARC SRIEAS Grant SR200100005 Securing
@@
 
-The underlying code for this study is available on github and can be
+The underlying code for this study is available on GitHub and can be
 accessed via this link https://github.com/eliocamp/access-s2\_ice-eval.
 
-Raw data of ACCESS-S1 and ACCESS-S2 forecast are not available due to
+Raw data of ACCESS-S1 and ACCESS-S2 forecasts are not available due to
 size. Derived datasets required to reproduce the results are available
-in this zenodo repository: XXXXX
+in this Zenodo repository: https://zenodo.org/records/17479538
+\citep{campitelli2025}
 
@@
 
-\caption{\label{fig-hindcast-extent-osi}Row a: Pan-Antarctic mean
+\caption{\label{fig-hindcast-extent-osi}Row a: Pan-Antarctic daily mean
 sea-ice extent for all hindcasts initialised on the first of each
-calendar month for ACCESS-S1 (column 1; green) and ACCESS-S1 (column 2;
+calendar month for ACCESS-S1 (column 1; green) and ACCESS-S2 (column 2;
 purple). Observed mean sea-ice extent in each corresponding hindcast
@@
 start of forecasts (i.e., the first of every month), and triangles
-represent the mean values at the end of forecasts (i.e., the longest
-possible lead time).}
+represent the mean values at the lead time corresponding to the maximum
+lead time in S1 (between 213 and 216 days, depending on the month)}
 
@@
 forecast of persistence of anomalies (black) and climatology (gray).
-Only the first 120 days are shown. In parenthersis, the shortest time at
+Only the first 120 days are shown. In parentheses, the shortest time at
 which ACCESS-S1 and ACCESS-S2 mean RMSE is not statistically different
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-osi-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-osi-1.png}}
 
@@
 forecasts with climatological forecast as reference computed on 15
-meridional slices 24º wide as a function of lead time and longitude.
+meridional slices 24° wide as a function of lead time and longitude.
 Antarctica's coastline is shown at the bottom of each panel for
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-osi-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-osi-1.png}}
 
@@
 but for ACCESS-S2.}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-3-osi-1.png}}
+
+}
+
+\caption{\label{fig-rmse_lon-3-osi}Same as Figure~\ref{fig-rmse_lon-1}
+but for the difference between ACCESS-S1 and ACCESS-S2.}
 

### 2025-11-19 analysis/paper/access-ice.qmd
Adds competing intests and author contributions

@@
 
-The underlying code for this study is available on GitHub and can be accessed via this link https://github.com/eliocamp/access-s2_ice-eval.
-
-Raw data of +S1 and +S2 forecasts are not available due to size. 
-Derived datasets required to reproduce the results are available in this Zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
+The underlying code for this study is available on GitHub: https://github.com/eliocamp/access-s2_ice-eval.
+
+Raw data of +S1 and +S2 hindcast are not available due to size. 
+Derived datasets required to reproduce the results, including extent timeseries and error measures, are available in this Zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
+
+# Competing interests
+
+All authors declare no financial or non-financial competing interests.
+
+# Author contributions
+
+EC performed the data analysis and wrote the manuscript draft. 
+AP, JA, EL, MW and PR, performed interpretation of the results, and reviewed and edited the draft. 
+All authors read and approved the final manuscript.
 

### 2025-11-19 analysis/paper/access-ice.tex
Renders

@@
 
-The underlying code for this study is available on GitHub and can be
-accessed via this link https://github.com/eliocamp/access-s2\_ice-eval.
-
-Raw data of ACCESS-S1 and ACCESS-S2 forecasts are not available due to
-size. Derived datasets required to reproduce the results are available
-in this Zenodo repository: https://zenodo.org/records/17479538
-\citep{campitelli2025}
+The underlying code for this study is available on GitHub:
+https://github.com/eliocamp/access-s2\_ice-eval.
+
+Raw data of ACCESS-S1 and ACCESS-S2 hindcast are not available due to
+size. Derived datasets required to reproduce the results, including
+extent timeseries and error measures, are available in this Zenodo
+repository: https://zenodo.org/records/17479538 \citep{campitelli2025}
+
+\section{Competing interests}\label{competing-interests}
+
+All authors declare no financial or non-financial competing interests.
+
+\section{Author contributions}\label{author-contributions}
+
+EC performed the data analysis and wrote the manuscript draft. AP, JA,
+EL, MW and PR, performed interpretation of the results, and reviewed and
+edited the draft. All authors read and approved the final manuscript.
 

### 2025-11-27 analysis/paper/access-ice.qmd
Fix journal technical checks

@@
   nature-pdf:
+    latex-clean: false
     # journal.cite-style is included in the tex file but ignored by pandoc if 
@@
     fig-dpi: 300    
-header-includes: | 
-   \usepackage[section]{placeins}
 author:
@@
     affiliations:
-      - department: Research, Bureau of Meteorology, Australia
+      - department: Research, Bureau of Meteorology, Melbourne, Australia
         id: 3
@@
 
+knitr::opts_chunk$set(
+  fig.path = "access-ice_files/figure-pdf/",
+  cache.path = here::here("analysis/paper/access-ice_cache/pdf/")
+)
+
 cdo_options_set("-L")
@@
 ```
+
+::: {.content-visible when-profile="main"}
 
@@
 
-@fig-hindcast-extent shows mean sea-ice extent of the +S1 and +S2 hindcasts (row a) and their differences from mean sea-ice extent of +CDR (row b).
+@fig-hindcast-extent (Figure \ref{fig-hindcast-extent-osi} for OSI) shows mean sea-ice extent of the +S1 and +S2 hindcasts (row a) and their differences from mean sea-ice extent of +CDR (row b).
 Mean extent at the first of every month is indicated with circles for the initial conditions and with triangles for the longest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
@@
 This is due primarily to the melt season being longer than in observations and with faster melt between January and March and the growing seasons being shorter with slower growth during March and April.
-This is then followed by faster growth between May and July (@fig-mean-growth).
+This is then followed by faster growth between May and July (@fig-mean-growth and Figure \ref{fig-mean-growth-osi}).
 Many sea-ice models exhibit this systematic underestimation during the sea-ice minimum and early freezing season [@massonnet2023], which could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
@@
 ```{r fig-bias-2}
-#| fig-cap: Same as @fig-bias-1 but for +S1.
+#| fig-cap: Same as Figure \ref{fig-bias-1} but for +S1.
 #| fig-height: 6
@@
 
-To further understand the bias in +S2, @fig-bias-1 shows spatial patterns of the differences of monthly mean sea-ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest monthly lead time.
+To further understand the bias in +S2, @fig-bias-1 (Figure \ref{fig-bias-1-osi}) shows spatial patterns of the differences of monthly mean sea-ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest monthly lead time.
 From October to May, the model underestimates sea-ice concentrations in most regions except for the inner Weddell Sea in April and May, where sea-ice concentrations saturate to 1 both in the observations and forecasts.
@@
 
-+S1 has a comparatively smaller overall bias (@fig-bias-2). 
++S1 has a comparatively smaller overall bias (@fig-bias-2 and Figure \ref{fig-bias-2-osi}). 
 The largest values are found between April and June, when the faster growth results in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Amundsen Seas.
@@
 
-@fig-extent-anom shows monthly sea-ice extent anomalies forecasted at selected lead times.
+@fig-extent-anom (Figure \ref{fig-extent-anom-osi}) shows monthly sea-ice extent anomalies forecasted at selected lead times.
 Compared with +S1, +S2 anomaly forecasts are relatively poor (large RMSE) even for the first month (lead time 0), whereas +S1 forecasts stay relatively skilful even at a lead time of three months.
@@
 The explanation seems to be the mentioned increased interannual variability.
-@fig-extent-sd shows the interannual standard deviation of monthly sea-ice extent of the forecasts as a function of lead time compared with observations.
+@fig-extent-sd (Figure \ref{fig-extent-sd-osi}) shows the interannual standard deviation of monthly sea-ice extent of the forecasts as a function of lead time compared with observations.
 +S1 standard deviation lies within the observed standard deviation regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at nine month lead time for most months.
@@
 
-Extending beyond the one case in @fig-forecast-example, @fig-mean-thickness shows monthly mean sea-ice thickness as a function of lead time for +S1 and +S2.
+Extending beyond the one case in @fig-forecast-example, @fig-mean-thickness (Figure \ref{fig-mean-thickness-osi}) shows monthly mean sea-ice thickness as a function of lead time for +S1 and +S2.
 Supporting the idea that thinner ice is what causes the increased extent variability in +S2, this system  simulates thinner sea-ice compared to +S1 overall at almost all lead times and in all months except for summer at short lead times (Dec-Jan, 0-1 months; Feb-Mar, 0-2 months).
@@
 To assess +S2 forecasts in more detail, we compute error measures for all hindcasts started on the 1st of every month.
-@fig-rmse shows the mean RMSE of sea-ice concentration anomalies for +S1 and +S2 hindcasts compared against persistence and climatological forecasts used as a benchmark.
+@fig-rmse (Figure \ref{fig-rmse-osi}) shows the mean RMSE of sea-ice concentration anomalies for +S1 and +S2 hindcasts compared against persistence and climatological forecasts used as a benchmark.
 Due to errors in the initial conditions, it is expected that persistence forecasts would be better than the model forecasts at very short lead times, but that the persistence forecast errors would grow faster and may eventually surpass the model forecast errors.
@@
 
-@fig-lead-time-window summarises the lead time window in which each hindcast is better than both the persistence forecast and the climatological forecast as a function of forecast month. 
+@fig-lead-time-window (Figure \ref{fig-lead-time-window-osi}) summarises the lead time window in which each hindcast is better than both the persistence forecast and the climatological forecast as a function of forecast month.
 +S1 forecasts have a wider lead time window in the summer than the other seasons and is not better than both benchmarks at forecasting June sea-ice concentration anomalies. 
@@
 #| label: fig-rmse_lon-2
-#| fig-cap: Same as @fig-rmse_lon-1 but for +S2.
+#| fig-cap: Same as Figure \ref{fig-rmse_lon-1} but for +S2.
 #| fig-height: 7
@@
 
-For +S1 forecasts (@fig-rmse_lon-1), skill tends to be lower off the coast of Eastern Antarctica even at short lead times; for instance, the skill score for forecasts initialised in May and June are negative between 0° and 120°E even at almost zero lead time.
+For +S1 forecasts (@fig-rmse_lon-1 and Figure \ref{fig-rmse_lon-1-osi}), skill tends to be lower off the coast of Eastern Antarctica even at short lead times; for instance, the skill score for forecasts initialised in May and June are negative between 0° and 120°E even at almost zero lead time.
 This mirrors @libera2022 findings of a “winter predictability barrier”, although they focus on the Weddell Sea and here we show that the effect seems to be stronger more to the east. 
@@
 
-+S2 forecasts  (@fig-rmse_lon-2) also have lower skill over East Antarctica. 
++S2 forecasts  (@fig-rmse_lon-2 and Figure \ref{fig-rmse_lon-2-osi}) also have lower skill over East Antarctica. 
 From July to December even though the pan-Antarctic average skill is negative at all lead times ([Fig. @fig-lead-time-window]), it is positive for up to a month in West Antarctica. 
@@
 #| label: fig-rmse_lon-3
-#| fig-cap: Same as @fig-rmse_lon-1 but for the difference between +S1 and +S2.
+#| fig-cap: Same as Figure \ref{fig-rmse_lon-1} but for the difference between +S1 and +S2.
 #| fig-height: 7
@@
 
-Finally, @fig-rmse_lon-3 shows the difference in skill between +S1 and +S2. 
+Finally, @fig-rmse_lon-3 (Figure \ref{fig-rmse_lon-3-osi}) shows the difference in skill between +S1 and +S2. 
 Large differences in skill indicate areas and months that are most affected by the data assimilation present in +S1.
@@
 In @fig-rmse the mean error was shown. 
-@fig-initial-spread column 1 shows the mean standard deviation of errors among ensemble members at various lead times.
+@fig-initial-spread (Figure \ref{fig-initial-spread-osi}) column 1 shows the mean standard deviation of errors among ensemble members at various lead times.
 At one day lead time ([Fig. @fig-initial-spread] a.1) +S2 has a slightly larger spread than +S1 due to the way that ensemble members are generated.
@@
 
+# Funding
+
+This work was supported by ARC SRIEAS Grant SR200100005 Securing Antarctica’s Environmental Future.
+
 # Acknowledgments
@@
 This research was undertaken with the assistance of resources from the National Computational Infrastructure (NCI Australia), an NCRIS enabled capability supported by the Australian Government.
-
-This work was supported by ARC SRIEAS Grant SR200100005 Securing Antarctica’s Environmental Future.
 
@@
 
+:::
+
+::: {.content-visible when-profile="supp"}
+
 \clearpage
@@
 ```
+
+:::

### 2025-11-27 build.sh

@@ -1,0 +1,31 @@
+set -eu
+
+dir=analysis/paper
+source=$dir/access-ice.qmd
+
+main=$dir/main.qmd 
+supp=$dir/supp.qmd 
+zip=access-ice.zip
+
+cp $source $main
+quarto render $main --profile main
+cp $dir/main.aux $dir/ref_main.aux
+
+cp $source $supp
+quarto render $supp --profile supp
+cp $dir/supp.aux $dir/ref_supp.aux
+
+rm $main $supp 
+
+cd $dir
+pdflatex main.tex
+pdflatex supp.tex
+
+
+
+rm $zip
+zip -r $zip *.bst *.csl references.bib main.tex ref_supp.aux access-ice_files/
+
+# Clean up everything 
+find . -maxdepth 1 -type f -name 'main*' ! -name 'main.pdf' ! -name 'main.tex' -exec rm {} +
+find . -maxdepth 1 -type f -name 'supp*' ! -name 'supp.pdf' ! -name 'supp.tex' -exec rm {} +

### 2025-11-27 analysis/paper/main.tex
Render

--- 2025-11-19 analysis/paper/access-ice.tex
@@
 \usepackage[section]{placeins}
+\usepackage{xr}
+\externaldocument{ref_supp}
 \makeatletter
@@
 Victoria, Australia}}
-\affil[3]{\orgdiv{Research, Bureau of Meteorology, Australia}}
+\affil[3]{\orgdiv{Research, Bureau of Meteorology, Melbourne,
+Australia}}
 
@@
 
-Figure~\ref{fig-hindcast-extent} shows mean sea-ice extent of the
-ACCESS-S1 and ACCESS-S2 hindcasts (row a) and their differences from
-mean sea-ice extent of NOAA/NSIDC CDRV4 (row b). Mean extent at the
-first of every month is indicated with circles for the initial
-conditions and with triangles for the longest lead time possible for
-each model (between 274 and 277 days for ACCESS-S2 and between 213 and
-216 days for ACCESS-S1). At this long lead time, information of the
-initial conditions is essentially lost and the forecast reverts close to
-each model's preferred equilibrium state.
+Figure~\ref{fig-hindcast-extent} (Figure \ref{fig-hindcast-extent-osi}
+for OSI) shows mean sea-ice extent of the ACCESS-S1 and ACCESS-S2
+hindcasts (row a) and their differences from mean sea-ice extent of
+NOAA/NSIDC CDRV4 (row b). Mean extent at the first of every month is
+indicated with circles for the initial conditions and with triangles for
+the longest lead time possible for each model (between 274 and 277 days
+for ACCESS-S2 and between 213 and 216 days for ACCESS-S1). At this long
+lead time, information of the initial conditions is essentially lost and
+the forecast reverts close to each model's preferred equilibrium state.
 
@@
 slower growth during March and April. This is then followed by faster
-growth between May and July (Figure~\ref{fig-mean-growth}). Many sea-ice
-models exhibit this systematic underestimation during the sea-ice
-minimum and early freezing season \citep{massonnet2023}, which could
-indicate problems in the representation of thermodynamics in the model
-\citep{zampieri2019}. It is also not surprising that both forecasting
-systems converge to a similar equilibrium state because they share the
-same model formulation.
+growth between May and July (Figure~\ref{fig-mean-growth} and Figure
+\ref{fig-mean-growth-osi}). Many sea-ice models exhibit this systematic
+underestimation during the sea-ice minimum and early freezing season
+\citep{massonnet2023}, which could indicate problems in the
+representation of thermodynamics in the model \citep{zampieri2019}. It
+is also not surprising that both forecasting systems converge to a
+similar equilibrium state because they share the same model formulation.
 
@@
 
-\caption{\label{fig-bias-2}Same as Figure~\ref{fig-bias-1} but for
+\caption{\label{fig-bias-2}Same as Figure \ref{fig-bias-1} but for
 ACCESS-S1.}
@@
 To further understand the bias in ACCESS-S2, Figure~\ref{fig-bias-1}
-shows spatial patterns of the differences of monthly mean sea-ice
-concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
-shortest monthly lead time. From October to May, the model
-underestimates sea-ice concentrations in most regions except for the
-inner Weddell Sea in April and May, where sea-ice concentrations
-saturate to 1 both in the observations and forecasts. In winter, the
-differences are limited to a narrow band around the sea-ice edge with
-slight positive biases in the African sector of East Antarctica and
-negative biases around the Indian Ocean sector which partially
-compensate, resulting in the near-zero extent bias seen in those months
-(Figure~\ref{fig-hindcast-extent}).
+(Figure \ref{fig-bias-1-osi}) shows spatial patterns of the differences
+of monthly mean sea-ice concentrations between NOAA/NSIDC CDRV4 and
+ACCESS-S2 hindcasts at the shortest monthly lead time. From October to
+May, the model underestimates sea-ice concentrations in most regions
+except for the inner Weddell Sea in April and May, where sea-ice
+concentrations saturate to 1 both in the observations and forecasts. In
+winter, the differences are limited to a narrow band around the sea-ice
+edge with slight positive biases in the African sector of East
+Antarctica and negative biases around the Indian Ocean sector which
+partially compensate, resulting in the near-zero extent bias seen in
+those months (Figure~\ref{fig-hindcast-extent}).
 
 ACCESS-S1 has a comparatively smaller overall bias
-(Figure~\ref{fig-bias-2}). The largest values are found between April
-and June, when the faster growth results in large positive bias along
-the sea-ice edge, and in January, when the faster melt leads to large
-negative bias in the Weddell and Amundsen Seas.
+(Figure~\ref{fig-bias-2} and Figure \ref{fig-bias-2-osi}). The largest
+values are found between April and June, when the faster growth results
+in large positive bias along the sea-ice edge, and in January, when the
+faster melt leads to large negative bias in the Weddell and Amundsen
+Seas.
 
@@
 
-Figure~\ref{fig-extent-anom} shows monthly sea-ice extent anomalies
-forecasted at selected lead times. Compared with ACCESS-S1, ACCESS-S2
-anomaly forecasts are relatively poor (large RMSE) even for the first
-month (lead time 0), whereas ACCESS-S1 forecasts stay relatively skilful
-even at a lead time of three months. ACCESS-S2 shows much larger
-interannual variability than observations, with dramatic lows between
-1995 and 2007, and highs between 2007 and 2015.
+Figure~\ref{fig-extent-anom} (Figure \ref{fig-extent-anom-osi}) shows
+monthly sea-ice extent anomalies forecasted at selected lead times.
+Compared with ACCESS-S1, ACCESS-S2 anomaly forecasts are relatively poor
+(large RMSE) even for the first month (lead time 0), whereas ACCESS-S1
+forecasts stay relatively skilful even at a lead time of three months.
+ACCESS-S2 shows much larger interannual variability than observations,
+with dramatic lows between 1995 and 2007, and highs between 2007 and
+2015.
 
@@
 explanation seems to be the mentioned increased interannual variability.
-Figure~\ref{fig-extent-sd} shows the interannual standard deviation of
-monthly sea-ice extent of the forecasts as a function of lead time
-compared with observations. ACCESS-S1 standard deviation lies within the
-observed standard deviation regardless of lead time, while ACCESS-S2
-standard deviation is more than twice that of observations at zero lead
-time and only approaches the observed value at nine month lead time for
-most months.
+Figure~\ref{fig-extent-sd} (Figure \ref{fig-extent-sd-osi}) shows the
+interannual standard deviation of monthly sea-ice extent of the
+forecasts as a function of lead time compared with observations.
+ACCESS-S1 standard deviation lies within the observed standard deviation
+regardless of lead time, while ACCESS-S2 standard deviation is more than
+twice that of observations at zero lead time and only approaches the
+observed value at nine month lead time for most months.
 
@@
 Extending beyond the one case in Figure~\ref{fig-forecast-example},
-Figure~\ref{fig-mean-thickness} shows monthly mean sea-ice thickness as
-a function of lead time for ACCESS-S1 and ACCESS-S2. Supporting the idea
-that thinner ice is what causes the increased extent variability in
-ACCESS-S2, this system simulates thinner sea-ice compared to ACCESS-S1
-overall at almost all lead times and in all months except for summer at
-short lead times (Dec-Jan, 0-1 months; Feb-Mar, 0-2 months). However, in
-both systems, forecasted sea-ice is thicker at shorter lead times and
-then decreases, particularly in the summer months. If thinner ice were a
-sufficient cause of increased variability, then we would expect
-variability to increase with lead time in both forecasting systems.
+Figure~\ref{fig-mean-thickness} (Figure \ref{fig-mean-thickness-osi})
+shows monthly mean sea-ice thickness as a function of lead time for
+ACCESS-S1 and ACCESS-S2. Supporting the idea that thinner ice is what
+causes the increased extent variability in ACCESS-S2, this system
+simulates thinner sea-ice compared to ACCESS-S1 overall at almost all
+lead times and in all months except for summer at short lead times
+(Dec-Jan, 0-1 months; Feb-Mar, 0-2 months). However, in both systems,
+forecasted sea-ice is thicker at shorter lead times and then decreases,
+particularly in the summer months. If thinner ice were a sufficient
+cause of increased variability, then we would expect variability to
+increase with lead time in both forecasting systems.
 
@@
 for all hindcasts started on the 1st of every month.
-Figure~\ref{fig-rmse} shows the mean RMSE of sea-ice concentration
-anomalies for ACCESS-S1 and ACCESS-S2 hindcasts compared against
-persistence and climatological forecasts used as a benchmark. Due to
-errors in the initial conditions, it is expected that persistence
-forecasts would be better than the model forecasts at very short lead
-times, but that the persistence forecast errors would grow faster and
-may eventually surpass the model forecast errors. The black line shows
-that the persistence forecast error indeed grows rapidly and reaches its
-maximum in about 30 days for most months except for February, when it
-grows much slower. The ACCESS-S1 forecast errors grow slower than
-persistence forecast errors and remain lower after less than 10 days on
-average. The ACCESS-S2 forecast error starts high in all months and is
-lower than the persistence forecast error after more than 15 days in
-most months except for forecast initialised in February, when it takes
-80 days.
+Figure~\ref{fig-rmse} (Figure \ref{fig-rmse-osi}) shows the mean RMSE of
+sea-ice concentration anomalies for ACCESS-S1 and ACCESS-S2 hindcasts
+compared against persistence and climatological forecasts used as a
+benchmark. Due to errors in the initial conditions, it is expected that
+persistence forecasts would be better than the model forecasts at very
+short lead times, but that the persistence forecast errors would grow
+faster and may eventually surpass the model forecast errors. The black
+line shows that the persistence forecast error indeed grows rapidly and
+reaches its maximum in about 30 days for most months except for
+February, when it grows much slower. The ACCESS-S1 forecast errors grow
+slower than persistence forecast errors and remain lower after less than
+10 days on average. The ACCESS-S2 forecast error starts high in all
+months and is lower than the persistence forecast error after more than
+15 days in most months except for forecast initialised in February, when
+it takes 80 days.
 
@@
 
-Figure~\ref{fig-lead-time-window} summarises the lead time window in
-which each hindcast is better than both the persistence forecast and the
+Figure~\ref{fig-lead-time-window} (Figure
+\ref{fig-lead-time-window-osi}) summarises the lead time window in which
+each hindcast is better than both the persistence forecast and the
 climatological forecast as a function of forecast month. ACCESS-S1
@@
 
-\caption{\label{fig-rmse_lon-2}Same as Figure~\ref{fig-rmse_lon-1} but
+\caption{\label{fig-rmse_lon-2}Same as Figure \ref{fig-rmse_lon-1} but
 for ACCESS-S2.}
@@
 
-For ACCESS-S1 forecasts (Figure~\ref{fig-rmse_lon-1}), skill tends to be
-lower off the coast of Eastern Antarctica even at short lead times; for
-instance, the skill score for forecasts initialised in May and June are
-negative between 0° and 120°E even at almost zero lead time. This
-mirrors \citet{libera2022} findings of a ``winter predictability
-barrier'', although they focus on the Weddell Sea and here we show that
-the effect seems to be stronger more to the east. In West Antarctica
-there is a hint of easterly-propagating skill in forecasts initialised
-in February and March. This is consistent with \citet{holland2013}
-findings that memory of sea-ice anomalies are stored in ocean heat
-content anomalies that are transported east by the Antarctic Circumpolar
-Current.
-
-ACCESS-S2 forecasts (Figure~\ref{fig-rmse_lon-2}) also have lower skill
-over East Antarctica. From July to December even though the
-pan-Antarctic average skill is negative at all lead times
-(Fig.~\ref{fig-lead-time-window}), it is positive for up to a month in
-West Antarctica. Since oceanic and atmospheric forcing is the only
-source of information, this suggests that sea-ice in this region is
-particularly sensitive to oceanic and atmospheric forcing and suggests a
-role of the Pacific-South American mode and the Amundsen Sea Low to
-shape sea-ice concentration anomalies. The fact that this is evident in
-the months in which El Niño--Southern Oscillation teleconnections are
-more important for atmospheric circulation also suggests the influence
-of tropical Pacific variability. February and March are the only two
-months that can be forecasted with marginally positive skill in large
-regions.
+For ACCESS-S1 forecasts (Figure~\ref{fig-rmse_lon-1} and Figure
+\ref{fig-rmse_lon-1-osi}), skill tends to be lower off the coast of
+Eastern Antarctica even at short lead times; for instance, the skill
+score for forecasts initialised in May and June are negative between 0°
+and 120°E even at almost zero lead time. This mirrors \citet{libera2022}
+findings of a ``winter predictability barrier'', although they focus on
+the Weddell Sea and here we show that the effect seems to be stronger
+more to the east. In West Antarctica there is a hint of
+easterly-propagating skill in forecasts initialised in February and
+March. This is consistent with \citet{holland2013} findings that memory
+of sea-ice anomalies are stored in ocean heat content anomalies that are
+transported east by the Antarctic Circumpolar Current.
+
+ACCESS-S2 forecasts (Figure~\ref{fig-rmse_lon-2} and Figure
+\ref{fig-rmse_lon-2-osi}) also have lower skill over East Antarctica.
+From July to December even though the pan-Antarctic average skill is
+negative at all lead times (Fig.~\ref{fig-lead-time-window}), it is
+positive for up to a month in West Antarctica. Since oceanic and
+atmospheric forcing is the only source of information, this suggests
+that sea-ice in this region is particularly sensitive to oceanic and
+atmospheric forcing and suggests a role of the Pacific-South American
+mode and the Amundsen Sea Low to shape sea-ice concentration anomalies.
+The fact that this is evident in the months in which El Niño--Southern
+Oscillation teleconnections are more important for atmospheric
+circulation also suggests the influence of tropical Pacific variability.
+February and March are the only two months that can be forecasted with
+marginally positive skill in large regions.
 
@@
 
-\caption{\label{fig-rmse_lon-3}Same as Figure~\ref{fig-rmse_lon-1} but
+\caption{\label{fig-rmse_lon-3}Same as Figure \ref{fig-rmse_lon-1} but
 for the difference between ACCESS-S1 and ACCESS-S2.}
@@
 
-Finally, Figure~\ref{fig-rmse_lon-3} shows the difference in skill
-between ACCESS-S1 and ACCESS-S2. Large differences in skill indicate
-areas and months that are most affected by the data assimilation present
-in ACCESS-S1. Between January and March, which are the months in which
-ACCESS-S1 is the most skilful (Figure~\ref{fig-lead-time-window}), most
-of the improvement compared with ACCESS-S2 is present in the Ross and
-Weddell Sea. In April and May, the improvement seems more homogeneous.
+Finally, Figure~\ref{fig-rmse_lon-3} (Figure \ref{fig-rmse_lon-3-osi})
+shows the difference in skill between ACCESS-S1 and ACCESS-S2. Large
+differences in skill indicate areas and months that are most affected by
+the data assimilation present in ACCESS-S1. Between January and March,
+which are the months in which ACCESS-S1 is the most skilful
+(Figure~\ref{fig-lead-time-window}), most of the improvement compared
+with ACCESS-S2 is present in the Ross and Weddell Sea. In April and May,
+the improvement seems more homogeneous.
 
@@
 In Figure~\ref{fig-rmse} the mean error was shown.
-Figure~\ref{fig-initial-spread} column 1 shows the mean standard
-deviation of errors among ensemble members at various lead times. At one
-day lead time (Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a
-slightly larger spread than ACCESS-S1 due to the way that ensemble
-members are generated. ACCESS-S1 ensemble members are generated by
-adding random field perturbations to the atmosphere only, which then are
-transferred to the other components via the coupled simulation
-\citep{hudson2017}. With this scheme, ensemble members are all but
-guaranteed to be underdispersed in the ocean and sea-ice components. The
-time-lag ensemble used for ACCESS-S2 ensures greater spread. This
-difference is gone after about just two days, and both systems have a
-comparable spread in ensemble member error afterwards
+Figure~\ref{fig-initial-spread} (Figure \ref{fig-initial-spread-osi})
+column 1 shows the mean standard deviation of errors among ensemble
+members at various lead times. At one day lead time
+(Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a slightly larger
+spread than ACCESS-S1 due to the way that ensemble members are
+generated. ACCESS-S1 ensemble members are generated by adding random
+field perturbations to the atmosphere only, which then are transferred
+to the other components via the coupled simulation \citep{hudson2017}.
+With this scheme, ensemble members are all but guaranteed to be
+underdispersed in the ocean and sea-ice components. The time-lag
+ensemble used for ACCESS-S2 ensures greater spread. This difference is
+gone after about just two days, and both systems have a comparable
+spread in ensemble member error afterwards
 (Fig.~\ref{fig-initial-spread} b1 and c1).
@@
 
+\section{Funding}\label{funding}
+
+This work was supported by ARC SRIEAS Grant SR200100005 Securing
+Antarctica's Environmental Future.
+
 \section{Acknowledgments}\label{acknowledgments}
@@
 capability supported by the Australian Government.
-
-This work was supported by ARC SRIEAS Grant SR200100005 Securing
-Antarctica's Environmental Future.
 
@@
 
-\clearpage
-
-\appendix  
-\renewcommand{\thefigure}{A\arabic{figure}}
-\setcounter{figure}{0}
-
-\section{Supplementary figures}\label{supplementary-figures}
-
-The following are the same figures from the main paper but using the OSI
-dataset instead of CDR.
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-hindcast-extent-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-hindcast-extent-osi}Row a: Pan-Antarctic daily mean
-sea-ice extent for all hindcasts initialised on the first of each
-calendar month for ACCESS-S1 (column 1; green) and ACCESS-S2 (column 2;
-purple). Observed mean sea-ice extent in each corresponding hindcast
-period is shown in black. Row b: Mean differences between the forecast
-and the observed values. Circles represent the initial conditions at the
-start of forecasts (i.e., the first of every month), and triangles
-represent the mean values at the lead time corresponding to the maximum
-lead time in S1 (between 213 and 216 days, depending on the month)}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-growth-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-mean-growth-osi}Mean daily sea-ice extent growth
-(\(10^6 km^2/day\)) in ACCESS-S1 (green) and ACCESS-S2 (purple)
-hindcasts and observations (black), computed as the mean daily
-differences in sea-ice extent between each date and the next for each
-forecast month. Values are smoothed with a 11-day running mean.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-bias-1-osi}Ensemble mean difference between monthly
-sea-ice concentration of ACCESS-S2 ensemble mean forecast at 0-month
-lead time (monthly mean values forecasted from the forecast initialised
-at the first of the month) and observations (OSI).}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-2-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-bias-2-osi}Same as Figure~\ref{fig-bias-1} but for
-ACCESS-S1.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-anom-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-extent-anom-osi}Monthly mean sea-ice extent
-anomalies of the observations (black) and forecasts from ACCESS-S1
-(right column; purple) and ACCESS-S2 (left column; green) at lead times
-of 0, 2, 4, and 6 months. The RMSE and correlation during the
-overlapping period of ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) are
-shown on the top left and bottom left of each panel respectively.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-sd-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-extent-sd-osi}Interannual standard deviation with
-95\% confidence interval of monthly mean sea-ice extent forecasted for
-each month divided by that month's sea-ice extent observation standard
-deviation. ACCESS-S1 and ACCESS-S2 at different lead times. Each panel
-indicates the target month. Note the reverse horizontal axis.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-rmse-osi}Mean RMSE of sea-ice concentration
-anomalies as a function of forecast lead time for all forecasts
-initialised on the first of each month compared with a reference
-forecast of persistence of anomalies (black) and climatology (gray).
-Only the first 120 days are shown. In parentheses, the shortest time at
-which ACCESS-S1 and ACCESS-S2 mean RMSE is not statistically different
-at the 99\% confidence level.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-lead-time-window-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-lead-time-window-osi}Minimum lead time at which each
-forecast's mean RMSE becomes larger than the lower bound of the 95\%
-confidence interval of persistence forecast RMSE (black lines) and
-maximum lead time at which each forecast's mean RMSE remains lower than
-the lower bound of the 95\% confidence interval of climatological
-forecast RMSE (gray lines). Green shading indicates the window where
-forecasts outperform both persistence (lead times longer than black
-line) and climatology (lead times shorter than gray line). Text labels
-show the date corresponding to the maximum lead time at which each
-forecast outperforms climatology.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-osi-1.png}}
-
-}
-
-\caption{\label{fig-rmse_lon-1-osi}RMSE skill score of ACCESS-S1
-forecasts with climatological forecast as reference computed on 15
-meridional slices 24° wide as a function of lead time and longitude.
-Antarctica's coastline is shown at the bottom of each panel for
-reference.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-osi-1.png}}
-
-}
-
-\caption{\label{fig-rmse_lon-2-osi}Same as Figure~\ref{fig-rmse_lon-1}
-but for ACCESS-S2.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-3-osi-1.png}}
-
-}
-
-\caption{\label{fig-rmse_lon-3-osi}Same as Figure~\ref{fig-rmse_lon-1}
-but for the difference between ACCESS-S1 and ACCESS-S2.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-initial-spread-osi-1.pdf}}
-
-}
-
-\caption{\label{fig-initial-spread-osi}Decomposition of forecast error
-spread at 1, 5 and 30 days lead time for ACCESS-S1 and ACCESS-S2
-hindcasts across initialization months. The left column shows the mean
-standard deviation of RMSE errors across ensemble members, while the
-right column shows the standard deviation of the ensemble mean RMSE
-error and the spread of the persistence and climatology forecasts
-errors.}
-
-\end{figure}%
-
 \end{document}

### 2025-11-27 analysis/paper/supp.tex
Render

--- 2025-11-19 analysis/paper/access-ice.tex
@@
 \usepackage[section]{placeins}
+\usepackage{xr}
+\externaldocument{ref_main}
 \makeatletter
@@
 Victoria, Australia}}
-\affil[3]{\orgdiv{Research, Bureau of Meteorology, Australia}}
+\affil[3]{\orgdiv{Research, Bureau of Meteorology, Melbourne,
+Australia}}
 
 % abstract
-
-\abstract{Accurate Antarctic sea-ice forecasts are crucial for climate
-monitoring and operational planning, yet they remain challenging due to
-model biases and complex ice-ocean-atmosphere interactions. The two
-versions of the Australian Bureau of Meteorology's ACCESS seasonal
-forecast system, ACCESS-S1 and ACCESS-S2, use identical model
-configuration and differ only in their initial conditions; primarily in
-that ACCESS-S2 does not assimilate sea-ice observations, whereas
-ACCESS-S1 does.\\
-This provides a convenient opportunistic experiment to assess the role
-of initial conditions on Antarctic sea-ice forecasts using more than 20
-years of fully coupled simulations with two 9-member ensembles. Our
-analysis reveals that both systems experience an extended melt season
-and delayed growth phase compared with observations. This leads to a
-significant negative sea-ice extent bias, which is corrected only in
-ACCESS-S1 by the data assimilation system. The impact of the differing
-initial conditions on forecast errors varies dramatically by season:
-summer and autumn initial conditions (January-April) provide predictive
-skill for up to three months, with February initial conditions being
-particularly crucial. In contrast, winter forecasts of the two systems
-are statistically indistinguishable after only two weeks. Regional
-analysis of forecast skill suggests that this winter predictability
-barrier is most dramatic over East Antarctica, where even ACCESS-S1
-shows negative skill. These findings highlight the critical importance
-of comprehensive year-round sampling in predictability studies and
-suggest that operational sea-ice data assimilation efforts should
-prioritize the summer-autumn period when initial conditions have maximum
-impact on forecast skill.}
 
@@
 \maketitle
-
-\section{Introduction}\label{introduction}
-
-Accurately modelling Antarctic sea ice is essential for understanding
-processes and improving climate projections to inform adaptation
-strategies. Accurate seasonal to sub-seasonal forecasts are also crucial
-for operation contingency planning in and around the Antarctic
-continent, including scientific missions, fisheries, and tourism
-\citep{desilva2020, wagner2020}. Improvements in modelled sea-ice might
-also help improve weather forecasts over and away from sea-ice regions
-\citep{rinke2006, wang2024, semmler2016}.
-
-However, progress in Antarctic sea-ice forecasting system has lagged
-behind Arctic sea-ice forecasts due to model biases, and inherent large
-variability and complexity \citep{zampieri2019, gao2024a}. Dynamical
-seasonal forecasts of summer Antarctic sea ice have been shown to
-perform worse than relatively simpler statistical methods
-\citep{massonnet2023} and machine learning approaches (e.g.
-\citet{dong2024}, \citet{lin2025}), which also underscores the need for
-better understanding and physical modelling of sea-ice dynamics, and
-drivers of its variability.
-
-Good initial conditions are generally required for a good forecast,
-however, it is not entirely known to what extent accurate sea-ice
-initial conditions affect the quality of the forecast and at what
-timescales. Exploring seasonal predictions of Arctic sea ice,
-\citet{guemas2016} found that sea-ice initial conditions are important
-in autumn to predict summer sea ice, but the impact wasn't as dramatic
-when predicting winter sea ice. \citet{day2014} also found
-seasonally-varying differences in the effect of initialisation, noting
-that accurate Arctic sea-ice thickness leads to improved sea-ice
-forecasts initialised in July but not when initialised in January.
-
-For the Antarctic, \citet{holland2013} studied the initial-value
-predictability of Antarctic sea ice in a perfect model study using the
-CCSM3 model. They found that sea-ice and ocean initial conditions
-provide predictive information to forecast sea-ice edge location several
-months in advance and that some predictability is retained for up to two
-years thanks to ocean heat content anomalies that are advected eastward.
-This is in contrast with \citet{marchi2020}, who ran perfect model
-experiments to argue that uncertainty in the predicted atmospheric state
-and evolution is the main driver of uncertainty in Antarctic sea-ice
-extent prediction on seasonal timescales, with sea-ice and ocean initial
-conditions having lesser importance. More recently, \citet{morioka2022}
-studied decadal forecasts of Antarctic sea ice and found that
-initialising ocean and sea ice improved the correlation between
-simulated and observed sea-ice concentration evolution in the
-Amundsen--Bellingshausen Sea. It is hard to compare these studies since
-they are based on forecasts initialised at different times of the year
-and different frameworks: \citet{holland2013} ran 20 ensemble members
-initialised on the 1st of January of a particular year,
-\citet{marchi2020} ran forecasts from the 1st of March and 1st of
-September, and \citet{morioka2022} ran forecasts only from the 1st of
-March. \citet{marchi2020} also used a coupled ocean--sea-ice model
-instead of a fully coupled model like \citet{holland2013} did.
-\citet{morioka2022} used observed sea-ice initial conditions and
-compared with observations, while \citet{marchi2020} and
-\citet{holland2013} were perfect model studies.
-
-In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the
-Australian Community Climate and Earth System Simulator -- Seasonal
-(ACCESS-S) from version S1 to S2. While the base model remained the
-same, the change in version was focused on using ocean, sea-ice and land
-initial conditions generated by the BoM instead of depending on the UK
-Met Office. Crucially, compared to ACCESS-S1, ACCESS-S2 does not
-assimilate sea-ice observations, so sea ice is only affected by the
-ocean and atmospheric data assimilation via the coupled integration.
-
-Since model configuration is identical between ACCESS-S1 and ACCESS-S2,
-they form a sort of ``opportunistic experiment'' where the same
-forecasting model was run over a long period of time with multiple
-ensemble forecasts initialised throughout the year, with the only
-difference being the initial conditions. This provides an opportunity to
-test the effect of sea-ice initial conditions on the forecast of sea-ice
-concentrations and the climate.
-
-In this study we compare sea-ice hindcasts produced by ACCESS-S1 and
-ACCESS-S2. We focus on seasonality of errors and biases and the effect
-of the data assimilation system. This comparison will inform future work
-with the prediction system as a research tool to better understand the
-dynamics and variability of the Antarctic sea ice and its impacts on the
-climate system as well as to explore the potential of using its sea-ice
-forecasts for decision-making. The work will also serve as a benchmark
-for future prediction systems to attempt to improve upon.
-
-\section{Data and methods}\label{data-and-methods}
-
-\subsection{ACCESS-S2}\label{s2}
-
-ACCESS-S2 \citep{wedd2022} is the Bureau of Meteorology's seasonal
-forecast system which became operational in October 2021, replacing the
-ACCESS-S1 system \citep{hudson2017}. The model components of both
-ACCESS-S2 and ACCESS-S1 are identical with the same numbers of levels
-and resolution. They consist of the Global Atmosphere 6.0 (GA6)
-\citep{williams2015, waters2017}, the Unified Model's Global Land 6.0
-\citep{best2011, waters2017}, NEMO Global Ocean 5.0
-\citep{gurvan2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
-\citet{rae2015}{]}. The atmosphere has a N216 horizontal resolution
-(\textasciitilde60 km in the mid-latitudes) with 85 vertical levels. The
-land model uses the same horizontal grid as the atmosphere with four
-soil levels. The ocean component has a nominal horizontal resolution of
-1/4° with 75 vertical levels. The sea-ice component, based on CICE
-version 4.1, has the same resolution as the ocean component and five
-sea-ice thickness categories as well as an open water category.
-
-Both systems take atmospheric initial conditions derived from
-ERA-interim \citep{dee2011} for their hindcasts. The main difference
-between the hindcasts of the two systems are the ocean and sea-ice
-initial conditions. ACCESS-S1's ocean and sea-ice initial conditions
-come from the Met Office FOAM system, which uses a multivariate,
-incremental three-dimensional variational (3D-Var),
-first-guess-at-appropriate-time (FGAT) data assimilation scheme
-\citep{waters2015} and assimilates sea surface temperature (SST), sea
-surface height (SSH), in situ temperature and salinity profiles, and
-satellite observations of sea-ice concentration using the EUMETSAT
-OSISAF product described in the next section. ACCESS-S2, on the other
-hand, is initialised from ocean conditions generated by the BoM weakly
-coupled ensemble data assimilation scheme described in \citet{wedd2022}.
-This scheme uses an optimal interpolation method and assimilates
-temperature and salinity profiles from EN4 \citep{good2013}. SSTs are
-nudged to Reynolds OISSTv2.1 \citep{reynolds2007} in areas where SSTs
-are over 0°C and Sea Surface Salinity is weakly nudged to the World
-Ocean Atlas 2013 climatology \citep{zweng2013}.
-
-Of most relevance for this work, sea-ice concentrations are not
-assimilated in ACCESS-S2. Assimilation cycles are performed daily. The
-coupled model runs for 24 hours initialised from the previous cycle.
-Then the restart file fields of the ocean component are used as first
-guess in the data assimilation cycle and the innovations are used to
-build the next ocean initial conditions for the following cycle. The
-atmosphere fields from that daily integration are not used and instead
-the model atmosphere is initialised using ERA-Interim. The sea-ice
-initial conditions for the next cycle are the unaltered output of the
-previous daily integration. Then the cycle starts again and the coupled
-model runs for another 24 hours. During this integration the sea-ice
-component is affected by the ocean innovations and the new atmosphere
-initial conditions via the coupler.
-
-The ACCESS-S1 hindcast set is made up of nine members created by
-perturbing the atmospheric fields only with a random field perturbation
-\citep{hudson2017} and runs for 217 days for the period 1990--2012
-initialised at the first of every month. The ACCESS-S2 hindcast set used
-in this study runs for the period 1981--2018. Ensemble members are
-created in the same manner as ACCESS-S1 members, however, due to
-computing cost limitations, only three members per forecast
-initialisation date were run for 279 days. Bigger ensembles were
-generated by aggregating several three-member ensembles initialised on
-successive days \citep{wedd2022}. Here, we build a nine-member
-time-lagged ensemble from three consecutive three-member forecasts
-initialised at the first of every month and the two previous days and
-run for 279 days. We analyse the ensemble mean hindcasts unless
-otherwise specified.
-
-Anomalies for each hindcast set are taken with respect to their own
-climatology specific to each initialisation date and forecast lead time,
-for the period 1990--2012. This serves as a first-order correction of
-model bias and drift. For monthly means, we define ``0 lead time
-months'' as the monthly mean forecast of the same month of
-initialisation.
-
-Besides sea-ice concentration, we also analyse mean sea-ice thickness,
-which we compute as total sea-ice volume divided by total sea-ice area.
-
-\subsection{Verification datasets}\label{verification-datasets}
-
-For verification we use satellite-derived sea-ice concentration, which
-estimates the proportion of each grid area that is covered with ice.
-Datasets derived using different algorithms and satellite platforms,
-each have their own biases and uncertainties. Estimates of inter-product
-uncertainty of sea-ice extent (SIE, defined here as the total region of
-the Southern Ocean with at least 15\% sea-ice cover) are of the order of
-0.5 million \(km^2\) \citep{meier2019}. As will be shown below, this
-spread is minimal compared with the typical errors in the ACCESS-S2 and
-ACCESS-S1 forecasts, so the overall conclusions of this study are
-independent of the verification dataset used.
-
-We use NOAA/NSIDC's Climate Data Record V4 {[}CDR; \citet{meier2014}{]}
-as the primary sea-ice verification dataset. It takes the maximum value
-of the NASA Team \citep{cavalieri1984} and NASA Bootstrap
-\citep{comiso2023} sea-ice concentration products to reduce their low
-concentration bias \citep{meier2014, meier2021}. Both source algorithms
-use data from the Scanning Multichannel Microwave Radiometer (SMMR) on
-the Nimbus-7 satellite and from the Special Sensor Microwave/Imager
-(SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP)
--F8, -F11, and -F13 satellites. The data have a spatial resolution of 25
-by 25 km and daily from November 1978 onwards.
-
-The European Organisation for the Exploitation of Meteorological
-Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility
-{[}OSI; \citet{OSISAF}{]} based on the SSMIS sensor is another
-satellite-derived sea-ice concentration product. It is based on mostly
-the same sensors as the NOAA CDR but computed independently using
-different algorithms. Figures prepared with this dataset are provided in
-the supplementary material and do not differ significantly from the ones
-prepared using CDR.
-
-\subsection{Error measures}\label{error-measures}
-
-For evaluation purposes, we use a series of measures. Sea-ice extent is
-defined as the area of the ocean at least 15\% covered by sea-ice. This
-threshold is motivated by the limitations in satellite retrieval, which
-is increasingly unreliable for lower sea-ice concentrations
-\citep{cavalieri1991}.
-
-Pan-Antarctic (net) sea-ice extent serves as a hemispheric measure of
-the amount of sea ice, but it does not take into account the spatial
-distribution. A model could have a relatively accurate extent of the net
-ice but with different regional distributions. To account for location
-errors, we computed the Root Mean Squared Error (RMSE) of grid-point
-sea-ice concentration anomalies.
-
-We compute RMSE as the square root of the area-averaged squared
-differences between grid-point forecasted and observed sea-ice
-concentration anomalies. We compute a pan-Antarctic RMSE by averaging
-over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a
-zonally-varying RMSE computed over 15 longitude slices 24° wide around
-Antarctica.
-
-All error measures were computed on the NOAA/NSIDC CDRV4 domain grid, to
-which model output was bilinearly interpolated. Note that the ACCESS
-CICE model grid has resolution between two and three times higher than
-NOAA/NSIDC CDRV4.
-
-Forecast errors are also compared with hypothetical forecasts based on
-the persistence of anomalies and on climatology. The persistence
-forecast is generated by extending the observed sea-ice concentration
-anomalies the day of the forecast initialisation and comparing it with
-the actual anomalies observed. The climatological forecast error is
-computed as the standard deviation of daily anomalies.
-
-As a measure of forecast improvement over the hypothetical forecast, we
-use the skill score \citep{murphy1985}, defined as
-
-\[
-S = 1 - \frac{RMSE_{f}}{RMSE_{r}}
-\]
-
-Where \(RMSE_{f}\) is the RMSE of the forecast, \(RMSE_{r}\) is the RMSE
-of the reference forecast. Negative skill score indicates that the
-forecast is worse than the reference forecast while positive values
-indicate an improvement. A perfect forecast would have zero RMSE and
-thus a skill score of 1.
-
-\subsection{Computational procedures}\label{computational-procedures}
-
-We performed all analyses in this paper using the R programming language
-\citep{rcoreteam2020}, using data.table \citep{dowle2020} and metR
-\citep{campitelli2020b} packages. Significant processing was performed
-using the CDO command line operators \citep{schulzweida2023}. All
-graphics are made using ggplot2 \citep{wickham2009}. The paper was
-rendered using knitr and Quarto \citep{xie2015, allaire2022}.
-
-\section{Results and discussion}\label{results-and-discussion}
-
-\subsection{Bias}\label{bias}
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}}
-
-}
-
-\caption{\label{fig-hindcast-extent}Row a: Pan-Antarctic daily mean
-sea-ice extent for all hindcasts initialised on the first of each
-calendar month for ACCESS-S1 (column 1; green) and ACCESS-S2 (column 2;
-purple). Observed mean sea-ice extent in each corresponding hindcast
-period is shown in black. Row b: Mean differences between the forecast
-and the observed values. Circles represent the initial conditions at the
-start of forecasts (i.e., the first of every month), and triangles
-represent the mean values at the lead time corresponding to the maximum
-lead time in S1 (between 213 and 216 days, depending on the month)}
-
-\end{figure}%
-
-Figure~\ref{fig-hindcast-extent} shows mean sea-ice extent of the
-ACCESS-S1 and ACCESS-S2 hindcasts (row a) and their differences from
-mean sea-ice extent of NOAA/NSIDC CDRV4 (row b). Mean extent at the
-first of every month is indicated with circles for the initial
-conditions and with triangles for the longest lead time possible for
-each model (between 274 and 277 days for ACCESS-S2 and between 213 and
-216 days for ACCESS-S1). At this long lead time, information of the
-initial conditions is essentially lost and the forecast reverts close to
-each model's preferred equilibrium state.
-
-ACCESS-S2 initial conditions (circles in Fig.~\ref{fig-hindcast-extent}
-column 2) show an overall negative bias, especially in the late
-summer-early autumn, while ACCESS-S1 initial conditions (circles in
-Fig.~\ref{fig-hindcast-extent} column 1) are very close to observations,
-as expected from the assimilation of sea-ice observations to produce the
-initial conditions of ACCESS-S1. Both systems' equilibrium states
-(triangles) show negative biases of sea-ice extent, particularly in the
-growth phase of late-autumn and winter months. This is due primarily to
-the melt season being longer than in observations and with faster melt
-between January and March and the growing seasons being shorter with
-slower growth during March and April. This is then followed by faster
-growth between May and July (Figure~\ref{fig-mean-growth}). Many sea-ice
-models exhibit this systematic underestimation during the sea-ice
-minimum and early freezing season \citep{massonnet2023}, which could
-indicate problems in the representation of thermodynamics in the model
-\citep{zampieri2019}. It is also not surprising that both forecasting
-systems converge to a similar equilibrium state because they share the
-same model formulation.
-
-The difference between the initial conditions (circles) and the model
-equilibrium state (triangles) can be mostly attributed to the effect of
-data assimilation, which in ACCESS-S2 is due solely to the coupling of
-sea-ice with the atmosphere and the ocean. From May to October, in
-ACCESS-S2 circles are closer to observations than the triangles are,
-indicating that the information from the ocean and atmosphere data
-assimilation is affecting sea ice and improving the initial conditions.
-During these months, ACCESS-S1 can overestimate the sea-ice extent at
-short lead time. For the rest of the year circles are overlaid with
-triangles in ACCESS-S2, indicating that the ocean and atmosphere data
-assimilation is not affecting sea ice and that this component of the
-model is virtually free-running.
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}}
-
-}
-
-\caption{\label{fig-mean-growth}Mean daily sea-ice extent growth
-(\(10^6 km^2/day\)) in ACCESS-S1 (green) and ACCESS-S2 (purple)
-hindcasts and observations (black), computed as the mean daily
-differences in sea-ice extent between each date and the next for each
-forecast month. Values are smoothed with a 11-day running mean.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1-1.pdf}}
-
-}
-
-\caption{\label{fig-bias-1}Ensemble mean difference between monthly
-sea-ice concentration of ACCESS-S2 ensemble mean forecast at 0-month
-lead time (monthly mean values forecasted from the forecast initialised
-at the first of the month) and observations (CDR).}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-2-1.pdf}}
-
-}
-
-\caption{\label{fig-bias-2}Same as Figure~\ref{fig-bias-1} but for
-ACCESS-S1.}
-
-\end{figure}%
-
-To further understand the bias in ACCESS-S2, Figure~\ref{fig-bias-1}
-shows spatial patterns of the differences of monthly mean sea-ice
-concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
-shortest monthly lead time. From October to May, the model
-underestimates sea-ice concentrations in most regions except for the
-inner Weddell Sea in April and May, where sea-ice concentrations
-saturate to 1 both in the observations and forecasts. In winter, the
-differences are limited to a narrow band around the sea-ice edge with
-slight positive biases in the African sector of East Antarctica and
-negative biases around the Indian Ocean sector which partially
-compensate, resulting in the near-zero extent bias seen in those months
-(Figure~\ref{fig-hindcast-extent}).
-
-ACCESS-S1 has a comparatively smaller overall bias
-(Figure~\ref{fig-bias-2}). The largest values are found between April
-and June, when the faster growth results in large positive bias along
-the sea-ice edge, and in January, when the faster melt leads to large
-negative bias in the Weddell and Amundsen Seas.
-
-\FloatBarrier
-
-\subsection{RMSE}\label{rmse}
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}}
-
-}
-
-\caption{\label{fig-extent-anom}Monthly mean sea-ice extent anomalies of
-the observations (black) and forecasts from ACCESS-S1 (right column;
-purple) and ACCESS-S2 (left column; green) at lead times of 0, 2, 4, and
-6 months. The RMSE and correlation during the overlapping period of
-ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) are shown on the top left
-and bottom left of each panel respectively.}
-
-\end{figure}%
-
-Figure~\ref{fig-extent-anom} shows monthly sea-ice extent anomalies
-forecasted at selected lead times. Compared with ACCESS-S1, ACCESS-S2
-anomaly forecasts are relatively poor (large RMSE) even for the first
-month (lead time 0), whereas ACCESS-S1 forecasts stay relatively skilful
-even at a lead time of three months. ACCESS-S2 shows much larger
-interannual variability than observations, with dramatic lows between
-1995 and 2007, and highs between 2007 and 2015.
-
-Unexpectedly, for ACCESS-S2, RMSE improves with lead time, even though
-the correlation degrades with lead time. This is puzzling behaviour that
-goes contrary to what is usually seen in prediction models. The
-explanation seems to be the mentioned increased interannual variability.
-Figure~\ref{fig-extent-sd} shows the interannual standard deviation of
-monthly sea-ice extent of the forecasts as a function of lead time
-compared with observations. ACCESS-S1 standard deviation lies within the
-observed standard deviation regardless of lead time, while ACCESS-S2
-standard deviation is more than twice that of observations at zero lead
-time and only approaches the observed value at nine month lead time for
-most months.
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-sd-1.pdf}}
-
-}
-
-\caption{\label{fig-extent-sd}Interannual standard deviation with 95\%
-confidence interval of monthly mean sea-ice extent forecasted for each
-month divided by that month's sea-ice extent observation standard
-deviation. ACCESS-S1 and ACCESS-S2 at different lead times. Each panel
-indicates the target month. Note the reverse horizontal axis.}
-
-\end{figure}%
-
-ACCESS-S2 forecasts of sea-ice extent anomalies seem to align moderately
-well with observations (leading to moderately high correlation) but
-their magnitude is overestimated (leading to large errors). This could
-be caused by ACCESS-S2 sea ice being much more sensitive to atmospheric
-and oceanic forcing, perhaps due to lower thickness.
-
-As an example, Figure~\ref{fig-forecast-example} shows sea-ice
-concentration anomalies (top row) and sea-ice thickness and the
-difference between the two models (bottom row) for 2 May 2008
-initialised one day prior; being that close to initialisation date,
-these are very approximately the initial conditions. ACCESS-S1 sea-ice
-concentrations anomalies are very close to observations as expected from
-the system assimilating these data. ACCESS-S2 sea-ice concentration
-anomalies, which are not assimilated, are not as close, but the
-large-scale pattern is aligned with observations. The system simulates
-large positive anomalies in the Weddell and Ross Seas and slight
-negative anomalies in the Amundsen and Bellingshausen Seas. The fact
-that ACCESS-S2 can simulate this pattern without assimilating sea-ice
-data suggests that atmospheric and oceanic forcing were the dominant
-drivers. However, the magnitude of the sea-ice anomalies is too big. It
-is plausible that this is due to the thinner ice simulated by ACCESS-S2
-(bottom row).
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-forecast-example-1.pdf}}
-
-}
-
-\caption{\label{fig-forecast-example}ACCESS-S1 and ACCESS-S2 hindcasts
-for 2 May 2008 at one day lead time. Top row shows sea-ice concentration
-anomalies forecasted by each system and the observations. Bottom row
-shows forecasted sea-ice thickness and the difference between ACCESS-S1
-and ACCESS-S2.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-thickness-1.pdf}}
-
-}
-
-\caption{\label{fig-mean-thickness}Mean and 95\% interval of monthly
-mean sea-ice thickness for ACCESS-S1 and ACCESS-S2 at different lead
-times. Each panel indicates the target month. Note the reverse
-horizontal axis.}
-
-\end{figure}%
-
-Extending beyond the one case in Figure~\ref{fig-forecast-example},
-Figure~\ref{fig-mean-thickness} shows monthly mean sea-ice thickness as
-a function of lead time for ACCESS-S1 and ACCESS-S2. Supporting the idea
-that thinner ice is what causes the increased extent variability in
-ACCESS-S2, this system simulates thinner sea-ice compared to ACCESS-S1
-overall at almost all lead times and in all months except for summer at
-short lead times (Dec-Jan, 0-1 months; Feb-Mar, 0-2 months). However, in
-both systems, forecasted sea-ice is thicker at shorter lead times and
-then decreases, particularly in the summer months. If thinner ice were a
-sufficient cause of increased variability, then we would expect
-variability to increase with lead time in both forecasting systems.
-
-The fact that ACCESS-S1 and ACCESS-S2 share the same model configuration
-and that the increased variability is more extreme at short lead times
-(Fig.~\ref{fig-extent-sd}) suggests that the data assimilation procedure
-is partly responsible. It is possible that sea-ice in the ACCESS-S2
-system is left in an unbalanced state after assimilating atmospheric and
-oceanic data but not sea-ice data, leading to large responses that are
-amplified by the thin ice in the initial states which then subside at
-longer lead times when the model is balanced.
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse-1.pdf}}
-
-}
-
-\caption{\label{fig-rmse}Mean RMSE of sea-ice concentration anomalies as
-a function of forecast lead time for all forecasts initialised on the
-first of each month compared with a reference forecast of persistence of
-anomalies (black) and climatology (gray). Only the first 120 days are
-shown. In parentheses, the shortest time at which ACCESS-S1 and
-ACCESS-S2 mean RMSE is not statistically different at the 99\%
-confidence level.}
-
-\end{figure}%
-
-To assess ACCESS-S2 forecasts in more detail, we compute error measures
-for all hindcasts started on the 1st of every month.
-Figure~\ref{fig-rmse} shows the mean RMSE of sea-ice concentration
-anomalies for ACCESS-S1 and ACCESS-S2 hindcasts compared against
-persistence and climatological forecasts used as a benchmark. Due to
-errors in the initial conditions, it is expected that persistence
-forecasts would be better than the model forecasts at very short lead
-times, but that the persistence forecast errors would grow faster and
-may eventually surpass the model forecast errors. The black line shows
-that the persistence forecast error indeed grows rapidly and reaches its
-maximum in about 30 days for most months except for February, when it
-grows much slower. The ACCESS-S1 forecast errors grow slower than
-persistence forecast errors and remain lower after less than 10 days on
-average. The ACCESS-S2 forecast error starts high in all months and is
-lower than the persistence forecast error after more than 15 days in
-most months except for forecast initialised in February, when it takes
-80 days.
-
-At longer lead times, it is more appropriate to compare errors with the
-climatological forecast error. The lead time at which ACCESS-S1 forecast
-error is higher than the climatological forecast error varies between
-more than 60 and less than 20 days depending on forecast initialisation
-month with the minimum in June. ACCESS-S2 forecasts never have lower
-error than climatology, on the other hand, except marginally in October
-forecasts.
-
-Figure~\ref{fig-lead-time-window} summarises the lead time window in
-which each hindcast is better than both the persistence forecast and the
-climatological forecast as a function of forecast month. ACCESS-S1
-forecasts have a wider lead time window in the summer than the other
-seasons and is not better than both benchmarks at forecasting June
-sea-ice concentration anomalies. Forecasts initialised in May and June
-are particularly poor, and July cannot be forecasted better than the
-benchmarks. This is consistent with the mid-winter loss of
-predictability observed by \citet{libera2022}, who attributed it to deep
-warm water entraining into the mixed layer.
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-lead-time-window-1.pdf}}
-
-}
-
-\caption{\label{fig-lead-time-window}Minimum lead time at which each
-forecast's mean RMSE becomes larger than the lower bound of the 95\%
-confidence interval of persistence forecast RMSE (black lines) and
-maximum lead time at which each forecast's mean RMSE remains lower than
-the lower bound of the 95\% confidence interval of climatological
-forecast RMSE (gray lines). Green shading indicates the window where
-forecasts outperform both persistence (lead times longer than black
-line) and climatology (lead times shorter than gray line). Text labels
-show the date corresponding to the maximum lead time at which each
-forecast outperforms climatology.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-1.png}}
-
-}
-
-\caption{\label{fig-rmse_lon-1}RMSE skill score of ACCESS-S1 forecasts
-with climatological forecast as reference computed on 15 meridional
-slices 24° wide as a function of lead time and longitude. Antarctica's
-coastline is shown at the bottom of each panel for reference.}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-1.png}}
-
-}
-
-\caption{\label{fig-rmse_lon-2}Same as Figure~\ref{fig-rmse_lon-1} but
-for ACCESS-S2.}
-
-\end{figure}%
-
-To analyse the spatial distribution of the model error, we computed the
-RMSE of zonal mean sea-ice concentration anomalies on 15 slices of 24°
-longitude span for each forecasting system. We control for some areas
-being naturally easier to forecast than others by computing the RMSE
-skill score with the climatological forecast RMSE as reference.
-
-For ACCESS-S1 forecasts (Figure~\ref{fig-rmse_lon-1}), skill tends to be
-lower off the coast of Eastern Antarctica even at short lead times; for
-instance, the skill score for forecasts initialised in May and June are
-negative between 0° and 120°E even at almost zero lead time. This
-mirrors \citet{libera2022} findings of a ``winter predictability
-barrier'', although they focus on the Weddell Sea and here we show that
-the effect seems to be stronger more to the east. In West Antarctica
-there is a hint of easterly-propagating skill in forecasts initialised
-in February and March. This is consistent with \citet{holland2013}
-findings that memory of sea-ice anomalies are stored in ocean heat
-content anomalies that are transported east by the Antarctic Circumpolar
-Current.
-
-ACCESS-S2 forecasts (Figure~\ref{fig-rmse_lon-2}) also have lower skill
-over East Antarctica. From July to December even though the
-pan-Antarctic average skill is negative at all lead times
-(Fig.~\ref{fig-lead-time-window}), it is positive for up to a month in
-West Antarctica. Since oceanic and atmospheric forcing is the only
-source of information, this suggests that sea-ice in this region is
-particularly sensitive to oceanic and atmospheric forcing and suggests a
-role of the Pacific-South American mode and the Amundsen Sea Low to
-shape sea-ice concentration anomalies. The fact that this is evident in
-the months in which El Niño--Southern Oscillation teleconnections are
-more important for atmospheric circulation also suggests the influence
-of tropical Pacific variability. February and March are the only two
-months that can be forecasted with marginally positive skill in large
-regions.
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-3-1.png}}
-
-}
-
-\caption{\label{fig-rmse_lon-3}Same as Figure~\ref{fig-rmse_lon-1} but
-for the difference between ACCESS-S1 and ACCESS-S2.}
-
-\end{figure}%
-
-Finally, Figure~\ref{fig-rmse_lon-3} shows the difference in skill
-between ACCESS-S1 and ACCESS-S2. Large differences in skill indicate
-areas and months that are most affected by the data assimilation present
-in ACCESS-S1. Between January and March, which are the months in which
-ACCESS-S1 is the most skilful (Figure~\ref{fig-lead-time-window}), most
-of the improvement compared with ACCESS-S2 is present in the Ross and
-Weddell Sea. In April and May, the improvement seems more homogeneous.
-
-\begin{figure}[!htb]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-initial-spread-1.pdf}}
-
-}
-
-\caption{\label{fig-initial-spread}Decomposition of forecast error
-spread at 1, 5 and 30 days lead time for ACCESS-S1 and ACCESS-S2
-hindcasts across initialization months. The left column shows the mean
-standard deviation of RMSE errors across ensemble members, while the
-right column shows the standard deviation of the ensemble mean RMSE
-error and the spread of the persistence and climatology forecasts
-errors.}
-
-\end{figure}%
-
-In Figure~\ref{fig-rmse} the mean error was shown.
-Figure~\ref{fig-initial-spread} column 1 shows the mean standard
-deviation of errors among ensemble members at various lead times. At one
-day lead time (Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a
-slightly larger spread than ACCESS-S1 due to the way that ensemble
-members are generated. ACCESS-S1 ensemble members are generated by
-adding random field perturbations to the atmosphere only, which then are
-transferred to the other components via the coupled simulation
-\citep{hudson2017}. With this scheme, ensemble members are all but
-guaranteed to be underdispersed in the ocean and sea-ice components. The
-time-lag ensemble used for ACCESS-S2 ensures greater spread. This
-difference is gone after about just two days, and both systems have a
-comparable spread in ensemble member error afterwards
-(Fig.~\ref{fig-initial-spread} b1 and c1).
-
-Figure~\ref{fig-initial-spread} column 2, on the other hand, shows the
-standard deviation of ensemble mean error of each hindcast and the
-persistence forecast. At one day lead time, ACCESS-S2 ensemble mean
-error standard deviation is much larger than ACCESS-S1's, which in turn
-is comparable to the persistence forecast error standard deviation. At
-longer lead times, the spread of ACCESS-S1 and persistence forecast
-standard deviation increases to eventually be comparable to ACCESS-S2
-and the standard deviation in climatological forecast errors. ACCESS-S2
-error standard deviation is fairly independent of lead time and similar
-to the climatological forecast error standard deviation at all lead
-times.
-
-\FloatBarrier
-
-\subsection{Conclusions}\label{conclusions}
-
-Sea-ice forecasts from the ACCESS-S2 system show a significant low
-extent bias, particularly during late summer and early autumn. This bias
-is attributed to a faster and longer melt season between January and
-March, and slower growth between March and April. This underestimation
-during the minimum and early freezing season is a common issue in many
-seasonal-to-subseasonal (S2S) systems, suggesting potential problems
-either with the model's thermodynamic representation or with short wave
-radiation forcing, as shown in other climate models
-\citep{zampieri2019, roach2020}. Even though ACCESS-S2 shares the same
-model components as ACCESS-S1, the latter does not suffer from this
-bias, indicating that assimilating sea-ice concentrations successfully
-corrects for the negative bias that exists in the free-running model.
-
-Ensemble spread grows quickly even when perturbations are only
-implemented in the atmosphere component (in ACCESS-S1), indicating that
-sea ice is indeed responding quickly to atmospheric perturbations.
-However, our analysis suggests that the atmosphere and ocean data
-assimilation implemented in ACCESS-S2 is only effectively influencing
-sea-ice initial conditions from June to October, while the rest of the
-year, the sea-ice component runs virtually free, reverting to its biased
-equilibrium state. \citet{zhou2022} had previously evaluated sea-ice
-forecasts in ACCESS-S2 and also highlighted the poor performance of this
-forecasting system attributed to the lack of good initial conditions.
-
-Analysis of the error spread shows that ACCESS-S2 initial conditions
-from December to May not only have large errors, but that the initial
-error spread is very large compared with ACCESS-S1. This spread is not
-due to the perturbation scheme, since the mean error variance for
-individual forecasts is low and comparable with ACCESS-S1. Instead, it
-is due to large variance of the mean error of individual forecasts,
-which is comparable to the climatology spread. This is further evidence
-that individual initial conditions are not being affected by the data
-assimilation scheme.
-
-Although ACCESS-S1 only assimilates sea-ice concentration, it is clear
-that sea-ice thickness is also affected through the assimilation
-process. ACCESS-S1 simulates significantly thicker ice than ACCESS-S2
-and in both systems sea-ice is thicker at shorter lead times than at
-longer lead times. Both the explicit data assimilation in ACCESS-S2 and
-the effects of atmospheric and oceanic data assimilation in ACCESS-S1
-might be nudging simulated sea ice to be thicker than the model
-equilibrium state. We suggest that the thinner sea ice in ACCESS-S2
-contributes to the large sea-ice extent variance, but other mechanisms,
-such as unbalanced initial conditions might also be important.
-
-Given that ACCESS-S2 sea-ice extent is not directly initialised by
-sea-ice observations, comparing its forecasts with those of ACCESS-S1
-allows us to estimate the time-scale over which initial conditions are
-important. We find that initial conditions affect Antarctic sea-ice
-forecasts in the order of a few months, but that effect is seasonally
-dependent. January to April initial conditions improve forecasts for up
-to three months. February initial conditions in particular are shown to
-be crucial for determining sea-ice evolution at least up to May. Arctic
-sea-ice forecasts also show greater sensitivity to initial conditions in
-boreal summer, compared with boreal winter \citep{day2014, bunzel2016},
-suggesting a similar mechanism might be playing a role.
-
-Forecasts initialised in winter have very little skill and ACCESS-S1 and
-ACCESS-S2 forecast errors are statistically indistinguishable after just
-two weeks. This is consistent with \citet{libera2022}'s finding of a
-``winter predictability barrier'' in the Weddell Sea, although they
-describe the barrier as a sharp loss of predictability in July, and here
-we find a gradual reduction in skill compared with climatology around
-June. This difference might be due to our use of pan-Antarctic RMSE,
-since our regional analysis indicates that the degraded skill is most
-dramatic in the King Haakon Sea.
-
-These findings have important implications for both operational
-forecasting, model development and predictability studies. For
-operational centers, our results suggest that efforts to improve sea-ice
-data assimilation should prioritize the summer and autumn months when
-initial conditions have the greatest impact on forecast skill.
-Additionally, the substantial bias in ACCESS-S2 highlights the need for
-improved model physics, particularly in the representation of sea-ice
-thermodynamics and radiation processes. Crucially, our results suggest
-dramatic seasonal variations in sea-ice predictability. Future studies
-should therefore use initial conditions through the whole year rather
-than focusing on a limited number of initialisation dates.
-
-\section{Acknowledgments}\label{acknowledgments}
-
-We thank the internal reviewers Bethan White and Xiaobing Zhou for their
-comments and feedback.
-
-This work benefited from earlier unpublished work by Laura Davies, Phil
-Reid, Andrew G. Marshall.
-
-This research was undertaken with the assistance of resources from the
-National Computational Infrastructure (NCI Australia), an NCRIS enabled
-capability supported by the Australian Government.
-
-This work was supported by ARC SRIEAS Grant SR200100005 Securing
-Antarctica's Environmental Future.
-
-\section{Code and data availability}\label{code-and-data-availability}
-
-The underlying code for this study is available on GitHub and can be
-accessed via this link https://github.com/eliocamp/access-s2\_ice-eval.
-
-Raw data of ACCESS-S1 and ACCESS-S2 forecasts are not available due to
-size. Derived datasets required to reproduce the results are available
-in this Zenodo repository: https://zenodo.org/records/17479538
-\citep{campitelli2025}
-
-\section{References}\label{references}
-
-\renewcommand{\bibsection}{}
-\bibliography{references.bib}
 
@@
 
-\caption{\label{fig-bias-2-osi}Same as Figure~\ref{fig-bias-1} but for
+\caption{\label{fig-bias-2-osi}Same as Figure \ref{fig-bias-1} but for
 ACCESS-S1.}
@@
 
-\caption{\label{fig-rmse_lon-2-osi}Same as Figure~\ref{fig-rmse_lon-1}
+\caption{\label{fig-rmse_lon-2-osi}Same as Figure \ref{fig-rmse_lon-1}
 but for ACCESS-S2.}
@@
 
-\caption{\label{fig-rmse_lon-3-osi}Same as Figure~\ref{fig-rmse_lon-1}
+\caption{\label{fig-rmse_lon-3-osi}Same as Figure \ref{fig-rmse_lon-1}
 but for the difference between ACCESS-S1 and ACCESS-S2.}
@@
 
+\bibliography{references.bib}
+
 \end{document}

### 2025-11-27 analysis/paper/access-ice.qmd
Renders supplemental in one file

--- 2025-11-19 analysis/paper/access-ice.qmd
@@
     fig-dpi: 300
+
 header-includes: | 
@@
 
-@fig-hindcast-extent shows mean sea-ice extent of the +S1 and +S2 hindcasts (row a) and their differences from mean sea-ice extent of +CDR (row b).
+@fig-hindcast-extent (@fig-hindcast-extent-osi for OSI) shows mean sea-ice extent of the +S1 and +S2 hindcasts (row a) and their differences from mean sea-ice extent of +CDR (row b).
 Mean extent at the first of every month is indicated with circles for the initial conditions and with triangles for the longest lead time possible for each model (between `r range_max_lag[model == "S2", min]` and `r range_max_lag[model == "S2", max]` days for +S2 and between `r range_max_lag[model == "S1", min]` and `r range_max_lag[model == "S1", max]` days for +S1).
@@
 This is due primarily to the melt season being longer than in observations and with faster melt between January and March and the growing seasons being shorter with slower growth during March and April.
-This is then followed by faster growth between May and July (@fig-mean-growth).
+This is then followed by faster growth between May and July (@fig-mean-growth and @fig-mean-growth-osi).
 Many sea-ice models exhibit this systematic underestimation during the sea-ice minimum and early freezing season [@massonnet2023], which could indicate problems in the representation of thermodynamics in the model [@zampieri2019].
@@
 
-To further understand the bias in +S2, @fig-bias-1 shows spatial patterns of the differences of monthly mean sea-ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest monthly lead time.
+To further understand the bias in +S2, @fig-bias-1 (@fig-bias-1-osi) shows spatial patterns of the differences of monthly mean sea-ice concentrations between NOAA/NSIDC CDRV4 and +S2 hindcasts at the shortest monthly lead time.
 From October to May, the model underestimates sea-ice concentrations in most regions except for the inner Weddell Sea in April and May, where sea-ice concentrations saturate to 1 both in the observations and forecasts.
@@
 
-+S1 has a comparatively smaller overall bias (@fig-bias-2). 
++S1 has a comparatively smaller overall bias (@fig-bias-2 and @fig-bias-2-osi). 
 The largest values are found between April and June, when the faster growth results in large positive bias along the sea-ice edge, and in January, when the faster melt leads to large negative bias in the Weddell and Amundsen Seas.
@@
 
-@fig-extent-anom shows monthly sea-ice extent anomalies forecasted at selected lead times.
+@fig-extent-anom (@fig-extent-anom-osi) shows monthly sea-ice extent anomalies forecasted at selected lead times.
 Compared with +S1, +S2 anomaly forecasts are relatively poor (large RMSE) even for the first month (lead time 0), whereas +S1 forecasts stay relatively skilful even at a lead time of three months.
@@
 The explanation seems to be the mentioned increased interannual variability.
-@fig-extent-sd shows the interannual standard deviation of monthly sea-ice extent of the forecasts as a function of lead time compared with observations.
+@fig-extent-sd (@fig-extent-sd-osi) shows the interannual standard deviation of monthly sea-ice extent of the forecasts as a function of lead time compared with observations.
 +S1 standard deviation lies within the observed standard deviation regardless of lead time, while +S2 standard deviation is more than twice that of observations at zero lead time and only approaches the observed value at nine month lead time for most months.
@@
 To assess +S2 forecasts in more detail, we compute error measures for all hindcasts started on the 1st of every month.
-@fig-rmse shows the mean RMSE of sea-ice concentration anomalies for +S1 and +S2 hindcasts compared against persistence and climatological forecasts used as a benchmark.
+@fig-rmse (@fig-rmse-osi) shows the mean RMSE of sea-ice concentration anomalies for +S1 and +S2 hindcasts compared against persistence and climatological forecasts used as a benchmark.
 Due to errors in the initial conditions, it is expected that persistence forecasts would be better than the model forecasts at very short lead times, but that the persistence forecast errors would grow faster and may eventually surpass the model forecast errors.
@@
 
-@fig-lead-time-window summarises the lead time window in which each hindcast is better than both the persistence forecast and the climatological forecast as a function of forecast month. 
+@fig-lead-time-window (@fig-lead-time-window-osi) summarises the lead time window in which each hindcast is better than both the persistence forecast and the climatological forecast as a function of forecast month.
 +S1 forecasts have a wider lead time window in the summer than the other seasons and is not better than both benchmarks at forecasting June sea-ice concentration anomalies. 
@@
 
-For +S1 forecasts (@fig-rmse_lon-1), skill tends to be lower off the coast of Eastern Antarctica even at short lead times; for instance, the skill score for forecasts initialised in May and June are negative between 0° and 120°E even at almost zero lead time.
+For +S1 forecasts (@fig-rmse_lon-1 and @fig-rmse_lon-1-osi), skill tends to be lower off the coast of Eastern Antarctica even at short lead times; for instance, the skill score for forecasts initialised in May and June are negative between 0° and 120°E even at almost zero lead time.
 This mirrors @libera2022 findings of a “winter predictability barrier”, although they focus on the Weddell Sea and here we show that the effect seems to be stronger more to the east. 
@@
 
-+S2 forecasts  (@fig-rmse_lon-2) also have lower skill over East Antarctica. 
++S2 forecasts  (@fig-rmse_lon-2 and @fig-rmse_lon-2-osi) also have lower skill over East Antarctica. 
 From July to December even though the pan-Antarctic average skill is negative at all lead times ([Fig. @fig-lead-time-window]), it is positive for up to a month in West Antarctica. 
@@
 
-Finally, @fig-rmse_lon-3 shows the difference in skill between +S1 and +S2. 
+Finally, @fig-rmse_lon-3 (@fig-rmse_lon-3-osi) shows the difference in skill between +S1 and +S2. 
 Large differences in skill indicate areas and months that are most affected by the data assimilation present in +S1.
@@
 In @fig-rmse the mean error was shown. 
-@fig-initial-spread column 1 shows the mean standard deviation of errors among ensemble members at various lead times.
+@fig-initial-spread (@fig-initial-spread-osi) column 1 shows the mean standard deviation of errors among ensemble members at various lead times.
 At one day lead time ([Fig. @fig-initial-spread] a.1) +S2 has a slightly larger spread than +S1 due to the way that ensemble members are generated.

### 2025-11-27 build.sh

@@ -2,30 +2,31 @@ set -eu
 
 dir=analysis/paper
 source=$dir/access-ice.qmd
-
-main=$dir/main.qmd 
-supp=$dir/supp.qmd 
 zip=access-ice.zip
 
-cp $source $main
-quarto render $main --profile main
-cp $dir/main.aux $dir/ref_main.aux
+# main=$dir/main.qmd 
+# supp=$dir/supp.qmd 
 
-cp $source $supp
-quarto render $supp --profile supp
-cp $dir/supp.aux $dir/ref_supp.aux
 
-rm $main $supp 
+# cp $source $main
+# quarto render $main --profile main
+# cp $dir/main.aux $dir/ref_main.aux
+
+# cp $source $supp
+# quarto render $supp --profile supp
+# cp $dir/supp.aux $dir/ref_supp.aux
+
+# rm $main $supp 
+
+
+# pdflatex main.tex
+# pdflatex supp.tex
+
+quarto render $source
 
 cd $dir
-pdflatex main.tex
-pdflatex supp.tex
+if [ -f $zip ]; then
+    rm $zip
+fi 
+zip -r $zip *.bst *.cls references.bib access-ice.tex access-ice_files/
 
-
-
-rm $zip
-zip -r $zip *.bst *.csl references.bib main.tex ref_supp.aux access-ice_files/
-
-# Clean up everything 
-find . -maxdepth 1 -type f -name 'main*' ! -name 'main.pdf' ! -name 'main.tex' -exec rm {} +
-find . -maxdepth 1 -type f -name 'supp*' ! -name 'supp.pdf' ! -name 'supp.tex' -exec rm {} +

### 2025-11-27 analysis/paper/access-ice.tex
Renders

@@
 
-Figure~\ref{fig-hindcast-extent} shows mean sea-ice extent of the
-ACCESS-S1 and ACCESS-S2 hindcasts (row a) and their differences from
-mean sea-ice extent of NOAA/NSIDC CDRV4 (row b). Mean extent at the
-first of every month is indicated with circles for the initial
-conditions and with triangles for the longest lead time possible for
-each model (between 274 and 277 days for ACCESS-S2 and between 213 and
-216 days for ACCESS-S1). At this long lead time, information of the
-initial conditions is essentially lost and the forecast reverts close to
-each model's preferred equilibrium state.
+Figure~\ref{fig-hindcast-extent} (Figure~\ref{fig-hindcast-extent-osi}
+for OSI) shows mean sea-ice extent of the ACCESS-S1 and ACCESS-S2
+hindcasts (row a) and their differences from mean sea-ice extent of
+NOAA/NSIDC CDRV4 (row b). Mean extent at the first of every month is
+indicated with circles for the initial conditions and with triangles for
+the longest lead time possible for each model (between 274 and 277 days
+for ACCESS-S2 and between 213 and 216 days for ACCESS-S1). At this long
+lead time, information of the initial conditions is essentially lost and
+the forecast reverts close to each model's preferred equilibrium state.
 
@@
 slower growth during March and April. This is then followed by faster
-growth between May and July (Figure~\ref{fig-mean-growth}). Many sea-ice
-models exhibit this systematic underestimation during the sea-ice
-minimum and early freezing season \citep{massonnet2023}, which could
-indicate problems in the representation of thermodynamics in the model
-\citep{zampieri2019}. It is also not surprising that both forecasting
-systems converge to a similar equilibrium state because they share the
-same model formulation.
+growth between May and July (Figure~\ref{fig-mean-growth} and
+Figure~\ref{fig-mean-growth-osi}). Many sea-ice models exhibit this
+systematic underestimation during the sea-ice minimum and early freezing
+season \citep{massonnet2023}, which could indicate problems in the
+representation of thermodynamics in the model \citep{zampieri2019}. It
+is also not surprising that both forecasting systems converge to a
+similar equilibrium state because they share the same model formulation.
 
@@
 To further understand the bias in ACCESS-S2, Figure~\ref{fig-bias-1}
-shows spatial patterns of the differences of monthly mean sea-ice
-concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
-shortest monthly lead time. From October to May, the model
-underestimates sea-ice concentrations in most regions except for the
-inner Weddell Sea in April and May, where sea-ice concentrations
-saturate to 1 both in the observations and forecasts. In winter, the
-differences are limited to a narrow band around the sea-ice edge with
-slight positive biases in the African sector of East Antarctica and
-negative biases around the Indian Ocean sector which partially
-compensate, resulting in the near-zero extent bias seen in those months
-(Figure~\ref{fig-hindcast-extent}).
+(Figure~\ref{fig-bias-1-osi}) shows spatial patterns of the differences
+of monthly mean sea-ice concentrations between NOAA/NSIDC CDRV4 and
+ACCESS-S2 hindcasts at the shortest monthly lead time. From October to
+May, the model underestimates sea-ice concentrations in most regions
+except for the inner Weddell Sea in April and May, where sea-ice
+concentrations saturate to 1 both in the observations and forecasts. In
+winter, the differences are limited to a narrow band around the sea-ice
+edge with slight positive biases in the African sector of East
+Antarctica and negative biases around the Indian Ocean sector which
+partially compensate, resulting in the near-zero extent bias seen in
+those months (Figure~\ref{fig-hindcast-extent}).
 
 ACCESS-S1 has a comparatively smaller overall bias
-(Figure~\ref{fig-bias-2}). The largest values are found between April
-and June, when the faster growth results in large positive bias along
-the sea-ice edge, and in January, when the faster melt leads to large
-negative bias in the Weddell and Amundsen Seas.
+(Figure~\ref{fig-bias-2} and Figure~\ref{fig-bias-2-osi}). The largest
+values are found between April and June, when the faster growth results
+in large positive bias along the sea-ice edge, and in January, when the
+faster melt leads to large negative bias in the Weddell and Amundsen
+Seas.
 
@@
 
-Figure~\ref{fig-extent-anom} shows monthly sea-ice extent anomalies
-forecasted at selected lead times. Compared with ACCESS-S1, ACCESS-S2
-anomaly forecasts are relatively poor (large RMSE) even for the first
-month (lead time 0), whereas ACCESS-S1 forecasts stay relatively skilful
-even at a lead time of three months. ACCESS-S2 shows much larger
-interannual variability than observations, with dramatic lows between
-1995 and 2007, and highs between 2007 and 2015.
+Figure~\ref{fig-extent-anom} (Figure~\ref{fig-extent-anom-osi}) shows
+monthly sea-ice extent anomalies forecasted at selected lead times.
+Compared with ACCESS-S1, ACCESS-S2 anomaly forecasts are relatively poor
+(large RMSE) even for the first month (lead time 0), whereas ACCESS-S1
+forecasts stay relatively skilful even at a lead time of three months.
+ACCESS-S2 shows much larger interannual variability than observations,
+with dramatic lows between 1995 and 2007, and highs between 2007 and
+2015.
 
@@
 explanation seems to be the mentioned increased interannual variability.
-Figure~\ref{fig-extent-sd} shows the interannual standard deviation of
-monthly sea-ice extent of the forecasts as a function of lead time
-compared with observations. ACCESS-S1 standard deviation lies within the
-observed standard deviation regardless of lead time, while ACCESS-S2
-standard deviation is more than twice that of observations at zero lead
-time and only approaches the observed value at nine month lead time for
-most months.
+Figure~\ref{fig-extent-sd} (Figure~\ref{fig-extent-sd-osi}) shows the
+interannual standard deviation of monthly sea-ice extent of the
+forecasts as a function of lead time compared with observations.
+ACCESS-S1 standard deviation lies within the observed standard deviation
+regardless of lead time, while ACCESS-S2 standard deviation is more than
+twice that of observations at zero lead time and only approaches the
+observed value at nine month lead time for most months.
 
@@
 for all hindcasts started on the 1st of every month.
-Figure~\ref{fig-rmse} shows the mean RMSE of sea-ice concentration
-anomalies for ACCESS-S1 and ACCESS-S2 hindcasts compared against
-persistence and climatological forecasts used as a benchmark. Due to
-errors in the initial conditions, it is expected that persistence
-forecasts would be better than the model forecasts at very short lead
-times, but that the persistence forecast errors would grow faster and
-may eventually surpass the model forecast errors. The black line shows
-that the persistence forecast error indeed grows rapidly and reaches its
-maximum in about 30 days for most months except for February, when it
-grows much slower. The ACCESS-S1 forecast errors grow slower than
-persistence forecast errors and remain lower after less than 10 days on
-average. The ACCESS-S2 forecast error starts high in all months and is
-lower than the persistence forecast error after more than 15 days in
-most months except for forecast initialised in February, when it takes
-80 days.
+Figure~\ref{fig-rmse} (Figure~\ref{fig-rmse-osi}) shows the mean RMSE of
+sea-ice concentration anomalies for ACCESS-S1 and ACCESS-S2 hindcasts
+compared against persistence and climatological forecasts used as a
+benchmark. Due to errors in the initial conditions, it is expected that
+persistence forecasts would be better than the model forecasts at very
+short lead times, but that the persistence forecast errors would grow
+faster and may eventually surpass the model forecast errors. The black
+line shows that the persistence forecast error indeed grows rapidly and
+reaches its maximum in about 30 days for most months except for
+February, when it grows much slower. The ACCESS-S1 forecast errors grow
+slower than persistence forecast errors and remain lower after less than
+10 days on average. The ACCESS-S2 forecast error starts high in all
+months and is lower than the persistence forecast error after more than
+15 days in most months except for forecast initialised in February, when
+it takes 80 days.
 
@@
 
-Figure~\ref{fig-lead-time-window} summarises the lead time window in
-which each hindcast is better than both the persistence forecast and the
-climatological forecast as a function of forecast month. ACCESS-S1
+Figure~\ref{fig-lead-time-window}
+(Figure~\ref{fig-lead-time-window-osi}) summarises the lead time window
+in which each hindcast is better than both the persistence forecast and
+the climatological forecast as a function of forecast month. ACCESS-S1
 forecasts have a wider lead time window in the summer than the other
@@
 
-For ACCESS-S1 forecasts (Figure~\ref{fig-rmse_lon-1}), skill tends to be
-lower off the coast of Eastern Antarctica even at short lead times; for
-instance, the skill score for forecasts initialised in May and June are
-negative between 0° and 120°E even at almost zero lead time. This
-mirrors \citet{libera2022} findings of a ``winter predictability
-barrier'', although they focus on the Weddell Sea and here we show that
-the effect seems to be stronger more to the east. In West Antarctica
-there is a hint of easterly-propagating skill in forecasts initialised
-in February and March. This is consistent with \citet{holland2013}
-findings that memory of sea-ice anomalies are stored in ocean heat
-content anomalies that are transported east by the Antarctic Circumpolar
-Current.
-
-ACCESS-S2 forecasts (Figure~\ref{fig-rmse_lon-2}) also have lower skill
-over East Antarctica. From July to December even though the
-pan-Antarctic average skill is negative at all lead times
-(Fig.~\ref{fig-lead-time-window}), it is positive for up to a month in
-West Antarctica. Since oceanic and atmospheric forcing is the only
-source of information, this suggests that sea-ice in this region is
-particularly sensitive to oceanic and atmospheric forcing and suggests a
-role of the Pacific-South American mode and the Amundsen Sea Low to
-shape sea-ice concentration anomalies. The fact that this is evident in
-the months in which El Niño--Southern Oscillation teleconnections are
-more important for atmospheric circulation also suggests the influence
-of tropical Pacific variability. February and March are the only two
-months that can be forecasted with marginally positive skill in large
-regions.
+For ACCESS-S1 forecasts (Figure~\ref{fig-rmse_lon-1} and
+Figure~\ref{fig-rmse_lon-1-osi}), skill tends to be lower off the coast
+of Eastern Antarctica even at short lead times; for instance, the skill
+score for forecasts initialised in May and June are negative between 0°
+and 120°E even at almost zero lead time. This mirrors \citet{libera2022}
+findings of a ``winter predictability barrier'', although they focus on
+the Weddell Sea and here we show that the effect seems to be stronger
+more to the east. In West Antarctica there is a hint of
+easterly-propagating skill in forecasts initialised in February and
+March. This is consistent with \citet{holland2013} findings that memory
+of sea-ice anomalies are stored in ocean heat content anomalies that are
+transported east by the Antarctic Circumpolar Current.
+
+ACCESS-S2 forecasts (Figure~\ref{fig-rmse_lon-2} and
+Figure~\ref{fig-rmse_lon-2-osi}) also have lower skill over East
+Antarctica. From July to December even though the pan-Antarctic average
+skill is negative at all lead times (Fig.~\ref{fig-lead-time-window}),
+it is positive for up to a month in West Antarctica. Since oceanic and
+atmospheric forcing is the only source of information, this suggests
+that sea-ice in this region is particularly sensitive to oceanic and
+atmospheric forcing and suggests a role of the Pacific-South American
+mode and the Amundsen Sea Low to shape sea-ice concentration anomalies.
+The fact that this is evident in the months in which El Niño--Southern
+Oscillation teleconnections are more important for atmospheric
+circulation also suggests the influence of tropical Pacific variability.
+February and March are the only two months that can be forecasted with
+marginally positive skill in large regions.
 
@@
 
-Finally, Figure~\ref{fig-rmse_lon-3} shows the difference in skill
-between ACCESS-S1 and ACCESS-S2. Large differences in skill indicate
-areas and months that are most affected by the data assimilation present
-in ACCESS-S1. Between January and March, which are the months in which
-ACCESS-S1 is the most skilful (Figure~\ref{fig-lead-time-window}), most
-of the improvement compared with ACCESS-S2 is present in the Ross and
-Weddell Sea. In April and May, the improvement seems more homogeneous.
+Finally, Figure~\ref{fig-rmse_lon-3} (Figure~\ref{fig-rmse_lon-3-osi})
+shows the difference in skill between ACCESS-S1 and ACCESS-S2. Large
+differences in skill indicate areas and months that are most affected by
+the data assimilation present in ACCESS-S1. Between January and March,
+which are the months in which ACCESS-S1 is the most skilful
+(Figure~\ref{fig-lead-time-window}), most of the improvement compared
+with ACCESS-S2 is present in the Ross and Weddell Sea. In April and May,
+the improvement seems more homogeneous.
 
@@
 In Figure~\ref{fig-rmse} the mean error was shown.
-Figure~\ref{fig-initial-spread} column 1 shows the mean standard
-deviation of errors among ensemble members at various lead times. At one
-day lead time (Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a
-slightly larger spread than ACCESS-S1 due to the way that ensemble
-members are generated. ACCESS-S1 ensemble members are generated by
-adding random field perturbations to the atmosphere only, which then are
-transferred to the other components via the coupled simulation
-\citep{hudson2017}. With this scheme, ensemble members are all but
-guaranteed to be underdispersed in the ocean and sea-ice components. The
-time-lag ensemble used for ACCESS-S2 ensures greater spread. This
-difference is gone after about just two days, and both systems have a
-comparable spread in ensemble member error afterwards
+Figure~\ref{fig-initial-spread} (Figure~\ref{fig-initial-spread-osi})
+column 1 shows the mean standard deviation of errors among ensemble
+members at various lead times. At one day lead time
+(Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a slightly larger
+spread than ACCESS-S1 due to the way that ensemble members are
+generated. ACCESS-S1 ensemble members are generated by adding random
+field perturbations to the atmosphere only, which then are transferred
+to the other components via the coupled simulation \citep{hudson2017}.
+With this scheme, ensemble members are all but guaranteed to be
+underdispersed in the ocean and sea-ice components. The time-lag
+ensemble used for ACCESS-S2 ensures greater spread. This difference is
+gone after about just two days, and both systems have a comparable
+spread in ensemble member error afterwards
 (Fig.~\ref{fig-initial-spread} b1 and c1).

### 2025-12-04 analysis/paper/access-ice.qmd
Uses copernicus template

@@
 title: The Importance of Initial Conditions in Seasonal Predictions of Antarctic Sea Ice
+pdf-engine: pdflatex
+journal:
+  name: bg
+author:
+  - given_name: Elio 
+    surname: Campitelli
+    affiliation: "1, 2"
+    email: elio.campitelli@monash.edu
+    corresponding: true
+  - given_name: Ariaan
+    surname: Purich
+    affiliation: "1, 2"
+  - given_name: Julie
+    surname: Arblaster
+    affiliation: "1, 2"
+  - given_name: Eun-Pa
+    surname: Lim
+    affiliation: "3"
+  - given_name: Matthew C.
+    surname: Wheeler
+    affiliation: "3"
+  - given_name: Phillip
+    surname: Reid
+    affiliation: "3"
+
+affiliation:
+  - code: 1
+    address: School of Earth, Atmosphere and Environment, Monash University, Kulin Nations, Clayton, Victoria, Australia. 
+  - code: 2
+    address: ARC Special Research Initiative for Securing Antarctica’s Environmental Future, Clayton, Kulin Nations, Victoria, Australia
+  - code: 3
+    address: Research, Bureau of Meteorology, Melbourne, Australia.
+
+bibliography: references.bib  
+running:
+  title: The Importance of Initial Conditions in Seasonal Predictions of Antarctic Sea Ice
+  author: Campitelli et al.
+# This section is mandatory even if you declare that no competing interests are present.
+competinginterests: |
+  The authors declare no competing interests.
+# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
+# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
+copyrightstatement: |
+  The author's copyright for this publication is transferred to institution/company. 
+### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
+### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
+### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
+availability:
+  #code: |
+  #  use this to add a statement when having only software code available
+  #data: |
+  #  use this to add a statement when having only data sets available
+  codedata: |
+    The underlying code for this study is available on GitHub: https://github.com/eliocamp/access-s2_ice-eval.
+    Raw data of +S1 and +S2 hindcast are not available due to size. 
+    Derived datasets required to reproduce the results, including extent timeseries and error measures, are available in this Zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
+authorcontribution: |
+    EC performed the data analysis and wrote the manuscript draft. 
+    AP, JA, EL, MW and PR, performed interpretation of the results, and reviewed and edited the draft. 
+    All authors read and approved the final manuscript.
+
+acknowledgements: |
+    We thank the internal reviewers Bethan White and Xiaobing Zhou for their comments and feedback. 
+    This work benefited from earlier unpublished work by Laura Davies, Phil Reid, Andrew G. Marshall. 
+    This research was undertaken with the assistance of resources from the National Computational Infrastructure (NCI Australia), an NCRIS enabled capability supported by the Australian Government.
+    This work was supported by ARC SRIEAS Grant SR200100005 Securing Antarctica’s Environmental Future.
+# appendix: |
+#   \section{Figures and tables in appendices}
+#   Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
+#   \subsection{Option 1}
+#   If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
+#   They will be correctly named automatically.
+#   \subsection{Option 2}
+#   If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
+
+#   To rename them correctly to A1, A2, etc., please add the following commands in front of them:
+#   `\appendixfigures` needs to be added in front of appendix figures
+#   `\appendixtables` needs to be added in front of appendix tables
+
+#   Please add `\clearpage` between each table and/or figure.
 format: 
-  nature-pdf:
-    # journal.cite-style is included in the tex file but ignored by pandoc if 
-    # cite-method is not `natbib`.
-    journal:
-      cite-style: sn-mathphys-num
-    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
-    # to use the bst styles from the upstream template.
-    cite-method: natbib
+  copernicus-pdf: 
       keep-tex: true
-    equal-margins: false
       classoption: referee,lineno
-    fig-width: 5  
+      fig-width: 6.5
       fig-pos: "!htb"
       fig-dpi: 300    
-
 header-includes: | 
    \usepackage[section]{placeins}
-author:
-  - name: Elio Campitelli
-    affiliations:
-      - name: Monash University, Kulin Nations, Clayton, Victoria, Australia.
-        id: 1
-        department: School of Earth, Atmosphere and Environment
-        # address: 1737 Cambridge Street
-        # city: Cambridge
-        # state: MA
-        # postal-code: 02138
-      - name: Clayton, Kulin Nations, Victoria, Australia
-        id: 2
-        department: SARC Special Research Initiative for Securing Antarctica’s Environmental Future
-        # address: 1 Oxford Street
-        # city: Cambridge
-        # state: MA
-        # postal-code: 02138
-    email: elio.campitelli@monash.edu
-    attributes:
-      corresponding: true
-  - name: Ariaan Purich
-    affiliations:
-      - ref: 1
-      - ref: 2
-  - name: Julie Arblaster
-    affiliations:
-      - ref: 1
-      - ref: 2
-  - name: Eun-Pa Lim
-    affiliations:
-      - department: Research, Bureau of Meteorology, Melbourne, Australia
-        id: 3
-  - name: Matthew C. Wheeler
-    affiliations:
-      - ref: 3      
-  - name: Phillip Reid
-    affiliations:
-      - ref: 3      
+
 keywords: [sea ice, seasonal predictability, initial conditions, forecasting]
-bibliography: references.bib
 
@@
   - _extensions/pandoc-ext/abstract-section/abstract-section.lua
-  # - ../../_extensions/mloubout/critic-markup/critic-markup.lua
 
@@
 It is based on mostly the same sensors as the NOAA CDR but computed independently using different algorithms. 
-Figures prepared with this dataset are provided in the supplementary material and do not differ significantly from the ones prepared using CDR.
+Figures prepared with this dataset are provided in the appendix and do not differ significantly from the ones prepared using CDR.
 
@@
 
-# Funding
-
-This work was supported by ARC SRIEAS Grant SR200100005 Securing Antarctica’s Environmental Future.
-
-# Acknowledgments
-
-We thank the internal reviewers Bethan White and Xiaobing Zhou for their comments and feedback.
-
-This work benefited from earlier unpublished work by Laura Davies, Phil Reid, Andrew G. Marshall.
-
-This research was undertaken with the assistance of resources from the National Computational Infrastructure (NCI Australia), an NCRIS enabled capability supported by the Australian Government.
-
-# Code and data availability
-
-The underlying code for this study is available on GitHub: https://github.com/eliocamp/access-s2_ice-eval.
-
-Raw data of +S1 and +S2 hindcast are not available due to size. 
-Derived datasets required to reproduce the results, including extent timeseries and error measures, are available in this Zenodo repository: https://zenodo.org/records/17479538 [@campitelli2025]
-
-# Competing interests
-
-All authors declare no financial or non-financial competing interests.
-
-# Author contributions
-
-EC performed the data analysis and wrote the manuscript draft. 
-AP, JA, EL, MW and PR, performed interpretation of the results, and reviewed and edited the draft. 
-All authors read and approved the final manuscript.
-
 # References
@@
 
-\appendix  
-\renewcommand{\thefigure}{A\arabic{figure}}
-\setcounter{figure}{0}
-
-# Supplementary figures {.appendix}
+# Appendix {.appendix}
+
+\appendixfigures
 
@@
 
+\clearpage
+
 ```{r, fig-mean-growth-osi, ref.label=I("fig-mean-growth")}
@@
 
+\clearpage
+
 ```{r, plot_bias-osi, ref.label=I("plot_bias")}
@@
 
+\clearpage
+
 ```{r, fig-bias-1-osi, ref.label=I("fig-bias-1")}
@@
 
+\clearpage
+
 ```{r, fig-bias-2-osi, ref.label=I("fig-bias-2")}
@@
 
+\clearpage
+
 ```{r, fig-extent-anom-osi, ref.label=I("fig-extent-anom")}
@@
 
+\clearpage
+
 ```{r, fig-extent-sd-osi, ref.label=I("fig-extent-sd")}
@@
 
+\clearpage
+
 ```{r, fig-rmse-osi, ref.label=I("fig-rmse")}
@@
 
+\clearpage
+
 ```{r, fig-lead-time-window-osi, ref.label=I("fig-lead-time-window")}
@@
 
+\clearpage
+
 ```{r, rmse_lon_plots-osi, ref.label=I("rmse_lon_plots")}
@@
 
+\clearpage
+
 ```{r, fig-rmse_lon-1-osi, ref.label=I("fig-rmse_lon-1")}
@@
 
+\clearpage
+
 ```{r, fig-rmse_lon-2-osi, ref.label=I("fig-rmse_lon-2")}
@@
 
+\clearpage
+
 ```{r, fig-rmse_lon-3-osi, ref.label=I("fig-rmse_lon-3")}
@@
 
+\clearpage
+
 ```{r, fig-initial-spread-osi, ref.label=I("fig-initial-spread")}

### 2025-12-04 analysis/paper/access-ice.tex
Renders

@@
 %
-\documentclass[
-  sn-mathphys-num,
-  referee,lineno]{sn-jnl}
-
+\documentclass[bg, manuscript]{copernicus}
 \usepackage{xcolor}
 \usepackage{amsmath,amssymb}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
 \usepackage{iftex}
@@
 
-%%%% Standard Packages
-
-\usepackage{graphicx}%
-\usepackage{multirow}%
-\usepackage{amsmath,amssymb,amsfonts}%
-\usepackage{amsthm}%
-\usepackage{mathrsfs}%
-\usepackage[title]{appendix}%
-\usepackage{xcolor}%
-\usepackage{textcomp}%
-\usepackage{manyfoot}%
-\usepackage{booktabs}%
-\usepackage{algorithm}%
-\usepackage{algorithmicx}%
-\usepackage{algpseudocode}%
-\usepackage{listings}%
-
-%%%%
-
-\raggedbottom
+\usepackage[inline]{trackchanges}
 \usepackage[section]{placeins}
@@
   pdftitle={The Importance of Initial Conditions in Seasonal Predictions of Antarctic Sea Ice},
-  pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Matthew C. Wheeler; Phillip Reid},
   pdfkeywords={sea ice, seasonal predictability, initial
@@
   citecolor={Blue},
-  urlcolor={Blue},
+  urlcolor={red},
   pdfcreator={LaTeX via pandoc}}
 
-\title[The Importance of Initial Conditions in Seasonal Predictions of
-Antarctic Sea Ice]{The Importance of Initial Conditions in Seasonal
+\title{The Importance of Initial Conditions in Seasonal Predictions of
+Antarctic Sea Ice}
+
+\Author[1, 2]{Elio}{Campitelli}
+\Author[1, 2]{Ariaan}{Purich}
+\Author[1, 2]{Julie}{Arblaster}
+\Author[3]{Eun-Pa}{Lim}
+\Author[3]{Matthew C.}{Wheeler}
+\Author[3]{Phillip}{Reid}
+
+\affil[1]{School of Earth, Atmosphere and Environment, Monash
+University, Kulin Nations, Clayton, Victoria, Australia.}
+\affil[2]{ARC Special Research Initiative for Securing Antarctica's
+Environmental Future, Clayton, Kulin Nations, Victoria, Australia}
+\affil[3]{Research, Bureau of Meteorology, Melbourne, Australia.}
+
+\runningtitle{The Importance of Initial Conditions in Seasonal
 Predictions of Antarctic Sea Ice}
 
-% author setup
-\author*[1,2]{\fnm{Elio} \sur{Campitelli}}\email{elio.campitelli@monash.edu}\author[1,2]{\fnm{Ariaan} \sur{Purich}}\author[1,2]{\fnm{Julie} \sur{Arblaster}}\author[3]{\fnm{Eun-Pa} \sur{Lim}}\author[3]{\fnm{Matthew C.} \sur{Wheeler}}\author[3]{\fnm{Phillip} \sur{Reid}}
-% affil setup
-\affil[1]{\orgdiv{School of Earth, Atmosphere and
-Environment}, \orgname{Monash University, Kulin Nations, Clayton,
-Victoria, Australia.}}
-\affil[2]{\orgdiv{SARC Special Research Initiative for Securing
-Antarctica's Environmental Future}, \orgname{Clayton, Kulin Nations,
-Victoria, Australia}}
-\affil[3]{\orgdiv{Research, Bureau of Meteorology, Melbourne,
-Australia}}
-
-% abstract
-
-\abstract{Accurate Antarctic sea-ice forecasts are crucial for climate
-monitoring and operational planning, yet they remain challenging due to
-model biases and complex ice-ocean-atmosphere interactions. The two
-versions of the Australian Bureau of Meteorology's ACCESS seasonal
-forecast system, ACCESS-S1 and ACCESS-S2, use identical model
-configuration and differ only in their initial conditions; primarily in
-that ACCESS-S2 does not assimilate sea-ice observations, whereas
-ACCESS-S1 does.\\
+\runningauthor{Campitelli et al.}
+
+\correspondence{Elio\ Campitelli\ (elio.campitelli@monash.edu)}
+
+\begin{document}
+
+\received{}
+\pubdiscuss{} %% only important for two-stage journals
+\revised{}
+\accepted{}
+\published{}
+
+%% These dates will be inserted by Copernicus Publications during the typesetting process.
+
+\firstpage{1}
+\maketitle
+\begin{abstract}
+Accurate Antarctic sea-ice forecasts are crucial for climate monitoring
+and operational planning, yet they remain challenging due to model
+biases and complex ice-ocean-atmosphere interactions. The two versions
+of the Australian Bureau of Meteorology's ACCESS seasonal forecast
+system, ACCESS-S1 and ACCESS-S2, use identical model configuration and
+differ only in their initial conditions; primarily in that ACCESS-S2
+does not assimilate sea-ice observations, whereas ACCESS-S1 does.\\
 This provides a convenient opportunistic experiment to assess the role
@@
 prioritize the summer-autumn period when initial conditions have maximum
-impact on forecast skill.}
-
-% keywords
-\keywords{sea ice,  seasonal predictability,  initial
-conditions,  forecasting}
-
-\begin{document}
-\maketitle
+impact on forecast skill.
+\end{abstract}
+\copyrightstatement{The author's copyright for this publication is
+transferred to institution/company.}
 
@@
 different algorithms. Figures prepared with this dataset are provided in
-the supplementary material and do not differ significantly from the ones
-prepared using CDR.
+the appendix and do not differ significantly from the ones prepared
+using CDR.
 
@@
 
-\section{Funding}\label{funding}
-
-This work was supported by ARC SRIEAS Grant SR200100005 Securing
-Antarctica's Environmental Future.
-
-\section{Acknowledgments}\label{acknowledgments}
-
-We thank the internal reviewers Bethan White and Xiaobing Zhou for their
-comments and feedback.
-
-This work benefited from earlier unpublished work by Laura Davies, Phil
-Reid, Andrew G. Marshall.
-
-This research was undertaken with the assistance of resources from the
-National Computational Infrastructure (NCI Australia), an NCRIS enabled
-capability supported by the Australian Government.
-
-\section{Code and data availability}\label{code-and-data-availability}
-
-The underlying code for this study is available on GitHub:
-https://github.com/eliocamp/access-s2\_ice-eval.
-
-Raw data of ACCESS-S1 and ACCESS-S2 hindcast are not available due to
-size. Derived datasets required to reproduce the results, including
-extent timeseries and error measures, are available in this Zenodo
-repository: https://zenodo.org/records/17479538 \citep{campitelli2025}
-
-\section{Competing interests}\label{competing-interests}
-
-All authors declare no financial or non-financial competing interests.
-
-\section{Author contributions}\label{author-contributions}
-
-EC performed the data analysis and wrote the manuscript draft. AP, JA,
-EL, MW and PR, performed interpretation of the results, and reviewed and
-edited the draft. All authors read and approved the final manuscript.
-
 \section{References}\label{references}
@@
 
-\appendix  
-\renewcommand{\thefigure}{A\arabic{figure}}
-\setcounter{figure}{0}
-
-\section{Supplementary figures}\label{supplementary-figures}
+\section{Appendix}\label{appendix}
+
+\appendixfigures
 
@@
 
+\clearpage
+
 \begin{figure}[!htb]
@@
 
+\clearpage
+
+\clearpage
+
 \begin{figure}[!htb]
@@
 \end{figure}%
+
+\clearpage
 
@@
 
+\codedataavailability{The underlying code for this study is available on
+GitHub: https://github.com/eliocamp/access-s2\_ice-eval. Raw data of +S1
+and +S2 hindcast are not available due to size. Derived datasets
+required to reproduce the results, including extent timeseries and error
+measures, are available in this Zenodo repository:
+https://zenodo.org/records/17479538
+\citep{campitelli2025}} %% use this section when having data sets and software code available
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontribution{EC performed the data analysis and wrote the
+manuscript draft. AP, JA, EL, MW and PR, performed interpretation of the
+results, and reviewed and edited the draft. All authors read and
+approved the final manuscript.} %% optional section
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\competinginterests{The authors declare no competing
+interests.} %% this section is mandatory even if you declare that no competing interests are present
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{acknowledgements}
+We thank the internal reviewers Bethan White and Xiaobing Zhou for their
+comments and feedback. This work benefited from earlier unpublished work
+by Laura Davies, Phil Reid, Andrew G. Marshall. This research was
+undertaken with the assistance of resources from the National
+Computational Infrastructure (NCI Australia), an NCRIS enabled
+capability supported by the Australian Government. This work was
+supported by ARC SRIEAS Grant SR200100005 Securing Antarctica's
+Environmental Future.
+\end{acknowledgements}
+
+\bibliographystyle{copernicus}
+
 \end{document}

### 2025-12-11 analysis/paper/access-ice.qmd
Adds section title to competing interests and etc...

@@
 competinginterests: |
+  # Competing interests
   The authors declare no competing interests.
@@
 # Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
-copyrightstatement: |
-  The author's copyright for this publication is transferred to institution/company. 
+# copyrightstatement: |
+#   The author's copyright for this publication is transferred to institution/company. 
 ### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
@@
   codedata: |
+    # Code/Data availability 
     The underlying code for this study is available on GitHub: https://github.com/eliocamp/access-s2_ice-eval.
@@
 authorcontribution: |
+    # Author contributions
     EC performed the data analysis and wrote the manuscript draft. 
@@
 acknowledgements: |
+    # Acknowledgements
     We thank the internal reviewers Bethan White and Xiaobing Zhou for their comments and feedback. 
@@
 Regional analysis of forecast skill suggests that this winter predictability barrier is most dramatic over East Antarctica, where even +S1 shows negative skill. 
-These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritize the summer-autumn period when initial conditions have maximum impact on forecast skill.
+These findings highlight the critical importance of comprehensive year-round sampling in predictability studies and suggest that operational sea-ice data assimilation efforts should prioritise the summer-autumn period when initial conditions have maximum impact on forecast skill.
 

### 2025-12-11 analysis/paper/access-ice.tex
Adds section title to competing interests and etc...

@@
 suggest that operational sea-ice data assimilation efforts should
-prioritize the summer-autumn period when initial conditions have maximum
+prioritise the summer-autumn period when initial conditions have maximum
 impact on forecast skill.
 \end{abstract}
-\copyrightstatement{The author's copyright for this publication is
-transferred to institution/company.}
 
@@
 
-\codedataavailability{The underlying code for this study is available on
-GitHub: https://github.com/eliocamp/access-s2\_ice-eval. Raw data of +S1
-and +S2 hindcast are not available due to size. Derived datasets
-required to reproduce the results, including extent timeseries and error
-measures, are available in this Zenodo repository:
+\codedataavailability{\section{Code/Data
+availability}\label{codedata-availability}
+
+The underlying code for this study is available on GitHub:
+https://github.com/eliocamp/access-s2\_ice-eval. Raw data of +S1 and +S2
+hindcast are not available due to size. Derived datasets required to
+reproduce the results, including extent timeseries and error measures,
+are available in this Zenodo repository:
 https://zenodo.org/records/17479538
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontribution{EC performed the data analysis and wrote the
-manuscript draft. AP, JA, EL, MW and PR, performed interpretation of the
-results, and reviewed and edited the draft. All authors read and
-approved the final manuscript.} %% optional section
+\authorcontribution{\section{Author
+contributions}\label{author-contributions}
+
+EC performed the data analysis and wrote the manuscript draft. AP, JA,
+EL, MW and PR, performed interpretation of the results, and reviewed and
+edited the draft. All authors read and approved the final
+manuscript.} %% optional section
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\competinginterests{The authors declare no competing
+\competinginterests{\section{Competing
+interests}\label{competing-interests}
+
+The authors declare no competing
 interests.} %% this section is mandatory even if you declare that no competing interests are present
@@
 \begin{acknowledgements}
+\section{Acknowledgements}\label{acknowledgements}
+
 We thank the internal reviewers Bethan White and Xiaobing Zhou for their

### 2025-12-11 build.sh

@@ -4,24 +4,6 @@ dir=analysis/paper
 source=$dir/access-ice.qmd
 zip=access-ice.zip
 
-# main=$dir/main.qmd 
-# supp=$dir/supp.qmd 
-
-
-# cp $source $main
-# quarto render $main --profile main
-# cp $dir/main.aux $dir/ref_main.aux
-
-# cp $source $supp
-# quarto render $supp --profile supp
-# cp $dir/supp.aux $dir/ref_supp.aux
-
-# rm $main $supp 
-
-
-# pdflatex main.tex
-# pdflatex supp.tex
-
 quarto render $source
 
 cd $dir
